### 2019-06-14 R/fluxes.Rmd
Create Rmarkdown report for fluxes

---
title: "Artificial Ecosystem Selection"
author: "Andrew Morris"
date: "`r Sys.Date()`"
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=8, fig.height=6, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE)
```

```{r load_packages, include=FALSE}

library(tidyverse)
library(lubridate)
library(broom)
library(broman)

```

```{r import_data}

top_jars_gen1 <- c(2, 3, 13, 16)

temp <- list.files(path="../Data", pattern="FluxData*", full.names = TRUE)
conc <- sapply(temp, read.csv, stringsAsFactors = FALSE,
                 simplify = FALSE, USE.NAMES = TRUE)
names(conc) <- substr(names(conc), 18, 25)
conc <- bind_rows(conc, .id = "Date")

# Code transfers
conc$transfer <- as.character(match(conc$Date, unique(conc$Date)))

initial_n1 <- length(unique(conc[conc$transfer == 1, ]$JAR))

initial_n2 <- length(unique(conc[conc$transfer == 2, ]$JAR))
```

# Overview

This is a pilot experiment with the following goals:

  1. Determine whether there is a response to selection on soil methane oxidation
  2. Determine the size of the inoculum and selection differential to use

A set of `r initial_n1` replicate ecosystems were generated
with 50 g of soil (sterile potting mix plus inoculum)
and sterile water to bring them to 60% of field capacity in 500 mL mason jars.
These jars were inoculated at 5% by mass and were initially kept at
atmospheric concentrations of methane. Later, they were capped and spiked to
a headspace concentration of approximately 1000 ppm CH4.  After about 8 weeks
of incubation, we finally observed significant (negative) methane fluxes. Jar
9 had the highest flux, but we observed a crack in the jar and so excluded that
data point. From the remaining `r initial_n1 - 1` jars, we selected the top 4 
(jars `r top_jars_gen1`) to homogenize and transfer to the next set of ecosystems.

```{r process_data}
# Jar broke so remove # 9 from gen 1
conc <- conc[!(conc$JAR == 9 & conc$Date == "20190527"), ]

# Jar # 1 and 2 from gen 2 had impossible ch4 values so remove
conc <- conc[!(conc$JAR == 1 & conc$Date == "20190607"), ]
conc <- conc[!(conc$JAR == 2 & conc$Date == "20190607"), ]

# Convert clock time to days and set t0 to 0 days
conc <- 
  conc  %>% 
  mutate(days = period_to_seconds(hms(paste0(conc$TIME, ':00'))) / 60 / 60) %>% 
  group_by(Date, JAR) %>% 
  mutate(days = days - min(days))

# Convert ppm to umol
conc$ch4_umol <- conc$CH4_CONC * 0.965 / ((0.08206 * (293)))

# Calculate fluxes from lm slopes
flux <- 
  conc %>% 
  group_by(transfer, JAR) %>%
  do(tidy(lm(ch4_umol ~ days, data = .))) %>% 
  filter(term == 'days')  %>% 
  select(transfer, JAR, flux_umold = estimate)

# Identify the jars that were used for selection
flux$selected[flux$JAR %in% top_jars_gen1 & flux$transfer == '1'] <- T
flux$selected[!(flux$JAR %in% top_jars_gen1) | flux$transfer != '1'] <- F

all_par <- flux$flux_umold[flux$transfer == '1']
offspr <- flux$flux_umold[flux$transfer == '2']
sel_par <- flux$flux_umold[flux$transfer == '1' & flux$selected == T]

```

For the first transfer, we used 10% inoculum intead of 5% inoculum and generated
`r initial_n2` jars, but two broke leaving `r initial_n2 -2`. After one week we observed
significant, negative methane fluxes that exceeded the first set of ecosystems. This
may have been a response to selection, a biomass effect due to the larger inoculum,
or both. So in order to truly demonstrate a response to selection we will generate at
least one more set of ecosystems through a transfer with the same inoculum size.

# Plot fluxes
```{r plot_fluxes}

ggplot(flux, aes(x = flux_umold, fill = selected)) +
  geom_histogram(aes(fill = interaction(transfer, selected))) +
  facet_grid(transfer ~ .) +
  scale_fill_manual(values = c('black', 'blue', 'red')) +
  geom_vline(xintercept = mean(all_par)) +
  geom_vline(xintercept = mean(offspr), color = 'blue') +
  geom_vline(xintercept = mean(sel_par), color = 'red') 
ggsave('../Figures/transfer12_hist.pdf', width = 5, height = 5)

ggplot(flux, aes(x = transfer, y = flux_umold)) +
  geom_boxplot(aes(color = transfer), outlier.shape = NA) +
  geom_jitter(aes(color = interaction(transfer, selected))) +
  scale_color_manual(values = c('black', 'black', 'red', 'blue', 'blue')) + 
  geom_hline(yintercept = mean(all_par)) +
  geom_hline(yintercept = mean(offspr), color = 'blue') +
  geom_hline(yintercept = mean(sel_par), color = 'red')

ggsave('../Figures/transfer12_box.pdf', width = 5, height = 5)
# Calculate heritability

```

# Calculate narrow-sense heritability

```{r calc_heritability}

ns_heritability <- function(all_par, offspr, sel_par) {
  R <- mean(all_par) - mean(offspr)
  S <- mean(all_par) - mean(sel_par)
  h2 <- R / S
  h2
  # Where:
  # h2 is the narrow-sense heritability
  # R is the realized average difference between the parent generation
  #     and the next generation
  # S is the average difference between the parent generation and the selected
  #     parents
}

nsh <- ns_heritability(all_par, offspr, sel_par) 
```

Based on the mean ecosystem-level trait values from the first and second set
of ecosystems, we calculated narrow-sense heritability as $h^2 = R / S$, which
was `r myround(nsh)`. Granted, this is also not a valid estimate, because of the difference
in inoculum size. We will calculate this again between transfer 2 and 3 to get a more
legitimate estimate.

# Identify selected individuals from transfer 2

```{r identify_top_jars} 
flux %>% 
  group_by(transfer) %>% 
  top_n(-4, flux_umold)
```

### 2019-06-16 R/fluxes.Rmd
Add new data and update analysis

@@
 conc <- bind_rows(conc, .id = "Date")
+# Code transfers
+#conc$transfer <- as.character(match(conc$Date, unique(conc$Date)))
+conc$transfer[conc$Date == 20190527] <- 1
+conc$transfer[conc$Date != 20190527] <- 2
 
-# Code transfers
-conc$transfer <- as.character(match(conc$Date, unique(conc$Date)))
+# Jar broke so remove # 9 from gen 1
+conc <- conc[!(conc$JAR == 9 & conc$Date == "20190527"), ]
 
-initial_n1 <- length(unique(conc[conc$transfer == 1, ]$JAR))
+# Jar # 1 and 2 from gen 2 had impossible ch4 values so remove
+conc <- conc[!(conc$JAR == 1 & conc$Date == "20190607"), ]
+conc <- conc[!(conc$JAR == 2 & conc$Date == "20190607"), ]
 
-initial_n2 <- length(unique(conc[conc$transfer == 2, ]$JAR))
+n1 <- length(unique(conc[conc$transfer == 1, ]$JAR))
+
+n2 <- length(unique(conc[conc$transfer == 2, ]$JAR))
 ```
@@
 
-A set of `r initial_n1` replicate ecosystems were generated
+A set of `r n1` replicate ecosystems were generated
 with 50 g of soil (sterile potting mix plus inoculum)
@@
 9 had the highest flux, but we observed a crack in the jar and so excluded that
-data point. From the remaining `r initial_n1 - 1` jars, we selected the top 4 
+data point. From the 'r n1' jars, we selected the top 4 
 (jars `r top_jars_gen1`) to homogenize and transfer to the next set of ecosystems.
@@
 ```{r process_data}
-# Jar broke so remove # 9 from gen 1
-conc <- conc[!(conc$JAR == 9 & conc$Date == "20190527"), ]
-
-# Jar # 1 and 2 from gen 2 had impossible ch4 values so remove
-conc <- conc[!(conc$JAR == 1 & conc$Date == "20190607"), ]
-conc <- conc[!(conc$JAR == 2 & conc$Date == "20190607"), ]
 
@@
   conc  %>% 
-  mutate(days = period_to_seconds(hms(paste0(conc$TIME, ':00'))) / 60 / 60) %>% 
+  mutate(days = as.numeric(difftime(ymd_hms(paste0(DATE, ' ', TIME, ':00')), +
+                       ISOdatetime(2019,1,1,0,0,0), unit="days"))) %>% 
   group_by(Date, JAR) %>% 
@@
 conc$ch4_umol <- conc$CH4_CONC * 0.965 / ((0.08206 * (293)))
+conc
+
+conc$selected[conc$JAR %in% top_jars_gen1 & conc$transfer == '1'] <- T
+conc$selected[!(conc$JAR %in% top_jars_gen1) | conc$transfer != '1'] <- F
+```
+
+```{r}
 
@@
   conc %>% 
-  group_by(transfer, JAR) %>%
+  group_by(Date, JAR) %>%
   do(tidy(lm(ch4_umol ~ days, data = .))) %>% 
-  filter(term == 'days')  %>% 
-  select(transfer, JAR, flux_umold = estimate)
+  filter(term == 'days')#  %>% 
+  select(Date, JAR, flux_umold = estimate)
 
-# Identify the jars that were used for selection
+flux <- 
+  conc %>% 
+  group_by(Date, JAR) %>%
+  #do(tidy(glm(ch4_umol ~ days, data = ., family = gaussian(link = "log")))) %>% 
+    nest() %>% 
+mutate(fit_glm = purrr::map(data, ~ glm(ch4_umol ~ days, data = ., family = gaussian(link = "log"))),
+         fit_lm = purrr::map(data, ~ lm(ch4_umol ~ days, data = .)))
+flux %>% 
+    mutate(aov = map2(fit_glm, fit_lm, stats::anova)) %>% 
+    pull(aov)
+
+out %>% 
+    mutate(aov = map2(fit_nls, fit_lm, stats::anova)) %>% 
+    pull(aov) %>% head(2)
+
+flux$transfer[flux$Date == 20190527] <- 1
+flux$transfer[flux$Date != 20190527] <- 2
+flux_log$transfer[flux_log$Date == 20190527] <- 1
+flux_log$transfer[flux_log$Date != 20190527] <- 2
+
 flux$selected[flux$JAR %in% top_jars_gen1 & flux$transfer == '1'] <- T
 flux$selected[!(flux$JAR %in% top_jars_gen1) | flux$transfer != '1'] <- F
-
-all_par <- flux$flux_umold[flux$transfer == '1']
-offspr <- flux$flux_umold[flux$transfer == '2']
+flux_log$selected[flux_log$JAR %in% top_jars_gen1 & flux_log$transfer == '1'] <- T
+flux_log$selected[!(flux_log$JAR %in% top_jars_gen1) | flux_log$transfer != '1'] <- F
+unique(flux$Date)
+all_par <- flux$flux_umold[flux$Date == '20190527']
+offspr1 <- flux$flux_umold[flux$Date == '20190607']
+offspr2 <- flux$flux_umold[flux$Date == '20190614']
 sel_par <- flux$flux_umold[flux$transfer == '1' & flux$selected == T]
 
+all_par_log <- flux_log$flux_umold[flux_log$Date == '20190527']
+offspr1_log <- flux_log$flux_umold[flux_log$Date == '20190607']
+offspr2_log <- flux_log$flux_umold[flux_log$Date == '20190614']
+sel_par_log <- flux_log$flux_umold[flux_log$transfer == '1' & flux_log$selected == T]
+```
+
+```{r plot_flux_curves}
+ggplot(conc, aes(x = days, y = ch4_umol, group = JAR)) + 
+  geom_point() + 
+  facet_wrap(~ Date, scales = "free_x")  + 
+    geom_smooth(method = "glm", se = F, 
+        method.args = list(family = gaussian(link = "log")))
 ```
@@
 
-ggplot(flux, aes(x = flux_umold, fill = selected)) +
-  geom_histogram(aes(fill = interaction(transfer, selected))) +
-  facet_grid(transfer ~ .) +
-  scale_fill_manual(values = c('black', 'blue', 'red')) +
+ggplot(flux, aes(x = flux_umold)) +
+  geom_histogram(aes(fill = selected)) +
+  facet_grid(Date ~ .) +
+  scale_fill_manual(values = c('black', 'red')) +
   geom_vline(xintercept = mean(all_par)) +
-  geom_vline(xintercept = mean(offspr), color = 'blue') +
+  geom_vline(xintercept = mean(offspr1), color = 'blue') +
+  geom_vline(xintercept = mean(offspr2), color = 'darkblue') +
   geom_vline(xintercept = mean(sel_par), color = 'red')
-ggsave('../Figures/transfer12_hist.pdf', width = 5, height = 5)
 
-ggplot(flux, aes(x = transfer, y = flux_umold)) +
-  geom_boxplot(aes(color = transfer), outlier.shape = NA) +
-  geom_jitter(aes(color = interaction(transfer, selected))) +
-  scale_color_manual(values = c('black', 'black', 'red', 'blue', 'blue')) + 
+ggplot(flux, aes(x = Date, y = flux_umold)) +
+  geom_boxplot(outlier.shape = NA) +
+  geom_jitter(aes(color =selected)) +
+  scale_color_manual(values = c('black', 'red')) + 
   geom_hline(yintercept = mean(all_par)) +
-  geom_hline(yintercept = mean(offspr), color = 'blue') +
+  geom_hline(yintercept = mean(offspr1), color = 'blue') +
+  geom_hline(yintercept = mean(offspr2), color = 'darkblue') +
   geom_hline(yintercept = mean(sel_par), color = 'red')
 
-ggsave('../Figures/transfer12_box.pdf', width = 5, height = 5)
-# Calculate heritability
+ggplot(flux_log, aes(x = flux_umold)) +
+  geom_histogram(aes(fill = selected)) +
+  facet_grid(Date ~ .) +
+  scale_fill_manual(values = c('black', 'red')) +
+  geom_vline(xintercept = mean(all_par_log)) +
+  geom_vline(xintercept = mean(offspr1_log), color = 'blue') +
+  geom_vline(xintercept = mean(offspr2_log), color = 'darkblue') +
+  geom_vline(xintercept = mean(sel_par_log), color = 'red')
 
+ggplot(flux_log, aes(x = Date, y = flux_umold)) +
+  geom_boxplot(outlier.shape = NA) +
+  geom_jitter(aes(color =selected)) +
+  scale_color_manual(values = c('black', 'red')) + 
+  geom_hline(yintercept = mean(all_par_log)) +
+  geom_hline(yintercept = mean(offspr1_log), color = 'blue') +
+  geom_hline(yintercept = mean(offspr2_log), color = 'darkblue') +
+  geom_hline(yintercept = mean(sel_par_log), color = 'red') 
 ```
@@
 
-nsh <- ns_heritability(all_par, offspr, sel_par) 
+nsh1 <- ns_heritability(all_par, offspr1, sel_par) 
+nsh2 <- ns_heritability(all_par, offspr2, sel_par) 
+nsh1_log <- ns_heritability(all_par_log, offspr1_log, sel_par_log)
+nsh2_log <- ns_heritability(all_par_log, offspr2_log, sel_par_log)
 ```
@@
 of ecosystems, we calculated narrow-sense heritability as $h^2 = R / S$, which
-was `r myround(nsh)`. Granted, this is also not a valid estimate, because of the difference
+was `r myround(nsh)` for the linear model and `r myround(nsh_log)` for the exponential
+model. Granted, this is also not a valid estimate, because of the difference
 in inoculum size. We will calculate this again between transfer 2 and 3 to get a more
@@
 flux %>% 
-  group_by(transfer) %>% 
+  group_by(Date) %>% 
+  top_n(-4, flux_umold)
+flux_log %>% 
+  group_by(Date) %>% 
   top_n(-4, flux_umold)

### 2019-06-16 R/fluxes.Rmd
Work on modeling fluxes

@@
 conc$ch4_umol <- conc$CH4_CONC * 0.965 / ((0.08206 * (293)))
-conc
 
@@
 
-```{r}
+```{r calc_flux}
 
-# Calculate fluxes from lm slopes
+## Calculate fluxes from lm slopes
+#flux <- 
+#  conc %>% 
+#  group_by(Date, JAR) %>%
+#  do(tidy(lm(ch4_umol ~ days, data = .))) %>% 
+#  filter(term == 'days')#  %>% 
+#  select(Date, JAR, flux_umold = estimate)
+ggplot(conc[conc$Date == "20190614" & conc$JAR == 8,], aes(x = days, y = ch4_umol)) +
+  geom_point() +
+  stat_smooth(method = "glm", method.args = list(family = gaussian(link = "log")))
 flux <- 
   conc %>% 
-  group_by(Date, JAR) %>%
-  do(tidy(lm(ch4_umol ~ days, data = .))) %>% 
-  filter(term == 'days')#  %>% 
-  select(Date, JAR, flux_umold = estimate)
-
-flux <- 
-  conc %>% 
-  group_by(Date, JAR) %>%
+  group_by(transfer, Date, JAR, selected) %>%
   #do(tidy(glm(ch4_umol ~ days, data = ., family = gaussian(link = "log")))) %>% 
@@
          fit_lm = purrr::map(data, ~ lm(ch4_umol ~ days, data = .)))
+results <-
 flux %>% 
-    mutate(aov = map2(fit_glm, fit_lm, stats::anova)) %>% 
-    pull(aov)
-
-out %>% 
-    mutate(aov = map2(fit_nls, fit_lm, stats::anova)) %>% 
-    pull(aov) %>% head(2)
+  mutate(glance_lm = map(fit_lm, glance),
+         glance_glm = map(fit_glm, glance)) %>% 
+        unnest(glance_lm, glance_glm) %>% 
+        select(transfer, Date, JAR, fit_lm, fit_glm, deviance, deviance1) %>% 
+        mutate(diff_deviance = deviance - deviance1) %>% 
+        select(-deviance, -deviance1)
 

### 2019-06-17 R/fluxes.Rmd
Update flux calculations

--- 2019-06-16 R/fluxes.Rmd
@@
 library(broman)
+library(purrr)
+library(kableExtra)
 
@@
 A set of `r n1` replicate ecosystems were generated
-with 50 g of soil (sterile potting mix plus inoculum)
+with 50 g of soil (5% sterile potting mix plus inoculum)
 and sterile water to bring them to 60% of field capacity in 500 mL mason jars.
-These jars were inoculated at 5% by mass and were initially kept at
+These jars were initially kept at
 atmospheric concentrations of methane. Later, they were capped and spiked to
 a headspace concentration of approximately 1000 ppm CH4.  After about 8 weeks
-of incubation, we finally observed significant (negative) methane fluxes. Jar
-9 had the highest flux, but we observed a crack in the jar and so excluded that
-data point. From the 'r n1' jars, we selected the top 4 
+of incubation (2 weeks at 1000 ppm), we finally observed significant (negative) methane fluxes. From the 'r n1' jars, we selected the top 4 
 (jars `r top_jars_gen1`) to homogenize and transfer to the next set of ecosystems.
@@
 
-```{r}
+```{r plot_flux_curves, fig.width=10, fig.height=15}
+ggplot(conc, aes(x = days, y = ch4_umol, group = JAR)) + 
+  geom_point() + 
+  facet_wrap(~ Date + JAR, scales = "free")  + 
+  stat_smooth(method = "glm", se = F, color = 'blue', alpha = 0.5, geom = 'line',
+              method.args = list(family = gaussian(link = "log"))) + 
+  stat_smooth(method = 'lm', se = F, color = 'red', alpha = 0.5, geom = 'line')
+```
 
-# Calculate fluxes from lm slopes
-flux <- 
-  conc %>% 
-  group_by(Date, JAR) %>%
-  do(tidy(lm(ch4_umol ~ days, data = .))) %>% 
-  filter(term == 'days')#  %>% 
-  select(Date, JAR, flux_umold = estimate)
+```{r calc_flux}
+
+#ggplot(conc[conc$Date == "20190614" & conc$JAR == 8,], aes(x = days, y = ch4_umol)) +
+#  geom_point() +
+#  stat_smooth(method = "glm", method.args = list(family = gaussian(link = "log"))) +
+#  stat_smooth(method = "lm", color = "red")
+#
+#ggplot(conc[conc$Date == "20190614" & conc$JAR == 20,], aes(x = days, y = ch4_umol)) +
+#  geom_point() +
+#  stat_smooth(method = "glm", method.args = list(family = gaussian(link = "log"))) +
+#  stat_smooth(method = "lm", color = "red")
+#
+#ggplot(conc[conc$Date == "20190527" & conc$JAR == 16,], aes(x = days, y = ch4_umol)) +
+#  geom_point() +
+#  stat_smooth(method = "glm", method.args = list(family = gaussian(link = "log"))) +
+#  stat_smooth(method = "lm", color = "red")
 
@@
   conc %>% 
-  group_by(Date, JAR) %>%
-  #do(tidy(glm(ch4_umol ~ days, data = ., family = gaussian(link = "log")))) %>% 
+  group_by(transfer, Date, JAR, selected) %>%
   nest() %>%
-mutate(fit_glm = purrr::map(data, ~ glm(ch4_umol ~ days, data = ., family = gaussian(link = "log"))),
-         fit_lm = purrr::map(data, ~ lm(ch4_umol ~ days, data = .)))
+  mutate(fit_lm = map(data, ~ lm(ch4_umol ~ days, data = .)),
+         fit_glm = map(data, ~ glm(ch4_umol ~ days, data = ., family = gaussian(link = "log"))))
+
+  comp_dev <- 
 flux  %>% 
-    mutate(aov = map2(fit_glm, fit_lm, stats::anova)) %>% 
-    pull(aov)
+  mutate(glance_lm = map(fit_lm, glance),
+         glance_glm = map(fit_glm, glance)) %>% 
+  unnest(glance_lm, glance_glm)  %>% 
+  select(transfer, Date, JAR, selected, data, fit_lm, fit_glm, df.null, deviance_lm = deviance, deviance_glm = deviance1) %>% 
+  mutate(dev_diff = deviance_lm - deviance_glm, dev_bool = deviance_lm >= deviance_glm)
 
-out %>% 
-    mutate(aov = map2(fit_nls, fit_lm, stats::anova)) %>% 
-    pull(aov) %>% head(2)
+flux <- 
+flux %>%
+  mutate(tidy_lm = map(fit_lm, tidy),
+         tidy_glm = map(fit_glm, tidy)) %>% 
+unnest(tidy_lm, tidy_glm) %>% 
+filter(term == 'days') %>% 
+select(transfer, Date, JAR, selected, lm_umold = estimate, glm_umold = estimate1)
 
-flux$transfer[flux$Date == 20190527] <- 1
-flux$transfer[flux$Date != 20190527] <- 2
-flux_log$transfer[flux_log$Date == 20190527] <- 1
-flux_log$transfer[flux_log$Date != 20190527] <- 2
+all_par <- flux$lm_umold[flux$Date == '20190527']
+offspr1 <- flux$lm_umold[flux$Date == '20190607']
+offspr2 <- flux$lm_umold[flux$Date == '20190614']
+sel_par <- flux$lm_umold[flux$transfer == '1' & flux$selected == T]
 
-flux$selected[flux$JAR %in% top_jars_gen1 & flux$transfer == '1'] <- T
-flux$selected[!(flux$JAR %in% top_jars_gen1) | flux$transfer != '1'] <- F
-flux_log$selected[flux_log$JAR %in% top_jars_gen1 & flux_log$transfer == '1'] <- T
-flux_log$selected[!(flux_log$JAR %in% top_jars_gen1) | flux_log$transfer != '1'] <- F
-unique(flux$Date)
-all_par <- flux$flux_umold[flux$Date == '20190527']
-offspr1 <- flux$flux_umold[flux$Date == '20190607']
-offspr2 <- flux$flux_umold[flux$Date == '20190614']
-sel_par <- flux$flux_umold[flux$transfer == '1' & flux$selected == T]
+all_par_log <- flux$glm_umold[flux$Date == '20190527']
+offspr1_log <- flux$glm_umold[flux$Date == '20190607']
+offspr2_log <- flux$glm_umold[flux$Date == '20190614']
+sel_par_log <- flux$glm_umold[flux$transfer == '1' & flux$selected == T]
 
-all_par_log <- flux_log$flux_umold[flux_log$Date == '20190527']
-offspr1_log <- flux_log$flux_umold[flux_log$Date == '20190607']
-offspr2_log <- flux_log$flux_umold[flux_log$Date == '20190614']
-sel_par_log <- flux_log$flux_umold[flux_log$transfer == '1' & flux_log$selected == T]
-```
-
-```{r plot_flux_curves}
-ggplot(conc, aes(x = days, y = ch4_umol, group = JAR)) + 
-  geom_point() + 
-  facet_wrap(~ Date, scales = "free_x")  + 
-    geom_smooth(method = "glm", se = F, 
-        method.args = list(family = gaussian(link = "log")))
 ```
@@
 For the first transfer, we used 10% inoculum intead of 5% inoculum and generated
-`r initial_n2` jars, but two broke leaving `r initial_n2 -2`. After one week we observed
+`r n2` jars. After one week we observed
 significant, negative methane fluxes that exceeded the first set of ecosystems. This
 may have been a response to selection, a biomass effect due to the larger inoculum,
-or both. So in order to truly demonstrate a response to selection we will generate at
-least one more set of ecosystems through a transfer with the same inoculum size.
+or both. After two weeks, we observed dampened flxues compared to week one, but still
+greater flux than before the firs transfer. The elevated fluxes in week 1 compared
+to week 2 could have been a result of using two different standard curves for t0 and t1
+and for t2 and t3 in week 1 or due to some temporal effect during the incubations 
+(e.g. response to water addition or initial available nutrients or growth).
+In order to truly demonstrate a response to selection we will generate at
+least one more set of ecosystems through a transfer with the same inoculum size
+(10%).
 
@@
 
-ggplot(flux, aes(x = flux_umold)) +
+ggplot(flux, aes(x = lm_umold)) +
   geom_histogram(aes(fill = selected)) +
@@
 
-ggplot(flux, aes(x = Date, y = flux_umold)) +
+ggplot(flux, aes(x = Date, y = lm_umold)) +
   geom_boxplot(outlier.shape = NA) +
@@
 
-ggplot(flux_log, aes(x = flux_umold)) +
+ggplot(flux, aes(x = glm_umold)) +
   geom_histogram(aes(fill = selected)) +
@@
 
-ggplot(flux_log, aes(x = Date, y = flux_umold)) +
+ggplot(flux, aes(x = Date, y = glm_umold)) +
   geom_boxplot(outlier.shape = NA) +
@@
 Based on the mean ecosystem-level trait values from the first and second set
-of ecosystems, we calculated narrow-sense heritability as $h^2 = R / S$, which
-was `r myround(nsh)` for the linear model and `r myround(nsh_log)` for the exponential
+of ecosystems (week 2), we calculated narrow-sense heritability as $h^2 = R / S$, which
+was `r myround(nsh2)` for the linear model and `r myround(nsh2_log)` for the exponential
 model. Granted, this is also not a valid estimate, because of the difference
-in inoculum size. We will calculate this again between transfer 2 and 3 to get a more
-legitimate estimate.
+in inoculum size between transfer 1 and 2. We will calculate this again between
+transfer 2 and 3 to get a more accurate estimate.
 
@@
   group_by(Date) %>% 
-  top_n(-4, flux_umold)
-flux_log %>% 
+  top_n(-4, lm_umold) %>% 
+  kable(caption = "Linear model fluxes") %>% 
+  kable_styling()
+flux %>% 
   group_by(Date) %>% 
-  top_n(-4, flux_umold)
+  top_n(-4, glm_umold) %>% 
+  kable(caption = "Generalized linear model link = log fluxes") %>% 
+  kable_styling()
 ```

### 2019-06-17 R/fluxes.Rmd
Add detail to markdown writeup

@@
 
-# Overview
+## Overview
 
-This is a pilot experiment with the following goals:
+This is a pilot experiment to demonstrate a response to selection on
+soil methane oxidation at the ecosystem/community level. We hope to demonstrate a
+change in mean trait value as a response to selection that is greater than 
+neutral selection, to calculate community heritability using the breeder's equation, 
+and to identify microbial taxa that change in abundance or presence/absence due to 
+selection. This pilot experiment has the following goals:
 
   1. Determine whether there is a response to selection on soil methane oxidation
-  2. Determine the size of the inoculum and selection differential to use
+  2. Determine the experimental setup including the size of the inoculum, the 
+     selection differential, the generation time, and the environmental conditions
+     (e.g. methane concentration) to achieve a response to selection
 
 A set of `r n1` replicate ecosystems were generated
-with 50 g of soil (5% sterile potting mix plus inoculum)
+with 50 g of soil (sterile potting mix plus 5% inoculum by mass)
 and sterile water to bring them to 60% of field capacity in 500 mL mason jars.
-These jars were initially kept at
-atmospheric concentrations of methane. Later, they were capped and spiked to
-a headspace concentration of approximately 1000 ppm CH4.  After about 8 weeks
-of incubation (2 weeks at 1000 ppm), we finally observed significant (negative) methane fluxes. From the 'r n1' jars, we selected the top 4 
+These jars were initially kept at atmospheric concentrations of methane, but did not 
+display significant methane oxidation rates after ~6 weeks. As a result, 
+they were capped and spiked to
+a headspace concentration of approximately 1000 ppm CH4. Both CH4 and CO2 were monitored
+daily and it was determined that flushing and respiking was required every three
+days in order to maintain aerobic conditions and an elevated CH4 concentration. After about 8 weeks
+of incubation (2 weeks at 1000 ppm), we finally observed significant (negative)
+methane fluxes. From the `r n1` jars, we selected the top 4 
 (jars `r top_jars_gen1`) to homogenize and transfer to the next set of ecosystems.
+
+For the first transfer, we used 10% inoculum intead of 5% inoculum and generated
+`r n2` jars. After one week we observed
+significant, negative methane fluxes that exceeded the first set of ecosystems. This
+may have been a response to selection, a biomass effect due to the larger inoculum,
+or both. After two weeks, we observed dampened fluxes compared to week one, but still
+greater flux than before the first transfer. The elevated fluxes in week 1 compared
+to week 2 could have been a result of using two different standard curves for t0 and t1
+and for t2 and t3 in week 1 or due to some temporal effect during the incubations 
+(e.g. response to water addition or initial available nutrients or growth).
+In order to truly demonstrate a response to selection we will generate at
+least one more set of ecosystems through a transfer with the same inoculum size
+(10%).
 
@@
 
+## Raw flux curves
+
+The raw methane concentration data for each methane flux measurement labelled by
+jar and the date the measurement was taken. The red lines indicate the regression
+line for OLS regression and the blue line indicates the regression line for 
+glm (link = log). Both x and y axes are allowed to vary.
+
 ```{r plot_flux_curves, fig.width=10, fig.height=15}
@@
   facet_wrap(~ Date + JAR, scales = "free")  + 
-  stat_smooth(method = "glm", se = F, color = 'blue', alpha = 0.5, geom = 'line',
+  stat_smooth(method = "glm", se = F, color = 'blue', alpha = 0.4, geom = 'line',
+              size = 1,
               method.args = list(family = gaussian(link = "log"))) + 
-  stat_smooth(method = 'lm', se = F, color = 'red', alpha = 0.5, geom = 'line')
+  stat_smooth(method = 'lm', se = F, color = 'red', alpha = 0.4, geom = 'line', size = 1) +
+  labs(x = "Time (days)", y = expression(CH[4] ~ "Concentration" ~ (mu * Mol)))
 ```
@@
 
-For the first transfer, we used 10% inoculum intead of 5% inoculum and generated
-`r n2` jars. After one week we observed
-significant, negative methane fluxes that exceeded the first set of ecosystems. This
-may have been a response to selection, a biomass effect due to the larger inoculum,
-or both. After two weeks, we observed dampened flxues compared to week one, but still
-greater flux than before the firs transfer. The elevated fluxes in week 1 compared
-to week 2 could have been a result of using two different standard curves for t0 and t1
-and for t2 and t3 in week 1 or due to some temporal effect during the incubations 
-(e.g. response to water addition or initial available nutrients or growth).
-In order to truly demonstrate a response to selection we will generate at
-least one more set of ecosystems through a transfer with the same inoculum size
-(10%).
+## Plot fluxes
 
-# Plot fluxes
+Histograms and boxplots of the distribution of methane oxidation rates across jars. 
+The first date is from the first set of jars with the four jars used as the
+inoculum in the first transfer indicated in red. The second date is Transfer 2, Week1
+and the third date is Transfer 2, Week 1. The black line is the
+mean methane flux for the parent generation, the red line is the mean methane
+flux for the selected jars, the light blue line is the mean methane flux
+for transfer 2 week 1 and the dark blue line is the mean metahne flux for
+transfer 2 week 2. Methane fluxes are negtaive, because it's a consumptive
+process (methane oxidation). The first two plots have fluxes calculated from simple
+linear regression. The second two plots have fluxes calculated from a glm
+(family = Gaussian, link = log) to better match the exponential decay function
+for methane oxidation.
+
 ```{r plot_fluxes}
@@
 ggplot(flux, aes(x = lm_umold)) +
-  geom_histogram(aes(fill = selected)) +
+  geom_histogram(aes(fill = interaction(selected, Date))) +
   facet_grid(Date ~ .) +
-  scale_fill_manual(values = c('black', 'red')) +
+  scale_fill_manual(values = c('black', 'red', 'blue', 'darkblue')) +
   geom_vline(xintercept = mean(all_par)) +
@@
   geom_vline(xintercept = mean(offspr2), color = 'darkblue') +
-  geom_vline(xintercept = mean(sel_par), color = 'red')
+  geom_vline(xintercept = mean(sel_par), color = 'red') +
+  labs(x = expression (CH[4] ~ "Flux" ~ (mu * "Mol"* ~ d^{-1})), y = "Number of Jars", title = "LM Fluxes")
 
@@
   geom_boxplot(outlier.shape = NA) +
-  geom_jitter(aes(color =selected)) +
-  scale_color_manual(values = c('black', 'red')) + 
+  geom_jitter(aes(color = interaction(selected, Date))) +
+  scale_color_manual(values = c('black', 'red', 'blue', 'darkblue')) + 
   geom_hline(yintercept = mean(all_par)) +
@@
   geom_hline(yintercept = mean(offspr2), color = 'darkblue') +
-  geom_hline(yintercept = mean(sel_par), color = 'red')
+  geom_hline(yintercept = mean(sel_par), color = 'red')  + 
+  labs(y = expression (CH[4] ~ "Flux" ~ (mu * "Mol"* ~ d^{-1})), x = "Measurement Date", title = "LM Fluxes")
 
 ggplot(flux, aes(x = glm_umold)) +
-  geom_histogram(aes(fill = selected)) +
+  geom_histogram(aes(fill = interaction(selected, Date))) +
   facet_grid(Date ~ .) +
-  scale_fill_manual(values = c('black', 'red')) +
+  scale_fill_manual(values = c('black', 'red', 'blue', 'darkblue')) +
   geom_vline(xintercept = mean(all_par_log)) +
@@
   geom_vline(xintercept = mean(offspr2_log), color = 'darkblue') +
-  geom_vline(xintercept = mean(sel_par_log), color = 'red')
+  geom_vline(xintercept = mean(sel_par_log), color = 'red') +
+  labs(x = expression (CH[4] ~ "Flux" ~ (mu * "Mol"* ~ d^{-1})), y = "Number of Jars", title = "GLM Fluxes")
 
@@
   geom_boxplot(outlier.shape = NA) +
-  geom_jitter(aes(color =selected)) +
-  scale_color_manual(values = c('black', 'red')) + 
+  geom_jitter(aes(color = interaction(selected, Date))) +
+  scale_color_manual(values = c('black', 'red', 'blue', 'darkblue')) + 
   geom_hline(yintercept = mean(all_par_log)) +
@@
   geom_hline(yintercept = mean(offspr2_log), color = 'darkblue') +
-  geom_hline(yintercept = mean(sel_par_log), color = 'red') 
+  geom_hline(yintercept = mean(sel_par_log), color = 'red') +
+  labs(y = expression (CH[4] ~ "Flux" ~ (mu * "Mol"* ~ d^{-1})), x = "Measurement Date", title = "GLM Fluxes")
 ```
 
-# Calculate narrow-sense heritability
+## Calculate narrow-sense heritability
 
@@
 
-# Identify selected individuals from transfer 2
+## Identify selected individuals from transfer 2
 
@@
   top_n(-4, lm_umold) %>% 
-  kable(caption = "Linear model fluxes") %>% 
+  kable(caption = "Linear model fluxes. transfer is the ecosystem generation,
+        Date is the flux measurement date, JAR lists the top 4 jars from that sample
+        date in terms of methane flux from a linear model, 
+        selected indicates whether they were the
+        selected parents for transfer 1, lm_umold is uMol/d flux based on a linear
+        regression and glm_umold is uMol/d flux based on a glm with family = Gaussian
+        and link = log") %>% 
   kable_styling()
@@
   top_n(-4, glm_umold) %>% 
-  kable(caption = "Generalized linear model link = log fluxes") %>% 
+  kable(caption = "Generalized linear model link = log fluxes. Same as above except
+        JAR indicates top 4 jars from a glm.") %>% 
   kable_styling()

### 2019-09-18 Documents/oxidation_protocol.Rmd
Update protocol

---
title: "Oxidation Protocol"
author: "Andrew Morris"
date: "9/9/2019"
output:
  html_document:
      toc: true
      toc_float: 
        collapsed: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,  fig.align="center")
library(tidyverse)
library(knitr)
library(kableExtra)
library(broom)
library(modelr)
library(lubridate)
```

## General notes

- Record everything in the lab notebook.
- Take note of any mistakes or errors.
- Ideally start flux measurements around 8:00 or 9:00 o'clock on Monday
  morning to allow time for three measurements on that day.

## Initial setup
1. Sterilize potting mix, DI water, and mason jar disks in the autoclave.
2. Sterilize jars with 70% ethanol.
3. Add 45 g of steril potting mix, 5 g of live soil, and 3.5 mL of sterile DI water to each jar and cap.
4. Spike each jar with 4.3 mL of 99% CH~4~.

Allow these to incubate for at least two weeks before taking a flux
measurements. Also, uncap them in the hood for ~30 minutes and respike with 4.3
mL of 99% CH~4~ every 3-4 days to maintain aerobic conditions and maintain
elevated CH~4~. So, for example, if the jars are spiked on Thursday, flush them
on Monday and Thursday of the following two weeks and then take the flux
measurements on the third Monday.

## Flux sampling

### Monday

1. Uncap jars in hood with blower on to return them to atmospheric gas
concentrations and to avoid cross-contamination.
2. Measure standard curve on GC with 5-12 points for CH~4~. Record the
   concentration of the standard in the bottle (ppm), the volume of the
   injected standard (mL), the
concentration of the injected standard (ppm), and the CH~4~ curve area. Also record the
slope and y-intercept for the resulting equation as well as the R^2^. Continue running standard curves until the R^2^ is 99% and ideally 99.9%.
3. Cap jars in the hood.
4. Spike jars to ~1000 ppm CH~4~ using 0.43 mL of 99% CH~4~ and record time for
   each jar to the minute (t~0~).
5. Measure t~1~ immediately after spiking all jars resulting in a t~1~ that is
20 to 30 minutes after t~0~. Measure jars in the order that they were spiked.
Record the time, CH~4~ curve area, and CO~2~ curve area.
6. Record t~2~ after 3-5 hours.
7. Record t~3~ after 5-8 hours.

### Tuesday

8. Record t~4~ after 24 hours.

### Wednesday

9. Record t~5~ after 48 hours.
10. Enter the standard curve and flux data into R. Visually inspect the graphs
    to determine how many points to include in the linear portion of the curve
    (the first 3 to 5 points) and calculate the linear decrease in CH~4~
    concentration (Vmax). 
11. Determine the top 2 (16.7%) or 3 (25%) of the ~12 replicate jars for the
    positive selection treatment (p). Homogenize and reserve the soil from
    those jars in the hood. Randomly select an equal number of jars from the
    neutral selection treatment (n). Homogenize and reserve the soil from those
    jars in the biosafety cabinet. Clearly label each of these containers.
12. Empty the rest of the soil from the jars into an autoclave container, cover
    it with aluminum foil, and label it "disposal". Clean the jars and lids
    with soap and rinse and dry thoroughly.
13. Add at least `r 45 * 24` grams of fresh potting mix to an autoclave tray
    and cover with aluminum foil. Wrap the disks in a foil packet. Set aside
    the rings, they do not need to be autoclaved.
14. Autoclave the potting soil, lids, and disposal soil on the gravity cycle for 40
    minutes. 
15. Remove from the autoclave, put the jars and lids into the hood, and allow
    to cool overnight.

### Thursday

16. Sterilize the inside of each jar with 70% ethanol.
17. Add 45 grams of sterile potting mix to each jar.
18. Add 5 grams of the selected soil homogenate to each 'p' jar and 5 grams of
    the random soil to each 'n' jar. 
19. Add 3.5 mL of sterile DI water to each jar. 
20. Cap each jar and spike with 4.3 mL CH~4~.

## Preferred data format

Dates should be recorded as `yyyy-mm-dd` and times should be recorded as
`yyyy-mm-dd hh:mm` (seconds are optional). Use standard column headers with no
caps and no spaces. Instead, simply use "variable_units" format.

Column Headers | Meaning
-------------- | ----------------------------------------------------------
`flux_date`    | The date flux measurements began (i.e. Monday)
`standard_ppm` | The concentration of the standard in the bottle. 
`injection_ml` | The volume of standard injected into the GC. 
`injectin_ppm` | The concentration of standard injected into the gc (standard_ppm multiplied by injection_ml) 
`ch4_area`     | The area-under-the-curve for the methane peak output from the GC.

### Standard Curve

```{r sc_data, fig.height=3}
example_sc <- data.frame(
  flux_date = c('2019-05-27', '2019-05-27', '2019-05-27', '2019-06-07', '2019-06-07', '2019-06-07', '2019-06-14', '2019-06-14', '2019-06-14'),
  standard_ppm = c(999, 999, 999, 999, 999, 999, 999, 999, 999),
  injection_ml = c(0.1, 0.5, 1, 0.1, 0.5, 1, 0.1, 0.5, 1),
  ch4_area = c(556.03, 3046.03, 5995.73, 493.256, 2720.52, 5259.6, 569.3, 3255.82, 6129.72))
example_sc <- 
  example_sc %>% 
  mutate(injection_ppm = standard_ppm * injection_ml) %>% 
  select(flux_date:injection_ml, injection_ppm, ch4_area)
example_sc %>% 
  kable() %>% 
  kable_styling()
ggplot(example_sc, aes(ch4_area, injection_ppm)) +
  facet_wrap(~ flux_date) + 
  geom_point() +
  geom_line() +
  coord_fixed(6)

sc_model <- function(df) {lm(injection_ppm ~ ch4_area, data = df)}
standard_curves <- 
  example_sc %>% 
  group_by(flux_date) %>% 
  nest() %>% 
  mutate(model = map(data, sc_model),
         glance = map(model, glance),
         tidy = map(model, tidy)) 
sc_r2 <- 
  standard_curves %>% 
  unnest(glance, .drop = TRUE) %>% 
  arrange(r.squared) %>% 
  select(flux_date, r.squared)
sc_equation <- 
  standard_curves %>% 
  unnest(tidy, .drop = TRUE) %>%
  select(flux_date, term, estimate) %>% 
  spread(term, estimate) %>% 
  rename(intercept = `(Intercept)`, slope = ch4_area)
sc_ch4 <- left_join(sc_r2, sc_equation, by = "flux_date")
sc_ch4 %>% 
  kable() %>% 
  kable_styling()
```

### Flux Measurements

```{r time_data}

time_data <- data.table::fread('../Data/eg_time_data.csv')

time_data[, c('t0', 't1', 't2', 't3', 't4')] <- 
  lapply(time_data[, c('t0', 't1', 't2', 't3', 't4')],
         ymd_hms)

as.days <- function(start, end) {as.numeric(interval(start, end))/60/60/24}

time_data[, c('t1', 't2', 't3', 't4')] <- 
  lapply(time_data[, c('t1', 't2', 't3', 't4')],
         function(x) as.days(time_data$t0, x))

time_data <- 
 time_data %>% 
  select(-t0) %>% 
  gather(t, days, t1:t4)

```

```{r conc_data, warning=FALSE}

conc_data <- data.table::fread('../Data/eg_conc_data.csv')

flux_data <- 
  conc_data %>% 
  gather(t, ch4_area, t1:t4) %>% 
  left_join(sc_ch4, by = 'flux_date') %>% 
  mutate(ch4_ppm = ch4_area * slope + intercept) %>% 
  select(flux_date, jar, t, ch4_ppm) %>% 
  left_join(time_data, by = c('flux_date', 'jar', 't'))

```

```{r plot_fluxes, fig.height=20}
flux_data <- 
  flux_data %>% 
  mutate(flux_date = ymd(flux_date), 
         jar = factor(jar),
         t = factor(t)) %>% 
  filter(ch4_ppm < 3000)

flux_data %>% 
  ggplot(aes(days, ch4_ppm)) + 
  facet_grid(jar ~ flux_date, scales = "free") +
  geom_point() + 
  geom_line()
```

### Calculate Flux

```{r calc_flux}

```

### Convert ppm to $\mu$grams

```{r conversion}

```

### Selection

### 2019-09-18 R/fluxes.Rmd
Update protocol

@@
 library(kableExtra)
+library(morris)
 
@@
 
-top_jars_gen1 <- c(2, 3, 13, 16)
+top_jars_pas1 <- c(2, 3, 13, 16)
+top_jars_pas2 <- c(6, 8, 17, 21)
 
@@
 n2 <- length(unique(conc[conc$transfer == 2, ]$JAR))
+
+# The two SCs from passage 2, week 1
+sc_1 <- function(x) {(0.1904*x)+58.823}
+sc_2 <- function(x) {(0.1427*x)+99.075}
+
+# Convert the saved concentrations (using two different curves) to using
+# one or the other
+conc <- 
+  conc  %>% 
+  mutate_cond(Date == "20190607", CH4_CONC = sc_1(CH4_AREA))
+# Confirmed that the weirdly high fluxes in Passage 2, Week 1 were due to 
+# using two different standard curves
+
 ```
@@
 methane fluxes. From the `r n1` jars, we selected the top 4 
-(jars `r top_jars_gen1`) to homogenize and transfer to the next set of ecosystems.
+(jars `r top_jars_pas1`) to homogenize and transfer to the next set of ecosystems.
 
@@
 least one more set of ecosystems through a transfer with the same inoculum size
-(10%).
+(10%) using jars `r top_jars_pas2` from passage 2.
 
@@
 
-conc$selected[conc$JAR %in% top_jars_gen1 & conc$transfer == '1'] <- T
-conc$selected[!(conc$JAR %in% top_jars_gen1) | conc$transfer != '1'] <- F
+conc$selected[conc$JAR %in% top_jars_pas1 & conc$transfer == '1'] <- T
+conc$selected[conc$JAR %in% top_jars_pas2 & conc$transfer == '2'] <- T
+conc$selected[!(conc$JAR %in% top_jars_pas1) & conc$transfer == '1'] <- F
+conc$selected[!(conc$JAR %in% top_jars_pas2) & conc$transfer == '2'] <- F
 ```
@@
 filter(term == 'days') %>% 
-select(transfer, Date, JAR, selected, lm_umold = estimate, glm_umold = estimate1)
+select(transfer, Date, JAR, selected, lm_umold = estimate, glm_umold = estimate1)  %>% 
+mutate(glm_umold_exp = exp(glm_umold))
 
@@
 offspr2 <- flux$lm_umold[flux$Date == '20190614']
-sel_par <- flux$lm_umold[flux$transfer == '1' & flux$selected == T]
+sel_par1 <- flux$lm_umold[flux$transfer == '1' & flux$selected == T]
+sel_par2 <- flux$lm_umold[flux$Date == '20190614' & flux$selected == T]
 
@@
 offspr2_log <- flux$glm_umold[flux$Date == '20190614']
-sel_par_log <- flux$glm_umold[flux$transfer == '1' & flux$selected == T]
+sel_par1_log <- flux$glm_umold[flux$transfer == '1' & flux$selected == T]
+sel_par2_log <- flux$glm_umold[flux$Date == '20190614' & flux$selected == T]
 
+all_par_exp <- flux$glm_umold_exp[flux$Date == '20190527']
+offspr1_exp <- flux$glm_umold_exp[flux$Date == '20190607']
+offspr2_exp <- flux$glm_umold_exp[flux$Date == '20190614']
+sel_par1_exp <- flux$glm_umold_exp[flux$transfer == '1' & flux$selected == T]
+sel_par2_exp <- flux$glm_umold_exp[flux$Date == '20190614' & flux$selected == T]
 ```
@@
   geom_histogram(aes(fill = interaction(selected, Date))) +
-  facet_grid(Date ~ .) +
-  scale_fill_manual(values = c('black', 'red', 'blue', 'darkblue')) +
-  geom_vline(xintercept = mean(all_par)) +
-  geom_vline(xintercept = mean(offspr1), color = 'blue') +
-  geom_vline(xintercept = mean(offspr2), color = 'darkblue') +
-  geom_vline(xintercept = mean(sel_par), color = 'red') +
+  facet_grid(Date + transfer ~ .) +
+  scale_fill_manual(values = c('darkred', 'red', 'darkblue', 'blue', 'darkorange', 'orange')) +
+  geom_vline(xintercept = mean(all_par), color = 'darkred') +
+  geom_vline(xintercept = mean(offspr1), color = 'darkblue') +
+  geom_vline(xintercept = mean(offspr2), color = 'darkorange') +
+  geom_vline(xintercept = mean(sel_par1), color = 'red') +
+  geom_vline(xintercept = mean(sel_par2), color = 'orange') +
   labs(x = expression (CH[4] ~ "Flux" ~ (mu * "Mol"* ~ d^{-1})), y = "Number of Jars", title = "LM Fluxes")
@@
 ggplot(flux, aes(x = Date, y = lm_umold)) +
+  facet_wrap(~ transfer + Date, scales = 'free_x') + 
   geom_boxplot(outlier.shape = NA) +
   geom_jitter(aes(color = interaction(selected, Date))) +
-  scale_color_manual(values = c('black', 'red', 'blue', 'darkblue')) + 
-  geom_hline(yintercept = mean(all_par)) +
-  geom_hline(yintercept = mean(offspr1), color = 'blue') +
-  geom_hline(yintercept = mean(offspr2), color = 'darkblue') +
-  geom_hline(yintercept = mean(sel_par), color = 'red')  + 
+  scale_color_manual(values = c('darkred', 'red', 'darkblue', 'blue', 'darkorange', 'orange')) + 
+  geom_hline(yintercept = mean(all_par), color = 'darkred') +
+  geom_hline(yintercept = mean(offspr1), color = 'darkblue') +
+  geom_hline(yintercept = mean(offspr2), color = 'darkorange') +
+  geom_hline(yintercept = mean(sel_par1), color = 'red')  + 
+  geom_hline(yintercept = mean(sel_par2), color = 'orange')  + 
   labs(y = expression (CH[4] ~ "Flux" ~ (mu * "Mol"* ~ d^{-1})), x = "Measurement Date", title = "LM Fluxes")
@@
   geom_histogram(aes(fill = interaction(selected, Date))) +
-  facet_grid(Date ~ .) +
-  scale_fill_manual(values = c('black', 'red', 'blue', 'darkblue')) +
-  geom_vline(xintercept = mean(all_par_log)) +
-  geom_vline(xintercept = mean(offspr1_log), color = 'blue') +
-  geom_vline(xintercept = mean(offspr2_log), color = 'darkblue') +
-  geom_vline(xintercept = mean(sel_par_log), color = 'red') +
+  facet_grid(Date + transfer ~ .) +
+  scale_fill_manual(values = c('darkred', 'red', 'darkblue', 'blue', 'darkorange', 'orange')) +
+  geom_vline(xintercept = mean(all_par_log), color = 'darkred') +
+  geom_vline(xintercept = mean(offspr1_log), color = 'darkblue') +
+  geom_vline(xintercept = mean(offspr2_log), color = 'darkorange') +
+  geom_vline(xintercept = mean(sel_par1_log), color = 'red') +
+  geom_vline(xintercept = mean(sel_par2_log), color = 'orange') +
   labs(x = expression (CH[4] ~ "Flux" ~ (mu * "Mol"* ~ d^{-1})), y = "Number of Jars", title = "GLM Fluxes")
@@
 ggplot(flux, aes(x = Date, y = glm_umold)) +
+  facet_wrap(~ transfer + Date, scales = 'free_x') + 
   geom_boxplot(outlier.shape = NA) +
   geom_jitter(aes(color = interaction(selected, Date))) +
-  scale_color_manual(values = c('black', 'red', 'blue', 'darkblue')) + 
-  geom_hline(yintercept = mean(all_par_log)) +
-  geom_hline(yintercept = mean(offspr1_log), color = 'blue') +
-  geom_hline(yintercept = mean(offspr2_log), color = 'darkblue') +
-  geom_hline(yintercept = mean(sel_par_log), color = 'red') +
+  scale_color_manual(values = c('darkred', 'red', 'darkblue', 'blue', 'darkorange', 'orange')) + 
+  geom_hline(yintercept = mean(all_par_log), color = 'darkred') +
+  geom_hline(yintercept = mean(offspr1_log), color = 'darkblue') +
+  geom_hline(yintercept = mean(offspr2_log), color = 'darkorange') +
+  geom_hline(yintercept = mean(sel_par1_log), color = 'red') +
+  geom_hline(yintercept = mean(sel_par2_log), color = 'orange') +
+  labs(y = expression (CH[4] ~ "Flux" ~ (mu * "Mol"* ~ d^{-1})), x = "Measurement Date", title = "GLM Fluxes")
+
+ggplot(flux, aes(x = glm_umold_exp)) +
+  geom_histogram(aes(fill = interaction(selected, Date))) +
+  facet_grid(Date + transfer ~ .) +
+  scale_fill_manual(values = c('darkred', 'red', 'darkblue', 'blue', 'darkorange', 'orange')) +
+  geom_vline(xintercept = mean(all_par_exp), color = 'darkred') +
+  geom_vline(xintercept = mean(offspr1_exp), color = 'darkblue') +
+  geom_vline(xintercept = mean(offspr2_exp), color = 'darkorange') +
+  geom_vline(xintercept = mean(sel_par1_exp), color = 'red') +
+  geom_vline(xintercept = mean(sel_par2_exp), color = 'orange') +
+  labs(x = expression (CH[4] ~ "Flux" ~ (mu * "Mol"* ~ d^{-1})), y = "Number of Jars", title = "GLM Fluxes")
+
+ggplot(flux, aes(x = Date, y = glm_umold_exp)) +
+  facet_wrap(~ transfer + Date, scales = 'free_x') + 
+  geom_boxplot(outlier.shape = NA) +
+  geom_jitter(aes(color = interaction(selected, Date))) +
+  scale_color_manual(values = c('darkred', 'red', 'darkblue', 'blue', 'darkorange', 'orange')) + 
+  geom_hline(yintercept = mean(all_par_exp), color = 'darkred') +
+  geom_hline(yintercept = mean(offspr1_exp), color = 'darkblue') +
+  geom_hline(yintercept = mean(offspr2_exp), color = 'darkorange') +
+  geom_hline(yintercept = mean(sel_par1_exp), color = 'red') +
+  geom_hline(yintercept = mean(sel_par2_exp), color = 'orange') +
   labs(y = expression (CH[4] ~ "Flux" ~ (mu * "Mol"* ~ d^{-1})), x = "Measurement Date", title = "GLM Fluxes")
@@
 
-nsh1 <- ns_heritability(all_par, offspr1, sel_par) 
-nsh2 <- ns_heritability(all_par, offspr2, sel_par) 
-nsh1_log <- ns_heritability(all_par_log, offspr1_log, sel_par_log)
-nsh2_log <- ns_heritability(all_par_log, offspr2_log, sel_par_log)
+nsh1 <- ns_heritability(all_par, offspr1, sel_par1) 
+nsh2 <- ns_heritability(all_par, offspr2, sel_par1) 
+nsh1_log <- ns_heritability(all_par_log, offspr1_log, sel_par1_log)
+nsh2_log <- ns_heritability(all_par_log, offspr2_log, sel_par1_log)
+nsh1_exp <- ns_heritability(all_par_exp, offspr1_exp, sel_par1_exp)
+nsh2_exp <- ns_heritability(all_par_exp, offspr2_exp, sel_par1_exp)
 ```
@@
 of ecosystems (week 2), we calculated narrow-sense heritability as $h^2 = R / S$, which
-was `r myround(nsh2)` for the linear model and `r myround(nsh2_log)` for the exponential
-model. Granted, this is also not a valid estimate, because of the difference
+was `r myround(nsh2, 2)` for the linear model and `r myround(nsh2_log, 2)` for the exponential
+model and `r myround(nsh2_exp, 2)` for the back-transformed exponential coefficient. 
+Granted, this is also not a valid estimate, because of the difference
 in inoculum size between transfer 1 and 2. We will calculate this again between
@@
   kable_styling()
+flux %>% 
+  group_by(Date) %>% 
+  top_n(-4, glm_umold_exp) %>% 
+  kable(caption = "Generalized linear model link = log fluxes. Same as above except
+        JAR indicates top 4 jars from a glm and instead of slopes it's the exp(coef)
+        so it's a multiplicative factor.") %>% 
+  kable_styling()
 ```

### 2019-09-18 clean.R

@@ -1,0 +1,20 @@
+library(tidyverse)
+library(lubridate)
+
+time <- read.csv('eg_time_data.csv')
+
+data_t0 <-
+  data.frame(flux_date = unique(time$flux_date),
+             t0 = c('2019-05-27 11:51:00', '2019-06-07 10:41:00', '2019-06-14 14:33:00'))
+
+as.days <- function(start, end) {as.numeric(interval(start, end))/60/60/24}
+time_data  <- 
+time %>% 
+  spread(t, time) %>% 
+  left_join(data_t0, by = 'flux_date') %>% 
+  select(flux_date:jar, t0, t1:t4)
+
+conc <- read.csv('eg_conc_data.csv')
+conc %>% 
+  select(-co2_area) %>% 
+  spread(t, ch4_area)

### 2019-11-01 R/process.Rmd
Save before switching to first-order decay

--- 2019-09-18 Documents/oxidation_protocol.Rmd
@@
 library(lubridate)
+library(morris)
 ```
@@
 ```{r sc_data, fig.height=3}
-example_sc <- data.frame(
-  flux_date = c('2019-05-27', '2019-05-27', '2019-05-27', '2019-06-07', '2019-06-07', '2019-06-07', '2019-06-14', '2019-06-14', '2019-06-14'),
-  standard_ppm = c(999, 999, 999, 999, 999, 999, 999, 999, 999),
-  injection_ml = c(0.1, 0.5, 1, 0.1, 0.5, 1, 0.1, 0.5, 1),
-  ch4_area = c(556.03, 3046.03, 5995.73, 493.256, 2720.52, 5259.6, 569.3, 3255.82, 6129.72))
-example_sc <- 
-  example_sc %>% 
-  mutate(injection_ppm = standard_ppm * injection_ml) %>% 
-  select(flux_date:injection_ml, injection_ppm, ch4_area)
-example_sc %>% 
-  kable() %>% 
-  kable_styling()
-ggplot(example_sc, aes(ch4_area, injection_ppm)) +
-  facet_wrap(~ flux_date) + 
+
+sc <- read.csv('../Data/standard_curve.csv')
+
+ggplot(sc, aes(area, injection_ppm)) +
+  facet_wrap(~ flux_date + molecule, scales = "free") + 
   geom_point() +
-  geom_line() +
-  coord_fixed(6)
+  stat_smooth(method = 'lm', se = FALSE)
 
-sc_model <- function(df) {lm(injection_ppm ~ ch4_area, data = df)}
+sc_model <- function(df) {lm(injection_ppm ~ area, data = df)}
 standard_curves <- 
-  example_sc %>% 
-  group_by(flux_date) %>% 
+  sc %>% 
+  group_by(flux_date, molecule) %>% 
   nest() %>% 
@@
   arrange(r.squared) %>% 
-  select(flux_date, r.squared)
+  select(flux_date, molecule, r.squared)
 sc_equation <- 
@@
   unnest(tidy, .drop = TRUE) %>%
-  select(flux_date, term, estimate) %>% 
+  select(flux_date, molecule, term, estimate) %>% 
   spread(term, estimate) %>% 
-  rename(intercept = `(Intercept)`, slope = ch4_area)
-sc_ch4 <- left_join(sc_r2, sc_equation, by = "flux_date")
+  rename(intercept = `(Intercept)`, slope = area)
+sc_ch4 <- left_join(sc_r2, sc_equation, by = c("flux_date", "molecule"))
 sc_ch4 %>% 
@@
   kable_styling()
+
 ```
@@
 
-time_data <- data.table::fread('../Data/eg_time_data.csv')
+time_data <- data.table::fread('../Data/time_data.csv')
 
-time_data[, c('t0', 't1', 't2', 't3', 't4')] <- 
-  lapply(time_data[, c('t0', 't1', 't2', 't3', 't4')],
+time_data[, c('t0', 't1', 't2', 't3', 't4', 't5')] <- 
+  lapply(time_data[, c('t0', 't1', 't2', 't3', 't4', 't5')],
          ymd_hms)
@@
 
-time_data[, c('t1', 't2', 't3', 't4')] <- 
-  lapply(time_data[, c('t1', 't2', 't3', 't4')],
+time_data[, c('t1', 't2', 't3', 't4', 't5')] <- 
+  lapply(time_data[, c('t1', 't2', 't3', 't4', 't5')],
          function(x) as.days(time_data$t0, x))
@@
   select(-t0) %>% 
-  gather(t, days, t1:t4)
+  gather(t, days, t1:t5)
 
@@
 
-conc_data <- data.table::fread('../Data/eg_conc_data.csv')
+conc_data <- data.table::fread('../Data/conc_data.csv')
 
@@
   conc_data %>% 
-  gather(t, ch4_area, t1:t4) %>% 
-  left_join(sc_ch4, by = 'flux_date') %>% 
-  mutate(ch4_ppm = ch4_area * slope + intercept) %>% 
-  select(flux_date, jar, t, ch4_ppm) %>% 
+  gather(t, area, t1:t5) %>% 
+  left_join(sc_ch4, by = c('flux_date', 'molecule')) %>% 
+  mutate(ppm = area * slope + intercept) %>% 
+  select(flux_date, jar, molecule, t, ppm)  %>% 
   left_join(time_data, by = c('flux_date', 'jar', 't'))
@@
 
-```{r plot_fluxes, fig.height=20}
+```{r plot_fluxes}
 flux_data <- 
   flux_data %>% 
-  mutate(flux_date = ymd(flux_date), 
+  mutate(flux_date = factor(ymd(flux_date)), 
          jar = factor(jar),
-         t = factor(t)) %>% 
-  filter(ch4_ppm < 3000)
+         molecule = factor(molecule),
+         t = factor(t))
+
+flux_data <- 
+  flux_data %>% 
+  mutate(rep = factor(substr(jar, 2, 3)),
+         treat = factor(substr(jar, 1, 1)))
+
+```
+
+# P1: CO2
+
+```{r, fig.height = 30, warning = FALSE}
 
 flux_data %>% 
-  ggplot(aes(days, ch4_ppm)) + 
+  filter(flux_date == '2019-09-24') %>% 
+  filter(molecule == 'co2') %>% 
+  ggplot(aes(days, ppm)) + 
+  facet_wrap(c('treat', 'rep'), scales = "free", ncol = 2) +
+  geom_point() + 
+  stat_smooth(method = "nls", formula = y ~ SSasymp(x, Asym, R0, lrc), se = FALSE) +
+  expand_limits(y = 0)
+
+```
+
+# P1: CH4
+
+```{r, fig.height = 30, warning = FALSE}
+
+flux_data %>% 
+  filter(flux_date == '2019-09-24') %>% 
+  filter(molecule == 'ch4') %>% 
+  ggplot(aes(days, ppm)) + 
+  facet_wrap(c('treat', 'rep'), scales = "free", ncol = 2) +
+  geom_point() + 
+  stat_smooth(method = "nls", formula = y ~ SSasymp(x, Asym, R0, lrc), se = FALSE) +
+  expand_limits(y = 0)
+
+```
+
+# P2: CO2
+
+```{r, fig.height = 30, warning = FALSE}
+
+flux_data %>% 
+  filter(flux_date == '2019-10-23') %>% 
+  filter(molecule == 'co2') %>% 
+  ggplot(aes(days, ppm)) + 
+  facet_wrap(c('treat', 'rep'), scales = "free", ncol = 2) +
+  geom_point() + 
+  stat_smooth(method = "nls", formula = y ~ SSasymp(x, Asym, R0, lrc), se = FALSE) +
+  expand_limits(y = 0)
+
+```
+
+# P2: CH4
+
+```{r, fig.height = 30, warning = FALSE}
+
+flux_data %>% 
+  filter(flux_date == '2019-10-23') %>% 
+  filter(molecule == 'ch4') %>% 
+  ggplot(aes(days, ppm)) + 
+  facet_wrap(c('treat', 'rep'), scales = "free", ncol = 2) +
+  geom_point() + 
+  stat_smooth(method = "nls", formula = y ~ SSasymp(x, Asym, R0, lrc), se = FALSE)
+
+```
+
+# P2: CH4 baseline 0
+
+```{r, fig.height = 30, warning = FALSE}
+
+flux_data %>% 
+  filter(flux_date == '2019-10-23') %>% 
+  filter(molecule == 'ch4') %>% 
+  ggplot(aes(days, ppm)) + 
+  facet_wrap(c('treat', 'rep'), scales = "free", ncol = 2) +
+  geom_point() + 
+  stat_smooth(method = "nls", formula = y ~ SSasymp(x, Asym, R0, lrc), se = FALSE) +
+  expand_limits(y = 0)
+
+```
+
+```{r flux_model}
+lin_model <- function(dt) {
+  lm(ppm ~ days, data = dt)
+}
+exp_model <- function(dt) {
+  t <- try(nls(ppm ~ SSasymp(days, Asym, R0, lrc), data = dt))
+  if(inherits(t, "try-error")) return('foo') else t
+}
+
+nested <- 
+flux_data %>% 
+  group_by(flux_date, jar, molecule, treat, rep) %>% 
+  nest() %>% 
+  mutate(model = map(data, exp_model))
+
+lin_nested <- 
+flux_data %>% 
+  group_by(flux_date, jar, molecule, treat, rep) %>% 
+  nest() %>% 
+  mutate(model = map(data, lin_model))
+
+fluxes <-
+  nested %>% 
+  mutate(class = map(model, class)) %>% 
+  unnest(class) %>% 
+  filter(class == 'nls') %>% 
+  mutate(tidy = map(model, tidy)) %>% 
+  unnest(tidy) %>% 
+  filter(term == 'lrc') %>% 
+  filter(molecule == 'ch4') %>% 
+  mutate(lambda = exp(estimate)) %>% 
+  group_by(flux_date) %>% 
+  arrange(lambda, .by_group = TRUE) %>% 
+  select(-data, -model, -class, -term)
+
+fluxes %>% 
+  kable() %>% 
+  kable_styling()
+
+lin_fluxes <- 
+lin_nested %>% 
+  mutate(tidy = map(model, tidy)) %>% 
+  unnest(tidy) %>% 
+  filter(term == 'days')  %>% 
+  group_by(flux_date) %>% 
+  arrange(estimate, p.value, .by_group=TRUE) %>% 
+  select(-data, -model)
+
+lin_fluxes %>% 
+  kable() %>% 
+  kable_styling()
+```
+
+```{r test_flux, eval = FALSE}
+
+summary(aov(lambda ~ treat + flux_date, data = fluxes))
+flux_data
+nls(ppm ~ SSasymp(days, Asym, R0, lrc), data = flux_data)
+nested
+fluxes %>% 
+  ggplot(aes(x = lambda, fill = treat)) +
+  facet_wrap(~ flux_date) +
+  geom_histogram(binwidth = 1, alpha = 0.5, position = 'identity')
+
+lin_fluxes %>% 
+  filter(molecule == 'ch4') %>% 
+  ggplot(aes(x = estimate, fill = treat)) +
+  facet_wrap(~ flux_date) +
+  geom_histogram(alpha = 0.5, position = 'identity')
+
+```
+
+```{r plot_resids, fig.height=20, eval=FALSE}
+
+by_date_jar %>% 
+  mutate(model = map(data, flux_model)) %>% 
+  mutate(resids = map2(data, model, add_residuals)) %>% 
+  unnest(resids) %>% 
+  ggplot(aes(days, abs(resid), group = interaction(flux_date, jar))) +
   facet_grid(jar ~ flux_date, scales = "free") +
@@
   geom_line()
-```
 
-### Calculate Flux
-
-```{r calc_flux}
+by_date_jar %>% 
+  unnest(data) %>% 
+  summarize(length(t)) %>% 
+  print(n = 100)
 
@@
 
-```{r conversion}
+```{r conversion, eval = FALSE}
 
+all_data <- read.csv('../Data/all_data.csv')
+head(all_data)
+all_data %>% 
+  gather(t, area, t1:t5) %>% 
+  gather(t2, time, `t1.1`:`t5.1`) %>% 
+  head()
+tail(all_data)
+flux_data %>% 
+  arrange(flux_date, molecule, jar, t) %>% 
+write.csv(., '../Data/all_data.csv')
 ```

### 2020-05-21 Manuscript/manuscript.Rmd
Input data, update analysis, begin manuscript

---
title: "Artificial ecosystem selection on soil methane oxidation reveals simple/complex mapping between microbial community structure and ecosystem function"
author: "Andrew H. Morris and Brendan J. M. Bohannan"
output:
  html_document: default
  word_document:
    reference_docx: reference.docx
---

```{r options, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
library(tidyverse)
library(captioner)
library(broom)
library(broman)
figures <- captioner()
figures("selection", "Rank CH~4~ oxidation rate for each jar in each passage.")
figures("herit", "Correlation between mid-parent CH~4~ flux and mid-offspring CH~4~ flux.")

```

# Abstract

Questions:

- What is the mapping between microbial community structure and ecosystem function?

- Is it simple? e.g., abundance of functional group, diversity of functional group (like in plants). Or is it complex? e.g., multiple functional groups, multiple taxa.

- How does it compare to macroogranisms? Have people even tested functional group diversity extensively? Is methanotrophy as deeply conserved in bacteria as photosynthesis in plants? i.e., is there a 16S sequence or other marker that maps to the methanotrophy guild? It seems like microbial BEF has focused  on Prokaryote diversity or functional group abundance, which are both different from how it's been carried out in plants.

- Does BEF theory apply? Niche theory: One limiting R -> one final competitor. Multiple Rs -> multiple peristent competitors.

# Introduction

Biodiversity-ecosystem function (BEF) research has the central goal of understanding whether and how biodiversity - the diversity and relative abundance of species in a community - determines the magnitude of ecosystem functions (cite Hooper, Naeem, Tilman, ...). Ecosystem functions are a broad range of processes, including pollination, water filtration, carbon storage, and nutrient cycling, that are driven by both abiotic and biotic factors. The ultimate goal of this research is to predict how changes in biodiversity due to climate change or land use change will alter ecosystem function and, in turn, understand how we should protect biodiversity in order to preserve essential ecosystem functions. The fundamental question for this research is what is the nature of the "map" between elements of biodiversity and elements of ecosystem function?

One could hypothesize a relatively simple map in which a single taxon determines the rate of an ecosystem function based entirely on its presence or abundance. (e.g. of this?) Alternatively, instead of defining biodiversity based on a taxonomic unit, biodiversity could be defined by functional groups or guilds. For example, photosynthesis could be determined by the abundace of photosynthesizers in a community. (Example of abundance of photosynthesizers?  measured as the Abundance of rubisco genes?). A somewhat more complex map could define biodiversity based on the diversity (i.e., richness or number of species) of photosynthesizers. Using this mapping, an abundance of literature has demonstrated that productivity increases asymptotically and variation in productivity decreases asymptotically with the richness of producers (Tilman 1997).

When microbial ecologists have attempted to define the mapping of ecosystem functions performed by microorganisms, these simple models rarely hold (Rocca et al., Graham ete al.). Most functions in most ecosystems do not exhibit a positive relationship between the abundance of functional marker genes and the rate of the associated function. Few studies test the diversity of functional groups. This could mean that the map is simple, but we are not accurately estimating the abundance of these functional groups. Alternatively, the mapping between microbial community structure and microbial processes could be more complex than the mapping between macrobial structure and ecosystem function.

To investigate the nature of the map between microbial community structure and ecosystem function, we will perform artificial ecosystem selection (*sensu* Swenson et al. 2000). With this approach, we will collect a single soil which we subsamble to generate multiple ecosystems. These ecosystems vary in both community structure and the rate of ecosystem function. From this variation, we select the top performing ecosystems, reshuffle their members through homogenization, and subsample this homogenate to generate more ecosystems. Through this iterative process, we test many different combinations of taxa and enrich for the top performers. This process of selection also reduces the genetic variation and diversity of the communities. By minimizing the environmental differences between jars and between passages, we can insure that the majority of functional variation from jar to jar and passage to passage is due to genetic variation in the community. As a result, genetic differences between the initial community and the selected community should be associated with the function of interest.

We hypothesize that the positive selection treatment will increase in methane oxidation rate while the neutral selection treatment will resemble a random walk over time demonstrating community heritability within our system. We also hypothesize that genetic changes over time will follow a predictable trajectory: the fastest changes in community copmosition will be due to species loss, slower changes will be due to changes in the relative abundance of species, and the slowest changes will be due to intraspecies selection (i.e., evolution). The soil communities will not only be adapting in response to selection on soil methane oxidation, but will also be adapting to survival and passaging between jars.

There are several alternatives as to the complexity of the "map" between microbes and methane. At the simplest level, methane oxidation rate could be a function of the abundance of methane oxidizers. A somewhat more complex map would be if methane flux is a function of the diversity of methane oxidizers as in other systems (Hooper et al.). If neither of these simple models hold, the mapping might be a more complex relationship between a broader level of organization, such as methane producers and consumers, a broader metabolic network, or a network of bacterial and archaeal interactions that's hard to predict.

The gold standard of molecular biology gene mapping studies is to not only identify a correlation, but also to confirm the association through knockout and overexpression studies. To approximate this approach for microbial ecology, we cultured the important strain(s) in pure culture and applied them to the neutral soil community to see if we could recapitulate the trait level for the selected treatment with only 1 or a few taxa or if it required the whole community. We used the neutral treatment, because these communities are already bottle-adapted and offer a more similar environment for the selected strains than a natural soil would.

Our goal is to understand the identity of the community members and their interactions that result in the greatest level of ecosystem function. Artificial ecosystem selection allows us to test many "teams" or communities of microbes in terms of their combined impact on function. By performing selection, we reduce the genetic variation in our natural community to the variation that is sufficient to perform a high rate of ecosystem function. Then, we can investigate the remaining variation to generate hypotheses about the relationship between community composition and ecosystem function, which mebmers are the best performers, and how simple or complex the map is between community and funtion. Finally, assuming we can culture these members, we can test those hypotheses through enrichment of one or a group of taxa to see which members are sufficient to recreate the highest rate of ecosystem function.

Questions:

1. What is the “map”? And how do we get to that map? What is the nature of the map? Taxonomy? Or something other than taxonomy? Simple – taxon. Diffuse – something other than taxonomy. Complicated interactions among organisms leads to “hitchhikers.”

2. Can we isolate 1 or a few strains and add them to the N community? Or does it take a whole complex community? The

3. Lewontin posited that higher levels of selection (species, community, ecosystem) would have weaker heritability. Do we observe this?

1. Does soil methane oxidation rate respond to selection at the ecosystem level?

2. Does a community that has been selected on for high methane oxidation rate differ in diversity or composition from a community that has been passaged without selection for methane oxidation rate?

3. What aspects of community composition differ between the positive and neutral lines? Changes in alpha diversity, beta diversity? Presence/absence of specific microbial taxa? Relative abundance of microbial taxa?

4. Does selection lead to greater differences in the bacterial/archaeal community (inferred from 16S sequences) or in the methanotroph community (inferred from pmoA sequences)?

5. Does a community with high methane oxidation rate dominate coalescence with a natural community compared to a community with average methane oxidation rate?

8. Many-to-one mapping

9. Crossing/mixing lines or competing lines or applying to natural soils for the application of the tool.

Hypotheses:

1. Soil methane oxidation rate responds to selection at the ecosystem level

2.

Questions to address in the Intro:

Is this “just enrichment” or is it community selection? i.e., is it dilution to extinction?

How do people enrich for methanotrophs?

Community-level selection? Paul Rainey, Martin Polz

# Materials & Methods

## Selection experiment
We performed an artificial ecosystem selection experiment similar to Swenson et al. (2000) by passaging replicate soil microbial communities. The trait we selected on was CH~4~ oxidation rate. Our experiment had two selection lines with twelve jars each for a total of twenty-four jars. One line was a positive selection line where the two or three jars with the highest CH~4~ oxidation rate were chosen to inoculate the next set of jars. The other line was a neutral selection where an equal number of jars as the positive line were chosen at random among the twelve. The number of jars chosen was based on the distribution of fluxes among the positive jars, i.e., we chose the top three jars unless only two jars had considerably greater oxidation rates based on visual inspection of histograms.

The initial soil microbial community was sampled from the top 10 cm of an upland mineral soil under a deciduous forest ecosystem near the University of Oregon campus. Incubations were performed in 500 mL mason jars (Ball Corporation, Broomfield, Colorado, USA) with rubber septa installed in the lids. Jars were incubated at ambient temperature on the benchtop. To create the first set of jars, twenty-four jars were created in an identical manner and then randomly assigned to either the positive or negative selection line. Each jar was sterilized with 70% ethanol to which was added 45 g of autoclaved potting mix, 5 g of live soil, and 3.5 mL of sterile deionized water to bring them to <mark> 60%??? of field capacity </mark>. The jars were then homogenized, capped, and injected with 4.3 mL of 99% CH~4~ to bring the headspace concentration to approximately 1000 ppm CH~4~. Jars were flushed and respiked twice per week to maintain elevated CH~4~ concentrations until travel, holidays, and COVID-19 got in the way and then we just did it as often as we could. Jars were incubated for approximately three weeks and then flux measurements were taken.

Once CH~4~ oxidation rates were determined, the top two or three jars were chosen for selection in the positive line and an equal number of jars were randomly chosen from the neutral line. The selected jars for each line were homogenized and used as the 5 g live soil for the subsequent passage. Twelve new jars were created for each line in the same manner as above using the inoculum for that line. This procedure continued until a divergence was observed in the mean flux of each line for two consecutive generations. Jars in passage number 2 and 5 were homogenized and a sample was collected and stored at -80$^\circ$C for later DNA extraction.

## Methane flux measurements
Methane oxidation rates were determined after flushing and spiking jars to approximately 1000 ppm CH~4~. Head-space samples of 1 mL were collected from each jar immediately after spiking and then at time points 3, 6, 24, and 48 hours for a 5-point curve. Samples were immediately injected into a <mark>???</mark> gas chromatograph fitted with an electron capture detector (company, city, state, USA) to determine the headspace CH~4~ concentration. Fluxes were calculated from a first-order exponential decay function as decay constant `k` with units<mark>???</mark>. Oxidation rates are presented as the additive inverse of `k` (i.e., `-k`) so that a more positive value represents a faster oxidation rate.

## Soil DNA extraction and sequencing

To reveal the mapping between soil microbial community structure and methane oxidation rate, we will estimate the abundance of methane oxidizers by quantifying pmoA, we will estimate the diversity of methane oxidizers by sequencing the pmoA gene, and we will estimate bacterial and archaeal diversity by sequencing the 16S gene.

Samples to extract:

- generation 2
- generation 5
- starting soil
- autoclaved potting mix negative control
- sterile water negative control

Soil DNA was extracted from a subsample of soil from the starting inoculum and from each jar in passage 2 and 5 after CH~4~ flux measurements. Negative controls were extracted from autoclaved potting mix and autoclaved, deinoized water. Extractions were performed using the DNeasy PowerSoil kit (QIAGEN, Hilden, Germany). <mark>Find example extraction and sequencing text.</mark> In order to estimate the diversity and relative abundance of the bacterial and archaeal community in our soil ecosystems, we sequenced the V3-V4 region of 16S rRNA gene using the Earth Microbiome Project protocol and primers<mark>???</mark>. In addition, to estimate the diversity of CH~4~ oxidizers, we sequenced the particulate methane monooxygenase subunit A gene (*pmoA*) using primers<mark>???</mark>.

## Bioinformatics
Bioinformatic analyses were conducted using the `QIIME2` bioinformatics platform (Bolyen et al. 2019). Raw sequencing reads were demultiplexed using ??? (citation), denoised using `DADA2` (citation). Taxonomic assignment for the 16S reads was performed using ??? based on the Silva database. All subsequence statistical analyses were performed in the `R` statistical programming environment (R Core Team 2020). Richness estimates were made using the `breakaway` package (Willis et al. ???).

## Statistical Analysis

### Response to selection

To test whether there was a significant change in CH~4~ oxidation rate (`k`) over multiple passages as a response to selection, we fit a multiple linear regression model using ordinary least squares:

$$
\hat{y}_i = \beta_0 + \beta_1 x_{1i} + \beta_2 x_{2i} + \beta_3 x_{1i} x_{2i}
$$

where $\hat{y}$ was the predicted CH~4~ oxidation rate in jar $i$ as a function of treatment $x_1$ (positive or neutral selection), passage number $x_2$, and the interaction between treatment and passage number $x_1 x_2$. We fit the model in `R` version `r paste(R.Version()$major, R.Version()$minor, sep = ".")` using the `lm` and `anova` functions (R Core Team 2020). Oxidation rates were highly skewed and residuals did not meet the assumptions of normality and heterogeneity of variance, therefore, fluxes were rank-transformed prior to analysis.

### Ecosystem heritability

We will estimate heritability of this trait as the correlation between parent and offspring jars.

To test the heritability of CH~4~ oxidation rate as an ecosystem property, we calculated narrow-sense heritability as the slope of the relationship between the mid-parent CH~4~ oxidation rate and the mid-offspring CH~4~ oxidation rate (Goodnight 2000). For example, the mid-parent of passage 1 was calculated as the mean of the three jars selected to inoculate passage 2 and the mid-offspring was calculated as the mean of all 12 jars produced from those parental jars. To estimate heritability, we fit the following model using ordinary least squares:

$$
\hat{y} = \beta_0 + \beta_1 x_{1}
$$

where $x$ was the parental mean, $\hat{y}$ was the offspring mean, and $\beta_{1}$ was heritability.

In addition, we calculated the inbreeding coefficient, $F$, as:

$$
F = 1 - \left(1 - \frac{1}{2N} \right) ^{T}
$$

where $N$ is the population size and $T$ is the number of generations (Goodnight 2000).

### Taxa differentially abundant in the positive selection treatment

To identify taxa that responded to selection on CH~4~ oxidation rate, we tested differential abundance using a beta-binomial model (Martin et al. 2020). We compared taxa in passage 5 between the positive and neutral treatment to identify taxa that

## Community over-expression assay

* Pick out specific crobes that respond to selection and are differentially abundant and try to isolate and culture them. Either on media or in soil??
* Add individual crobes back to a lab soil community to "overexpress" and see if the same "phenotype" is achieved. Compared that to an extraction of the full community. Use water as a control.
Should we inoculate them onto sterile potting mix or onto neutral potting mix or onto natural potting mix?

## Depth of trait conservation - What is the taxonomic depth or % 16S sequence similarity that best explains variation in methane oxidation rates?

# Results

```{r sel_model}
fluxes <- read_tsv('../Output/fluxes.tsv')
rank_fit <- lm(estimate_rank ~ treat * passage, data = fluxes)

sel_sum <- tidy(summary(rank_fit))
sel_aov <- tidy(anova(rank_fit))
```

Methane oxidation rates increased over the course of the experiment in the positive selection treatment and did not change in the neutral selection treatment (`r figures("selection", display="cite")`). This can be seen in the significant interaction between treatment and passage number for the rank of CH~4~ oxidation rate (F~`r sel_aov[3,2]`,`r sel_aov[4,2]`~ = `r myround(sel_aov[3,5], 2)`, p = `r myround(sel_aov[3,6], 3)`). While there was no difference in the overall mean of the two treatments (F~`r sel_aov[1,2]`,`r sel_aov[4,2]`~ = `r myround(sel_aov[1,5], 2)`, p = `r myround(sel_aov[1,6], 2)`), the mean rank CH~4~ oxidation rate increased by `r myround(sel_sum[4,2],1)`  per passage for the positive selection treatment (t = `r myround(sel_sum[4, 4], 2)`, p = `r myround(sel_sum[4, 5], 3)`), but the neutral treatment did not change ($\beta$ = `r myround(sel_sum[3,2],1)`, t = `r myround(sel_sum[3,4], 2)`, p = `r myround(sel_sum[3,5], 2)`).

```{r sel_fig, fig.width=5, fig.height=4, fig.cap=figures("selection")}
fluxes %>% 
  ggplot(mapping = aes(x = passage, y = estimate_rank, color = treat)) +
  geom_jitter(width = 0.2) +
  geom_smooth(method = 'lm') +
  labs(x = "Passage Number", y = "Rank Methane Oxidation Rate (-k)") +
  scale_color_discrete(name = "Selection\nTreatment", labels = c('Neutral', 'Positive')) +
  theme_bw()
```

```{r herit_model}
heritability <- read_tsv('../Output/heritability.tsv')
fit <- lm(offspring ~ parental, data = heritability)
her_sum <- tidy(summary(fit))
her_aov <- tidy(anova(fit))
```

Methane oxidation rate was heritable between consecutive generations (`r figures("herit", display="cite")`). Specifically, the narrow-sense heritability ($h^2$), calculated as the slope of the mid-parent and mid-offspring, was `r myround(her_sum[2, 2], 2)`.

```{r herit_fig, fig.width=5, fig.height=4, fig.cap=figures("herit")}
ggplot(heritability, aes(x = parental, y = offspring)) +
  geom_point(aes(color = treat)) +
  stat_smooth(method = 'lm') + 
  labs(x = "Parental Mean Flux (-k)", y = "Offspring Mean Flux (-k)") +
  scale_color_discrete(name = "Selection\nTreatment", labels = c('Neutral', 'Positive')) +
  theme_bw()
```

```{r inbreeding_coef}
N = 12 # population size
T = 5  # number of passages (generations)
F = 1 - (1 - 1 / (2 * N)) ^ T
F
```

# Discussion

# Acknowledgments
This project was supported by the National Science Foundation Graduate Research Fellowship Program (grant no. DGE 1255832) and the ARCS Foundation Florence and Mike Nudelman Scholarship.

# Competing Interests
We declare we have no competing interests.

# References

### 2020-05-21 R/process.Rmd
Input data, update analysis, begin manuscript

--- 2019-09-18 Documents/oxidation_protocol.Rmd
@@
 ---
-title: "Oxidation Protocol"
+title: "Artificial Ecosystem Selection: Methane Oxidation"
 author: "Andrew Morris"
-date: "9/9/2019"
 output:
@@
 library(lubridate)
+library(morris)
+library(magrittr)
 ```
-
-## General notes
-
-- Record everything in the lab notebook.
-- Take note of any mistakes or errors.
-- Ideally start flux measurements around 8:00 or 9:00 o'clock on Monday
-  morning to allow time for three measurements on that day.
-
-## Initial setup
-1. Sterilize potting mix, DI water, and mason jar disks in the autoclave.
-2. Sterilize jars with 70% ethanol.
-3. Add 45 g of steril potting mix, 5 g of live soil, and 3.5 mL of sterile DI water to each jar and cap.
-4. Spike each jar with 4.3 mL of 99% CH~4~.
-
-Allow these to incubate for at least two weeks before taking a flux
-measurements. Also, uncap them in the hood for ~30 minutes and respike with 4.3
-mL of 99% CH~4~ every 3-4 days to maintain aerobic conditions and maintain
-elevated CH~4~. So, for example, if the jars are spiked on Thursday, flush them
-on Monday and Thursday of the following two weeks and then take the flux
-measurements on the third Monday.
-
-## Flux sampling
-
-### Monday
-
-1. Uncap jars in hood with blower on to return them to atmospheric gas
-concentrations and to avoid cross-contamination.
-2. Measure standard curve on GC with 5-12 points for CH~4~. Record the
-   concentration of the standard in the bottle (ppm), the volume of the
-   injected standard (mL), the
-concentration of the injected standard (ppm), and the CH~4~ curve area. Also record the
-slope and y-intercept for the resulting equation as well as the R^2^. Continue running standard curves until the R^2^ is 99% and ideally 99.9%.
-3. Cap jars in the hood.
-4. Spike jars to ~1000 ppm CH~4~ using 0.43 mL of 99% CH~4~ and record time for
-   each jar to the minute (t~0~).
-5. Measure t~1~ immediately after spiking all jars resulting in a t~1~ that is
-20 to 30 minutes after t~0~. Measure jars in the order that they were spiked.
-Record the time, CH~4~ curve area, and CO~2~ curve area.
-6. Record t~2~ after 3-5 hours.
-7. Record t~3~ after 5-8 hours.
-
-### Tuesday
-
-8. Record t~4~ after 24 hours.
-
-### Wednesday
-
-9. Record t~5~ after 48 hours.
-10. Enter the standard curve and flux data into R. Visually inspect the graphs
-    to determine how many points to include in the linear portion of the curve
-    (the first 3 to 5 points) and calculate the linear decrease in CH~4~
-    concentration (Vmax). 
-11. Determine the top 2 (16.7%) or 3 (25%) of the ~12 replicate jars for the
-    positive selection treatment (p). Homogenize and reserve the soil from
-    those jars in the hood. Randomly select an equal number of jars from the
-    neutral selection treatment (n). Homogenize and reserve the soil from those
-    jars in the biosafety cabinet. Clearly label each of these containers.
-12. Empty the rest of the soil from the jars into an autoclave container, cover
-    it with aluminum foil, and label it "disposal". Clean the jars and lids
-    with soap and rinse and dry thoroughly.
-13. Add at least `r 45 * 24` grams of fresh potting mix to an autoclave tray
-    and cover with aluminum foil. Wrap the disks in a foil packet. Set aside
-    the rings, they do not need to be autoclaved.
-14. Autoclave the potting soil, lids, and disposal soil on the gravity cycle for 40
-    minutes. 
-15. Remove from the autoclave, put the jars and lids into the hood, and allow
-    to cool overnight.
-
-### Thursday
-
-16. Sterilize the inside of each jar with 70% ethanol.
-17. Add 45 grams of sterile potting mix to each jar.
-18. Add 5 grams of the selected soil homogenate to each 'p' jar and 5 grams of
-    the random soil to each 'n' jar. 
-19. Add 3.5 mL of sterile DI water to each jar. 
-20. Cap each jar and spike with 4.3 mL CH~4~.
-
-## Preferred data format
-
-Dates should be recorded as `yyyy-mm-dd` and times should be recorded as
-`yyyy-mm-dd hh:mm` (seconds are optional). Use standard column headers with no
-caps and no spaces. Instead, simply use "variable_units" format.
-
-Column Headers | Meaning
--------------- | ----------------------------------------------------------
-`flux_date`    | The date flux measurements began (i.e. Monday)
-`standard_ppm` | The concentration of the standard in the bottle. 
-`injection_ml` | The volume of standard injected into the GC. 
-`injectin_ppm` | The concentration of standard injected into the gc (standard_ppm multiplied by injection_ml) 
-`ch4_area`     | The area-under-the-curve for the methane peak output from the GC.
 
@@
 ```{r sc_data, fig.height=3}
-example_sc <- data.frame(
-  flux_date = c('2019-05-27', '2019-05-27', '2019-05-27', '2019-06-07', '2019-06-07', '2019-06-07', '2019-06-14', '2019-06-14', '2019-06-14'),
-  standard_ppm = c(999, 999, 999, 999, 999, 999, 999, 999, 999),
-  injection_ml = c(0.1, 0.5, 1, 0.1, 0.5, 1, 0.1, 0.5, 1),
-  ch4_area = c(556.03, 3046.03, 5995.73, 493.256, 2720.52, 5259.6, 569.3, 3255.82, 6129.72))
-example_sc <- 
-  example_sc %>% 
-  mutate(injection_ppm = standard_ppm * injection_ml) %>% 
-  select(flux_date:injection_ml, injection_ppm, ch4_area)
-example_sc %>% 
-  kable() %>% 
-  kable_styling()
-ggplot(example_sc, aes(ch4_area, injection_ppm)) +
-  facet_wrap(~ flux_date) + 
+
+sc <- read.csv('../Data/standard_curve.csv')
+
+sc %>% 
+  filter(molecule == 'ch4') %>% 
+ggplot(mapping = aes(area, injection_ppm)) +
+  facet_wrap(~ flux_date + molecule, scales = "free") + 
   geom_point() +
-  geom_line() +
-  coord_fixed(6)
+  stat_smooth(method = 'lm', se = FALSE)
 
-sc_model <- function(df) {lm(injection_ppm ~ ch4_area, data = df)}
+sc_model <- function(df) {lm(injection_ppm ~ area, data = df)}
 standard_curves <- 
-  example_sc %>% 
+  sc %>% 
+  filter(molecule == 'ch4') %>% 
   group_by(flux_date) %>% 
@@
   standard_curves %>% 
-  unnest(glance, .drop = TRUE) %>% 
-  arrange(r.squared) %>% 
+  unnest(glance) %>% 
   select(flux_date, r.squared)
@@
   standard_curves %>% 
-  unnest(tidy, .drop = TRUE) %>%
+  unnest(tidy) %>%
   select(flux_date, term, estimate) %>% 
   spread(term, estimate) %>% 
-  rename(intercept = `(Intercept)`, slope = ch4_area)
-sc_ch4 <- left_join(sc_r2, sc_equation, by = "flux_date")
+  rename(intercept = `(Intercept)`, slope = area)
+sc_ch4 <- left_join(sc_r2, sc_equation, by = c("flux_date"))
 sc_ch4 %>% 
@@
 
-time_data[, c('t1', 't2', 't3', 't4')] <- 
-  lapply(time_data[, c('t1', 't2', 't3', 't4')],
-         function(x) as.days(time_data$t0, x))
-
+time_data[, c('t1', 't2', 't3', 't4', 't5')] <- 
+  lapply(time_data[, c('t1', 't2', 't3', 't4', 't5')],
+         function(x) as.days(time_data$t1, x))
 time_data <- 
@@
   select(-t0) %>% 
-  gather(t, days, t1:t4)
+  gather(t, days, t1:t5) %>% 
+  mutate(flux_date = factor(flux_date))
 
@@
 
-```{r conc_data, warning=FALSE}
+```{r}
 
-conc_data <- data.table::fread('../Data/eg_conc_data.csv')
+conc_data <- data.table::fread('../Data/conc_data.csv')
+
+conc_data <- 
+  conc_data %>% 
+  filter(molecule == 'ch4') %>% 
+  mutate(flux_date = factor(flux_date)) %>% 
+  gather(t, area, t1:t5) %>% 
+  left_join(sc_ch4, by = c('flux_date')) %>% 
+  mutate(ppm = area * slope + intercept) %>% 
+  select(flux_date, jar, t, ppm) %>% 
+  spread(t, ppm)
+
+ln_ch0_chn <- function(t0, tn) {
+  log(t0/tn)
+}
+
+conc_data[, c('t1', 't2', 't3', 't4', 't5')] <- 
+  lapply(conc_data[, c('t1', 't2', 't3', 't4', 't5')],
+         function(x) ln_ch0_chn(conc_data$t1, x))
+flux_data <-
+conc_data %>% 
+  gather(t, ppm, t1:t5) %>% 
+  left_join(time_data, by = c('flux_date', 'jar', 't'))
 
 flux_data <- 
-  conc_data %>% 
-  gather(t, ch4_area, t1:t4) %>% 
-  left_join(sc_ch4, by = 'flux_date') %>% 
-  mutate(ch4_ppm = ch4_area * slope + intercept) %>% 
-  select(flux_date, jar, t, ch4_ppm) %>% 
-  left_join(time_data, by = c('flux_date', 'jar', 't'))
+  flux_data %>% 
+  mutate(flux_date = factor(ymd(flux_date)), 
+         jar = factor(jar),
+         t = factor(t)) %>% 
+  mutate(rep = factor(substr(jar, 2, 3)),
+         treat = factor(substr(jar, 1, 1)))
 
@@
 
-```{r plot_fluxes, fig.height=20}
-flux_data <- 
-  flux_data %>% 
-  mutate(flux_date = ymd(flux_date), 
-         jar = factor(jar),
-         t = factor(t)) %>% 
-  filter(ch4_ppm < 3000)
+# Flux
+
+```{r flux_curves, warning = FALSE, fig.height=8}
 
 flux_data %>% 
-  ggplot(aes(days, ch4_ppm)) + 
-  facet_grid(jar ~ flux_date, scales = "free") +
+  ggplot(aes(days, ppm)) + 
+  facet_grid(flux_date + treat ~ rep, scales = "free") +
   geom_point() + 
-  geom_line()
-```
-
-### Calculate Flux
-
-```{r calc_flux}
+  stat_smooth(method = 'lm')
 
@@
 
-### Convert ppm to $\mu$grams
+# Fluxes
 
-```{r conversion}
+```{r flux_model}
+lin_model <- function(dt) {
+  lm(ppm ~ days, data = dt)
+}
+
+nested <- 
+flux_data %>% 
+  group_by(flux_date, jar, treat, rep) %>% 
+  nest() %>% 
+  mutate(model = map(data, lin_model))
+
+fluxes <-
+  nested %>% 
+  mutate(tidy = map(model, tidy)) %>% 
+  unnest(tidy) %>% 
+  filter(term == 'days') %>% 
+  group_by(flux_date) %>% 
+  arrange(estimate, .by_group = TRUE) %>% 
+  select(-data, -model, -term) %>% 
+  ungroup() %>% 
+  mutate(passage = as.numeric(flux_date)) %>% 
+  mutate(estimate_rank = rank(estimate))
+
+write_tsv(fluxes, '../Output/fluxes.tsv')
 
@@
 
-### Selection
+```{r plot_estimates}
+ggplot(fluxes, aes(x = estimate, fill = treat)) +
+  facet_wrap(~ flux_date, scales = "free") +
+  geom_histogram(position = 'identity', alpha = 0.5, bins = 20)
+
+ggplot(fluxes, aes(y = estimate, x = treat, color = treat)) +
+  facet_wrap(~ flux_date, scales = "free") +
+  geom_boxplot(outlier.shape = NA) +
+  geom_label(aes(label = rep))
+```
+
+```{r test_response}
+
+fluxes$passage <- as.numeric(fluxes$flux_date)
+
+fit <- lm(estimate ~ treat * passage, data = fluxes)
+summary(fit)
+selection_table <- tidy(anova(fit))
+
+selection_table[1, 1] <- "Treatment"
+selection_table[2, 1] <- "Passage"
+selection_table[3, 1] <- "Treatment:Passage"
+selection_table[4, 1] <- "Residuals"
+
+selection_table %>% 
+  kable(col.names = c('Term', 'Df', 'Sum Sq', 'Mean Sq', 'F Value', 'p-value')) %>% 
+  kable_styling()
+
+```
+
+```{r plot_selection_data, eval=FALSE, include=FALSE}
+
+fluxes %>% 
+  ggplot(fluxes, mapping = aes(x = passage, y = estimate, group = passage)) +
+  geom_boxplot() +
+  geom_jitter(width = 0.2) +
+  labs(x = "Passage Number", y = "Methane Oxidation Rate (-k)") +
+  theme_bw()
+
+fluxes %>% 
+  ggplot(mapping = aes(x = treat, y = estimate)) +
+  geom_boxplot() +
+  geom_jitter(width = 0.2) +
+  labs(x = "Treatment", y = "Methane Oxidation Rate (-k)") +
+  theme_bw()
+
+library(rstanarm)
+
+post1 <- stan_glm(estimate ~ treat, data = fluxes, prior = cauchy(), seed = 12345)
+post2 <- update(post1, formula = . ~ passage)
+post3 <- update(post1, formula = . ~ treat + passage)
+(post4 <- update(post1, formula = . ~ treat * passage))
+
+base <-  ggplot(fluxes, mapping = aes(x = passage, y = estimate, group = passage)) +
+  geom_boxplot() +
+  geom_jitter(width = 0.2) +
+  labs(x = "Passage Number", y = "Methane Oxidation Rate (-k)") +
+  theme_bw()
+
+base + geom_abline(intercept = coef(post1)[1], slope = coef(post1)[2], 
+                   color = "skyblue4", size = 1)
+draws <- as.data.frame(post1)
+colnames(draws)[1:2] <- c("a", "b")
+
+base + 
+  geom_abline(data = draws, aes(intercept = a, slope = b), 
+              color = "skyblue", size = 0.2, alpha = 0.25) + 
+  geom_abline(intercept = coef(post1)[1], slope = coef(post1)[2], 
+              color = "skyblue4", size = 1)
+
+draws <- as.data.frame(as.matrix(post2))
+colnames(draws)[1:2] <- c("a", "b")
+ggplot(fluxes, mapping = aes(x = treat, y = estimate)) +
+  geom_boxplot() +
+  geom_jitter(width = 0.2) +
+  labs(x = "Treatment", y = "Methane Oxidation Rate (-k)") +
+  theme_bw() +
+  geom_abline(data = draws, aes(intercept = a, slope = b), 
+              color = "skyblue", size = 0.2, alpha = 0.25) + 
+  geom_abline(intercept = coef(post2)[1], slope = coef(post2)[2], 
+              color = "skyblue4", size = 1)
+
+reg0 <- function(x, ests) cbind(1, 0, x) %*% ests 
+reg1 <- function(x, ests) cbind(1, 1, x) %*% ests
+
+args <- list(ests = coef(post3))
+lgnd <- guide_legend(title = NULL)
+base2 <- ggplot(fluxes, aes(x = passage, fill = treat)) + 
+  geom_jitter(aes(y = estimate), shape = 21, stroke = .2, size = 1) + 
+  guides(color = lgnd, fill = lgnd) + 
+  theme(legend.position = "right")
+base2 + 
+  stat_function(fun = reg0, args = args, aes(color = "Neutral"), size = 1.5) +
+  stat_function(fun = reg1, args = args, aes(color = "Positive"), size = 1.5)
+
+reg0 <- function(x, ests) cbind(1, 0, x, 0 * x) %*% ests 
+reg1 <- function(x, ests) cbind(1, 1, x, 1 * x) %*% ests
+args <- list(ests = coef(post4))
+base2 +
+  stat_function(fun = reg0, args = args, aes(color = "Neutral"), size = 1.5) + 
+  stat_function(fun = reg1, args = args, aes(color = "Positive"), size = 1.5)
+
+loo1 <- loo(post1, cores = 2, k_threshold = 0.7)
+loo2 <- loo(post2, cores = 2, k_threshold = 0.7)
+loo3 <- loo(post3, cores = 2, k_threshold = 0.7)
+loo4 <- loo(post4, cores = 2, k_threshold = 0.7)
+(comp <- loo_compare(loo1, loo2, loo3, loo4))
+######################
+
+post2 <- stan_lmer(estimate ~ 1 + (1|treat) + (1|passage) + (1|treat:passage),
+                   data = fluxes, prior_intercept = cauchy(),
+                   prior_covariance = decov(shape = 2, scale = 2),
+                   adapt_delta = 0.999, seed = 12345)
+
+```
+
+```{r selection}
+
+fluxes %>% 
+  ggplot(mapping = aes(x = passage, y = estimate, color = treat)) +
+  geom_jitter(width = 0.2) +
+  geom_smooth(method = 'lm') +
+  labs(x = "Passage Number", y = "Methane Oxidation Rate (-k)") +
+  scale_color_discrete(name = "Selection Treatment", labels = c('Neutral', 'Positive')) +
+  theme_bw()
+
+#ggsave('selection.pdf', width = 6, height = 4)
+```
+
+```{r selection_log_plot}
+
+log_fluxes <- fluxes
+log_fluxes$estimate[log_fluxes$estimate < 0] <- 0.01
+
+fit <- lm(estimate ~ treat * passage, data = log_fluxes)
+
+selection_table <- tidy(anova(fit))
+
+#selection_table[1, 1] <- "Neutral0"
+#selection_table[2, 1] <- "Positive0"
+#selection_table[3, 1] <- "Passage#"
+#selection_table[4, 1] <- "Positive:Passage#"
+
+selection_table[1, 1] <- "Treatment"
+selection_table[2, 1] <- "Passage"
+selection_table[3, 1] <- "Treatment:Passage"
+selection_table[4, 1] <- "Residuals"
+
+selection_table %>% 
+  kable(col.names = c('Term', 'Df', 'Sum Sq', 'Mean Sq', 'F Value', 'p-value')) %>% 
+  kable_styling()
+
+log_fluxes %>% 
+  ggplot(mapping = aes(x = passage, y = estimate, color = treat)) +
+  geom_jitter(width = 0.2) +
+  geom_smooth(method = 'lm') +
+  labs(x = "Passage Number", y = "log10(Methane Oxidation Rate (-k))") +
+  scale_color_discrete(name = "Selection Treatment", labels = c('Neutral', 'Positive')) +
+  theme_bw() +
+  scale_y_log10()
+
+#ggsave('selection.pdf', width = 6, height = 4)
+```
+
+```{r test_response_rank}
+
+fluxes$estimate_rank <- rank(fluxes$estimate)
+rank_fit <- lm(estimate_rank ~ treat * passage, data = fluxes)
+
+rank_fit
+summary(rank_fit)
+anova(rank_fit)
+
+```
+
+```{r}
+selection_table <- tidy(anova(rank_fit))
+selection_table[1, 1] <- "Treatment"
+selection_table[2, 1] <- "Passage"
+selection_table[3, 1] <- "Treatment:Passage"
+selection_table[4, 1] <- "Residuals"
+selection_table %>% 
+  kable(col.names = c('Term', 'Df', 'Sum Sq', 'Mean Sq', 'F Value', 'p-value')) %>% 
+  kable_styling()
+
+selection_table <- tidy(car::Anova(rank_fit, type = "II"))
+selection_table[1, 1] <- "Treatment"
+selection_table[2, 1] <- "Passage"
+selection_table[3, 1] <- "Treatment:Passage"
+selection_table[4, 1] <- "Residuals"
+selection_table %>% 
+  kable(col.names = c('Term', 'Sum Sq', 'Df', 'F Value', 'p-value')) %>% 
+  kable_styling()
+
+selection_table <- tidy(car::Anova(rank_fit, type = "III"))
+selection_table[1, 1] <- "Intercept"
+selection_table[2, 1] <- "Treatment"
+selection_table[3, 1] <- "Passage"
+selection_table[4, 1] <- "Treatment:Passage"
+selection_table[5, 1] <- "Residuals"
+selection_table %>% 
+  kable(col.names = c('Term', 'Sum Sq', 'Df', 'F Value', 'p-value')) %>% 
+  kable_styling()
+
+selection_table <- tidy(summary(rank_fit))
+
+selection_table[1, 1] <- "Neutral"
+selection_table[2, 1] <- "Positive"
+selection_table[3, 1] <- "Passage"
+selection_table[4, 1] <- "Positive:Passage"
+
+selection_table %>% 
+  kable(col.names = c('Term', 'Estimate', 'Std Error', 't-statistic', 'p-value')) %>% 
+  kable_styling()
+
+```
+
+```{r selection_rank}
+
+fluxes %>% 
+  ggplot(mapping = aes(x = passage, y = estimate_rank, color = treat)) +
+  geom_jitter(width = 0.2) +
+  geom_smooth(method = 'lm') +
+  labs(x = "Passage Number", y = "Rank Methane Oxidation Rate (-k)") +
+  scale_color_discrete(name = "Selection Treatment", labels = c('Neutral', 'Positive')) +
+  theme_bw()
+
+ggsave('selection.pdf', width = 6, height = 4)
+ggsave('selection.png', width = 6, height = 4)
+```
+
+# Deviance from mean
+
+```{r}
+fluxes <- 
+fluxes %>% 
+  group_by(passage) %>% 
+  mutate(passage_mean = mean(estimate)) %>% 
+  mutate(deviance = estimate - passage_mean)
+
+ggplot(fluxes, aes(x = passage, y = deviance, color = treat)) + 
+  geom_point() +
+  stat_smooth(method = 'lm')
+
+fit <- lm(deviance ~ treat * passage, data = fluxes)
+summary(fit)
+```
+
+# Heritability
+
+```{r heritability, message=FALSE, warning=FALSE}
+selected_jars <- read_csv('../Data/selected.csv')
+
+parental <- 
+  fluxes[paste0(fluxes$passage, fluxes$jar) %in% paste0(selected_jars$passage, selected_jars$jar), ] %>% 
+  ungroup() %>% 
+  select(treat, rep, estimate, passage)
+
+parental <-
+  parental %>% 
+  group_by(passage, treat) %>% 
+  summarize(parental = mean(estimate)) %>% 
+  ungroup() %>% 
+  filter(passage != 5)
+
+offspring <- 
+  fluxes %>% 
+  ungroup() %>%
+  select(treat, rep, estimate, passage)
+
+ggplot(offspring, aes(x = estimate, fill = factor(passage))) +
+  geom_histogram(position = "identity", alpha = 0.5)
+
+offspring <-
+  offspring %>% 
+  group_by(passage, treat) %>%
+  summarize(offspring = mean(estimate)) %>%
+  ungroup() %>%
+  # mutate(offspring = estimate) %>%
+  filter(passage != 1) %>% 
+  mutate(passage = passage - 1)
+
+ggplot(heritability, aes(x = offspring)) +
+  geom_histogram(position = 'identity', alpha = 0.5)
+
+ggplot(heritability, aes(x = parental)) +
+  geom_histogram(position = 'identity', alpha = 0.5)
+
+# offspring <-
+#   fluxes %>%
+#   ungroup() %>%
+#   select(estimate, passage, treat) %>%
+#   mutate(offspring = rank(estimate))
+
+heritability <- left_join(parental, offspring, by = c("passage", "treat"))
+heritability$passage <- factor(heritability$passage)
+heritability$treat <- factor(heritability$treat)
+# heritability$parental <- rank(heritability$parental)
+
+write_tsv(heritability, '../Output/heritability.tsv')
+
+ggplot(heritability, aes(x = parental, y = offspring)) +
+  geom_point(aes(color = treat)) +
+  stat_smooth(method = 'lm') + 
+  labs(x = "Parental Median Flux (-k)", y = "Offspring Flux (-k)") +
+  scale_color_discrete(name = "Selection\nTreatment", labels = c('Neutral', 'Positive')) +
+  theme_bw()
+
+ggsave('heritability.pdf', width = 4, height = 3)
+ggsave('heritability.png', width = 4, height = 3)
+heritability$parental_sq <- heritability$parental^2
+fit <- lm(sqrt(offspring) ~ parental, data = heritability)
+
+summary(fit)
+tidy(fit) %>% 
+  kable() %>% 
+  kable_styling()
+glance(fit) %>% 
+  kable() %>% 
+  kable_styling()
+ggplot(mapping = aes(fit$residuals, fit$fitted.values)) +
+  geom_point()
+ggplot(mapping = aes(fit$residuals, heritability$parental)) +
+  geom_point()
+
+ggplot(heritability, aes(x = parental, y = offspring)) +
+  geom_point()
+
+```
+
+```{r}
+library(rstan)
+
+srr_block <- "
+data { 
+    int<lower=1> N;               // Number of observations
+    int<lower=1> nX;               // Number of predictors
+    matrix [N,nX] X;
+    vector[N] y;                  // Methane flux values
+}
+parameters {
+    vector[nX] beta;
+    real<lower=0> sigma;
+}
+transformed parameters {
+  vector[N] mu;
+  mu = X*beta;
+  }
+model {
+    y ~ cauchy(mu, sigma);
+    sigma ~ normal(0, 0.1);
+}"
+
+Xmat <- model.matrix(~parental, heritability)
+data.list <- with(heritability, list(y = offspring, X = Xmat, nX = ncol(Xmat), N = nrow(heritability)))
+
+data.rstan <- stan(model_code = srr_block,
+                            data = data.list, 
+                            iter=3e3)
+
+stan_trace(data.rstan)
+
+mcmc = as.data.frame(data.rstan) %>% dplyr:::select(contains("beta"),
+                                                    sigma) %>% as.matrix
+## get median parameter estimates
+coefs = apply(mcmc[, 1:2], 2, median)
+fit = as.vector(coefs %*% t(Xmat))
+resid = heritability$offspring - fit
+ggplot() + geom_point(data = NULL, aes(y = resid, x = fit))
+newdata = heritability %>% cbind(fit, resid)
+ggplot(newdata) + geom_point(aes(y = resid, x = parental))
+
+print(data.rstan, par = c("beta", "sigma"))
+
+tidyMCMC(data.rstan, conf.int = TRUE, conf.method = "HPDinterval", pars = c("beta", "sigma"))
+
+library(bayesplot)
+posterior <- as.matrix(data.rstan)
+
+plot_title <- ggtitle("Posterior distributions",
+                      "with medians and 80% intervals")
+mcmc_areas(posterior,
+           pars = c("beta[2]"),
+           prob = 0.8) + plot_title
+```

### 2020-05-21 protocol.Rmd
Input data, update analysis, begin manuscript

--- 2019-09-18 Documents/oxidation_protocol.Rmd
@@
----
-title: "Oxidation Protocol"
-author: "Andrew Morris"
-date: "9/9/2019"
-output:
-  html_document:
-      toc: true
-      toc_float: 
-        collapsed: false
----
-
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE,  fig.align="center")
-library(tidyverse)
-library(knitr)
-library(kableExtra)
-library(broom)
-library(modelr)
-library(lubridate)
-```
-
 ## General notes
@@
 `ch4_area`     | The area-under-the-curve for the methane peak output from the GC.
-
-### Standard Curve
-
-```{r sc_data, fig.height=3}
-example_sc <- data.frame(
-  flux_date = c('2019-05-27', '2019-05-27', '2019-05-27', '2019-06-07', '2019-06-07', '2019-06-07', '2019-06-14', '2019-06-14', '2019-06-14'),
-  standard_ppm = c(999, 999, 999, 999, 999, 999, 999, 999, 999),
-  injection_ml = c(0.1, 0.5, 1, 0.1, 0.5, 1, 0.1, 0.5, 1),
-  ch4_area = c(556.03, 3046.03, 5995.73, 493.256, 2720.52, 5259.6, 569.3, 3255.82, 6129.72))
-example_sc <- 
-  example_sc %>% 
-  mutate(injection_ppm = standard_ppm * injection_ml) %>% 
-  select(flux_date:injection_ml, injection_ppm, ch4_area)
-example_sc %>% 
-  kable() %>% 
-  kable_styling()
-ggplot(example_sc, aes(ch4_area, injection_ppm)) +
-  facet_wrap(~ flux_date) + 
-  geom_point() +
-  geom_line() +
-  coord_fixed(6)
-
-sc_model <- function(df) {lm(injection_ppm ~ ch4_area, data = df)}
-standard_curves <- 
-  example_sc %>% 
-  group_by(flux_date) %>% 
-  nest() %>% 
-  mutate(model = map(data, sc_model),
-         glance = map(model, glance),
-         tidy = map(model, tidy)) 
-sc_r2 <- 
-  standard_curves %>% 
-  unnest(glance, .drop = TRUE) %>% 
-  arrange(r.squared) %>% 
-  select(flux_date, r.squared)
-sc_equation <- 
-  standard_curves %>% 
-  unnest(tidy, .drop = TRUE) %>%
-  select(flux_date, term, estimate) %>% 
-  spread(term, estimate) %>% 
-  rename(intercept = `(Intercept)`, slope = ch4_area)
-sc_ch4 <- left_join(sc_r2, sc_equation, by = "flux_date")
-sc_ch4 %>% 
-  kable() %>% 
-  kable_styling()
-```
-
-### Flux Measurements
-
-```{r time_data}
-
-time_data <- data.table::fread('../Data/eg_time_data.csv')
-
-time_data[, c('t0', 't1', 't2', 't3', 't4')] <- 
-  lapply(time_data[, c('t0', 't1', 't2', 't3', 't4')],
-         ymd_hms)
-
-as.days <- function(start, end) {as.numeric(interval(start, end))/60/60/24}
-
-time_data[, c('t1', 't2', 't3', 't4')] <- 
-  lapply(time_data[, c('t1', 't2', 't3', 't4')],
-         function(x) as.days(time_data$t0, x))
-
-time_data <- 
- time_data %>% 
-  select(-t0) %>% 
-  gather(t, days, t1:t4)
-
-```
-
-```{r conc_data, warning=FALSE}
-
-conc_data <- data.table::fread('../Data/eg_conc_data.csv')
-
-flux_data <- 
-  conc_data %>% 
-  gather(t, ch4_area, t1:t4) %>% 
-  left_join(sc_ch4, by = 'flux_date') %>% 
-  mutate(ch4_ppm = ch4_area * slope + intercept) %>% 
-  select(flux_date, jar, t, ch4_ppm) %>% 
-  left_join(time_data, by = c('flux_date', 'jar', 't'))
-
-```
-
-```{r plot_fluxes, fig.height=20}
-flux_data <- 
-  flux_data %>% 
-  mutate(flux_date = ymd(flux_date), 
-         jar = factor(jar),
-         t = factor(t)) %>% 
-  filter(ch4_ppm < 3000)
-
-flux_data %>% 
-  ggplot(aes(days, ch4_ppm)) + 
-  facet_grid(jar ~ flux_date, scales = "free") +
-  geom_point() + 
-  geom_line()
-```
-
-### Calculate Flux
-
-```{r calc_flux}
-
-```
-
-### Convert ppm to $\mu$grams
-
-```{r conversion}
-
-```
-
-### Selection

### 2020-05-21 R/stan.R

@@ -1,0 +1,100 @@
+library(rstan)
+library(tidyverse)
+
+data <- read_tsv('./Output/fluxes.tsv')
+rank_fit <- lm(estimate_rank ~ treat * passage, data = data)
+
+sel_sum <- tidy(summary(rank_fit))
+sel_aov <- tidy(anova(rank_fit))
+
+ggplot(data, aes(y = estimate, x = passage, color = treat)) +
+  geom_point() +
+  geom_smooth(method = "lm", se = FALSE)
+
+srr_block <- "
+data { 
+    int<lower=1> N;               // Number of observations
+    int<lower=1> nX;               // Number of predictors
+    matrix [N,nX] X;
+    vector[N] y;                  // Methane flux values
+}
+parameters {
+    vector[nX] beta;
+    real<lower=0> sigma;
+}
+transformed parameters {
+  vector[N] mu;
+
+  mu = X*beta;
+  }
+model {
+    y ~ cauchy(mu, sigma);
+    sigma ~ normal(0, 10);
+}"
+
+Xmat <- model.matrix(~treat * passage, data)
+data.list <- with(data, list(y = estimate, X = Xmat, nX = ncol(Xmat), N = nrow(data)))
+
+data.rstan <- stan(model_code = srr_block,
+                            data = data.list, 
+                            iter=1e3)
+
+print(data.rstan, par = c("beta", "sigma"))
+
+stan_trace(data.rstan)
+
+mcmc = as.data.frame(data.rstan) %>% dplyr:::select(contains("beta"),
+                                                    sigma) %>% as.matrix
+# generate a model matrix
+newdata = data
+Xmat = model.matrix(~treat * passage, newdata)
+## get median parameter estimates
+coefs = apply(mcmc[, 1:4], 2, median)
+fit = as.vector(coefs %*% t(Xmat))
+resid = data$estimate - fit
+ggplot() + geom_point(data = NULL, aes(y = resid, x = fit))
+
+
+mcmc = as.data.frame(data.rstan) %>% dplyr:::select(contains("beta"),
+                                                    sigma) %>% as.matrix
+# generate a model matrix
+newdata = newdata
+Xmat = model.matrix(~treat * passage, newdata)
+## get median parameter estimates
+coefs = apply(mcmc[, 1:4], 2, median)
+fit = as.vector(coefs %*% t(Xmat))
+resid = data$estimate - fit
[…]

### 2020-07-15 R/process.Rmd
add deviance plot and model

@@
 summary(fit)
+plot(fit$residuals ~ fluxes$estimate)
 selection_table <- tidy(anova(fit))
@@
   geom_jitter(width = 0.2) +
-  geom_smooth(method = 'lm') +
+  #geom_smooth(method = 'lm') +
   labs(x = "Passage Number", y = "Methane Oxidation Rate (-k)") +
   scale_color_discrete(name = "Selection Treatment", labels = c('Neutral', 'Positive')) +
-  theme_bw()
+  theme_bw() + 
+  geom_boxplot(aes(x = passage, y = estimate, color = treat, group = interaction(treat, passage)))
 
@@
 log_fluxes <- fluxes
-log_fluxes$estimate[log_fluxes$estimate < 0] <- 0.01
+#log_fluxes$estimate[log_fluxes$estimate < 0] <- 0.01
+log_fluxes$estimate <- log10(log_fluxes$estimate + abs(min(log_fluxes$estimate)) + 0.01)
 
 fit <- lm(estimate ~ treat * passage, data = log_fluxes)
+summary(fit)
+plot(fit$residuals ~ log_fluxes$estimate)
 
@@
 #ggsave('selection.pdf', width = 6, height = 4)
+
+log_fluxes %>% 
+  ggplot(mapping = aes(x = passage, y = estimate, color = treat)) +
+  geom_jitter(width = 0.2) +
+  #geom_smooth(method = 'lm') +
+  labs(x = "Passage Number", y = "Methane Oxidation Rate (-k)") +
+  scale_color_discrete(name = "Selection Treatment", labels = c('Neutral', 'Positive')) +
+  theme_bw() + 
+  geom_boxplot(aes(x = passage, y = estimate, color = treat, group = interaction(treat, passage)))
 ```
@@
 anova(rank_fit)
+
+plot(fit$residuals ~ fluxes$estimate_rank)
 
@@
 
+```{r}
+
+gm_mean = function(x, na.rm=TRUE){
+  exp(sum(log(x[x > 0]), na.rm=na.rm) / length(x))
+}
+
+ggplot(fluxes, aes(x = estimate)) + 
+  facet_wrap(~ passage + treat) +
+  geom_dotplot(method = 'histodot') +
+  geom_vline(aes(xintercept = mean(estimate)), color = 'blue') +
+  geom_vline(aes(xintercept = median(estimate)), color = 'red') +
+    geom_vline(aes(xintercept = gm_mean(estimate)), color = 'green')
+
+```
+
 # Deviance from mean
@@
 ```{r}
-fluxes <- 
+deviance_se <- 
   fluxes %>% 
-  group_by(passage) %>% 
-  mutate(passage_mean = mean(estimate)) %>% 
-  mutate(deviance = estimate - passage_mean)
+  select(treat, passage, estimate) %>% 
+  group_by(passage, treat) %>% 
+  summarize(se = se(estimate)) %>% 
+  pivot_wider(names_from = treat, values_from = se) %>% 
+  mutate(deviance_se = p + n) %>% 
+  pull(deviance_se)
 
-ggplot(fluxes, aes(x = passage, y = deviance, color = treat)) + 
-  geom_point() +
-  stat_smooth(method = 'lm')
+deviance <-
+fluxes %>% 
+  select(treat, passage, estimate) %>% 
+  group_by(passage, treat) %>% 
+  summarize(mean = median(estimate)) %>% 
+  pivot_wider(names_from = treat, values_from = mean) %>% 
+  mutate(deviance = p - n)
 
-fit <- lm(deviance ~ treat * passage, data = fluxes)
+deviance$se <- deviance_se
+
+fit <- lm(deviance ~ passage, data = deviance)
 summary(fit)
+
+fit <- rma(deviance~passage,se^2,method="FE",data = deviance)
+fit
+
+summary(fit)
+plot(fit)
+
+ggplot(deviance, aes(x = passage, y = deviance, ymin = deviance - se, ymax = deviance + se)) + 
+  geom_pointrange() + 
+  geom_abline(intercept = fit$beta[1], slope = fit$beta[2], color = 'blue')
+
 ```

### 2020-07-16 Manuscript/manuscript.Rmd
Add deviance figure and model results

@@
 ---
-title: "Artificial ecosystem selection on soil methane oxidation reveals simple/complex mapping between microbial community structure and ecosystem function"
+title: Artificial ecosystem selection on soil methane oxidation reveals simple/complex
+  mapping between microbial community structure and ecosystem function
 author: "Andrew H. Morris and Brendan J. M. Bohannan"
@@
     reference_docx: reference.docx
+editor_options:
+  chunk_output_type: inline
 ---
@@
 library(broman)
+library(morris)
+library(metafor)
 figures <- captioner()
 figures("selection", "Rank CH~4~ oxidation rate for each jar in each passage.")
-figures("herit", "Correlation between mid-parent CH~4~ flux and mid-offspring CH~4~ flux.")
+figures("deviance", "Deviance of CH~4~ oxidation rate for the Positive selection treatment relative to the Neutral selection treatment.
+Deviance is calculated as the mean of P within each passage minus the mean of N within each passage.
+Errorbars are the sums of the standard errors for P and N for each passage.
+The best fit line is an ordinary least squares fit taking into account the standard errors of the deviations.
+The gray ribbon is the 95% confidence interval for the regression coefficient.")
+figures("herit", "Correlation between mid-parent CH~4~ flux and mid-offspring CH~4~ flux.
+The mid-parents are the mean of the three (or two) jars selected to inoculate the next passage.
+The mid-offsprings are the mean of all twelve jars produced in one passage.
+The best fit line is an ordinary least squares regression line and the gray ribbon is the 95% confidence interval for the regression coefficient.")
 
@@
 
+## Response to selection
+
 ```{r sel_model}
@@
 
-```{r sel_fig, fig.width=5, fig.height=4, fig.cap=figures("selection")}
+```{r sel_fig, fig.width=5, fig.height=3, fig.cap=figures("selection")}
 fluxes %>% 
@@
   geom_jitter(width = 0.2) +
-  geom_smooth(method = 'lm') +
+  geom_smooth(method = 'lm', formula = y ~ x, se = FALSE) +
   labs(x = "Passage Number", y = "Rank Methane Oxidation Rate (-k)") +
-  scale_color_discrete(name = "Selection\nTreatment", labels = c('Neutral', 'Positive')) +
+  #scale_color_discrete(name = "Selection\nTreatment", labels = c('Neutral', 'Positive')) +
+    scale_color_manual(name = "Selection\nTreatment", labels = c('Neutral', 'Positive'), values = c("gray60", "darkorange2")) +
   theme_bw()
@@
 
+## Deviance
+
+```{r dev_model}
+deviance_se <- 
+  fluxes %>% 
+  select(treat, passage, estimate) %>% 
+  group_by(passage, treat) %>% 
+  summarize(se = se(estimate)) %>% 
+  pivot_wider(names_from = treat, values_from = se) %>% 
+  mutate(deviance_se = p + n) %>% 
+  pull(deviance_se)
+
+deviance <-
+fluxes %>% 
+  select(treat, passage, estimate) %>% 
+  group_by(passage, treat) %>% 
+  summarize(mean = mean(estimate)) %>% 
+  pivot_wider(names_from = treat, values_from = mean) %>% 
+  mutate(deviance = p - n)
+
+deviance$se <- deviance_se
+
+dev_fit <- rma(deviance~passage, se^2, method="FE", data = deviance)
+dev_fit1 <- lm(deviance~passage, data = deviance)
+
+flux_per_psg <-
+fluxes %>% 
+  group_by(passage) %>% 
+  summarize(mean = mean(estimate), se = se(estimate)) %>% 
+  pull(mean)
+
+fluxes %>% 
+  group_by(passage, treat) %>% 
+  summarize(mean = mean(estimate), se = se(estimate))
+
+dev_pred <- predict(dev_fit)
+```
+
+Methane oxidation rate increased as a response to selection in the Positive selection treatment (`r figures("deviance", display="cite")`).
+This can be seen in the increase in deviation between the Positive and Neutral selection treatments over the course of the experiment (z = `r myround(dev_fit$zval[2], 2)`, p = `r myround(dev_fit$pval[2], 3)`).
+Mean CH~4~ oxidation rate for the Positive treatment increased by `r myround(dev_fit$beta[2], 3)` per passage relative to the Neutral treatment.
+The mean flux across all jars in all treatments in the first passage was `r myround(flux_per_psg[1], 3)` so this effectively represents a doubling of CH~4~ consumption in every passage due to selection.
+
+```{r dev_fig, fig.width=4, fig.height=3, fig.cap=figures("deviance")}
+
+ggplot(deviance, aes(x = passage, y = deviance, ymin = deviance - se, ymax = deviance + se)) + 
+  geom_ribbon(aes(ymin = dev_pred$ci.lb, ymax = dev_pred$ci.ub), alpha = 0.2) +
+  geom_pointrange(color = 'darkorange2') + 
+  #geom_abline(intercept = 0.1 + fit$beta[1], slope = fit$beta[2], color = 'blue') +
+      geom_segment(aes(x = 1, xend = 5, y = dev_fit$beta[1] + dev_fit$beta[2], yend = dev_fit$beta[1] + dev_fit$beta[2]*5), color = 'darkorange2', size = 1) +
+  labs(x = "Passage Number", y = "Deviation from Control (P - N)")+
+  theme_bw()
+```
+
+## Ecosystem heritability
+
 ```{r herit_model}
 heritability <- read_tsv('../Output/heritability.tsv')
-fit <- lm(offspring ~ parental, data = heritability)
-her_sum <- tidy(summary(fit))
-her_aov <- tidy(anova(fit))
-```
+her_fit <- lm(offspring ~ parental, data = heritability)
+her_sum <- tidy(summary(her_fit))
+her_aov <- tidy(anova(her_fit))
 
-Methane oxidation rate was heritable between consecutive generations (`r figures("herit", display="cite")`).
-Specifically, the narrow-sense heritability ($h^2$), calculated as the slope of the mid-parent and mid-offspring, was `r myround(her_sum[2, 2], 2)`.
-
-```{r herit_fig, fig.width=5, fig.height=4, fig.cap=figures("herit")}
-ggplot(heritability, aes(x = parental, y = offspring)) +
-  geom_point(aes(color = treat)) +
-  stat_smooth(method = 'lm') + 
-  labs(x = "Parental Mean Flux (-k)", y = "Offspring Mean Flux (-k)") +
-  scale_color_discrete(name = "Selection\nTreatment", labels = c('Neutral', 'Positive')) +
-  theme_bw()
 ```
@@
 F = 1 - (1 - 1 / (2 * N)) ^ T
-F
+```
+
+Methane oxidation rate was heritable between consecutive passages (`r figures("herit", display="cite")`).
+Specifically, the narrow-sense heritability, calculated as the slope of the mid-parent and mid-offspring, was $h^2$ = `r myround(her_sum[2, 2], 2)` $\pm$ `r myround(her_sum[2, 3], 2)`.
+With an inbreeding coefficient of `r myround(F, 2)`,
+
+```{r herit_fig, fig.width=5, fig.height=4, fig.cap=figures("herit")}
+ggplot(heritability, aes(x = parental, y = offspring)) +
+  geom_point(aes(color = treat)) +
+  stat_smooth(method = 'lm', formula = y ~ x, color = "black") + 
+  labs(x = "Parental Mean Flux (-k)", y = "Offspring Mean Flux (-k)") +
+    scale_color_manual(name = "Selection\nTreatment", labels = c('Neutral', 'Positive'), values = c("gray60", "darkorange2")) +
+  theme_bw()
+
+```
+
+```{r}
+R = dev_fit$beta[2]
+h2 = pull(her_sum[2, 2])
+S = R/h2
+```
+
+```{r}
+fluxes
+heritability
 ```

### 2020-07-16 R/process.Rmd
Add deviance figure and model results

@@
 library(magrittr)
+library(metafor)
 ```
@@
   geom_point() +
-  stat_smooth(method = 'lm', se = FALSE)
+  stat_smooth(formula = y ~ x, method = 'lm', se = FALSE)
 
@@
 fit <- lm(estimate ~ treat * passage, data = fluxes)
-summary(fit)
 plot(fit$residuals ~ fluxes$estimate)
@@
   ggplot(mapping = aes(x = passage, y = estimate, color = treat)) + 
+  geom_boxplot(aes(x = passage, y = estimate, color = treat, group = interaction(treat, passage))) +
   geom_jitter(width = 0.2) +
@@
   scale_color_discrete(name = "Selection Treatment", labels = c('Neutral', 'Positive')) +
-  theme_bw() + 
-  geom_boxplot(aes(x = passage, y = estimate, color = treat, group = interaction(treat, passage)))
+  theme_bw()
 
@@
   group_by(passage, treat) %>% 
-  summarize(mean = median(estimate)) %>% 
+  summarize(mean = mean(estimate)) %>% 
   pivot_wider(names_from = treat, values_from = mean) %>% 
@@
 ggplot(deviance, aes(x = passage, y = deviance, ymin = deviance - se, ymax = deviance + se)) + 
-  geom_pointrange() + 
+  geom_pointrange(color = 'darkorange2') + 
   geom_abline(intercept = fit$beta[1], slope = fit$beta[2], color = 'blue')
@@
   mutate(passage = passage - 1)
-
-ggplot(heritability, aes(x = offspring)) +
-  geom_histogram(position = 'identity', alpha = 0.5)
-
-ggplot(heritability, aes(x = parental)) +
-  geom_histogram(position = 'identity', alpha = 0.5)
 
@@
 
+ggplot(heritability, aes(x = offspring)) +
+  geom_histogram(position = 'identity', alpha = 0.5)
+
+ggplot(heritability, aes(x = parental)) +
+  geom_histogram(position = 'identity', alpha = 0.5)
+
 ```
 
-```{r}
+```{r eval=FALSE, include=FALSE}
 library(rstan)

### 2020-07-16 Manuscript/manuscript.Rmd
Remove old rank-based response to selection

@@
 
-```{r sel_model}
+```{r dev_model}
 fluxes <- read_tsv('../Output/fluxes.tsv')
-rank_fit <- lm(estimate_rank ~ treat * passage, data = fluxes)
 
-sel_sum <- tidy(summary(rank_fit))
-sel_aov <- tidy(anova(rank_fit))
-```
-
-Methane oxidation rates increased over the course of the experiment in the positive selection treatment and did not change in the neutral selection treatment (`r figures("selection", display="cite")`).
-This can be seen in the significant interaction between treatment and passage number for the rank of CH~4~ oxidation rate (F~`r sel_aov[3,2]`,`r sel_aov[4,2]`~ = `r myround(sel_aov[3,5], 2)`, p = `r myround(sel_aov[3,6], 3)`).
-While there was no difference in the overall mean of the two treatments (F~`r sel_aov[1,2]`,`r sel_aov[4,2]`~ = `r myround(sel_aov[1,5], 2)`, p = `r myround(sel_aov[1,6], 2)`), the mean rank CH~4~ oxidation rate increased by `r myround(sel_sum[4,2],1)`  per passage for the positive selection treatment (t = `r myround(sel_sum[4, 4], 2)`, p = `r myround(sel_sum[4, 5], 3)`), but the neutral treatment did not change ($\beta$ = `r myround(sel_sum[3,2],1)`, t = `r myround(sel_sum[3,4], 2)`, p = `r myround(sel_sum[3,5], 2)`).
-
-```{r sel_fig, fig.width=5, fig.height=3, fig.cap=figures("selection")}
-fluxes %>% 
-  ggplot(mapping = aes(x = passage, y = estimate_rank, color = treat)) +
-  geom_jitter(width = 0.2) +
-  geom_smooth(method = 'lm', formula = y ~ x, se = FALSE) +
-  labs(x = "Passage Number", y = "Rank Methane Oxidation Rate (-k)") +
-  #scale_color_discrete(name = "Selection\nTreatment", labels = c('Neutral', 'Positive')) +
-    scale_color_manual(name = "Selection\nTreatment", labels = c('Neutral', 'Positive'), values = c("gray60", "darkorange2")) +
-  theme_bw()
-```
-
-## Deviance
-
-```{r dev_model}
 deviance_se <- 

### 2020-07-28 Manuscript/manuscript.Rmd
use multiple standard curves per flux date

@@
 ---
-title: Artificial ecosystem selection on soil methane oxidation reveals simple/complex
-  mapping between microbial community structure and ecosystem function
+title: Artificial ecosystem selection to deduce the mapping between microbial community structure and ecosystem function
 author: "Andrew H. Morris and Brendan J. M. Bohannan"
@@
 figures <- captioner()
-figures("selection", "Rank CH~4~ oxidation rate for each jar in each passage.")
 figures("deviance", "Deviance of CH~4~ oxidation rate for the Positive selection treatment relative to the Neutral selection treatment.
@@
 The best fit line is an ordinary least squares fit taking into account the standard errors of the deviations.
-The gray ribbon is the 95% confidence interval for the regression coefficient.")
-figures("herit", "Correlation between mid-parent CH~4~ flux and mid-offspring CH~4~ flux.
-The mid-parents are the mean of the three (or two) jars selected to inoculate the next passage.
+The gray ribbon is the 95% confidence interval for the regression.")
+figures("herit", "Regression of offspring CH~4~ flux on mid-parent CH~4~ flux.
+Mid-parent is the mean of jars selected to inoculate the next passage.
 The mid-offsprings are the mean of all twelve jars produced in one passage.
@@
 
-Questions:
-
-- What is the mapping between microbial community structure and ecosystem function?
-
-- Is it simple? e.g., abundance of functional group, diversity of functional group (like in plants). Or is it complex? e.g., multiple functional groups, multiple taxa.
-
-- How does it compare to macroogranisms?
-Have people even tested functional group diversity extensively?
-Is methanotrophy as deeply conserved in bacteria as photosynthesis in plants?
-i.e., is there a 16S sequence or other marker that maps to the methanotrophy guild?
-It seems like microbial BEF has focused  on Prokaryote diversity or functional group abundance, which are both different from how it's been carried out in plants.
-
-- Does BEF theory apply? Niche theory: One limiting R -> one final competitor. Multiple Rs -> multiple peristent competitors.
-
 # Introduction
-
-Biodiversity-ecosystem function (BEF) research has the central goal of understanding whether and how biodiversity - the diversity and relative abundance of species in a community - determines the magnitude of ecosystem functions (cite Hooper, Naeem, Tilman, ...).
-Ecosystem functions are a broad range of processes, including pollination, water filtration, carbon storage, and nutrient cycling, that are driven by both abiotic and biotic factors.
-The ultimate goal of this research is to predict how changes in biodiversity due to climate change or land use change will alter ecosystem function and, in turn, understand how we should protect biodiversity in order to preserve essential ecosystem functions.
-The fundamental question for this research is what is the nature of the "map" between elements of biodiversity and elements of ecosystem function?
-
-One could hypothesize a relatively simple map in which a single taxon determines the rate of an ecosystem function based entirely on its presence or abundance.
-(e.g.
-of this?) Alternatively, instead of defining biodiversity based on a taxonomic unit, biodiversity could be defined by functional groups or guilds.
-For example, photosynthesis could be determined by the abundace of photosynthesizers in a community.
-(Example of abundance of photosynthesizers?
-measured as the Abundance of rubisco genes?).
-A somewhat more complex map could define biodiversity based on the diversity (i.e., richness or number of species) of photosynthesizers.
-Using this mapping, an abundance of literature has demonstrated that productivity increases asymptotically and variation in productivity decreases asymptotically with the richness of producers (Tilman 1997).
-
-When microbial ecologists have attempted to define the mapping of ecosystem functions performed by microorganisms, these simple models rarely hold (Rocca et al., Graham ete al.).
-Most functions in most ecosystems do not exhibit a positive relationship between the abundance of functional marker genes and the rate of the associated function.
-Few studies test the diversity of functional groups.
-This could mean that the map is simple, but we are not accurately estimating the abundance of these functional groups.
-Alternatively, the mapping between microbial community structure and microbial processes could be more complex than the mapping between macrobial structure and ecosystem function.
-
-To investigate the nature of the map between microbial community structure and ecosystem function, we will perform artificial ecosystem selection (*sensu* Swenson et al.
-2000).
-With this approach, we will collect a single soil which we subsamble to generate multiple ecosystems.
-These ecosystems vary in both community structure and the rate of ecosystem function.
-From this variation, we select the top performing ecosystems, reshuffle their members through homogenization, and subsample this homogenate to generate more ecosystems.
-Through this iterative process, we test many different combinations of taxa and enrich for the top performers.
-This process of selection also reduces the genetic variation and diversity of the communities.
-By minimizing the environmental differences between jars and between passages, we can insure that the majority of functional variation from jar to jar and passage to passage is due to genetic variation in the community.
-As a result, genetic differences between the initial community and the selected community should be associated with the function of interest.
-
-We hypothesize that the positive selection treatment will increase in methane oxidation rate while the neutral selection treatment will resemble a random walk over time demonstrating community heritability within our system.
-We also hypothesize that genetic changes over time will follow a predictable trajectory: the fastest changes in community copmosition will be due to species loss, slower changes will be due to changes in the relative abundance of species, and the slowest changes will be due to intraspecies selection (i.e., evolution).
-The soil communities will not only be adapting in response to selection on soil methane oxidation, but will also be adapting to survival and passaging between jars.
-
-There are several alternatives as to the complexity of the "map" between microbes and methane.
-At the simplest level, methane oxidation rate could be a function of the abundance of methane oxidizers.
-A somewhat more complex map would be if methane flux is a function of the diversity of methane oxidizers as in other systems (Hooper et al.).
-If neither of these simple models hold, the mapping might be a more complex relationship between a broader level of organization, such as methane producers and consumers, a broader metabolic network, or a network of bacterial and archaeal interactions that's hard to predict.
-
-The gold standard of molecular biology gene mapping studies is to not only identify a correlation, but also to confirm the association through knockout and overexpression studies.
-To approximate this approach for microbial ecology, we cultured the important strain(s) in pure culture and applied them to the neutral soil community to see if we could recapitulate the trait level for the selected treatment with only 1 or a few taxa or if it required the whole community.
-We used the neutral treatment, because these communities are already bottle-adapted and offer a more similar environment for the selected strains than a natural soil would.
-
-Our goal is to understand the identity of the community members and their interactions that result in the greatest level of ecosystem function.
-Artificial ecosystem selection allows us to test many "teams" or communities of microbes in terms of their combined impact on function.
-By performing selection, we reduce the genetic variation in our natural community to the variation that is sufficient to perform a high rate of ecosystem function.
-Then, we can investigate the remaining variation to generate hypotheses about the relationship between community composition and ecosystem function, which mebmers are the best performers, and how simple or complex the map is between community and funtion.
-Finally, assuming we can culture these members, we can test those hypotheses through enrichment of one or a group of taxa to see which members are sufficient to recreate the highest rate of ecosystem function.
-
-Questions:
-
-1.
-What is the “map”?
-And how do we get to that map?
-What is the nature of the map?
-Taxonomy?
-Or something other than taxonomy?
-Simple – taxon.
-Diffuse – something other than taxonomy.
-Complicated interactions among organisms leads to “hitchhikers.”
-
-2. Can we isolate 1 or a few strains and add them to the N community? Or does it take a whole complex community? The
-
-3. Lewontin posited that higher levels of selection (species, community, ecosystem) would have weaker heritability. Do we observe this?
-
-1. Does soil methane oxidation rate respond to selection at the ecosystem level?
-
-2. Does a community that has been selected on for high methane oxidation rate differ in diversity or composition from a community that has been passaged without selection for methane oxidation rate?
-
-3.
-What aspects of community composition differ between the positive and neutral lines?
-Changes in alpha diversity, beta diversity?
-Presence/absence of specific microbial taxa?
-Relative abundance of microbial taxa?
-
-4. Does selection lead to greater differences in the bacterial/archaeal community (inferred from 16S sequences) or in the methanotroph community (inferred from pmoA sequences)?
-
-5. Does a community with high methane oxidation rate dominate coalescence with a natural community compared to a community with average methane oxidation rate?
-
-8. Many-to-one mapping
-
-9. Crossing/mixing lines or competing lines or applying to natural soils for the application of the tool.
-
-Hypotheses:
-
-1. Soil methane oxidation rate responds to selection at the ecosystem level
-
-2.
-
-Questions to address in the Intro:
-
-Is this “just enrichment” or is it community selection? i.e., is it dilution to extinction?
-
-How do people enrich for methanotrophs?
-
-Community-level selection? Paul Rainey, Martin Polz
 
@@
 The trait we selected on was CH~4~ oxidation rate.
-Our experiment had two selection lines with twelve jars each for a total of twenty-four jars.
+Our experiment had two selection lines with twelve jars each for a total of twenty-four jars per passage.
 One line was a positive selection line where the two or three jars with the highest CH~4~ oxidation rate were chosen to inoculate the next set of jars.
-The other line was a neutral selection where an equal number of jars as the positive line were chosen at random among the twelve.
+The other line was a neutral selection line where an equal number of jars as the positive line were chosen at random among the twelve.
 The number of jars chosen was based on the distribution of fluxes among the positive jars, i.e., we chose the top three jars unless only two jars had considerably greater oxidation rates based on visual inspection of histograms.
@@
 
-To reveal the mapping between soil microbial community structure and methane oxidation rate, we will estimate the abundance of methane oxidizers by quantifying pmoA, we will estimate the diversity of methane oxidizers by sequencing the pmoA gene, and we will estimate bacterial and archaeal diversity by sequencing the 16S gene.
-
-Samples to extract:
-
-- generation 2
-- generation 5
-- starting soil
-- autoclaved potting mix negative control
-- sterile water negative control
-
-Soil DNA was extracted from a subsample of soil from the starting inoculum and from each jar in passage 2 and 5 after CH~4~ flux measurements.
-Negative controls were extracted from autoclaved potting mix and autoclaved, deinoized water.
-Extractions were performed using the DNeasy PowerSoil kit (QIAGEN, Hilden, Germany).
-<mark>Find example extraction and sequencing text.</mark> In order to estimate the diversity and relative abundance of the bacterial and archaeal community in our soil ecosystems, we sequenced the V3-V4 region of 16S rRNA gene using the Earth Microbiome Project protocol and primers<mark>???</mark>.
-In addition, to estimate the diversity of CH~4~ oxidizers, we sequenced the particulate methane monooxygenase subunit A gene (*pmoA*) using primers<mark>???</mark>.
+Soil DNA was extracted from a 0.25 g subsample of soil from the starting inoculum and from each jar in passage 2 and 5.
+Negative controls were extracted from autoclaved potting mix and DNase-free water.
+Extractions were performed using the DNeasy PowerSoil kit (QIAGEN, Hilden, Germany) and quantified using Qubit (company, city, state).
+In order to estimate the diversity and relative abundance of the bacterial and archaeal taxa in our soil ecosystems, we sequenced the V3-V4 region of the 16S rRNA gene using the Earth Microbiome Project protocol and primers<mark>???</mark>.
+In addition, to estimate the diversity and relative abundance of CH~4~ oxidizing taxa, we sequenced the particulate methane monooxygenase subunit A gene (*pmoA*) using primers<mark>???</mark>.
+Sequencing was performed on the Illumina MiSeq in v3 mode with 300 bp paired-end reads.
 
@@
 based on the Silva database.
-All subsequence statistical analyses were performed in the `R` statistical programming environment (R Core Team 2020).
-Richness estimates were made using the `breakaway` package (Willis et al.
+All subsequent tatistical analyses were performed in the `R` statistical programming environment (R Core Team 2020).
+Richness estimates were made using the `breakaway` package and diversity indices using the `DivNet` package (Willis et al.
 ???).
@@
 
-To test whether there was a significant change in CH~4~ oxidation rate (`k`) over multiple passages as a response to selection, we fit a multiple linear regression model using ordinary least squares:
-
-$$
-\hat{y}_i = \beta_0 + \beta_1 x_{1i} + \beta_2 x_{2i} + \beta_3 x_{1i} x_{2i}
-$$
-
-where $\hat{y}$ was the predicted CH~4~ oxidation rate in jar $i$ as a function of treatment $x_1$ (positive or neutral selection), passage number $x_2$, and the interaction between treatment and passage number $x_1 x_2$.
-We fit the model in `R` version `r paste(R.Version()$major, R.Version()$minor, sep = ".")` using the `lm` and `anova` functions (R Core Team 2020).
-Oxidation rates were highly skewed and residuals did not meet the assumptions of normality and heterogeneity of variance, therefore, fluxes were rank-transformed prior to analysis.
+To test whether there was a significant change in CH~4~ oxidation rate (k) as a response to selection, we tested the deviation between the positive and neutral selection lines over the five passages.
+We calculated deviation as the difference in the mean oxidation rate between the positive and neutral line within each passage with standard errors equal to the sum of the errors for each of the positive and neutral lines.
+We then fit an ordinary least squares regression model taking into account the standard errors of the deviations.
+The model was fit using the `rma` function from the `metafor` package in `r` (Viechtbauer 2010).
 
@@
 
-We will estimate heritability of this trait as the correlation between parent and offspring jars.
-
-To test the heritability of CH~4~ oxidation rate as an ecosystem property, we calculated narrow-sense heritability as the slope of the relationship between the mid-parent CH~4~ oxidation rate and the mid-offspring CH~4~ oxidation rate (Goodnight 2000).
-For example, the mid-parent of passage 1 was calculated as the mean of the three jars selected to inoculate passage 2 and the mid-offspring was calculated as the mean of all 12 jars produced from those parental jars.
-To estimate heritability, we fit the following model using ordinary least squares:
-
-$$
-\hat{y} = \beta_0 + \beta_1 x_{1}
-$$
-
-where $x$ was the parental mean, $\hat{y}$ was the offspring mean, and $\beta_{1}$ was heritability.
+To test the heritability of CH~4~ oxidation rate as an ecosystem property, we calculated narrow-sense heritability ($h^2$) as the regression of offspring on mid-parent (Goodnight 2000, Falconer 1997).
+The mid-parent was the mean for all three selected jars and the offspring was the mean for all jars produced by those parents.
 
@@
 
-where $N$ is the population size and $T$ is the number of generations (Goodnight 2000).
+where $N$ is the population size and $T$ is the number of generations (Goodnight 2000 or Falconer 1997).
 
@@
 We compared taxa in passage 5 between the positive and neutral treatment to identify taxa that
-
-## Community over-expression assay
-
-* Pick out specific crobes that respond to selection and are differentially abundant and try to isolate and culture them. Either on media or in soil??
-* Add individual crobes back to a lab soil community to "overexpress" and see if the same "phenotype" is achieved. Compared that to an extraction of the full community. Use water as a control.
-Should we inoculate them onto sterile potting mix or onto neutral potting mix or onto natural potting mix?
-
-## Depth of trait conservation - What is the taxonomic depth or % 16S sequence similarity that best explains variation in methane oxidation rates?
 
@@
 dev_fit <- rma(deviance~passage, se^2, method="FE", data = deviance)
-dev_fit1 <- lm(deviance~passage, data = deviance)
+dev_fit2 <- lm(deviance~passage, data = deviance)
 
@@
   group_by(passage) %>% 
-  summarize(mean = mean(estimate), se = se(estimate)) %>% 
-  pull(mean)
-
-fluxes %>% 
-  group_by(passage, treat) %>% 
   summarize(mean = mean(estimate), se = se(estimate))
@@
 Methane oxidation rate increased as a response to selection in the Positive selection treatment (`r figures("deviance", display="cite")`). 
-This can be seen in the increase in deviation between the Positive and Neutral selection treatments over the course of the experiment (z = `r myround(dev_fit$zval[2], 2)`, p = `r myround(dev_fit$pval[2], 3)`).
-Mean CH~4~ oxidation rate for the Positive treatment increased by `r myround(dev_fit$beta[2], 3)` per passage relative to the Neutral treatment.
-The mean flux across all jars in all treatments in the first passage was `r myround(flux_per_psg[1], 3)` so this effectively represents a doubling of CH~4~ consumption in every passage due to selection.
+This can be seen in the increase in deviation between the Positive and Neutral selection lines over the course of the experiment (z = `r myround(dev_fit$zval[2], 2)`, p = `r myround(dev_fit$pval[2], 3)`).
+Mean CH~4~ oxidation rate for the Positive treatment increased by `r myround(dev_fit$beta[2], 3)` $\pm$ `r myround(dev_fit$se[2], 3)` per passage relative to the Neutral treatment.
+The mean flux across all jars in all treatments in the first passage was `r myround(flux_per_psg[1, 2], 3)` $\pm$ `r myround(flux_per_psg[1, 3], 3)` so this represents a doubling of CH~4~ consumption in every passage due to selection.
 
@@
 
-```{r}
-fluxes
-heritability
-```
+# Discussion
 
-# Discussion
+## Future Directions
+
+This study focused on the practical implications of using AES as a tool to deduce how microbes map onto functions at the whole-ecosystem level.
+If one were interested in the evolutionary dynamics underlying this process, one could design a study to distinguish between individual selection on microbial strains and group selection or multilevel selection.
+That is, selection that cannot be recapitulated through selection at the individual level (Williams and Lenton 2007).
+This could be accomplished by culturing differentially abundant taxa in the positive selection treatment and then inoculating them in isolation (or in a non-selected background community) to see if the same level of ecosystem function could be recapitulated as in the final generation of the positive selection treatment.
+
+Another avenue for research could be to look at correlated responses of ecosystem-level traits to selection.
+This would address questions such as, "Is methane oxidation rate correlated with other ecosystem processes due to pleiotropy or linkage disequilibrium?" At the community level, pleiotropy might be defined as "the production by a single taxon of two or more unrelated effects." Further, linkage disequilbrium would represent correlated traits due to taxa who tend to co-associate due to shared environmental preferences or coupled metabolic pathways.
+
+These experiments would further help lus understand the nature of the "map" between community structure and ecosystem function.
+This would not only reveal how best to predict ecosystem function in earth system models, but also demonstrate the mechanism underlying that relationship.
+This is akin to deducing the genetic control of trait variation in molecular biology.
+
+Community epistasis is analogous to functional redundancy.
 

### 2020-07-28 R/process.Rmd
use multiple standard curves per flux date

@@
         collapsed: false
+editor_options: 
+  chunk_output_type: inline
 ---
@@
 sc <- read.csv('../Data/standard_curve.csv')
+sc_dates <- read.csv('../Data/sc_dates.csv')
 
@@
 ggplot(mapping = aes(area, injection_ppm)) +
-  facet_wrap(~ flux_date + molecule, scales = "free") + 
+  facet_wrap(~ sc_date + molecule, scales = "free") + 
   geom_point() +
@@
   filter(molecule == 'ch4') %>% 
-  group_by(flux_date) %>% 
+  group_by(flux_date, sc_date) %>% 
   nest() %>% 
@@
   unnest(glance) %>% 
-  select(flux_date, r.squared)
+  select(flux_date, sc_date, r.squared)
 sc_equation <- 
@@
   unnest(tidy) %>%
-  select(flux_date, term, estimate) %>% 
+  select(flux_date, sc_date, term, estimate) %>% 
   spread(term, estimate) %>% 
   rename(intercept = `(Intercept)`, slope = area)
-sc_ch4 <- left_join(sc_r2, sc_equation, by = c("flux_date"))
+sc_ch4 <- left_join(sc_r2, sc_equation, by = c("flux_date", "sc_date"))
 sc_ch4 %>% 
@@
   kable_styling()
+
+sc_ch4 <- 
+sc_ch4 %>% 
+  left_join(sc_dates) %>% 
+  select(flux_date:sc_date, t, r.squared:slope)
 
@@
   gather(t, area, t1:t5) %>% 
-  left_join(sc_ch4, by = c('flux_date')) %>% 
+  left_join(sc_ch4, by = c('flux_date', 't')) %>% 
   mutate(ppm = area * slope + intercept) %>% 

### 2020-07-28 R/process.Rmd
remove stan chunk

@@
   kable_styling()
-
-```
-
-```{r plot_selection_data, eval=FALSE, include=FALSE}
-
-fluxes %>% 
-  ggplot(fluxes, mapping = aes(x = passage, y = estimate, group = passage)) +
-  geom_boxplot() +
-  geom_jitter(width = 0.2) +
-  labs(x = "Passage Number", y = "Methane Oxidation Rate (-k)") +
-  theme_bw()
-
-fluxes %>% 
-  ggplot(mapping = aes(x = treat, y = estimate)) +
-  geom_boxplot() +
-  geom_jitter(width = 0.2) +
-  labs(x = "Treatment", y = "Methane Oxidation Rate (-k)") +
-  theme_bw()
-
-library(rstanarm)
-
-post1 <- stan_glm(estimate ~ treat, data = fluxes, prior = cauchy(), seed = 12345)
-post2 <- update(post1, formula = . ~ passage)
-post3 <- update(post1, formula = . ~ treat + passage)
-(post4 <- update(post1, formula = . ~ treat * passage))
-
-base <-  ggplot(fluxes, mapping = aes(x = passage, y = estimate, group = passage)) +
-  geom_boxplot() +
-  geom_jitter(width = 0.2) +
-  labs(x = "Passage Number", y = "Methane Oxidation Rate (-k)") +
-  theme_bw()
-
-base + geom_abline(intercept = coef(post1)[1], slope = coef(post1)[2], 
-                   color = "skyblue4", size = 1)
-draws <- as.data.frame(post1)
-colnames(draws)[1:2] <- c("a", "b")
-
-base + 
-  geom_abline(data = draws, aes(intercept = a, slope = b), 
-              color = "skyblue", size = 0.2, alpha = 0.25) + 
-  geom_abline(intercept = coef(post1)[1], slope = coef(post1)[2], 
-              color = "skyblue4", size = 1)
-
-draws <- as.data.frame(as.matrix(post2))
-colnames(draws)[1:2] <- c("a", "b")
-ggplot(fluxes, mapping = aes(x = treat, y = estimate)) +
-  geom_boxplot() +
-  geom_jitter(width = 0.2) +
-  labs(x = "Treatment", y = "Methane Oxidation Rate (-k)") +
-  theme_bw() +
-  geom_abline(data = draws, aes(intercept = a, slope = b), 
-              color = "skyblue", size = 0.2, alpha = 0.25) + 
-  geom_abline(intercept = coef(post2)[1], slope = coef(post2)[2], 
-              color = "skyblue4", size = 1)
-
-reg0 <- function(x, ests) cbind(1, 0, x) %*% ests 
-reg1 <- function(x, ests) cbind(1, 1, x) %*% ests
-
-args <- list(ests = coef(post3))
-lgnd <- guide_legend(title = NULL)
-base2 <- ggplot(fluxes, aes(x = passage, fill = treat)) + 
-  geom_jitter(aes(y = estimate), shape = 21, stroke = .2, size = 1) + 
-  guides(color = lgnd, fill = lgnd) + 
-  theme(legend.position = "right")
-base2 + 
-  stat_function(fun = reg0, args = args, aes(color = "Neutral"), size = 1.5) +
-  stat_function(fun = reg1, args = args, aes(color = "Positive"), size = 1.5)
-
-reg0 <- function(x, ests) cbind(1, 0, x, 0 * x) %*% ests 
-reg1 <- function(x, ests) cbind(1, 1, x, 1 * x) %*% ests
-args <- list(ests = coef(post4))
-base2 +
-  stat_function(fun = reg0, args = args, aes(color = "Neutral"), size = 1.5) + 
-  stat_function(fun = reg1, args = args, aes(color = "Positive"), size = 1.5)
-
-loo1 <- loo(post1, cores = 2, k_threshold = 0.7)
-loo2 <- loo(post2, cores = 2, k_threshold = 0.7)
-loo3 <- loo(post3, cores = 2, k_threshold = 0.7)
-loo4 <- loo(post4, cores = 2, k_threshold = 0.7)
-(comp <- loo_compare(loo1, loo2, loo3, loo4))
-######################
-
-post2 <- stan_lmer(estimate ~ 1 + (1|treat) + (1|passage) + (1|treat:passage),
-                   data = fluxes, prior_intercept = cauchy(),
-                   prior_covariance = decov(shape = 2, scale = 2),
-                   adapt_delta = 0.999, seed = 12345)
 

### 2020-07-28 R/process.Rmd
remove second stan chunk

@@
 ```
-
-```{r eval=FALSE, include=FALSE}
-library(rstan)
-
-srr_block <- "
-data { 
-    int<lower=1> N;               // Number of observations
-    int<lower=1> nX;               // Number of predictors
-    matrix [N,nX] X;
-    vector[N] y;                  // Methane flux values
-}
-parameters {
-    vector[nX] beta;
-    real<lower=0> sigma;
-}
-transformed parameters {
-  vector[N] mu;
-  mu = X*beta;
-  }
-model {
-    y ~ cauchy(mu, sigma);
-    sigma ~ normal(0, 0.1);
-}"
-
-Xmat <- model.matrix(~parental, heritability)
-data.list <- with(heritability, list(y = offspring, X = Xmat, nX = ncol(Xmat), N = nrow(heritability)))
-
-data.rstan <- stan(model_code = srr_block,
-                            data = data.list, 
-                            iter=3e3)
-
-stan_trace(data.rstan)
-
-mcmc = as.data.frame(data.rstan) %>% dplyr:::select(contains("beta"),
-                                                    sigma) %>% as.matrix
-## get median parameter estimates
-coefs = apply(mcmc[, 1:2], 2, median)
-fit = as.vector(coefs %*% t(Xmat))
-resid = heritability$offspring - fit
-ggplot() + geom_point(data = NULL, aes(y = resid, x = fit))
-newdata = heritability %>% cbind(fit, resid)
-ggplot(newdata) + geom_point(aes(y = resid, x = parental))
-
-print(data.rstan, par = c("beta", "sigma"))
-
-tidyMCMC(data.rstan, conf.int = TRUE, conf.method = "HPDinterval", pars = c("beta", "sigma"))
-
-library(bayesplot)
-posterior <- as.matrix(data.rstan)
-
-plot_title <- ggtitle("Posterior distributions",
-                      "with medians and 80% intervals")
-mcmc_areas(posterior,
-           pars = c("beta[2]"),
-           prob = 0.8) + plot_title
-```

### 2020-07-28 Manuscript/manuscript.Rmd
add p6

@@
 dev_fit <- rma(deviance~passage, se^2, method="FE", data = deviance)
-dev_fit2 <- lm(deviance~passage, data = deviance)
 
@@
   geom_pointrange(color = 'darkorange2') + 
-  #geom_abline(intercept = 0.1 + fit$beta[1], slope = fit$beta[2], color = 'blue') +
-      geom_segment(aes(x = 1, xend = 5, y = dev_fit$beta[1] + dev_fit$beta[2], yend = dev_fit$beta[1] + dev_fit$beta[2]*5), color = 'darkorange2', size = 1) +
+      geom_segment(aes(x = 1, xend = 6, y = dev_fit$beta[1] + dev_fit$beta[2], yend = dev_fit$beta[1] + dev_fit$beta[2]*5), color = 'darkorange2', size = 1) +
   labs(x = "Passage Number", y = "Deviation from Control (P - N)")+
   theme_bw()
+
 ```
@@
 Methane oxidation rate was heritable between consecutive passages (`r figures("herit", display="cite")`).
-Specifically, the narrow-sense heritability, calculated as the slope of the mid-parent and mid-offspring, was $h^2$ = `r myround(her_sum[2, 2], 2)` $\pm$ `r myround(her_sum[2, 3], 2)`.
+Specifically, the narrow-sense heritability, calculated as the slope of the mid-parent and mid-offspring, was $h^2$ = `r myround(her_sum[2, 2], 2)` $\pm$ `r myround(her_sum[2, 3], 2)` (t = `r myround(her_sum[2, 4], 2)`, p = `r myround(her_sum[2, 5], 3)`).
 With an inbreeding coefficient of `r myround(F, 2)`,

### 2020-07-28 R/process.Rmd
add p6

@@
 
-### Standard Curve
+# Standard Curves
 
-```{r sc_data, fig.height=3}
+Standard curves for each sc_date. (Some flux_dates have multiple standard curves if the instrument drifted from day to day as indicated by check standards.)
+
+```{r sc_data, fig.height=4}
 
@@
 sc_ch4 %>% 
-  left_join(sc_dates) %>% 
+  left_join(sc_dates, by = "sc_date") %>% 
   select(flux_date:sc_date, t, r.squared:slope)
@@
 
-### Flux Measurements
+# Fluxes
+
+Each column is jar by number (the jar numbers are arbitrary - there is no relationship between the communities of jar 01, passage 1 and jar 01, passage 2).
+Rows are each selection treatment (positive = p, neutral = n) within each passage (indicated by the date flux was measured).
+Lines are proportion of ppm CH4 consumed per day.
 
@@
 
-# Flux
-
 ```{r flux_curves, warning = FALSE, fig.height=8}
@@
   geom_point() + 
-  stat_smooth(method = 'lm')
+  stat_smooth(method = 'lm', formula = y ~ x)
 
@@
 
-# Fluxes
+Histograms and boxplots of flux estimates for each jar within each treatment. The numbered points give an indication of which jars were the greatest methane consumers
 
@@
 
-```{r test_response}
+# Response to Selection
+
+```{r test_response, eval=FALSE, include=FALSE}
 
@@
 
-```{r selection}
+```{r selection, eval=FALSE, include=FALSE}
 
@@
 
-```{r selection_log_plot}
+```{r selection_log_plot, eval=FALSE, include=FALSE}
 
@@
 fit <- lm(estimate ~ treat * passage, data = log_fluxes)
-summary(fit)
 plot(fit$residuals ~ log_fluxes$estimate)
@@
 
-```{r test_response_rank}
+```{r test_response_rank, eval=FALSE, include=FALSE}
 
@@
 
-```{r}
+```{r eval=FALSE, include=FALSE}
 selection_table <- tidy(anova(rank_fit))
@@
 
-```{r selection_rank}
+```{r selection_rank, eval=FALSE, include=FALSE}
 
@@
 
-```{r}
-
-gm_mean = function(x, na.rm=TRUE){
-  exp(sum(log(x[x > 0]), na.rm=na.rm) / length(x))
-}
-
-ggplot(fluxes, aes(x = estimate)) + 
-  facet_wrap(~ passage + treat) +
-  geom_dotplot(method = 'histodot') +
-  geom_vline(aes(xintercept = mean(estimate)), color = 'blue') +
-  geom_vline(aes(xintercept = median(estimate)), color = 'red') +
-    geom_vline(aes(xintercept = gm_mean(estimate)), color = 'green')
-
-```
-
 # Deviance from mean
@@
   group_by(passage, treat) %>% 
-  summarize(se = se(estimate)) %>% 
+  summarize(se = se(estimate), .groups = "drop") %>% 
   pivot_wider(names_from = treat, values_from = se) %>% 
@@
   group_by(passage, treat) %>% 
-  summarize(mean = mean(estimate)) %>% 
+  summarize(mean = mean(estimate), .groups = "drop") %>% 
   pivot_wider(names_from = treat, values_from = mean) %>% 
@@
 
-fit <- lm(deviance ~ passage, data = deviance)
-summary(fit)
-
 fit <- rma(deviance~passage,se^2,method="FE",data = deviance)
-fit
-
-summary(fit)
-plot(fit)
 
@@
 
+rma_output <- data.frame(
+Estimate = fit$beta,
+se = fit$se,
+z.value = fit$zval,
+p.value = fit$pval,
+upper.ci = fit$ci.lb,
+lower.ci = fit$ci.ub
+)
+rma_output %>% 
+  kable() %>% 
+  kable_styling()
 ```
@@
   group_by(passage, treat) %>% 
-  summarize(parental = mean(estimate)) %>% 
-  ungroup() %>% 
-  filter(passage != 5)
+  summarize(parental = mean(estimate), .groups = "drop")
 
@@
   select(treat, rep, estimate, passage)
+
+```
+
+```{r}
 
@@
   group_by(passage, treat) %>%
-  summarize(offspring = mean(estimate)) %>%
-  ungroup() %>%
+  summarize(offspring = mean(estimate), .groups = "drop") %>%
   # mutate(offspring = estimate) %>%
@@
 ggsave('heritability.png', width = 4, height = 3)
+
+```
+
+```{r}
+
 heritability$parental_sq <- heritability$parental^2

### 2020-07-31 Manuscript/manuscript.Rmd
create new process script and add G6 data and write new Intro.

@@
 
+Ecosystem functions that humans depend on are largely controlled by biodiversity in terms of the diversity and function of organisms in an ecosystem.
+Therefore, to understand the impact of human alterations of the environment and to manage environments for future generations requires that we understand the effects of species on ecosystem processes (Hooper et al.
+2005).
+The vast majority of research on the connections between biodiversity and ecosystem function have been conducted in temperate grassland ecosystems evaluating the connection between manipulated plant richness and productivity.
+This research has demonstrated a positive, asymptotic relationship between species richness and productivity.
+However, it remains unclear whether this mapping is generalizable to other functions or to other taxonomic groups.
+
+In evolutionary biology, genotype-phenotype mapping provides a way to understand the fitness effects of a genotype.
+While selection typically acts on phenotype, the genotypic changes are the fundamental driver of phenotypic differences and thus fitness differences (Stadler and Stephes 2005).
+A similar analogy can be drawn for biodiversity-ecosystem function relationships.
+Traits of a community are what drive community assembly, not the taxa or genotypes themselves (Inkpen et al.
+2017).
+However, as ecologists, we wish to understand the effects of species on ecosystem function.
+In a similar way that evolutionary biologists want to deduce the genotype-phenoype map to understand the genetic basis of fitness, community ecologists must describe the biodiversity-ecoystem function map to understand what drives communty assembly, and ultimately, ecosystem function.
+
+The relationship that ecologists have described between plant diversity and productivity represents a relatively simple map between taxonomy and function.
+The number of unique members of a functional group (i.e., plant species) is correlated with the rate of the function performed by that group.
+However, for microbial communities that mapping is not quite as clear.
+For many microbial traits, there is not a clear coupling between taxonomy and function, except for in the case of highly conserved traits involving many genes (Martiny et al.
+2013).
+In addition, most microbial taxa have not been sampled and so we don't have a full picture of which taxa are members of a particular functional group (citation).
+It's much harder to identify methanotrophic prokaryotes in soil than it is to identify photosynthetic eukaryotes in a prairie.
+Finally, even for highly conserved traits with a putative functional marker, for example methane oxidation and the gene *pmoA*, functional group abundance is not correlated with the rate of the function performed by that group in most ecosystems (Rocca et al.
+2016).
+Because of these differences between macroorganisms and microorganisms, it may be more useful to approach microbial biodiversity-ecosystem function in a way analogus to genotype-phenotype mapping in organisms (Morris et al.
+2020).
+
+One approach to deducing the genotype-phenotype map in organisms is through artificial selection (Fuller et al.
+2005).
+In this approach, selection for a trait is imposed in a contrived laboratory environment and compared to laboratory environment without selection on that trait.
+By comparing the genetic respones between these two treatments, you can remove the environment-fitness connection that would be present in a comparative study and identify genes that are evolving in response to selection only on the trait of interest.
+It has been shown that whole communities can respond to artifical selection on traits of the ecosystem (Goodnight et al.
+1997, Williams and Lenton 2007).
+And it has been proposed that this could be a suitable approach for identify communities of organisms that are capable of performing a particular function at a high rate (Swenson et al.
+2000).
+Therefore, artificial ecosystem selection is a potentially fruitful approach to identifying the map between biodiversity and ecosystem function in microbial communities.
+
+To identify the map between microbial community composition and ecosystem function, we imposed artificial ecosystem selection on methane oxidation rate of whole soil ecosystems.
+We prepared two lines of replicate ecosystems: one in which we imposed positive selection on methane oxidation by only using the top performing ecosystems to generate the next set of ecosystems and one in which we imposed neutral selection where a random set of ecosystems were chosen to generate the next set of ecosystems.
+We repated this process over multiple passages and then compared the resultant metacommunities in terms of their rate of ecosystem function and the composition of their communities.
+
+With this experiment, we asked several questions.
+First, is there a response to selection on methane oxidation by soil ecosystems?
+And how strong is the response to selection?
+(i.e., how "heritable" is methane oxidation as a community trait in terms of the phenotypic resemblence of ecosystems to the ecosystems that generated them?) If methane oxidation is heritable, then what is the nature of the map between microbial community composition and methane oxidation?
+Is it the diversity of methane oxidizers, as it is for plants and productivity?
+Or is there a more complex map involving upstream metabolic regulation?
+
+Things this Intro doesn't yet discuss:
+
+1. What a more complex map might look like. For example, maybe pmoA and methanotrophy aren't correlated, but maybe understand methanogenesis or nitrogen cycling is helpful.
+
+2. The idea of "heritability" for communities/ecosystems and why that relates to environmental vs. genetic control over ecosystem function.
+
 # Materials & Methods
@@
 Samples were immediately injected into a <mark>???</mark> gas chromatograph fitted with an electron capture detector (company, city, state, USA) to determine the headspace CH~4~ concentration.
-Fluxes were calculated from a first-order exponential decay function as decay constant `k` with units<mark>???</mark>.
+Fluxes were calculated from a first-order exponential decay function as decay constant `k`.
 Oxidation rates are presented as the additive inverse of `k` (i.e., `-k`) so that a more positive value represents a faster oxidation rate.
@@
 
-To test whether there was a significant change in CH~4~ oxidation rate (k) as a response to selection, we tested the deviation between the positive and neutral selection lines over the five passages.
-We calculated deviation as the difference in the mean oxidation rate between the positive and neutral line within each passage with standard errors equal to the sum of the errors for each of the positive and neutral lines.
+To test whether there was a significant change in CH~4~ oxidation rate (k) as a response to selection, we tested the deviation between the positive and neutral selection lines over multiple passages.
+We calculated deviation as the difference in the mean oxidation rate between the positive and neutral treatment within each passage.
+The standard errors of the deviations were calculated as the sum of the standard errors of the mean for each of the positive and neutral lines.
 We then fit an ordinary least squares regression model taking into account the standard errors of the deviations.
@@
 
+ggsave('new_select_fig.pdf', width = 4, height = 3)
+
 ```
@@
   theme_bw()
+
+ggsave('new_herit_fig.pdf', width = 5, height = 4)
 

### 2020-07-31 R/process.R

@@ -1,17 +1,4 @@
----
-title: "Artificial Ecosystem Selection: Methane Oxidation"
-author: "Andrew Morris"
-output:
-  html_document:
-      toc: true
-      toc_float: 
-        collapsed: false
-editor_options: 
-  chunk_output_type: inline
----
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE,  fig.align="center")
 library(tidyverse)
 library(knitr)
 library(kableExtra)
@@ -21,20 +8,18 @@ library(lubridate)
 library(morris)
 library(magrittr)
 library(metafor)
-```
+
 
 # Standard Curves
 
-Standard curves for each sc_date. (Some flux_dates have multiple standard curves if the instrument drifted from day to day as indicated by check standards.)
+# Standard curves for each sc_date. (Some flux_dates have multiple standard curves if the instrument drifted from day to day as indicated by check standards.)
 
-```{r sc_data, fig.height=4}
-
-sc <- read.csv('../Data/standard_curve.csv')
-sc_dates <- read.csv('../Data/sc_dates.csv')
+sc <- read.csv('Data/standard_curve.csv')
+sc_dates <- read.csv('Data/sc_dates.csv')
 
 sc %>% 
   filter(molecule == 'ch4') %>% 
-ggplot(mapping = aes(area, injection_ppm)) +
+  ggplot(mapping = aes(area, injection_ppm)) +
   facet_wrap(~ sc_date + molecule, scales = "free") + 
   geom_point() +
   stat_smooth(formula = y ~ x, method = 'lm', se = FALSE)
@@ -64,20 +49,15 @@ sc_ch4 %>%
   kable_styling()
 
 sc_ch4 <- 
-sc_ch4 %>% 
+  sc_ch4 %>% 
   left_join(sc_dates, by = "sc_date") %>% 
   select(flux_date:sc_date, t, r.squared:slope)
 
-```
-
-
 # Fluxes
 
-Each column is jar by number (the jar numbers are arbitrary - there is no relationship between the communities of jar 01, passage 1 and jar 01, passage 2). Rows are each selection treatment (positive = p, neutral = n) within each passage (indicated by the date flux was measured). Lines are proportion of ppm CH4 consumed per day.
[…]

### 2020-09-18 Manuscript/manuscript.Rmd
update deviance errors to be the standard error of the difference

@@
 Deviance is calculated as the mean of P within each passage minus the mean of N within each passage.
-Errorbars are the sums of the standard errors for P and N for each passage.
+Errorbars are the standard error of the difference between P and N for each passage.
 The best fit line is an ordinary least squares fit taking into account the standard errors of the deviations.
@@
 
-Ecosystem functions that humans depend on are largely controlled by biodiversity in terms of the diversity and function of organisms in an ecosystem.
-Therefore, to understand the impact of human alterations of the environment and to manage environments for future generations requires that we understand the effects of species on ecosystem processes (Hooper et al.
-2005).
-The vast majority of research on the connections between biodiversity and ecosystem function have been conducted in temperate grassland ecosystems evaluating the connection between manipulated plant richness and productivity.
-This research has demonstrated a positive, asymptotic relationship between species richness and productivity.
-However, it remains unclear whether this mapping is generalizable to other functions or to other taxonomic groups.
+## Determining the mapping between microbial community structure and ecosystem function
+
+Microbial communities mediate ecosystem nutrient cycling via decomposition, greenhouse gas cycling,
+carbon storage, and other processes.
+Despite the central role of bacteria in performing these processes, we have not found a consistent relationship between the composition of a microbial community and the rate of an ecosystem process across habitats.
+Typically, research in this area has followed the model of macroorganismal biodiversity-ecosystem function studies such as grassland productivity studies that measure or manipulate the diversity of plant species and correlate that with the rate of primary productivity.
+By analogy, microbial ecologists use microbial sequencing data to create taxonomic relative abundance tables and then infer diversity or relative abundance from the whole archaeal/bacterial community and correlate that with the rate of a microbial process.
+Alternatively, they estimate the abundance of a putative functional gene (such as *pmoA* for methane oxidation) and correlate abundance with function.
+These approaches only rarely show significant, positive relationships between biodiversity and ecosystem function.
+In my first chapter published in Phil Trans I lay out philosophically why this approach shouldn’t work – namely that marker gene sequences are not analogous to plant species counts, but are more analogous to measurements of gene markers as in evolutionary biology studies.
 
@@
 However, as ecologists, we wish to understand the effects of species on ecosystem function.
-In a similar way that evolutionary biologists want to deduce the genotype-phenoype map to understand the genetic basis of fitness, community ecologists must describe the biodiversity-ecoystem function map to understand what drives communty assembly, and ultimately, ecosystem function.
+Therefore, in a similar way that evolutionary biologists want to deduce the genotype-phenoype map to understand the genetic basis of fitness, community ecologists must describe the "mapping" between community genotype and ecoystem function to understand what drives communty assembly, and ultimately, ecosystem function.
 
@@
 In addition, most microbial taxa have not been sampled and so we don't have a full picture of which taxa are members of a particular functional group (citation).
-It's much harder to identify methanotrophic prokaryotes in soil than it is to identify photosynthetic eukaryotes in a prairie.
+Put another way, it's much harder to identify methanotrophic prokaryotes in soil than it is to identify plants in a prairie.
 Finally, even for highly conserved traits with a putative functional marker, for example methane oxidation and the gene *pmoA*, functional group abundance is not correlated with the rate of the function performed by that group in most ecosystems (Rocca et al.
@@
 We calculated deviation as the difference in the mean oxidation rate between the positive and neutral treatment within each passage.
-The standard errors of the deviations were calculated as the sum of the standard errors of the mean for each of the positive and neutral lines.
+The standard errors of the deviations were calculated as the square root of the sum of the squared errors of the positive and neutral lines.
 We then fit an ordinary least squares regression model taking into account the standard errors of the deviations.
@@
 fluxes <- read_tsv('../Output/fluxes.tsv')
-
-deviance_se <- 
-  fluxes %>% 
-  select(treat, passage, estimate) %>% 
-  group_by(passage, treat) %>% 
-  summarize(se = se(estimate)) %>% 
-  pivot_wider(names_from = treat, values_from = se) %>% 
-  mutate(deviance_se = p + n) %>% 
-  pull(deviance_se)
-
 deviance <-
@@
 
-deviance$se <- deviance_se
+se <- function(x, na.rm = FALSE){
+  sqrt(var(if (is.vector(x) || is.factor(x)) x else as.double(x), 
+    na.rm = na.rm)) / sqrt(length(if (na.rm == TRUE) na.omit(x) else x))
+}
+
+sed <- function(p, n, na.rm = FALSE) {
+  sqrt(se(p, na.rm = na.rm) ^ 2 + se(n, na.rm = na.rm)^ 2)
+}
+
+deviance$se <-
+  fluxes %>% 
+  select(treat, passage, estimate) %>% 
+  mutate(x = 1:nrow(fluxes)) %>% 
+  pivot_wider(names_from = treat, values_from = estimate) %>% 
+  group_by(passage) %>% 
+  summarize(sed = sed(p, n, na.rm = TRUE), .groups = "drop") %>% 
+  pull(sed)
 
@@
   group_by(passage) %>% 
-  summarize(mean = mean(estimate), se = se(estimate))
+  summarize(mean = mean(estimate), se = morris::se(estimate))
 
@@
 ```{r dev_fig, fig.width=4, fig.height=3, fig.cap=figures("deviance")}
-
+passages <- length(deviance$passage)
 ggplot(deviance, aes(x = passage, y = deviance, ymin = deviance - se, ymax = deviance + se)) + 
@@
   geom_pointrange(color = 'darkorange2') + 
-      geom_segment(aes(x = 1, xend = 6, y = dev_fit$beta[1] + dev_fit$beta[2], yend = dev_fit$beta[1] + dev_fit$beta[2]*5), color = 'darkorange2', size = 1) +
+      geom_segment(aes(x = 1, xend = passages, y = dev_fit$beta[1] + dev_fit$beta[2], yend = dev_fit$beta[1] + dev_fit$beta[2]*passages), color = 'darkorange2', size = 1) +
   labs(x = "Passage Number", y = "Deviation from Control (P - N)")+
@@
 heritability <- read_tsv('../Output/heritability.tsv')
+
 her_fit <- lm(offspring ~ parental, data = heritability)
@@
 
-```{r herit_fig, fig.width=5, fig.height=4, fig.cap=figures("herit")}
+```{r herit_fig, fig.width=5, fig.height=3, fig.cap=figures("herit")}
 ggplot(heritability, aes(x = parental, y = offspring)) +
@@
 
-ggsave('new_herit_fig.pdf', width = 5, height = 4)
+ggsave('new_herit_fig.pdf', width = 5, height = 3)
 

### 2020-09-18 R/process.Rmd
update deviance errors to be the standard error of the difference

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE,  fig.align="center")
+knitr::opts_chunk$set(echo = TRUE,  fig.align="center")
 library(tidyverse)
@@
 library(metafor)
+library(lme4)
 ```
@@
 
-sc %>% 
-  filter(molecule == 'ch4') %>% 
-ggplot(mapping = aes(area, injection_ppm)) +
-  facet_wrap(~ sc_date + molecule, scales = "free") + 
-  geom_point() +
-  stat_smooth(formula = y ~ x, method = 'lm', se = FALSE)
+# sc %>% 
+#   filter(molecule == 'ch4') %>% 
+# ggplot(mapping = aes(area, injection_ppm)) +
+#   facet_wrap(~ sc_date + molecule, scales = "free") + 
+#   geom_point() +
+#   stat_smooth(formula = y ~ x, method = 'lm', se = FALSE)
 
@@
 sc_ch4 <- left_join(sc_r2, sc_equation, by = c("flux_date", "sc_date"))
-sc_ch4 %>% 
-  kable() %>% 
-  kable_styling()
+
+# sc_ch4 %>% 
+#   kable() %>% 
+#   kable_styling()
 
@@
 
-Each column is jar by number (the jar numbers are arbitrary - there is no relationship between the communities of jar 01, passage 1 and jar 01, passage 2).
-Rows are each selection treatment (positive = p, neutral = n) within each passage (indicated by the date flux was measured).
-Lines are proportion of ppm CH4 consumed per day.
-
 ```{r time_data}
-
 time_data <- data.table::fread('../Data/time_data.csv')
@@
   mutate(flux_date = factor(flux_date))
-
 ```
@@
 ```{r}
-
 conc_data <- data.table::fread('../Data/conc_data.csv')
@@
          function(x) ln_ch0_chn(conc_data$t1, x))
+
 flux_data <-
@@
          treat = factor(substr(jar, 1, 1)))
+```
 
-```
+Each column is jar by number (the jar numbers are arbitrary - there is no relationship between the communities of jar 01, passage 1 and jar 01, passage 2).
+Rows are each selection treatment (positive = p, neutral = n) within each passage (indicated by the date flux was measured).
+Lines are proportion of ppm CH4 consumed per day.
 
@@
 ```
-
-Histograms and boxplots of flux estimates for each jar within each treatment. The numbered points give an indication of which jars were the greatest methane consumers
 
@@
   mutate(passage = as.numeric(flux_date)) %>% 
-  mutate(estimate_rank = rank(estimate))
+  mutate(estimate_rank = rank(estimate)) %>% 
+  filter(passage != 6)
 
@@
 
+Histograms and boxplots of flux estimates for each jar within each treatment. The numbered points give an indication of which jars were the greatest methane consumers
+
 ```{r plot_estimates}
-ggplot(fluxes, aes(x = estimate, fill = treat)) +
-  facet_wrap(~ flux_date, scales = "free") +
-  geom_histogram(position = 'identity', alpha = 0.5, bins = 20)
+# ggplot(fluxes, aes(x = estimate, fill = treat)) +
+#   facet_wrap(~ flux_date, scales = "free") +
+#   geom_histogram(position = 'identity', alpha = 0.5, bins = 20)
 
@@
 
-```{r test_response, eval=FALSE, include=FALSE}
-
-fluxes$passage <- as.numeric(fluxes$flux_date)
-
-fit <- lm(estimate ~ treat * passage, data = fluxes)
-plot(fit$residuals ~ fluxes$estimate)
-selection_table <- tidy(anova(fit))
-
-selection_table[1, 1] <- "Treatment"
-selection_table[2, 1] <- "Passage"
-selection_table[3, 1] <- "Treatment:Passage"
-selection_table[4, 1] <- "Residuals"
-
-selection_table %>% 
-  kable(col.names = c('Term', 'Df', 'Sum Sq', 'Mean Sq', 'F Value', 'p-value')) %>% 
-  kable_styling()
-
-```
-
-```{r selection, eval=FALSE, include=FALSE}
-
-fluxes %>% 
-  ggplot(mapping = aes(x = passage, y = estimate, color = treat)) + 
-  geom_boxplot(aes(x = passage, y = estimate, color = treat, group = interaction(treat, passage))) +
-  geom_jitter(width = 0.2) +
-  #geom_smooth(method = 'lm') +
-  labs(x = "Passage Number", y = "Methane Oxidation Rate (-k)") +
-  scale_color_discrete(name = "Selection Treatment", labels = c('Neutral', 'Positive')) +
-  theme_bw()
-
-#ggsave('selection.pdf', width = 6, height = 4)
-```
-
-```{r selection_log_plot, eval=FALSE, include=FALSE}
-
-log_fluxes <- fluxes
-#log_fluxes$estimate[log_fluxes$estimate < 0] <- 0.01
-log_fluxes$estimate <- log10(log_fluxes$estimate + abs(min(log_fluxes$estimate)) + 0.01)
-
-fit <- lm(estimate ~ treat * passage, data = log_fluxes)
-plot(fit$residuals ~ log_fluxes$estimate)
-
-selection_table <- tidy(anova(fit))
-
-#selection_table[1, 1] <- "Neutral0"
-#selection_table[2, 1] <- "Positive0"
-#selection_table[3, 1] <- "Passage#"
-#selection_table[4, 1] <- "Positive:Passage#"
-
-selection_table[1, 1] <- "Treatment"
-selection_table[2, 1] <- "Passage"
-selection_table[3, 1] <- "Treatment:Passage"
-selection_table[4, 1] <- "Residuals"
-
-selection_table %>% 
-  kable(col.names = c('Term', 'Df', 'Sum Sq', 'Mean Sq', 'F Value', 'p-value')) %>% 
-  kable_styling()
-
-log_fluxes %>% 
-  ggplot(mapping = aes(x = passage, y = estimate, color = treat)) +
-  geom_jitter(width = 0.2) +
-  geom_smooth(method = 'lm') +
-  labs(x = "Passage Number", y = "log10(Methane Oxidation Rate (-k))") +
-  scale_color_discrete(name = "Selection Treatment", labels = c('Neutral', 'Positive')) +
-  theme_bw() +
-  scale_y_log10()
-
-#ggsave('selection.pdf', width = 6, height = 4)
-
-log_fluxes %>% 
-  ggplot(mapping = aes(x = passage, y = estimate, color = treat)) +
-  geom_jitter(width = 0.2) +
-  #geom_smooth(method = 'lm') +
-  labs(x = "Passage Number", y = "Methane Oxidation Rate (-k)") +
-  scale_color_discrete(name = "Selection Treatment", labels = c('Neutral', 'Positive')) +
-  theme_bw() + 
-  geom_boxplot(aes(x = passage, y = estimate, color = treat, group = interaction(treat, passage)))
-```
-
-```{r test_response_rank, eval=FALSE, include=FALSE}
-
-fluxes$estimate_rank <- rank(fluxes$estimate)
-rank_fit <- lm(estimate_rank ~ treat * passage, data = fluxes)
-
-rank_fit
-summary(rank_fit)
-anova(rank_fit)
-
-plot(fit$residuals ~ fluxes$estimate_rank)
-
-```
-
 ```{r eval=FALSE, include=FALSE}
-selection_table <- tidy(anova(rank_fit))
-selection_table[1, 1] <- "Treatment"
-selection_table[2, 1] <- "Passage"
-selection_table[3, 1] <- "Treatment:Passage"
-selection_table[4, 1] <- "Residuals"
-selection_table %>% 
-  kable(col.names = c('Term', 'Df', 'Sum Sq', 'Mean Sq', 'F Value', 'p-value')) %>% 
-  kable_styling()
-
-selection_table <- tidy(car::Anova(rank_fit, type = "II"))
-selection_table[1, 1] <- "Treatment"
-selection_table[2, 1] <- "Passage"
-selection_table[3, 1] <- "Treatment:Passage"
-selection_table[4, 1] <- "Residuals"
-selection_table %>% 
-  kable(col.names = c('Term', 'Sum Sq', 'Df', 'F Value', 'p-value')) %>% 
-  kable_styling()
-
-selection_table <- tidy(car::Anova(rank_fit, type = "III"))
-selection_table[1, 1] <- "Intercept"
-selection_table[2, 1] <- "Treatment"
-selection_table[3, 1] <- "Passage"
-selection_table[4, 1] <- "Treatment:Passage"
-selection_table[5, 1] <- "Residuals"
-selection_table %>% 
-  kable(col.names = c('Term', 'Sum Sq', 'Df', 'F Value', 'p-value')) %>% 
-  kable_styling()
-
-selection_table <- tidy(summary(rank_fit))
-
-selection_table[1, 1] <- "Neutral"
-selection_table[2, 1] <- "Positive"
-selection_table[3, 1] <- "Passage"
-selection_table[4, 1] <- "Positive:Passage"
-
-selection_table %>% 
-  kable(col.names = c('Term', 'Estimate', 'Std Error', 't-statistic', 'p-value')) %>% 
-  kable_styling()
-
-```
-
-```{r selection_rank, eval=FALSE, include=FALSE}
-
-fluxes %>% 
-  ggplot(mapping = aes(x = passage, y = estimate_rank, color = treat)) +
-  geom_jitter(width = 0.2) +
-  geom_smooth(method = 'lm') +
-  labs(x = "Passage Number", y = "Rank Methane Oxidation Rate (-k)") +
-  scale_color_discrete(name = "Selection Treatment", labels = c('Neutral', 'Positive')) +
-  theme_bw()
-
-ggsave('selection.pdf', width = 6, height = 4)
-ggsave('selection.png', width = 6, height = 4)
-```
-
-# Deviance from mean
-
-```{r}
-deviance_se <- 
-  fluxes %>% 
-  select(treat, passage, estimate) %>% 
-  group_by(passage, treat) %>% 
-  summarize(se = se(estimate), .groups = "drop") %>% 
-  pivot_wider(names_from = treat, values_from = se) %>% 
-  mutate(deviance_se = p + n) %>% 
-  pull(deviance_se)
-
 deviance <-
@@
 
-deviance$se <- deviance_se
+se <- function(x, na.rm = FALSE){
+  sqrt(var(if (is.vector(x) || is.factor(x)) x else as.double(x), 
+    na.rm = na.rm)) / sqrt(length(if (na.rm == TRUE) na.omit(x) else x))
+}
 
-fit <- rma(deviance~passage,se^2,method="FE",data = deviance)
+sed <- function(p, n, na.rm = FALSE) {
+  sqrt(se(p, na.rm = na.rm) ^ 2 + se(n, na.rm = na.rm)^ 2)
+}
+
+deviance$se <-
+  fluxes %>% 
+  select(treat, passage, estimate) %>% 
+  mutate(x = 1:nrow(fluxes)) %>% 
+  pivot_wider(names_from = treat, values_from = estimate) %>% 
+  group_by(passage) %>% 
+  summarize(sed = sed(p, n, na.rm = TRUE), .groups = "drop") %>% 
+  pull(sed)
+
+fit <- rma(deviance ~ passage, se^2, method="FE", data = deviance)
 
@@
   kable_styling()
+
+write.csv(deviance, 
 ```
@@
 
+# Mean parental (selected) flux for each passage
 parental <-
@@
 
+# Mean offspring flux for each passage
+offspring <-
+  offspring %>% 
+  group_by(passage, treat) %>%
+  summarize(offspring = mean(estimate), .groups = "drop") %>%
+  filter(passage != 1) %>% 
+  mutate(passage = passage - 1)
 ```
@@
 
-ggplot(offspring, aes(x = estimate, fill = factor(passage))) +
-  geom_histogram(position = "identity", alpha = 0.5)
-
-offspring <-
-  offspring %>% 
-  group_by(passage, treat) %>%
-  summarize(offspring = mean(estimate), .groups = "drop") %>%
-  # mutate(offspring = estimate) %>%
-  filter(passage != 1) %>% 
-  mutate(passage = passage - 1)
+###################################################
+# This is the nest place you stopped
+# You need to calculate the slope of midoffspring
+# on midparent for the Positive treatments to get R
+# and then calculate the selection differential
+# as above S. S doesn't make sense for the neutral
+# treatment.
+###################################################
 
@@
 heritability$treat <- factor(heritability$treat)
+heritability <-
+  filter(heritability, passage != 5)
 # heritability$parental <- rank(heritability$parental)
@@
   stat_smooth(method = 'lm') + 
-  labs(x = "Parental Median Flux (-k)", y = "Offspring Flux (-k)") +
+  labs(x = "Parental Mean Flux (-k)", y = "Offspring Flux (-k)") +
   scale_color_discrete(name = "Selection\nTreatment", labels = c('Neutral', 'Positive')) +
@@
 
-ggsave('heritability.pdf', width = 4, height = 3)
-ggsave('heritability.png', width = 4, height = 3)
+# ggsave('heritability.pdf', width = 4, height = 3)
+# ggsave('heritability.png', width = 4, height = 3)
 
 ```
+```{r}
+# This chunk calculates the slope of mid-offspring on mid-parent for all generations and all treatments
 
-```{r}
+fluxes
+```
 
-heritability$parental_sq <- heritability$parental^2
-fit <- lm(sqrt(offspring) ~ parental, data = heritability)
+```{r breeders}
 
-summary(fit)
-tidy(fit) %>% 
-  kable() %>% 
-  kable_styling()
-glance(fit) %>% 
-  kable() %>% 
-  kable_styling()
-ggplot(mapping = aes(fit$residuals, fit$fitted.values)) +
+h2 <- function(R, S) {
+  R / S
+}
+
+# R = Response to selection, difference in mean between offspring of selected parents and all of the parents
+
+# S = Selection differental, difference in mean between selected parents and all of the parents
+
+parental_all <-
+  fluxes %>% 
+  select(passage, treat, estimate) %>% 
+  group_by(passage, treat) %>% 
+  summarize(all = mean(estimate), .groups = "drop")
+
+parental_selected <- parental
+
+selection_diff <- 
+  left_join(parental_selected, parental_all) %>% 
+  rename(selected = parental) %>% 
+  mutate(S = selected - all)
+
+heritability_ns <- 
+  parental_all %>% 
+  left_join(parental_selected) %>% 
+  left_join(offspring) %>% 
+  filter(passage != 5)
+
+heritability_only_p <- 
+  heritability_ns %>% 
+  filter(treat == 'p')
+
+R <- tidy(lm(offspring ~ all, data = heritability_only_p))[[2, 2]]
+S <- tidy(lm(parental ~ all, data = heritability_only_p))[[2, 2]]
+h2(R, S)
+
+# Generational R, S, global h2, 0.82
+heritability_only_p %>% 
+  mutate(R = offspring - all,
+         S = parental - all) %>% 
+  summarize(R = mean(R),
+            S = mean(S)) %>%
+  mutate(h2 = h2(R, S))
+
+# Generational R, S, h2, 0.73
+heritability_only_p %>% 
+  #filter(passage != 2) %>% 
+  mutate(R = offspring - all,
+         S = parental - all,
+         h2 = h2(R, S)) %>% 
+  summarize(se = se(h2),
+            h2 = mean(h2))
+
+# Global R, S, h2, 0.82
+heritability_only_p %>% 
+  summarize(offspring = mean(offspring),
+         all = mean(all),
+         parental = mean(parental)) %>% 
+  mutate(R = offspring - all,
+         S = parental - all,
+         h2 = h2(R, S))
+heritability_ns %>% 
+  pivot_longer(all:offspring, names_to = "group", values_to = "mean") %>% 
+ggplot(., aes(x = passage, y = mean, color = group)) +
+  facet_wrap(~ treat) +
   geom_point()
-ggplot(mapping = aes(fit$residuals, heritability$parental)) +
-  geom_point()
-
-ggplot(heritability, aes(x = parental, y = offspring)) +
-  geom_point()
-
-ggplot(heritability, aes(x = offspring)) +
-  geom_histogram(position = 'identity', alpha = 0.5)
-
-ggplot(heritability, aes(x = parental)) +
-  geom_histogram(position = 'identity', alpha = 0.5)
-
 ```

### 2020-09-18 R/process.R

@@ -155,6 +155,13 @@ ggplot(fluxes, aes(x = estimate, fill = treat)) +
   facet_wrap(~ flux_date, scales = "free") +
   geom_histogram(position = 'identity', alpha = 0.5, bins = 20)
 
+fluxes %>% 
+  filter(passage == 4) %>% 
+  ggplot(., aes(x = estimate)) +
+  geom_histogram(bins = 10, fill = "darkorange2") +
+  labs(x = "Methane oxidation rate (-k)", y = "Count") +
+  theme_bw()
+
 ggplot(fluxes, aes(y = estimate, x = treat, color = treat)) +
   facet_wrap(~ flux_date, scales = "free") +
   geom_boxplot(outlier.shape = NA) +
@@ -163,7 +170,43 @@ ggplot(fluxes, aes(y = estimate, x = treat, color = treat)) +
 
 # Response to Selection
 
-fluxes$passage <- as.numeric(fluxes$flux_date)
+# Mixed model response to selection
+
+library(lme4)
+
+fit_full <- lmer(estimate ~ passage * treat + (passage|treat), data = fluxes)
+fit_no_int <- lmer(estimate ~ passage * treat + (0+passage|treat), data = fluxes)
+anova(fit_no_int, fit_full, test = "LRT")
+fit_lm <- lm(estimate ~ passage * treat, data = fluxes)
+anova(fit_no_int, fit_lm)
+summary(fit)
+fit_lm_no_interaction <- lm(estimate ~ passage + treat, data = fluxes)
+anova(fit_lm, fit_lm_no_interaction, test = "LRT")
+coef(fit)
+tidy(fit)
+fit_no_treat <- lm(estimate ~ passage, data = fluxes)
+fit_treat <- lm(estimate ~ passage * treat, data = fluxes)
+anova(fit_no_treat, fit_treat, test = "LRT")
+tidy(fit_treat)
+
+ggplot(fluxes, aes(x = passage, y = estimate, color = treat)) +
+  theme_bw() +
+  geom_point() +
+  scale_color_manual(values = c('gray60', 'darkorange2')) +
+  geom_abline(slope = tidy(fit_treat)[[2, 2]], intercept = tidy(fit_treat)[[1, 2]],
+              color = "gray60") +
+  geom_abline(slope = tidy(fit_treat)[[4, 2]], intercept = tidy(fit_treat)[[3, 2]],
[…]

### 2020-10-14 clean.R
clean up old files

@@ -1,20 +1,0 @@
-library(tidyverse)
-library(lubridate)
-
-time <- read.csv('eg_time_data.csv')
-
-data_t0 <-
-  data.frame(flux_date = unique(time$flux_date),
-             t0 = c('2019-05-27 11:51:00', '2019-06-07 10:41:00', '2019-06-14 14:33:00'))
-
-as.days <- function(start, end) {as.numeric(interval(start, end))/60/60/24}
-time_data  <- 
-time %>% 
-  spread(t, time) %>% 
-  left_join(data_t0, by = 'flux_date') %>% 
-  select(flux_date:jar, t0, t1:t4)
-
-conc <- read.csv('eg_conc_data.csv')
-conc %>% 
-  select(-co2_area) %>% 
-  spread(t, ch4_area)

### 2020-10-15 Manuscript/manuscript.Rmd
Update Makefile, remove old scripts, create new numbered scripts

@@
 output:
-  html_document: default
   word_document:
@@
 her_fit <- lm(offspring ~ parental, data = heritability)
-her_sum <- tidy(summary(her_fit))
+her_sum <- tidy(her_fit)
 her_aov <- tidy(anova(her_fit))

### 2020-10-15 R/01-process-flux.R

@@ -1,0 +1,129 @@
+# Input: standard_curve.csv, sc_dates.csv, time_data.csv, conc_data.csv, selected.csv
+# Output: fluxes.tsv
+
+
+library(tidyverse)
+#library(knitr)
+#library(kableExtra)
+library(broom)
+#library(modelr)
+library(lubridate)
+#library(morris)
+#library(magrittr)
+#library(metafor)
+#library(lme4)
+
+# Standard Curves
+
+sc <- read.csv('../Data/standard_curve.csv')
+sc_dates <- read.csv('../Data/sc_dates.csv')
+
+sc_model <- function(df) {lm(injection_ppm ~ area, data = df)}
+standard_curves <- 
+  sc %>% 
+  filter(molecule == 'ch4') %>% 
+  group_by(flux_date, sc_date) %>% 
+  nest() %>% 
+  mutate(model = map(data, sc_model),
+         glance = map(model, glance),
+         tidy = map(model, tidy)) 
+sc_r2 <- 
+  standard_curves %>% 
+  unnest(glance) %>% 
+  select(flux_date, sc_date, r.squared)
+sc_equation <- 
+  standard_curves %>% 
+  unnest(tidy) %>%
+  select(flux_date, sc_date, term, estimate) %>% 
+  spread(term, estimate) %>% 
+  rename(intercept = `(Intercept)`, slope = area)
+sc_ch4 <- left_join(sc_r2, sc_equation, by = c("flux_date", "sc_date"))
+
+sc_ch4 <- 
+  sc_ch4 %>% 
+  left_join(sc_dates, by = "sc_date") %>% 
+  select(flux_date:sc_date, t, r.squared:slope)
+
+
+# Fluxes
+
+time_data <- data.table::fread('../Data/time_data.csv')
+
+time_data[, c('t0', 't1', 't2', 't3', 't4', 't5')] <- 
+  lapply(time_data[, c('t0', 't1', 't2', 't3', 't4', 't5')],
+         ymd_hms)
+
+as.days <- function(start, end) {as.numeric(interval(start, end))/60/60/24}
+
+time_data[, c('t1', 't2', 't3', 't4', 't5')] <- 
+  lapply(time_data[, c('t1', 't2', 't3', 't4', 't5')],
+         function(x) as.days(time_data$t1, x))
+time_data <- 
+ time_data %>% 
+  select(-t0) %>% 
[…]

### 2020-10-15 R/02-process-herit.R

@@ -1,0 +1,55 @@
+# input: fluxes.tsv, selected.csv
+# output: heritability.tsv
+
+library(tidyverse)
+
+selected_jars <- read_csv('../Data/selected.csv')
+
+# Pull out the selected parental jars
+selected <- 
+  fluxes[paste0(fluxes$passage, fluxes$jar) %in% paste0(selected_jars$passage, selected_jars$jar), ] %>%
+  select(treat, estimate, passage)
+
+# Pull out all parental jars
+parental <-
+  fluxes %>% 
+  select(passage, treat, estimate)
+
+# Pull out all offspring jars
+offspring <- 
+  fluxes %>%
+  select(treat, estimate, passage) %>% 
+  filter(passage != 1) %>% 
+  mutate(passage = passage - 1)
+```
+
+
+```{r heritability_mean}
+
+# Calculate means for east generation
+parental_mean <-
+  parental %>% 
+  group_by(passage, treat) %>% 
+  summarize(parental = mean(estimate), .groups = "drop")
+
+offspring_mean <-
+  offspring %>% 
+  group_by(passage, treat) %>% 
+  summarize(offspring = mean(estimate), .groups = "drop")
+
+selected_mean <-
+  selected %>% 
+  group_by(passage, treat) %>% 
+  summarize(selected = mean(estimate), .groups = "drop")
+
+# Combine all parent/offspring data into one data frame
+heritability <- 
+  parental_mean %>% 
+  left_join(offspring_mean, by = c("passage", "treat")) %>% 
+  left_join(selected_mean, by = c("passage", "treat")) %>% 
+  #filter(treat == "p") %>% 
+  filter(passage != 5) %>% # Remove passage 5, which has no offspring
+  ungroup()
+
+
+write_tsv(heritability, '../Output/heritability.tsv')

### 2020-10-15 R/03-response.R

@@ -1,0 +1,91 @@
+# input: fluxes.tsv
+# output: deviance.tsv, deviance figure
+
+library(tidyverse)
+library(metafor)
+
+# Response to Selection
+
+fluxes <- read_tsv('../Output/fluxes.tsv')
+
+# Calculate deviance as p - n
+deviance <-
+  fluxes %>% 
+  select(treat, passage, estimate) %>% 
+  group_by(passage, treat) %>% 
+  summarize(mean = mean(estimate), .groups = "drop") %>% 
+  pivot_wider(names_from = treat, values_from = mean) %>% 
+  mutate(deviance = p - n)
+
+# Create se function that can handle NAs
+se <- function(x, na.rm = FALSE){
+  sqrt(var(if (is.vector(x) || is.factor(x)) x else as.double(x), 
+    na.rm = na.rm)) / sqrt(length(if (na.rm == TRUE) na.omit(x) else x))
+}
+
+# Standard error of the difference function
+sed <- function(p, n, na.rm = FALSE) {
+  sqrt(se(p, na.rm = na.rm) ^ 2 + se(n, na.rm = na.rm)^ 2)
+}
+
+# Calculate variance of deviance as SED
+deviance$se <-
+  fluxes %>% 
+  select(treat, passage, estimate) %>% 
+  mutate(x = 1:nrow(fluxes)) %>% 
+  pivot_wider(names_from = treat, values_from = estimate) %>% 
+  group_by(passage) %>% 
+  summarize(sed = sed(p, n, na.rm = TRUE), .groups = "drop") %>%
+  pull(sed)
+
+# Fit deviance model
+fit <- rma(deviance ~ passage, se^2, method="FE", data = deviance)
+
+# Plot deviance
+ggplot(deviance, aes(x = passage, y = deviance, ymin = deviance - se, ymax = deviance + se)) + 
+  geom_pointrange(color = 'darkorange2') + 
+  geom_abline(intercept = fit$beta[1], slope = fit$beta[2])
+
+# Take deviance model parameters and export into a data.frame
+rma_output <- data.frame(
+Estimate = fit$beta,
+se = fit$se,
+z.value = fit$zval,
+p.value = fit$pval,
+upper.ci = fit$ci.lb,
+lower.ci = fit$ci.ub
+)
+
+# Fit deviance model without passage 2
+fit_no2 <- 
+filter(deviance, passage != 2) %>% 
[…]

### 2020-10-15 R/04-heritability.R

@@ -1,0 +1,183 @@
+# Inpot: selected.csv, fluxes.tsv
+# Output: heritability, heritability plot???
+
+
+# Heritability
+
+```{r create_heritability_data, message=FALSE, warning=FALSE}
+selected_jars <- read_csv('../Data/selected.csv')
+
+# Pull out the selected parental jars
+selected <- 
+  fluxes[paste0(fluxes$passage, fluxes$jar) %in% paste0(selected_jars$passage, selected_jars$jar), ] %>%
+  select(treat, estimate, passage)
+
+# Pull out all parental jars
+parental <-
+  fluxes %>% 
+  select(passage, treat, estimate)
+
+# Pull out all offspring jars
+offspring <- 
+  fluxes %>%
+  select(treat, estimate, passage) %>% 
+  filter(passage != 1) %>% 
+  mutate(passage = passage - 1)
+```
+
+
+```{r heritability_mean}
+
+# Calculate means for east generation
+parental_mean <-
+  parental %>% 
+  group_by(passage, treat) %>% 
+  summarize(parental = mean(estimate), .groups = "drop")
+
+offspring_mean <-
+  offspring %>% 
+  group_by(passage, treat) %>% 
+  summarize(offspring = mean(estimate), .groups = "drop")
+
+selected_mean <-
+  selected %>% 
+  group_by(passage, treat) %>% 
+  summarize(selected = mean(estimate), .groups = "drop")
+
+# Combine all parent/offspring data into one data frame
+heritability <- 
+  parental_mean %>% 
+  left_join(offspring_mean, by = c("passage", "treat")) %>% 
+  left_join(selected_mean, by = c("passage", "treat")) %>% 
+  #filter(treat == "p") %>% 
+  filter(passage != 5) %>% # Remove passage 5, which has no offspring
+  ungroup()
+
+
+write_tsv(heritability, '../Output/heritability.tsv')
+
+```
+
+
+``` {r generation_h2}
+
+
+######################################################################
+# This is where you stopped
+# Figure out why h2 is different from h2_calc_method
+# Finish writing your email to brendan about heritability
+######################################################################
+
+# Function for the Breeder's equation
+
+breeders <- function(R, S) {
+  R / S
+}
+
[…]

### 2020-10-15 R/exp_decay.R

@@ -1,27 +1,0 @@
-library(tidyverse)
-library(broom)
-
-t = 1:100
-y1 = 22 + (53 - 22) * exp(-0.02 * t) %>% jitter(10)
-y2 = 24 + (60 - 22) * exp(-0.01 * t) %>% jitter(10)
-
-
-df <- data.frame(t = t, y = y1, sensor = 'sensor1') %>% 
-  rbind(. , data.frame(t = t, y = y2, sensor = 'sensor2'))
-sensor1 <- df %>% filter(sensor == 'sensor1')
-
-
-qplot(t, y, data = df, colour = sensor)
-nls(y ~ yf + (y0 - yf) * exp(-alpha * t), 
-    data = sensor1,
-    start = list(y0 = 54, yf = 25, alpha = 1))
-fit <- nls(y ~ SSasymp(t, yf, y0, log_alpha), data = sensor1)
-fit
-qplot(t, y, data = augment(fit)) + geom_line(aes(y = .fitted))
-df %>% 
-  group_by(sensor) %>% 
-  do(fit = nls(y ~ SSasymp(t, yf, y0, log_alpha), data = .)) %>% 
-  tidy(fit) %>% 
-  select(sensor, term, estimate) %>% 
-  spread(term, estimate) %>% 
-  mutate(alpha = exp(log_alpha))

### 2020-10-15 R/fluxes.R

@@ -1,94 +1,0 @@
-library(tidyverse)
-library(lubridate)
-library(broom)
-
-top_jars_gen1 <- c(2, 3, 13, 16)
-
-temp <- list.files(path="../Data", pattern="FluxData*", full.names = TRUE)
-conc <- sapply(temp, read.csv, stringsAsFactors = FALSE,
-                 simplify = FALSE, USE.NAMES = TRUE)
-names(conc) <- substr(names(conc), 18, 25)
-conc <- bind_rows(conc, .id = "Date")
-
-# Code transfers
-conc$transfer <- as.character(match(conc$Date, unique(conc$Date)))
-
-
-# Jar broke so remove # 9 from gen 1
-conc <- conc[!(conc$JAR == 9 & conc$Date == "20190527"), ]
-
-# Jar # 1 and 2 from gen 2 had impossible ch4 values so remove
-conc <- conc[!(conc$JAR == 1 & conc$Date == "20190607"), ]
-conc <- conc[!(conc$JAR == 2 & conc$Date == "20190607"), ]
-
-# Convert clock time to days and set t0 to 0 days
-conc <- 
-  conc  %>% 
-  mutate(days = period_to_seconds(hms(paste0(conc$TIME, ':00'))) / 60 / 60) %>% 
-  group_by(Date, JAR) %>% 
-  mutate(days = days - min(days))
-
-# Convert ppm to umol
-conc$ch4_umol <- conc$CH4_CONC * 0.965 / ((0.08206 * (293)))
-
-# Calculate fluxes from lm slopes
-flux <- 
-  conc %>% 
-  group_by(transfer, JAR) %>%
-  do(tidy(lm(ch4_umol ~ days, data = .))) %>% 
-  filter(term == 'days')  %>% 
-  select(transfer, JAR, flux_umold = estimate)
-
-# Identify the jars that were used for selection
-flux$selected[flux$JAR %in% top_jars_gen1 & flux$transfer == '1'] <- T
-flux$selected[!(flux$JAR %in% top_jars_gen1) | flux$transfer != '1'] <- F
-
-all_par <- flux$flux_umold[flux$transfer == '1']
-offspr <- flux$flux_umold[flux$transfer == '2']
-sel_par <- flux$flux_umold[flux$transfer == '1' & flux$selected == T]
-# Plot fluxes
-
-ggplot(flux, aes(x = flux_umold, fill = selected)) +
-  geom_histogram(aes(fill = interaction(transfer, selected))) +
-  facet_grid(transfer ~ .) +
[…]

### 2020-10-15 R/oxidation.R

@@ -1,61 +1,0 @@
-
-##### load packages
-library(tidyverse)
-library(lubridate)
-library(stringr)
-library(broom)
-
-#### import and cleaning data
-prod <- read_tsv('data/production.tsv') # import data
-prod <- prod[complete.cases(prod), ] # remove missing data
-
-prod <- 
-	prod %>% 
-	mutate(rate = factor(substr(vial, 1, 1))) %>% 
-	mutate(date = factor(date)) %>% 
-	mutate(rep = str_sub(vial, start=-1)) %>% 
-	mutate(seconds = period_to_seconds(hms(time))) %>% 
-	mutate(days = seconds / 60 / 60 / 24) %>% 
-  mutate(days = days - min(days)) %>%  
-	filter(vial != 'check_std')
-
-levels(prod$date) <- c('Day 9', 'Day 21')
-
-ggplot(prod, aes(x = days, y = conc, group = vial, color = rate)) +
-  geom_point() +
-  facet_wrap(~ date) +
-  stat_smooth(method = 'lm', se = F)
-
-fluxes <- prod %>% 
-  group_by(vial, date) %>%
-  do(fit = lm(conc ~ days, .)) %>% 
-  tidy(fit) %>% 
-  select(vial, date, term, estimate) %>% 
-  filter(term == 'days')
-
-flux_sum <- fluxes %>% 
-  mutate(size = substr(vial, 1, 1)) %>% 
-  group_by(size, date) %>% 
-  summarize_at(vars(estimate), funs(mean, sd))
-
-ggplot(flux_sum, aes(y = mean, ymax = mean + sd, 
-                     ymin = mean - sd, x = size, color = size)) +
-	geom_pointrange() +
-	facet_wrap(~ date) +
-	theme_classic() +
-  labs(y = expression('Methane Flux' ~ (mu* L ~ L ^ -1 ~ day ^ -1))) +
-  scale_color_discrete(name = 'Inoculum\nSize (%)', labels = c('0.1', '1.0', '5.0')) +
-  scale_x_discrete(name = 'Inoculum Size (%)', labels = c('0.1', '1.0', '5.0'))
-ggsave(file = 'inoculum_size.pdf', width = 5, height = 4)
-fluxes <- fluxes %>% 
-  mutate(size = substr(vial, 1, 1))
-ggplot(fluxes, aes(y = estimate, x = size)) +
-  geom_boxplot() +
-	geom_point() +
[…]

### 2020-10-15 R/power_analysis.R

@@ -1,30 +1,0 @@
-## Power Analysis
-## Andrew Morris
-## 10/8/18
-
-library(tidyverse)
-library(pwr)
-
-sam_siz <- NULL
-eff_siz <- .10
-sig <- 0.05
-pow = 0.80
-
-
-pwr.t2n.test(n1 = 8, n2 = 8*3, power = 0.8, sig.level = 0.05, d = NULL)
-# d = 1.2
-
-pwr.t2n.test(n1 = 8, n2 = 8, power = 0.8, sig.level = 0.05, d = NULL)
-# d = 1.5
-
-pwr.t.test(n = NULL, power = 0.8, sig.level = 0.05, d = 0.5)
-
-pwr.anova.test(k = 4, n = 8, f = NULL, sig.level = 0.05, power = 0.8)
-
-pwr.t.test(n = 8, d = NULL, sig.level = 0.05, power = 0.8)
-
-qplot(rlnorm(1000, meanlog = 1.6))
-
-(2.25-(2.25*1.75))/ 1.11
-(20-19)/10
-20-(1.5*10)

### 2020-10-15 R/process.R

@@ -1,335 +1,0 @@
-
-library(tidyverse)
-library(knitr)
-library(kableExtra)
-library(broom)
-library(modelr)
-library(lubridate)
-library(morris)
-library(magrittr)
-library(metafor)
-
-
-# Standard Curves
-
-# Standard curves for each sc_date. (Some flux_dates have multiple standard curves if the instrument drifted from day to day as indicated by check standards.)
-
-sc <- read.csv('Data/standard_curve.csv')
-sc_dates <- read.csv('Data/sc_dates.csv')
-
-sc %>% 
-  filter(molecule == 'ch4') %>% 
-  ggplot(mapping = aes(area, injection_ppm)) +
-  facet_wrap(~ sc_date + molecule, scales = "free") + 
-  geom_point() +
-  stat_smooth(formula = y ~ x, method = 'lm', se = FALSE)
-
-sc_model <- function(df) {lm(injection_ppm ~ area, data = df)}
-standard_curves <- 
-  sc %>% 
-  filter(molecule == 'ch4') %>% 
-  group_by(flux_date, sc_date) %>% 
-  nest() %>% 
-  mutate(model = map(data, sc_model),
-         glance = map(model, glance),
-         tidy = map(model, tidy)) 
-sc_r2 <- 
-  standard_curves %>% 
-  unnest(glance) %>% 
-  select(flux_date, sc_date, r.squared)
-sc_equation <- 
-  standard_curves %>% 
-  unnest(tidy) %>%
-  select(flux_date, sc_date, term, estimate) %>% 
-  spread(term, estimate) %>% 
-  rename(intercept = `(Intercept)`, slope = area)
-sc_ch4 <- left_join(sc_r2, sc_equation, by = c("flux_date", "sc_date"))
-sc_ch4 %>% 
-  kable() %>% 
-  kable_styling()
-
-sc_ch4 <- 
-  sc_ch4 %>% 
-  left_join(sc_dates, by = "sc_date") %>% 
-  select(flux_date:sc_date, t, r.squared:slope)
-
-# Fluxes
-
[…]

### 2020-10-15 R/stan.R

@@ -1,100 +1,0 @@
-library(rstan)
-library(tidyverse)
-
-data <- read_tsv('./Output/fluxes.tsv')
-rank_fit <- lm(estimate_rank ~ treat * passage, data = data)
-
-sel_sum <- tidy(summary(rank_fit))
-sel_aov <- tidy(anova(rank_fit))
-
-ggplot(data, aes(y = estimate, x = passage, color = treat)) +
-  geom_point() +
-  geom_smooth(method = "lm", se = FALSE)
-
-srr_block <- "
-data { 
-    int<lower=1> N;               // Number of observations
-    int<lower=1> nX;               // Number of predictors
-    matrix [N,nX] X;
-    vector[N] y;                  // Methane flux values
-}
-parameters {
-    vector[nX] beta;
-    real<lower=0> sigma;
-}
-transformed parameters {
-  vector[N] mu;
-
-  mu = X*beta;
-  }
-model {
-    y ~ cauchy(mu, sigma);
-    sigma ~ normal(0, 10);
-}"
-
-Xmat <- model.matrix(~treat * passage, data)
-data.list <- with(data, list(y = estimate, X = Xmat, nX = ncol(Xmat), N = nrow(data)))
-
-data.rstan <- stan(model_code = srr_block,
-                            data = data.list, 
-                            iter=1e3)
-
-print(data.rstan, par = c("beta", "sigma"))
-
-stan_trace(data.rstan)
-
-mcmc = as.data.frame(data.rstan) %>% dplyr:::select(contains("beta"),
-                                                    sigma) %>% as.matrix
-# generate a model matrix
-newdata = data
-Xmat = model.matrix(~treat * passage, newdata)
-## get median parameter estimates
-coefs = apply(mcmc[, 1:4], 2, median)
-fit = as.vector(coefs %*% t(Xmat))
-resid = data$estimate - fit
-ggplot() + geom_point(data = NULL, aes(y = resid, x = fit))
-
-
-mcmc = as.data.frame(data.rstan) %>% dplyr:::select(contains("beta"),
-                                                    sigma) %>% as.matrix
-# generate a model matrix
-newdata = newdata
-Xmat = model.matrix(~treat * passage, newdata)
-## get median parameter estimates
-coefs = apply(mcmc[, 1:4], 2, median)
-fit = as.vector(coefs %*% t(Xmat))
-resid = data$estimate - fit
[…]

### 2020-10-15 Manuscript/manuscript.Rmd
move model out of manuscript.Rmd, update makefile

@@
 fluxes <- read_tsv('../Output/fluxes.tsv')
-deviance <-
-  fluxes %>% 
-  select(treat, passage, estimate) %>% 
-  group_by(passage, treat) %>% 
-  summarize(mean = mean(estimate), .groups = "drop") %>% 
-  pivot_wider(names_from = treat, values_from = mean) %>% 
-  mutate(deviance = p - n)
-
-se <- function(x, na.rm = FALSE){
-  sqrt(var(if (is.vector(x) || is.factor(x)) x else as.double(x), 
-    na.rm = na.rm)) / sqrt(length(if (na.rm == TRUE) na.omit(x) else x))
-}
-
-sed <- function(p, n, na.rm = FALSE) {
-  sqrt(se(p, na.rm = na.rm) ^ 2 + se(n, na.rm = na.rm)^ 2)
-}
-
-deviance$se <-
-  fluxes %>% 
-  select(treat, passage, estimate) %>% 
-  mutate(x = 1:nrow(fluxes)) %>% 
-  pivot_wider(names_from = treat, values_from = estimate) %>% 
-  group_by(passage) %>% 
-  summarize(sed = sed(p, n, na.rm = TRUE), .groups = "drop") %>% 
-  pull(sed)
-
-dev_fit <- rma(deviance~passage, se^2, method="FE", data = deviance)
+deviance <- read_tsv('../Output/deviance.tsv')
+rma_output <- readRDS('../Output/response.rds')
+dev_fit <- readRDS('../Output/dev_fit.rds')
+dev_pred <- predict(dev_fit)
 
@@
 
-dev_pred <- predict(dev_fit)
+passages <- length(deviance$passage)
+
 ```
@@
 ```{r dev_fig, fig.width=4, fig.height=3, fig.cap=figures("deviance")}
-passages <- length(deviance$passage)
 ggplot(deviance, aes(x = passage, y = deviance, ymin = deviance - se, ymax = deviance + se)) + 
@@
   theme_bw()
-
-ggsave('new_select_fig.pdf', width = 4, height = 3)
 

### 2020-10-15 R/02-process-herit.R

@@ -4,6 +4,7 @@
 library(tidyverse)
 
 selected_jars <- read_csv('../Data/selected.csv')
+fluxes <- read_tsv('../Output/fluxes.tsv')
 
 # Pull out the selected parental jars
 selected <- 
@@ -21,10 +22,6 @@ offspring <-
   select(treat, estimate, passage) %>% 
   filter(passage != 1) %>% 
   mutate(passage = passage - 1)
-```
-
-
-```{r heritability_mean}
 
 # Calculate means for east generation
 parental_mean <-

### 2020-10-15 R/03-response.R

@@ -64,15 +64,15 @@ rma(deviance ~ passage, se^2, method="FE", data = .)
 # Take deviance model withou passage 2 parameters and export into a data.frame
 
 rma_output_no2 <- data.frame(
-Estimate = fit$beta,
-se = fit$se,
-z.value = fit$zval,
-p.value = fit$pval,
-upper.ci = fit$ci.lb,
-lower.ci = fit$ci.ub
+Estimate = fit_no2$beta,
+se = fit_no2$se,
+z.value = fit_no2$zval,
+p.value = fit_no2$pval,
+upper.ci = fit_no2$ci.lb,
+lower.ci = fit_no2$ci.ub
 )
 
-# R (response to selection) for the no2 model
+# R (response to selection) for the model w/o sample 2
 R_no2 <- rma_output_no2[[2, 1]]
 R <- rma_output[[2, 1]]
 
@@ -82,10 +82,6 @@ ggplot(., aes(x = passage, y = deviance, ymin = deviance - se, ymax = deviance +
   geom_pointrange(color = 'darkorange2') + 
   geom_abline(intercept = fit_no2$beta[1], slope = fit_no2$beta[2])
 
-# Print the full model
-rma_output %>% 
-  kable() %>% 
-  kable_styling()
-
-write.csv(deviance, '../Output/deviance.tsv')
-```
+write_tsv(deviance, '../Output/deviance.tsv')
+saveRDS(rma_output, '../Output/response.rds')
+saveRDS(fit, '../Output/dev_fit.rds')

### 2020-10-16 R/04-heritability.R
update heritability script, add clean to makefile

@@ -1,10 +1,12 @@
 # Inpot: selected.csv, fluxes.tsv
-# Output: heritability, heritability plot???
+# Output: heritability.tsv
 
+library(tidyverse)
+library(morris)
+
+fluxes <- read_tsv('../Output/fluxes.tsv')
 
 # Heritability
-
-```{r create_heritability_data, message=FALSE, warning=FALSE}
 selected_jars <- read_csv('../Data/selected.csv')
 
 # Pull out the selected parental jars
@@ -23,10 +25,6 @@ offspring <-
   select(treat, estimate, passage) %>% 
   filter(passage != 1) %>% 
   mutate(passage = passage - 1)
-```
-
-
-```{r heritability_mean}
 
 # Calculate means for east generation
 parental_mean <-
@@ -55,11 +53,6 @@ heritability <-
 
 
 write_tsv(heritability, '../Output/heritability.tsv')
-
-```
-
-
-``` {r generation_h2}
 
 
 ######################################################################
@@ -106,9 +99,6 @@ h2_no2 <-
   mutate(method = "calc p no 2")
 
 h2 <- rbind(h2, h2_no2)
-```
-
-```{r h2_slope_method}
 
 # Calculate S (selection differential) with/without passage 2
 
@@ -160,9 +150,6 @@ ggplot(data = heritability_scaled, aes(x = selected, y = offspring)) +
     geom_text(label = "S", x = S, y = -0.2)
 
 
-```
-
-```{r}
 heritability %>% 
   ggplot(aes(parental, offspring, color = treat)) +
   geom_point() +
@@ -171,13 +158,11 @@ heritability %>%
   scale_color_discrete(name = "Treatment", labels = c('Neutral', 'Positive'))
 
 summary(lm(offspring ~ parental, data = heritability))
-```
-```{r}
+
 heritability %>% 
   ggplot(aes(selected, offspring, color = treat)) +
   geom_point() +
   labs(x = "Selected", y = "Offspring") +
   theme_bw() +
   scale_color_discrete(name = "Treatment", labels = c('Neutral', 'Positive'))
-```
 

### 2020-10-17 R/01-process-flux.R
add descriptions of R scripts, remove duplicate code.

@@ -1,17 +1,18 @@
+# This scripts imports the raw methane oxidation rate data as manually entered.
+# This includes the date of each standard curve and the measures Areas for each
+# Concentration of the standard curve. It then calculates the slope of the
+# relationship Methane ~ Area for each date a standard curve was taken.
+#
+# Next, this script imports the time point data and concentration (Area) data
+# for each jar in each passage. It then computes an exponential decay function
+#
+# and extracts the exponential decay rate (k).
 # Input: standard_curve.csv, sc_dates.csv, time_data.csv, conc_data.csv, selected.csv
 # Output: fluxes.tsv
 
-
 library(tidyverse)
-#library(knitr)
-#library(kableExtra)
 library(broom)
-#library(modelr)
 library(lubridate)
-#library(morris)
-#library(magrittr)
-#library(metafor)
-#library(lme4)
 
 # Standard Curves
 

### 2020-10-17 R/02-process-herit.R

@@ -1,3 +1,8 @@
+# This script takes in the flux data from script 01 and the data on which jars
+# were selected in each passage. It then aligns these data so that for each
+# passage there is a mean flux for all parents, the selected parents, and all
+# offspring. 
+
 # input: fluxes.tsv, selected.csv
 # output: heritability.tsv
 

### 2020-10-17 R/03-response.R

@@ -1,5 +1,12 @@
+# This script intakes the flux data from script 01. It then calculates deviance
+# as the difference between the positive line and the neutral line in each
+# passage. The errors can either be calculated as the sum of the errors for
+# each line or the standard error of difference, that is, the square root of
+# the sum of the squared errors, which is slightly smaller than the sum. In
+# this case, I have chosen the SED.
+
 # input: fluxes.tsv
-# output: deviance.tsv, deviance figure
+# output: deviance.tsv, response.rds
 
 library(tidyverse)
 library(metafor)

### 2020-10-17 R/04-heritability.R

@@ -1,65 +1,28 @@
-# Inpot: selected.csv, fluxes.tsv
-# Output: heritability.tsv
+# This script imports the heritability data, which includes, for each passage,
+# the mean for all parents, the selected parents, and the offspring (of the
+# selected parents). It then calculates S, the selection differential, as the
+# mean difference between the selected parents and all parents. Because passage
+# 2 was accidentaly selected for the lowest methane oxidation rate, I take the
+# absolute value of this difference, i.e. |S|. 
+#
+# Next, it imports the deviance data and calculates R, the respones to
+# selection, as the slope of mean difference in flux (positive - neutral) over
+# time. 
+# 
+# An alternative approach would be to calculate R directly instead of
+# approximating it using the slope as in the equation offspring - parent. In
+# this case, I would not be using the deviance, but instead the raw fluxes. I
+# can do that with the heritability dataset.
+# 
+# Once I have R through one of the above approaches, 
+# 
+# Inpot: selected.csv, fluxes.tsv, heritability.tsv
+# Output: ???
 
 library(tidyverse)
 library(morris)
 
-fluxes <- read_tsv('../Output/fluxes.tsv')
-
-# Heritability
-selected_jars <- read_csv('../Data/selected.csv')
-
-# Pull out the selected parental jars
-selected <- 
-  fluxes[paste0(fluxes$passage, fluxes$jar) %in% paste0(selected_jars$passage, selected_jars$jar), ] %>%
-  select(treat, estimate, passage)
-
-# Pull out all parental jars
-parental <-
-  fluxes %>% 
-  select(passage, treat, estimate)
-
-# Pull out all offspring jars
-offspring <- 
-  fluxes %>%
-  select(treat, estimate, passage) %>% 
-  filter(passage != 1) %>% 
-  mutate(passage = passage - 1)
-
-# Calculate means for east generation
-parental_mean <-
-  parental %>% 
-  group_by(passage, treat) %>% 
-  summarize(parental = mean(estimate), .groups = "drop")
-
-offspring_mean <-
-  offspring %>% 
-  group_by(passage, treat) %>% 
-  summarize(offspring = mean(estimate), .groups = "drop")
-
-selected_mean <-
-  selected %>% 
[…]

### 2020-10-17 Presentation/Artificial_Ecosystem_Selection.Rmd
update gitignore in subdirectories and add presentation to makefile

---
title: "Artificial Ecosystem Selection"
author: "Andrew"
date: "November 8, 2018"
output:
  beamer_presentation:
    fig_width: 2
    fig_height: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
library(pwr)
```

## What I'm hoping to get out of this

- Feedback and criticism on questions and experimental design
- Identifying any pitfalls: In 6 months what will I be thinking "Dang, I wish I had done that."?
- Your ideas about how to think about the results I might get
- Statistical approaches to analyzing the results

## Overarching Question

- How do the members of a microbial community relate to the functions their ecosystem exhibits?

- How can we use genetic information to understand ecosystem function?

- What members of the community are statisticaly correlated with a particular function?

- Why might those taxa be controlling function?

## Approaches to structure-function studies

> - **Comparative approach**
>     - Sample a natural gradient in process rate and compare the communities along that gradient.
> - **Manipulative approach**
>     - Alter the microbial community by testing specific combinations of microbes.
>     - Label the substrate and measure the community that incorporates the label (SIP, BrdU).
> - **Experimental approach**
>     - Create many replicate communities and select the top performers to inoculate the next generation.

- Sampling along a natural gradient imposes a problem of covariation with other aspects of the environmet.
- By incubating replicate communities under virtually identical conditions and selecting for the particular function of interest 
- Randomization at the community level breaks the association between confounding variables and our explanatory variables of interest.
- Plus, replication across multiple lines could potentially break up correlations between taxa that are unrelated to the process

## Artificial Ecosystem Selection

```{r, fig.align='center', fig.width=1}
knitr::include_graphics("emi_140_f1.png")
```

Swenson, W., J. Arendt, and D. S. Wilson. 2000. Artificial selection of microbial ecosystems for 3-chloroaniline biodegradation. Environmental Microbiology 2:564–571.

- Create many replicate communities with variation in the microbiota
- Measure the function you're selecting on
- Take the top performers and use them to inoculate the next generation
- Variation is introduced through bottlenecks or subsampling effects on the community
- Akin to a quantitative genetics study where you would take backcross F1s to see what variation in the genome is correlated with variation in a quantitative trait

# Examples of Artificial Ecosystem Selection

## Artificial Ecosystem Selection

- Selected on plant biomass in soil and pH in aquatic microbiota
- Did not look at the microbes
- Found samples separated by soil nutrients (NH4+) (1st Discriminate function)
- Circles are controls, triangles are selected

```{r, fig.align='center'}
knitr::include_graphics("F1_large.jpg")
```

Swenson, W., D. S. Wilson, and R. Elias. 2000. Artificial ecosystem selection. Proceedings of the National Academy of Sciences 97:9110–9114.

## Response to selection

```{r, fig.align='center'}
knitr::include_graphics("F2_large.jpg")
```

Swenson, W., D. S. Wilson, and R. Elias. 2000. Artificial ecosystem selection. Proceedings of the National Academy of Sciences 97:9110–9114.

- High/low plant biomass
- Two transmission rates: 0.6 g and 6.0 g
- High/low aquatic pH

## Selection on soil microbiomes reveals reproducible impacts on plant function

- Selected on early/late flowering time
- Found reproducible effect on multiple species
- Found no differences in nutrients, but nitrogen enzyme activity was greater
- They did not test the effect of nitrogen additions

```{r, fig.align='center'}
knitr::include_graphics("ismej2014196f4.jpg")
```

Panke-Buisse, K., A. C. Poole, J. K. Goodrich, R. E. Ley, and J. Kao-Kniffin. 2015. Selection on soil microbiomes reveals reproducible impacts on plant function. The ISME Journal 9:980–989.

## Community analysis

```{r, fig.align='center'}
knitr::include_graphics("ismej2014196f2.jpg")
```

Panke-Buisse, K., A. C. Poole, J. K. Goodrich, R. E. Ley, and J. Kao-Kniffin. 2015. Selection on soil microbiomes reveals reproducible impacts on plant function. The ISME Journal 9:980–989.

## The actual study

No immigration and (assuming) speciation will be negligible.
Control should be entirely drift (and selection for surviving passage).
Experimental should be drift + selection (experimental drift = control drift).

### I expect changes in the community to be:

1. Changes in relative abundance of taxa
2. Changes in intraspecific variation in traits

### Why replicate lines?
1. Potential to separate correlations between taxa that are unrelated to oxidation rates
2. Reproducibility in selection: do the same taxa respond?

##

### Control
One line with neutral selection (random reps inoculate the next gen)

## Pilot Study

```{r, fig.align='center'}
knitr::include_graphics("select_bottles.png")
```

## Things to test with a pilot study

- Is there a response to selection?
- What transmssion rate is sufficient to see a response to selection?
- Do selection on forest and wetland soils result in different responses/communities? 
    - i.e. To what extent does the initial community impact the response to selection?

## Plan

- Soils from West Eugene Wetlands
- 8 bottles per line
- 4 lines
- Each line with different transmission rate (0.1%, 1%, 10%)
- 1 control line
- Maybe a follow-up with 1 transmission rate and separate Wetland/Forest lines to see whether source affects the response to selection

# Analysis

## Detecting a significant change in oxidation rates

- ANOVA to compare individual lines with control
- Repeated Measures ANOVA if I want to incorporate all generations in this test
- Power analysis with $\alpha$ = 0.05, power = 1 - $\beta$ = 0.8, and n = 8 suggests I would need to see a 75% increase in oxidation rates in the experimental line over the control line to detect an effect.

## Community analysis

- Differentially abundant taxa

### 2020-10-17 Presentation/test.Rmd
update gitignore in subdirectories and add presentation to makefile

--- 2020-10-17 Presentation/Artificial_Ecosystem_Selection.Rmd
@@
 ---
-title: "Artificial Ecosystem Selection"
+title: "Untitled"
 author: "Andrew"
-date: "November 8, 2018"
-output:
-  beamer_presentation:
-    fig_width: 2
-    fig_height: 3
+date: "December 4, 2018"
+output: beamer_presentation
 ---
@@
 
-- Sampling along a natural gradient imposes a problem of covariation with other aspects of the environmet.
-- By incubating replicate communities under virtually identical conditions and selecting for the particular function of interest 
-- Randomization at the community level breaks the association between confounding variables and our explanatory variables of interest.
-- Plus, replication across multiple lines could potentially break up correlations between taxa that are unrelated to the process
-
 ## Artificial Ecosystem Selection
 
-```{r, fig.align='center', fig.width=1}
-knitr::include_graphics("emi_140_f1.png")
+```{r, fig.align='center'}
+knitr::include_graphics("emi_140_f1.gif")
 ```
 
-Swenson, W., J. Arendt, and D. S. Wilson. 2000. Artificial selection of microbial ecosystems for 3-chloroaniline biodegradation. Environmental Microbiology 2:564–571.
+## Slide with Plot
 
-- Create many replicate communities with variation in the microbiota
-- Measure the function you're selecting on
-- Take the top performers and use them to inoculate the next generation
-- Variation is introduced through bottlenecks or subsampling effects on the community
-- Akin to a quantitative genetics study where you would take backcross F1s to see what variation in the genome is correlated with variation in a quantitative trait
-
-# Examples of Artificial Ecosystem Selection
-
-## Artificial Ecosystem Selection
-
-- Selected on plant biomass in soil and pH in aquatic microbiota
-- Did not look at the microbes
-- Found samples separated by soil nutrients (NH4+) (1st Discriminate function)
-- Circles are controls, triangles are selected
-
-```{r, fig.align='center'}
-knitr::include_graphics("F1_large.jpg")
+```{r pressure}
+plot(pressure)
 ```
-
-Swenson, W., D. S. Wilson, and R. Elias. 2000. Artificial ecosystem selection. Proceedings of the National Academy of Sciences 97:9110–9114.
-
-## Response to selection
-
-```{r, fig.align='center'}
-knitr::include_graphics("F2_large.jpg")
-```
-
-Swenson, W., D. S. Wilson, and R. Elias. 2000. Artificial ecosystem selection. Proceedings of the National Academy of Sciences 97:9110–9114.
-
-- High/low plant biomass
-- Two transmission rates: 0.6 g and 6.0 g
-- High/low aquatic pH
-
-## Selection on soil microbiomes reveals reproducible impacts on plant function
-
-- Selected on early/late flowering time
-- Found reproducible effect on multiple species
-- Found no differences in nutrients, but nitrogen enzyme activity was greater
-- They did not test the effect of nitrogen additions
-
-```{r, fig.align='center'}
-knitr::include_graphics("ismej2014196f4.jpg")
-```
-
-Panke-Buisse, K., A. C. Poole, J. K. Goodrich, R. E. Ley, and J. Kao-Kniffin. 2015. Selection on soil microbiomes reveals reproducible impacts on plant function. The ISME Journal 9:980–989.
-
-## Community analysis
-
-```{r, fig.align='center'}
-knitr::include_graphics("ismej2014196f2.jpg")
-```
-
-Panke-Buisse, K., A. C. Poole, J. K. Goodrich, R. E. Ley, and J. Kao-Kniffin. 2015. Selection on soil microbiomes reveals reproducible impacts on plant function. The ISME Journal 9:980–989.
-
-## The actual study
-
-No immigration and (assuming) speciation will be negligible.
-Control should be entirely drift (and selection for surviving passage).
-Experimental should be drift + selection (experimental drift = control drift).
-
-### I expect changes in the community to be:
-
-1. Changes in relative abundance of taxa
-2. Changes in intraspecific variation in traits
-
-### Why replicate lines?
-1. Potential to separate correlations between taxa that are unrelated to oxidation rates
-2. Reproducibility in selection: do the same taxa respond?
-
-##
-
-### Control
-One line with neutral selection (random reps inoculate the next gen)
-
-## Pilot Study
-
-```{r, fig.align='center'}
-knitr::include_graphics("select_bottles.png")
-```
-
-## Things to test with a pilot study
-
-- Is there a response to selection?
-- What transmssion rate is sufficient to see a response to selection?
-- Do selection on forest and wetland soils result in different responses/communities? 
-    - i.e. To what extent does the initial community impact the response to selection?
-
-## Plan
-
-- Soils from West Eugene Wetlands
-- 8 bottles per line
-- 4 lines
-- Each line with different transmission rate (0.1%, 1%, 10%)
-- 1 control line
-- Maybe a follow-up with 1 transmission rate and separate Wetland/Forest lines to see whether source affects the response to selection
-
-# Analysis
-
-## Detecting a significant change in oxidation rates
-
-- ANOVA to compare individual lines with control
-- Repeated Measures ANOVA if I want to incorporate all generations in this test
-- Power analysis with $\alpha$ = 0.05, power = 1 - $\beta$ = 0.8, and n = 8 suggests I would need to see a 75% increase in oxidation rates in the experimental line over the control line to detect an effect.
-
-## Community analysis
-
-- Differentially abundant taxa

### 2020-10-20 Presentation/presentation.Rmd
start presentation for lab meeting

---
title: "Artificial Ecosystem Selection Update"
author: "Andrew Morris"
date: "10/21/2020"
output: 
  revealjs::revealjs_presentation:
    transition: fade
    reveal_options:
      slideNumber: true
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE, message = FALSE)
library(tidyverse)
library(metafor)
library(morris)
library(broman)
```

# Outline

## What I'm hoping to get out of this

- Show you what I've been working on
- Hear your questions to see if that guides me down any path
- ???

# The Question

## 
One of the fundamental questions in ecology is, "How does the biodiversity
  that we observe influence ecosystem functions?"

## 
Historically, attempts to address this question have focused on alpha
  diversity as it relates to the *rate* or *stability* of an ecosystem function

##
For certain functions, this works pretty well. For example, the stability of
  primary productivity in temperate plant communities in the face of drought is
  correlated with species richness of plants.

## 
However, microbial community data exhibit certain limitations that cause this
  approach to fail:

  1. Without a consistent species concept, it is not obvious at what level of
     biological organization to evaluate this relationship. (the ASV? the
     genus? the family?)
  2. It is not always clear which members of a community are members of a
     particular functional group. (This can be determiend morpholigically for
     plants, but must be determined genetically in natural microbial
     communities.)
  3. Microbial communities are immensely diverse and without solutions to the
     previous two problems, sifting through that diversity is challenging.

## When we apply this BEF approach to microbial communities, there are limited
  gains in terms of our knowlegde of the controls on ecosystem function

# A new approach to linking microbial communities to ecosystem function

## An alternative approach

- One way to think about microbial communities and ecosystem functions is as assemblages of genes within an ecosystem and a trait at the ecosystem level.
- In fact, this often the only way we can measure microbial communities in natural ecosystems.

## Experiment

I collected a single soil core from near UO campus and used sterile potting mix as the subtrate. Ecosystems were created in 500 mL mason jars with 45 g of sterile potting mix and 5 g of living soil and were brought to 60% of field capacity (the optimal level for aerobic processes such as methane oxidation). I created 24 jars in this way and randomly assigned them to a Positive or Neutral selection line.

# Is there a response to selection at the ecosystem level on soil methane oxidation rate?

##

```{r, fig.width=4, fig.height=3, out.width=6e2, out.height=4.5e2}

fluxes <- read_tsv('../Output/fluxes.tsv')
ggplot(fluxes, aes(x = passage, y = estimate, color = treat)) + 
  geom_jitter() +
  scale_color_manual(name = 'Treatment', labels = c('Neutral', 'Positive'),
                     values = c('gray60', 'darkorange2')) +
  labs(x = "Passage", y = 'Flux (-k)') +
  theme_bw()

```

##

```{r, fig.width=4, fig.height=3, out.width=6e2, out.height=4.5e2}

fluxes %>% 
  select(treat, passage, estimate) %>% 
  group_by(passage, treat) %>% 
  summarize(mean = mean(estimate),
            se = se(estimate)) %>% 
ggplot(mapping = aes(x = passage, y = mean, ymin = mean - se, ymax = mean + se,
                     color = treat)) + 
  geom_pointrange() +
  scale_color_manual(name = 'Treatment', values = c('gray60', 'darkorange2')) +
  theme_bw()
```

##

There was a response to selection on soil methane oxidation rate in the positive treatment relative to the neutral treatment.

```{r, fig.width=4, fig.height=3, out.width=6e2, out.height=4.5e2}

deviance <- read_tsv('../Output/deviance.tsv')
dev_fit <- readRDS('../Output/dev_fit.rds')
dev_pred <- predict(dev_fit)
passages <- length(deviance$passage)

ggplot(deviance, aes(x = passage, y = deviance, ymin = deviance - se, ymax = deviance + se)) + 
  geom_ribbon(aes(ymin = dev_pred$ci.lb, ymax = dev_pred$ci.ub), alpha = 0.2) +
  geom_pointrange(color = 'darkorange2') + 
      geom_segment(aes(x = 1, xend = passages, y = dev_fit$beta[1] + dev_fit$beta[2], yend = dev_fit$beta[1] + dev_fit$beta[2]*passages), color = 'darkorange2', size = 1) +
  labs(x = "Passage Number", y = "Deviation from Control (P - N)")+
  theme_bw()

```

```{r}
init <- 
  fluxes %>% 
  group_by(passage) %>% 
  summarize(mean = mean(estimate)) %>% 
  filter(passage == 1) %>% 
  pull()

init_se <- 
  fluxes %>% 
  group_by(passage) %>% 
  summarize(se = se(estimate)) %>% 
  filter(passage == 1) %>% 
  pull()
```

z = `r myround(dev_fit$zval[2], 2)`, p = `r myround(dev_fit$pval[2], 3)`

slope = `r myround(dev_fit$beta[2], 3)` $\pm$ `r myround(dev_fit$se[2], 3)`

(initial mean flux  = `r myround(init, 3)` $\pm$ `r myround(init_se, 3)`)

### 2020-10-20 R/04-heritability.R

@@ -21,8 +21,10 @@
 
 library(tidyverse)
 library(morris)
+library(broom)
 
 heritability <- read_tsv('../Output/heritability.tsv')
+fluxes <- read_tsv('../Output/fluxes.tsv')
 
 # Function for the Breeder's equation
 
@@ -30,6 +32,16 @@ breeders <- function(R, S) {
   R / S
 }
 
+# Matt's approach: h2 for each generation
+h2_per_gen  <- 
+heritability %>% 
+  filter(treat == 'p') %>% 
+  mutate(R = offspring - parental) %>% 
+  mutate(S = selected - parental) %>% 
+  mutate(h2 = breeders(R, S))
+
+ggplot(h2_per_gen, aes(x = passage, y = h2)) +
+  geom_point()
 
 # Calculate R using regression of deviance over time.
 # Calculate S using math.
@@ -41,7 +53,19 @@ S <-
   heritability %>%
   filter(treat == "p") %>%
   mutate(S = abs(selected - parental)) %>%
-  summarize(S = mean(S))
+  summarize(S = mean(S))  %>% 
+  pull()
+
+h2_reg_dev <- breeders(R, S)
+
+lm_fit <- 
+  fluxes %>% 
+  filter(treat == 'p') %>% 
+  select(treat, passage, estimate) %>% 
+  lm(estimate ~ passage, data = .) 
+
+R_not_dev <- pull(tidy(lm_fit)[2, 2])
+h2_reg_abs <- breeders(R_not_dev, S)
 
 # Calculate h2 (heritability) using math, not regression
 # For each generation and then calculate the mean/se
@@ -75,6 +99,7 @@ h2 <-
 h2_no2 <- 
   h2_calc_method_no_2 %>% 
   mutate(method = "calc p no 2")
+
 
 h2 <- rbind(h2, h2_no2)
 

### 2020-10-21 Presentation/presentation.Rmd
update heritability analysis and presentation

@@
 
-# Outline
-
 ## What I'm hoping to get out of this
 
-- Show you what I've been working on
-- Hear your questions to see if that guides me down any path
-- ???
+- Present my results so far and see if you have any questions that might guide future steps.
+- Talk about what "heritability" means in a community context.
 
-# The Question
-
-## 
-One of the fundamental questions in ecology is, "How does the biodiversity
-  that we observe influence ecosystem functions?"
+# How does biodiversity influence ecosystem functions?
 
@@
 
+```{r, fig.width=4, fig.height=3, out.width=6e2, out.height=4.5e2}
+knitr::include_graphics('draught-stability.png')
+```
+
 ##
-For certain functions, this works pretty well. For example, the stability of
-  primary productivity in temperate plant communities in the face of drought is
-  correlated with species richness of plants.
+Reasons this works:
+
+1. There's a clear unit of biological organization - the species. 
+2. It's obvious what a "plant" is.
+3. Communities are relatively simple - (1-25 species).
 
@@
   2. It is not always clear which members of a community are members of a
-     particular functional group. (This can be determiend morpholigically for
-     plants, but must be determined genetically in natural microbial
-     communities.)
-  3. Microbial communities are immensely diverse and without solutions to the
+     particular functional group. Further, in a metabolic network, it's not
+     always obvious which functional group is limiting the end product.
+  3. Microbial communities are immensely diverse (100s-1000s of ASVs) and
+     without solutions to the
      previous two problems, sifting through that diversity is challenging.
-
-## When we apply this BEF approach to microbial communities, there are limited
-  gains in terms of our knowlegde of the controls on ecosystem function
 
@@
 - In fact, this often the only way we can measure microbial communities in natural ecosystems.
+- Considering them this way, they resemble an organism with a phenotype. (In a way that is dangerously Clementsian) 
+- Using that metaphor, we can think about ways of identifying genes or organisms that are associated with the presence or rate of particular functions.
+
+## Artificial Ecosystem Selection
+
+A *comparative* approach to this problem would be to go out and sample lets of
+ecosystems for their functions and their communities. And do lots of
+correlations. That's fine, but it's more of a hypothesis generating tool. A
+somewhat more *experimental* approach would be to manipulate the community by
+generating lots of different communities under very similar environmental
+conditions. And then testing which combinations produce the greatest function.
+
+## Artificial Ecosystem Selection
+
+To do this, I generated 24 ecosystems.
 
@@
   geom_pointrange() +
-  scale_color_manual(name = 'Treatment', values = c('gray60', 'darkorange2')) +
+  scale_color_manual(name = 'Treatment', labels = c('Neutral', 'Positive'),
+                     values = c('gray60', 'darkorange2')) +
+  labs(x = "Passage", y = 'Flux (-k)') +
   theme_bw()
@@
 (initial mean flux  = `r myround(init, 3)` $\pm$ `r myround(init_se, 3)`)
+
+# How *heritable* is soil methane oxidation rate as an ecosystem trait?
+
+##
+
+Heritability is "the relative importance of heredity in determining phenotypic
+values"
+
+"A character can be 'hereditary' in the sense of being **determined by
+the genotype**[1] or in the sense of being **transmitted from parents to offspring**[2]."
+
+[1] Broad-sense heritability $= V_G / V_P$
+
+[2] Narrow-sense heritability $= V_A / V_P$
+
+<small>Falconer and MacKay, 1996</small>
+
+##
+
+$V_G$ is quite are to estimate, but is of considerable theoretical interest.
+Estimating $V_G$ would answer the question, "How much of an ecosystem function is determined by the genes in that ecosystem vs.
+the environmetal conditions in that ecosystem?"
+
+$V_A$ is easier to estimate and is of more practical value in population genetics for its use in breeding programs.
+It is the ratio of breeding value to phenotype value and determines whether selecting a parent for their traits can produce offspring with those traits.
+
+##
+
+```{r, fig.width=2, fig.height=3, out.width=3e2, out.height=4.5e2}
+h2_per_gen <- read_tsv('../Output/h2_per_gen.tsv')
+
+h2_per_gen_sum  <-
+  h2_per_gen %>% 
+  summarize(h2_med = median(h2), h2_mad = mad(h2))
+
+ggplot(h2_per_gen, aes(x = passage, y = h2)) +
+  geom_point() +
+  geom_pointrange(aes(x = 2.5, y = h2_med, 
+                                  ymax = h2_med + h2_mad,
+                                  ymin = h2_med - h2_mad),
+  data = h2_per_gen_sum) +
+  labs(x = 'Passage Number', y = expression(paste("Narrow-sense Heritability (",
+                                                h^2, ")"))) +
+  theme_bw()
+
+```
+
+Each small point is the h2 for that passage and its offspring. The larger central point is the median h2 (+/- median absolute difference) across all four passages.

### 2020-10-21 R/03-response.R

@@ -10,6 +10,13 @@
 
 library(tidyverse)
 library(metafor)
+library(morris)
+
+# Standard error of the difference function
+sed <- function(p, n, na.rm = FALSE) {
+  sqrt(se(p, na.rm = na.rm) ^ 2 + se(n, na.rm = na.rm)^ 2)
+}
+
 
 # Response to Selection
 
@@ -24,17 +31,6 @@ deviance <-
   pivot_wider(names_from = treat, values_from = mean) %>% 
   mutate(deviance = p - n)
 
-# Create se function that can handle NAs
-se <- function(x, na.rm = FALSE){
-  sqrt(var(if (is.vector(x) || is.factor(x)) x else as.double(x), 
-    na.rm = na.rm)) / sqrt(length(if (na.rm == TRUE) na.omit(x) else x))
-}
-
-# Standard error of the difference function
-sed <- function(p, n, na.rm = FALSE) {
-  sqrt(se(p, na.rm = na.rm) ^ 2 + se(n, na.rm = na.rm)^ 2)
-}
-
 # Calculate variance of deviance as SED
 deviance$se <-
   fluxes %>% 
@@ -47,6 +43,8 @@ deviance$se <-
 
 # Fit deviance model
 fit <- rma(deviance ~ passage, se^2, method="FE", data = deviance)
+
+plot(fit)
 
 # Plot deviance
 ggplot(deviance, aes(x = passage, y = deviance, ymin = deviance - se, ymax = deviance + se)) + 
@@ -92,3 +90,4 @@ ggplot(., aes(x = passage, y = deviance, ymin = deviance - se, ymax = deviance +
 write_tsv(deviance, '../Output/deviance.tsv')
 saveRDS(rma_output, '../Output/response.rds')
 saveRDS(fit, '../Output/dev_fit.rds')
+

### 2020-10-21 R/04-heritability.R

@@ -40,8 +40,7 @@ heritability %>%
   mutate(S = selected - parental) %>% 
   mutate(h2 = breeders(R, S))
 
-ggplot(h2_per_gen, aes(x = passage, y = h2)) +
-  geom_point()
+write_tsv(h2_per_gen, '../Output/h2_per_gen.tsv')
 
 # Calculate R using regression of deviance over time.
 # Calculate S using math.

### 2020-10-21 Presentation/presentation.Rmd
update presentation

--- 2020-10-20 Presentation/presentation.Rmd
@@
 
-# Outline
-
 ## What I'm hoping to get out of this
 
-- Show you what I've been working on
-- Hear your questions to see if that guides me down any path
-- ???
+- Give the rationale/motivation for this project, what I've learned so far, and what I could learn from future data.
+- Hear  your questions and comments that might guide future steps.
+- Talk about what "heritability" means in a community context.
 
-# The Question
+# How does biodiversity influence ecosystem function?
 
-## 
-One of the fundamental questions in ecology is, "How does the biodiversity
-  that we observe influence ecosystem functions?"
+## One approach
+Relating alpha diversity within a functional group to the *rate* or *stability* of an ecosystem function
 
-## 
-Historically, attempts to address this question have focused on alpha
-  diversity as it relates to the *rate* or *stability* of an ecosystem function
+```{r, fig.width=4, fig.height=3, out.width=6e2, out.height=4.5e2}
+knitr::include_graphics('draught-stability.png')
+```
 
-##
-For certain functions, this works pretty well. For example, the stability of
-  primary productivity in temperate plant communities in the face of drought is
-  correlated with species richness of plants.
+## Reasons this works
 
-## 
-However, microbial community data exhibit certain limitations that cause this
-  approach to fail:
+1. There's a clear unit of biological organization - the species. 
+2. It's obvious what a "plant" is.
+3. Communities are relatively simple - (1-25 species).
+
+## Limitations of microbial community data
 
@@
   2. It is not always clear which members of a community are members of a
-     particular functional group. (This can be determiend morpholigically for
-     plants, but must be determined genetically in natural microbial
-     communities.)
-  3. Microbial communities are immensely diverse and without solutions to the
+     particular functional group. Further, in a metabolic network, it's not
+     always obvious which functional group is limiting the end product.
+  3. Microbial communities are immensely diverse (100s-1000s of ASVs) and
+     without solutions to the
      previous two problems, sifting through that diversity is challenging.
 
-## When we apply this BEF approach to microbial communities, there are limited
-  gains in terms of our knowlegde of the controls on ecosystem function
+## An alternative conceptualization of communities
 
-# A new approach to linking microbial communities to ecosystem function
+A community is a pool of genes analogous to a genome (the metagenome) and that community exhibits a phenotype (ecosystem function).
 
-## An alternative approach
+Thinking about it this way, an ecosystem function might exhibit similar properties to an organismal trait.
 
-- One way to think about microbial communities and ecosystem functions is as assemblages of genes within an ecosystem and a trait at the ecosystem level.
-- In fact, this often the only way we can measure microbial communities in natural ecosystems.
+$V_P = V_G + V_E$
 
-## Experiment
+## Heritability is "the relative importance of heredity in determining phenotypic values"
 
-I collected a single soil core from near UO campus and used sterile potting mix as the subtrate.
-Ecosystems were created in 500 mL mason jars with 45 g of sterile potting mix and 5 g of living soil and were brought to 60% of field capacity (the optimal level for aerobic processes such as methane oxidation).
-I created 24 jars in this way and randomly assigned them to a Positive or Neutral selection line.
+"A character can be 'hereditary' in the sense of being **determined by
+the genotype**[1] or in the sense of being **transmitted from parents to offspring**[2]."
+
+[1] Broad-sense heritability,  $H^2 = V_G / V_P$
+
+[2] Narrow-sense heritability, $h^2 = V_A / V_P$
+
+<small>Falconer and MacKay, 1996</small>
+
+# An alternative approach to linking microbial communities to ecosystem function
+
+## Artificial Ecosystem Selection
+
+- Generate lots of communities with different combinations of taxa.
+- Put them into a common environment to minimize environmental variation.
+- Test which combinations generate the greatest function.
+- Reshuffle the best communities and test again.
+- Similar to dilution to extinction or enrichment culturing.
+
+## Method
+
+To do this, I generated 24 ecosystems.
+
+```{r, fig.width=4, fig.height=3, out.width=9e2*0.8, out.height=4.5e2*0.8}
+knitr::include_graphics('aes_design.png')
+```
 
@@
                      values = c('gray60', 'darkorange2')) +
-  labs(x = "Passage", y = 'Flux (-k)') +
-  theme_bw()
+  labs(x = "Passage", y = 'Methane Oxidation Rate (-k)') +
+  theme_classic()
 
@@
   geom_pointrange() +
-  scale_color_manual(name = 'Treatment', values = c('gray60', 'darkorange2')) +
-  theme_bw()
+  scale_color_manual(name = 'Treatment', labels = c('Neutral', 'Positive'),
+                     values = c('gray60', 'darkorange2')) +
+  labs(x = "Passage", y = 'Methane Oxidation Rate (-k)') +
+  theme_classic()
 ```
@@
 There was a response to selection on soil methane oxidation rate in the positive treatment relative to the neutral treatment.
+<style>
+.container{
+    display: flex;
+}
+.col{
+    flex: 1;
+}
+</style>
 
-```{r, fig.width=4, fig.height=3, out.width=6e2, out.height=4.5e2}
+<div class="container">
+
+<div class="col">
+```{r, fig.width=4, fig.height=3, out.width=6e2*0.8, out.height=4.5e2*0.8}
 
@@
   labs(x = "Passage Number", y = "Deviation from Control (P - N)")+
-  theme_bw()
+  theme_classic()
 
 ```
+</div>
 
+<div class="col">
 ```{r}
@@
 (initial mean flux  = `r myround(init, 3)` $\pm$ `r myround(init_se, 3)`)
+</div>
+
+</div>
+
+# How *heritable* is soil methane oxidation rate as an ecosystem trait?
+
+## Broad- vs. Narrow-sense heritability
+
+$V_G/V_P$ is the extent to which phenotype is determined by the genotype.
+
+$V_A/V_P$ is the extent to which phenotype is determined by the genes transmitted from the parents.
+
+## Broad- vs. Narrow-sense heritability
+
+$V_G/V_P$ is the extent to which phenotype is determined by the genotype.
+
+**$V_A/V_P$ is the extent to which phenotype is determined by the genes transmitted from the parents.**
+
+## Narrow-sense heritability
+
+$h^2 = R / S$
+
+$R$ = Response to selection, difference between parents and offspring
+
+$S$ = Selection differential, difference between selected parents and all parents
+
+##
+
+```{r}
+h2_per_gen <- read_tsv('../Output/h2_per_gen.tsv')
+
+h2_per_gen_sum  <-
+  h2_per_gen %>% 
+  summarize(h2_med = median(h2), h2_mad = mad(h2))
+```
+
+```{r, fig.width=2.5, fig.height=2.5, out.width=3e2, out.height=3e2}
+ggplot(h2_per_gen, aes(x = passage, y = h2)) +
+  geom_point() +
+  geom_pointrange(aes(x = 2.5, y = h2_med, 
+                                  ymax = h2_med + h2_mad,
+                                  ymin = h2_med - h2_mad),
+  data = h2_per_gen_sum) +
+  labs(x = 'Passage Number', y = expression(paste("Narrow-sense Heritability (",
+                                                h^2, ")"))) +
+  theme_classic()
+
+```
+
+Each small point is the $h^2$ for that passage and its offspring. The
+larger central point is the median $h^2$ (+/- median absolute difference)
+across all four passages. $h^2$ =  `r myround(h2_per_gen_sum$h2_med, 2)` $\pm$ `r myround(h2_per_gen_sum$h2_mad, 2)`.
+
+# Conclusions and interpetation
+
+## Community composition impacts the rate of this ecosystem function
+
+```{r, fig.width=4, fig.height=3, out.width=6e2*.8, out.height=4.5e2*.8}
+
+deviance <- read_tsv('../Output/deviance.tsv')
+dev_fit <- readRDS('../Output/dev_fit.rds')
+dev_pred <- predict(dev_fit)
+passages <- length(deviance$passage)
+
+ggplot(deviance, aes(x = passage, y = deviance, ymin = deviance - se, ymax = deviance + se)) + 
+  geom_ribbon(aes(ymin = dev_pred$ci.lb, ymax = dev_pred$ci.ub), alpha = 0.2) +
+  geom_pointrange(color = 'darkorange2') + 
+      geom_segment(aes(x = 1, xend = passages, y = dev_fit$beta[1] + dev_fit$beta[2], yend = dev_fit$beta[1] + dev_fit$beta[2]*passages), color = 'darkorange2', size = 1) +
+  labs(x = "Passage Number", y = "Deviation from Control (P - N)")+
+  theme_classic()
+
+```
+
+## `r myround(h2_per_gen_sum$h2_med, 2)` of the variation in methane oxidation rate is attributable to the additive effect of taxa passed from the parents
+
+The remaining `r 1 - as.numeric(myround(h2_per_gen_sum$h2_med, 2))` is due to
+environmental variation not controlled for experimentally or with the neutral
+treatment, interactions among taxa, ...?
+
+Caveat: this doesn't account for cor
+
+```{r, fig.width=2.5, fig.height=2.5, out.width=3e2, out.height=3e2}
+ggplot(h2_per_gen, aes(x = passage, y = h2)) +
+  geom_point() +
+  geom_pointrange(aes(x = 2.5, y = h2_med, 
+                                  ymax = h2_med + h2_mad,
+                                  ymin = h2_med - h2_mad),
+  data = h2_per_gen_sum) +
+  labs(x = 'Passage Number', y = expression(paste("Narrow-sense Heritability (",
+                                                h^2, ")"))) +
+  theme_classic()
+
+```
+
+## Next: Which taxa changed as a response to selection?

### 2020-10-22 Presentation/presentation.Rmd
update presentation for lab meeting. update heritability calc to include neutral line.

@@
 
-- Give the rationale/motivation for this project, what I've learned so far, and what I could learn from future data.
+- Practice talking about why I'm doing what I'm doing and what we can learn from it.
 - Hear your questions and comments that might guide future steps.
@@
 
-```{r, fig.width=4, fig.height=3, out.width=6e2, out.height=4.5e2}
+```{r, fig.width=4, fig.height=3, out.width=6e2*.8, out.height=4.5e2*0.8}
 knitr::include_graphics('draught-stability.png')
@@
 2. It's obvious what a "plant" is.
-3. Communities are relatively simple - (1-25 species).
+3. Communities are relatively simple (1-25 species).
 
@@
 
-  1. Without a consistent species concept, it is not obvious at what level of
-     biological organization to evaluate this relationship. (the ASV? the
-     genus? the family?)
-  2. It is not always clear which members of a community are members of a
-     particular functional group. Further, in a metabolic network, it's not
-     always obvious which functional group is limiting the end product.
-  3. Microbial communities are immensely diverse (100s-1000s of ASVs) and
-     without solutions to the
-     previous two problems, sifting through that diversity is challenging.
+  1. No consistent species concept.
+  2. What's a plant?
+  3. Diversity can be 100s to 10,000s.
 
@@
 
-- Generate lots of communities with different combinations of taxa.
-- Put them into a common environment to minimize environmental variation.
-- Test which combinations generate the greatest function.
-- Reshuffle the best communities and test again.
-- Similar to dilution to extinction or enrichment culturing.
+1. Generate lots of communities with different combinations of taxa.
+2. Put them into a common environment to minimize environmental variation.
+3. Test which combinations generate the greatest function.
+4. Reshuffle the best communities and test again.
+5. Investigate which taxa respond to selection.
 
-## Method
-
-To do this, I generated 24 ecosystems.
+## Design
 
@@
 
-##
+## Community composition impacts the rate of this ecosystem function
 
-There was a response to selection on soil methane oxidation rate in the positive treatment relative to the neutral treatment.
 <style>
@@
 
+There was a response to selection on soil methane oxidation rate in the positive treatment relative to the neutral treatment.
+
 z = `r myround(dev_fit$zval[2], 2)`, p = `r myround(dev_fit$pval[2], 3)`
@@
 
-##
-
 ```{r}
@@
   h2_per_gen %>% 
-  summarize(h2_med = median(h2), h2_mad = mad(h2))
+  group_by(treat) %>% 
+  summarize(h2_med = median(h2), h2_mad = mad(h2), .groups = 'drop')
 ```
 
-```{r, fig.width=2.5, fig.height=2.5, out.width=3e2, out.height=3e2}
-ggplot(h2_per_gen, aes(x = passage, y = h2)) +
+## Half of the variation in methane oxidation rate is attributable to the additive effect of genes passed from the parents
+
+Positive: `r as.numeric(myround(h2_per_gen_sum$h2_med[2], 2))` 
+$\pm$
+`r as.numeric(myround(h2_per_gen_sum$h2_mad[2], 2))`
+
+Neutral: `r as.numeric(myround(h2_per_gen_sum$h2_med[1], 2))` 
+$\pm$
+`r as.numeric(myround(h2_per_gen_sum$h2_mad[1], 2))`
+
+The remainder is due to
+environmental variation not controlled for experimentally or with the neutral
+treatment, interactions among taxa, ...?
+
+Caveat: $h^2$ may be affected by correlated traits also under selection.
+
+## Heritability for the positive and neutral lines
+
+<style>
+.container{
+    display: flex;
+}
+.col{
+    flex: 1;
+}
+</style>
+
+<div class="container">
+<div class="col">
+```{r, warning=FALSE, fig.width=4, fig.height=3, out.width=6e2*0.8, out.height=4.5e2*0.8}
+ggplot(h2_per_gen, aes(x = passage, y = h2, color = treat)) +
   geom_point() +
-  geom_pointrange(aes(x = 2.5, y = h2_med, 
-                                  ymax = h2_med + h2_mad,
-                                  ymin = h2_med - h2_mad),
-  data = h2_per_gen_sum) +
   labs(x = 'Passage Number', y = expression(paste("Narrow-sense Heritability (",
                                                 h^2, ")"))) +
+  scale_color_manual(name = 'Treatment', labels = c('Neutral', 'Positive'),
+                     values = c('gray60', 'darkorange2')) +
   theme_classic()
@@
 
-Each small point is the $h^2$ for that passage and its offspring. The
-larger central point is the median $h^2$ (+/- median absolute difference)
-across all four passages. $h^2$ =  `r myround(h2_per_gen_sum$h2_med, 2)` $\pm$ `r myround(h2_per_gen_sum$h2_mad, 2)`.
+</div>
 
-# Conclusions and interpetation
+<div class="col">
 
-## Community composition impacts the rate of this ecosystem function
-
-```{r, fig.width=4, fig.height=3, out.width=6e2*.8, out.height=4.5e2*.8}
-
-deviance <- read_tsv('../Output/deviance.tsv')
-dev_fit <- readRDS('../Output/dev_fit.rds')
-dev_pred <- predict(dev_fit)
-passages <- length(deviance$passage)
-
-ggplot(deviance, aes(x = passage, y = deviance, ymin = deviance - se, ymax = deviance + se)) + 
-  geom_ribbon(aes(ymin = dev_pred$ci.lb, ymax = dev_pred$ci.ub), alpha = 0.2) +
-  geom_pointrange(color = 'darkorange2') + 
-      geom_segment(aes(x = 1, xend = passages, y = dev_fit$beta[1] + dev_fit$beta[2], yend = dev_fit$beta[1] + dev_fit$beta[2]*passages), color = 'darkorange2', size = 1) +
-  labs(x = "Passage Number", y = "Deviation from Control (P - N)")+
-  theme_classic()
+```{r, warning=FALSE, fig.width=4, fig.height=3, out.width=6e2*0.8, out.height=4.5e2*0.8}
+ggplot(h2_per_gen, aes(x = passage, y = h2, color = treat)) +
+  geom_point() +
+  labs(x = 'Passage Number', y = expression(paste("Narrow-sense Heritability (",
+                                                h^2, ")"))) +
+  scale_color_manual(name = 'Treatment', labels = c('Neutral', 'Positive'),
+                     values = c('gray60', 'darkorange2')) +
+  theme_classic() +
+  ylim(-0.683, 1.56)
 
 ```
+</div>
 
-## `r myround(h2_per_gen_sum$h2_med, 2)` of the variation in methane oxidation rate is attributable to the additive effect of taxa passed from the parents
+</div>
 
-The remaining `r 1 - as.numeric(myround(h2_per_gen_sum$h2_med, 2))` is due to
-environmental variation not controlled for experimentally or with the neutral
-treatment, interactions among taxa, ...?
+# Which taxa respond to selection? Are correlated with flux?
 
-Caveat: this doesn't account for cor
+## In progress:
 
-```{r, fig.width=2.5, fig.height=2.5, out.width=3e2, out.height=3e2}
-ggplot(h2_per_gen, aes(x = passage, y = h2)) +
-  geom_point() +
-  geom_pointrange(aes(x = 2.5, y = h2_med, 
-                                  ymax = h2_med + h2_mad,
-                                  ymin = h2_med - h2_mad),
-  data = h2_per_gen_sum) +
-  labs(x = 'Passage Number', y = expression(paste("Narrow-sense Heritability (",
-                                                h^2, ")"))) +
-  theme_classic()
-
-```
-
-## Next: Which taxa changed as a response to selection?
+- Sequencing *pmoA* to characterize the diversity of methane oxidizers.
+- Sequencing 16S to look broadly at bacterial/archaeal taxa.

### 2020-10-22 R/04-heritability.R

@@ -35,12 +35,14 @@ breeders <- function(R, S) {
 # Matt's approach: h2 for each generation
 h2_per_gen  <- 
 heritability %>% 
-  filter(treat == 'p') %>% 
   mutate(R = offspring - parental) %>% 
   mutate(S = selected - parental) %>% 
   mutate(h2 = breeders(R, S))
 
 write_tsv(h2_per_gen, '../Output/h2_per_gen.tsv')
+
+ggplot(h2_per_gen, aes(x = passage, color = treat, y = R)) +
+  geom_point()
 
 # Calculate R using regression of deviance over time.
 # Calculate S using math.

### 2020-10-22 R/process.Rmd
recover processing rmd file to look at fluxes

@@
 
-```{r eval=FALSE, include=FALSE}
+```{r deviance, eval=FALSE, include=FALSE}
+# Calculate deviance as p - n
 deviance <-
@@
 
+# Create se function that can handle NAs
 se <- function(x, na.rm = FALSE){
@@
 
+# Standard error of the difference function
 sed <- function(p, n, na.rm = FALSE) {
@@
 
+# Calculate variance of deviance as SED
 deviance$se <-
@@
 
+# Fit deviance model
 fit <- rma(deviance ~ passage, se^2, method="FE", data = deviance)
 
+# Plot deviance
 ggplot(deviance, aes(x = passage, y = deviance, ymin = deviance - se, ymax = deviance + se)) + 
   geom_pointrange(color = 'darkorange2') + 
-  geom_abline(intercept = fit$beta[1], slope = fit$beta[2], color = 'blue')
+  geom_abline(intercept = fit$beta[1], slope = fit$beta[2])
 
+# Take deviance model parameters and export into a data.frame
 rma_output <- data.frame(
@@
 )
+
+# Fit deviance model without passage 2
+fit_no2 <- 
+filter(deviance, passage != 2) %>% 
+rma(deviance ~ passage, se^2, method="FE", data = .)
+
+# Take deviance model withou passage 2 parameters and export into a data.frame
+
+rma_output_no2 <- data.frame(
+Estimate = fit$beta,
+se = fit$se,
+z.value = fit$zval,
+p.value = fit$pval,
+upper.ci = fit$ci.lb,
+lower.ci = fit$ci.ub
+)
+
+# R (response to selection) for the no2 model
+R_no2 <- rma_output_no2[[2, 1]]
+R <- rma_output[[2, 1]]
+
+# Plot the no2 model
+filter(deviance, passage != 2) %>% 
+ggplot(., aes(x = passage, y = deviance, ymin = deviance - se, ymax = deviance + se)) + 
+  geom_pointrange(color = 'darkorange2') + 
+  geom_abline(intercept = fit_no2$beta[1], slope = fit_no2$beta[2])
+
+# Print the full model
 rma_output %>% 
@@
 
-write.csv(deviance, 
+write.csv(deviance, '../Output/deviance.tsv')
 ```
@@
 
-```{r heritability, message=FALSE, warning=FALSE}
+```{r create_heritability_data, message=FALSE, warning=FALSE}
 selected_jars <- read_csv('../Data/selected.csv')
 
+# Pull out the selected parental jars
+selected <- 
+  fluxes[paste0(fluxes$passage, fluxes$jar) %in% paste0(selected_jars$passage, selected_jars$jar), ] %>%
+  select(treat, estimate, passage)
+
+# Pull out all parental jars
 parental <-
-  fluxes[paste0(fluxes$passage, fluxes$jar) %in% paste0(selected_jars$passage, selected_jars$jar), ] %>% 
-  ungroup() %>% 
-  select(treat, rep, estimate, passage)
+  fluxes %>% 
+  select(passage, treat, estimate)
 
-# Mean parental (selected) flux for each passage
-parental <-
+# Pull out all offspring jars
+offspring <- 
+  fluxes %>%
+  select(treat, estimate, passage) %>% 
+  filter(passage != 1) %>% 
+  mutate(passage = passage - 1)
+```
+
+```{r heritability_mean}
+
+# Calculate means for east generation
+parental_mean <-
   parental %>% 
@@
 
-offspring <- 
-  fluxes %>% 
-  ungroup() %>%
-  select(treat, rep, estimate, passage)
-
-# Mean offspring flux for each passage
-offspring <-
+offspring_mean <-
   offspring %>% 
   group_by(passage, treat) %>% 
-  summarize(offspring = mean(estimate), .groups = "drop") %>%
-  filter(passage != 1) %>% 
-  mutate(passage = passage - 1)
+  summarize(offspring = mean(estimate), .groups = "drop")
+
+selected_mean <-
+  selected %>% 
+  group_by(passage, treat) %>% 
+  summarize(selected = mean(estimate), .groups = "drop")
+
+# Combine all parent/offspring data into one data frame
+heritability <- 
+  parental_mean %>% 
+  left_join(offspring_mean, by = c("passage", "treat")) %>% 
+  left_join(selected_mean, by = c("passage", "treat")) %>% 
+  #filter(treat == "p") %>% 
+  filter(passage != 5) %>% # Remove passage 5, which has no offspring
+  ungroup()
+
+write_tsv(heritability, '../Output/heritability.tsv')
+
+```
+
+``` {r generation_h2}
+
+######################################################################
+# This is where you stopped
+# Figure out why h2 is different from h2_calc_method
+# Finish writing your email to brendan about heritability
+######################################################################
+
+# Function for the Breeder's equation
+
+breeders <- function(R, S) {
+  R / S
+}
+
+# Calculate h2 (heritability) using math, not regression
+# For each generation and then calculate the mean/se
+h2_calc_method <-
+  heritability %>%
+  filter(treat == "p") %>%
+  mutate(S = selected - parental) %>%
+  mutate(R = offspring - parental) %>%
+  mutate(h2 = breeders(R, S)) %>%
+  summarize(S = mean(S),
+            R = mean(R),
+            se = se(h2),
+            h2 = mean(h2))
+
+h2_calc_method_no_2 <-
+  heritability %>%
+  filter(treat == "p") %>%  
+  filter(passage != 2) %>%
+  mutate(S = selected - parental) %>%
+  mutate(R = offspring - parental) %>%
+  mutate(h2 = breeders(R, S)) %>%
+  summarize(S = mean(S),
+            R = mean(R),
+            se = se(h2),
+            h2 = mean(h2))
+h2 <- 
+  h2_calc_method %>% 
+  mutate(method = "calc p")
+h2_no2 <- 
+  h2_calc_method_no_2 %>% 
+  mutate(method = "calc p no 2")
+
+h2 <- rbind(h2, h2_no2)
+```
+
+```{r h2_slope_method}
+
+# Calculate S (selection differential) with/without passage 2
+
+S <-
+  heritability %>%
+  filter(treat == "p") %>%
+  mutate(S = abs(selected - parental)) %>%
+  summarize(S = mean(S)) %>% 
+  pull()
+S_no2 <-
+  heritability %>%
+  filter(treat == "p") %>%
+  filter(passage != 2) %>% 
+  mutate(S = abs(selected - parental)) %>%
+  summarize(S = mean(S)) %>% 
+  pull()
+
+# For the equivalent plot of Falconer and MacKay
+# Scale and center all fluxes
+heritability_scaled <-
+  heritability %>% 
+  pivot_longer(parental:selected) %>% 
+  mutate(value = scale(value)) %>% 
+  pivot_wider(names_from = "name", values_from = "value")
+
+# Calculate R and S 
+R_and_S <-
+  heritability_scaled %>% 
+  filter(treat == "p") %>% 
+  summarize(R = mean(offspring),
+            S = mean(selected))
+
+# Calculate scaled h2
+R <- R_and_S[[1, 'R']]
+S <- R_and_S[[1, 'S']]
+h2 <- breeders(R, S)
+
+# Plot scaled h2
+ggplot(data = heritability_scaled, aes(x = selected, y = offspring)) +
+  geom_point(aes(shape = treat)) +
+  scale_shape_manual(values = c(1,19), name = "Treatment", labels = c('Neutral', 'Positive')) +
+  stat_smooth(se = FALSE, method = "lm", formula = y ~ x, color = 'black') +
+  labs(x = "Selected Parents", y = "Offspring") +
+  geom_segment(aes(x = 0, y = R, xend = S, yend = R), linetype = 3) +
+  geom_segment(aes(x = S, y = 0, xend = S, yend = R), linetype = 3) +
+  geom_hline(yintercept = 0) +
+  geom_vline(xintercept = 0) +
+  geom_text(label = "R", x = -0.1, y = R) +
+    geom_text(label = "S", x = S, y = -0.2)
+
 ```
@@
 ```{r}
+heritability %>% 
+  ggplot(aes(parental, offspring, color = treat)) +
+  geom_point() +
+  labs(x = "Parental", y = "Offspring") +
+  theme_bw() +
+  scale_color_discrete(name = "Treatment", labels = c('Neutral', 'Positive'))
 
-###################################################
-# This is the nest place you stopped
-# You need to calculate the slope of midoffspring
-# on midparent for the Positive treatments to get R
-# and then calculate the selection differential
-# as above S. S doesn't make sense for the neutral
-# treatment.
-###################################################
-
-# offspring <-
-#   fluxes %>%
-#   ungroup() %>%
-#   select(estimate, passage, treat) %>%
-#   mutate(offspring = rank(estimate))
-
-heritability <- left_join(parental, offspring, by = c("passage", "treat"))
-heritability$passage <- factor(heritability$passage)
-heritability$treat <- factor(heritability$treat)
-heritability <-
-  filter(heritability, passage != 5)
-# heritability$parental <- rank(heritability$parental)
-
-write_tsv(heritability, '../Output/heritability.tsv')
-
-ggplot(heritability, aes(x = parental, y = offspring)) +
-  geom_point(aes(color = treat)) +
-  stat_smooth(method = 'lm') + 
-  labs(x = "Parental Mean Flux (-k)", y = "Offspring Flux (-k)") +
-  scale_color_discrete(name = "Selection\nTreatment", labels = c('Neutral', 'Positive')) +
-  theme_bw()
-
-# ggsave('heritability.pdf', width = 4, height = 3)
-# ggsave('heritability.png', width = 4, height = 3)
-
+summary(lm(offspring ~ parental, data = heritability))
 ```
 ```{r}
-# This chunk calculates the slope of mid-offspring on mid-parent for all generations and all treatments
-
-fluxes
+heritability %>% 
+  ggplot(aes(selected, offspring, color = treat)) +
+  geom_point() +
+  labs(x = "Selected", y = "Offspring") +
+  theme_bw() +
+  scale_color_discrete(name = "Treatment", labels = c('Neutral', 'Positive'))
 ```
-
-```{r breeders}
-
-h2 <- function(R, S) {
-  R / S
-}
-
-# R = Response to selection, difference in mean between offspring of selected parents and all of the parents
-
-# S = Selection differental, difference in mean between selected parents and all of the parents
-
-parental_all <-
-  fluxes %>% 
-  select(passage, treat, estimate) %>% 
-  group_by(passage, treat) %>% 
-  summarize(all = mean(estimate), .groups = "drop")
-
-parental_selected <- parental
-
-selection_diff <- 
-  left_join(parental_selected, parental_all) %>% 
-  rename(selected = parental) %>% 
-  mutate(S = selected - all)
-
-heritability_ns <- 
-  parental_all %>% 
-  left_join(parental_selected) %>% 
-  left_join(offspring) %>% 
-  filter(passage != 5)
-
-heritability_only_p <- 
-  heritability_ns %>% 
-  filter(treat == 'p')
-
-R <- tidy(lm(offspring ~ all, data = heritability_only_p))[[2, 2]]
-S <- tidy(lm(parental ~ all, data = heritability_only_p))[[2, 2]]
-h2(R, S)
-
-# Generational R, S, global h2, 0.82
-heritability_only_p %>% 
-  mutate(R = offspring - all,
-         S = parental - all) %>% 
-  summarize(R = mean(R),
-            S = mean(S)) %>%
-  mutate(h2 = h2(R, S))
-
-# Generational R, S, h2, 0.73
-heritability_only_p %>% 
-  #filter(passage != 2) %>% 
-  mutate(R = offspring - all,
-         S = parental - all,
-         h2 = h2(R, S)) %>% 
-  summarize(se = se(h2),
-            h2 = mean(h2))
-
-# Global R, S, h2, 0.82
-heritability_only_p %>% 
-  summarize(offspring = mean(offspring),
-         all = mean(all),
-         parental = mean(parental)) %>% 
-  mutate(R = offspring - all,
-         S = parental - all,
-         h2 = h2(R, S))
-heritability_ns %>% 
-  pivot_longer(all:offspring, names_to = "group", values_to = "mean") %>% 
-ggplot(., aes(x = passage, y = mean, color = group)) +
-  facet_wrap(~ treat) +
-  geom_point()
-```

### 2020-10-23 R/03-response.R
try robust regression for the test of response to selection

@@ -11,6 +11,7 @@
 library(tidyverse)
 library(metafor)
 library(morris)
+library(broom)
 
 # Standard error of the difference function
 sed <- function(p, n, na.rm = FALSE) {
@@ -91,3 +92,66 @@ write_tsv(deviance, '../Output/deviance.tsv')
 saveRDS(rma_output, '../Output/response.rds')
 saveRDS(fit, '../Output/dev_fit.rds')
 
+
+
+# Calculate deviance as percentage of p - n
+deviance <-
+  fluxes %>% 
+  select(treat, passage, estimate) %>% 
+  group_by(passage, treat) %>% 
+  summarize(mean = mean(estimate), .groups = "drop") %>% 
+  pivot_wider(names_from = treat, values_from = mean) %>% 
+  mutate(deviance = (p - n) / n)
+
+# Calculate variance of deviance as SED
+deviance$se <-
+  fluxes %>% 
+  select(treat, passage, estimate) %>% 
+  mutate(x = 1:nrow(fluxes)) %>% 
+  pivot_wider(names_from = treat, values_from = estimate) %>% 
+  group_by(passage) %>% 
+  summarize(sed = sed(p, n, na.rm = TRUE), .groups = "drop") %>%
+  pull(sed)
+
+# Fit deviance model
+fit <- rma(deviance ~ passage, se^2, method="FE", data = deviance)
+
+plot(fit)
+
+# Plot deviance
+ggplot(deviance, aes(x = passage, y = deviance, ymin = deviance - se, ymax = deviance + se)) + 
+  geom_pointrange(color = 'darkorange2') + 
+  geom_abline(intercept = fit$beta[1], slope = fit$beta[2])
+
+# Take deviance model parameters and export into a data.frame
+rma_output <- data.frame(
+Estimate = fit$beta,
+se = fit$se,
+z.value = fit$zval,
+p.value = fit$pval,
+upper.ci = fit$ci.lb,
+lower.ci = fit$ci.ub
+)
+
+# Fit separte slopes for Positive and Neutral with robust M-estimator regression
+
+library(MASS)
+
+fit <- rlm(estimate ~ passage * treat, data = fluxes)
+
+summary(fit)
+anova(fit)
+new_data <- cbind(fluxes, predict(fit, interval = 'confidence'))
+colnames(new_data) 
[…]

### 2020-10-23 R/04-heritability.R
start cumulative response calculation for heritabiality

@@ -104,3 +104,19 @@ h2_no2 <-
 
 h2 <- rbind(h2, h2_no2)
 
+# Cumulative response to selection
+
+parent <- heritability[[1, 3]]
+offspr <- heritability[[8, 4]]
+select <- heritability[[8, 5]]
+R = offspr - parent
+S = select - parent
+h2 = breeders(R, S)
+
+heritability$cum_res <- NULL
+heritability$cum_sel <- c()
+for (i in nrow(heritability)) {
+  cum_res <- cbind(cum_res, heritability[i, 4] - heritability[i, 3])
+  cum_sel <- cbind(cum_sel, heritability[i, 5] - heritability[i, 3])
+}
+cum_res

### 2020-10-23 Manuscript/manuscript.Rmd
update presentation
update manuscript text

@@
 
-Microbial communities mediate ecosystem nutrient cycling via decomposition, greenhouse gas cycling,
-carbon storage, and other processes.
-Despite the central role of bacteria in performing these processes, we have not found a consistent relationship between the composition of a microbial community and the rate of an ecosystem process across habitats.
-Typically, research in this area has followed the model of macroorganismal biodiversity-ecosystem function studies such as grassland productivity studies that measure or manipulate the diversity of plant species and correlate that with the rate of primary productivity.
-By analogy, microbial ecologists use microbial sequencing data to create taxonomic relative abundance tables and then infer diversity or relative abundance from the whole archaeal/bacterial community and correlate that with the rate of a microbial process.
-Alternatively, they estimate the abundance of a putative functional gene (such as *pmoA* for methane oxidation) and correlate abundance with function.
-These approaches only rarely show significant, positive relationships between biodiversity and ecosystem function.
-In my first chapter published in Phil Trans I lay out philosophically why this approach shouldn’t work – namely that marker gene sequences are not analogous to plant species counts, but are more analogous to measurements of gene markers as in evolutionary biology studies.
+Microbial communities mediate ecosystem nutrient cycling via decomposition,
+greenhouse gas cycling, carbon storage, and other processes. Despite the
+central role of bacteria in performing these processes, we have not found a
+consistent relationship between the composition of a microbial community and
+the rate of an ecosystem process across habitats. Typically, research in this
+area has followed the model of macroorganismal biodiversity-ecosystem function
+studies such as grassland productivity studies that measure or manipulate the
+diversity of plant species and correlate that with the rate of primary
+productivity. By analogy, microbial ecologists use microbial sequencing data to
+create taxonomic relative abundance tables and then infer diversity or relative
+abundance from the whole archaeal/bacterial community and correlate that with
+the rate of a microbial process. Alternatively, they estimate the abundance of
+a putative functional gene (such as *pmoA* for methane oxidation) and correlate
+abundance with function. These approaches only rarely show significant,
+positive relationships between biodiversity and ecosystem function. In my first
+chapter published in Phil Trans I lay out philosophically why this approach
+shouldn’t work – namely that marker gene sequences are not analogous to plant
+species counts, but are more analogous to measurements of gene markers as in
+evolutionary biology studies.
 
-In evolutionary biology, genotype-phenotype mapping provides a way to understand the fitness effects of a genotype.
-While selection typically acts on phenotype, the genotypic changes are the fundamental driver of phenotypic differences and thus fitness differences (Stadler and Stephes 2005).
-A similar analogy can be drawn for biodiversity-ecosystem function relationships.
-Traits of a community are what drive community assembly, not the taxa or genotypes themselves (Inkpen et al.
-2017).
-However, as ecologists, we wish to understand the effects of species on ecosystem function.
-Therefore, in a similar way that evolutionary biologists want to deduce the genotype-phenoype map to understand the genetic basis of fitness, community ecologists must describe the "mapping" between community genotype and ecoystem function to understand what drives communty assembly, and ultimately, ecosystem function.
+In evolutionary biology, genotype-phenotype mapping provides a way to
+understand the fitness effects of a genotype. While selection typically acts on
+phenotype, the genotypic changes are the fundamental driver of phenotypic
+differences and thus fitness differences (Stadler and Stephes 2005). A similar
+analogy can be drawn for biodiversity-ecosystem function relationships. Traits
+of a community are what drive community assembly, not the taxa or genotypes
+themselves (Inkpen et al. 2017). However, as ecologists, we wish to understand
+the effects of species on ecosystem function. Therefore, in a similar way that
+evolutionary biologists want to deduce the genotype-phenoype map to understand
+the genetic basis of fitness, community ecologists must describe the "mapping"
+between community genotype and ecoystem function to understand what drives
+communty assembly, and ultimately, ecosystem function.
 
-The relationship that ecologists have described between plant diversity and productivity represents a relatively simple map between taxonomy and function.
-The number of unique members of a functional group (i.e., plant species) is correlated with the rate of the function performed by that group.
-However, for microbial communities that mapping is not quite as clear.
-For many microbial traits, there is not a clear coupling between taxonomy and function, except for in the case of highly conserved traits involving many genes (Martiny et al.
-2013).
-In addition, most microbial taxa have not been sampled and so we don't have a full picture of which taxa are members of a particular functional group (citation).
-Put another way, it's much harder to identify methanotrophic prokaryotes in soil than it is to identify plants in a prairie.
-Finally, even for highly conserved traits with a putative functional marker, for example methane oxidation and the gene *pmoA*, functional group abundance is not correlated with the rate of the function performed by that group in most ecosystems (Rocca et al.
-2016).
-Because of these differences between macroorganisms and microorganisms, it may be more useful to approach microbial biodiversity-ecosystem function in a way analogus to genotype-phenotype mapping in organisms (Morris et al.
-2020).
+The relationship that ecologists have described between plant diversity and
+productivity represents a relatively simple map between taxonomy and function.
+The number of unique members of a functional group (i.e., plant species) is
+correlated with the rate of the function performed by that group. However, for
+microbial communities that mapping is not quite as clear. For many microbial
+traits, there is not a clear coupling between taxonomy and function, except for
+in the case of highly conserved traits involving many genes (Martiny et al.
+2013). In addition, most microbial taxa have not been sampled and so we don't
+have a full picture of which taxa are members of a particular functional group
+(citation). Put another way, it's much harder to identify methanotrophic
+prokaryotes in soil than it is to identify plants in a prairie. Finally, even
+for highly conserved traits with a putative functional marker, for example
+methane oxidation and the gene *pmoA*, functional group abundance is not
+correlated with the rate of the function performed by that group in most
+ecosystems (Rocca et al. 2016). Because of these differences between
+macroorganisms and microorganisms, it may be more useful to approach microbial
+biodiversity-ecosystem function in a way analogus to genotype-phenotype mapping
+in organisms (Morris et al. 2020).
 
-One approach to deducing the genotype-phenotype map in organisms is through artificial selection (Fuller et al.
-2005).
-In this approach, selection for a trait is imposed in a contrived laboratory environment and compared to laboratory environment without selection on that trait.
-By comparing the genetic respones between these two treatments, you can remove the environment-fitness connection that would be present in a comparative study and identify genes that are evolving in response to selection only on the trait of interest.
-It has been shown that whole communities can respond to artifical selection on traits of the ecosystem (Goodnight et al.
-1997, Williams and Lenton 2007).
-And it has been proposed that this could be a suitable approach for identify communities of organisms that are capable of performing a particular function at a high rate (Swenson et al.
-2000).
-Therefore, artificial ecosystem selection is a potentially fruitful approach to identifying the map between biodiversity and ecosystem function in microbial communities.
+One approach to deducing the genotype-phenotype map in organisms is through
+artificial selection (Fuller et al. 2005). In this approach, selection for a
+trait is imposed in a contrived laboratory environment and compared to
+laboratory environment without selection on that trait. By comparing the
+genetic respones between these two treatments, you can remove the
+environment-fitness connection that would be present in a comparative study and
+identify genes that are evolving in response to selection only on the trait of
+interest. It has been shown that whole communities can respond to artifical
+selection on traits of the ecosystem (Goodnight et al. 1997, Williams and
+Lenton 2007). And it has been proposed that this could be a suitable approach
+for identify communities of organisms that are capable of performing a
+particular function at a high rate (Swenson et al. 2000). Therefore, artificial
+ecosystem selection is a potentially fruitful approach to identifying the map
+between biodiversity and ecosystem function in microbial communities.
 
-To identify the map between microbial community composition and ecosystem function, we imposed artificial ecosystem selection on methane oxidation rate of whole soil ecosystems.
-We prepared two lines of replicate ecosystems: one in which we imposed positive selection on methane oxidation by only using the top performing ecosystems to generate the next set of ecosystems and one in which we imposed neutral selection where a random set of ecosystems were chosen to generate the next set of ecosystems.
-We repated this process over multiple passages and then compared the resultant metacommunities in terms of their rate of ecosystem function and the composition of their communities.
+To identify the map between microbial community composition and ecosystem
+function, we imposed artificial ecosystem selection on methane oxidation rate
+of whole soil ecosystems. We prepared two lines of replicate ecosystems: one in
+which we imposed positive selection on methane oxidation by only using the top
+performing ecosystems to generate the next set of ecosystems and one in which
+we imposed neutral selection where a random set of ecosystems were chosen to
+generate the next set of ecosystems. We repated this process over multiple
+passages and then compared the resultant metacommunities in terms of their rate
+of ecosystem function and the composition of their communities.
 
-With this experiment, we asked several questions.
-First, is there a response to selection on methane oxidation by soil ecosystems?
-And how strong is the response to selection?
-(i.e., how "heritable" is methane oxidation as a community trait in terms of the phenotypic resemblence of ecosystems to the ecosystems that generated them?) If methane oxidation is heritable, then what is the nature of the map between microbial community composition and methane oxidation?
-Is it the diversity of methane oxidizers, as it is for plants and productivity?
-Or is there a more complex map involving upstream metabolic regulation?
+With this experiment, we asked several questions. First, is there a response to
+selection on methane oxidation by soil ecosystems? And how strong is the
+response to selection? (i.e., how "heritable" is methane oxidation as a
+community trait in terms of the phenotypic resemblence of ecosystems to the
+ecosystems that generated them?) If methane oxidation is heritable, then what
+is the nature of the map between microbial community composition and methane
+oxidation? Is it the diversity of methane oxidizers, as it is for plants and
+productivity? Or is there a more complex map involving upstream metabolic
+regulation?
 
@@
 ## Selection experiment
-We performed an artificial ecosystem selection experiment similar to Swenson et al.
-(2000) by passaging replicate soil microbial communities.
-The trait we selected on was CH~4~ oxidation rate.
-Our experiment had two selection lines with twelve jars each for a total of twenty-four jars per passage.
-One line was a positive selection line where the two or three jars with the highest CH~4~ oxidation rate were chosen to inoculate the next set of jars.
-The other line was a neutral selection line where an equal number of jars as the positive line were chosen at random among the twelve.
-The number of jars chosen was based on the distribution of fluxes among the positive jars, i.e., we chose the top three jars unless only two jars had considerably greater oxidation rates based on visual inspection of histograms.
+We performed an artificial ecosystem selection experiment similar to Swenson et
+al. (2000) by passaging replicate soil microbial communities. The trait we
+selected on was CH~4~ oxidation rate. Our experiment had two selection lines
+with twelve jars each for a total of twenty-four jars per passage. One line was
+a positive selection line where the two or three jars with the highest CH~4~
+oxidation rate were chosen to inoculate the next set of jars. The other line
+was a neutral selection line where an equal number of jars as the positive line
+were chosen at random among the twelve. The number of jars chosen was based on
+the distribution of fluxes among the positive jars, i.e., we chose the top
+three jars unless only two jars had considerably greater oxidation rates based
+on visual inspection of histograms.
 
-The initial soil microbial community was sampled from the top 10 cm of an upland mineral soil under a deciduous forest ecosystem near the University of Oregon campus.
-Incubations were performed in 500 mL mason jars (Ball Corporation, Broomfield, Colorado, USA) with rubber septa installed in the lids.
-Jars were incubated at ambient temperature on the benchtop.
-To create the first set of jars, twenty-four jars were created in an identical manner and then randomly assigned to either the positive or negative selection line.
-Each jar was sterilized with 70% ethanol to which was added 45 g of autoclaved potting mix, 5 g of live soil, and 3.5 mL of sterile deionized water to bring them to <mark> 60%???
-of field capacity </mark>.
-The jars were then homogenized, capped, and injected with 4.3 mL of 99% CH~4~ to bring the headspace concentration to approximately 1000 ppm CH~4~.
-Jars were flushed and respiked twice per week to maintain elevated CH~4~ concentrations until travel, holidays, and COVID-19 got in the way and then we just did it as often as we could.
-Jars were incubated for approximately three weeks and then flux measurements were taken.
+The initial soil microbial community was sampled from the top 10 cm of an
+upland mineral soil under a deciduous forest ecosystem near the University of
+Oregon campus. Incubations were performed in 500 mL mason jars (Ball
+Corporation, Broomfield, Colorado, USA) with rubber septa installed in the
+lids. Jars were incubated at ambient temperature on the benchtop. To create the
+first set of jars, twenty-four jars were created in an identical manner and
+then randomly assigned to either the positive or negative selection line. Each
+jar was sterilized with 70% ethanol to which was added 45 g of autoclaved
+potting mix, 5 g of live soil, and 3.5 mL of sterile deionized water to bring
+them to <mark> 60%??? of field capacity </mark>. The jars were then
+homogenized, capped, and injected with 4.3 mL of 99% CH~4~ to bring the
+headspace concentration to approximately 1000 ppm CH~4~. Jars were flushed and
+respiked twice per week to maintain elevated CH~4~ concentrations until travel,
+holidays, and COVID-19 got in the way and then we just did it as often as we
+could. Jars were incubated for approximately three weeks and then flux
+measurements were taken.
 
-Once CH~4~ oxidation rates were determined, the top two or three jars were chosen for selection in the positive line and an equal number of jars were randomly chosen from the neutral line.
-The selected jars for each line were homogenized and used as the 5 g live soil for the subsequent passage.
-Twelve new jars were created for each line in the same manner as above using the inoculum for that line.
-This procedure continued until a divergence was observed in the mean flux of each line for two consecutive generations.
-Jars in passage number 2 and 5 were homogenized and a sample was collected and stored at -80$^\circ$C for later DNA extraction.
+Once CH~4~ oxidation rates were determined, the top two or three jars were
+chosen for selection in the positive line and an equal number of jars were
+randomly chosen from the neutral line. The selected jars for each line were
+homogenized and used as the 5 g live soil for the subsequent passage. Twelve
+new jars were created for each line in the same manner as above using the
+inoculum for that line. This procedure continued until a divergence was
+observed in the mean flux of each line for two consecutive generations. Jars in
+passage number 2 and 5 were homogenized and a sample was collected and stored
+at -80$^\circ$C for later DNA extraction.
 
-## Methane flux measurements
-Methane oxidation rates were determined after flushing and spiking jars to approximately 1000 ppm CH~4~.
-Head-space samples of 1 mL were collected from each jar immediately after spiking and then at time points 3, 6, 24, and 48 hours for a 5-point curve.
-Samples were immediately injected into a <mark>???</mark> gas chromatograph fitted with an electron capture detector (company, city, state, USA) to determine the headspace CH~4~ concentration.
-Fluxes were calculated from a first-order exponential decay function as decay constant `k`.
-Oxidation rates are presented as the additive inverse of `k` (i.e., `-k`) so that a more positive value represents a faster oxidation rate.
+## Methane flux measurements Methane oxidation rates were determined after
+flushing and spiking jars to approximately 1000 ppm CH~4~. Head-space samples
+of 1 mL were collected from each jar immediately after spiking and then at time
+points 3, 6, 24, and 48 hours for a 5-point curve. Samples were immediately
+injected into a <mark>???</mark> gas chromatograph fitted with an electron
+capture detector (company, city, state, USA) to determine the headspace CH~4~
+concentration. Fluxes were calculated from a first-order exponential decay
+function as decay constant `k`. Oxidation rates are presented as the additive
+inverse of `k` (i.e., `-k`) so that a more positive value represents a faster
+oxidation rate.
 
@@
 
-Soil DNA was extracted from a 0.25 g subsample of soil from the starting inoculum and from each jar in passage 2 and 5.
-Negative controls were extracted from autoclaved potting mix and DNase-free water.
-Extractions were performed using the DNeasy PowerSoil kit (QIAGEN, Hilden, Germany) and quantified using Qubit (company, city, state).
-In order to estimate the diversity and relative abundance of the bacterial and archaeal taxa in our soil ecosystems, we sequenced the V3-V4 region of the 16S rRNA gene using the Earth Microbiome Project protocol and primers<mark>???</mark>.
-In addition, to estimate the diversity and relative abundance of CH~4~ oxidizing taxa, we sequenced the particulate methane monooxygenase subunit A gene (*pmoA*) using primers<mark>???</mark>.
-Sequencing was performed on the Illumina MiSeq in v3 mode with 300 bp paired-end reads.
+Soil DNA was extracted from a 0.25 g subsample of soil from the starting
+inoculum and from each jar in passage 2 and 5. Negative controls were extracted
+from autoclaved potting mix and DNase-free water. Extractions were performed
+using the DNeasy PowerSoil kit (QIAGEN, Hilden, Germany) and quantified using
+Qubit (company, city, state). In order to estimate the diversity and relative
+abundance of the bacterial and archaeal taxa in our soil ecosystems, we
+sequenced the V3-V4 region of the 16S rRNA gene using the Earth Microbiome
+Project protocol and primers<mark>???</mark>. In addition, to estimate the
+diversity and relative abundance of CH~4~ oxidizing taxa, we sequenced the
+particulate methane monooxygenase subunit A gene (*pmoA*) using
+primers<mark>???</mark>. Sequencing was performed on the Illumina MiSeq in v3
+mode with 300 bp paired-end reads.
 
-## Bioinformatics
-Bioinformatic analyses were conducted using the `QIIME2` bioinformatics platform (Bolyen et al.
-2019).
-Raw sequencing reads were demultiplexed using ???
-(citation), denoised using `DADA2` (citation).
-Taxonomic assignment for the 16S reads was performed using ???
-based on the Silva database.
-All subsequent tatistical analyses were performed in the `R` statistical programming environment (R Core Team 2020).
-Richness estimates were made using the `breakaway` package and diversity indices using the `DivNet` package (Willis et al.
-???).
+## Bioinformatics Bioinformatic analyses were conducted using the `QIIME2`
+bioinformatics platform (Bolyen et al. 2019). Raw sequencing reads were
+demultiplexed using ??? (citation), denoised using `DADA2` (citation).
+Taxonomic assignment for the 16S reads was performed using ??? based on the
+Silva database. All subsequent tatistical analyses were performed in the `R`
+statistical programming environment (R Core Team 2020). Richness estimates were
+made using the `breakaway` package and diversity indices using the `DivNet`
+package (Willis et al. ???).
 
@@
 
-To test whether there was a significant change in CH~4~ oxidation rate (k) as a response to selection, we tested the deviation between the positive and neutral selection lines over multiple passages.
-We calculated deviation as the difference in the mean oxidation rate between the positive and neutral treatment within each passage.
-The standard errors of the deviations were calculated as the square root of the sum of the squared errors of the positive and neutral lines.
-We then fit an ordinary least squares regression model taking into account the standard errors of the deviations.
-The model was fit using the `rma` function from the `metafor` package in `r` (Viechtbauer 2010).
+To test whether there was a significant change in CH~4~ oxidation rate (k) as a
+response to selection, we tested the deviation between the positive and neutral
+selection lines over multiple passages. We calculated deviation as the
+difference in the mean oxidation rate between the positive and neutral
+treatment within each passage. The standard errors of the difference were
+calculated as the square root of the sum of the squared errors of the positive
+and neutral lines. We then fit an ordinary least squares regression model
+taking into account the standard errors of the deviations. The model was fit
+using the `rma` function from the `metafor` package in `r` (Viechtbauer 2010).
 
@@
 
-To test the heritability of CH~4~ oxidation rate as an ecosystem property, we calculated narrow-sense heritability ($h^2$) as the regression of offspring on mid-parent (Goodnight 2000, Falconer 1997).
-The mid-parent was the mean for all three selected jars and the offspring was the mean for all jars produced by those parents.
+To estimate the proportion of variance in ecosystem function due to additive
+genetic variance of the microbial communtiy, we applied the Breeder's equation
+to the results of our selection experiment. the heritability of
+CH~4~ oxidation rate as an ecosystem property, we calculated narrow-sense
+heritability ($h^2$) as the regression of offspring on mid-parent (Goodnight
+2000, Falconer 1997). The mid-parent was the mean for all three selected jars
+and the offspring was the mean for all jars produced by those parents.
 
@@
 
-To identify taxa that responded to selection on CH~4~ oxidation rate, we tested differential abundance using a beta-binomial model (Martin et al.
-2020).
-We compared taxa in passage 5 between the positive and neutral treatment to identify taxa that
+To identify taxa that responded to selection on CH~4~ oxidation rate, we tested
+differential abundance using a beta-binomial model (Martin et al. 2020). We
+compared taxa in passage 5 between the positive and neutral treatment to
+identify taxa that
+
+Possible tests:
+
+- Random forest or gradient boosting to predict Positive vs. Neutral.
 
@@
 
-ggsave('new_herit_fig.pdf', width = 5, height = 3)
-
 ```

### 2020-10-26 R/04-heritability.R
calculate cumulative hertiability

@@ -68,6 +68,12 @@ lm_fit <-
 R_not_dev <- pull(tidy(lm_fit)[2, 2])
 h2_reg_abs <- breeders(R_not_dev, S)
 
+# Calculate R as regression of offspring on parents
+
+ggplot(heritability, aes(x = parental, y = offspring)) +
+  geom_point() + 
+  stat_smooth(method = 'lm', se = FALSE)
+summary(lm(offspring ~ parental, data = heritability, subset = treat == 'p'))
 # Calculate h2 (heritability) using math, not regression
 # For each generation and then calculate the mean/se
 h2_calc_method <-
@@ -113,10 +119,22 @@ R = offspr - parent
 S = select - parent
 h2 = breeders(R, S)
 
-heritability$cum_res <- NULL
-heritability$cum_sel <- c()
-for (i in nrow(heritability)) {
-  cum_res <- cbind(cum_res, heritability[i, 4] - heritability[i, 3])
-  cum_sel <- cbind(cum_sel, heritability[i, 5] - heritability[i, 3])
+heritability[,'cum_res'] <- NA
+heritability$cum_sel <- NA
+herit_p <- filter(heritability, treat == 'p')
+for (i in seq(nrow(herit_p))) {
+  off <- sum(herit_p$offspring[1:i])
+  par <- sum(herit_p$parental[1:i])
+  sel <- sum(herit_p$selected[1:i])
+  herit_p$cum_res[i] <- off - par
+  herit_p$cum_sel[i] <- sel - par
 }
-cum_res
+herit_p$cum_h2 <- herit_p$cum_res / herit_p$cum_sel
+
+ggplot(herit_p, aes(x = cum_sel, y = cum_res)) + 
+  geom_point() + 
+  stat_smooth(method = 'lm', se = FALSE)
+
+summary(lm(cum_res ~ cum_sel, data = herit_p))
+
+

### 2020-11-04 R/01-process-flux.R
clean up fluxes dataframe

@@ -125,6 +125,6 @@ fluxes <-
   ungroup() %>% 
   mutate(passage = as.numeric(flux_date)) %>% 
   mutate(estimate_rank = rank(estimate)) %>% 
-  filter(passage != 6)
-
+  filter(passage != 6) %>% 
+  select(-std.error, -statistic, -p.value)
 write_tsv(fluxes, '../Output/fluxes.tsv')

### 2020-11-04 Manuscript/manuscript.Rmd
add peter's suggestions

@@
   theme_bw()
-
+ggsave('deviance.pdf')
 ```

### 2020-11-04 R/03-response.R

@@ -136,8 +136,9 @@ lower.ci = fit$ci.ub
 # Fit separte slopes for Positive and Neutral with robust M-estimator regression
 
 library(MASS)
-
-fit <- rlm(estimate ~ passage * treat, data = fluxes)
+anova(
+fit <- rlm(estimate ~ passage * treat, data = fluxes),
+fit <- rlm(estimate ~ passage + treat, data = fluxes))
 
 summary(fit)
 anova(fit)
@@ -155,3 +156,7 @@ ggplot(new_data, aes(x = passage, y = estimate)) +
   geom_line(aes(passage, fit, color = treat)) +
   geom_ribbon(aes(ymin=lwr,ymax=upr, group = treat), alpha=0.3) +
   ylim(-0.01, 0.08)
+ggplot(fluxes, aes(x = passage, y = estimate, color = treat)) + 
+  geom_point() + 
+  stat_smooth(method='lm')
+ggsave('raw_fluxes.pdf')

### 2020-11-04 R/04-heritability.R

@@ -175,5 +175,4 @@ ggplot(heritability, aes(parental, offspring, color = treat)) +
   theme_classic() +
   geom_point() + 
   stat_smooth(method = 'lm', se = FALSE)
-summary(lm(offspring ~ parental, data = heritability, subset = treat == 'p'))
-summary(lm(offspring ~ parental, data = heritability, subset = treat == 'n'))
+summary(lm(offspring ~ parental * treat, data = heritability))

### 2020-11-05 R/01-process-flux.R
rerun the analysis with log-transformed value and generate new figures and statistics

@@ -127,4 +127,8 @@ fluxes <-
   mutate(estimate_rank = rank(estimate)) %>% 
   filter(passage != 6) %>% 
   select(-std.error, -statistic, -p.value)
+
+
+fluxes$estimate_log10 <- log10(fluxes$estimate + abs(min(fluxes$estimate)) + 0.001)
+
 write_tsv(fluxes, '../Output/fluxes.tsv')

### 2020-11-05 R/03-response.R

@@ -23,6 +23,70 @@ sed <- function(p, n, na.rm = FALSE) {
 
 fluxes <- read_tsv('../Output/fluxes.tsv')
 
+#####################################################################
+# Okay here to deviance is how I should really be doing it
+###############################################################
+
+# 1. Plot the raw data
+ratio.values <- (max(fluxes$passage)-min(fluxes$passage))/(max(fluxes$estimate)-min(fluxes$estimate))
+
+ggplot(fluxes, aes(x = passage, y = estimate, color = treat)) +
+  theme_classic() +
+  geom_jitter(width = 0.1) +
+  stat_smooth(method = 'lm', se = FALSE, formula = 'y ~ x') +
+#  coord_fixed(ratio.values) +
+  labs(x = "Passage Number", y = "Methane Oxidation Rate (-k)") +
+  scale_color_manual(name = "Treatment", labels = c('Neutral', 'Positive'), values = c('gray40', 'darkorange2'))
+
+ggsave('fluxes_raw.pdf', width = 5, height = 4)
+
+# 2. Transform it to make it more normal
+
+# 3. Plot the transformed data
+
+ratio.values <- (max(fluxes$passage)-min(fluxes$passage))/(max(fluxes$estimate_log10)-min(fluxes$estimate_log10))
+
+ggplot(fluxes, aes(x = passage, y = estimate_log10, color = treat)) +
+  theme_classic() +
+  geom_jitter(width = 0.1) +
+  stat_smooth(method = 'lm', se = FALSE, formula = 'y ~ x') +
+#  coord_fixed(ratio.values) +
+  labs(x = "Passage Number", y = "log10(Methane Oxidation Rate (-k))") +
+  scale_color_manual(name = "Treatment", labels = c('Neutral', 'Positive'), values = c('gray40', 'darkorange2'))
+
+ggsave('fluxes_log.pdf', width = 5, height = 4)
+
+
+
+# 3. Test a difference of slopes
+
+fit <- lm(estimate_log10 ~ passage * treat, data = fluxes)
+summary(fit)
+
+fit1 <- lm(estimate_log10 ~ passage + treat, data = fluxes)
+summary(fit1)
+anova(
+      lm(estimate_log10 ~ passage * treat, data = fluxes),
+      lm(estimate_log10 ~ passage + treat, data = fluxes)
+)
+
+# 4. Plot the back-transformed slope
+
+summary(fit)
[…]

### 2020-11-05 R/04-heritability.R

@@ -25,6 +25,61 @@ library(broom)
 
 heritability <- read_tsv('../Output/heritability.tsv')
 fluxes <- read_tsv('../Output/fluxes.tsv')
+selected_jars <- read_csv('../Data/selected.csv')
+
+#######################################################################
+# The way I should be doing
+########################################################################
+# Pull out the selected parental jars
+selected <- 
+  fluxes[paste0(fluxes$passage, fluxes$jar) %in% paste0(selected_jars$passage, selected_jars$jar), ] %>%
+  select(treat, estimate_log10, passage) %>% 
+  group_by(treat, passage) %>% 
+  summarize(parental = mean(estimate_log10), .groups = 'drop')
+
+offspring <- 
+  fluxes %>% 
+  select(passage, treat, estimate_log10) %>% 
+  filter(passage != 1) %>% 
+  mutate(passage = passage - 1) %>% 
+  rename(offspring = estimate_log10)
+
+heritability <- 
+  offspring %>% 
+  left_join(selected, by = c('passage', 'treat'))
+
+
+ggplot(heritability, aes(parental, offspring)) +
+  theme_classic() +
+  geom_jitter(width = 0.01) + 
+  stat_smooth(method = 'lm', se = FALSE, formula = 'y ~ x') +
+  labs(x = "Mid-parent (mean(log10(Flux (-k))))", y = "Offspring (log10(Flux (-k)))")
+ggsave('herit_all.pdf', width = 5, height = 4)
+
+summary(lm(offspring ~ parental, data = heritability))
+
+ggplot(heritability, aes(parental, offspring, color = treat)) +
+  theme_classic() +
+  geom_jitter(width = 0.01) + 
+  stat_smooth(method = 'lm', se = FALSE, formula = 'y ~ x') +
+  labs(x = "Mid-parent (mean(log10(Flux (-k))))", y = "Offspring (log10(Flux (-k)))") +
+  scale_color_manual(name = "Treatment", labels = c('Neutral', 'Positive'), values = c('gray40', 'darkorange2'))
+
+ggsave('herit_treat.pdf', width = 5, height = 4)
+
+summary(lm(offspring ~ parental * treat, data = heritability))
+
+anova(
+      lm(offspring ~ parental * treat, data = heritability),
+      lm(offspring ~ parental + treat, data = heritability)
+)
+
+      summary(lm(offspring ~ parental * treat, data = heritability))
+      summary(lm(offspring ~ parental + treat, data = heritability))
+
+######################################################################
[…]

### 2021-06-16 R/01-process-flux.Rmd
convert processing script to rmd

---
title: "Process Flux Date"
author: "Andrew"
date: "6/15/2021"
output: html_document
---

```{r initialize_packages}
library(tidyverse)
library(broom)
library(lubridate)
library(knitr)
```

```{r initialize_variables}
raw_dir <- '../Data/'
der_dir <- '../Output/'
```

```{r import_data}
sc <- read.csv(paste0(raw_dir, 'standard_curve.csv'))
sc_dates <- read.csv(paste0(raw_dir, 'sc_dates.csv'))
time_data <- data.table::fread(paste0(raw_dir, 'time_data.csv'))
conc_data <- data.table::fread(paste0(raw_dir, 'conc_data.csv'))
selected <- read_csv(paste0(raw_dir, 'selected.csv'))
```

# Overview

This script processes the standard curve data from the GC on each sample date. It then converts area to methane concentration using the standard curves for every sample. Finally, it computes fluxes (change in concentration over time) as a first-order exponential decay function. These are exported into the `r der_dir` directory.

This dataset includes `r length(unique(substr(conc_data$jar, 1, 1)))` treatments across `r length(unique(conc_data$flux_date))` passages (generations). This includes a total of n = `r nrow(conc_data)` samples. However, passage 6 will not be included in the analyzed dataset, because that passage was sampled much later due to covid and used different potting mix so has too many confounding factors. The samples are broken down across passages and treatments as in the following table:

```{r echo=FALSE}
conc_data %>% 
  mutate(treat = substr(conc_data$jar, 1, 1),
         rep = substr(conc_data$jar, 2, 3)) %>% 
  group_by(flux_date, treat) %>% 
  summarize(n = n_distinct(rep), .groups='drop') %>% 
  kable()
```

# Standard Curves

This script first imports the values for the standard curves. It then calculates the slope of Methane on Area for each sample date.

```{r standard_curves}
sc_model <- function(df) {lm(injection_ppm ~ area, data = df)}
standard_curves <- 
  sc %>% 
  filter(molecule == 'ch4') %>% 
  group_by(flux_date, sc_date) %>% 
  nest() %>% 
  mutate(model = map(data, sc_model),
         glance = map(model, glance),
         tidy = map(model, tidy)) 
sc_r2 <- 
  standard_curves %>% 
  unnest(glance) %>% 
  select(flux_date, sc_date, r.squared)
sc_equation <- 
  standard_curves %>% 
  unnest(tidy) %>%
  select(flux_date, sc_date, term, estimate) %>% 
  spread(term, estimate) %>% 
  rename(intercept = `(Intercept)`, slope = area)
sc_ch4 <- left_join(sc_r2, sc_equation, by = c("flux_date", "sc_date"))

sc_ch4 <- 
  sc_ch4 %>% 
  left_join(sc_dates, by = "sc_date") %>% 
  select(flux_date:sc_date, t, r.squared:slope)
```

# Compute Fluxes

```{r compute_fluxes}

time_data[, c('t0', 't1', 't2', 't3', 't4', 't5')] <- 
  lapply(time_data[, c('t0', 't1', 't2', 't3', 't4', 't5')],
         ymd_hms)

as.days <- function(start, end) {as.numeric(interval(start, end))/60/60/24}

time_data[, c('t1', 't2', 't3', 't4', 't5')] <- 
  lapply(time_data[, c('t1', 't2', 't3', 't4', 't5')],
         function(x) as.days(time_data$t1, x))
time_data <- 
 time_data %>% 
  select(-t0) %>% 
  gather(t, days, t1:t5) %>% 
  mutate(flux_date = factor(flux_date))

conc_data <- 
  conc_data %>% 
  filter(molecule == 'ch4') %>% 
  mutate(flux_date = factor(flux_date)) %>% 
  gather(t, area, t1:t5) %>% 
  left_join(sc_ch4, by = c('flux_date', 't')) %>% 
  mutate(ppm = area * slope + intercept) %>% 
  select(flux_date, jar, t, ppm) %>% 
  spread(t, ppm)

ln_ch0_chn <- function(t0, tn) {
  log(t0/tn)
}

conc_data[, c('t1', 't2', 't3', 't4', 't5')] <- 
  lapply(conc_data[, c('t1', 't2', 't3', 't4', 't5')],
         function(x) ln_ch0_chn(conc_data$t1, x))

flux_data <-
  conc_data %>% 
  gather(t, ppm, t1:t5) %>% 
  left_join(time_data, by = c('flux_date', 'jar', 't'))

flux_data <- 
  flux_data %>% 
  mutate(flux_date = factor(ymd(flux_date)), 
         jar = factor(jar),
         t = factor(t)) %>% 
  mutate(rep = factor(substr(jar, 2, 3)),
         treat = factor(substr(jar, 1, 1)))

#flux_data %>% 
#  ggplot(aes(days, ppm)) + 
#  facet_grid(flux_date + treat ~ rep, scales = "free") +
#  geom_point() + 
#  stat_smooth(method = 'lm', formula = y ~ x)

lin_model <- function(dt) {
  lm(ppm ~ days, data = dt)
}

nested <- 
  flux_data %>% 
  group_by(flux_date, jar, treat, rep) %>% 
  nest() %>% 
  mutate(model = map(data, lin_model))

fluxes <-
  nested %>% 
  mutate(tidy = map(model, tidy)) %>% 
  unnest(tidy) %>% 
  filter(term == 'days') %>% 
  group_by(flux_date) %>% 
  arrange(estimate, .by_group = TRUE) %>% 
  select(-data, -model, -term) %>% 
  ungroup() %>% 
  mutate(passage = as.numeric(flux_date)) %>% 
  mutate(estimate_rank = rank(estimate)) %>% 
  filter(passage != 6) %>% 
  select(-std.error, -statistic, -p.value)

fluxes$estimate_log10 <- log10(fluxes$estimate + abs(min(fluxes$estimate)) + 0.001)

selected$selected <- TRUE

fluxes <- left_join(fluxes, selected, by = c('passage', 'jar'))
fluxes$selected <- !is.na(fluxes$selected)

write_tsv(fluxes, paste0(der_dir, 'fluxes.tsv'))
```

The final dataset includes `r length(unique(fluxes$treat))` treatments across `r length(unique(fluxes$passage))` passages (generations). This includes a total of n = `r nrow(fluxes)` samples. The samples are broken down across passages and treatments as in the following table:

```{r echo=FALSE}
fluxes %>% 
  group_by(flux_date, treat) %>% 
  summarize(n = n_distinct(rep), .groups='drop') %>% 
  kable()
```

### 2021-06-17 R/01-process-flux.Rmd
continue converting R to Rmd

@@
 ---
-title: "Process Flux Date"
-author: "Andrew"
-date: "6/15/2021"
+title: "Process Flux Data"
+author: "Andrew Morris"
+date: "`r Sys.Date()`"
 output: html_document
@@
 
-```{r initialize_packages}
+This script processes the standard curve data from the GC on each sample date.
+It then converts area to methane concentration using the standard curves for every sample.
+Finally, it computes fluxes (change in concentration over time) as a first-order exponential decay function.
+
+```{r initialize_packages, message=FALSE}
 library(tidyverse)
@@
 ```{r initialize_variables}
-raw_dir <- '../Data/'
-der_dir <- '../Output/'
+source('functions.R')
 ```
@@
 conc_data <- data.table::fread(paste0(raw_dir, 'conc_data.csv'))
-selected <- read_csv(paste0(raw_dir, 'selected.csv'))
+selected <- read.csv(paste0(raw_dir, 'selected.csv'))
 ```
-
-# Overview
-
-This script processes the standard curve data from the GC on each sample date.
-It then converts area to methane concentration using the standard curves for every sample.
-Finally, it computes fluxes (change in concentration over time) as a first-order exponential decay function.
-These are exported into the `r der_dir` directory.
 
@@
 
-# Standard Curves
+## Standard Curves
 
-This script first imports the values for the standard curves. It then calculates the slope of Methane on Area for each sample date.
+This step first imports the values for the standard curves then calculates the slope of Methane on Area for each sample date.
 
@@
 sc_model <- function(df) {lm(injection_ppm ~ area, data = df)}
+
+# fit a separate model for each date
 standard_curves <- 
@@
          tidy = map(model, tidy))
+
 sc_r2 <- 
@@
   select(flux_date, sc_date, r.squared)
+
 sc_equation <- 
@@
   rename(intercept = `(Intercept)`, slope = area)
+
 sc_ch4 <- left_join(sc_r2, sc_equation, by = c("flux_date", "sc_date"))
 
+# These are the final slopes for each Methane~Area standard curve 
+# model for each date
 sc_ch4 <- 
@@
 
-# Compute Fluxes
+## Compute Fluxes
 
-```{r compute_fluxes}
+This step turns the time point data to time intervals in units of days.
+
+```{r compute_time_intervals}
 
@@
 
+```
+
+This step takes the raw gas chromatograph area data and calculates methane concentration for each time point in units of ppm.
+
+```{r compute_ch4_conc}
 conc_data <- 
@@
   spread(t, ppm)
+```
 
+Finally, this step takes the time intervals and ch4 concentrations and fits a first-order exponential decay function to compute methane flux as `k` which is the rate-constant for the first-order exponential decay function.
+
+```{r fit_flux}
 ln_ch0_chn <- function(t0, tn) {
@@
 
+# Identify the selected jars in the fluxes dataframe
 selected$selected <- TRUE
-
 fluxes <- left_join(fluxes, selected, by = c('passage', 'jar'))
 fluxes$selected <- !is.na(fluxes$selected)
+```
 
+```{r export_data}
 write_tsv(fluxes, paste0(der_dir, 'fluxes.tsv'))

### 2021-06-17 R/02-process-herit.Rmd
continue converting R to Rmd

---
title: "Process Heritability Data"
author: "Andrew Morris"
date: "`r Sys.Date()`"
output: html_document
---

This script takes in the flux data from script 01 and the data on which jars
were selected in each passage. It then aligns these data so that for each
passage there is a mean flux for all parents, the selected parents, and all
offspring.

```{r initialize_packages, message=FALSE}
library(tidyverse)
```

```{r initialize_variables}
source('functions.R')
```

```{r import_data, message=FALSE}
selected_jars <- read_csv(paste0(raw_dir, 'selected.csv'))
fluxes <- read_tsv(paste0(der_dir, 'fluxes.tsv'))
```

## Organize parent/offspring values

```{r process_herit_data}
# Pull out the selected parental jars
selected <- 
  fluxes[paste0(fluxes$passage, fluxes$jar) %in% paste0(selected_jars$passage, selected_jars$jar), ] %>%
  select(treat, estimate, passage)

# Pull out all parental jars
parental <-
  fluxes %>% 
  select(passage, treat, estimate)

# Pull out all offspring jars
# The `passage - 1` function sets the value of generation `n + 1` 
# as the "offspring" value for generation `n`. 
offspring <- 
  fluxes %>%
  select(treat, estimate, passage) %>% 
  filter(passage != 1) %>% 
  mutate(passage = passage - 1)

# Calculate means within each generation for all parental jars,
# all offspring jars, and all selected parental jars
parental_mean <-
  parental %>% 
  group_by(passage, treat) %>% 
  summarize(parental = mean(estimate), .groups = "drop")

offspring_mean <-
  offspring %>% 
  group_by(passage, treat) %>% 
  summarize(offspring = mean(estimate), .groups = "drop")

selected_mean <-
  selected %>% 
  group_by(passage, treat) %>% 
  summarize(selected = mean(estimate), .groups = "drop")

# Combine all parent/offspring data into one data frame
heritability <- 
  parental_mean %>% 
  left_join(offspring_mean, by = c("passage", "treat")) %>% 
  left_join(selected_mean, by = c("passage", "treat")) %>% 
  #filter(treat == "p") %>% 
  filter(passage != 5) %>% # Remove passage 5, which has no offspring,
  # because it's the last generation
  ungroup()

write_tsv(heritability, paste0(der_dir, 'heritability.tsv'))
```

### 2021-06-17 R/03-response.Rmd
continue converting R to Rmd

---
title: "Estimate Response to Selection"
author: "Andrew Morris"
date: "`r Sys.Date()`"
output: html_document
---

This script intakes the flux data from script 01. It then calculates deviance
as the difference between the positive line and the neutral line in each
passage. The errors can either be calculated as the sum of the errors for
each line or the standard error of difference, that is, the square root of
the sum of the squared errors, which is slightly smaller than the sum. In
this case, I have chosen the SED.
```{r initialize_packages, message=FALSE}
library(tidyverse)
library(metafor)
library(morris)
library(broom)
```

```{r initialize_variables}
source('functions.R')
```

```{r import_data, message=FALSE}
fluxes <- read_tsv(paste0(der_dir, 'fluxes.tsv'))
```

Standard error of the difference function

```{r functions}
# Standard error of the difference
sed <- function(p, n, na.rm = FALSE) {
  sqrt(se(p, na.rm = na.rm) ^ 2 + se(n, na.rm = na.rm)^ 2)
}
```

## Plot the raw data

```{r}
ratio.values <- (max(fluxes$passage)-min(fluxes$passage))/(max(fluxes$estimate)-min(fluxes$estimate))

plot_flux <- function(fluxes, passage, estimate) {
ggplot(fluxes, aes(x = {{passage}}, y = {{estimate}}, color = treat)) +
  theme_classic() +
  geom_jitter(aes(alpha = selected), width = 0.1) +
  stat_smooth(method = 'lm', se = FALSE, formula = 'y ~ x') +
  labs(x = "Passage Number", y = "Methane Oxidation Rate (-k)") +
  scale_color_manual(name = "Treatment", labels = c('Neutral', 'Positive'), values = c('gray40', 'darkorange2')) +
  scale_alpha_manual(name = "Selected", labels = c('Not', 'Selected'), values = c(0.3, 1))
}

plot_log_flux <- function(fluxes, passage, estimate_log10) {
ggplot(fluxes, aes(x = {{passage}}, y = {{estimate_log10}}, color = treat)) +
  theme_classic() +
  geom_jitter(aes(alpha = selected), width = 0.1) +
  stat_smooth(aes(group = treat), method = 'lm', se = FALSE, formula = 'y ~ x') +
  #  coord_fixed(ratio.values) +
  labs(x = "Passage Number", y = expression(log[10] * "(Methane Consumption Rate (-k))")) +
  scale_color_manual(name = "Treatment", labels = c('Neutral', 'Positive'), values = c('gray40', 'darkorange2')) +
  scale_alpha_manual(name = "Selected", labels = c('Not', 'Selected'), values = c(0.3, 1))
}

plot_flux(fluxes, passage, estimate)

#ggsave('fluxes_raw.pdf', width = 5, height = 4)

```

## 3. Plot the log-transformed data

```{r}
# ratio.values <- (max(fluxes$passage)-min(fluxes$passage))/(max(fluxes$estimate_log10)-min(fluxes$estimate_log10))
# 
# selected$selected <- TRUE
# 
# selected_data <- 
# fluxes %>% 
#   left_join(selected, by = c('passage', 'jar')) %>% 
#   filter(selected == TRUE)

plot_log_flux(fluxes, passage, estimate_log10)

ggplot(fluxes, aes(x = passage, y = estimate_log10, color = treat)) +
  theme_classic() +
  geom_jitter(aes(alpha = selected), width = 0.2) +
  stat_smooth(aes(group = treat), method = 'lm', se = FALSE, formula = 'y ~ x') +
#  coord_fixed(ratio.values) +
  labs(x = "Passage Number", y = "log10(Methane Oxidation Rate (-k))") +
  scale_color_manual(name = "Treatment", labels = c('Neutral', 'Positive'), values = c('gray40', 'darkorange2')) +
  scale_alpha_manual(name = "Selected", labels = c('Not', 'Selected'), values = c(0.5, 1))

#ggsave('fluxes_log.pdf', width = 5, height = 4)
```

```{r}
ggplot(fluxes, aes(x = passage, y = estimate_log10, color = treat)) +
  theme_classic() +
  geom_jitter(width = 0.2) +
  stat_smooth(aes(group = treat), method = 'lm', se = FALSE, formula = 'y ~ x') +
  #  coord_fixed(ratio.values) +
  labs(x = "Passage Number", y = expression(log[10] * "(Methane Consumption Rate (-k))")) +
  scale_color_manual(name = "Treatment", labels = c('Neutral', 'Positive'), values = c('gray40', 'darkorange2'))

#ggsave('fluxes_log.png', width = 5, height = 4)
```

# 3. Test a difference of slopes

```{r}
fit <- lm(estimate_log10 ~ 0 + passage * treat, data = fluxes)
summary(fit)
fit
10^fit$coef["passage:treatp"]

fit1 <- lm(estimate_log10 ~ passage + treat, data = fluxes)
summary(fit1)
anova(
      lm(estimate_log10 ~ passage * treat, data = fluxes),
      lm(estimate_log10 ~ passage + treat, data = fluxes)
)
```

4. Plot the back-transformed slope

```{r}
summary(fit)
fit$coef[[3]]
fit$coef[[4]]

ggplot(fluxes, aes(x = passage, y = estimate, color = treat)) +
  theme_classic() +
  geom_jitter() +
  geom_abline(intercept = fit$coef[[3]], slope = fit$coef[[4]]^10) +
  labs(x = "Passage Number", y = "Methane Oxidation Rate (-k)") +
  scale_color_manual(name = "Treatment", labels = c('Neutral', 'Positive'), values = c('gray40', 'darkorange2'))
```

```{r}
####################################################################
# End the correct way to do it
################################################################

# Calculate deviance as p - n
deviance <-
  fluxes %>% 
  select(treat, passage, estimate) %>% 
  group_by(passage, treat) %>% 
  summarize(mean = mean(estimate), .groups = "drop") %>% 
  pivot_wider(names_from = treat, values_from = mean) %>% 
  mutate(deviance = p - n)

# Calculate variance of deviance as SED
deviance$se <-
  fluxes %>% 
  select(treat, passage, estimate) %>% 
  mutate(x = 1:nrow(fluxes)) %>% 
  pivot_wider(names_from = treat, values_from = estimate) %>% 
  group_by(passage) %>% 
  summarize(sed = sed(p, n, na.rm = TRUE), .groups = "drop") %>%
  pull(sed)

# Fit deviance model
fit <- rma(deviance ~ passage, se^2, method="FE", data = deviance)

plot(fit)

# Plot deviance
ggplot(deviance, aes(x = passage, y = deviance, ymin = deviance - se, ymax = deviance + se)) + 
  geom_pointrange(color = 'darkorange2') + 
  geom_abline(intercept = fit$beta[1], slope = fit$beta[2])

# Take deviance model parameters and export into a data.frame
rma_output <- data.frame(
Estimate = fit$beta,
se = fit$se,
z.value = fit$zval,
p.value = fit$pval,
upper.ci = fit$ci.lb,
lower.ci = fit$ci.ub
)

# Fit deviance model without passage 2
fit_no2 <- 
filter(deviance, passage != 2) %>% 
rma(deviance ~ passage, se^2, method="FE", data = .)

# Take deviance model withou passage 2 parameters and export into a data.frame

rma_output_no2 <- data.frame(
Estimate = fit_no2$beta,
se = fit_no2$se,
z.value = fit_no2$zval,
p.value = fit_no2$pval,
upper.ci = fit_no2$ci.lb,
lower.ci = fit_no2$ci.ub
)

# R (response to selection) for the model w/o sample 2
R_no2 <- rma_output_no2[[2, 1]]
R <- rma_output[[2, 1]]

# Plot the no2 model
filter(deviance, passage != 2) %>% 
ggplot(., aes(x = passage, y = deviance, ymin = deviance - se, ymax = deviance + se)) + 
  geom_pointrange(color = 'darkorange2') + 
  geom_abline(intercept = fit_no2$beta[1], slope = fit_no2$beta[2])

#write_tsv(deviance, '../Output/deviance.tsv')
#saveRDS(rma_output, '../Output/response.rds')
#saveRDS(fit, '../Output/dev_fit.rds')

# Calculate deviance as percentage of p - n
deviance <-
  fluxes %>% 
  select(treat, passage, estimate) %>% 
  group_by(passage, treat) %>% 
  summarize(mean = mean(estimate), .groups = "drop") %>% 
  pivot_wider(names_from = treat, values_from = mean) %>% 
  mutate(deviance = (p - n) / n)

# Calculate variance of deviance as SED
deviance$se <-
  fluxes %>% 
  select(treat, passage, estimate) %>% 
  mutate(x = 1:nrow(fluxes)) %>% 
  pivot_wider(names_from = treat, values_from = estimate) %>% 
  group_by(passage) %>% 
  summarize(sed = sed(p, n, na.rm = TRUE), .groups = "drop") %>%
  pull(sed)

# Fit deviance model
fit <- rma(deviance ~ passage, se^2, method="FE", data = deviance)

plot(fit)

# Plot deviance
ggplot(deviance, aes(x = passage, y = deviance, ymin = deviance - se, ymax = deviance + se)) + 
  geom_pointrange(color = 'darkorange2') + 
  geom_abline(intercept = fit$beta[1], slope = fit$beta[2])

# Take deviance model parameters and export into a data.frame
rma_output <- data.frame(
Estimate = fit$beta,
se = fit$se,
z.value = fit$zval,
p.value = fit$pval,
upper.ci = fit$ci.lb,
lower.ci = fit$ci.ub
)

# Fit separte slopes for Positive and Neutral with robust M-estimator regression

library(MASS)
anova(
fit <- rlm(estimate ~ passage * treat, data = fluxes),
fit <- rlm(estimate ~ passage + treat, data = fluxes))

summary(fit)
anova(fit)
new_data <- cbind(fluxes, predict(fit, interval = 'confidence'))
colnames(new_data) 
ggplot(new_data, aes(x = passage, y = estimate, color = treat)) +
  theme_bw() +
  geom_jitter() +
  geom_line(aes(passage, fit)) +
  geom_ribbon(aes(ymin=lwr,ymax=upr), alpha=0.3)

ggplot(new_data, aes(x = passage, y = estimate)) +
  theme_bw() +
  geom_jitter(aes(color = treat)) +
  geom_line(aes(passage, fit, color = treat)) +
  geom_ribbon(aes(ymin=lwr,ymax=upr, group = treat), alpha=0.3) +
  ylim(-0.01, 0.08)
ggplot(fluxes, aes(x = passage, y = estimate, color = treat)) + 
  geom_point() + 
  stat_smooth(method='lm')
#ggsave('raw_fluxes.pdf')
```

### 2021-06-17 R/functions.R

@@ -1,0 +1,2 @@
+raw_dir <- '../Data/' # Raw Data Directory
+der_dir <- '../Output/' # Derived/Modified Data Directory

### 2021-06-18 R/03-response.Rmd
update the response script and move some repeated operations over to functions.R

@@
 library(tidyverse)
-library(metafor)
 library(morris)
 library(broom)
+library(knitr)
 ```
@@
 
-Standard error of the difference function
+## Plot the raw data
 
-```{r functions}
+```{r calc_plot_ratios}
+
+```
+
+```{r plot_raw}
+ratio <- calc_plot_ratios(fluxes$passage, fluxes$estimate)
+
+plot_flux(fluxes, passage, estimate, ratio)
+#ggsave('fluxes_raw.pdf', width = 5, height = 4)
+```
+
+## Plot the log-transformed data
+
+```{r plot_log10}
+ratio_log10 <- calc_plot_ratios(fluxes$passage, fluxes$estimate_log10)
+
+plot_flux(fluxes, passage, estimate_log10, ratio_log10, log10 = TRUE)
+
+#ggsave('fluxes_log.pdf', width = 5, height = 4)
+```
+
+## Test a difference of slopes
+
+```{r response_model}
+# Fit full interaction model
+inter_model <- lm(estimate_log10 ~ 0 + passage * treat, data = fluxes)
+
+# Fit additive model
+addit_model <- lm(estimate_log10 ~ passage + treat, data = fluxes)
+
+# Back calculate the slope of the positive treatment
+pos_slope <- 10^inter_model$coef["passage:treatp"]
+```
+
+Here I fit two models with flux as the dependent variable and passage and treatment as the independent variables.
+The first is the interaction model, which includes an interaction between passage and treatment.
+The second is an additive model that lacks this interaction.
+Here is the model comparison using `anova`, which indicates the full model is somewhat better:
+
+```{r compare_models}
+# Compare interaction to no interaction
+compare_mods <- anova(inter_model, addit_model)
+kable(tidy(compare_mods))
+```
+
+Here is the full interaction model:
+
+```{r inter_model}
+kable(tidy(inter_model))
+```
+
+The term `passage` was not significant, which is the y-intercept of the neutral treatment So This treatment did not change.
+
+## Previous Method
+
+Previously, I was using the `metafor` package to test the slope of the difference between the positive and neutral line.
+
+```{r eval=FALSE}
+
+library(metafor)
 # Standard error of the difference
@@
 }
-```
-
-## Plot the raw data
-
-```{r}
-ratio.values <- (max(fluxes$passage)-min(fluxes$passage))/(max(fluxes$estimate)-min(fluxes$estimate))
-
-plot_flux <- function(fluxes, passage, estimate) {
-ggplot(fluxes, aes(x = {{passage}}, y = {{estimate}}, color = treat)) +
-  theme_classic() +
-  geom_jitter(aes(alpha = selected), width = 0.1) +
-  stat_smooth(method = 'lm', se = FALSE, formula = 'y ~ x') +
-  labs(x = "Passage Number", y = "Methane Oxidation Rate (-k)") +
-  scale_color_manual(name = "Treatment", labels = c('Neutral', 'Positive'), values = c('gray40', 'darkorange2')) +
-  scale_alpha_manual(name = "Selected", labels = c('Not', 'Selected'), values = c(0.3, 1))
-}
-
-plot_log_flux <- function(fluxes, passage, estimate_log10) {
-ggplot(fluxes, aes(x = {{passage}}, y = {{estimate_log10}}, color = treat)) +
-  theme_classic() +
-  geom_jitter(aes(alpha = selected), width = 0.1) +
-  stat_smooth(aes(group = treat), method = 'lm', se = FALSE, formula = 'y ~ x') +
-  #  coord_fixed(ratio.values) +
-  labs(x = "Passage Number", y = expression(log[10] * "(Methane Consumption Rate (-k))")) +
-  scale_color_manual(name = "Treatment", labels = c('Neutral', 'Positive'), values = c('gray40', 'darkorange2')) +
-  scale_alpha_manual(name = "Selected", labels = c('Not', 'Selected'), values = c(0.3, 1))
-}
-
-plot_flux(fluxes, passage, estimate)
-
-#ggsave('fluxes_raw.pdf', width = 5, height = 4)
-
-```
-
-## 3. Plot the log-transformed data
-
-```{r}
-# ratio.values <- (max(fluxes$passage)-min(fluxes$passage))/(max(fluxes$estimate_log10)-min(fluxes$estimate_log10))
-# 
-# selected$selected <- TRUE
-# 
-# selected_data <- 
-# fluxes %>% 
-#   left_join(selected, by = c('passage', 'jar')) %>% 
-#   filter(selected == TRUE)
-
-plot_log_flux(fluxes, passage, estimate_log10)
-
-ggplot(fluxes, aes(x = passage, y = estimate_log10, color = treat)) +
-  theme_classic() +
-  geom_jitter(aes(alpha = selected), width = 0.2) +
-  stat_smooth(aes(group = treat), method = 'lm', se = FALSE, formula = 'y ~ x') +
-#  coord_fixed(ratio.values) +
-  labs(x = "Passage Number", y = "log10(Methane Oxidation Rate (-k))") +
-  scale_color_manual(name = "Treatment", labels = c('Neutral', 'Positive'), values = c('gray40', 'darkorange2')) +
-  scale_alpha_manual(name = "Selected", labels = c('Not', 'Selected'), values = c(0.5, 1))
-
-#ggsave('fluxes_log.pdf', width = 5, height = 4)
-```
-
-```{r}
-ggplot(fluxes, aes(x = passage, y = estimate_log10, color = treat)) +
-  theme_classic() +
-  geom_jitter(width = 0.2) +
-  stat_smooth(aes(group = treat), method = 'lm', se = FALSE, formula = 'y ~ x') +
-  #  coord_fixed(ratio.values) +
-  labs(x = "Passage Number", y = expression(log[10] * "(Methane Consumption Rate (-k))")) +
-  scale_color_manual(name = "Treatment", labels = c('Neutral', 'Positive'), values = c('gray40', 'darkorange2'))
-
-#ggsave('fluxes_log.png', width = 5, height = 4)
-```
-
-# 3. Test a difference of slopes
-
-```{r}
-fit <- lm(estimate_log10 ~ 0 + passage * treat, data = fluxes)
-summary(fit)
-fit
-10^fit$coef["passage:treatp"]
-
-fit1 <- lm(estimate_log10 ~ passage + treat, data = fluxes)
-summary(fit1)
-anova(
-      lm(estimate_log10 ~ passage * treat, data = fluxes),
-      lm(estimate_log10 ~ passage + treat, data = fluxes)
-)
-```
-
-4. Plot the back-transformed slope
-
-```{r}
-summary(fit)
-fit$coef[[3]]
-fit$coef[[4]]
-
-ggplot(fluxes, aes(x = passage, y = estimate, color = treat)) +
-  theme_classic() +
-  geom_jitter() +
-  geom_abline(intercept = fit$coef[[3]], slope = fit$coef[[4]]^10) +
-  labs(x = "Passage Number", y = "Methane Oxidation Rate (-k)") +
-  scale_color_manual(name = "Treatment", labels = c('Neutral', 'Positive'), values = c('gray40', 'darkorange2'))
-```
-
-```{r}
-####################################################################
-# End the correct way to do it
-################################################################
-
 # Calculate deviance as p - n

### 2021-06-18 R/functions.R

@@ -1,2 +1,30 @@
+library(tidyverse)
+
 raw_dir <- '../Data/' # Raw Data Directory
 der_dir <- '../Output/' # Derived/Modified Data Directory
+
+
+plot_flux <- function(fluxes, passage, estimate, ratio = NULL, log10 = FALSE) {
+  p <- ggplot(fluxes, aes(x = {{passage}}, y = {{estimate}}, color = treat)) +
+    theme_classic() +
+    stat_smooth(method = 'lm', se = FALSE, formula = 'y ~ x') +
+    scale_color_manual(name = "Treatment", labels = c('Neutral', 'Positive'), values = c('gray40', 'darkorange2')) +
+    geom_jitter(aes(alpha = selected), width = 0.1) +
+    scale_alpha_manual(name = "Selected", labels = c('Not', 'Selected'), values = c(0.3, 1))
+  if (log10 == FALSE) {
+    p <- p + labs(x = "Passage Number", y = "Methane Oxidation Rate (-k)")
+  } 
+  else if (log10 == TRUE) {
+    p <- p + labs(x = "Passage Number", y = expression(log[10] * "(Methane Consumption Rate (-k))"))
+  } 
+  
+  if (!is.null(ratio)) {
+    p <- p + coord_fixed({{ratio}})
+  }
+  
+  return(p)
+}
+
+calc_plot_ratios <- function(x, y) {
+  (max(x)-min(x))/(max(y)-min(y))
+}

### 2021-06-24 R/03-response.Rmd
convert response to selection script to rmd

@@
 
-This script intakes the flux data from script 01. It then calculates deviance
-as the difference between the positive line and the neutral line in each
-passage. The errors can either be calculated as the sum of the errors for
-each line or the standard error of difference, that is, the square root of
-the sum of the squared errors, which is slightly smaller than the sum. In
-this case, I have chosen the SED.
+This script intakes the flux data from script 01.
+It then fits a linear model to the flux data over time per treatment and tests a difference of slope between the positive and the neutral treatment.
+Finally, it
 ```{r initialize_packages, message=FALSE}
@@
 library(knitr)
+#library(broman)
 ```
@@
 fluxes <- read_tsv(paste0(der_dir, 'fluxes.tsv'))
+
+fluxes <-
+  fluxes %>% 
+  mutate(passage = passage - 1)
 ```
@@
 
-```{r calc_plot_ratios}
+```{r plot_raw}
+ratio <- calc_plot_ratios(fluxes$passage, fluxes$ch4)
 
-```
-
-```{r plot_raw}
-ratio <- calc_plot_ratios(fluxes$passage, fluxes$estimate)
-
-plot_flux(fluxes, passage, estimate, ratio)
+plot_flux(fluxes, passage, ch4, ratio)
 #ggsave('fluxes_raw.pdf', width = 5, height = 4)
@@
 ```{r plot_log10}
-ratio_log10 <- calc_plot_ratios(fluxes$passage, fluxes$estimate_log10)
+ratio_log10 <- calc_plot_ratios(fluxes$passage, fluxes$log_ch4)
 
-plot_flux(fluxes, passage, estimate_log10, ratio_log10, log10 = TRUE)
+plot_flux(fluxes, passage, log_ch4, ratio_log10, log10 = TRUE)
 
@@
 # Fit full interaction model
-inter_model <- lm(estimate_log10 ~ 0 + passage * treat, data = fluxes)
+inter_model <- lm(log_ch4 ~ passage * treat, data = fluxes)
 
-# Fit additive model
-addit_model <- lm(estimate_log10 ~ passage + treat, data = fluxes)
+# No treatment model
+no_treat_model <- lm(log_ch4 ~ passage, data = fluxes)
 
@@
 pos_slope <- 10^inter_model$coef["passage:treatp"]
+
 ```
@@
 # Compare interaction to no interaction
-compare_mods <- anova(inter_model, addit_model)
+compare_mods <- anova(inter_model, no_treat_model)
 kable(tidy(compare_mods))
@@
 kable(tidy(inter_model))
+
 ```
 
-The term `passage` was not significant, which is the y-intercept of the neutral treatment So This treatment did not change.
+The intercept for the Neutral treatment was significantly different from zero at `r round(tidy(inter_model)[[1, 2]], 2)` (SE = `r round(tidy(inter_model)[[1, 3]], 2)`, t = `r round(tidy(inter_model)[[1, 4]], 2)`, p = `r round(tidy(inter_model)[[1, 5]], 2)`).
+The Positive treatment started at a significantly lower methane oxidation rate than the Neutral treatment with a y-intercept of `r round(tidy(inter_model)[[3, 2]], 2)` relative to Neutral (SE = `r round(tidy(inter_model)[[3, 3]], 2)`, t = `r round(tidy(inter_model)[[3, 4]], 2)`, p = `r round(tidy(inter_model)[[3, 5]], 2)`).
+There was no change in methane oxidation rate of the Neutral treatment over the five passages (slope = `r round(tidy(inter_model)[[2, 2]], 2)`, SE = `r round(tidy(inter_model)[[2, 3]], 2)`, t = `r round(tidy(inter_model)[[2, 4]], 2)`, p = `r round(tidy(inter_model)[[2, 5]], 2)`).
+By contrast, there was a significant increase in the Positive treatment of `r round(tidy(inter_model)[[4, 2]], 2)` per passage which is approximately a `r round((pos_slope - 1) * 100, 0)`% increase in methane oxidation rate per passage (SE = `r round(tidy(inter_model)[[4, 3]], 2)`, t = `r round(tidy(inter_model)[[4, 4]], 2)`, p = `r round(tidy(inter_model)[[4, 5]], 2)`).
 
@@
 Previously, I was using the `metafor` package to test the slope of the difference between the positive and neutral line.
+For this model, I calculate the difference between the Positive and Neutral treatment and fit that difference as the response.
+I also calculate the standard error of the difference and included that variance in the model.
+I then fit the model `difference ~ passage` to test the slope of the difference over time.
 

### 2021-06-24 R/01-process-flux.Rmd
update processing rmd, remove old r scripts, update makefile with new rmd

@@
   mutate(passage = as.numeric(flux_date)) %>% 
-  mutate(estimate_rank = rank(estimate)) %>% 
+  mutate(ch4 = estimate,
+         rank_ch4 = rank(ch4),
+         log_ch4 = log10(ch4 + abs(min(ch4)) + 0.001)) %>% 
   filter(passage != 6) %>% 
-  select(-std.error, -statistic, -p.value)
+  select(-std.error, -statistic, -p.value, -estimate)
+```
 
-fluxes$estimate_log10 <- log10(fluxes$estimate + abs(min(fluxes$estimate)) + 0.001)
+```{r identify_selected}
 
@@
 selected$selected <- TRUE
-fluxes <- left_join(fluxes, selected, by = c('passage', 'jar'))
-fluxes$selected <- !is.na(fluxes$selected)
+fluxes <- 
+  fluxes %>% 
+  left_join(selected, by = c('passage', 'jar')) %>% 
+  mutate(selected = !is.na(selected)) %>% 
+  select(-jar) %>% 
+  select(flux_date, passage, treat, rep, ch4:selected)
+```
+
+```{r print_data}
+kable(fluxes)
 ```

### 2021-06-24 R/01-process-flux.R

@@ -1,141 +1,0 @@
-# This scripts imports the raw methane oxidation rate data as manually entered.
-# This includes the date of each standard curve and the measures Areas for each
-# Concentration of the standard curve. It then calculates the slope of the
-# relationship Methane ~ Area for each date a standard curve was taken.
-#
-# Next, this script imports the time point data and concentration (Area) data
-# for each jar in each passage. It then computes an exponential decay function
-#
-# and extracts the exponential decay rate (k).
-# Input: standard_curve.csv, sc_dates.csv, time_data.csv, conc_data.csv, selected.csv
-# Output: fluxes.tsv
-
-library(tidyverse)
-library(broom)
-library(lubridate)
-
-# Standard Curves
-
-sc <- read.csv('../Data/standard_curve.csv')
-sc_dates <- read.csv('../Data/sc_dates.csv')
-
-sc_model <- function(df) {lm(injection_ppm ~ area, data = df)}
-standard_curves <- 
-  sc %>% 
-  filter(molecule == 'ch4') %>% 
-  group_by(flux_date, sc_date) %>% 
-  nest() %>% 
-  mutate(model = map(data, sc_model),
-         glance = map(model, glance),
-         tidy = map(model, tidy)) 
-sc_r2 <- 
-  standard_curves %>% 
-  unnest(glance) %>% 
-  select(flux_date, sc_date, r.squared)
-sc_equation <- 
-  standard_curves %>% 
-  unnest(tidy) %>%
-  select(flux_date, sc_date, term, estimate) %>% 
-  spread(term, estimate) %>% 
-  rename(intercept = `(Intercept)`, slope = area)
-sc_ch4 <- left_join(sc_r2, sc_equation, by = c("flux_date", "sc_date"))
-
-sc_ch4 <- 
-  sc_ch4 %>% 
-  left_join(sc_dates, by = "sc_date") %>% 
-  select(flux_date:sc_date, t, r.squared:slope)
-
-
-# Fluxes
-
-time_data <- data.table::fread('../Data/time_data.csv')
-
-time_data[, c('t0', 't1', 't2', 't3', 't4', 't5')] <- 
-  lapply(time_data[, c('t0', 't1', 't2', 't3', 't4', 't5')],
[…]

### 2021-06-24 R/02-process-herit.R

@@ -1,64 +1,0 @@
----
-title: "Process Heritability Data"
-author: "Andrew Morris"
-date: "6/17/2021"
-output: html_document
----
-
-# This script takes in the flux data from script 01 and the data on which jars
-# were selected in each passage. It then aligns these data so that for each
-# passage there is a mean flux for all parents, the selected parents, and all
-# offspring. 
-
-# input: fluxes.tsv, selected.csv
-# output: heritability.tsv
-
-library(tidyverse)
-
-selected_jars <- read_csv('../Data/selected.csv')
-fluxes <- read_tsv('../Output/fluxes.tsv')
-
-# Pull out the selected parental jars
-selected <- 
-  fluxes[paste0(fluxes$passage, fluxes$jar) %in% paste0(selected_jars$passage, selected_jars$jar), ] %>%
-  select(treat, estimate, passage)
-
-# Pull out all parental jars
-parental <-
-  fluxes %>% 
-  select(passage, treat, estimate)
-
-# Pull out all offspring jars
-offspring <- 
-  fluxes %>%
-  select(treat, estimate, passage) %>% 
-  filter(passage != 1) %>% 
-  mutate(passage = passage - 1)
-
-# Calculate means for east generation
-parental_mean <-
-  parental %>% 
-  group_by(passage, treat) %>% 
-  summarize(parental = mean(estimate), .groups = "drop")
-
-offspring_mean <-
-  offspring %>% 
-  group_by(passage, treat) %>% 
-  summarize(offspring = mean(estimate), .groups = "drop")
-
-selected_mean <-
-  selected %>% 
-  group_by(passage, treat) %>% 
-  summarize(selected = mean(estimate), .groups = "drop")
-
-# Combine all parent/offspring data into one data frame
-heritability <- 
-  parental_mean %>% 
-  left_join(offspring_mean, by = c("passage", "treat")) %>% 
-  left_join(selected_mean, by = c("passage", "treat")) %>% 
-  #filter(treat == "p") %>% 
-  filter(passage != 5) %>% # Remove passage 5, which has no offspring
-  ungroup()
-
-
-write_tsv(heritability, '../Output/heritability.tsv')

### 2021-06-24 R/03-response.R

@@ -1,245 +1,0 @@
-# This script intakes the flux data from script 01. It then calculates deviance
-# as the difference between the positive line and the neutral line in each
-# passage. The errors can either be calculated as the sum of the errors for
-# each line or the standard error of difference, that is, the square root of
-# the sum of the squared errors, which is slightly smaller than the sum. In
-# this case, I have chosen the SED.
-
-# input: fluxes.tsv
-# output: deviance.tsv, response.rds
-
-library(tidyverse)
-library(metafor)
-library(morris)
-library(broom)
-
-# Standard error of the difference function
-sed <- function(p, n, na.rm = FALSE) {
-  sqrt(se(p, na.rm = na.rm) ^ 2 + se(n, na.rm = na.rm)^ 2)
-}
-
-
-# Response to Selection
-
-fluxes <- read_tsv('Output/fluxes.tsv')
-
-#####################################################################
-# Okay here to deviance is how I should really be doing it
-###############################################################
-
-# 1. Plot the raw data
-ratio.values <- (max(fluxes$passage)-min(fluxes$passage))/(max(fluxes$estimate)-min(fluxes$estimate))
-
-ggplot(fluxes, aes(x = passage, y = estimate, color = treat)) +
-  theme_classic() +
-  geom_jitter(width = 0.1) +
-  stat_smooth(method = 'lm', se = FALSE, formula = 'y ~ x') +
-#  coord_fixed(ratio.values) +
-  labs(x = "Passage Number", y = "Methane Oxidation Rate (-k)") +
-  scale_color_manual(name = "Treatment", labels = c('Neutral', 'Positive'), values = c('gray40', 'darkorange2'))
-
-ggsave('fluxes_raw.pdf', width = 5, height = 4)
-
-# 2. Transform it to make it more normal
-
-# 3. Plot the transformed data
-
-# ratio.values <- (max(fluxes$passage)-min(fluxes$passage))/(max(fluxes$estimate_log10)-min(fluxes$estimate_log10))
-# 
-# selected$selected <- TRUE
-# 
-# selected_data <- 
-# fluxes %>% 
-#   left_join(selected, by = c('passage', 'jar')) %>% 
[…]

### 2021-06-25 R/05_price.R
remove old script

@@ -1,4 +1,0 @@
-library(tidyverse)
-
-fluxes <- read_tsv('../Output/fluxes.tsv')
-

### 2021-06-25 R/02-process-herit.Rmd
convert heritability script to Rmd

@@
 library(tidyverse)
+library(knitr)
 ```
@@
 selected <- 
-  fluxes[paste0(fluxes$passage, fluxes$jar) %in% paste0(selected_jars$passage, selected_jars$jar), ] %>%
-  select(treat, estimate, passage)
+  fluxes %>% 
+  filter(selected) %>% 
+  select(passage, treat, log_ch4)
 
@@
   fluxes %>% 
-  select(passage, treat, estimate)
+  select(passage, treat, log_ch4)
 
@@
   fluxes %>%
-  select(treat, estimate, passage) %>% 
+  select(treat, log_ch4, passage) %>% 
   filter(passage != 1) %>% 
@@
   group_by(passage, treat) %>% 
-  summarize(parental = mean(estimate), .groups = "drop")
+  summarize(parental = mean(log_ch4), .groups = "drop")
 
@@
   group_by(passage, treat) %>% 
-  summarize(offspring = mean(estimate), .groups = "drop")
+  summarize(offspring = mean(log_ch4), .groups = "drop")
 
@@
   group_by(passage, treat) %>% 
-  summarize(selected = mean(estimate), .groups = "drop")
+  summarize(selected = mean(log_ch4), .groups = "drop")
 
@@
   ungroup()
+```
 
+```{r print_data}
+kable(heritability)
+```
+
+```{r export_data}
 write_tsv(heritability, paste0(der_dir, 'heritability.tsv'))

### 2021-06-25 R/04-heritability.Rmd
convert heritability script to Rmd

---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

This script imports the heritability data. This dataset includes the mean flux withn treatment*generation for 1) all parents, 2) selected parents, and 3) offspring. Then, it plots offspring on the selection parents with the two treatments combined and then separately for the two treatments.

```{r initialize_packages}
library(tidyverse)
library(morris)
library(broom)
library(knitr)
library(broman)
```

```{r initialize_variables}
source('functions.R')
```

```{r import_data}
heritability <- read_tsv(paste0(der_dir, 'heritability.tsv'))

```
# Test heritability as mid-offspring on mid-parent

```{r test_treatment}
fit_no_treat <- lm(offspring ~ selected, data = heritability)
kable(tidy(fit_no_treat))
fit_treat <- lm(offspring ~ selected * treat, data = heritability)
kable(tidy(fit_treat))
kable(tidy(anova(fit_no_treat, fit_treat)))
```

Here, I fit two models. The first is simply offspring flux on selected parent flux, ignoring treatment. There is no correlation `r print_model(fit_no_treat, 2)`. Next, I added treatment and the interaction between treatment and parental flux. Here, there is a significant correlation between parent and offspring for both the Neutral treatment `r print_model(fit_treat, 2)` and the Positive treatment `r print_model(fit_treat, 4)`. Notably, the sign of the effect is reversed: negative for the Neutral treatment and positive for the Positive treatment.

The negative heritability of methane oxidation rate in the Neutral selection treatment may result from selection on alternative characters in those populations. The strong, positive heritability in the Positive treatment ... . It seems strange that the

```{r plot_herit}

plot_herit(heritability)

```

######################################################################
# End of the way I should be doing it
################################################################

The Old way:

It then calculates S, the selection differential, as the mean difference between the selected parents and all parents. Because passage 2 was accidentaly selected for the lowest methane oxidation rate, I take the absolute value of this difference, i.e. |S|. Next, it imports the deviance data and calculates R, the respones to selection, as the slope of mean difference in flux (positive - neutral) over time.

An alternative approach would be to calculate R directly instead of approximating it using the slope as in the equation offspring - parent. In this case, I would not be using the deviance, but instead the raw fluxes. I can do that with the heritability dataset.

# Function for the Breeder's equation

```{r eval=FALSE, include=FALSE}
breeders <- function(R, S) {
  R / S
}

#

selected <- 
  fluxes[paste0(fluxes$passage, fluxes$jar) %in% paste0(selected_jars$passage, selected_jars$jar), ] %>%
  select(treat, log_ch4, passage) %>% 
  group_by(treat, passage) %>% 
  summarize(selected = mean(log_ch4), .groups = 'drop')

parents <- 
  fluxes %>%
  select(treat, log_ch4, passage) %>% 
  group_by(treat, passage) %>% 
  summarize(parental = mean(log_ch4), .groups = 'drop')

offspring <- 
  fluxes %>% 
  select(passage, treat, log_ch4) %>% 
  filter(passage != 1) %>% 
  mutate(passage = passage - 1) %>% 
  group_by(treat, passage) %>% 
  summarize(offspring = mean(log_ch4), .groups = 'drop')

heritability <- 
  offspring %>% 
  left_join(selected, by = c('passage', 'treat')) %>% 
  left_join(parents, by = c('passage', 'treat'))

# Matt's approach: h2 for each generation
h2_per_gen  <- 
heritability %>% 
  mutate(R = offspring - parental) %>% 
  mutate(S = selected - parental) %>% 
  mutate(h2 = breeders(R, S))

write_tsv(h2_per_gen, 'Output/h2_per_gen.tsv')

ggplot(h2_per_gen, aes(x = passage, color = treat, y = h2)) +
  geom_point()

h2_per_gen %>% 
  select(treat, passage, h2)

h2_per_gen %>% 
  select(treat, passage, h2) %>% 
  group_by(treat) %>% 
  summarize(mean_h2 = mean(h2), sd_h2 = sd(h2), .groups = 'drop')

# Calculate R using regression of deviance over time.
# Calculate S using math.

dev_fit <- readRDS('../Output/dev_fit.rds')
R <- dev_fit$beta[2]

S <-
  heritability %>%
  filter(treat == "p") %>%
  mutate(S = abs(selected - parental)) %>%
  summarize(S = mean(S))  %>% 
  pull()

h2_reg_dev <- breeders(R, S)

lm_fit <- 
  fluxes %>% 
  filter(treat == 'p') %>% 
  select(treat, passage, ch4) %>% 
  lm(ch4 ~ passage, data = .)

R_not_dev <- pull(tidy(lm_fit)[2, 2])
h2_reg_abs <- breeders(R_not_dev, S)

# Calculate R as regression of offspring on parents

ggplot(heritability, aes(x = parental, y = offspring)) +
  geom_point() + 
  stat_smooth(method = 'lm', se = FALSE)
summary(lm(offspring ~ parental, data = heritability, subset = treat == 'p'))
# Calculate h2 (heritability) using math, not regression
# For each generation and then calculate the mean/se
h2_calc_method <-
  heritability %>%
  filter(treat == "p") %>%
  mutate(S = abs(selected - parental)) %>%
  mutate(R = offspring - parental) %>%
  mutate(h2 = breeders(R, S)) %>%
  summarize(S = mean(S),
            R = mean(R),
            se = se(h2),
            h2 = mean(h2))

h2_calc_method_no_2 <-
  heritability %>%
  filter(treat == "p") %>%  
  filter(passage != 2) %>%
  mutate(S = abs(selected - parental)) %>%
  mutate(R = offspring - parental) %>%
  mutate(h2 = breeders(R, S)) %>%
  summarize(S = mean(S),
            R = mean(R),
            se = se(h2),
            h2 = mean(h2))

h2 <- 
  h2_calc_method %>% 
  mutate(method = "calc p")

h2_no2 <- 
  h2_calc_method_no_2 %>% 
  mutate(method = "calc p no 2")

h2 <- rbind(h2, h2_no2)

# Cumulative response to selection

parent <- heritability[[1, 3]]
offspr <- heritability[[8, 4]]
select <- heritability[[8, 5]]
R = offspr - parent
S = select - parent
h2 = breeders(R, S)

heritability[,'cum_res'] <- NA
heritability$cum_sel <- NA
herit_p <- filter(heritability, treat == 'p')
for (i in seq(nrow(herit_p))) {
  off <- sum(herit_p$offspring[1:i])
  par <- sum(herit_p$parental[1:i])
  sel <- sum(herit_p$selected[1:i])
  herit_p$cum_res[i] <- off - par
  herit_p$cum_sel[i] <- sel - par
}
herit_p$cum_h2 <- herit_p$cum_res / herit_p$cum_sel

ggplot(herit_p, aes(x = cum_sel, y = cum_res)) + 
  theme_classic() +
  labs(x = 'Cumulative selection differential',
       y = 'Cumulative response to selection') +
  geom_point() + 
  stat_smooth(method = 'lm', se = FALSE)
summary(lm(cum_res ~ cum_sel, data = herit_p))

# Regression of offspring on mid-parent

selected_jars <- read_csv('../Data/selected.csv')
fluxes <- read_tsv('../Output/fluxes.tsv')

# Pull out the selected parental jars
selected <- 
  fluxes[paste0(fluxes$passage, fluxes$jar) %in% paste0(selected_jars$passage, selected_jars$jar), ] %>%
  select(treat, ch4, passage) %>% 
  group_by(treat, passage) %>% 
  summarize(parental = mean(ch4), .groups = 'drop')

offspring <- 
  fluxes %>% 
  select(passage, treat, ch4) %>% 
  filter(passage != 1) %>% 
  mutate(passage = passage - 1) %>% 
  rename(offspring = ch4)

heritability <- 
  offspring %>% 
  left_join(selected, by = c('passage', 'treat'))

heritability

ggplot(heritability, aes(parental, offspring)) +
  theme_classic() +
  geom_point() + 
  stat_smooth(method = 'lm', se = FALSE)

summary(lm(offspring ~ parental, data = heritability))

ggplot(heritability, aes(parental, offspring, color = treat)) +
  theme_classic() +
  geom_point() + 
  stat_smooth(method = 'lm', se = FALSE)
summary(lm(offspring ~ parental * treat, data = heritability))
```

### 2021-06-25 R/functions.R

@@ -1,4 +1,5 @@
 library(tidyverse)
+library(broman)
 
 raw_dir <- '../Data/' # Raw Data Directory
 der_dir <- '../Output/' # Derived/Modified Data Directory
@@ -19,7 +20,7 @@ plot_flux <- function(fluxes, passage, estimate, ratio = NULL, log10 = FALSE) {
   } 
   
   if (!is.null(ratio)) {
-    p <- p + coord_fixed({{ratio}})
+    p <- p + coord_fixed(ratio)
   }
   
   return(p)
@@ -27,4 +28,27 @@ plot_flux <- function(fluxes, passage, estimate, ratio = NULL, log10 = FALSE) {
 
 calc_plot_ratios <- function(x, y) {
   (max(x)-min(x))/(max(y)-min(y))
+}
+
+
+print_model <- function(fit, term = 1) {
+  paste0('(slope = ', myround(tidy(fit)[term, 2], 2), 
+        ', SE = ', myround(tidy(fit)[term, 3], 2), 
+        ', t = ', myround(tidy(fit)[term, 4], 2), 
+        ', p = ', myround(tidy(fit)[term, 5], 2), ')')
+}
+
+
+plot_herit <- function(heritability, ratio) {
+  ratio <- calc_plot_ratios(heritability$selected, heritability$offspring)
+  ggplot(heritability, aes(selected, offspring, color = treat)) +
+    theme_classic() +
+    geom_jitter(width = 0.01) + 
+    stat_smooth(method = 'lm', se = FALSE, formula = 'y ~ x') +
+    labs(x = "Mid-parent (mean(log10(Flux (-k))))", 
+         y = "Offspring (log10(Flux (-k)))") +
+    scale_color_manual(name = "Treatment", 
+                       labels = c('Neutral', 'Positive'), 
+                       values = c('gray40', 'darkorange2')) +
+    coord_fixed(ratio)
 }

### 2021-06-29 Manuscript/manuscript.Rmd
update manuscript with new analyses and update makefile

--- 2020-10-23 Manuscript/manuscript.Rmd
@@
 knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
+```
+
+```{r initialize_packages}
 library(tidyverse)
@@
 library(morris)
-library(metafor)
+#library(metafor)
+```
+```{r load_data}
+heritability <- read_tsv('../Output/heritability.tsv')
+fluxes <- read_tsv('../Output/fluxes.tsv')
+
+load(file = '../Output/response_model.Rdata')
+load(file = '../Output/herit_model.Rdata')
+
+source('../R/functions.R')
+
+```
+
+```{r figure_captions, include=FALSE}
+
 figures <- captioner()
-figures("deviance", "Deviance of CH~4~ oxidation rate for the Positive selection treatment relative to the Neutral selection treatment.
+figures("response", "Deviance of CH~4~ oxidation rate for the Positive selection treatment relative to the Neutral selection treatment.
 Deviance is calculated as the mean of P within each passage minus the mean of N within each passage.
@@
 
-```{r dev_model}
-fluxes <- read_tsv('../Output/fluxes.tsv')
-deviance <- read_tsv('../Output/deviance.tsv')
-rma_output <- readRDS('../Output/response.rds')
-dev_fit <- readRDS('../Output/dev_fit.rds')
-dev_pred <- predict(dev_fit)
+The intercept for the Neutral treatment was significantly different from zero
+at `r myround(tidy(response_model)[[1, 2]], 2)` (SE = `r
+myround(tidy(response_model)[[1, 3]], 2)`, t = `r myround(tidy(response_model)[[1,
+4]], 2)`, p = `r myround(tidy(response_model)[[1, 5]], 2)`). The Positive
+treatment started at a significantly lower methane oxidation rate than the
+Neutral treatment with a y-intercept of `r myround(tidy(response_model)[[3, 2]],
+2)` relative to Neutral (SE = `r myround(tidy(response_model)[[3, 3]], 2)`, t = `r
+myround(tidy(response_model)[[3, 4]], 2)`, p = `r myround(tidy(response_model)[[3,
+5]], 2)`). There was no change in methane oxidation rate of the Neutral
+treatment over the five passages (slope = `r myround(tidy(response_model)[[2, 2]],
+2)`, SE = `r myround(tidy(response_model)[[2, 3]], 2)`, t = `r
+myround(tidy(response_model)[[2, 4]], 2)`, p = `r myround(tidy(response_model)[[2,
+5]], 2)`). By contrast, there was a significant increase in the Positive
+treatment of `r myround(tidy(response_model)[[4, 2]], 2)` per passage which is
+approximately a `r round((pos_slope - 1) * 100, 0)`% increase in methane
+oxidation rate per passage (SE = `r myround(tidy(response_model)[[4, 3]], 2)`, t =
+`r myround(tidy(response_model)[[4, 4]], 2)`, p = `r myround(tidy(response_model)[[4,
+5]], 2)`).
 
-flux_per_psg <-
-  fluxes %>% 
-  group_by(passage) %>% 
-  summarize(mean = mean(estimate), se = morris::se(estimate))
+Methane oxidation rate increased as a response to selection in the Positive
+selection treatment (`r figures("deviance", display="cite")`).  This can be
+seen in the increase in deviation between the Positive and Neutral selection
+lines over the course of the experiment (z = , p = ).  Mean CH~4~ oxidation
+rate for the Positive treatment increased by  $\pm$ per passage relative to the
+Neutral treatment. The mean flux across all jars in all treatments in the first
+passage was  $\pm$  so this represents a doubling of CH~4~ consumption in every passage due
+to selection.
 
-passages <- length(deviance$passage)
+```{r response_figure, fig.width=4, fig.height=3, fig.cap=figures("response")}
+ratio_log10 <- calc_plot_ratios(fluxes$passage, fluxes$log_ch4)
 
-```
-
-Methane oxidation rate increased as a response to selection in the Positive selection treatment (`r figures("deviance", display="cite")`). 
-This can be seen in the increase in deviation between the Positive and Neutral selection lines over the course of the experiment (z = `r myround(dev_fit$zval[2], 2)`, p = `r myround(dev_fit$pval[2], 3)`).
-Mean CH~4~ oxidation rate for the Positive treatment increased by `r myround(dev_fit$beta[2], 3)` $\pm$ `r myround(dev_fit$se[2], 3)` per passage relative to the Neutral treatment.
-The mean flux across all jars in all treatments in the first passage was `r myround(flux_per_psg[1, 2], 3)` $\pm$ `r myround(flux_per_psg[1, 3], 3)` so this represents a doubling of CH~4~ consumption in every passage due to selection.
-
-```{r dev_fig, fig.width=4, fig.height=3, fig.cap=figures("deviance")}
-ggplot(deviance, aes(x = passage, y = deviance, ymin = deviance - se, ymax = deviance + se)) + 
-  geom_ribbon(aes(ymin = dev_pred$ci.lb, ymax = dev_pred$ci.ub), alpha = 0.2) +
-  geom_pointrange(color = 'darkorange2') + 
-      geom_segment(aes(x = 1, xend = passages, y = dev_fit$beta[1] + dev_fit$beta[2], yend = dev_fit$beta[1] + dev_fit$beta[2]*passages), color = 'darkorange2', size = 1) +
-  labs(x = "Passage Number", y = "Deviation from Control (P - N)")+
-  theme_bw()
+plot_flux(fluxes, passage, log_ch4, ratio_log10, log10 = TRUE)
 
@@
 
-```{r herit_model}
-heritability <- read_tsv('../Output/heritability.tsv')
+Here, I fit two models. The first is simply offspring flux on selected parent
+flux, ignoring treatment. There is no correlation. Next, I added treatment and the interaction between treatment and parental
+flux. Here, there is a significant correlation between parent and offspring for
+both the Neutral treatment `r print_model(herit_model, 2)` and the Positive
+treatment `r print_model(herit_model, 4)`. Notably, the sign of the effect is
+reversed: negative for the Neutral treatment and positive for the Positive
+treatment.
 
-her_fit <- lm(offspring ~ parental, data = heritability)
-her_sum <- tidy(her_fit)
-her_aov <- tidy(anova(her_fit))
-
-```
-
-```{r inbreeding_coef}
-N = 12 # population size
-T = 5  # number of passages (generations)
-F = 1 - (1 - 1 / (2 * N)) ^ T
-```
-
-Methane oxidation rate was heritable between consecutive passages (`r figures("herit", display="cite")`).
-Specifically, the narrow-sense heritability, calculated as the slope of the mid-parent and mid-offspring, was $h^2$ = `r myround(her_sum[2, 2], 2)` $\pm$ `r myround(her_sum[2, 3], 2)` (t = `r myround(her_sum[2, 4], 2)`, p = `r myround(her_sum[2, 5], 3)`).
-With an inbreeding coefficient of `r myround(F, 2)`,
+The negative heritability of methane oxidation rate in the Neutral selection
+treatment may result from selection on alternative characters in those
+populations. The strong, positive heritability in the Positive treatment ... .
+It seems strange that the Methane oxidation rate was heritable between
+consecutive passages (`r figures("herit", display="cite")`). Specifically, the
+narrow-sense heritability, calculated as the slope of the mid-parent and
+mid-offspring, was $h^2$ =  $\pm$  (t = , p = ). With an inbreeding coefficient of ,
 
 ```{r herit_fig, fig.width=5, fig.height=3, fig.cap=figures("herit")}
-ggplot(heritability, aes(x = parental, y = offspring)) +
-  geom_point(aes(color = treat)) +
-  stat_smooth(method = 'lm', formula = y ~ x, color = "black") + 
-  labs(x = "Parental Mean Flux (-k)", y = "Offspring Mean Flux (-k)") +
-    scale_color_manual(name = "Selection\nTreatment", labels = c('Neutral', 'Positive'), values = c("gray60", "darkorange2")) +
-  theme_bw()
-
-```
-
-```{r}
-R = dev_fit$beta[2]
-h2 = pull(her_sum[2, 2])
-S = R/h2
+plot_herit(heritability)
 ```

### 2021-06-29 R/03-response.Rmd
update manuscript with new analyses and update makefile

@@
 
-This script intakes the flux data from script 01.
-It then fits a linear model to the flux data over time per treatment and tests a difference of slope between the positive and the neutral treatment.
-Finally, it
+This script takes in the flux data from script 01. It then plots the raw data
+and the log-transformed data. Finally, it fits a linear model to the flux data
+over time per treatment and tests a difference of slope between the positive
+and the neutral treatment.
+
 ```{r initialize_packages, message=FALSE}
@@
 library(knitr)
-#library(broman)
+library(broman)
 ```
@@
 
+# Shift the passage numbers so the first passage becomes "Passage 0"
 fluxes <-
@@
 # Fit full interaction model
-inter_model <- lm(log_ch4 ~ passage * treat, data = fluxes)
+response_model <- lm(log_ch4 ~ passage * treat, data = fluxes)
 
@@
 # Back calculate the slope of the positive treatment
-pos_slope <- 10^inter_model$coef["passage:treatp"]
+pos_slope <- 10^response_model$coef["passage:treatp"]
 
@@
 
-Here I fit two models with flux as the dependent variable and passage and treatment as the independent variables.
-The first is the interaction model, which includes an interaction between passage and treatment.
-The second is an additive model that lacks this interaction.
-Here is the model comparison using `anova`, which indicates the full model is somewhat better:
+Here I fit two models.
+The first is the full model which includes passage number, treatment, and the interaction between the two.
+The second model includes only passage with no effect of treatment.
+Comparing the models using `anova`
 
@@
 # Compare interaction to no interaction
-compare_mods <- anova(inter_model, no_treat_model)
+compare_mods <- anova(response_model, no_treat_model)
 kable(tidy(compare_mods))
@@
 
-```{r inter_model}
-kable(tidy(inter_model))
-
+```{r print_model}
+kable(tidy(response_model))
 ```
 
-The intercept for the Neutral treatment was significantly different from zero at `r round(tidy(inter_model)[[1, 2]], 2)` (SE = `r round(tidy(inter_model)[[1, 3]], 2)`, t = `r round(tidy(inter_model)[[1, 4]], 2)`, p = `r round(tidy(inter_model)[[1, 5]], 2)`).
-The Positive treatment started at a significantly lower methane oxidation rate than the Neutral treatment with a y-intercept of `r round(tidy(inter_model)[[3, 2]], 2)` relative to Neutral (SE = `r round(tidy(inter_model)[[3, 3]], 2)`, t = `r round(tidy(inter_model)[[3, 4]], 2)`, p = `r round(tidy(inter_model)[[3, 5]], 2)`).
-There was no change in methane oxidation rate of the Neutral treatment over the five passages (slope = `r round(tidy(inter_model)[[2, 2]], 2)`, SE = `r round(tidy(inter_model)[[2, 3]], 2)`, t = `r round(tidy(inter_model)[[2, 4]], 2)`, p = `r round(tidy(inter_model)[[2, 5]], 2)`).
-By contrast, there was a significant increase in the Positive treatment of `r round(tidy(inter_model)[[4, 2]], 2)` per passage which is approximately a `r round((pos_slope - 1) * 100, 0)`% increase in methane oxidation rate per passage (SE = `r round(tidy(inter_model)[[4, 3]], 2)`, t = `r round(tidy(inter_model)[[4, 4]], 2)`, p = `r round(tidy(inter_model)[[4, 5]], 2)`).
+```{r save_model}
+save(response_model, pos_slope, file = '../Output/response_model.Rdata')
+```
+
+The intercept for the Neutral treatment was significantly different from zero
+at `r myround(tidy(response_model)[[1, 2]], 2)` (SE = `r
+myround(tidy(response_model)[[1, 3]], 2)`, t = `r myround(tidy(response_model)[[1,
+4]], 2)`, p = `r myround(tidy(response_model)[[1, 5]], 2)`). The Positive
+treatment started at a significantly lower methane oxidation rate than the
+Neutral treatment with a y-intercept of `r myround(tidy(response_model)[[3, 2]],
+2)` relative to Neutral (SE = `r myround(tidy(response_model)[[3, 3]], 2)`, t = `r
+myround(tidy(response_model)[[3, 4]], 2)`, p = `r myround(tidy(response_model)[[3,
+5]], 2)`). There was no change in methane oxidation rate of the Neutral
+treatment over the five passages (slope = `r myround(tidy(response_model)[[2, 2]],
+2)`, SE = `r myround(tidy(response_model)[[2, 3]], 2)`, t = `r
+myround(tidy(response_model)[[2, 4]], 2)`, p = `r myround(tidy(response_model)[[2,
+5]], 2)`). By contrast, there was a significant increase in the Positive
+treatment of `r myround(tidy(response_model)[[4, 2]], 2)` per passage which is
+approximately a `r round((pos_slope - 1) * 100, 0)`% increase in methane
+oxidation rate per passage (SE = `r myround(tidy(response_model)[[4, 3]], 2)`, t =
+`r myround(tidy(response_model)[[4, 4]], 2)`, p = `r myround(tidy(response_model)[[4,
+5]], 2)`).
 

### 2021-06-29 R/04-heritability.Rmd
update manuscript with new analyses and update makefile

@@
 ---
-title: "R Notebook"
-output:
-  html_document:
-    df_print: paged
+title: "Heritability"
+author: "Andrew Morris"
+date: "`r Sys.Date()`"
+output: html_document
 ---
 
-This script imports the heritability data.
-This dataset includes the mean flux withn treatment*generation for 1) all parents, 2) selected parents, and 3) offspring.
-Then, it plots offspring on the selection parents with the two treatments combined and then separately for the two treatments.
+This script imports the heritability data. This dataset includes the mean flux
+withn treatmentxgeneration for 1) all parents, 2) selected parents, and 3)
+offspring. Then, it plots offspring on the selection parents with the two
+treatments combined and then separately for the two treatments.
 
@@
 kable(tidy(fit_no_treat))
-fit_treat <- lm(offspring ~ selected * treat, data = heritability)
-kable(tidy(fit_treat))
-kable(tidy(anova(fit_no_treat, fit_treat)))
+herit_model <- lm(offspring ~ selected * treat, data = heritability)
+kable(tidy(herit_model))
+kable(tidy(anova(fit_no_treat, herit_model)))
 ```
@@
 Next, I added treatment and the interaction between treatment and parental flux.
-Here, there is a significant correlation between parent and offspring for both the Neutral treatment `r print_model(fit_treat, 2)` and the Positive treatment `r print_model(fit_treat, 4)`.
+Here, there is a significant correlation between parent and offspring for both the Neutral treatment `r print_model(herit_model, 2)` and the Positive treatment `r print_model(herit_model, 4)`.
 Notably, the sign of the effect is reversed: negative for the Neutral treatment and positive for the Positive treatment.
@@
 
+```
+
+```{r save_heritability}
+save(herit_model, file = '../Output/herit_model.Rdata')
 ```

### 2021-07-01 Manuscript/manuscript.Rmd
revise introduction

@@
 
-## Determining the mapping between microbial community structure and ecosystem function
+One of the fundamental questions in Ecology is To what extent does biodiversity
+drive variation in ecosystem functioning? Variation in abiotic conditions
+regulates the rate of ecosystem processes, but it remains unclear to what
+extent variation in community structure regulates the rate of ecosystem
+processes. Many ecosystem functions are mediated by organisms. In particular,
+microorganisms such as bacteria mediate the major biogeochemical cycles such
+as the carbon and nitrogen cycles. Despite the central role of bacteria in
+performing these processes, we have not found a consistent relationship between
+microbial community structure and the rate of any ecosystem process. In order
+to understand the importance of biodiversity to ecosystem functioning and to
+better predict and manage specific ecosystem processes, we need to determine
+the relationship between microbial community structure and ecosystem function.
 
-Microbial communities mediate ecosystem nutrient cycling via decomposition,
-greenhouse gas cycling, carbon storage, and other processes. Despite the
-central role of bacteria in performing these processes, we have not found a
-consistent relationship between the composition of a microbial community and
-the rate of an ecosystem process across habitats. Typically, research in this
-area has followed the model of macroorganismal biodiversity-ecosystem function
-studies such as grassland productivity studies that measure or manipulate the
-diversity of plant species and correlate that with the rate of primary
-productivity. By analogy, microbial ecologists use microbial sequencing data to
-create taxonomic relative abundance tables and then infer diversity or relative
-abundance from the whole archaeal/bacterial community and correlate that with
-the rate of a microbial process. Alternatively, they estimate the abundance of
-a putative functional gene (such as *pmoA* for methane oxidation) and correlate
-abundance with function. These approaches only rarely show significant,
-positive relationships between biodiversity and ecosystem function. In my first
-chapter published in Phil Trans I lay out philosophically why this approach
-shouldn’t work – namely that marker gene sequences are not analogous to plant
-species counts, but are more analogous to measurements of gene markers as in
-evolutionary biology studies.
+In macroorganismal communities, there appears to be a relationship between the
+diversity of a community and the resilience of ecosystem functioning to
+environmental perturbations (Tilman, ????... Hooper). Typically, this is the
+diversity (or more specifically, the richness) of a particular functional
+group, such as photosynthetic plants. Similar studies in microbial communities
+rarely find a relationship between diversity and function (Graham, ). However,
+these studies typically define diversity much more broadly than plant diversity
+studies by looking at the phylogenetic diversity of the entire bacterial and
+archaeal community, not just of a particular functional group. Studies that
+focus more narrowly on a particular ecosystem process and a functional group
+involved in that process still rarely find a significant correlation between
+the abundance of a functional group (or a functional protein-coding gene) and
+the rate of the corresponding process (Rocca, ????). For certain ecosystem
+functions in certain ecosystems, these relationships can be informative. For
+example, the abundance of a single methanogenic species (*G. species*) is
+correlated with methane emissions in a permafrost ecosystem (cite). In addition,
+bacterial and archaeal community composition is correlated with methane
+emissions in forests and pastures of the Brazilian Amazon (Meyer et al. 2020).
+Therefore, there appears to be some indication that microbial community
+structure drives variation in methane cycling in soils.
 
-In evolutionary biology, genotype-phenotype mapping provides a way to
-understand the fitness effects of a genotype. While selection typically acts on
-phenotype, the genotypic changes are the fundamental driver of phenotypic
-differences and thus fitness differences (Stadler and Stephes 2005). A similar
-analogy can be drawn for biodiversity-ecosystem function relationships. Traits
-of a community are what drive community assembly, not the taxa or genotypes
-themselves (Inkpen et al. 2017). However, as ecologists, we wish to understand
-the effects of species on ecosystem function. Therefore, in a similar way that
-evolutionary biologists want to deduce the genotype-phenoype map to understand
-the genetic basis of fitness, community ecologists must describe the "mapping"
-between community genotype and ecoystem function to understand what drives
-communty assembly, and ultimately, ecosystem function.
+To further investigate the relationship between microbial community structure
+and the rate of methane cycling in soils, we want to determine which organisms
+are important for regulating methane cycling. This ecological question is
+analogous to the problem of genotype-phenotype mapping in quantitative genetics
+(Morris 2020). Variation in an organimsal trait is partially determined by
+environmental conditions and partially determined by genetic varition. One goal
+of quantitative genetics is to determine the proportion of the variation
+attributable to genetics and to identify which alleles regulate the phenotypic
+trait. These goals are commonly achieved through artificial selection
+experiments and association mapping studies, such as genome-wide association
+studies. To determine the mapping between microbial community structure and
+ecosystem function, we applied these approaches of genotype-phenotype mapping
+to whole microbial communities. In this case, the "alleles" are different
+microbial taxa and the "phenotype" is the rate of ecosystem function at the
+whole-ecosystem level.
 
-The relationship that ecologists have described between plant diversity and
-productivity represents a relatively simple map between taxonomy and function.
-The number of unique members of a functional group (i.e., plant species) is
-correlated with the rate of the function performed by that group. However, for
-microbial communities that mapping is not quite as clear. For many microbial
-traits, there is not a clear coupling between taxonomy and function, except for
-in the case of highly conserved traits involving many genes (Martiny et al.
-2013). In addition, most microbial taxa have not been sampled and so we don't
-have a full picture of which taxa are members of a particular functional group
-(citation). Put another way, it's much harder to identify methanotrophic
-prokaryotes in soil than it is to identify plants in a prairie. Finally, even
-for highly conserved traits with a putative functional marker, for example
-methane oxidation and the gene *pmoA*, functional group abundance is not
-correlated with the rate of the function performed by that group in most
-ecosystems (Rocca et al. 2016). Because of these differences between
-macroorganisms and microorganisms, it may be more useful to approach microbial
-biodiversity-ecosystem function in a way analogus to genotype-phenotype mapping
-in organisms (Morris et al. 2020).
+To do this, we performed an artificial ecosystem selection experiment to
+identify communities of bacteria and archaea that collectively perform a high
+rate of methane oxidation (Swenson et al. 2000).  (Goodnight et al. 1997,
+Williams and Lenton 2007)we generated twenty-four soil ecosystems in the lab
+using a sterilized potting mix as the substrate. We then inoculated these
+ecosystems with a small amount of living soil to generate variation in
+community structure across the twenty-four ecosystems. These ecosystems were
+then assigned to one of two selection treatments: neutral selection with
+ecosystems chosen for reproduction at random or positive selection with
+ecosystems chosen for reproduction based on their methane oxidation rate. Each
+ecosystem was maintained at 1000 ppm CH4 over several weeks to allow for
+colonization and growth of the microbial community and to enrich for methane
+oxidizers. We then determined the methane oxidation rate for each ecosystem and
+selected three jars in each treatment to inoculate the next set of twenty-four
+jars. This process continued for five passages. At the end of the experiment,
+we extracted DNA from soils in the second and fifth passage for both the
+neutral and positive treatments. By comparing the response to selection within
+the community between the two treatments, we can identify microbial species
+that contribute to variation in ecosystem methane emissions.
 
-One approach to deducing the genotype-phenotype map in organisms is through
-artificial selection (Fuller et al. 2005). In this approach, selection for a
-trait is imposed in a contrived laboratory environment and compared to
-laboratory environment without selection on that trait. By comparing the
-genetic respones between these two treatments, you can remove the
-environment-fitness connection that would be present in a comparative study and
-identify genes that are evolving in response to selection only on the trait of
-interest. It has been shown that whole communities can respond to artifical
-selection on traits of the ecosystem (Goodnight et al. 1997, Williams and
-Lenton 2007). And it has been proposed that this could be a suitable approach
-for identify communities of organisms that are capable of performing a
-particular function at a high rate (Swenson et al. 2000). Therefore, artificial
-ecosystem selection is a potentially fruitful approach to identifying the map
-between biodiversity and ecosystem function in microbial communities.
-
-To identify the map between microbial community composition and ecosystem
-function, we imposed artificial ecosystem selection on methane oxidation rate
-of whole soil ecosystems. We prepared two lines of replicate ecosystems: one in
-which we imposed positive selection on methane oxidation by only using the top
-performing ecosystems to generate the next set of ecosystems and one in which
-we imposed neutral selection where a random set of ecosystems were chosen to
-generate the next set of ecosystems. We repated this process over multiple
-passages and then compared the resultant metacommunities in terms of their rate
-of ecosystem function and the composition of their communities.
-
-With this experiment, we asked several questions. First, is there a response to
-selection on methane oxidation by soil ecosystems? And how strong is the
-response to selection? (i.e., how "heritable" is methane oxidation as a
-community trait in terms of the phenotypic resemblence of ecosystems to the
-ecosystems that generated them?) If methane oxidation is heritable, then what
-is the nature of the map between microbial community composition and methane
-oxidation? Is it the diversity of methane oxidizers, as it is for plants and
-productivity? Or is there a more complex map involving upstream metabolic
-regulation?
-
-Things this Intro doesn't yet discuss:
-
-1. What a more complex map might look like. For example, maybe pmoA and methanotrophy aren't correlated, but maybe understand methanogenesis or nitrogen cycling is helpful.
-
-2. The idea of "heritability" for communities/ecosystems and why that relates to environmental vs. genetic control over ecosystem function.
+This approach is powerful because by using a common soil substrate and
+maintaining a constant headspace concentration of methane, we can eliminate
+much of the environmental variation that would be present in an observational
+study in the wild. In addition, by applying selection on methane oxidation rate
+in the positive treatment, we can enrich for taxa involved in methane cycling
+making it easier to perform association mapping between microbial taxa and
+ecosystem methane emissions. In this paper, we address whether there is a
+response to selection on methane oxidation rate at the whole ecosystem level. We
+then quantify the amount of variation in methane oxidation attributable to
+variation in community composition within our laboratory environment. Finally,
+we identify microbial markers of soil methane cycling using association mapping
+in order to deduce the mapping between microbial community structure and
+ecosystem function.
 

### 2021-07-01 Manuscript/manuscript.Rmd
update methods and response to selection results

@@
 One of the fundamental questions in Ecology is To what extent does biodiversity
-drive variation in ecosystem functioning? Variation in abiotic conditions
+drive variation in ecosystem function? Variation in abiotic conditions
 regulates the rate of ecosystem processes, but it remains unclear to what
@@
 performing these processes, we have not found a consistent relationship between
-microbial community structure and the rate of any ecosystem process. In order
+microbial community structure and the rate of an ecosystem process. In order
 to understand the importance of biodiversity to ecosystem functioning and to
@@
 In macroorganismal communities, there appears to be a relationship between the
-diversity of a community and the resilience of ecosystem functioning to
+diversity of a community and the resilience of ecosystem functions to
 environmental perturbations (Tilman, ????... Hooper). Typically, this is the
@@
 studies by looking at the phylogenetic diversity of the entire bacterial and
-archaeal community, not just of a particular functional group. Studies that
-focus more narrowly on a particular ecosystem process and a functional group
-involved in that process still rarely find a significant correlation between
-the abundance of a functional group (or a functional protein-coding gene) and
-the rate of the corresponding process (Rocca, ????). For certain ecosystem
+archaeal community. Studies that focus more narrowly on a particular ecosystem
+process and a functional group involved in that process still rarely find a
+significant correlation between the abundance of a functional group and the
+rate of the corresponding process (Rocca, ????). For certain ecosystem
 functions in certain ecosystems, these relationships can be informative. For
 example, the abundance of a single methanogenic species (*G. species*) is
-correlated with methane emissions in a permafrost ecosystem (cite). In addition,
-bacterial and archaeal community composition is correlated with methane
-emissions in forests and pastures of the Brazilian Amazon (Meyer et al. 2020).
-Therefore, there appears to be some indication that microbial community
+correlated with methane emissions in a permafrost ecosystem (cite). In
+addition, bacterial and archaeal community composition is correlated with
+methane emissions in forests and pastures of the Brazilian Amazon (Meyer et al.
+2020).  Therefore, there appears to be some indication that microbial community
 structure drives variation in methane cycling in soils.
@@
 analogous to the problem of genotype-phenotype mapping in quantitative genetics
-(Morris 2020). Variation in an organimsal trait is partially determined by
-environmental conditions and partially determined by genetic varition. One goal
+(Morris 2020). Variation in an organismal trait is partially determined by
+environmental conditions and partially determined by genetic variation. One goal
 of quantitative genetics is to determine the proportion of the variation
@@
 
-To do this, we performed an artificial ecosystem selection experiment to
-identify communities of bacteria and archaea that collectively perform a high
-rate of methane oxidation (Swenson et al. 2000).  (Goodnight et al. 1997,
-Williams and Lenton 2007)we generated twenty-four soil ecosystems in the lab
-using a sterilized potting mix as the substrate. We then inoculated these
-ecosystems with a small amount of living soil to generate variation in
-community structure across the twenty-four ecosystems. These ecosystems were
-then assigned to one of two selection treatments: neutral selection with
-ecosystems chosen for reproduction at random or positive selection with
-ecosystems chosen for reproduction based on their methane oxidation rate. Each
-ecosystem was maintained at 1000 ppm CH4 over several weeks to allow for
-colonization and growth of the microbial community and to enrich for methane
-oxidizers. We then determined the methane oxidation rate for each ecosystem and
-selected three jars in each treatment to inoculate the next set of twenty-four
-jars. This process continued for five passages. At the end of the experiment,
-we extracted DNA from soils in the second and fifth passage for both the
-neutral and positive treatments. By comparing the response to selection within
-the community between the two treatments, we can identify microbial species
-that contribute to variation in ecosystem methane emissions.
+To identify communities of bacteria and archaea that collectively perform a
+high rate of methane oxidation, we performed an artificial ecosystem selection
+experiment  (Swenson et al. 2000).  We generated twenty-four soil ecosystems in the lab using a sterilized
+potting mix as the substrate. We then inoculated these ecosystems with a small
+amount of living soil to generate variation in community structure across the
+twenty-four ecosystems. These ecosystems were then assigned to one of two
+selection treatments: neutral selection or positive selection. Each ecosystem
+was maintained at 1000 ppm CH~4~ over several weeks to allow for colonization and
+growth of the microbial community and to enrich for methane oxidizers. We then
+determined the methane oxidation rate for each ecosystem. To impose selection,
+we selected three jars from the positive treatment with the highest rate of
+methane oxidation, homogenized them, and used that soil to inoculate the next
+set of jars. For the neutral treatment, we selected three jars at random to
+inoculate the next set of jars. This process continued for five passages. At
+the end of the experiment, we extracted DNA from soils in the second and fifth
+passage for both the neutral and positive treatments. By comparing the response
+to selection within the community between the two treatments, we can identify
+microbial species that contribute to variation in ecosystem methane emissions.
 
@@
 much of the environmental variation that would be present in an observational
-study in the wild. In addition, by applying selection on methane oxidation rate
+study. In addition, by applying selection on methane oxidation rate
 in the positive treatment, we can enrich for taxa involved in methane cycling
@@
 
+(Goodnight et al. 1997, Williams and Lenton
+2007)
+
 # Materials & Methods
@@
 a positive selection line where the two or three jars with the highest CH~4~
-oxidation rate were chosen to inoculate the next set of jars. The other line
-was a neutral selection line where an equal number of jars as the positive line
-were chosen at random among the twelve. The number of jars chosen was based on
-the distribution of fluxes among the positive jars, i.e., we chose the top
-three jars unless only two jars had considerably greater oxidation rates based
-on visual inspection of histograms.
+oxidation rate were chosen to inoculate the next set of positive jars. The
+other line was a neutral selection line where an equal number of jars as the
+positive line were chosen at random to inoculate the next set of neutral jars.
+The number of jars chosen was based on the distribution of fluxes among the
+positive jars, i.e., we chose the top three jars unless only two jars had
+considerably greater oxidation rates based on visual inspection of histograms.
+The experiment was carried out over five passages.
 
@@
 upland mineral soil under a deciduous forest ecosystem near the University of
-Oregon campus. Incubations were performed in 500 mL mason jars (Ball
-Corporation, Broomfield, Colorado, USA) with rubber septa installed in the
-lids. Jars were incubated at ambient temperature on the benchtop. To create the
-first set of jars, twenty-four jars were created in an identical manner and
-then randomly assigned to either the positive or negative selection line. Each
-jar was sterilized with 70% ethanol to which was added 45 g of autoclaved
-potting mix, 5 g of live soil, and 3.5 mL of sterile deionized water to bring
-them to <mark> 60%??? of field capacity </mark>. The jars were then
-homogenized, capped, and injected with 4.3 mL of 99% CH~4~ to bring the
-headspace concentration to approximately 1000 ppm CH~4~. Jars were flushed and
-respiked twice per week to maintain elevated CH~4~ concentrations until travel,
-holidays, and COVID-19 got in the way and then we just did it as often as we
-could. Jars were incubated for approximately three weeks and then flux
-measurements were taken.
+Oregon campus in Eugene, OR, USA. Incubations were performed in 500 mL mason
+jars (Ball Corporation, Broomfield, Colorado, USA) with rubber septa installed
+in the lids. Each jar was sterilized with 70% ethanol to which was added 45 g
+of autoclaved potting mix, 5 g of living soil, and 3.5 mL of sterile deionized
+water to bring them to <mark> 60%??? of field capacity</mark>. The jars were
+then homogenized, capped, and injected with 4.3 mL of 99% CH~4~ to bring the
+headspace concentration to approximately 1000 ppm CH~4~. To create the two
+treatment groups, twenty-four jars were created in an identical manner and then
+randomly assigned to either the positive or neutral selection line. Jars were
+flushed and respiked twice per week to maintain elevated CH~4~ concentrations
+and were incubated at ambient temperature for approximately three weeks.
+Methane oxidation rates were determined at the end of the incubation period.
+These values were then used to determine the jars selected to inoculate the
+next set of jars.
 
-Once CH~4~ oxidation rates were determined, the top two or three jars were
-chosen for selection in the positive line and an equal number of jars were
-randomly chosen from the neutral line. The selected jars for each line were
-homogenized and used as the 5 g live soil for the subsequent passage. Twelve
-new jars were created for each line in the same manner as above using the
-inoculum for that line. This procedure continued until a divergence was
-observed in the mean flux of each line for two consecutive generations. Jars in
-passage number 2 and 5 were homogenized and a sample was collected and stored
-at -80$^\circ$C for later DNA extraction.
+## Methane oxidation rate measurements
 
-## Methane flux measurements Methane oxidation rates were determined after
-flushing and spiking jars to approximately 1000 ppm CH~4~. Head-space samples
-of 1 mL were collected from each jar immediately after spiking and then at time
-points 3, 6, 24, and 48 hours for a 5-point curve. Samples were immediately
-injected into a <mark>???</mark> gas chromatograph fitted with an electron
-capture detector (company, city, state, USA) to determine the headspace CH~4~
-concentration. Fluxes were calculated from a first-order exponential decay
-function as decay constant `k`. Oxidation rates are presented as the additive
+Methane oxidation rates were determined after flushing and spiking jars to
+approximately 1000 ppm CH~4~. Headspace samples of 1 mL were collected from
+each jar immediately after spiking and then at time points 3, 6, 24, and 48
+hours for a 5-point curve. Samples were immediately injected into a
+<mark>???</mark> gas chromatograph fitted with an electron capture detector
+(company, city, state, USA) to determine the headspace CH~4~ concentration.
+Fluxes were calculated from a first-order exponential decay function as decay
+constant `k` with units day^-1^. Oxidation rates are presented as the additive
 inverse of `k` (i.e., `-k`) so that a more positive value represents a faster
-oxidation rate.
+oxidation rate. The jars selected for the positive treatment in passage 2 had
+the lowest methane oxidation rate of the twelve jars due to a calculation error
+in the methane oxidation rate. All other passages correctly used the top three jars.
 
@@
 
-Soil DNA was extracted from a 0.25 g subsample of soil from the starting
-inoculum and from each jar in passage 2 and 5. Negative controls were extracted
+A subsample of soil from the starting inoculum and from every jar in passages 2
+and 5 was collected and stored at -80$^\circ$C for later DNA extraction.
+Soil DNA was extracted from a 0.25 g subsample of soil. Negative controls were extracted
 from autoclaved potting mix and DNase-free water. Extractions were performed
@@
 sequenced the V3-V4 region of the 16S rRNA gene using the Earth Microbiome
-Project protocol and primers<mark>???</mark>. In addition, to estimate the
-diversity and relative abundance of CH~4~ oxidizing taxa, we sequenced the
-particulate methane monooxygenase subunit A gene (*pmoA*) using
-primers<mark>???</mark>. Sequencing was performed on the Illumina MiSeq in v3
-mode with 300 bp paired-end reads.
+Project protocol and primers<mark>???</mark>. Sequencing was performed on the
+Illumina NovaSeq 6000 with paired-end 150 bp reads.
 
-## Bioinformatics Bioinformatic analyses were conducted using the `QIIME2`
+## Bioinformatics
+
+Bioinformatic analyses were conducted using the `QIIME2`
 bioinformatics platform (Bolyen et al. 2019). Raw sequencing reads were
@@
 To test whether there was a significant change in CH~4~ oxidation rate (k) as a
-response to selection, we tested the deviation between the positive and neutral
-selection lines over multiple passages. We calculated deviation as the
-difference in the mean oxidation rate between the positive and neutral
-treatment within each passage. The standard errors of the difference were
-calculated as the square root of the sum of the squared errors of the positive
-and neutral lines. We then fit an ordinary least squares regression model
-taking into account the standard errors of the deviations. The model was fit
-using the `rma` function from the `metafor` package in `r` (Viechtbauer 2010).
+response to selection, we tested a difference in slopes between the positive
+and neutral selection lines. The CH~4~ oxidation rates were strongly right
+skewed with most values close to zero and few large, positive values. This
+resulted in residuals that did not meet the assumptions of constant variance
+and normal distribution. Therefore, methane oxidation rates were log~10~
+transformed to better meet the assumptions of a linear model and to make
+figures easier to interpret. First, we tested the effect of treatment by
+fitting two nested models with and without treatment using `lm` in `R`. The
+model formulation for the full model was `flux ~ passage * treatment` and the
+model formulation for the reduced model was `flux ~ passage`. We compared these
+models using the likelihood ratio test with the `anova` function in `R`. We
+then fit the full model to determine the slope of the positive line, which
+represents the change in methane oxidation rate per passage as a response to
+selection.
 
@@
 
-The intercept for the Neutral treatment was significantly different from zero
-at `r myround(tidy(response_model)[[1, 2]], 2)` (SE = `r
-myround(tidy(response_model)[[1, 3]], 2)`, t = `r myround(tidy(response_model)[[1,
-4]], 2)`, p = `r myround(tidy(response_model)[[1, 5]], 2)`). The Positive
-treatment started at a significantly lower methane oxidation rate than the
-Neutral treatment with a y-intercept of `r myround(tidy(response_model)[[3, 2]],
-2)` relative to Neutral (SE = `r myround(tidy(response_model)[[3, 3]], 2)`, t = `r
-myround(tidy(response_model)[[3, 4]], 2)`, p = `r myround(tidy(response_model)[[3,
-5]], 2)`). There was no change in methane oxidation rate of the Neutral
-treatment over the five passages (slope = `r myround(tidy(response_model)[[2, 2]],
-2)`, SE = `r myround(tidy(response_model)[[2, 3]], 2)`, t = `r
-myround(tidy(response_model)[[2, 4]], 2)`, p = `r myround(tidy(response_model)[[2,
-5]], 2)`). By contrast, there was a significant increase in the Positive
-treatment of `r myround(tidy(response_model)[[4, 2]], 2)` per passage which is
-approximately a `r round((pos_slope - 1) * 100, 0)`% increase in methane
-oxidation rate per passage (SE = `r myround(tidy(response_model)[[4, 3]], 2)`, t =
-`r myround(tidy(response_model)[[4, 4]], 2)`, p = `r myround(tidy(response_model)[[4,
-5]], 2)`).
-
-Methane oxidation rate increased as a response to selection in the Positive
-selection treatment (`r figures("deviance", display="cite")`).  This can be
-seen in the increase in deviation between the Positive and Neutral selection
-lines over the course of the experiment (z = , p = ).  Mean CH~4~ oxidation
-rate for the Positive treatment increased by  $\pm$ per passage relative to the
-Neutral treatment. The mean flux across all jars in all treatments in the first
-passage was  $\pm$  so this represents a doubling of CH~4~ consumption in every passage due
-to selection.
+We observed a response to selection on soil methane oxidation rate at the whole
+ecosystem level (`r figures("response", display="cite")`).
+At the start of the experiment, the Positive
+treatment had a mean methane oxidation rate that was
+`r percent_less(initial = tidy(response_model)[[1, 2]], final = tidy(response_model)[[3, 2]] + tidy(response_model)[[1, 2]])`%
+lower than the
+Neutral treatment (SE = 
+`r myround(tidy(response_model)[[3, 3]], 2)`, t = 
+`r myround(tidy(response_model)[[3, 4]], 2)`, p = 
+`r myround(tidy(response_model)[[3, 5]], 2)`). There was no change in methane oxidation rate of the Neutral
+treatment over the five passages (slope = 
+`r myround(tidy(response_model)[[2, 2]], 2)`, SE = 
+`r myround(tidy(response_model)[[2, 3]], 2)`, t = 
+`r myround(tidy(response_model)[[2, 4]], 2)`, p = 
+`r myround(tidy(response_model)[[2, 5]], 2)`). By contrast, the Positive treatment had a
+`r round(per_slope)`% increase in methane
+oxidation rate per passage (slope = 
+`r myround(tidy(response_model)[[4, 2]], 2)`, SE = 
+`r myround(tidy(response_model)[[4, 3]], 2)`, t =
+`r myround(tidy(response_model)[[4, 4]], 2)`, p = 
+`r myround(tidy(response_model)[[4, 5]], 2)`).
 

### 2021-07-01 R/03-response.Rmd
update methods and response to selection results

@@
 
+```{r raw_model}
+raw_model <- lm(ch4 ~ passage * treat, data = fluxes)
+
+ggplot(raw_model, aes(x = .fitted, y = .resid)) + geom_point()
+
+```
+
 ```{r response_model}
-# Fit full interaction model
 response_model <- lm(log_ch4 ~ passage * treat, data = fluxes)
 
+ggplot(response_model, aes(x = .fitted, y = .resid)) + geom_point()
+
+```
+
+```{r no_treat_model}
 # No treatment model
@@
 pos_slope <- 10^response_model$coef["passage:treatp"]
-
+per_slope <- (10^response_model$coef["passage:treatp"] - 1) * 100
 ```
@@
 ```{r save_model}
-save(response_model, pos_slope, file = '../Output/response_model.Rdata')
+save(response_model, per_slope, file = '../Output/response_model.Rdata')
 ```

### 2021-07-01 R/functions.R

@@ -16,7 +16,7 @@ plot_flux <- function(fluxes, passage, estimate, ratio = NULL, log10 = FALSE) {
     p <- p + labs(x = "Passage Number", y = "Methane Oxidation Rate (-k)")
   } 
   else if (log10 == TRUE) {
-    p <- p + labs(x = "Passage Number", y = expression(log[10] * "(Methane Consumption Rate (-k))"))
+    p <- p + labs(x = "Passage Number", y = expression(log[10] * "(Methane Oxidation Rate (-k))"))
   } 
   
   if (!is.null(ratio)) {
@@ -51,4 +51,8 @@ plot_herit <- function(heritability, ratio) {
                        labels = c('Neutral', 'Positive'), 
                        values = c('gray40', 'darkorange2')) +
     coord_fixed(ratio)
+}
+
+percent_less <- function(initial, final) {
+  round(-100 * (final - initial)/abs(initial))
 }

### 2021-07-27 Manuscript/manuscript.Rmd
update manuscript text

@@
 structure drives variation in methane cycling in soils.
+
+Why haven't we been more successful in identifying relationships between
+microbial community structure and ecosystem functions? Microbial communities
+are super diverse, particularly in soil environments, and
+looking at the whole community using broad phylogenetic markers such as 16S
+makes it difficult to sift through all of that diversity to find informative
+markers. Looking at individual protein-coding genes manages to reduce that
+diversity, but it also assumes we know what limits the rate of a process in a
+particular ecosystem. Typically, we choose the gene that encodes the final step
+in a pathway - for example, the mcrA gene, which encodes a subunit of the
+enzyme the performs the final step in the methanogenesis pathway. However,
+regulation may happen further up stream within the pathway and therefore the
+abundance of the final gene wouldn't correlate with the rate of the process.
+In fact, the limitation may be in substrate creation by other functional groups
+unrelated to methanogens. To overcome these challenges, we chose a
 

### 2021-08-17 Manuscript/manuscript.Rmd
update manuscript and analyses and figures plus add citations

--- 2021-07-01 Manuscript/manuscript.Rmd
@@
     reference_docx: reference.docx
+bibliography: bibliography.bib
 editor_options:
@@
 library(morris)
-#library(metafor)
 ```
+
 ```{r load_data}
@@
 figures <- captioner()
-figures("response", "Deviance of CH~4~ oxidation rate for the Positive selection treatment relative to the Neutral selection treatment.
-Deviance is calculated as the mean of P within each passage minus the mean of N within each passage.
-Errorbars are the standard error of the difference between P and N for each passage.
-The best fit line is an ordinary least squares fit taking into account the standard errors of the deviations.
-The gray ribbon is the 95% confidence interval for the regression.")
-figures("herit", "Regression of offspring CH~4~ flux on mid-parent CH~4~ flux.
-Mid-parent is the mean of jars selected to inoculate the next passage.
-The mid-offsprings are the mean of all twelve jars produced in one passage.
-The best fit line is an ordinary least squares regression line and the gray ribbon is the 95% confidence interval for the regression coefficient.")
+
+figures("response", "Response to selection on soil CH~4~ oxidation rate fit by ordinary least squares regression.
+The y-axis is the log~10~ of the additive inverse of the first-order exponential decay constant `k` (i.e., `-k`) with units day^-1^.
+The x-axis is the passage number.
+The orange line and points are the Positive treatment and the gray line and points are the Neutral treatment.
+The dark points are the selected jars in each passage and the light points are the non-selected jars.
+Regression lines aIn passage 2, the jars with the small `-k` were selected due to an error in the flux calculations.")
+
+figures("herit", "Ordinary least squares regression of mid-offspring CH~4~ flux on mid-parent CH~4~ flux.
+Mid-parent is the mean of the jars selected to inoculate the next passage.
+Mid-offspring are the mean of all twelve jars produced in one passage.
+The orange lines and points are the Positive treatment and the gray lines and points are the Neutral treatment.")
 
@@
 
-One of the fundamental questions in Ecology is To what extent does biodiversity
-drive variation in ecosystem function? Variation in abiotic conditions
-regulates the rate of ecosystem processes, but it remains unclear to what
-extent variation in community structure regulates the rate of ecosystem
-processes. Many ecosystem functions are mediated by organisms. In particular,
-microorganisms such as bacteria mediate the major biogeochemical cycles such
-as the carbon and nitrogen cycles. Despite the central role of bacteria in
-performing these processes, we have not found a consistent relationship between
-microbial community structure and the rate of an ecosystem process. In order
-to understand the importance of biodiversity to ecosystem functioning and to
-better predict and manage specific ecosystem processes, we need to determine
-the relationship between microbial community structure and ecosystem function.
+Soil microbial communities are central to the functioning of ecosystems
+[@crowther2019]. And yet, microbes have historically been excluded from
+ecosystems models in favor of abiotic variables [@schimel1995]. This is in part
+because attempts to establish quantitative relationships between microbial
+marker genes and ecosystem functions have not been fruitful [@graham2016,
+@rocca2015]. Here we use a novel approach to estimate how much of the variation
+in ecosystem function can be attributed to variation in the microbial community.
+We then identify microbial markers of ecosystem function. These experiments will
+expand our ability to model microbial ecosystem-function relationships and allow
+us to manage ecosystems for particular ecosystem functions [@fuhrman2009,
+@schimel1998].
 
@@
 diversity of a community and the resilience of ecosystem functions to
-environmental perturbations (Tilman, ????... Hooper). Typically, this is the
-diversity (or more specifically, the richness) of a particular functional
-group, such as photosynthetic plants. Similar studies in microbial communities
-rarely find a relationship between diversity and function (Graham, ). However,
-these studies typically define diversity much more broadly than plant diversity
-studies by looking at the phylogenetic diversity of the entire bacterial and
-archaeal community. Studies that focus more narrowly on a particular ecosystem
-process and a functional group involved in that process still rarely find a
-significant correlation between the abundance of a functional group and the
-rate of the corresponding process (Rocca, ????). For certain ecosystem
-functions in certain ecosystems, these relationships can be informative. For
-example, the abundance of a single methanogenic species (*G. species*) is
-correlated with methane emissions in a permafrost ecosystem (cite). In
-addition, bacterial and archaeal community composition is correlated with
-methane emissions in forests and pastures of the Brazilian Amazon (Meyer et al.
-2020).  Therefore, there appears to be some indication that microbial community
-structure drives variation in methane cycling in soils.
+environmental perturbations [@hooper2005]. Typically, diversity is represented
+as the richness of a particular functional group, such as photosynthetic plants.
+Similar studies in microbial communities rarely find a relationship between
+diversity and function [@graham2016]. However, these studies define
+diversity much more broadly by looking at the richness of the entire bacterial
+and archaeal community using 16S rRNA gene variants. Studies that focus more
+narrowly on a particular functional group still rarely find a significant
+correlation between the abundance of a functional group and the rate of the
+corresponding process [@rocca2015]. For certain functions in certain
+ecosystems, these relationships can be informative. For example, the abundance
+of a single methanogenic species (Candidatus ‘*Methanoflorens
+stordalenmirensis*’) is correlated with the CH~4~:CO~2~ ratio in a permafrost
+ecosystem [@mccalley2014]. In addition, bacterial and archaeal community composition is
+correlated with methane emissions in forests and pastures of the Brazilian
+Amazon [@meyer2020]. Therefore, there appears to be some indication that
+microbial community structure drives variation in methane cycling in soils.
+
+Why have we been unsuccessful at identifying relationships between
+microbial community structure and ecosystem functions? Microbial communities are
+extremely diverse, particularly in soil environments [@whitman1998].
+Characterizing these communities using broad phylogenetic markers, such as 16S,
+reveals thousands of mostly uninformative taxa which makes it difficult to
+identify markers of ecosystem function. Some studies constrain this diversity by
+characterizing individual protein-coding genes. However, this approach assumes
+investigators already know what limits the rate of a process in a particular
+ecosystem. Typically, this is assumed to be the gene that encodes the final step
+in a pathway--for example, the *mcrA* gene, which encodes a subunit of the
+enzyme that performs the final step in methanogenesis. However, regulation may
+happen further upstream (or downstream) within the pathway and therefore the
+abundance of the final gene would not correlate with the rate of the process. In
+the case of methane emissions, the limitation may be in generating the
+substrates for methanogenesis or may be in the consumption of methane by
+methanotrophs.
 
 To further investigate the relationship between microbial community structure
-and the rate of methane cycling in soils, we want to determine which organisms
-are important for regulating methane cycling. This ecological question is
-analogous to the problem of genotype-phenotype mapping in quantitative genetics
-(Morris 2020). Variation in an organismal trait is partially determined by
-environmental conditions and partially determined by genetic variation. One goal
-of quantitative genetics is to determine the proportion of the variation
-attributable to genetics and to identify which alleles regulate the phenotypic
-trait. These goals are commonly achieved through artificial selection
-experiments and association mapping studies, such as genome-wide association
-studies. To determine the mapping between microbial community structure and
-ecosystem function, we applied these approaches of genotype-phenotype mapping
-to whole microbial communities. In this case, the "alleles" are different
-microbial taxa and the "phenotype" is the rate of ecosystem function at the
-whole-ecosystem level.
+and ecosystem function, we want to determine how much of the variation in soil
+methane cycling is attributable to variation in the microbial community. This
+ecological question is analogous to the problem of estimating heritability in
+quantitative genetics [@morris2020]. Variation in an organismal trait is
+determined by the sum of genetic variation, environmental variation, and the interaction between the two [@falconer1996]. One goal of quantitative genetics is to determine the
+proportion of the variation attributable to genetics. This is commonly
+estimated as the narrow-sense heritability (h^2^) defined as the additive
+genetic variance. This goal can be achieved through artificial selection
+experiments with selection on the character of interest [@falconer1996]. We
+applied artificial ecosystem selection to whole microbial communities by
+selecting on the ecosystem-scale methane oxidation rate in order to estimate the
+additive genetic variance of soil methane oxidation [@swenson2000].
+
+To identify which organisms are important for regulating methane cycling, we
+compared the composition of the artificially selected microbial
+communities and a control set of communities without selection. Artificial
+ecosystem selection has a similar effect to enrichment culturing by amplifying
+the population of interest - in this case methanotrophs [@swenson2000]. This will reduce the
+diversity of soil microbes and allow for greater power in detecting significant
+markers of methane oxidation rate. To do this, we extracted DNA from several of
+the ecosystems within the selection experiment and sequenced the 16S rRNA gene
+in order to perform a differential abundance test on methane oxidation rate.
+
+# Materials & Methods
+
+## Selection experiment
+We performed an artificial ecosystem selection experiment [*sensu* @swenson2000]
+by passaging replicate soil microbial communities. The trait we selected on was
+CH~4~ oxidation rate. Our experiment had two selection lines with twelve jars
+each for a total of twenty-four jars per passage. One line was a positive
+selection line where the two or three jars with the highest CH~4~ oxidation rate
+were chosen to inoculate the next set of positive jars. The other line was a
+neutral selection line where an equal number of jars as the positive line were
+chosen at random to inoculate the next set of neutral jars. The number of jars
+chosen was based on the distribution of fluxes among the positive jars, i.e., we
+chose the top three jars unless only two jars had considerably greater oxidation
+rates based on visual inspection of histograms. The experiment was carried out
+over five passages.
+
+The initial soil microbial community was sampled from the top 10 cm of an upland
+mineral soil under a deciduous forest ecosystem near the University of Oregon
+campus in Eugene, OR, USA. Incubations were performed in 500 mL mason jars with
+rubber septa installed in the lids. Each jar was sterilized with 70% ethanol to
+which was added 45 g of autoclaved potting mix, 5 g of living soil, and 3.5 mL
+of sterile deionized water to bring them to 60% of field capacity. The jars were
+then homogenized, capped, and injected with 4.3 mL of 99% CH~4~ to bring the
+headspace concentration to approximately 1000 ppm CH~4~. To create the two
+treatment groups, twenty-four jars were created in an identical manner and then
+randomly assigned to either the positive or neutral selection line. Jars were
+flushed and respiked twice per week to maintain aerobic conditions and elevated
+CH~4~ concentrations and were incubated at ambient temperature for approximately
+three weeks. Methane oxidation rates were determined at the end of the
+incubation period. These values were then used to determine the jars selected to
+inoculate the next set of jars.
+
+## Methane oxidation rate measurements
+
+Methane oxidation rates were determined after flushing and spiking jars to
+approximately 1000 ppm CH~4~. Headspace samples of 1 mL were collected from each
+jar immediately after spiking and then at time points 3, 6, 24, and 48 hours for
+a 5-point curve. Samples were immediately injected into a SRI model 8610C gas
+chromatograph (Torrance, CA, USA) to determine the headspace CH~4~
+concentration. Fluxes were calculated from a first-order exponential decay
+function as decay constant `k` with units day^-1^. Oxidation rates are presented
+as the additive inverse of `k` (i.e., `-k`) so that a more positive value
+represents a greater oxidation rate. The jars selected for the positive
+treatment in passage 2 had the lowest methane oxidation rate of the twelve jars
+due to a calculation error in the methane oxidation rate. All other passages
+correctly used the top three jars.
+
+## Soil DNA extraction and sequencing
+
+A subsample of soil from the starting inoculum and from every jar in passages 2
+and 5 was collected and stored at -80$^\circ$C. Soil DNA was extracted from 0.25
+g soil. Negative controls were extracted from autoclaved potting mix and
+DNase-free water. Extractions were performed using the DNeasy PowerSoil kit
+(QIAGEN, Germany) and quantified using Qubit dsDNA HS Assay Kit (Thermo Fisher
+Scientific, USA). In order to estimate the diversity and relative abundance of
+the bacterial and archaeal taxa in our soil ecosystems, we sequenced the V4
+region of the 16S rRNA gene using the using the 515F − 806R primer combination
+[@caporaso2011]. PCR mixtures were: 10 $\mu$l NEBNext Q5 Hot Start HiFi PCR
+master mix, 9.2 $\mu$l primer mixture (1.09 $\mu$M concentration), and 0.8
+$\mu$l of DNA template. Reaction conditions were: 98$^\circ$C for 30 s
+(initialization), 35 cycles of 98$^\circ$C for 10 s (denaturation), 61$^\circ$C
+for 20 s (annealing), and 72$^\circ$C for 20 s (extension), and 72$^\circ$C for
+2 m (final extension). Amplicons were purified twice using 0.8x ratio Mag-Bind
+RxnPure Plus isolation beads (Omega Bio-Tek, USA). Sequencing libraries were
+prepared using a dual-indexing approach (Fadrosh et al., 2014; Kozich et al.,
+2013). Amplicon concentrations were quantified using Qubit (Thermo Fisher
+Scientific Technologies, USA) multiplexed at equimolar concentration. Sequencing
+was performed at the University of Oregon Genomics Core Facility on the Illumina
+NovaSeq 6000 with paired-end 150 bp reads.
+
+## Bioinformatics
+
+Bioinformatics processing was performed in `R` (R Core Team 2021). Demultiplexed
+sequencing reads were denoised using `DADA2` [@callahan2016]. Taxonomic
+assignment for the 16S reads was performed using the RDP naive Bayesian
+classifier [@wang2007].
+
+## Statistical Analysis
+
+### Response to selection
+
+To test whether there was a significant change in CH~4~ oxidation rate (k) as a
+response to selection, we tested a difference in slopes between the positive
+and neutral selection lines. The CH~4~ oxidation rates were strongly right
+skewed with most values close to zero and few large, positive values. This
+resulted in residuals that did not meet the assumptions of constant variance
+and normal distribution. Therefore, CH~4~ oxidation rates were log~10~
+transformed to better meet the assumptions of a linear model and to make
+figures easier to interpret. First, we tested the effect of treatment by
+fitting two nested models with and without treatment using `lm` in `R`. The
+model formulation for the full model was `flux ~ passage * treatment` and the
+model formulation for the reduced model was `flux ~ passage`. We compared these
+models using the likelihood ratio test with the `anova` function in `R`. We
+then fit the full model to estimate the slope of the positive line, which
+represents the change in CH~4~ oxidation rate per passage as a response to
+selection.
+
+### Ecosystem heritability
+
+To estimate the proportion of variance in ecosystem function due to variation in
+the microbial community, we estimated narrow-sense heritability ($h^2$) as the
+regression of offspring on mid-parent [@falconer1996]. The mid-parent was the
+mean for all three selected jars and the offspring was the mean for all jars
+produced by those parents. First, we tested if there was an effect of treatment
+on the heritability estiamte. The model formulation for the full model was
+`offspring ~ parent * treatment` and the model formulation for the reduced model
+was `offspring ~ parent`. We compared these models using the likelihood ratio
+test with the `anova` function `R`. We then fit the full model to estimate the
+heritability of the Positive and Neutral treatments.
+
+### Community response to selection
+
+Diversity estimates and tests were performed using `DivNet` [@willis2020]. We
+tested a difference in Shannon diversity between the Positive and Neutral
+treatment and between the beginning and the end of the experiment with model
+formulation `diversity ~ treatment * passage`. We also tested a difference in
+beta-diversity as both Bray-Curtis distance and Jaccard distance between the
+Positive and Neutral treatment and the beginning and end of the experiment with
+model formulation `diversity ~ treatment * passage`.
+
+To identify taxa that responded to selection on CH~4~ oxidation rate, we tested
+differential abundance using a beta-binomial model [@martin2020]. We compared
+taxa in passage 5 between the positive and neutral treatment to identify taxa
+that were enriched or depleted as a response to selection. We also tested
+differential abundance with differences in CH~4~ oxidation rate to identify taxa
+correlated with CH~4~ flux.
+
+Possible tests:
+
+- Random forest or gradient boosting to predict Positive vs. Neutral.
+
+Test if random forest or gradient boosting can predict whether this is the positive line or the neutral line based solely on the microbial community?
+
+# Results
+
+## Methane oxidation
+
+## Response to selection
+
+We observed a response to selection on soil methane oxidation rate at the whole
+ecosystem level (`r figures("response", display="cite")`). In addition, the
+response to selection varied with treatment 
+`r print_lrt(compare_response_mods, 2, 'LRT of nested models with and without treatment')`. At the start of the experiment, the Positive
+treatment had a mean methane oxidation rate that was
+`r percent_less(initial = tidy(response_model)[[1, 2]], final = tidy(response_model)[[3, 2]] + tidy(response_model)[[1, 2]])`% lower than the
+Neutral treatment `r print_lm(response_model, 3, 'difference of y-intercepts')`. 
+There was no change in methane oxidation rate of the Neutral
+treatment over the five passages `r print_lm(response_model, 2)`. By contrast,
+the Positive treatment had a `r round(per_slope)`% increase in methane
+oxidation rate per passage `r print_lm(response_model, 4)`.
+
+```{r response_figure, fig.width=4, fig.height=3, fig.cap=figures("response")}
+ratio_log10 <- calc_plot_ratios(fluxes$passage, fluxes$log_ch4)
+
+plot_flux(fluxes, passage, log_ch4, ratio_log10, log10 = TRUE)
+
+```
+
+## Ecosystem heritability
+
+We estimated heritability as the regression of mid-offspring on mid-parent.
+Offspring CH~4~ oxidation rates were correlated with parental CH~4~ oxidation
+rates in both the Positive treatment `r print_lm(herit_model, 4)` and the
+Neutral treatment `r print_lm(herit_model, 2)`. Notably, the sign of the effect
+was reversed with a positive heritability for the Positive treatment and a negative heritability for the Neutral treatment.
+
+```{r herit_fig, fig.width=5, fig.height=3, fig.cap=figures("herit")}
+plot_herit(heritability)
+```
+
+## Community Response
+
+Alpha diversity
+- treatment
+- flux
+
+Beta diversity
+- treatment
+- flux
+
+Differential abundance
+- treatment
+- flux
+
+# Discussion
+
+We observed a response to selection on soil CH~4~ oxidation rate of the whole soil microbial community.
+This type of response has been observed for other functions, such as chloroaniline degradation in water [@swenson2000], but is the first, to our knowledge, of selection being performed on soil biogeochemical cycling.
+
+A previous study near our sample site observed zero CH~4~ emissions from soils (cite laurel's paper).
+This suggests that little or no methanogenesis is happening in soils in this region and so the soil microbial communities are not "primed" to consume CH~4~.
+With no soil methanogenesis, the only type of methanotrophy happening would be high-affinity CH~4~ oxidation or consumption of atmospheric CH~4~, which is on the order of *some number* ppm CH~4~.
+By contrast, our incubations were at ~1000 ppm CH~4~.
+This is disparity is reflected in our results in that within the first passage we generated, only four of the 24 jars consumed more than 10% of the headspace CH~4~ within 48 hours and the biggest consumer only consumed 60% By contrast, in the final generation there were eight jars that consumed more than 10% and the top jar consumed over 95% of the headspace CH~4~.
+
+If this experiment had been run for longer, we might be expected to see crashes in methane oxidation as the process of random sampling at each generation
+
+Ecosystem heritability sounds confusing because it is difficult to imagine heritability without discrete generations.
+And how could a soil have "generations" (in the absence of coalescence events (cite Williams and Lenton))?
+However, this definition of heritability is conflated with the colloquial use of the term as referring to *inheritance*.
+But looking at the definition of narrow-sense heritability, it is simply the additive genetic variation associated with a trait [@falconer1996].
+In this case, additive genetic effect would refer to the identity and relative abundance of microbial taxa in the metagenome.
+This notably excludes any interactions among taxa and interactions with the environment (i.e., GxE effects or context-dependent effects).
+
+The heritability we observed was quite large at h^2^ = `r myround(tidy(herit_model)[4, 2], 2)`.
+However, given the small sample size and and wide confidence intervals (95% CI: `r myround(confint(herit_model)[4, ], 2)`) it is likely that a larger study could more precisely estimate the true heritability.
+Given these data, though, it would seem that genetic variation within the soil microbial community has a very strong effect on soil CH~4~ oxidation rates independent of any environmental variation.
+Granted, we intentionally eliminated much of the environmental variation that would be present in natural ecosystems.
+
+The negative heritability of CH~4~ oxidation rate in the Neutral selection
+treatment may result from selection on alternative characters in those
+populations.
+
+## Future Directions
+
+This study focused on the practical implications of using AES as a tool to deduce how microbes map onto functions at the whole-ecosystem level.
+If one were interested in the evolutionary dynamics underlying this process, one could design a study to distinguish between individual selection on microbial strains and group selection or multilevel selection.
+That is, selection that cannot be recapitulated through selection at the individual level (Williams and Lenton 2007).
+This could be accomplished by culturing differentially abundant taxa in the positive selection treatment and then inoculating them in isolation (or in a non-selected background community) to see if the same level of ecosystem function could be recapitulated as in the final generation of the positive selection treatment.
+
+Another avenue for research could be to look at correlated responses of ecosystem-level traits to selection.
+This would address questions such as, "Is methane oxidation rate correlated with other ecosystem processes due to pleiotropy or linkage disequilibrium?" At the community level, pleiotropy might be defined as "the production by a single taxon of two or more unrelated effects." Further, linkage disequilbrium would represent correlated traits due to taxa who tend to co-associate due to shared environmental preferences or coupled metabolic pathways.
+
+These experiments would further help us understand the nature of the "map" between community structure and ecosystem function.
+This would not only reveal how best to predict ecosystem function in earth system models, but also demonstrate the mechanism underlying that relationship.
+This is akin to deducing the genetic control of trait variation in molecular biology.
+
+Community epistasis is analogous to functional redundancy.
+
+# Acknowledgments
+This project was supported by the National Science Foundation Graduate Research Fellowship Program (grant no. DGE 1255832) and the ARCS Foundation Florence and Mike Nudelman Scholarship.
+
+# Competing Interests
+We declare we have no competing interests.
+
+# Supplement
+
+First, we found an effect of treatment by comparing the full model to the reduced model `r print_lrt(compare_response_mods)`.
+
+# Junk
 
@@
 
-# Materials & Methods
-
-## Selection experiment
-We performed an artificial ecosystem selection experiment similar to Swenson et
-al. (2000) by passaging replicate soil microbial communities. The trait we
-selected on was CH~4~ oxidation rate. Our experiment had two selection lines
-with twelve jars each for a total of twenty-four jars per passage. One line was
-a positive selection line where the two or three jars with the highest CH~4~
-oxidation rate were chosen to inoculate the next set of positive jars. The
-other line was a neutral selection line where an equal number of jars as the
-positive line were chosen at random to inoculate the next set of neutral jars.
-The number of jars chosen was based on the distribution of fluxes among the
-positive jars, i.e., we chose the top three jars unless only two jars had
-considerably greater oxidation rates based on visual inspection of histograms.
-The experiment was carried out over five passages.
-
-The initial soil microbial community was sampled from the top 10 cm of an
-upland mineral soil under a deciduous forest ecosystem near the University of
-Oregon campus in Eugene, OR, USA. Incubations were performed in 500 mL mason
-jars (Ball Corporation, Broomfield, Colorado, USA) with rubber septa installed
-in the lids. Each jar was sterilized with 70% ethanol to which was added 45 g
-of autoclaved potting mix, 5 g of living soil, and 3.5 mL of sterile deionized
-water to bring them to <mark> 60%??? of field capacity</mark>. The jars were
-then homogenized, capped, and injected with 4.3 mL of 99% CH~4~ to bring the
-headspace concentration to approximately 1000 ppm CH~4~. To create the two
-treatment groups, twenty-four jars were created in an identical manner and then
-randomly assigned to either the positive or neutral selection line. Jars were
-flushed and respiked twice per week to maintain elevated CH~4~ concentrations
-and were incubated at ambient temperature for approximately three weeks.
-Methane oxidation rates were determined at the end of the incubation period.
-These values were then used to determine the jars selected to inoculate the
-next set of jars.
-
-## Methane oxidation rate measurements
-
-Methane oxidation rates were determined after flushing and spiking jars to
-approximately 1000 ppm CH~4~. Headspace samples of 1 mL were collected from
-each jar immediately after spiking and then at time points 3, 6, 24, and 48
-hours for a 5-point curve. Samples were immediately injected into a
-<mark>???</mark> gas chromatograph fitted with an electron capture detector
-(company, city, state, USA) to determine the headspace CH~4~ concentration.
-Fluxes were calculated from a first-order exponential decay function as decay
-constant `k` with units day^-1^. Oxidation rates are presented as the additive
-inverse of `k` (i.e., `-k`) so that a more positive value represents a faster
-oxidation rate. The jars selected for the positive treatment in passage 2 had
-the lowest methane oxidation rate of the twelve jars due to a calculation error
-in the methane oxidation rate. All other passages correctly used the top three jars.
-
-## Soil DNA extraction and sequencing
-
-A subsample of soil from the starting inoculum and from every jar in passages 2
-and 5 was collected and stored at -80$^\circ$C for later DNA extraction.
-Soil DNA was extracted from a 0.25 g subsample of soil. Negative controls were extracted
-from autoclaved potting mix and DNase-free water. Extractions were performed
-using the DNeasy PowerSoil kit (QIAGEN, Hilden, Germany) and quantified using
-Qubit (company, city, state). In order to estimate the diversity and relative
-abundance of the bacterial and archaeal taxa in our soil ecosystems, we
-sequenced the V3-V4 region of the 16S rRNA gene using the Earth Microbiome
-Project protocol and primers<mark>???</mark>. Sequencing was performed on the
-Illumina NovaSeq 6000 with paired-end 150 bp reads.
-
-## Bioinformatics
-
-Bioinformatic analyses were conducted using the `QIIME2`
-bioinformatics platform (Bolyen et al. 2019). Raw sequencing reads were
-demultiplexed using ??? (citation), denoised using `DADA2` (citation).
-Taxonomic assignment for the 16S reads was performed using ??? based on the
-Silva database. All subsequent tatistical analyses were performed in the `R`
-statistical programming environment (R Core Team 2020). Richness estimates were
-made using the `breakaway` package and diversity indices using the `DivNet`
-package (Willis et al. ???).
-
-## Statistical Analysis
-
-### Response to selection
-
-To test whether there was a significant change in CH~4~ oxidation rate (k) as a
-response to selection, we tested a difference in slopes between the positive
-and neutral selection lines. The CH~4~ oxidation rates were strongly right
-skewed with most values close to zero and few large, positive values. This
-resulted in residuals that did not meet the assumptions of constant variance
-and normal distribution. Therefore, methane oxidation rates were log~10~
-transformed to better meet the assumptions of a linear model and to make
-figures easier to interpret. First, we tested the effect of treatment by
-fitting two nested models with and without treatment using `lm` in `R`. The
-model formulation for the full model was `flux ~ passage * treatment` and the
-model formulation for the reduced model was `flux ~ passage`. We compared these
-models using the likelihood ratio test with the `anova` function in `R`. We
-then fit the full model to determine the slope of the positive line, which
-represents the change in methane oxidation rate per passage as a response to
-selection.
-
-### Ecosystem heritability
-
-To estimate the proportion of variance in ecosystem function due to additive
-genetic variance of the microbial communtiy, we applied the Breeder's equation
-to the results of our selection experiment. the heritability of
-CH~4~ oxidation rate as an ecosystem property, we calculated narrow-sense
-heritability ($h^2$) as the regression of offspring on mid-parent (Goodnight
-2000, Falconer 1997). The mid-parent was the mean for all three selected jars
-and the offspring was the mean for all jars produced by those parents.
-
-In addition, we calculated the inbreeding coefficient, $F$, as:
-
-$$
-F = 1 - \left(1 - \frac{1}{2N} \right) ^{T}
-$$
-
-where $N$ is the population size and $T$ is the number of generations (Goodnight 2000 or Falconer 1997).
-
-### Taxa differentially abundant in the positive selection treatment
-
-To identify taxa that responded to selection on CH~4~ oxidation rate, we tested
-differential abundance using a beta-binomial model (Martin et al. 2020). We
-compared taxa in passage 5 between the positive and neutral treatment to
-identify taxa that
-
-Possible tests:
-
-- Random forest or gradient boosting to predict Positive vs. Neutral.
-
-# Results
-
-## Response to selection
-
-We observed a response to selection on soil methane oxidation rate at the whole
-ecosystem level (`r figures("response", display="cite")`).
-At the start of the experiment, the Positive
-treatment had a mean methane oxidation rate that was
-`r percent_less(initial = tidy(response_model)[[1, 2]], final = tidy(response_model)[[3, 2]] + tidy(response_model)[[1, 2]])`%
-lower than the
-Neutral treatment (SE = 
-`r myround(tidy(response_model)[[3, 3]], 2)`, t = 
-`r myround(tidy(response_model)[[3, 4]], 2)`, p = 
-`r myround(tidy(response_model)[[3, 5]], 2)`). There was no change in methane oxidation rate of the Neutral
-treatment over the five passages (slope = 
-`r myround(tidy(response_model)[[2, 2]], 2)`, SE = 
-`r myround(tidy(response_model)[[2, 3]], 2)`, t = 
-`r myround(tidy(response_model)[[2, 4]], 2)`, p = 
-`r myround(tidy(response_model)[[2, 5]], 2)`). By contrast, the Positive treatment had a
-`r round(per_slope)`% increase in methane
-oxidation rate per passage (slope = 
-`r myround(tidy(response_model)[[4, 2]], 2)`, SE = 
-`r myround(tidy(response_model)[[4, 3]], 2)`, t =
-`r myround(tidy(response_model)[[4, 4]], 2)`, p = 
-`r myround(tidy(response_model)[[4, 5]], 2)`).
-
-```{r response_figure, fig.width=4, fig.height=3, fig.cap=figures("response")}
-ratio_log10 <- calc_plot_ratios(fluxes$passage, fluxes$log_ch4)
-
-plot_flux(fluxes, passage, log_ch4, ratio_log10, log10 = TRUE)
-
-```
-
-## Ecosystem heritability
-
-Here, I fit two models. The first is simply offspring flux on selected parent
-flux, ignoring treatment. There is no correlation. Next, I added treatment and the interaction between treatment and parental
-flux. Here, there is a significant correlation between parent and offspring for
-both the Neutral treatment `r print_model(herit_model, 2)` and the Positive
-treatment `r print_model(herit_model, 4)`. Notably, the sign of the effect is
-reversed: negative for the Neutral treatment and positive for the Positive
-treatment.
-
-The negative heritability of methane oxidation rate in the Neutral selection
-treatment may result from selection on alternative characters in those
-populations. The strong, positive heritability in the Positive treatment ... .
-It seems strange that the Methane oxidation rate was heritable between
-consecutive passages (`r figures("herit", display="cite")`). Specifically, the
-narrow-sense heritability, calculated as the slope of the mid-parent and
-mid-offspring, was $h^2$ =  $\pm$  (t = , p = ). With an inbreeding coefficient of ,
-
-```{r herit_fig, fig.width=5, fig.height=3, fig.cap=figures("herit")}
-plot_herit(heritability)
-```
-
-# Discussion
-
-## Future Directions
-
-This study focused on the practical implications of using AES as a tool to deduce how microbes map onto functions at the whole-ecosystem level.
-If one were interested in the evolutionary dynamics underlying this process, one could design a study to distinguish between individual selection on microbial strains and group selection or multilevel selection.
-That is, selection that cannot be recapitulated through selection at the individual level (Williams and Lenton 2007).
-This could be accomplished by culturing differentially abundant taxa in the positive selection treatment and then inoculating them in isolation (or in a non-selected background community) to see if the same level of ecosystem function could be recapitulated as in the final generation of the positive selection treatment.
-
-Another avenue for research could be to look at correlated responses of ecosystem-level traits to selection.
-This would address questions such as, "Is methane oxidation rate correlated with other ecosystem processes due to pleiotropy or linkage disequilibrium?" At the community level, pleiotropy might be defined as "the production by a single taxon of two or more unrelated effects." Further, linkage disequilbrium would represent correlated traits due to taxa who tend to co-associate due to shared environmental preferences or coupled metabolic pathways.
-
-These experiments would further help lus understand the nature of the "map" between community structure and ecosystem function.
-This would not only reveal how best to predict ecosystem function in earth system models, but also demonstrate the mechanism underlying that relationship.
-This is akin to deducing the genetic control of trait variation in molecular biology.
-
-Community epistasis is analogous to functional redundancy.
-
-# Acknowledgments
-This project was supported by the National Science Foundation Graduate Research Fellowship Program (grant no. DGE 1255832) and the ARCS Foundation Florence and Mike Nudelman Scholarship.
-
-# Competing Interests
-We declare we have no competing interests.
-
 # References

### 2021-08-17 R/01-process-flux.Rmd
update manuscript and analyses and figures plus add citations

@@
 
+```{r remove_p6}
+conc_data <- 
+  conc_data %>% 
+  filter(flux_date != '2020-07-27')
+time_data <- 
+  time_data %>% 
+  filter(flux_date != '2020-07-27')
+
+```
+
 ```{r compute_time_intervals}
@@
 
+```{r plot_fractions}
+
+conc_data %>% 
+  select(flux_date, jar, t1, t4) %>% 
+  mutate(frac_cons = t4/t1) %>% 
+  mutate(treat = substring(jar,1,1)) %>% 
+  ggplot(aes(x = flux_date, y = frac_cons, color = treat)) +
+  geom_jitter(width = 0.2) +
+  geom_hline(yintercept = 0.90) +
+  geom_hline(yintercept = 0.05)
+
+```
+
 Finally, this step takes the time intervals and ch4 concentrations and fits a first-order exponential decay function to compute methane flux as `k` which is the rate-constant for the first-order exponential decay function.
@@
 
-#flux_data %>% 
-#  ggplot(aes(days, ppm)) + 
-#  facet_grid(flux_date + treat ~ rep, scales = "free") +
-#  geom_point() + 
-#  stat_smooth(method = 'lm', formula = y ~ x)
-
 lin_model <- function(dt) {
@@
          log_ch4 = log10(ch4 + abs(min(ch4)) + 0.001)) %>% 
-  filter(passage != 6) %>% 
   select(-std.error, -statistic, -p.value, -estimate)
+```
+
+```{r plot_fluxes, fig.height=50}
+flux_data %>%
+ ggplot(aes(days, ppm)) +
+ facet_grid(passage + treat ~ rep, scales = "free") +
+ geom_point() +
+ stat_smooth(method = 'lm', formula = y ~ x)
 ```

### 2021-08-17 R/03-response.Rmd
update manuscript and analyses and figures plus add citations

@@
 # Compare interaction to no interaction
-compare_mods <- anova(response_model, no_treat_model)
-kable(tidy(compare_mods))
+compare_response_mods <- anova(no_treat_model, response_model, test = 'LRT')
+kable(tidy(compare_response_mods))
 ```
@@
 ```{r save_model}
-save(response_model, per_slope, file = '../Output/response_model.Rdata')
+save(response_model, compare_response_mods, per_slope, file = '../Output/response_model.Rdata')
 ```

### 2021-08-17 R/04-heritability.Rmd
update manuscript and analyses and figures plus add citations

@@
 The first is simply offspring flux on selected parent flux, ignoring treatment.
-There is no correlation `r print_model(fit_no_treat, 2)`.
+There is no correlation `r print_lm(fit_no_treat, 2)`.
 Next, I added treatment and the interaction between treatment and parental flux.
-Here, there is a significant correlation between parent and offspring for both the Neutral treatment `r print_model(herit_model, 2)` and the Positive treatment `r print_model(herit_model, 4)`.
+Here, there is a significant correlation between parent and offspring for both the Neutral treatment `r print_lm(herit_model, 2)` and the Positive treatment `r print_lm(herit_model, 4)`.
 Notably, the sign of the effect is reversed: negative for the Neutral treatment and positive for the Positive treatment.

### 2021-08-17 R/functions.R

@@ -31,11 +31,17 @@ calc_plot_ratios <- function(x, y) {
 }
 
 
-print_model <- function(fit, term = 1) {
-  paste0('(slope = ', myround(tidy(fit)[term, 2], 2), 
+print_lm <- function(fit, term = 1, estimate = 'slope') {
+  paste0('(', estimate, ' = ', myround(tidy(fit)[term, 2], 2), 
         ', SE = ', myround(tidy(fit)[term, 3], 2), 
         ', t = ', myround(tidy(fit)[term, 4], 2), 
         ', p = ', myround(tidy(fit)[term, 5], 2), ')')
+}
+
+print_lrt <- function(fit, term = 2, test = "Likelihood ratio test") {
+  paste0('(', test, ': df = ', round(tidy(fit)[term, 3]), 
+         ', ss = ', myround(tidy(fit)[term, 4], 2), 
+         ', p = ', myround(tidy(fit)[term, 5], 2), ')')
 }
 
 
@@ -46,7 +52,7 @@ plot_herit <- function(heritability, ratio) {
     geom_jitter(width = 0.01) + 
     stat_smooth(method = 'lm', se = FALSE, formula = 'y ~ x') +
     labs(x = "Mid-parent (mean(log10(Flux (-k))))", 
-         y = "Offspring (log10(Flux (-k)))") +
+         y = "Mid-offspring (log10(Flux (-k)))") +
     scale_color_manual(name = "Treatment", 
                        labels = c('Neutral', 'Positive'), 
                        values = c('gray40', 'darkorange2')) +

### 2022-04-07 R/03-response.Rmd
clean up R directory

@@
 
-library(metafor)
 # Standard error of the difference

### 2022-04-07 R/07-generate_physeq.Rmd
clean up R directory

---
title: "Sequence data processing"
author: "Andrew Morris"
date: "9/13/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
knitr::opts_knit$set(root.dir=normalizePath('../'))
```

```{r initialize_packages}
library(tidyverse)
library(phyloseq)
library(decontam)
library(microViz)
library(knitr)
```

```{r import_data}
load(file = "Output/dada2_output.RData")

samples <- read_tsv('Output/barcode_master.tsv')
fluxes <- read_tsv('Output/fluxes.tsv')
qubit <- read_tsv('Data/post_pcr_qubit.tsv')

```

```{r setup_phyloseq}
samples <-
  samples %>% 
  left_join(qubit, by = c("pcr_well" = "well"))

samples <- as.data.frame(samples)
rownames(samples) <- samples$sample_id
rownames(seqtab.nochim) <- samples$sample_id
samples$treatment <- substr(samples$sample_id, 1, 2)
samples$treatment[59] <- samples$sample_id[59]
samples$treatment[60] <- samples$sample_id[60] 
samples$col <- as.character(samples$col)
sample_subset <- 
  as.data.frame(select(samples, sample_id, row, col, treatment, qubit_ngml))
rownames(sample_subset) <- samples$sample_id

ps <- phyloseq(otu_table(seqtab.nochim, taxa_are_rows=FALSE), 
               sample_data(sample_subset),
               tax_table(taxa))

dna <- Biostrings::DNAStringSet(taxa_names(ps))
names(dna) <- taxa_names(ps)
ps <- merge_phyloseq(ps, dna)
taxa_names(ps) <- paste0("ASV", seq(ntaxa(ps)))

```

# Alpha Diversity for All Data

Here is alpha diversity for each treatment, positive/negative control, etc.
In addition, there is alpha diversity by row and column of the PCR plate. I don't see any relationship between PCR plate location and alpha diversity, which is good. It looks like the PCR positive control, which was a pure culture of E. Coli, has particularly low diversity, which is good. Let's pull that sample out and look at it.

```{r plot_richness_all_data}

plot_richness(ps, x = "treatment", color="treatment", measures=c("Observed", "Shannon", "Simpson"), title = "Alpha diversity by treatment and blanks")
plot_richness(ps, x = "row", measures=c("Observed", "Shannon", "Simpson"), title = "Alpha diversity by PCR plate row")
plot_richness(ps, x = "col", measures=c("Observed", "Shannon", "Simpson"), title = "Alpha diversity by PCR plate col")

```

## Specific Controls

```{r}
ps_pm <-
ps %>% 
  ps_filter(treatment == "PM")

ps_pcrp_rich <- estimate_richness(ps_pm, measures="Observed")

top20 <- names(sort(taxa_sums(ps_pm), decreasing=TRUE))[1:20]
ps_pm.top20 <- transform_sample_counts(ps_pm, function(OTU) OTU/sum(OTU))
ps_pm.top20 <- prune_taxa(top20, ps_pm.top20)
plot_bar(ps_pm.top20, x="sample_id", fill="Family") + facet_wrap(~treatment, scales="free_x")
tax_table(ps_pm.top20) %>% 
  kable(caption = "Top 20 taxa from autoclaved potting mix")

```

```{r}
ps_nc <-
ps %>% 
  ps_filter(treatment == "NC")

ps_pcrp_rich <- estimate_richness(ps_nc, measures="Observed")

top20 <- names(sort(taxa_sums(ps_nc), decreasing=TRUE))[1:20]
ps_nc.top20 <- transform_sample_counts(ps_nc, function(OTU) OTU/sum(OTU))
ps_nc.top20 <- prune_taxa(top20, ps_nc.top20)
plot_bar(ps_nc.top20, x="sample_id", fill="Family") + facet_wrap(~treatment, scales="free_x")
tax_table(ps_nc.top20) %>% 
  kable(caption = "Top 20 taxa from negative controls")

```

## PCR Positive Control: E. Coli

Here is a stacked bar chat and taxonomy of the top 20 OTUs from the positive control

```{r pcrp_top_20}

ps_pcrp <- ps_filter(ps, treatment == "PCRP")

ps_pcrp_rich <- estimate_richness(ps_pcrp, measures="Observed")

top20 <- names(sort(taxa_sums(ps_pcrp), decreasing=TRUE))[1:20]
ps_pcrp.top20 <- transform_sample_counts(ps_pcrp, function(OTU) OTU/sum(OTU))
ps_pcrp.top20 <- prune_taxa(top20, ps_pcrp.top20)
plot_bar(ps_pcrp.top20, x="sample_id", fill="Family") + facet_wrap(~treatment, scales="free_x")
tax_table(ps_pcrp.top20) %>% 
  kable(caption = "Top 20 taxa from positive control")

top1 <- names(sort(taxa_sums(ps_pcrp), decreasing=TRUE))[1]
ps_pcrp.top1 <- prune_taxa(top1, ps_pcrp)
tax_table(ps_pcrp.top1) %>% 
  kable(caption = "Top 1 taxon from positive control")

```

One Enterobacteriaceae is dominating the community. This is great: that OTU is E. Coli. However, there are `r ps_pcrp_rich[1,1]` OTUs in this sample. Looking at the top 20, many of them are also E coli., so it's probably just multiple 16S copies or variants. The rest are contaminants and we will hopefully remove these contaminates using the negative controls with DECONTAM in the next step. But first, let's remove the positive control.

# Alpha + Beta Diversity for all minus the PCR positive control

It looks like diversity decreases between Generation 2 and 5, which makes sense. The negative controls and PCR negative have close to zero ASVs, but some have closer to 500-1000. The potting mix is slightly higher at around 1000. The soil cores have the highest alpha diversity.

```{r ps_no_pcrp}

ps_no_PCRP <- ps_filter(ps, treatment != "PCRP")

ps <- ps_no_PCRP

plot_richness(ps, x = "treatment", color="treatment", measures=c("Observed","Shannon", "Simpson"))
ps.prop <- transform_sample_counts(ps, function(otu) otu/sum(otu))
ord.nmds.bray <- ordinate(ps.prop, method="NMDS", distance="bray")
plot_ordination(ps.prop, ord.nmds.bray, color="treatment", title="Bray NMDS")
plot_ordination(ps.prop, ord.nmds.bray, color="row", title="Bray NMDS")
plot_ordination(ps.prop, ord.nmds.bray, color="col", title="Bray NMDS")

```

## Top 20 most dominant ASVs

Interestingly, none of the most dominant ASVs are in the soil cores. The jars look quite different from the negative controls and the potting mix.

```{r ps_top20}

top20 <- names(sort(taxa_sums(ps), decreasing=TRUE))[1:20]
ps.top20 <- transform_sample_counts(ps, function(OTU) OTU/sum(OTU))
ps.top20 <- prune_taxa(top20, ps.top20)
plot_bar(ps.top20, x="sample_id", fill="Family") + facet_wrap(~treatment, scales="free_x") + labs(title = "Stacked bars of top 20 taxa in all samples (minus PCRP)")

ps_sc <- subset_samples(ps, treatment == "SC")

top20 <- names(sort(taxa_sums(ps_sc), decreasing=TRUE))[1:20]
ps_sc.top20 <- transform_sample_counts(ps_sc, function(OTU) OTU/sum(OTU))
ps_sc.top20 <- prune_taxa(top20, ps_sc.top20)
plot_bar(ps_sc.top20, x="sample_id", fill="Family") + facet_wrap(~treatment, scales="free_x") + labs(title = "Top 20 taxa from the soil cores")

tax_table(ps_sc.top20) %>% 
  kable(caption = "Top 20 taxa from soil cores")

ps_nc <- ps_filter(ps, treatment == "NC")

top20 <- names(sort(taxa_sums(ps_nc), decreasing=TRUE))[1:20]
ps_nc.top20 <- transform_sample_counts(ps_nc, function(OTU) OTU/sum(OTU))
ps_nc.top20 <- prune_taxa(top20, ps_nc.top20)
plot_bar(ps_nc.top20, x="sample_id", fill="Family") + facet_wrap(~treatment, scales="free_x") + labs(title = "Top 20 taxa from the negative controls")

tax_table(ps_nc.top20) %>% 
  kable(caption = "Top 20 taxa from negative controls.")

```

# DECONTAM

```{r}

ps_decontam <- ps_filter(ps, !treatment  %in% c('PM', 'SC', 'PCRN'))
ps_decontam <-
ps_decontam %>% 
  ps_select(sample_id:treatment, quant_reading = qubit_ngml) %>% 
       ps_mutate(Sample_or_Control = case_when(treatment == "NC" ~ "Control", 
                             TRUE ~ "Sample"))

df <- as.data.frame(sample_data(ps_decontam)) # Put sample_data into a ggplot-friendly data.frame
df$LibrarySize <- sample_sums(ps_decontam)
df <- df[order(df$LibrarySize),]
df$Index <- seq(nrow(df))
ggplot(data=df, aes(x=Index, y=LibrarySize, color=Sample_or_Control)) + geom_point()

```

My negative controls have a lower library size than my samples.

## Decontam Prevalance

First I'll run decontam's prevalence approach, which identifies contaminates based on which ASVs are more prevalent in negative controls than positive controls. It doesn't identify any that are more prevalent in the negatives.

```{r decontam_prevalence}
sample_data(ps_decontam)$is.neg <- sample_data(ps_decontam)$Sample_or_Control == "Control Sample"
contamdf.prev <- isContaminant(ps_decontam, method="prevalence", neg="is.neg")
table(contamdf.prev$contaminant)

contamdf.prev05 <- isContaminant(ps_decontam, method="prevalence", neg="is.neg", threshold=0.5)
table(contamdf.prev05$contaminant)

# Make phyloseq object of presence-absence in negative controls and true samples
ps.pa <- transform_sample_counts(ps_decontam, function(abund) 1*(abund>0))
ps.pa.neg <- prune_samples(sample_data(ps.pa)$Sample_or_Control == "Control", ps.pa)
ps.pa.pos <- prune_samples(sample_data(ps.pa)$Sample_or_Control == "Sample", ps.pa)
# Make data.frame of prevalence in positive and negative samples
df.pa <- data.frame(pa.pos=taxa_sums(ps.pa.pos), pa.neg=taxa_sums(ps.pa.neg),
                      contaminant=contamdf.prev$contaminant)
ggplot(data=df.pa, aes(x=pa.neg, y=pa.pos, color=contaminant)) + geom_point() +
  xlab("Prevalence (Negative Controls)") + ylab("Prevalence (True Samples)")

```

## Decontam Frequency

Next I'll try decontam's frequency approach which identifies contaminants based on which taxa are present at higher frequency in low concentration samples. i.e., their abundance is inversely proportional to sample DNA concentration.

```{r decontam_frequency}
ps_decontam <-
  ps_decontam %>% 
  ps_filter(quant_reading > 0)

contamdf.freq <- isContaminant(ps_decontam, method="frequency", conc="quant_reading")
head(contamdf.freq)
table(contamdf.freq$contaminant)
head(which(contamdf.freq$contaminant))
plot_frequency(ps_decontam, taxa_names(ps_decontam)[c(1, 184)], conc="quant_reading") + 
  xlab("DNA Concentration (Qubit)")

set.seed(99)
plot_frequency(ps_decontam, taxa_names(ps_decontam)[sample(which(contamdf.freq$contaminant),3)], conc="quant_reading") +
    xlab("DNA Concentration (Qubit)")

```

There were `r table(contamdf.freq$contaminant)[2]` contaminants identified by DECONTAM. However, I'm missing Qubit values for 3 negative controls, which were below detection limit. And looking at the plots of Frequency ~ Qubit, most contaminants are driven by just the two remaining negative controls. It's not clear that their abundance is correlated with DNA concentration. Therefore, I don't feel comfortable removing them as contaminants.

I will try rerunning decontam with just the samples (no controls).

```{r decontam_frequency_no_controls}

ps_decontam <-
  ps_decontam %>% 
  ps_filter(treatment != "NC")

contamdf.freq <- isContaminant(ps_decontam, method="frequency", conc="quant_reading")
head(contamdf.freq)
table(contamdf.freq$contaminant)
head(which(contamdf.freq$contaminant))
plot_frequency(ps_decontam, taxa_names(ps_decontam)[c(1, 2216, 2882, 3057)], conc="quant_reading") + 
  xlab("DNA Concentration (Qubit)")

set.seed(99)
plot_frequency(ps_decontam, taxa_names(ps_decontam)[sample(which(contamdf.freq$contaminant),3)], conc="quant_reading") +
    xlab("DNA Concentration (Qubit)")

```

This time there were only `r table(contamdf.freq$contaminant)[2]` contaminants identified. However, they're all such low prevalence (present in maybe 3-5 samples based on the handful I plotted) that it's hard to draw any conclusions. To build on this, I'll remove low prevalence ASVs - that is, ASVs present in <10% of samples.

```{r decontam_frequency_prevalence}
ps_decontam_prev <- tax_filter(ps_decontam, min_prevalence = 0.10)

contamdf.freq <- isContaminant(ps_decontam_prev, method="frequency", conc="quant_reading")
head(contamdf.freq)
table(contamdf.freq$contaminant)
head(which(contamdf.freq$contaminant))
plot_frequency(ps_decontam_prev, taxa_names(ps_decontam_prev)[c(1, 2856, 4687, 4780, 5093)], conc="quant_reading") + 
  xlab("DNA Concentration (Qubit)")

taxa_names(ps_decontam_prev)[head(which(contamdf.freq$contaminant))]

corncob::otu_to_taxonomy(taxa_names(ps_decontam_prev)[head(which(contamdf.freq$contaminant))], data = ps_decontam_prev)

```

After removing low-prevalence ASVs, only `r table(contamdf.freq$contaminant)[2]` were identified as contaminants.

Based on the above analyses, I have decided not to remove any contaminants.

# Create final phyloseq

```{r create_physeq}

flux_sd <- 
  fluxes %>% 
  mutate(sample_id = paste0('G', passage, toupper(treat), rep)) %>% 
  select(sample_id, flux_date:selected) %>% 
  mutate(passage = as.character(passage))

samples_filtered <- 
filter(samples, sample_id %in% flux_sd$sample_id)

fluxes_filtered <- 
  filter(flux_sd, sample_id %in% samples$sample_id) %>% 
  as.data.frame()

rownames(fluxes_filtered) <- fluxes_filtered$sample_id

sample_data(ps_decontam) <- sample_data(fluxes_filtered)

ps <- ps_decontam

save(ps, file = 'Output/phyloseq.RData')

# 
# 
# seqtab.nochim.filtered <-
#   seqtab.nochim[rownames(seqtab.nochim) %in% rownames(fluxes_filtered),]
# 
# ps <- phyloseq(otu_table(seqtab.nochim.filtered, taxa_are_rows=FALSE), 
#                                      sample_data(fluxes_filtered),
# 
#                tax_table(taxa))
# 
# 
# 
# dna <- Biostrings::DNAStringSet(taxa_names(ps))
# names(dna) <- taxa_names(ps)
# ps <- merge_phyloseq(ps, dna)
# taxa_names(ps) <- paste0("ASV", seq(ntaxa(ps)))

```

### 2022-04-07 R/08-community.Rmd
clean up R directory

---
title: "Community Analysis"
author: "Andrew Morris"
date: "9/15/2021"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
knitr::opts_knit$set(root.dir=normalizePath('../'))
```

```{r initialize_packages}
library(tidyverse)
library(phyloseq)
library(decontam)
library(microViz)
library(knitr)
library(corncob)
library(DivNet)
library(magrittr)
library(vegan)
library(ggpubr)
library(SpiecEasi)
library(ggthemes)
library(broom)

source('R/functions.R')
```

```{r ggplot-settings}
```

```{r import-data}
load(file = "Output/phyloseq.RData")
ps <-
  ps %>% 
  ps_mutate(pas_by_treat = paste0(passage, treat))
ps_melt <-
  ps %>% 
  ps_melt()
```

# Exploratory

```{r plot-depth}

depth_data <-
  ps %>% 
  ps_melt %>% 
  group_by(Sample, passage, treat) %>% 
  summarize(depth = sum(Abundance))

summary(aov(depth ~ passage * treat, data = depth_data))
depth_data %>% 
  ggplot(aes(x = interaction(treat, passage), y = depth)) +
  theme_tufte() +
  geom_rangeframe() +
  # coord_cartesian(clip="off") +
  geom_tufteboxplot()

```
There is no systematic difference in sequencing depth/library size between treatments. It does appear that sequencing depth decreases between passage 2 and 5.

```{r plot-richness}

plot_richness(ps, x = "pas_by_treat", color="treat", measures=c("Observed", "Shannon", "Simpson"), title = "Alpha Diversity ASV") +
    scale_color_manual(name = "Treatment", 
                         labels = c('Neutral', 'Positive'),
                         values = c('gray40', 'darkorange2'))

```

Richness decreases with passage, but seems unchanged between 5n and 5p.
Initial richness appears similar between the treatments.

```{r stacked_bar, eval=FALSE}
level = "Family"
ps_phylum <- speedyseq::tax_glom(ps, taxrank=level)
top10 = names(sort(taxa_sums(ps_phylum), TRUE)[1:10])
ps_phylum10 <- prune_taxa(top10, ps_phylum)
ps_phylum10 %>% 
  ps_filter(passage == 5) %>% 
  speedyseq::plot_bar(facet_grid = 'treat', fill = level) + 
  geom_bar(aes_string( color=level, fill=level), stat="identity", position="stack") +
  geom_bar(aes(x = reorder(rep, rank_ch4)))

```

# Hypothesis Testing

## Richness

```{r, eval=FALSE}
richness_soil <- ps %>% breakaway
plot(richness_soil, physeq=ps, color="passage")
summary(richness_soil) %>% as_tibble

meta <- ps %>%
  sample_data %>%
  as_tibble %>%
  mutate("sample_names" = ps %>% sample_names )

combined_richness <- meta %>%
  left_join(summary(richness_soil),
            by = "sample_names")

bt_passage <- betta(chats = combined_richness$estimate,
                      ses = combined_richness$error,
                      X = model.matrix(~passage, data = combined_richness))
bt_passage$table

save(bt_passage, file = 'Output/richness_model_passage.Rdata')

ps_5 <- ps_filter(ps, passage == "5")
richness_soil <- ps_5 %>% breakaway
plot(richness_soil, physeq=ps_5, color="treat")
summary(richness_soil) %>% as_tibble

meta <- ps_5 %>%
  sample_data %>%
  as_tibble %>%
  mutate("sample_names" = ps_5 %>% sample_names )

combined_richness <- meta %>%
  left_join(summary(richness_soil),
            by = "sample_names")

bt_treat5 <- betta(chats = combined_richness$estimate,
                      ses = combined_richness$error,
                      X = model.matrix(~treat, data = combined_richness))
bt_treat5$table

save(bt_treat5, file = 'Output/richness_model_treat5.Rdata')

bt_ch4 <- betta(chats = combined_richness$estimate,
                      ses = combined_richness$error,
                      X = model.matrix(~ch4, data = combined_richness))
bt_ch4$table

save(bt_ch4, file = 'Output/richness_model_ch4.Rdata')

```

Test confirms the above observation: Richness decreases between Passage 2
and Passage 5. No difference in richness by treatment in passage 5 or by
methane.

# Beta Diversity

```{r}

ps_fixed <-
ps %>%
 tax_fix(
  min_length = 4,
  unknowns = c(""),
  sep = " ", anon_unique = TRUE,
  suffix_rank = "classified"
 )

ps_fixed <- 
  ps_fixed %>% 
  ps_mutate(treat_bin = factor(ifelse(treat=='p', 1,0))) %>% 
  ps_mutate(passage_bin = factor(ifelse(passage==5, 1,0))) %>% 
  ps_mutate(treat_pas = interaction(treat_bin, passage_bin))

ps_fixed
```

## Beta diversity test passage by treatment

```{r}
plot_beta(ps_fixed)
```

```{r calc-beta, results=FALSE}

ord_calc_aitch <-
  ps_fixed %>% 
  tax_transform(trans = 'clr') %>% 
  ord_calc(method = "PCA")

ord_calc_bray <-
  ps_fixed %>%
  dist_calc("bray") %>%
  ord_calc(method = "NMDS")

ord_calc_jacc <-
  ps_fixed %>%
  dist_calc("jaccard", binary = TRUE) %>%
  ord_calc(method = "NMDS")

```

```{r plot-beta, results=FALSE, fig.width = 6, fig.height = 2}
bray <-
  ord_calc_bray %>% 
  ord_plot(color = "treat") +
  theme_tufte() +
  geom_rangeframe() +
  coord_cartesian(clip="off") +
  scale_color_manual(name = "Treatment", 
                       labels = c('Neutral', 'Positive'),
                       values = c('gray40', 'darkorange2'))

jacc <- 
  ord_calc_jacc %>% 
  ord_plot(color = "treat") +
  theme_tufte() +
  geom_rangeframe() +
  coord_cartesian(clip="off") +
  scale_color_manual(name = "Treatment", 
                       labels = c('Neutral', 'Positive'),
                       values = c('gray40', 'darkorange2'))

ggarrange(bray, jacc)

```

```{r plot-aitch, fig.width = 4, fig.height = 4}

ord_calc_aitch %>% 
  ord_plot(color = "treat",
           auto_caption = FALSE) +
  theme_tufte() +
  geom_rangeframe() +
  coord_fixed(5.5/43.6, clip="off") +
  scale_color_manual(name = "Treatment", 
                       labels = c('Neutral', 'Positive'),
                       values = c('gray40', 'darkorange2'))

```

```{r permanova-aitch, eval=FALSE}
dist_aitch <-
  ps %>% 
  dist_calc(dist = "aitchison")

dist_permanova(dist_aitch, c("passage", "treat"), "passage * treat")

passage_groups <- factor(c(rep(1,22), rep(2,24)), labels = c("2","5"))

dist_aitch %>% 
  dist_get %>% 
  vegan::betadisper(passage_groups) %>% 
  permutest()

treat_groups <- factor(c(rep(1,12), rep(2,10), rep(1,12), rep(2,12)), labels = c("n","p"))

dist_aitch %>% 
  dist_get %>% 
  vegan::betadisper(treat_groups) %>% 
  permutest()

ps %>%
  ps_filter(passage == 2) %>% 
  dist_calc(dist = "aitchison") %>% 
  dist_permanova("treat")

ps %>%
  ps_filter(passage == 5) %>% 
  dist_calc(dist = "aitchison") %>% 
  dist_permanova("treat")

adonis2(dist_get(dist_aitch) ~ passage * treat, data = as(sample_data(ps_fixed), 'data.frame'))

dist_aitch2 <-
  ps %>% 
  ps_filter(passage == 2) %>% 
  dist_calc(dist = "aitchison")
dist_permanova(dist_aitch2, "treat")
groups <- factor(c(rep(1,12), rep(2,10)), labels = c("N","P"))

dist_aitch2 %>% 
  dist_get %>% 
  vegan::betadisper(groups) %>% 
  permutest()
dist_aitch5 <-
  ps %>% 
  ps_filter(passage == 5) %>% 
  dist_calc(dist = "aitchison")
dist_permanova(dist_aitch5, "treat")
groups <- factor(c(rep(1,12), rep(2,12)), labels = c("N","P"))

dist_aitch5 %>% 
  dist_get %>% 
  vegan::betadisper(groups) %>% 
  permutest()

```

There is a significant main effect and interaction of passage and treatment on
taxonomic dissimilarity of the soil microbiome. In addition, I reject the null
hypothesis of homogeneity of group dispersions for passage, but not for treatment.
Inspecting the NMDS plot, from this I conclude that passaging between Passage 2
and 5 both altered the soil microbiome taxonomic composition and reduced its 
variability. In addition, there is a difference in centroid between the two
treatments in both Passage 2 and Passage 5, but no difference in dispersion.

```{r beta-test}
# ps_prop <- transform_sample_counts(ps, function(otu) otu/sum(otu))
# 
# fit_bray <- adonis2(otu_table(ps_prop) ~ passage * treat, data = as(sample_data(ps_fixed), 'data.frame'), method = "bray")
# 
# fit_bray
# 
# fit_jacc <- adonis2(otu_table(ps_prop) ~ passage * treat, data = as(sample_data(ps_fixed), 'data.frame'), method = "jacc", binary = TRUE)
# 
# fit_jacc
ps_clr <- microbiome::transform(ps, "clr")

fit_bin_jacc <- adonis2(otu_table(ps) ~ passage * treat, data = as(sample_data(ps), 'data.frame'), method = "jaccard", binary = TRUE)

fit_aitch <- adonis2(otu_table(ps_clr) ~ passage * treat, data = as(sample_data(ps_clr), 'data.frame'), method = "euclidean")

fit_adonis <- rbind(tidy(fit_aitch), tidy(fit_bin_jacc))
fit_adonis$term <- c('Passage', 'Treatment', 'Pass:Treat', 'Residual', 'Total',
      'Passage', 'Treatment', 'Pass:Treat', 'Residual', 'Total')
save(fit_adonis, file = 'Output/beta_model.Rdata')

############

clr_dist_matrix <- phyloseq::distance(ps_clr, method = "euclidean")

dispr <- vegan::betadisper(clr_dist_matrix, phyloseq::sample_data(ps_clr)$passage)
disper_test <- permutest(dispr)

dispr <- vegan::betadisper(clr_dist_matrix, phyloseq::sample_data(ps_clr)$treat)
disper_test <- permutest(dispr)

dist_aitch5 <-
  ps %>% 
  ps_filter(passage == 5) %>% 
  dist_calc(dist = "aitchison")
dist_permanova(dist_aitch5, "treat")
groups <- factor(c(rep(1,12), rep(2,12)), labels = c("N","P"))

dist_aitch5 %>% 
  dist_get %>% 
  vegan::betadisper(groups) %>% 
  permutest()

dist_aitch2 <-
  ps %>% 
  ps_filter(passage == 2) %>% 
  dist_calc(dist = "aitchison")
dist_permanova(dist_aitch5, "treat")
groups <- factor(c(rep(1,12), rep(2,10)), labels = c("N","P"))

dist_aitch2 %>% 
  dist_get %>% 
  vegan::betadisper(groups) %>% 
  permutest()

###########3

# adonis2(otu_table(ps_filter(ps_prop, passage == 5)) ~ treat, 
#         data = 
#           as(
#             sample_data(
#               ps_filter(
#                 ps_prop, 
#                 passage == 5)
#             ), 
#             'data.frame'), 
#         method = "bray")
# adonis2(otu_table(ps_filter(ps_prop, passage == 5)) ~ treat, 
#         data = 
#           as(
#             sample_data(
#               ps_filter(
#                 ps_prop, 
#                 passage == 5)
#             ), 
#             'data.frame'), 
#         method = "jaccard",
#         binary = TRUE)

```

# Differential Abundance

## ALDex2

```{r}
library(ALDEx2)
data(selex)
#subset only the last 400 features for efficiency
selex.sub <- selex[1:400,]
conds <- c(rep("NS", 7), rep("S", 7))
x <- aldex.clr(selex.sub, conds, mc.samples=16, denom="iqlr", verbose=F)
x.tt <- aldex.ttest(x, paired.test=FALSE, verbose=FALSE)
x.effect <- aldex.effect(x, CI=T, verbose=FALSE)
head(x.effect)

```

```{r}
ps_fixed_family <- speedyseq::tax_glom(ps_fixed, taxrank="Family")

ps_fam_pas5 <- microViz::ps_filter(ps_fixed_family, passage == 5)

ps_fam_pas5_prev <- microViz::tax_filter(ps_fam_pas5, min_prevalence = 0.10)
ps_pass_5 <- microViz::ps_filter(ps_fixed, passage == 5)
otu_table <- as(t(otu_table(ps_fam_pas5_prev)), 'matrix')
conds <- c(rep("n", 12), rep("p", 12))

# Based on effect confidence interval

x <- aldex.clr(otu_table, conds, mc.samples=128, denom="all", verbose=F)
x.tt <- aldex.ttest(x, paired.test=FALSE, verbose=FALSE)
x.effect <- aldex.effect(x, CI=T, verbose=FALSE)

sgn <- sign(x.effect$effect.low) == sign(x.effect$effect.high)
par(mfrow=c(1,2))
plot(x.effect$rab.all, x.effect$diff.btw, pch=19, cex=0.3, col="grey", xlab="Abundance", ylab="Difference", main="Bland-Altman")
points(x.effect$rab.all[abs(x.effect$effect) >=2], x.effect$diff.btw[abs(x.effect$effect) >=2], pch=19, cex=0.5, col="red")
points(x.effect$rab.all[sgn], x.effect$diff.btw[sgn], cex=0.7, col="blue")

plot(x.effect$diff.win, x.effect$diff.btw, pch=19, cex=0.3, col="grey",xlab="Dispersion", ylab="Difference", main="Effect")
points(x.effect$diff.win[abs(x.effect$effect) >=2], x.effect$diff.btw[abs(x.effect$effect) >=2], pch=19, cex=0.5, col="red")
points(x.effect$diff.win[sgn], x.effect$diff.btw[sgn], cex=0.7, col="blue")
abline(0,2, lty=2, col="grey")
abline(0,-2, lty=2, col="grey")

merge(as(tax_table(ps_fam_pas5_prev)[rownames(x.effect[sgn,]), ], 'data.frame'), x.effect[sgn,], by = 0)
```

```{r}

# Based on effect size

x.all <- aldex(otu_table, conds, mc.samples=128, test="t", effect=TRUE,
     include.sample.summary=FALSE, denom="all", verbose=FALSE)
effect_gt_2 <- merge(as(tax_table(ps_pass_5)[rownames(x.all[abs(x.all$effect)>2,]), ], 'data.frame'), x.all[abs(x.all$effect)>2,], by = 0)

effect_gt_2 %>% 
  ggplot(aes(x = effect, y = reorder(Family,effect))) +
  geom_point()

par(mfrow=c(1,2))
aldex.plot(x.all, type="MA", test="welch", xlab="Log-ratio abundance",
    ylab="Difference")
aldex.plot(x.all, type="MW", test="welch", xlab="Dispersion",
    ylab="Difference")

par(mfrow=c(1,2))
plot(x.all$effect, x.all$we.ep, log="y", cex=0.7, col=rgb(0,0,1,0.2),
  pch=19, xlab="Effect size", ylab="P value", main="Effect size plot")
points(x.all$effect, x.all$we.eBH, cex=0.7, col=rgb(1,0,0,0.2),
  pch=19)
abline(h=0.05, lty=2, col="grey")
legend(15,1, legend=c("P value", "BH-adjusted"), pch=19, col=c("blue", "red"))

plot(x.all$diff.btw, x.all$we.ep, log="y", cex=0.7, col=rgb(0,0,1,0.2),
  pch=19, xlab="Difference", ylab="P value", main="Volcano plot")
points(x.all$diff.btw, x.all$we.eBH, cex=0.7, col=rgb(1,0,0,0.2),
  pch=19)
abline(h=0.05, lty=2, col="grey")

# identify which values are significant in both the t-test and glm tests
found.by.all <- which(x.all$we.eBH < 0.05 & x.all$wi.eBH < 0.05)

# identify which values are significant in fewer than all tests
found.by.one <- which(x.all$we.eBH < 0.05 | x.all$wi.eBH < 0.05)

# plot the within and between variation of the data
plot(x.all$diff.win, x.all$diff.btw, pch=19, cex=0.3, col=rgb(0,0,0,0.3),
 xlab="Dispersion", ylab="Difference")
points(x.all$diff.win[found.by.one], x.all$diff.btw[found.by.one], pch=19,
 cex=0.7, col=rgb(0,0,1,0.5))
points(x.all$diff.win[found.by.all], x.all$diff.btw[found.by.all], pch=19,
 cex=0.7, col=rgb(1,0,0,1))
abline(0,1,lty=2)
abline(0,-1,lty=2)

#View(tax_table(ps_fam_pas5_prev)[rownames(otu_table[found.by.all, ]), ])
```

```{r, eval=FALSE}
boxData <-
  ps_fixed %>% 
  ps_filter(passage ==5) %>% 
  tax_agg('Family') %>% 
  ps_get %>% 
  ps_melt()

Family_order<-
  boxData %>% 
  group_by(Family) %>% 
  summarize(median = median(Abundance)) %>% 
  arrange(median)
boxData$Family <- factor(boxData$Family, levels = Family_order$Family)
boxData$treat <- factor(boxData$treat, levels = c('p', 'n'))
```

```{r, eval=FALSE, fig.height=3, fig.width=12}

boxData %>% 
  ggplot(aes(x = Family, 
             y = Abundance)) +
  facet_grid(treat ~ .) +
  theme_tufte() +
  geom_rangeframe(color = "black") +
  # scale_x_reordered() +
  scale_y_continuous(trans = "pseudo_log") +
  geom_tufteboxplot() +
  theme(axis.text.x = element_text(angle = 90))
  # geom_boxplot() +
  # scale_color_discrete(guide = 'none') +
  # coord_flip()

```

```{r, eval=FALSE}
g5p<- 
  ps_fixed %>% 
  ps_filter(treat == "p", passage == 5) %>% 
  tax_agg('Family') %>% 
  ps_get() %>% 
  ps_melt()
g5n<- 
  ps_fixed %>% 
  ps_filter(treat == "n", passage == 5) %>% 
  tax_agg('Family') %>% 
  ps_get() %>% 
  ps_melt()

# Families unique to p in passage 5
g5p_unique <-
  g5p %>% 
  filter(Family %in% unique(g5p$Family)[!(unique(g5p$Family) %in% unique(g5n$Family))]) %>% 
  mutate(Prevalence = as.numeric(Abundance > 0))

# Abundance
Family_order<-
  g5p_unique %>% 
  group_by(OTU) %>% 
  summarize(median = median(Abundance), max = max(Abundance)) %>% 
  arrange(median, max)
g5p_unique$OTU <- factor(g5p_unique$OTU, levels = Family_order$OTU)

g5p_unique %>% 
  ggplot(aes(y = Abundance, x = OTU)) +
  geom_tufteboxplot() +
  theme_tufte() +
  geom_rangeframe() +
  theme(axis.text.x = element_text(angle = 90)) +
  scale_y_continuous(breaks = c(0, 5, 10, 15, 20, 24))

# Prevalence
Family_order<-
  g5p_unique %>% 
  group_by(OTU) %>% 
  summarize(sum = sum(Prevalence)) %>% 
  arrange(sum)
g5p_unique$OTU <- factor(g5p_unique$OTU, levels = Family_order$OTU)

g5p_unique %>% 
  group_by(OTU) %>% 
  summarize(Prevalence = sum(Prevalence)) %>% 
  ggplot(aes(y = Prevalence, x = OTU)) +
  geom_point() +
  theme_tufte() +
  geom_rangeframe() +
  theme(axis.text.x = element_text(angle = 90)) +
  scale_y_continuous(breaks = c(1, 2, 3, 5, 10, 12))

# Families unique to n in passage 5
g5n_unique <-
  g5n %>% 
  filter(Family %in% unique(g5n$Family)[!(unique(g5n$Family) %in% unique(g5p$Family))]) %>% 
  mutate(Prevalence = as.numeric(Abundance > 0))

# Abundance
Family_order<-
  g5n_unique %>% 
  group_by(OTU) %>% 
  summarize(median = median(Abundance), max = max(Abundance)) %>% 
  arrange(median, max)
g5n_unique$OTU <- factor(g5n_unique$OTU, levels = Family_order$OTU)

g5n_unique %>% 
  ggplot(aes(y = Abundance, x = OTU)) +
  geom_tufteboxplot() +
  theme_tufte() +
  geom_rangeframe() +
  theme(axis.text.x = element_text(angle = 90)) +
  scale_y_continuous(breaks = c(0, 10, 20, 30, 37))

# Prevalence
Family_order<-
  g5n_unique %>% 
  group_by(OTU) %>% 
  summarize(sum = sum(Prevalence)) %>% 
  arrange(sum)
g5n_unique$OTU <- factor(g5n_unique$OTU, levels = Family_order$OTU)

g5n_unique %>% 
  group_by(OTU) %>% 
  summarize(Prevalence = sum(Prevalence)) %>% 
  ggplot(aes(y = Prevalence, x = OTU)) +
  geom_point() +
  theme_tufte() +
  geom_rangeframe() +
  theme(axis.text.x = element_text(angle = 90)) +
  scale_y_continuous(breaks = c(1, 2, 6))
```

First, I looked at which taxa are unique to treatments N and P. Several families
present in P and absent in N. Most of them had low prevalance (present in 3 or
fewer samples) and low abundance (median < 5 reads across samples). There were two families unique
to the Positive treatment with relatively high prevalence. This included an ASV
that was a member of the Bacteroidia Class with no lower taxonomic designation. 
This ASV had a prevalence of 10/12 and a median abundance of ~2 reads. The other
prevalent family was a member of the Silvanigrellaceae, a newly described family places in
its own order. This family was present in
all 12 samples and was represented by three OTUs: one with no lower designation, one from the genus
Silvanigrella and one from the genus Candidatus Turbobacter. Silvanigrella was
isolated from a temperate fresh water lake (https://www.microbiologyresearch.org/content/journal/ijsem/10.1099/ijsem.0.001965).

Of the families unique to the Neutral treatment, only one had a prevalence greater than 2/12. 
This family, Armatimonadaceae, was present in half of the Neutral samples (6/12). This
family was represented by two ASVs in the genus Armatimonas. The type strain for
Armatimonas was isolated from the rhizoplane of Phragmites australis (https://www.microbiologyresearch.org/content/journal/ijsem/10.1099/ijs.0.025643-0)

```{r, eval=FALSE}

g5<- 
  ps_fixed %>% 
  ps_filter(passage == 5) %>% 
  tax_agg('Family') %>% 
  ps_get() %>% 
  ps_melt()

g2<- 
  ps_fixed %>% 
  ps_filter(passage == 2) %>% 
  tax_agg('Family') %>% 
  ps_get() %>% 
  ps_melt()

unique(g5$Family)[!(unique(g5$Family) %in% unique(g2$Family))]
unique(g5$Family)[unique(g5$Family) %in% unique(g2$Family)]

unique(g2$Family)[!(unique(g2$Family) %in% unique(g5$Family))]
unique(g2$Family)[unique(g2$Family) %in% unique(g5$Family)]

```

## Differential Abundance Treatment in passage 5

```{r corncob_treatment, fig.hight=40}
ps_fixed_family <- speedyseq::tax_glom(ps_fixed, taxrank="Family")

ps_fam_pas5 <- microViz::ps_filter(ps_fixed_family, passage == 5)

ps_fam_pas5_prev <- microViz::tax_filter(ps_fam_pas5, min_prevalence = 0.10)

sample_data(ps_fam_pas5_prev) =
  sample_data(ps_fam_pas5_prev)[,c('sample_id', 'treat')] %>% 
  sample_data()
ps = ps_fam_pas5_prev

save(ps, file = 'ps.RData')
# Differential abundance test
corncob <- bbdml(formula = ASV2954 ~ treat,
                                 phi.formula = ~ 1,
                                 formula_null = ~ 1,
                                 phi.formula_null = ~ 1,
                                 test = "lrt", boot = TRUE,
                                 data = ps_fam_pas5_prev,
                                 fdr_cutoff = 0.05)
da_analysis <- differentialTest(formula = ~ treat,
                                 phi.formula = ~ 1,
                                 formula_null = ~ 1,
                                 phi.formula_null = ~ 1,
                                 test = "Wald", boot = FALSE,
                                 data = ps_fam_pas5_prev,
                                 fdr_cutoff = 0.05, B = 50)
plot(da_analysis, level = c("Family"))
da_analysis$significant_models
dt_data <- plot(da_analysis, level = c("Family"), data_only = TRUE)

sig_taxa_treat <- tax_select(ps_fam_pas5_prev, da_analysis$significant_taxa)
ntaxa(sig_taxa_treat)
save(da_analysis, sig_taxa_treat, file = 'Output/da_corncob.Rdata')

save(dt_data, file = 'Output/dt_data.Rdata')

```

```{r corncob-plot, fig.width=6, fig.height=6}
dt_plot(dt_data)

```

## Faprotax on DA taxa

```{r export_to_faprotax, eval=FALSE}
faprotax_taxonomy <-
  sig_taxa_treat %>% 
  tax_table %>% 
  as_tibble %>% 
  mutate(.taxonomy = 
           faprotax_taxonomy %>% 
           select(Kingdom:Species) %>% 
           apply(1 , paste , collapse = "; " )) %>% 
  select(.otu, .taxonomy)

otu<-t(as(otu_table(sig_taxa_treat),"matrix")) 
  # 't' to transform if taxa_are_rows=FALSE
  #if taxa_are_rows=TRUE
  #otu<-as(otu_table(GlobalPatterns),"matrix"))
rownames(otu) <- faprotax_taxonomy$.taxonomy
otu_biom <- biomformat::make_biom(data=otu)
biomformat::write_biom(otu_biom,"otu_biom.biom")

```

```{bash faprotax, eval=FALSE}
conda activate faprotax

rm Output/func_table_from_biom.biom \
    Output/faprotax_report_from_biom.txt

python python/FAPROTAX_1.2.4/collapse_table.py \
    -i otu_biom.biom \
    -g python/FAPROTAX_1.2.4/FAPROTAX.txt \
    -o Output/func_table_from_biom.biom \
    -r Output/faprotax_report_from_biom.txt \
    -v

biom convert \
    -i Output/func_table_from_biom.biom \
    -o Output/func_table_from_biom_json.biom \
    --table-type="OTU table" \
    --to-json

conda deactivate
```

```{r}
biom_file <- "Output/func_table_from_biom_json.biom"

ps_faprotax <-
  biom_file %>% 
  biomformat::read_biom() %>% 
  biomformat::biom_data() %>% 
  as("matrix") %>% 
  otu_table(taxa_are_rows=TRUE) %>% 
  phyloseq(sample_data(ps_gen_pas5_prev))

da_faprotax <- differentialTest(formula = ~ treat,
                                 phi.formula = ~ treat,
                                 formula_null = ~ 1,
                                 phi.formula_null = ~ treat,
                                 test = "Wald", boot = FALSE,
                                 data = ps_faprotax,
                                 fdr_cutoff = 0.05)
plot(da_faprotax)

```

## Picrust on DA taxa

```{r}
x = NULL
for (i in seq_along(da_analysis$significant_models)) {
  x <- c(x, da_analysis$significant_models[[i]]$coefficients[2, 1])
}
x > 0
sig_taxa_pos_neg <-
  da_analysis$significant_taxa %>% 
  as_tibble_col(column_name = ".otu") %>% 
  mutate(estimate = x) %>% 
  mutate(pos_or_neg = if_else(x > 0, "pos", "neg"))

top_da_taxa <- 
  sig_taxa_treat %>% 
  otu_table %>% 
  as_tibble %>% 
  filter(.abundance > 2000) %>% 
  distinct(.otu) %$% .otu

taxonomy <-
  sig_taxa_treat %>% 
  tax_table %>% 
  as_tibble %>%
  filter(.otu %in% top_da_taxa)

taxonomy$.taxonomy <- taxonomy %>% 
           select(Kingdom:Species) %>% 
           apply(1 , paste , collapse = "; " )

taxonomy <-
  taxonomy %>% 
  select(.otu, .taxonomy)

df_top_da_taxa <-
  sig_taxa_treat %>% 
  otu_table %>% 
  as_tibble %>%
  left_join(as_tibble(sample_data(sig_taxa_treat))) %>% 
  left_join(sig_taxa_pos_neg, by = ".otu") %>% 
  filter(.otu != 'ASV4174') %>% 
  filter(.otu %in% top_da_taxa) %>% 
  left_join(taxonomy)

df_top_da_taxa %>% 
  ggplot(., aes(y = Genus, x = .abundance, color = treat)) +
  geom_jitter(height = 0.1) + 
  #theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  facet_wrap(~ pos_or_neg)
ggsave('Output/dif_abund_results.png', width = 6, height = 4)

major_otus <- 
sig_taxa_treat %>% 
  otu_table %>% 
  as_tibble %>%
  left_join(as_tibble(sample_data(sig_taxa_treat))) %>% 
  left_join(sig_taxa_pos_neg, by = ".otu") %>% 
  filter(.otu != 'ASV4174') %>% 
  filter(.abundance > 300) %>% 
  select(.otu, pos_or_neg) %>% 
  distinct %>% 
  mutate(tax = otu_to_taxonomy(.otu, sig_taxa_treat))

otu_to_taxonomy(major_otus$.otu, sig_taxa_treat)

otu_to_taxonomy('ASV192', sig_taxa_treat)

sig_taxa_treat %>% 
  otu_table %>% 
  as_tibble %>%
  left_join(as_tibble(sample_data(sig_taxa_treat))) %>% 
  left_join(sig_taxa_pos_neg, by = ".otu") %>% 
  filter(.otu == 'ASV192') %>% 
  ggplot(., aes(x = .otu, y = .abundance, color = treat)) +
  geom_point() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  facet_wrap(~ pos_or_neg)

beijer_otu <-
  ps %>% 
  otu_table %>% 
  as_tibble %>% 
  filter(.otu == 'ASV192')

beijer_metadata <-
  ps %>% 
  sample_data %>% 
  as_tibble %>% 
    left_join(beijer_otu)

ggplot(beijer_metadata, aes(x = .abundance, y = log_ch4, color = treat)) +
  geom_point() +
  facet_wrap(~ passage)

df_top_da_taxa %>% 
filter(pos_or_neg == "pos") %>% 
select(Phylum:Genus) %>% 
distinct %>% 
  write_tsv(., 'top_positive_taxa.txt')
```

```{bash, eval=FALSE}
mkdir picrust2
```

```{r export_to_picrust}

top_da_taxa_pos <-
  df_top_da_taxa %>% 
  select(.otu, pos_or_neg) %>% 
  distinct(.otu, pos_or_neg) %>% 
  filter(pos_or_neg == "pos") %$%
  .otu
sig_taxa_treat %>% 
  tax_select(top_da_taxa_pos) %>% 
  otu_table %>% 
  as("matrix") %>% 
  t %>% 
  biomformat::make_biom() %>% 
  biomformat::write_biom("Output/study_seqs.biom")

sequences <-
  sig_taxa_treat %>% 
  tax_select(top_da_taxa_pos) %>% 
  refseq %>% 
  as.vector %>% 
  as.list
seqinr::write.fasta(sequences, names(sequences), file.out = "Output/study_seqs.fna")

```

```{bash picrust, eval=FALSE}
conda activate picrust2
cd picrust3c

place_seqs.py -s ../Output/study_seqs.fna \
    -o out.tre \
    -p 1 \
    --intermediate intermediate/place_seqs \
    --chunk_size 500
hsp.py -i 16S -t out.tre -o marker_predicted_and_nsti.tsv.gz -p 1 -n

hsp.py -i EC -t out.tre -o EC_predicted.tsv.gz -p 1

metagenome_pipeline.py -i ../Output/study_seqs.biom -m marker_predicted_and_nsti.tsv.gz -f EC_predicted.tsv.gz \
                       -o EC_metagenome_out --strat_out

pathway_pipeline.py -i EC_metagenome_out/pred_metagenome_contrib.tsv.gz \
                    -o pathways_out -p 1

add_descriptions.py -i EC_metagenome_out/pred_metagenome_unstrat.tsv.gz -m EC \
                    -o EC_metagenome_out/pred_metagenome_unstrat_descrip.tsv.gz

add_descriptions.py -i pathways_out/path_abun_unstrat.tsv.gz -m METACYC \
                    -o pathways_out/path_abun_unstrat_descrip.tsv.gz

#####

picrust2_pipeline.py -s Output/study_seqs.fna -i Output/study_seqs.biom -o picrust2_out_pipeline -p 4 --stratified

conda deactivate
```

```{r}
se.mb.amgut <- spiec.easi(amgut1.filt, method='mb', lambda.min.ratio=1e-2,
                          nlambda=20, pulsar.params=list(rep.num=50))
se.gl.amgut <- spiec.easi(amgut1.filt, method='glasso', lambda.min.ratio=1e-2,
                          nlambda=20, pulsar.params=list(rep.num=50))
sparcc.amgut <- sparcc(amgut1.filt)
## Define arbitrary threshold for SparCC correlation matrix for the graph
sparcc.graph <- abs(sparcc.amgut$Cor) >= 0.3
diag(sparcc.graph) <- 0
library(Matrix)
sparcc.graph <- Matrix(sparcc.graph, sparse=TRUE)
## Create igraph objects
ig.mb     <- adj2igraph(getRefit(se.mb.amgut))
ig.gl     <- adj2igraph(getRefit(se.gl.amgut))
ig.sparcc <- adj2igraph(sparcc.graph)

```

```{r}
library(igraph)
## set size of vertex proportional to clr-mean
vsize    <- rowMeans(clr(amgut1.filt, 1))+6
am.coord <- layout.fruchterman.reingold(ig.mb)

par(mfrow=c(1,3))
plot(ig.mb, layout=am.coord, vertex.size=vsize, vertex.label=NA, main="MB")
plot(ig.gl, layout=am.coord, vertex.size=vsize, vertex.label=NA, main="glasso")
plot(ig.sparcc, layout=am.coord, vertex.size=vsize, vertex.label=NA, main="sparcc")

```

```{r}

library(phyloseq)
## Load round 2 of American gut project
data('amgut2.filt.phy')
se.mb.amgut2 <- spiec.easi(amgut2.filt.phy, method='mb', lambda.min.ratio=1e-2,
                           nlambda=20, pulsar.params=list(rep.num=50))
ig2.mb <- adj2igraph(getRefit(se.mb.amgut2),  vertex.attr=list(name=taxa_names(amgut2.filt.phy)))
plot_network(ig2.mb, amgut2.filt.phy, type='taxa', color="Rank3")

ps_p <-
  test %>% 
  speedyseq::tax_glom("Family") %>% 
  ps_filter(passage == "5") %>% 
  ps_filter(treat == "p")

se.mb.ps_p <- spiec.easi(ps_p, method='mb', lambda.min.ratio=1e-2,
                           nlambda=200, pulsar.params=list(rep.num=50))
ig2.mb.ps_p <- adj2igraph(getRefit(se.mb.ps_p),  vertex.attr=list(name=taxa_names(ps_p)))
plot_network(ig2.mb.ps_p, ps_p, type='taxa', color = "Class")

ps_n <-
  test %>% 
  speedyseq::tax_glom("Family") %>% 
  ps_filter(passage == "5") %>% 
  ps_filter(treat == "n")

se.mb.ps_n <- spiec.easi(ps_n, method='mb', lambda.min.ratio=1e-2,
                           nlambda=200, pulsar.params=list(rep.num=50))
ig2.mb.ps_n <- adj2igraph(getRefit(se.mb.ps_n),  vertex.attr=list(name=taxa_names(ps_n)))
plot_network(ig2.mb.ps_n, ps_n, type='taxa', color="Class")
getOptMerge(se.mb.ps_p)
getOptMerge(se.mb.ps_n)
getStability(se.mb.ps_n)
###

se1.mb.amgut <- spiec.easi(amgut1.filt, method='mb', lambda.min.ratio=1e-2,
                          nlambda=20, pulsar.params=list(rep.num=20, ncores=4))
se2.mb.amgut <- spiec.easi(amgut2.filt.phy, method='mb', lambda.min.ratio=1e-2,
                          nlambda=20, pulsar.params=list(rep.num=20, ncores=4))

se1.slr.amgut <- spiec.easi(amgut1.filt, method='slr', r=10, lambda.min.ratio=1e-2,
                          nlambda=20, pulsar.params=list(rep.num=20, ncores=4))
se2.slr.amgut <- spiec.easi(amgut2.filt.phy, method='slr', r=10, lambda.min.ratio=1e-2,
                          nlambda=20, pulsar.params=list(rep.num=20, ncores=4))

otu1 <- colnames(amgut1.filt)
otu2 <- taxa_names(amgut2.filt.phy)
edge.diss(getRefit(se1.mb.amgut), getRefit(se2.mb.amgut), 'jaccard', otu1, otu2)
edge.diss(getRefit(se1.slr.amgut), getRefit(se2.slr.amgut), 'jaccard', otu1, otu2)

X <- se2.slr.amgut$est$data
L <- se2.slr.amgut$est$resid[[getOptInd(se2.slr.amgut)]]
pca <- robustPCA(X, L)
age <- as.numeric(as.character(sample_data(amgut2.filt.phy)$AGE))
bmi <- as.numeric(as.character(sample_data(amgut2.filt.phy)$BMI))
depth <- colSums(otu_table(amgut2.filt.phy))

cor(age, pca$scores, use='pairwise')
cor(bmi, pca$scores, use='pairwise')
cor(depth, pca$scores, use='pairwise')

```

## Differential Abundance Methane

```{r corncob_ch4}

ps_gen_pas5_trep <- ps_filter(ps_gen_pas5, treat == "p")

ps_gen_prev_trep_prev <- tax_filter(ps_gen_pas5_trep, min_prevalence = 0.10)

da_analysis <- differentialTest(formula = ~ ch4,
                                 phi.formula = ~ ch4,
                                 formula_null = ~ 1,
                                 phi.formula_null = ~ ch4,
                                 test = "Wald", boot = FALSE,
                                 data = ps_gen_prev_trep_prev,
                                 fdr_cutoff = 0.05)
# otu_to_taxonomy(OTU = da_analysis$significant_taxa, data = ps_gen_prev)

plot(da_analysis, level = c("Family", "Genus"))

# Check unfitted taxa
# otu_to_taxonomy(which(is.na(da_analysis$p)) %>% names, data = ps_gen_prev)
# otu_table(ps_gen_prev)[,which(is.na(da_analysis$p))]

# Differential variability test

dv_analysis <- differentialTest(formula = ~ passage + treat + ch4,
                                 phi.formula = ~ passage + treat + ch4,
                                 formula_null = ~ passage + treat + ch4,
                                 phi.formula_null = ~ passage + treat,
                                 data = ps_gen_prev,
                                 test = "Wald", boot = FALSE,
                                 fdr_cutoff = 0.05)
# otu_to_taxonomy(OTU = dv_analysis$significant_taxa, data = ps_gen_prev)
plot(dv_analysis, level = c("Family", "Genus"))

# Check unfitted taxa

# otu_to_taxonomy(which(is.na(dv_analysis$p)) %>% names, data = ps_gen_prev)
# otu_table(ps_gen_prev)[,which(is.na(dv_analysis$p))]

```

## Alpha Diversity Methane

```{r diversity_ch4, eval = FALSE}
ps_phylum <- speedyseq::tax_glom(ps, taxrank="Phylum")

divnet_phylum_ch4 <- ps_phylum %>%
  divnet(ncores = 4, tuning = "careful", formula = ~ rank_ch4)

divnet_phylum_ch4$shannon %>% 
  plot(ps_phylum, color = "rank_ch4") +
  xlab("Rank Methane") +
  ylab("Shannon diversity estimate\n(phylum level)") #+
  #coord_cartesian(ylim = c(0,2))

estimates <- divnet_phylum_ch4$shannon %>% summary %$% estimate
ses <- sqrt(divnet_phylum_ch4$`shannon-variance`)
X <- breakaway::make_design_matrix(ps_phylum, "rank_ch4")
betta(estimates, ses, X)$table

betta(estimates, ses, X)$global[2]

```

## Alpha Diversity Treatment in Passage 2

```{r, results=FALSE, eval = FALSE}
ps_class <- speedyseq::tax_glom(ps, taxrank="Class")
ps_class

ps_c_2 <- ps_filter(ps_class, passage == 2)

divnet_c_2 <- ps_c_2 %>%
  divnet(ncores = 4, tuning = "careful", formula = ~ treat)
```

```{r, eval = FALSE}
divnet_c_2$shannon %>% 
  plot(ps_c_2, color = "treat") +
  xlab("Selection treatment") +
  ylab("Shannon diversity estimate\n(class level)") +
  labs(title="Test alpha diversity between P and N treatments within Passage 2")

estimates <- divnet_c_2$shannon %>% summary %$% estimate
ses <- sqrt(divnet_c_2$`shannon-variance`)
X <- breakaway::make_design_matrix(ps_c_2, "treat")
betta(estimates, ses, X)$table
betta(estimates, ses, X)$global

divnet_c_2$simpson %>% 
  plot(ps_c_2, color = "treat") +
  xlab("Selection treatment") +
  ylab("Simpson diversity estimate\n(class level)") +
  labs(title="Test alpha diversity between P and N treatments within Passage 2")

estimates <- divnet_c_2$simpson %>% summary %$% estimate
ses <- sqrt(divnet_c_2$`simpson-variance`)
X <- breakaway::make_design_matrix(ps_c_2, "treat")
betta(estimates, ses, X)$table
betta(estimates, ses, X)$global

```

## Alpha Diversity Passage 2 vs 5 (all treatment)

```{r, results=FALSE, eval=FALSE}
ps_phylum <- speedyseq::tax_glom(ps, taxrank="Phylum")

divnet_p <- ps_phylum %>%
  divnet(ncores = 4, tuning = "careful", formula = ~ passage)
```

```{r, eval = FALSE}
divnet_p$shannon %>% 
  plot(ps_phylum, color = "passage") +
  xlab("Selection treatment") +
  ylab("Shannon diversity estimate\n(phylum level)") +
  labs(title="Test alpha diversity between Passage 2 and 5 all treatments")

estimates <- divnet_p$shannon %>% summary %$% estimate
ses <- sqrt(divnet_p$`shannon-variance`)
X <- breakaway::make_design_matrix(ps_phylum, "passage")
betta(estimates, ses, X)$table
betta(estimates, ses, X)$global[2]

divnet_p$simpson %>% 
  plot(ps_phylum, color = "passage") +
  xlab("Selection treatment") +
  ylab("Simpson diversity estimate\n(phylum level)") +
  labs(title="Test alpha diversity between Passage 2 and 5 all treatments")

estimates <- divnet_p$simpson %>% summary %$% estimate
ses <- sqrt(divnet_p$`simpson-variance`)
X <- breakaway::make_design_matrix(ps_phylum, "passage")
betta(estimates, ses, X)$table
betta(estimates, ses, X)$global[2]

```

## Alpha Diversity Treatment in Passage 5

```{r, results=FALSE, eval = FALSE}

ps_class <- speedyseq::tax_glom(ps, taxrank="Class")

ps_c_5 <- ps_filter(ps_class, passage == 5)

ps_5 <- ps_filter(ps_phylum, passage == 5)

divnet_p_5 <- ps_5 %>%
  divnet(ncores = 4, tuning = "careful", formula = ~ treat)
```

```{r, eval = FALSE}
divnet_p_5$shannon %>% 
  plot(ps_5, color = "treat") +
  xlab("Selection treatment") +
  ylab("Shannon diversity estimate\n(phylum level)") +
  labs(title="Test alpha diversity between P and N treatments within Passage 5")

estimates <- divnet_p_5$shannon %>% summary %$% estimate
ses <- sqrt(divnet_p_5$`shannon-variance`)
X <- breakaway::make_design_matrix(ps_5, "treat")
betta(estimates, ses, X)$table
betta(estimates, ses, X)$global
#betta_pic(estimates, ses)

divnet_p_5$simpson %>% 
  plot(ps_5, color = "treat") +
  xlab("Selection treatment") +
  ylab("Simpson diversity estimate\n(phylum level)") +
  labs(title="Test alpha diversity between P and N treatments within Passage 5")

estimates <- divnet_p_5$simpson %>% summary %$% estimate
ses <- sqrt(divnet_p_5$`simpson-variance`)
X <- breakaway::make_design_matrix(ps_5, "treat")
betta(estimates, ses, X)$table
betta(estimates, ses, X)$global
#betta_pic(estimates, ses)

```

## Beta Diversity passage + treatment

```{r, eval=FALSE}

ps_f_c <- speedyseq::tax_glom(ps_fixed, "Class")

X <- diag(nrow(sample_data(ps_f_c)))
colnames(X) <- rownames(X) <-  as.character(1:ncol(X))

divnet_class <-  divnet(W = ps_f_c,
                         X = X,
                         variance = "none")

# Median composition
set.seed(454)
bc_test <- testBetaDiversity(divnet_class, 
                             h0 = "bray-curtis",
                             n_boot = 1000, # set this to 10000
                             sample_specimen_matrix = divnet_class$X,
                             groups = sample_data(ps_f_c)$treat_pas)
bc_test$p_value

# Euclidean distance
set.seed(3532)
euc_test <- testBetaDiversity(dv = divnet_class , 
                              h0 = "euclidean",
                              n_boot = 1000,
                              sample_specimen_matrix = divnet_class$X,
                              groups = sample_data(ps_f_c)$treat_pas)
euc_test$p_value

# Aitchison Distance
set.seed(3423)
ait_test <- testBetaDiversity(dv = divnet_class, 
                              sample_specimen_matrix = divnet_class$X,
                              h0 = "aitchison",
                              n_boot = 1000,
                    groups = sample_data(ps_f_c)$treat_pas)

ait_test$p_value
```

## Beta Diversity gen 2 treatment

```{r, results=FALSE, eval = FALSE}
ps_f_2 <- ps_filter(ps_fixed, passage == "2")
ps_f_2_c <- speedyseq::tax_glom(ps_f_2, "Class")
X <- diag(nrow(sample_data(ps_f_2_c)))
colnames(X) <- rownames(X) <-  as.character(1:ncol(X))

divnet_c_2 <-  divnet(W = ps_f_2_c,
                         X = X,
                         variance = "none")

```

```{r, eval = FALSE}

# Median composition
set.seed(454)
bc_test <- testBetaDiversity(dv = divnet_c_2, 
                             h0 = "bray-curtis",
                             n_boot = 1000, # set this to 10000
                             sample_specimen_matrix = divnet_c_2$X,
                             groups = sample_data(ps_f_2_c)$treat_bin)
bc_test$p_value
set.seed(3532)
# Euclidean distance
euc_test <- testBetaDiversity(dv = divnet_c_2 , 
                              h0 = "euclidean",
                              n_boot = 1000,
                              sample_specimen_matrix = divnet_c_2$X,
                              groups = sample_data(ps_f_2_c)$treat_bin)
euc_test$p_value

# Aitchison Distance
set.seed(3423)
ait_test <- testBetaDiversity(dv = divnet_c_2, 
                              sample_specimen_matrix = divnet_c_2$X,
                              h0 = "aitchison",
                              n_boot = 1000,
                              groups = sample_data(ps_f_2_c)$treat_bin)

ait_test$p_value

```

## Beta Diversity gen 5 treatment

```{r, eval = FALSE}
ps_f_5 <- ps_filter(ps_fixed, passage == "5")
ps_f_5_c <- speedyseq::tax_glom(ps_f_5, "Class")
X <- diag(nrow(sample_data(ps_f_5_c)))
colnames(X) <- rownames(X) <-  as.character(1:ncol(X))

divnet_c_5 <-  divnet(W = ps_f_5_c,
                         X = X,
                         variance = "none")

# Median composition
set.seed(454)
bc_test <- testBetaDiversity(dv = divnet_c_5, 
                             h0 = "bray-curtis",
                             n_boot = 1000, # set this to 10000
                             sample_specimen_matrix = divnet_c_5$X,
                             groups = sample_data(ps_f_5_c)$treat_bin)
bc_test$p_value
set.seed(3532)
# Euclidean distance
euc_test <- testBetaDiversity(dv = divnet_c_5 , 
                              h0 = "euclidean",
                              n_boot = 1000,
                              sample_specimen_matrix = divnet_c_5$X,
                              groups = sample_data(ps_f_5_c)$treat_bin)
euc_test$p_value

# Aitchison Distance
set.seed(3423)
ait_test <- testBetaDiversity(dv = divnet_c_5, 
                              sample_specimen_matrix = divnet_c_5$X,
                              h0 = "aitchison",
                              n_boot = 1000,
                              groups = sample_data(ps_f_5_c)$treat_bin)

ait_test$p_value

```

### 2022-04-07 R/06-dada2.R

@@ -118,16 +118,4 @@ taxa.print <- taxa # Removing sequence rownames for display only
 rownames(taxa.print) <- NULL
 head(taxa.print)
 
-save.image(file='myEnvironment.RData')
-#load('myEnvironment.RData')
-
-library(phyloseq); packageVersion("phyloseq")
-library(Biostrings); packageVersion("Biostrings")
-
-samples <- read_tsv('Output/barcode_master.tsv')
-fluxes <- read_tsv('Output/fluxes.tsv')
-
-ps <- phyloseq(otu_table(seqtab.nochim, taxa_are_rows=FALSE), 
-               sample_data(samdf), 
-               tax_table(taxa))
-
+save(seqtab.nochim, taxa, file = "Output/dada2_output.RData")

### 2022-04-07 R/functions.R

@@ -1,22 +1,32 @@
-library(tidyverse)
-library(broman)
 
 raw_dir <- '../Data/' # Raw Data Directory
 der_dir <- '../Output/' # Derived/Modified Data Directory
 
+# theme_set(theme_tufte()+
+#             theme(axis.title.y = element_text(angle = 0, vjust = 0.5)))
+theme_set(theme_classic() +
+            theme(panel.border = element_rect(fill = NA, size = 1.0625),
+                  axis.line = element_blank()))
 
 plot_flux <- function(fluxes, passage, estimate, ratio = NULL, log10 = FALSE) {
   p <- ggplot(fluxes, aes(x = {{passage}}, y = {{estimate}}, color = treat)) +
-    theme_classic() +
+    # theme_classic() +
+    # geom_rangeframe(color = 'black') +
     stat_smooth(method = 'lm', se = FALSE, formula = 'y ~ x') +
-    scale_color_manual(name = "Treatment", labels = c('Neutral', 'Positive'), values = c('gray40', 'darkorange2')) +
-    geom_jitter(aes(alpha = selected), width = 0.1) +
-    scale_alpha_manual(name = "Selected", labels = c('Not', 'Selected'), values = c(0.3, 1))
+    scale_color_manual(
+      name = "Treatment", 
+      labels = c('Neutral', 'Positive'), 
+      values = c('gray40', 'darkorange2')) +
+    geom_jitter(#aes(alpha = selected), 
+      width = 0.1) #+
+#    scale_alpha_manual(name = "Selected", labels = c('Not', 'Selected'), values = c(0.3, 1))
   if (log10 == FALSE) {
-    p <- p + labs(x = "Passage Number", y = "Methane Oxidation Rate (-k)")
+    p <- p + labs(x = "Passage Number", 
+                  y = "Methane Oxidation Rate (-k)")
   } 
   else if (log10 == TRUE) {
-    p <- p + labs(x = "Passage Number", y = expression(log[10] * "(Methane Oxidation Rate (-k))"))
+    p <- p + labs(x = "Passage Number", 
+                  y = expression(log[10] ~ CH[4] ~ "(-k)"))
   } 
   
   if (!is.null(ratio)) {
@@ -48,11 +58,12 @@ print_lrt <- function(fit, term = 2, test = "Likelihood ratio test") {
 plot_herit <- function(heritability, ratio) {
   ratio <- calc_plot_ratios(heritability$selected, heritability$offspring)
[…]

### 2022-04-07 Manuscript/manuscript.Rmd
Update manuscript

--- 2021-07-01 Manuscript/manuscript.Rmd
@@
 ---
-title: Artificial ecosystem selection to deduce the mapping between microbial community structure and ecosystem function
-author: "Andrew H. Morris and Brendan J. M. Bohannan"
+title: Artificial ecosystem selection reveals relationships between microbiome composition and ecosystem function
+author: "Andrew H. Morris^1^* and Brendan J. M. Bohannan^1^"
 output:
@@
     reference_docx: reference.docx
-editor_options:
-  chunk_output_type: inline
+bibliography: bibliography.bib
+csl: vancouver.csl
 ---
+
+^1^ Institute of Ecology and Evolution, University of Oregon, Eugene, OR, USA
+
+\* Correspondence: Andrew H. Morris, amorris3@uoregon.edu
+
+Competing Interests: We declare we have no competing interests.
 
@@
 knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
+
 ```
 
-```{r initialize_packages}
+```{r initialize-packages}
 library(tidyverse)
-library(captioner)
 library(broom)
@@
 library(morris)
-#library(metafor)
+library(phyloseq)
+library(ggpubr)
+library(knitr)
+library(corncob)
+library(ggthemes)
+library(microViz)
+library(captioner)
+figures <- captioner()
+tables <- captioner(prefix = "Table")
 ```
-```{r load_data}
+
+```{r load-data}
 heritability <- read_tsv('../Output/heritability.tsv')
@@
 
+load(file = '../Output/richness_model_passage.Rdata')
+load(file = '../Output/richness_model_treat5.Rdata')
+load(file = '../Output/richness_model_ch4.Rdata')
+load(file = '../Output/phyloseq.RData')
+load(file = '../Output/beta_model.Rdata')
+load(file = '../Output/da_corncob.Rdata')
+load(file = '../Output/dt_data.Rdata')
+
 source('../R/functions.R')
-
-```
-
-```{r figure_captions, include=FALSE}
-
-figures <- captioner()
-figures("response", "Deviance of CH~4~ oxidation rate for the Positive selection treatment relative to the Neutral selection treatment.
-Deviance is calculated as the mean of P within each passage minus the mean of N within each passage.
-Errorbars are the standard error of the difference between P and N for each passage.
-The best fit line is an ordinary least squares fit taking into account the standard errors of the deviations.
-The gray ribbon is the 95% confidence interval for the regression.")
-figures("herit", "Regression of offspring CH~4~ flux on mid-parent CH~4~ flux.
-Mid-parent is the mean of jars selected to inoculate the next passage.
-The mid-offsprings are the mean of all twelve jars produced in one passage.
-The best fit line is an ordinary least squares regression line and the gray ribbon is the 95% confidence interval for the regression coefficient.")
 
@@
 
+Microbiomes mediate important ecosystem functions, yet it has proven difficult
+to determine the relationship between microbiome composition and the rate of
+any ecosystem function. This challenge remains because microbiomes are often
+complex, it is difficult to know a priori which microbiome members limit the
+rate of an ecosystem function, and it is usually challenging to manipulate
+microbiome composition directly. We propose that artificial ecosystem selection
+can overcome these challenges. Artificial ecosystem selection involves creating
+replicate ecosystems, selecting replicates with a desired ecosystem-level
+trait, and using the microbiomes from the selected ecosystems to inoculate a
+new set of ecosystems. We performed artificial ecosystem selection on replicate
+soil microcosms to select for increased soil methane oxidation rate. We
+observed a strong response of methane oxidation rate to artificial ecosystem
+selection with a 51% increase in methane oxidation rate per round over five
+rounds of selection. This suggests that a large portion of the variation in
+methane oxidation rate in these soils can be attributed to microbiome
+variation. We also found that selection did not enrich for known methane
+oxidizers; rather microbes that were enriched after selection included 24
+families such as Puniceicoccaceae, Elsteraceae, and Cytophagaceae. This result
+is in contrast to the typical assumption that the rate of an ecosystem function
+is limited by the final step in the associated microbial pathway. Our study
+demonstrates that artificial ecosystem selection is a useful strategy for
+revealing relationships between microbiome composition and ecosystem function.
+
 # Introduction
 
-One of the fundamental questions in Ecology is To what extent does biodiversity
-drive variation in ecosystem function? Variation in abiotic conditions
-regulates the rate of ecosystem processes, but it remains unclear to what
-extent variation in community structure regulates the rate of ecosystem
-processes. Many ecosystem functions are mediated by organisms. In particular,
-microorganisms such as bacteria mediate the major biogeochemical cycles such
-as the carbon and nitrogen cycles. Despite the central role of bacteria in
-performing these processes, we have not found a consistent relationship between
-microbial community structure and the rate of an ecosystem process. In order
-to understand the importance of biodiversity to ecosystem functioning and to
-better predict and manage specific ecosystem processes, we need to determine
-the relationship between microbial community structure and ecosystem function.
+Microbiomes mediate a variety of important functions in ecosystems, and there
+is great interest in understanding how attributes of microbiomes may influence
+variation in ecosystem functions [@crowther2019]. Biodiversity-ecosystem
+function relationships have been described for a variety of macroorganismal
+communities. For example, plant species richness correlates with productivity
+and marine community diversity correlates with functions such as production,
+consumption, and nutrient release
+[@cardinale2006; @hooper2012; @gamfeldt2015]. While there is evidence that
+variation in the microbiome contributes to variation in ecosystem functions,
+describing this relationship has been elusive [@rocca2015; @graham2016].
 
-In macroorganismal communities, there appears to be a relationship between the
-diversity of a community and the resilience of ecosystem functions to
-environmental perturbations (Tilman, ????... Hooper). Typically, this is the
-diversity (or more specifically, the richness) of a particular functional
-group, such as photosynthetic plants. Similar studies in microbial communities
-rarely find a relationship between diversity and function (Graham, ). However,
-these studies typically define diversity much more broadly than plant diversity
-studies by looking at the phylogenetic diversity of the entire bacterial and
-archaeal community. Studies that focus more narrowly on a particular ecosystem
-process and a functional group involved in that process still rarely find a
-significant correlation between the abundance of a functional group and the
-rate of the corresponding process (Rocca, ????). For certain ecosystem
-functions in certain ecosystems, these relationships can be informative. For
-example, the abundance of a single methanogenic species (*G. species*) is
-correlated with methane emissions in a permafrost ecosystem (cite). In
-addition, bacterial and archaeal community composition is correlated with
-methane emissions in forests and pastures of the Brazilian Amazon (Meyer et al.
-2020).  Therefore, there appears to be some indication that microbial community
-structure drives variation in methane cycling in soils.
+One reason for this difficulty is that it is challenging to directly manipulate
+microbiome composition and diversity the way we can for macroorganismal
+communities. Instead, we must use genetics through marker gene or metagenomic
+sequencing to characterize the microbiome.  However, studies that measure the
+taxonomic or functional diversity of a microbiome through these approaches
+rarely find a correlation between diversity and function [@graham2016]. Another
+approach is to quantify cells or genes using  quantitative
+PCR. Studies such as this still rarely find a correlation between the abundance
+of a functional group and the rate of the corresponding process [@rocca2015].
+As a result, it remains unclear how and when microbiome variation contributes
+to variation in the rate of ecosystem functions.
 
-To further investigate the relationship between microbial community structure
-and the rate of methane cycling in soils, we want to determine which organisms
-are important for regulating methane cycling. This ecological question is
-analogous to the problem of genotype-phenotype mapping in quantitative genetics
-(Morris 2020). Variation in an organismal trait is partially determined by
-environmental conditions and partially determined by genetic variation. One goal
-of quantitative genetics is to determine the proportion of the variation
-attributable to genetics and to identify which alleles regulate the phenotypic
-trait. These goals are commonly achieved through artificial selection
-experiments and association mapping studies, such as genome-wide association
-studies. To determine the mapping between microbial community structure and
-ecosystem function, we applied these approaches of genotype-phenotype mapping
-to whole microbial communities. In this case, the "alleles" are different
-microbial taxa and the "phenotype" is the rate of ecosystem function at the
-whole-ecosystem level.
+Another likely reason that microbiome-ecosystem function relationships have been
+difficult to document is that the most common approaches for looking for a
+relationship requires some prior knowledge of the likely causal links between
+microbiomes and ecosystem functions. For example, many studies have looked for
+relationships with the abundance or diversity of a "marker gene" for a
+particular function. Marker genes are microbial genes that code for an enzyme or
+enzyme-subunit known to be involved in a particular function. But this approach
+assumes that one can identify the "right" marker gene a priori which, given
+how little is understood about microbial diversity, seems highly unlikely. What
+is needed are approaches to microbiome-ecosystem
+function mapping that do not require this degree of prior knowledge.
 
-To identify communities of bacteria and archaea that collectively perform a
-high rate of methane oxidation, we performed an artificial ecosystem selection
-experiment  (Swenson et al. 2000).  We generated twenty-four soil ecosystems in the lab using a sterilized
-potting mix as the substrate. We then inoculated these ecosystems with a small
-amount of living soil to generate variation in community structure across the
-twenty-four ecosystems. These ecosystems were then assigned to one of two
-selection treatments: neutral selection or positive selection. Each ecosystem
-was maintained at 1000 ppm CH~4~ over several weeks to allow for colonization and
-growth of the microbial community and to enrich for methane oxidizers. We then
-determined the methane oxidation rate for each ecosystem. To impose selection,
-we selected three jars from the positive treatment with the highest rate of
-methane oxidation, homogenized them, and used that soil to inoculate the next
-set of jars. For the neutral treatment, we selected three jars at random to
-inoculate the next set of jars. This process continued for five passages. At
-the end of the experiment, we extracted DNA from soils in the second and fifth
-passage for both the neutral and positive treatments. By comparing the response
-to selection within the community between the two treatments, we can identify
-microbial species that contribute to variation in ecosystem methane emissions.
+Yet another likely reason that microbiome-ecosystem function relationships have
+been difficult to document is that microbiomes can be related to ecosystem
+functions in two related but distinct ways that historically have been difficult
+to separate. One way is that microbiomes may simply be conduits through which
+the environment drives function. That is to say that the environmental
+conditions completely regulate the attributes of the microbiome and a shift in
+the environment causes a shift in microbiome attributes, such as the abundance
+of a microbial functional group. This shift results in a change in the rate of
+an ecosystem function. In this case, microbiome attributes would be
+statistically associated with variation in ecosystem function, but also strongly
+covary with environmental conditions. Therefore, there would be no association
+between microbiomes and ecosystem functions independent of the environmental
+conditions. Under this scenario, knowing anything about the microbiome would not
+necessarily provide you with information regarding the causes of the change in
+ecosystem function or improve one's ability to predict changes in ecosystem
+function. Furthermore, attempting to alter ecosystem function by altering the
+microbiome without changing the underlying environmental conditions would likely
+be fruitless.
 
-This approach is powerful because by using a common soil substrate and
-maintaining a constant headspace concentration of methane, we can eliminate
-much of the environmental variation that would be present in an observational
-study. In addition, by applying selection on methane oxidation rate
-in the positive treatment, we can enrich for taxa involved in methane cycling
-making it easier to perform association mapping between microbial taxa and
-ecosystem methane emissions. In this paper, we address whether there is a
-response to selection on methane oxidation rate at the whole ecosystem level. We
-then quantify the amount of variation in methane oxidation attributable to
-variation in community composition within our laboratory environment. Finally,
-we identify microbial markers of soil methane cycling using association mapping
-in order to deduce the mapping between microbial community structure and
-ecosystem function.
+Another possibility is that the microbiome itself alters the rate of ecosystem
+function independent of the environment. In this case, one cannot predict the
+rate of ecosystem function without understanding changes in the microbiome. In
+addition, managing the microbiome through inoculation or selective antibiotics
+would potentially be an effective approach for enhancing or mitigating certain
+changes to ecosystem functions. These two scenarios are not mutually
+exclusive, nor are they likely unique to microbes, but methods to determine the
+relative important of these two scenarios have not until recently been employed.
 
-(Goodnight et al. 1997, Williams and Lenton
-2007)
+The problem of connecting microbiome attributes to the rate of an ecosystem
+function is analogous to the problem of connecting genomic variation to
+phenotypic traits in organisms, and solutions to this problem may be found by
+exploring this analogy [@morris2020]. For example, one approach to the problem of
+genotype-phenotype mapping is to perform artificial selection on a trait of
+interest and identify genes that respond to selection. Such genes are likely to
+be causally linked to the trait under selection. An analogous approach for
+exploring microbiome-ecosystem function relationships is gaining acceptance
+among microbiologists [@swenson2000; @panke-buisse2015]. Microbiomes have been
+shown to respond to selection at the ecosystem-trait level [@swenson2000],
+and some ecosystem-scale microbiome traits have been shown to be transferable
+through microbiome inoculations [@panke-buisse2015]. Applying artificial
+selection to whole microbiomes could be a novel approach for quantifying how
+much of the variation in the rate of an ecosystem function could be attributed
+independently to microbiome variation [@goodnight2000]. In addition, this
+approach could be used to identify microbiome attributes, such as genes or taxa,
+associated with the rate of an ecosystem function, although, to my knowledge
+this has not previously been accomplished [@arias-sanchez2019; @morris2020]. 
+Such an approach has the potential to expand our ability to model
+microbial ecosystem-function relationships more accurately and allow us to
+manage ecosystems for particular outcomes [@fuhrman2009; @schimel1998]. Here I
+use this approach to estimate how much of the variation in ecosystem function
+(the flux of CH~4~ gas between the soil and atmosphere) can be attributed to
+variation in the microbiome and to identify microbial markers associated with
+this ecosystem function.
 
-# Materials & Methods
+Soil CH~4~ oxidation is a suitable function for this study because there is
+evidence that it may vary with microbiome attributes. For example, variation in
+CH~4~ emissions in arctic permafrost is correlated with the transcriptional
+activity of certain methanogens [@freitag2009; @freitag2010]. In addition,
+soil microbiome phylogenetic variation is a strong predictor of CH~4~ oxidation in
+forests and pastures of the Brazilian Amazon [@meyer2020]. Finally,
+methanogenesis and methanotrophy are two of the most deeply conserved microbial
+physiologies and are represented in a narrow range of taxa and so the taxonomic
+composition of the microbiome is more likely to be associated with the rate of
+CH~4~ flux than other broader or more shallowly conserved functions
+[@schimel1998; @martiny2013]. This suggests that microbiome variation might
+drive variation in CH~4~ oxidation in soils.
 
-## Selection experiment
-We performed an artificial ecosystem selection experiment similar to Swenson et
-al. (2000) by passaging replicate soil microbial communities. The trait we
-selected on was CH~4~ oxidation rate. Our experiment had two selection lines
-with twelve jars each for a total of twenty-four jars per passage. One line was
-a positive selection line where the two or three jars with the highest CH~4~
-oxidation rate were chosen to inoculate the next set of positive jars. The
-other line was a neutral selection line where an equal number of jars as the
-positive line were chosen at random to inoculate the next set of neutral jars.
-The number of jars chosen was based on the distribution of fluxes among the
-positive jars, i.e., we chose the top three jars unless only two jars had
-considerably greater oxidation rates based on visual inspection of histograms.
-The experiment was carried out over five passages.
+For the current project, we applied artificial ecosystem selection on soil
+microbiomes by selecting on the ecosystem-scale CH~4~ oxidation rate in order
+to estimate the amount of microbiome variation underlying variation in soil CH~4~ oxidation rate
+[@swenson2000]. To identify which microorganisms could be markers of CH~4~
+oxidation rate, we compared the composition of the artificially selected
+microbiomes to a control set of microbiomes without selection. Artificial
+ecosystem selection has a similar effect to enrichment culturing by amplifying
+the population of interest [@swenson2000]. This will reduce the diversity of
+the soil microbiome and allow for greater power in detecting significant
+markers of CH~4~ oxidation rate. We then evaluated whether these markers meet
+our underlying assumptions about which taxa limit the rate of ecosystem
+function.
 
-The initial soil microbial community was sampled from the top 10 cm of an
-upland mineral soil under a deciduous forest ecosystem near the University of
-Oregon campus in Eugene, OR, USA. Incubations were performed in 500 mL mason
-jars (Ball Corporation, Broomfield, Colorado, USA) with rubber septa installed
-in the lids. Each jar was sterilized with 70% ethanol to which was added 45 g
-of autoclaved potting mix, 5 g of living soil, and 3.5 mL of sterile deionized
-water to bring them to <mark> 60%??? of field capacity</mark>. The jars were
-then homogenized, capped, and injected with 4.3 mL of 99% CH~4~ to bring the
-headspace concentration to approximately 1000 ppm CH~4~. To create the two
-treatment groups, twenty-four jars were created in an identical manner and then
-randomly assigned to either the positive or neutral selection line. Jars were
-flushed and respiked twice per week to maintain elevated CH~4~ concentrations
-and were incubated at ambient temperature for approximately three weeks.
-Methane oxidation rates were determined at the end of the incubation period.
-These values were then used to determine the jars selected to inoculate the
-next set of jars.
+# Materials and Methods
 
-## Methane oxidation rate measurements
+## Experimental design
+
+We performed an artificial ecosystem selection experiment [sensu @swenson2000]
+by passaging replicate soil microbiomes. The trait we selected on was
+CH~4~ oxidation rate. This experiment had two selection lines with twelve jars
+each for a total of twenty-four jars per passage. One line was a positive
+selection line where the two or three jars with the highest CH~4~ oxidation rate
+were chosen to inoculate the next set of positive jars. The other line was a
+neutral selection line where an equal number of jars as the positive line were
+chosen at random to inoculate the next set of neutral jars. The number of jars
+chosen was based on the distribution of fluxes among the positive jars, i.e., we
+chose the top three jars unless only two jars had considerably greater CH~4~ oxidation
+rates based on visual inspection of histograms. The experiment was carried out
+over five passages until a significant divergence in functional rates was 
+observed between the two selection treatments.
+
+The initial soil microbiome was sampled from the top 10 cm of an upland mineral
+soil under a deciduous forest ecosystem near the University of Oregon campus in
+Eugene, OR, USA. Incubations were performed in 500 mL mason jars with rubber
+septa installed in the lids. Each jar was sterilized with 70\% ethanol to which
+was added 45 g of autoclaved potting mix, 5 g of living soil, and 3.5 mL of
+sterile deionized water to bring them to 60\% of field capacity. The potting mix was
+Lane Potting Mix from Lane Forest Products (Eugene, OR, USA). The jars were
+then homogenized, capped, and injected with 4.3 mL of 99\% CH~4~ to bring
+the headspace concentration to approximately 1000 ppm CH~4~. To create the
+two treatment groups, twenty-four jars were created in an identical manner and
+then randomly assigned to either the positive or neutral selection line. Jars
+were flushed and respiked twice per week to maintain aerobic conditions and
+elevated CH~4~ concentrations and were incubated at ambient temperature for
+approximately three weeks. Methane oxidation rates were determined at the end
+of the incubation period. For the positive treatment, the three jars with the
+greatest CH~4~ oxidation rates were chosen to inoculate the next generation.
+For the neutral treatment, three jars were randomly selected to inoculate the
+next generation. For each treatment, these three jars were homogenized and 5 g 
+of this homogenized soil was used to inoculate the next set of jars and which
+represents a 10% subsampling. The
+next set of jars were created in an identical manner to the first generation
+with fresh autoclaved potting mix and the same moisture and CH~4~ content.
+
+## Methane oxidation rate
 
 Methane oxidation rates were determined after flushing and spiking jars to
-approximately 1000 ppm CH~4~. Headspace samples of 1 mL were collected from
-each jar immediately after spiking and then at time points 3, 6, 24, and 48
-hours for a 5-point curve. Samples were immediately injected into a
-<mark>???</mark> gas chromatograph fitted with an electron capture detector
-(company, city, state, USA) to determine the headspace CH~4~ concentration.
-Fluxes were calculated from a first-order exponential decay function as decay
-constant `k` with units day^-1^. Oxidation rates are presented as the additive
-inverse of `k` (i.e., `-k`) so that a more positive value represents a faster
-oxidation rate. The jars selected for the positive treatment in passage 2 had
-the lowest methane oxidation rate of the twelve jars due to a calculation error
-in the methane oxidation rate. All other passages correctly used the top three jars.
+approximately 1000 ppm CH~4~. Headspace samples of 1 mL were collected from each
+jar immediately after spiking and then at time points 3, 6, 24, and 48 hours for
+a 5-point curve. Samples were immediately injected into a SRI model 8610C gas
+chromatograph equipped with a flame ionization detector (SRI Instruments, Torrance, CA, USA) to determine the headspace CH~4~
+concentration. Fluxes were calculated from a first-order exponential decay
+function as decay constant $k$ with units day^-1^. Oxidation rates are presented
+as the additive inverse of $k$ (i.e., $-k$) so that a more positive value
+represents a greater oxidation rate. The jars selected for the positive
+treatment in passage 2 had the lowest CH~4~ oxidation rate of the twelve jars
+due to a calculation error in the CH~4~ oxidation rate. All other passages
+correctly used the top three jars.
 
@@
 A subsample of soil from the starting inoculum and from every jar in passages 2
-and 5 was collected and stored at -80$^\circ$C for later DNA extraction.
-Soil DNA was extracted from a 0.25 g subsample of soil. Negative controls were extracted
-from autoclaved potting mix and DNase-free water. Extractions were performed
-using the DNeasy PowerSoil kit (QIAGEN, Hilden, Germany) and quantified using
-Qubit (company, city, state). In order to estimate the diversity and relative
-abundance of the bacterial and archaeal taxa in our soil ecosystems, we
-sequenced the V3-V4 region of the 16S rRNA gene using the Earth Microbiome
-Project protocol and primers<mark>???</mark>. Sequencing was performed on the
-Illumina NovaSeq 6000 with paired-end 150 bp reads.
+and 5 was collected and stored at $-80^\circ$C. Soil DNA was extracted from 0.25
+g soil. Negative controls were extracted from autoclaved potting mix and
+DNase-free water. Extractions were performed using the DNeasy PowerSoil kit
+(QIAGEN, Düsseldorf, Germany) and quantified using Qubit dsDNA HS Assay Kit (Thermo Fisher
+Scientific, Inc., Waltham, MA, USA). To estimate the diversity and relative abundance of
+the bacterial and archaeal taxa in our soil ecosystems, we sequenced the V4
+region of the 16S rRNA gene using the using the 515F - 806R primer combination
+[@caporaso2011]. PCR mixtures were: 10 $\mu$l NEBNext Q5 Hot Start HiFi PCR
+master mix, 9.2 $\mu$l primer mixture (1.09 $\mu$M concentration), and 0.8
+$\mu$l of DNA template. Reaction conditions were: $98^\circ$C for 30 s
+(initialization); 35 cycles of $98^\circ$C for 10 s (denaturation), $61^\circ$C
+for 20 s (annealing), and $72^\circ$C for 20 s (extension); and $72^\circ$C for
+2 m (final extension). Reactions were performed in triplicate and then combined.
+Amplicons were purified twice using 0.8x ratio Mag-Bind
+RxnPure Plus isolation beads (Omega Bio-Tek, Norcross, GA, USA). Sequencing libraries were
+prepared using a dual-indexing approach [@fadrosh2014; @kozich2013]. Amplicon
+concentrations were quantified using Qubit and multiplexed at equimolar concentration. Sequencing
+was performed at the University of Oregon Genomics Core Facility on the Illumina
+NovaSeq 6000 with paired-end 150 bp reads (Illumina, Inc., San Diego, CA, USA).
 
@@
 
-Bioinformatic analyses were conducted using the `QIIME2`
-bioinformatics platform (Bolyen et al. 2019). Raw sequencing reads were
-demultiplexed using ??? (citation), denoised using `DADA2` (citation).
-Taxonomic assignment for the 16S reads was performed using ??? based on the
-Silva database. All subsequent tatistical analyses were performed in the `R`
-statistical programming environment (R Core Team 2020). Richness estimates were
-made using the `breakaway` package and diversity indices using the `DivNet`
-package (Willis et al. ???).
+Bioinformatics processing was performed in 'R' [@rcoreteam2018].
+Demultiplexed sequencing reads were denoised using 'DADA2' to generate a table
+of amplicon sequence variants (ASVs)
+[@callahan2016]. Taxonomic assignment for the 16S reads was performed
+using the Ribosomal Database Project naive Bayesian classifier [@wang2007]. The presence of
+contaminants was evaluated using both the prevalence and frequency methods from
+'DECONTAM' by comparing samples to negative controls [@davis2018]. We did not identify
+any obvious contamination using these methods.
 
@@
 
-### Response to selection
-
-To test whether there was a significant change in CH~4~ oxidation rate (k) as a
+Statistical analyses were performed in 'R' [@rcoreteam2018]. To test
+whether there was a significant change in CH~4~ oxidation rate ($k$) as a
 response to selection, we tested a difference in slopes between the positive
-and neutral selection lines. The CH~4~ oxidation rates were strongly right
-skewed with most values close to zero and few large, positive values. This
+and neutral selection lines. The CH~4~ oxidation rates were strongly 
+right-skewed with most values close to zero and few large, positive values. This
 resulted in residuals that did not meet the assumptions of constant variance
-and normal distribution. Therefore, methane oxidation rates were log~10~
+and normal distribution. Therefore, CH~4~ oxidation rates were log~10~
 transformed to better meet the assumptions of a linear model and to make
 figures easier to interpret. First, we tested the effect of treatment by
-fitting two nested models with and without treatment using `lm` in `R`. The
-model formulation for the full model was `flux ~ passage * treatment` and the
-model formulation for the reduced model was `flux ~ passage`. We compared these
-models using the likelihood ratio test with the `anova` function in `R`. We
-then fit the full model to determine the slope of the positive line, which
-represents the change in methane oxidation rate per passage as a response to
+fitting two nested models with and without treatment using 'lm'. We compared these
+models using the likelihood ratio test with the 'anova' function. We
+then fit the full model to estimate the slope of the positive line, which
+represents the change in CH~4~ oxidation rate per passage as a response to
 selection.
 
-### Ecosystem heritability
+To estimate the proportion of variance in ecosystem function due to variation in
+the microbiome, we estimated narrow-sense heritability ($h^2$) as the
+regression of mid-offspring on mid-parent [@falconer1996]. The mid-parent was the
+mean for all three selected jars and the mid-offspring was the mean for all jars
+produced by those parents. First, we tested if there was an effect of treatment
+on the heritability estimate. We compared these models using the likelihood ratio
+test with the 'anova' function. We then fit the full model to estimate the
+heritability of the Positive and Neutral treatments.
 
-To estimate the proportion of variance in ecosystem function due to additive
-genetic variance of the microbial communtiy, we applied the Breeder's equation
-to the results of our selection experiment. the heritability of
-CH~4~ oxidation rate as an ecosystem property, we calculated narrow-sense
-heritability ($h^2$) as the regression of offspring on mid-parent (Goodnight
-2000, Falconer 1997). The mid-parent was the mean for all three selected jars
-and the offspring was the mean for all jars produced by those parents.
-
-In addition, we calculated the inbreeding coefficient, $F$, as:
-
-$$
-F = 1 - \left(1 - \frac{1}{2N} \right) ^{T}
-$$
-
-where $N$ is the population size and $T$ is the number of generations (Goodnight 2000 or Falconer 1997).
-
-### Taxa differentially abundant in the positive selection treatment
+Richness estimates and tests were performed using 'breakaway' [@willis2015]. We
+tested a difference in richness between Passages 2 and 5 with both treatments
+combined. Then we subset the samples from Passage 5 and compared richness between
+the Positive and Neutral treatment. Next, we compared beta-diversity as both
+Aitchison dissimilarity, which is the Euclidean distance after center-log ratio
+transformation, and the binary Jaccard distance for presence or absence [@aitchison1982]. 
+We tested a difference in centroid and dispersion of beta
+diversity between the Positive and Neutral treatment and Passage 2 and 5 with
+permutational analysis of variance (PERMANOVA) using the 'adonis2' function from
+'vegan' [@mcardle2001; @oksanen2019].
 
 To identify taxa that responded to selection on CH~4~ oxidation rate, we tested
-differential abundance using a beta-binomial model (Martin et al. 2020). We
-compared taxa in passage 5 between the positive and neutral treatment to
-identify taxa that
-
-Possible tests:
-
-- Random forest or gradient boosting to predict Positive vs. Neutral.
+differential abundance using a beta-binomial model [@martin2020]. We first
+grouped ASVs at the family level keeping ASVs that lacked a family-level taxonomic
+assignment. We
+then subset the samples in Passage 5 and removed all families with a prevalence
+of less than 10\%. We compared families in passage 5 between the Positive and
+Neutral treatment to identify taxa that were enriched or depleted as a response
+to selection. we used the likelihood ratio test to estimate p-values with a
+significance threshold of false-discovery rate < 0.05 [@martin2020]. Estimates
+are presented as the expected relative abundance in the positive treatment
+relative to the neutral treatment.
 
@@
 
-## Response to selection
+We observed a response to selection on whole-ecosystem soil CH~4~ oxidation
+rate (`r figures("response", display = "cite")`). In addition, the
+response to selection varied with treatment 
+`r print_lrt(compare_response_mods, 2, 'LRT of nested models with and without treatment')`. At the start of the experiment, the Positive
+treatment had a mean CH~4~ oxidation rate that was
+`r percent_less(initial = tidy(response_model)[[1, 2]], final = tidy(response_model)[[3, 2]] + tidy(response_model)[[1, 2]])`\% lower than the
+Neutral treatment `r print_lm(response_model, 3, 'difference of y-intercepts')`. 
+There was no change in CH~4~ oxidation rate of the Neutral
+treatment over the five passages `r print_lm(response_model, 2)`. By contrast,
+the Positive treatment had a `r round(per_slope)`\% increase in CH~4~
+oxidation rate per passage `r print_lm(response_model, 4)`.
 
-We observed a response to selection on soil methane oxidation rate at the whole
-ecosystem level (`r figures("response", display="cite")`).
-At the start of the experiment, the Positive
-treatment had a mean methane oxidation rate that was
-`r percent_less(initial = tidy(response_model)[[1, 2]], final = tidy(response_model)[[3, 2]] + tidy(response_model)[[1, 2]])`%
-lower than the
-Neutral treatment (SE = 
-`r myround(tidy(response_model)[[3, 3]], 2)`, t = 
-`r myround(tidy(response_model)[[3, 4]], 2)`, p = 
-`r myround(tidy(response_model)[[3, 5]], 2)`). There was no change in methane oxidation rate of the Neutral
-treatment over the five passages (slope = 
-`r myround(tidy(response_model)[[2, 2]], 2)`, SE = 
-`r myround(tidy(response_model)[[2, 3]], 2)`, t = 
-`r myround(tidy(response_model)[[2, 4]], 2)`, p = 
-`r myround(tidy(response_model)[[2, 5]], 2)`). By contrast, the Positive treatment had a
-`r round(per_slope)`% increase in methane
-oxidation rate per passage (slope = 
-`r myround(tidy(response_model)[[4, 2]], 2)`, SE = 
-`r myround(tidy(response_model)[[4, 3]], 2)`, t =
-`r myround(tidy(response_model)[[4, 4]], 2)`, p = 
-`r myround(tidy(response_model)[[4, 5]], 2)`).
-
-```{r response_figure, fig.width=4, fig.height=3, fig.cap=figures("response")}
+```{r response, fig.width=4, fig.height=3}
 ratio_log10 <- calc_plot_ratios(fluxes$passage, fluxes$log_ch4)
 
-plot_flux(fluxes, passage, log_ch4, ratio_log10, log10 = TRUE)
+p <- plot_flux(fluxes, passage, log_ch4, ratio_log10, log10 = TRUE)
+
+ggsave('fig1.jpg', p)
 
@@
 
-## Ecosystem heritability
+We estimated heritability as the regression of mid-offspring on mid-parent (`r figures("herit", display = "cite")`).
+Offspring CH~4~ oxidation rates were correlated with parental CH~4~ oxidation
+rates in both the Positive treatment `r print_lm(herit_model, 4)` and the
+Neutral treatment `r print_lm(herit_model, 2)`. Notably, the sign of the effect
+was reversed with a positive heritability for the Positive treatment and a negative heritability for the Neutral treatment.
 
-Here, I fit two models. The first is simply offspring flux on selected parent
-flux, ignoring treatment. There is no correlation. Next, I added treatment and the interaction between treatment and parental
-flux. Here, there is a significant correlation between parent and offspring for
-both the Neutral treatment `r print_model(herit_model, 2)` and the Positive
-treatment `r print_model(herit_model, 4)`. Notably, the sign of the effect is
-reversed: negative for the Neutral treatment and positive for the Positive
-treatment.
+```{r herit, fig.width=5, fig.height=3}
 
-The negative heritability of methane oxidation rate in the Neutral selection
-treatment may result from selection on alternative characters in those
-populations. The strong, positive heritability in the Positive treatment ... .
-It seems strange that the Methane oxidation rate was heritable between
-consecutive passages (`r figures("herit", display="cite")`). Specifically, the
-narrow-sense heritability, calculated as the slope of the mid-parent and
-mid-offspring, was $h^2$ =  $\pm$  (t = , p = ). With an inbreeding coefficient of ,
+p <- plot_herit(heritability)
 
-```{r herit_fig, fig.width=5, fig.height=3, fig.cap=figures("herit")}
-plot_herit(heritability)
+ggsave('fig2.jpg', p)
+
 ```
+
+Sequencing of 16S rRNA amplicons revealed that there were `r ntaxa(ps)` unique
+ASVs across all `r nsamples(ps)` jars sampled.
+Richness of ASVs decreased among all of the jars between Passage 2 and 5
+(Richness difference between Passage 5 and Passage 2 = `r myround(bt_passage$table[2, 1], 1)`, SE =
+`r myround(bt_passage$table[2, 2], 2)`, p < 0.001). However, there was no
+difference in richness between the Positive and the Neutral treatment in
+Passage 5 (Difference = `r myround(bt_treat5$table[2, 1], 1)`, SE =
+`r myround(bt_treat5$table[2, 2], 2)`, p =
+`r myround(bt_treat5$table[2, 3], 3)`).  In addition, there was no
+correlation between richness and CH~4~ oxidation rate across all passages
+and treatments.  (slope = `r myround(bt_ch4$table[2, 1], 1)`, SE =
+`r myround(bt_ch4$table[2, 2], 2)`, p = `r myround(bt_ch4$table[2, 3], 3)`).
+
+Taxonomic dissimilarity of the soil microbiome varied strongly by passage and
+weakly by treatment (`r figures("beta", display = "cite")`). Permutational
+analysis of variance demonstrated an effect of passage, treatment,
+and their interaction on Aitchison dissimilarity (`r tables("adonis", display = "cite")`). 
+Passage explained `r myround(100*fit_adonis[1, 4])`\% of the
+variation in Aitchison dissimilarity and `r myround(100*fit_adonis[6, 4])`\% in
+Jaccard dissimilarity.  Treatment explains `r myround(100*fit_adonis[2, 4])`\%
+for Aitchison and `r myround(100*fit_adonis[7, 4])`\% for Jaccard.  Variation
+in microbiome composition decreased as a result of passaging in jars (F = 80.5,
+p < 0.001), but did not differ between selection treatments (F = 0.40, p =
+0.54).
+
+```{r beta, results=FALSE, fig.height=2.5}
+
+p <- plot_beta(ps)
+
+ggsave('fig3.jpg', p)
+
+```
+
+```{r adonis}
+
+write.table(fit_adonis, 'table_1.tsv')
+# kable(col.names = c('Term', 'df', 'SS', 'R2', 'F', 'p'), 
+#       digits = c(2,2,2,2,2,3),
+## 	format = "Table",
+## 	booktabs = T,
+# 	caption = "PERMANOVA on Bray-Curtis and Jaccard dissimilarities.",
+#   	label = "beta") %>%
+#   pack_rows("Bray-Curtis", 1, 5) %>%
+#   pack_rows("Jaccard", 6, 10)
+
+```
+
+To investigate which taxa responded to selection on soil CH~4~ oxidation rate, we
+first looked at which taxa were unique to the Positive and Neutral treatments
+aggregated at the family level. Several taxa were present in the Positive
+treatment and absent in the Neutral treatment. Most of these families had low
+prevalence (present in 3 or fewer samples) and low abundance (median < 5 reads).
+There were two families unique to the Positive treatment with relatively high
+prevalence. This included an ASV that was a member of the Bacteroidia Class with
+no lower taxonomic designation. This ASV had a prevalence of 10/12 and a median
+abundance of ~2 reads. The other prevalent family was a member of the
+Silvanigrellaceae, a newly described family placed in its own order. This family
+was present in all 12 samples and had a median abundance of 8 reads.
+Silvanigrella is the only cultivated member of
+Silvanigrellaceae and was isolated from a temperate fresh water lake
+[@hahn2017]. Of the families unique to the Neutral treatment, only one had a
+prevalence greater than 2/12. This family, Armatimonadaceae, was present in half
+of the Neutral samples (prevalence = 6/12). The type strain for Armatimonadaceae was isolated from
+the rhizosphere of *Phragmites australis* [@tamaki2011].
+
+The remaining families were shared between the Positive and Neutral treatment,
+but some differed in their relative abundance. To identify taxa that responded
+to selection on soil CH~4~ oxidation rate, we tested the differential relative
+abundance of families in the Positive jars relative to the Neutral jars within
+Passage 5 using a beta-binomial model. We identified `r ntaxa(sig_taxa_treat)`
+families that were enriched or depleted in the Positive treatmeant relative to
+the Neutral treatment (`r figures("da", display = "cite")`).
+
+```{r da, results=FALSE}
+
+p <- dt_plot(dt_data)
+ggsave('fig4.jpg', p)
+
+```
+
+Overall, none of the families enriched in the Positive selection treatment
+are known methanotrophs. Several taxa identified had a 
+higher taxonomic designation that contains methanotrophs, for example, the
+Gammaproteobacteria class had a large effect size.
+This class is one of the groups containing methanotrophic families such as
+Methylococcaceae. However, gammaproteobacteria is among the most diverse 
+groups in the Prokaryotes, so this is not strong evidence for a selection response
+by methanotrophs [@garrity2005].
+The next family
+was Puniceicoccaceae, which is a member of the phylum Verrucomicrobia and
+contains marine microbes. The Verrucomicrobia is a diverse group that contain
+known methanotrophs as well as ammonia-oxidizing bacteria [@freitag2003]. 
+An ASV from the order Armatimonadales was also enriched in the Positive treatment.
+This order also contains the family Armatimonadaceae, which was
+found to be unique to the Neutral treatment. Cytophagaceae was also enriched in
+the Positive treatment and contains a number of mainly aerobic heterotrophs
+that can digest a variety of macromolecules [@mcbride2014]. Notably, the family
+Gemmatimonadaceae has a single bacterial bin, Candidatus '*Methylotropicum kingii*',
+that contains methanotrophy genes, but this family was depleted in the Positive
+selection treatment [@bay2021].
 
@@
 
-## Future Directions
+We observed a response to
+selection on whole-ecosystem soil CH~4~ oxidation rate. This type of response
+has been observed for other functions, such as chloroaniline degradation in
+water [@swenson2000], but this is the first instance, to our knowledge, of
+selection being performed directly on soil biogeochemical cycling. With this
+experimental design, we attempted to create multiple ecosystems that were very
+similar with respect to the type and quantity of soil, soil moisture content,
+and the headspace concentration of CH~4~. we further controlled for environmental
+variation across replicates by having twelve replicates per treatment, using
+controlled laboratory conditions, and by having a random selection treatment as
+a control. Therefore, the observed response to selection on
+CH~4~ oxidation rate is likely due to changes in the microbiome.
 
-This study focused on the practical implications of using AES as a tool to deduce how microbes map onto functions at the whole-ecosystem level.
-If one were interested in the evolutionary dynamics underlying this process, one could design a study to distinguish between individual selection on microbial strains and group selection or multilevel selection.
-That is, selection that cannot be recapitulated through selection at the individual level (Williams and Lenton 2007).
-This could be accomplished by culturing differentially abundant taxa in the positive selection treatment and then inoculating them in isolation (or in a non-selected background community) to see if the same level of ecosystem function could be recapitulated as in the final generation of the positive selection treatment.
+To investigate the relationship between microbiome structure and
+ecosystem function, we wanted to determine how much of the variation in soil CH4
+oxidation was attributable to variation in the microbiome. This ecological
+question is analogous to the problem of estimating heritability in quantitative
+genetics [@morris2020]. Variation in an organismal trait is
+determined by the sum of genetic variation, environmental variation, and the
+interaction between the two [@falconer1996]. One goal of quantitative
+genetics is to determine the proportion of phenotypic variation attributable to
+genetic variation. This is commonly estimated as the narrow-sense heritability
+defined as the additive genetic variance. This goal can be achieved through
+artificial selection experiments with selection on the character of interest
+[@falconer1996].
 
-Another avenue for research could be to look at correlated responses of ecosystem-level traits to selection.
-This would address questions such as, "Is methane oxidation rate correlated with other ecosystem processes due to pleiotropy or linkage disequilibrium?" At the community level, pleiotropy might be defined as "the production by a single taxon of two or more unrelated effects." Further, linkage disequilbrium would represent correlated traits due to taxa who tend to co-associate due to shared environmental preferences or coupled metabolic pathways.
+The heritability we observed in the Positive treatment was large ($h^2$ =
+`r myround(tidy(herit_model)[4, 2], 2)`) relative to other studies of biological
+heritability [@visscher2008]. However, given the small sample
+size and wide confidence intervals (95\% CI: 
+`r myround(confint(herit_model)[4, ], 2)`) it is likely that a larger study could
+more precisely estimate the true heritability. We intentionally eliminated much of the
+environmental variation that would be present in natural ecosystems, but it is
+unclear how much jar-to-jar variation there is in the composition of the soil
+substrate. Given these data, though, we conclude that variation within
+the soil microbiome has a very strong effect on soil CH~4~
+oxidation rates independent of any environmental variation.
 
-These experiments would further help lus understand the nature of the "map" between community structure and ecosystem function.
-This would not only reveal how best to predict ecosystem function in earth system models, but also demonstrate the mechanism underlying that relationship.
-This is akin to deducing the genetic control of trait variation in molecular biology.
+The negative heritability observed in the Neutral treatment could indicate a
+negative genetic correlation between CH~4~ oxidation rate and traits
+associated with persistence in a jar [@conner2003]. The conditions in the
+jar favor organisms that can grow in potting mix, reproduce within a three week
+incubation, and survive passaging between jars. These traits might be
+negatively correlated with a community's ability to oxidize CH~4~ indicating
+a potential trade-off [@conner2003]. This trade-off would result in a
+negative heritability of CH~4~ oxidation rate in the Neutral treatment,
+which did not experience selection on CH~4~ oxidation rate.
 
-Community epistasis is analogous to functional redundancy.
+Since there was a response to selection, we wanted to identify
+which taxa responded to selection. That is, the taxa that
+were enriched in the Positive treatment relative to the Neutral treatment. 
+Studies of microbiome structure-function
+relationships typically quantify marker genes for the final enzyme in a pathway
+to try to predict the rate of flux through that pathway. However, of the 
+`r ntaxa(sig_taxa_treat)` taxa identified by the differential abundance test, none
+were members of a family known to oxidize CH~4~. This suggests
+that, in this system, the taxa that perform the final step in the CH~4~ oxidation pathway
+do not
+limit the rate of flux through the pathway. Traits that we assume are important
+for regulating function (such as the diversity or relative abundance of
+methanotrophs estimated from phylogenetic marker genes) may not be important for
+determining variation in function.
+Perhaps once a sufficient number of methanotrophs are present, the
+process itself is limited by other metabolic processes in the ecosystem, such
+as nitrogen cycling mediated by non-methanotrophic organisms.
+
+We found that `r nrow(dt_data[dt_data$x > 0,])` taxa were enriched in the Positive treatment,
+which might indicate that the soil CH~4~ oxidation rate in these jars is controlled by
+multiple taxa. 
+This conclusion is analogous to the conclusion in quantitative genetics that
+most traits at the organismal level are the result of the interaction between
+many independent genes. Indeed, very few traits or genetic diseases are the
+result of a single gene or mutation [@kerem1989; @macdonald1992]. This results in the use of polygenic risk scores for predicting
+phenotype or disease risk in humans [@reich2001; @hindorff2009]. An analogous conclusion can be drawn for predicting
+the rate of ecosystem function from the presence or relative abundance of 
+microbial taxa. The rate of an ecosystem function at the whole-ecosystem level is 
+the result of interactions among a variety of disparate taxa with different traits.
+To better understand microbiome structure-function relationships, our results suggest that we
+should move away from single marker genes and instead investigate the role of
+multiple functional groups in determining the function of ecosystems whether it
+is CH~4~ flux from soil or host-microbiome health.
+
+Taxonomic richness dropped precipitously
+between Passage 2 and Passage 5. There are several explanations for this drop
+in diversity. The initial community was a diverse microbiome sampled from a natural soil.
+With the initial inoculation and at
+each subsequent passaging, the ecosystem was subsampled to 10\% (5 g of living
+soil was combined with 45 g of sterilized substrate). This subsampling likely
+explains a large part of the drop in diversity. In addition, beyond the imposed
+selection regime based on CH~4~ oxidation rate, these soil microbiomes were under
+a variety of selection pressures such as persistence in a jar, survival during the transfer
+from one jar to the next, and the ability to colonize a new jar in a couple of
+weeks, to name a few. These selection pressures would have further filtered
+the microbiome for taxa that could persist in this laboratory environment.
+
+In addition to the drop in richness, there was also a decrease in beta
+diversity among the jars as a result of passaging. This could be due to the
+homogenization step between each passage. Once the jars were selected to generate
+the next set of jars, the soil from those jars was homogenized and this homogenate
+was used to inoculate the next set of jars. The goal of this step was to "shuffle"
+membership in the microbiomes among the jars in order to test different
+combinations and relative abundances of taxa in terms of their effect on CH~4~
+oxidation rate. This likely also made the communities more similar in their
+membership by increasing the likelihood that each taxon was represented in each
+jar. Another explanation for this biotic homogenization is the aforementioned
+selection for persistence in a jar. These selection pressures would have 
+favored the subset of taxa suited to these laboratory conditions regardless
+of whether they were in the Positive or Neutral selection line. This can be
+seen by the relative similarity of the Positive and Neutral jars in Passage 5.
+Despite the Positive treatment undergoing selection on CH~4~ oxidation rate,
+both the Positive and Neutral jars were experiencing selection on a variety
+of other traits that made the overall soil microbiomes appear similar.
+
+One caveat that is important to keep in mind is that 16S rRNA genes are not
+markers of functional characteristics of microorganisms. They are simply
+phylogenetic markers useful for determining the relative position of taxa in a
+phylogenetic tree. In addition, assigning functions to taxa based solely on
+their taxonomy is a fraught exercise because many microbial functions are not
+very deeply conserved, are spread across disparate lineages of the tree of life,
+or are easily transferred between distantly related taxa. However,
+methanotrophy and methanogenesis are two of the most deeply conserved microbial
+traits [@martiny2013]. This results in part from the fact that
+these functions require multiple genes to perform and are therefore not easy to
+evolve independently or to transfer between lineages through horizontal gene
+transfer. However, in order to be confident that the response we observed was not
+simply due to a relative increase in methanotrophs, we would need to apply
+a functional approach to characterizing the microbiome, for example, by sequencing
+whole metagenomes.
+
+Future research in this area should use methods that are able to detect
+functions and metabolic pathways within the community that respond to
+selection. For example, genome-resolved metagenomics could identify pathways
+that are enriched in the Positive selection treatment. This would further
+advance our understanding of the exact traits that are enriched in high-flux
+ecosystems and therefore could be markers of ecosystem function. In addition,
+future studies could build on experimental results such as these to sample a
+natural gradient of ecosystem CH~4~ fluxes to determine whether the taxa identified
+in our selection experiment could be useful indicators of CH~4~ flux in the
+field. These studies would allow us to build a picture of the multiple traits
+that drive variation in ecosystem function at the whole-ecosystem level.
+
+# Conclusion
+
+We performed an artificial ecosystem selection experiment on whole-ecosystem
+CH~4~ oxidation rate to determine whether variation in microbiome
+composition can contribute to variation in ecosystem function. We observed
+a significant response to selection on CH~4~ oxidation rate - the first
+example of an experiment performing selection on soil biogeochemical cycling.
+In addition, we observed strong heritability of CH~4~ flux between
+passages, suggesting that variation in microbiome composition could be
+a major source of variation in CH~4~ flux in ecosystems. Surprisingly, 
+the taxonomic groups that experienced selection on CH~4~
+oxidation rate were not enriched in methanotrophs. This suggests that
+understanding a single functional group is insufficient for predicting the
+effect of microbiome composition on the rate of ecosystem function.
+Instead, we may need to investigate alternative functional groups or multiple
+interacting taxa to understand the role of microbiomes in ecosystem function. While it has often been assumed that
+microorganisms play a major role in determining variation in ecosystem function, 
+this has not previously been shown empirically. This experiment demonstrates that
+variation in microbiome composition can contribute to considerable variation
+in ecosystem function.
 
 # Acknowledgments
+
 This project was supported by the National Science Foundation Graduate Research Fellowship Program (grant no. DGE 1255832) and the ARCS Foundation Florence and Mike Nudelman Scholarship.
@@
 # Competing Interests
+
 We declare we have no competing interests.
 
+# Data Availability Statement
+
+The 16S rRNA sequencing data generated during the current study are available in the NCBI Sequence Read Archive (SRA).
+
+# Supplement
+
 # References
+
+::: {#refs}
+:::
+
+# Figures
+
+`r figures("response", "Response to selection on soil CH~4~ oxidation rate fit by ordinary least squares regression.
+The y-axis is the log~10~ of the additive inverse of the first-order exponential decay constant $k$ (i.e., $-k$) with units day^-1^.
+The x-axis is the passage number.
+The orange line and points are the Positive treatment and the gray line and points are the Neutral treatment.")`
+
+`r figures("herit", "Ordinary least squares regression of mid-offspring CH~4~ flux on mid-parent CH~4~ flux ($-k$ with units day^-1^).
+Mid-parent is the mean of the jars selected to inoculate the next passage.
+Mid-offspring is the mean of all twelve jars produced in one passage.
+The orange lines and points are the Positive treatment and the gray lines and points are the Neutral treatment")`
+
+`r figures("beta", "Principal component analysis plot of beta diversity for all jars.
+Dissimilarities are based on Aitchison dissimilarity, which is Euclidean distance after center-log ratio transformation.
+Colors represent treatment and shapes represent passage number.
+The ratio of the axes is proportional to the variance explained by each principal component in order to accurately represent the distance between the points.")`
+
+`r figures("da", "Differentially abundant family-level taxa based on a beta-binomial model.
+Values on the x-axis are estimates and standard errors of the expected difference in the logit-transformed relative abundance between the two treatments.
+Positive values are enriched in the Positive treatment and negative values are enriched in the Neutral treatment.
+Taxa presented here are significant at FDR < 0.5")`
+
+`r tables("adonis", "PERMANOVA on Aitchison and Jaccard dissimilarities.")`

### 2022-04-07 bash/00_import.sh
cleanup bash directory

@@ -1,9 +1,0 @@
-#!/bin/bash
-qiime tools import \
-  --type 'SampleData[PairedEndSequencesWithQuality]' \
-  --input-path 00_raw_seqs \
-  --input-format CasavaOneEightSingleLanePerSampleDirFmt \
-  --output-path 01_qiime_seqs/demux-paired-end.qza
-qiime demux summarize \
-  --i-data 01_qiime_seqs/demux-paired-end.qza \
-  --o-visualization 01_qiime_seqs/demux-paired-end.qzv

### 2022-04-07 bash/01_dada2.sh

@@ -1,18 +1,0 @@
-#!/bin/bash
-#SBATCH --partition=short       ### queue to submit to
-#SBATCH --job-name=dada2      ### job name
-#SBATCH --output=dada2.out   ### file in which to store job stdout
-#SBATCH --error=dada2.err    ### file in which to store job stderr
-#SBATCH --time=1440                ### wall-clock time limit, in minutes
-#SBATCH --nodes=1
-#SBATCH --ntasks-per-node=1
-#SBATCH --cpus-per-task=28
-#SBATCH --mem=117G
-#SBATCH -A crobe
-#SBATCH --mail-user=andrewmorris@mailbox.org
-#SBATCH --mail-type=ALL
-module load miniconda
-module load R/4.0.2
-
-Rscript R/dada2.R --save
-

### 2022-04-07 bash/02_taxonomy.sh

@@ -1,22 +1,0 @@
-#!/bin/bash
-#SBATCH --partition=short       ### queue to submit to
-#SBATCH --job-name=tax      ### job name
-#SBATCH --output=tax.out   ### file in which to store job stdout
-#SBATCH --error=tax.err    ### file in which to store job stderr
-#SBATCH --time=180                ### wall-clock time limit, in minutes
-#SBATCH --nodes=1
-#SBATCH --ntasks-per-node=1
-#SBATCH --cpus-per-task=4
-#SBATCH --mem=128G
-#SBATCH -A crobe
-#SBATCH --mail-user=amorris28@gmail.com
-#SBATCH --mail-type=ALL
-cd projects/Phylo_Cons_Traits/Gabon
-module load miniconda
-conda activate qiime2-2019.10
-
-qiime feature-classifier classify-sklearn \
-  --i-classifier silva-132-99-515-806-nb-classifier.qza \
-  --i-reads rep-seqs-dn-70.qza \
-  --o-classification taxonomy-dn-70.qza \
-  --p-n-jobs 4

### 2022-04-08 Manuscript/manuscript.Rmd
update manuscript text

@@
 library(captioner)
+library(knitr)
+library(kableExtra)
+library(flextable)
 figures <- captioner()
@@
 
-```{r response, fig.width=4, fig.height=3}
-ratio_log10 <- calc_plot_ratios(fluxes$passage, fluxes$log_ch4)
-
-p <- plot_flux(fluxes, passage, log_ch4, ratio_log10, log10 = TRUE)
-
-ggsave('fig1.jpg', p)
-
-```
-
 We estimated heritability as the regression of mid-offspring on mid-parent (`r figures("herit", display = "cite")`).
@@
 was reversed with a positive heritability for the Positive treatment and a negative heritability for the Neutral treatment.
-
-```{r herit, fig.width=5, fig.height=3}
-
-p <- plot_herit(heritability)
-
-ggsave('fig2.jpg', p)
-
-```
 
@@
 
-```{r beta, results=FALSE, fig.height=2.5}
-
-p <- plot_beta(ps)
-
-ggsave('fig3.jpg', p)
-
-```
-
-```{r adonis}
-
-write.table(fit_adonis, 'table_1.tsv')
-# kable(col.names = c('Term', 'df', 'SS', 'R2', 'F', 'p'), 
-#       digits = c(2,2,2,2,2,3),
-## 	format = "Table",
-## 	booktabs = T,
-# 	caption = "PERMANOVA on Bray-Curtis and Jaccard dissimilarities.",
-#   	label = "beta") %>%
-#   pack_rows("Bray-Curtis", 1, 5) %>%
-#   pack_rows("Jaccard", 6, 10)
-
-```
-
 To investigate which taxa responded to selection on soil CH~4~ oxidation rate, we
@@
 the Neutral treatment (`r figures("da", display = "cite")`).
-
-```{r da, results=FALSE}
-
-p <- dt_plot(dt_data)
-ggsave('fig4.jpg', p)
-
-```
 
@@
 
+```{r response, fig.width=4, fig.height=3}
+ratio_log10 <- calc_plot_ratios(fluxes$passage, fluxes$log_ch4)
+
+p <- plot_flux(fluxes, passage, log_ch4, ratio_log10, log10 = TRUE)
+p
+ggsave('fig1.jpg', p)
+
+```
+
 `r figures("response", "Response to selection on soil CH~4~ oxidation rate fit by ordinary least squares regression.
@@
 The orange line and points are the Positive treatment and the gray line and points are the Neutral treatment.")`
+
+```{r herit, fig.width=5, fig.height=3}
+
+p <- plot_herit(heritability)
+p
+ggsave('fig2.jpg', p)
+
+```
 
@@
 
+```{r beta, results=FALSE, fig.height=2.5}
+
+p <- plot_beta(ps)
+p
+ggsave('fig3.jpg', p)
+
+```
+
 `r figures("beta", "Principal component analysis plot of beta diversity for all jars.
@@
 The ratio of the axes is proportional to the variance explained by each principal component in order to accurately represent the distance between the points.")`
+
+```{r da, results=FALSE}
+
+p <- dt_plot(dt_data)
+p
+ggsave('fig4.jpg', p)
+
+```
 
@@
 `r tables("adonis", "PERMANOVA on Aitchison and Jaccard dissimilarities.")`
+
+```{r adonis}
+
+write.table(fit_adonis, 'table_1.tsv')
+fit_adonis %>%
+  mutate(metric = c(rep("aitchison", 5), rep("jaccard", 5))) %>% 
+  select(metric, term:p.value) %>% 
+	flextable()
+# kable(col.names = c('Term', 'df', 'SS', 'R2', 'F', 'p'), 
+#       digits = c(2,2,2,2,2,3),
+## 	format = "Table",
+## 	booktabs = T,
+# 	caption = "PERMANOVA on Bray-Curtis and Jaccard dissimilarities.",
+#   	label = "beta") %>%
+#   pack_rows("Bray-Curtis", 1, 5) %>%
+#   pack_rows("Jaccard", 6, 10)
+
+```

### 2022-04-08 R/01-process-flux.Rmd
update analysis scripts

@@
 
-```{r plot_fluxes, fig.height=50}
+```{r plot_fluxes, fig.height=20}
 flux_data %>%
  ggplot(aes(days, ppm)) +
- facet_grid(passage + treat ~ rep, scales = "free") +
+ facet_grid(flux_date + treat ~ rep, scales = "free") +
  geom_point() +

### 2022-04-08 R/08-community.Rmd
update analysis scripts

@@
 knitr::opts_knit$set(root.dir=normalizePath('../'))
+
 ```
@@
 source('R/functions.R')
+
 ```

### 2022-04-09 R/08-community.Rmd
Trying to update community analysis to work smoothly

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
+knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = TRUE)
 knitr::opts_knit$set(root.dir=normalizePath('../'))
@@
 library(ggpubr)
-library(SpiecEasi)
+# library(SpiecEasi)
 library(ggthemes)
@@
 
-```
-
-```{r ggplot-settings}
 ```
@@
 
-```{r stacked_bar, eval=FALSE}
-level = "Family"
-ps_phylum <- speedyseq::tax_glom(ps, taxrank=level)
-top10 = names(sort(taxa_sums(ps_phylum), TRUE)[1:10])
-ps_phylum10 <- prune_taxa(top10, ps_phylum)
-ps_phylum10 %>% 
-  ps_filter(passage == 5) %>% 
-  speedyseq::plot_bar(facet_grid = 'treat', fill = level) + 
-  geom_bar(aes_string( color=level, fill=level), stat="identity", position="stack") +
-  geom_bar(aes(x = reorder(rep, rank_ch4)))
-
-```
-
 # Hypothesis Testing
@@
   dist_calc("jaccard", binary = TRUE) %>%
-  ord_calc(method = "NMDS")
+  ord_calc(method = "PCoA")
 
@@
   geom_rangeframe() +
-  coord_cartesian(clip="off") +
+  coord_fixed(3.9/48.3) +
+  # coord_cartesian(clip="off") +
   scale_color_manual(name = "Treatment", 
@@
 
-## ALDex2
-
-```{r}
-library(ALDEx2)
-data(selex)
-#subset only the last 400 features for efficiency
-selex.sub <- selex[1:400,]
-conds <- c(rep("NS", 7), rep("S", 7))
-x <- aldex.clr(selex.sub, conds, mc.samples=16, denom="iqlr", verbose=F)
-x.tt <- aldex.ttest(x, paired.test=FALSE, verbose=FALSE)
-x.effect <- aldex.effect(x, CI=T, verbose=FALSE)
-head(x.effect)
-
-```
-
-```{r}
-ps_fixed_family <- speedyseq::tax_glom(ps_fixed, taxrank="Family")
-
-ps_fam_pas5 <- microViz::ps_filter(ps_fixed_family, passage == 5)
-
-ps_fam_pas5_prev <- microViz::tax_filter(ps_fam_pas5, min_prevalence = 0.10)
-ps_pass_5 <- microViz::ps_filter(ps_fixed, passage == 5)
-otu_table <- as(t(otu_table(ps_fam_pas5_prev)), 'matrix')
-conds <- c(rep("n", 12), rep("p", 12))
-
-# Based on effect confidence interval
-
-x <- aldex.clr(otu_table, conds, mc.samples=128, denom="all", verbose=F)
-x.tt <- aldex.ttest(x, paired.test=FALSE, verbose=FALSE)
-x.effect <- aldex.effect(x, CI=T, verbose=FALSE)
-
-sgn <- sign(x.effect$effect.low) == sign(x.effect$effect.high)
-par(mfrow=c(1,2))
-plot(x.effect$rab.all, x.effect$diff.btw, pch=19, cex=0.3, col="grey", xlab="Abundance", ylab="Difference", main="Bland-Altman")
-points(x.effect$rab.all[abs(x.effect$effect) >=2], x.effect$diff.btw[abs(x.effect$effect) >=2], pch=19, cex=0.5, col="red")
-points(x.effect$rab.all[sgn], x.effect$diff.btw[sgn], cex=0.7, col="blue")
-
-plot(x.effect$diff.win, x.effect$diff.btw, pch=19, cex=0.3, col="grey",xlab="Dispersion", ylab="Difference", main="Effect")
-points(x.effect$diff.win[abs(x.effect$effect) >=2], x.effect$diff.btw[abs(x.effect$effect) >=2], pch=19, cex=0.5, col="red")
-points(x.effect$diff.win[sgn], x.effect$diff.btw[sgn], cex=0.7, col="blue")
-abline(0,2, lty=2, col="grey")
-abline(0,-2, lty=2, col="grey")
-
-merge(as(tax_table(ps_fam_pas5_prev)[rownames(x.effect[sgn,]), ], 'data.frame'), x.effect[sgn,], by = 0)
-```
-
-```{r}
-
-# Based on effect size
-
-x.all <- aldex(otu_table, conds, mc.samples=128, test="t", effect=TRUE,
-     include.sample.summary=FALSE, denom="all", verbose=FALSE)
-effect_gt_2 <- merge(as(tax_table(ps_pass_5)[rownames(x.all[abs(x.all$effect)>2,]), ], 'data.frame'), x.all[abs(x.all$effect)>2,], by = 0)
-
-effect_gt_2 %>% 
-  ggplot(aes(x = effect, y = reorder(Family,effect))) +
-  geom_point()
-
-par(mfrow=c(1,2))
-aldex.plot(x.all, type="MA", test="welch", xlab="Log-ratio abundance",
-    ylab="Difference")
-aldex.plot(x.all, type="MW", test="welch", xlab="Dispersion",
-    ylab="Difference")
-
-par(mfrow=c(1,2))
-plot(x.all$effect, x.all$we.ep, log="y", cex=0.7, col=rgb(0,0,1,0.2),
-  pch=19, xlab="Effect size", ylab="P value", main="Effect size plot")
-points(x.all$effect, x.all$we.eBH, cex=0.7, col=rgb(1,0,0,0.2),
-  pch=19)
-abline(h=0.05, lty=2, col="grey")
-legend(15,1, legend=c("P value", "BH-adjusted"), pch=19, col=c("blue", "red"))
-
-plot(x.all$diff.btw, x.all$we.ep, log="y", cex=0.7, col=rgb(0,0,1,0.2),
-  pch=19, xlab="Difference", ylab="P value", main="Volcano plot")
-points(x.all$diff.btw, x.all$we.eBH, cex=0.7, col=rgb(1,0,0,0.2),
-  pch=19)
-abline(h=0.05, lty=2, col="grey")
-
-# identify which values are significant in both the t-test and glm tests
-found.by.all <- which(x.all$we.eBH < 0.05 & x.all$wi.eBH < 0.05)
-
-# identify which values are significant in fewer than all tests
-found.by.one <- which(x.all$we.eBH < 0.05 | x.all$wi.eBH < 0.05)
-
-# plot the within and between variation of the data
-plot(x.all$diff.win, x.all$diff.btw, pch=19, cex=0.3, col=rgb(0,0,0,0.3),
- xlab="Dispersion", ylab="Difference")
-points(x.all$diff.win[found.by.one], x.all$diff.btw[found.by.one], pch=19,
- cex=0.7, col=rgb(0,0,1,0.5))
-points(x.all$diff.win[found.by.all], x.all$diff.btw[found.by.all], pch=19,
- cex=0.7, col=rgb(1,0,0,1))
-abline(0,1,lty=2)
-abline(0,-1,lty=2)
-
-#View(tax_table(ps_fam_pas5_prev)[rownames(otu_table[found.by.all, ]), ])
-```
-
-```{r, eval=FALSE}
-boxData <-
-  ps_fixed %>% 
-  ps_filter(passage ==5) %>% 
-  tax_agg('Family') %>% 
-  ps_get %>% 
-  ps_melt()
-
-Family_order<-
-  boxData %>% 
-  group_by(Family) %>% 
-  summarize(median = median(Abundance)) %>% 
-  arrange(median)
-boxData$Family <- factor(boxData$Family, levels = Family_order$Family)
-boxData$treat <- factor(boxData$treat, levels = c('p', 'n'))
-```
-
-```{r, eval=FALSE, fig.height=3, fig.width=12}
-
-boxData %>% 
-  ggplot(aes(x = Family, 
-             y = Abundance)) +
-  facet_grid(treat ~ .) +
-  theme_tufte() +
-  geom_rangeframe(color = "black") +
-  # scale_x_reordered() +
-  scale_y_continuous(trans = "pseudo_log") +
-  geom_tufteboxplot() +
-  theme(axis.text.x = element_text(angle = 90))
-  # geom_boxplot() +
-  # scale_color_discrete(guide = 'none') +
-  # coord_flip()
-
-```
-
-```{r, eval=FALSE}
-g5p<- 
-  ps_fixed %>% 
-  ps_filter(treat == "p", passage == 5) %>% 
-  tax_agg('Family') %>% 
-  ps_get() %>% 
-  ps_melt()
-g5n<- 
-  ps_fixed %>% 
-  ps_filter(treat == "n", passage == 5) %>% 
-  tax_agg('Family') %>% 
-  ps_get() %>% 
-  ps_melt()
-
-# Families unique to p in passage 5
-g5p_unique <-
-  g5p %>% 
-  filter(Family %in% unique(g5p$Family)[!(unique(g5p$Family) %in% unique(g5n$Family))]) %>% 
-  mutate(Prevalence = as.numeric(Abundance > 0))
-
-# Abundance
-Family_order<-
-  g5p_unique %>% 
-  group_by(OTU) %>% 
-  summarize(median = median(Abundance), max = max(Abundance)) %>% 
-  arrange(median, max)
-g5p_unique$OTU <- factor(g5p_unique$OTU, levels = Family_order$OTU)
-
-g5p_unique %>% 
-  ggplot(aes(y = Abundance, x = OTU)) +
-  geom_tufteboxplot() +
-  theme_tufte() +
-  geom_rangeframe() +
-  theme(axis.text.x = element_text(angle = 90)) +
-  scale_y_continuous(breaks = c(0, 5, 10, 15, 20, 24))
-
-# Prevalence
-Family_order<-
-  g5p_unique %>% 
-  group_by(OTU) %>% 
-  summarize(sum = sum(Prevalence)) %>% 
-  arrange(sum)
-g5p_unique$OTU <- factor(g5p_unique$OTU, levels = Family_order$OTU)
-
-g5p_unique %>% 
-  group_by(OTU) %>% 
-  summarize(Prevalence = sum(Prevalence)) %>% 
-  ggplot(aes(y = Prevalence, x = OTU)) +
-  geom_point() +
-  theme_tufte() +
-  geom_rangeframe() +
-  theme(axis.text.x = element_text(angle = 90)) +
-  scale_y_continuous(breaks = c(1, 2, 3, 5, 10, 12))
-
-# Families unique to n in passage 5
-g5n_unique <-
-  g5n %>% 
-  filter(Family %in% unique(g5n$Family)[!(unique(g5n$Family) %in% unique(g5p$Family))]) %>% 
-  mutate(Prevalence = as.numeric(Abundance > 0))
-
-# Abundance
-Family_order<-
-  g5n_unique %>% 
-  group_by(OTU) %>% 
-  summarize(median = median(Abundance), max = max(Abundance)) %>% 
-  arrange(median, max)
-g5n_unique$OTU <- factor(g5n_unique$OTU, levels = Family_order$OTU)
-
-g5n_unique %>% 
-  ggplot(aes(y = Abundance, x = OTU)) +
-  geom_tufteboxplot() +
-  theme_tufte() +
-  geom_rangeframe() +
-  theme(axis.text.x = element_text(angle = 90)) +
-  scale_y_continuous(breaks = c(0, 10, 20, 30, 37))
-
-# Prevalence
-Family_order<-
-  g5n_unique %>% 
-  group_by(OTU) %>% 
-  summarize(sum = sum(Prevalence)) %>% 
-  arrange(sum)
-g5n_unique$OTU <- factor(g5n_unique$OTU, levels = Family_order$OTU)
-
-g5n_unique %>% 
-  group_by(OTU) %>% 
-  summarize(Prevalence = sum(Prevalence)) %>% 
-  ggplot(aes(y = Prevalence, x = OTU)) +
-  geom_point() +
-  theme_tufte() +
-  geom_rangeframe() +
-  theme(axis.text.x = element_text(angle = 90)) +
-  scale_y_continuous(breaks = c(1, 2, 6))
-```
-
 First, I looked at which taxa are unique to treatments N and P. Several families
@@
 
-```{r corncob_treatment, fig.hight=40}
+```{r corncob_treatment, results=FALSE, fig.hight=40}
 ps_fixed_family <- speedyseq::tax_glom(ps_fixed, taxrank="Family")
@@
 
+```{r stahp}
+knitr::opts_chunk$set(eval = FALSE)
+
+```
+
+```{r stacked_bar, eval=FALSE}
+level = "Family"
+ps_phylum <- speedyseq::tax_glom(ps, taxrank=level)
+top10 = names(sort(taxa_sums(ps_phylum), TRUE)[1:10])
+ps_phylum10 <- prune_taxa(top10, ps_phylum)
+ps_phylum10 %>% 
+  ps_filter(passage == 5) %>% 
+  speedyseq::plot_bar(facet_grid = 'treat', fill = level) + 
+  geom_bar(aes_string( color=level, fill=level), stat="identity", position="stack") +
+  geom_bar(aes(x = reorder(rep, rank_ch4)))
+
+```
+
+## ALDex2
+
+```{r, eval = FALSE}
+library(ALDEx2)
+data(selex)
+#subset only the last 400 features for efficiency
+selex.sub <- selex[1:400,]
+conds <- c(rep("NS", 7), rep("S", 7))
+x <- aldex.clr(selex.sub, conds, mc.samples=16, denom="iqlr", verbose=F)
+x.tt <- aldex.ttest(x, paired.test=FALSE, verbose=FALSE)
+x.effect <- aldex.effect(x, CI=T, verbose=FALSE)
+head(x.effect)
+
+```
+
+```{r, eval=FALSE}
+ps_fixed_family <- speedyseq::tax_glom(ps_fixed, taxrank="Family")
+
+ps_fam_pas5 <- microViz::ps_filter(ps_fixed_family, passage == 5)
+
+ps_fam_pas5_prev <- microViz::tax_filter(ps_fam_pas5, min_prevalence = 0.10)
+ps_pass_5 <- microViz::ps_filter(ps_fixed, passage == 5)
+otu_table <- as(t(otu_table(ps_fam_pas5_prev)), 'matrix')
+conds <- c(rep("n", 12), rep("p", 12))
+
+# Based on effect confidence interval
+
+x <- aldex.clr(otu_table, conds, mc.samples=128, denom="all", verbose=F)
+x.tt <- aldex.ttest(x, paired.test=FALSE, verbose=FALSE)
+x.effect <- aldex.effect(x, CI=T, verbose=FALSE)
+
+sgn <- sign(x.effect$effect.low) == sign(x.effect$effect.high)
+par(mfrow=c(1,2))
+plot(x.effect$rab.all, x.effect$diff.btw, pch=19, cex=0.3, col="grey", xlab="Abundance", ylab="Difference", main="Bland-Altman")
+points(x.effect$rab.all[abs(x.effect$effect) >=2], x.effect$diff.btw[abs(x.effect$effect) >=2], pch=19, cex=0.5, col="red")
+points(x.effect$rab.all[sgn], x.effect$diff.btw[sgn], cex=0.7, col="blue")
+
+plot(x.effect$diff.win, x.effect$diff.btw, pch=19, cex=0.3, col="grey",xlab="Dispersion", ylab="Difference", main="Effect")
+points(x.effect$diff.win[abs(x.effect$effect) >=2], x.effect$diff.btw[abs(x.effect$effect) >=2], pch=19, cex=0.5, col="red")
+points(x.effect$diff.win[sgn], x.effect$diff.btw[sgn], cex=0.7, col="blue")
+abline(0,2, lty=2, col="grey")
+abline(0,-2, lty=2, col="grey")
+
+merge(as(tax_table(ps_fam_pas5_prev)[rownames(x.effect[sgn,]), ], 'data.frame'), x.effect[sgn,], by = 0)
+```
+
+```{r, eval=FALSE}
+
+# Based on effect size
+
+x.all <- aldex(otu_table, conds, mc.samples=128, test="t", effect=TRUE,
+     include.sample.summary=FALSE, denom="all", verbose=FALSE)
+effect_gt_2 <- merge(as(tax_table(ps_pass_5)[rownames(x.all[abs(x.all$effect)>2,]), ], 'data.frame'), x.all[abs(x.all$effect)>2,], by = 0)
+
+effect_gt_2 %>% 
+  ggplot(aes(x = effect, y = reorder(Family,effect))) +
+  geom_point()
+
+par(mfrow=c(1,2))
+aldex.plot(x.all, type="MA", test="welch", xlab="Log-ratio abundance",
+    ylab="Difference")
+aldex.plot(x.all, type="MW", test="welch", xlab="Dispersion",
+    ylab="Difference")
+
+par(mfrow=c(1,2))
+plot(x.all$effect, x.all$we.ep, log="y", cex=0.7, col=rgb(0,0,1,0.2),
+  pch=19, xlab="Effect size", ylab="P value", main="Effect size plot")
+points(x.all$effect, x.all$we.eBH, cex=0.7, col=rgb(1,0,0,0.2),
+  pch=19)
+abline(h=0.05, lty=2, col="grey")
+legend(15,1, legend=c("P value", "BH-adjusted"), pch=19, col=c("blue", "red"))
+
+plot(x.all$diff.btw, x.all$we.ep, log="y", cex=0.7, col=rgb(0,0,1,0.2),
+  pch=19, xlab="Difference", ylab="P value", main="Volcano plot")
+points(x.all$diff.btw, x.all$we.eBH, cex=0.7, col=rgb(1,0,0,0.2),
+  pch=19)
+abline(h=0.05, lty=2, col="grey")
+
+# identify which values are significant in both the t-test and glm tests
+found.by.all <- which(x.all$we.eBH < 0.05 & x.all$wi.eBH < 0.05)
+
+# identify which values are significant in fewer than all tests
+found.by.one <- which(x.all$we.eBH < 0.05 | x.all$wi.eBH < 0.05)
+
+# plot the within and between variation of the data
+plot(x.all$diff.win, x.all$diff.btw, pch=19, cex=0.3, col=rgb(0,0,0,0.3),
+ xlab="Dispersion", ylab="Difference")
+points(x.all$diff.win[found.by.one], x.all$diff.btw[found.by.one], pch=19,
+ cex=0.7, col=rgb(0,0,1,0.5))
+points(x.all$diff.win[found.by.all], x.all$diff.btw[found.by.all], pch=19,
+ cex=0.7, col=rgb(1,0,0,1))
+abline(0,1,lty=2)
+abline(0,-1,lty=2)
+
+#View(tax_table(ps_fam_pas5_prev)[rownames(otu_table[found.by.all, ]), ])
+```
+
+```{r, eval=FALSE}
+boxData <-
+  ps_fixed %>% 
+  ps_filter(passage ==5) %>% 
+  tax_agg('Family') %>% 
+  ps_get %>% 
+  ps_melt()
+
+Family_order<-
+  boxData %>% 
+  group_by(Family) %>% 
+  summarize(median = median(Abundance)) %>% 
+  arrange(median)
+boxData$Family <- factor(boxData$Family, levels = Family_order$Family)
+boxData$treat <- factor(boxData$treat, levels = c('p', 'n'))
+```
+
+```{r, eval=FALSE, fig.height=3, fig.width=12}
+
+boxData %>% 
+  ggplot(aes(x = Family, 
+             y = Abundance)) +
+  facet_grid(treat ~ .) +
+  theme_tufte() +
+  geom_rangeframe(color = "black") +
+  # scale_x_reordered() +
+  scale_y_continuous(trans = "pseudo_log") +
+  geom_tufteboxplot() +
+  theme(axis.text.x = element_text(angle = 90))
+  # geom_boxplot() +
+  # scale_color_discrete(guide = 'none') +
+  # coord_flip()
+
+```
+
+```{r, eval=FALSE}
+g5p<- 
+  ps_fixed %>% 
+  ps_filter(treat == "p", passage == 5) %>% 
+  tax_agg('Family') %>% 
+  ps_get() %>% 
+  ps_melt()
+g5n<- 
+  ps_fixed %>% 
+  ps_filter(treat == "n", passage == 5) %>% 
+  tax_agg('Family') %>% 
+  ps_get() %>% 
+  ps_melt()
+
+# Families unique to p in passage 5
+g5p_unique <-
+  g5p %>% 
+  filter(Family %in% unique(g5p$Family)[!(unique(g5p$Family) %in% unique(g5n$Family))]) %>% 
+  mutate(Prevalence = as.numeric(Abundance > 0))
+
+# Abundance
+Family_order<-
+  g5p_unique %>% 
+  group_by(OTU) %>% 
+  summarize(median = median(Abundance), max = max(Abundance)) %>% 
+  arrange(median, max)
+g5p_unique$OTU <- factor(g5p_unique$OTU, levels = Family_order$OTU)
+
+g5p_unique %>% 
+  ggplot(aes(y = Abundance, x = OTU)) +
+  geom_tufteboxplot() +
+  theme_tufte() +
+  geom_rangeframe() +
+  theme(axis.text.x = element_text(angle = 90)) +
+  scale_y_continuous(breaks = c(0, 5, 10, 15, 20, 24))
+
+# Prevalence
+Family_order<-
+  g5p_unique %>% 
+  group_by(OTU) %>% 
+  summarize(sum = sum(Prevalence)) %>% 
+  arrange(sum)
+g5p_unique$OTU <- factor(g5p_unique$OTU, levels = Family_order$OTU)
+
+g5p_unique %>% 
+  group_by(OTU) %>% 
+  summarize(Prevalence = sum(Prevalence)) %>% 
+  ggplot(aes(y = Prevalence, x = OTU)) +
+  geom_point() +
+  theme_tufte() +
+  geom_rangeframe() +
+  theme(axis.text.x = element_text(angle = 90)) +
+  scale_y_continuous(breaks = c(1, 2, 3, 5, 10, 12))
+
+# Families unique to n in passage 5
+g5n_unique <-
+  g5n %>% 
+  filter(Family %in% unique(g5n$Family)[!(unique(g5n$Family) %in% unique(g5p$Family))]) %>% 
+  mutate(Prevalence = as.numeric(Abundance > 0))
+
+# Abundance
+Family_order<-
+  g5n_unique %>% 
+  group_by(OTU) %>% 
+  summarize(median = median(Abundance), max = max(Abundance)) %>% 
+  arrange(median, max)
+g5n_unique$OTU <- factor(g5n_unique$OTU, levels = Family_order$OTU)
+
+g5n_unique %>% 
+  ggplot(aes(y = Abundance, x = OTU)) +
+  geom_tufteboxplot() +
+  theme_tufte() +
+  geom_rangeframe() +
+  theme(axis.text.x = element_text(angle = 90)) +
+  scale_y_continuous(breaks = c(0, 10, 20, 30, 37))
+
+# Prevalence
+Family_order<-
+  g5n_unique %>% 
+  group_by(OTU) %>% 
+  summarize(sum = sum(Prevalence)) %>% 
+  arrange(sum)
+g5n_unique$OTU <- factor(g5n_unique$OTU, levels = Family_order$OTU)
+
+g5n_unique %>% 
+  group_by(OTU) %>% 
+  summarize(Prevalence = sum(Prevalence)) %>% 
+  ggplot(aes(y = Prevalence, x = OTU)) +
+  geom_point() +
+  theme_tufte() +
+  geom_rangeframe() +
+  theme(axis.text.x = element_text(angle = 90)) +
+  scale_y_continuous(breaks = c(1, 2, 6))
+```
+
 ## Faprotax on DA taxa
@@
 
-```{r}
+```{r, eval=FALSE}
 biom_file <- "Output/func_table_from_biom_json.biom"

### 2022-04-18 R/08-community.Rmd
update Readme; prevent saving of phyloseq RData file

@@
 
-save(ps, file = 'ps.RData')
+#save(ps, file = 'ps.RData')
 # Differential abundance test

### 2022-04-26 Manuscript/manuscript.Rmd
Cleanup directory for publication

@@
 
-load(file = '../Output/response_model.Rdata')
-load(file = '../Output/herit_model.Rdata')
+# CH4 models
+load(file = '../Output/response_model.RData')
+load(file = '../Output/herit_model.RData')
 
-load(file = '../Output/richness_model_passage.Rdata')
-load(file = '../Output/richness_model_treat5.Rdata')
-load(file = '../Output/richness_model_ch4.Rdata')
+# 16S models
+load(file = '../Output/richness_models.RData')
 load(file = '../Output/phyloseq.RData')
-load(file = '../Output/beta_model.Rdata')
-load(file = '../Output/da_corncob.Rdata')
-load(file = '../Output/dt_data.Rdata')
+load(file = '../Output/beta_model.RData')
+load(file = '../Output/da_corncob.RData')
 

### 2022-04-26 R/01-process-flux.Rmd
Cleanup directory for publication

@@
 
-This script processes the standard curve data from the GC on each sample date.
-It then converts area to methane concentration using the standard curves for every sample.
-Finally, it computes fluxes (change in concentration over time) as a first-order exponential decay function.
+This script processes the standard curve data from the gas chromatograph on
+each sample date. It then converts area-under-the-curve to methane
+concentration using the standard curves for every sample. Finally, it
+computes fluxes (change in concentration over time) as a first-order
+exponential decay function.
 
@@
 
+where `flux_date` is the date flux was measured, `treat` is the selection
+treatment, either `p` Positive or `n` Neutral, and `n` is the number of jars
+for that treatment and sample date. Dates with fewer than 12 samples were due
+to broken jars.
+
 ## Standard Curves
@@
 This step first imports the values for the standard curves then calculates the slope of Methane on Area for each sample date.
+On one flux date, a new standard curve was established on the second day of measurement due to drift in the check standards so t1 and t2 had a different standard curve than t3-t5.
+`flux_date` indicates the main sample data while `sc_date` indicates which standard curve was used on that day.
 
@@
 
+sc_ch4 %>% 
+	kable()
+
 # These are the final slopes for each Methane~Area standard curve 
-# model for each date
+# model for each date and time point
 sc_ch4 <- 
@@
   select(flux_date:sc_date, t, r.squared:slope)
+
 ```
@@
 
-This step turns the time point data to time intervals in units of days.
+Remove Passage 6 as explained earlier.
 
@@
 
+Converts raw data and time time interval in units `days`.
+The first time point, t0 is when the jars were capped and is ignored because t1 is the first sample injected into the gas chromatograph and so is the first time point with concentrations.
+Decay constants are calculated between t1 and t5.
+
 ```{r compute_time_intervals}
+# Standardize time format
+time_data[, c('t0', 't1', 't2', 't3', 't4', 't5')] <- 
+  lapply(time_data[, c('t0', 't1', 't2', 't3', 't4', 't5')], ymd_hms)
 
-time_data[, c('t0', 't1', 't2', 't3', 't4', 't5')] <- 
-  lapply(time_data[, c('t0', 't1', 't2', 't3', 't4', 't5')],
-         ymd_hms)
-
+# Convert time to 'days'
 as.days <- function(start, end) {as.numeric(interval(start, end))/60/60/24}
@@
          function(x) as.days(time_data$t1, x))
+
 time_data <- 
@@
 
-This step takes the raw gas chromatograph area data and calculates methane concentration for each time point in units of ppm.
+This step takes the raw gas chromatograph area data and calculates methane concentration for each time point in units of ppm using the standard curves calculated above.
 
@@
 
+calc_frac_remain <- function(x) {
+  if (is.na(x[3])) {
+    x[2]/x[1]
+  }
+  else {
+    x[3]/x[1]
+  }
+}
+
 conc_data %>% 
-  select(flux_date, jar, t1, t4) %>% 
-  mutate(frac_cons = t4/t1) %>% 
+  select(flux_date, jar, t1, t4, t5) %>% 
+  mutate(frac_remain = apply(conc_data[,c('t1', 't4', 't5')], 1, calc_frac_remain)) %>% 
   mutate(treat = substring(jar,1,1)) %>% 
-  ggplot(aes(x = flux_date, y = frac_cons, color = treat)) +
+  ggplot(aes(x = flux_date, y = frac_remain, color = treat)) +
   geom_jitter(width = 0.2) +
   geom_hline(yintercept = 0.90) +
-  geom_hline(yintercept = 0.05)
+  geom_hline(yintercept = 0.05) +
+  labs(x = "Flux Date", y = "Fraction of Methane Remaining")
 
@@
   gather(t, ppm, t1:t5) %>% 
-  left_join(time_data, by = c('flux_date', 'jar', 't'))
-
-flux_data <- 
-  flux_data %>% 
+  left_join(time_data, by = c('flux_date', 'jar', 't')) %>% 
   mutate(flux_date = factor(ymd(flux_date)), 
@@
 
-# Identify the selected jars in the fluxes dataframe
+# Identify the selected jars in the fluxes data frame
 selected$selected <- TRUE

### 2022-04-26 R/03-response.Rmd
Cleanup directory for publication

@@
 ```{r save_model}
-save(response_model, compare_response_mods, per_slope, file = '../Output/response_model.Rdata')
+save(response_model, compare_response_mods, per_slope, file = '../Output/response_model.RData')
 ```
@@
 5]], 2)`).
-
-## Previous Method
-
-Previously, I was using the `metafor` package to test the slope of the difference between the positive and neutral line.
-For this model, I calculate the difference between the Positive and Neutral treatment and fit that difference as the response.
-I also calculate the standard error of the difference and included that variance in the model.
-I then fit the model `difference ~ passage` to test the slope of the difference over time.
-
-```{r eval=FALSE}
-
-# Standard error of the difference
-sed <- function(p, n, na.rm = FALSE) {
-  sqrt(se(p, na.rm = na.rm) ^ 2 + se(n, na.rm = na.rm)^ 2)
-}
-# Calculate deviance as p - n
-deviance <-
-  fluxes %>% 
-  select(treat, passage, estimate) %>% 
-  group_by(passage, treat) %>% 
-  summarize(mean = mean(estimate), .groups = "drop") %>% 
-  pivot_wider(names_from = treat, values_from = mean) %>% 
-  mutate(deviance = p - n)
-
-# Calculate variance of deviance as SED
-deviance$se <-
-  fluxes %>% 
-  select(treat, passage, estimate) %>% 
-  mutate(x = 1:nrow(fluxes)) %>% 
-  pivot_wider(names_from = treat, values_from = estimate) %>% 
-  group_by(passage) %>% 
-  summarize(sed = sed(p, n, na.rm = TRUE), .groups = "drop") %>%
-  pull(sed)
-
-# Fit deviance model
-fit <- rma(deviance ~ passage, se^2, method="FE", data = deviance)
-
-plot(fit)
-
-# Plot deviance
-ggplot(deviance, aes(x = passage, y = deviance, ymin = deviance - se, ymax = deviance + se)) + 
-  geom_pointrange(color = 'darkorange2') + 
-  geom_abline(intercept = fit$beta[1], slope = fit$beta[2])
-
-# Take deviance model parameters and export into a data.frame
-rma_output <- data.frame(
-Estimate = fit$beta,
-se = fit$se,
-z.value = fit$zval,
-p.value = fit$pval,
-upper.ci = fit$ci.lb,
-lower.ci = fit$ci.ub
-)
-
-# Fit deviance model without passage 2
-fit_no2 <- 
-filter(deviance, passage != 2) %>% 
-rma(deviance ~ passage, se^2, method="FE", data = .)
-
-# Take deviance model withou passage 2 parameters and export into a data.frame
-
-rma_output_no2 <- data.frame(
-Estimate = fit_no2$beta,
-se = fit_no2$se,
-z.value = fit_no2$zval,
-p.value = fit_no2$pval,
-upper.ci = fit_no2$ci.lb,
-lower.ci = fit_no2$ci.ub
-)
-
-# R (response to selection) for the model w/o sample 2
-R_no2 <- rma_output_no2[[2, 1]]
-R <- rma_output[[2, 1]]
-
-# Plot the no2 model
-filter(deviance, passage != 2) %>% 
-ggplot(., aes(x = passage, y = deviance, ymin = deviance - se, ymax = deviance + se)) + 
-  geom_pointrange(color = 'darkorange2') + 
-  geom_abline(intercept = fit_no2$beta[1], slope = fit_no2$beta[2])
-
-#write_tsv(deviance, '../Output/deviance.tsv')
-#saveRDS(rma_output, '../Output/response.rds')
-#saveRDS(fit, '../Output/dev_fit.rds')
-
-# Calculate deviance as percentage of p - n
-deviance <-
-  fluxes %>% 
-  select(treat, passage, estimate) %>% 
-  group_by(passage, treat) %>% 
-  summarize(mean = mean(estimate), .groups = "drop") %>% 
-  pivot_wider(names_from = treat, values_from = mean) %>% 
-  mutate(deviance = (p - n) / n)
-
-# Calculate variance of deviance as SED
-deviance$se <-
-  fluxes %>% 
-  select(treat, passage, estimate) %>% 
-  mutate(x = 1:nrow(fluxes)) %>% 
-  pivot_wider(names_from = treat, values_from = estimate) %>% 
-  group_by(passage) %>% 
-  summarize(sed = sed(p, n, na.rm = TRUE), .groups = "drop") %>%
-  pull(sed)
-
-# Fit deviance model
-fit <- rma(deviance ~ passage, se^2, method="FE", data = deviance)
-
-plot(fit)
-
-# Plot deviance
-ggplot(deviance, aes(x = passage, y = deviance, ymin = deviance - se, ymax = deviance + se)) + 
-  geom_pointrange(color = 'darkorange2') + 
-  geom_abline(intercept = fit$beta[1], slope = fit$beta[2])
-
-# Take deviance model parameters and export into a data.frame
-rma_output <- data.frame(
-Estimate = fit$beta,
-se = fit$se,
-z.value = fit$zval,
-p.value = fit$pval,
-upper.ci = fit$ci.lb,
-lower.ci = fit$ci.ub
-)
-
-# Fit separte slopes for Positive and Neutral with robust M-estimator regression
-
-library(MASS)
-anova(
-fit <- rlm(estimate ~ passage * treat, data = fluxes),
-fit <- rlm(estimate ~ passage + treat, data = fluxes))
-
-summary(fit)
-anova(fit)
-new_data <- cbind(fluxes, predict(fit, interval = 'confidence'))
-colnames(new_data) 
-ggplot(new_data, aes(x = passage, y = estimate, color = treat)) +
-  theme_bw() +
-  geom_jitter() +
-  geom_line(aes(passage, fit)) +
-  geom_ribbon(aes(ymin=lwr,ymax=upr), alpha=0.3)
-
-ggplot(new_data, aes(x = passage, y = estimate)) +
-  theme_bw() +
-  geom_jitter(aes(color = treat)) +
-  geom_line(aes(passage, fit, color = treat)) +
-  geom_ribbon(aes(ymin=lwr,ymax=upr, group = treat), alpha=0.3) +
-  ylim(-0.01, 0.08)
-ggplot(fluxes, aes(x = passage, y = estimate, color = treat)) + 
-  geom_point() + 
-  stat_smooth(method='lm')
-#ggsave('raw_fluxes.pdf')
-```

### 2022-04-26 R/04-heritability.Rmd
Cleanup directory for publication

@@
 This script imports the heritability data. This dataset includes the mean flux
-withn treatmentxgeneration for 1) all parents, 2) selected parents, and 3)
-offspring. Then, it plots offspring on the selection parents with the two
+within treatment by generation for 1) all parents, 2) selected parents, and 3)
+offspring. Then, it plots offspring on the selected parents with the two
 treatments combined and then separately for the two treatments.
@@
 
-The negative heritability of methane oxidation rate in the Neutral selection treatment may result from selection on alternative characters in those populations.
-The strong, positive heritability in the Positive treatment ...
-.
-It seems strange that the
-
 ```{r plot_herit}
@@
 ```{r save_heritability}
-save(herit_model, file = '../Output/herit_model.Rdata')
+save(herit_model, file = '../Output/herit_model.RData')
 ```
-
-######################################################################
-# End of the way I should be doing it
-################################################################
-
-The Old way:
-
-It then calculates S, the selection differential, as the mean difference between the selected parents and all parents.
-Because passage 2 was accidentaly selected for the lowest methane oxidation rate, I take the absolute value of this difference, i.e.
-|S|.
-Next, it imports the deviance data and calculates R, the respones to selection, as the slope of mean difference in flux (positive - neutral) over time.
-
-An alternative approach would be to calculate R directly instead of approximating it using the slope as in the equation offspring - parent.
-In this case, I would not be using the deviance, but instead the raw fluxes.
-I can do that with the heritability dataset.
-
-# Function for the Breeder's equation
-
-```{r eval=FALSE, include=FALSE}
-breeders <- function(R, S) {
-  R / S
-}
-
-#
-
-selected <- 
-  fluxes[paste0(fluxes$passage, fluxes$jar) %in% paste0(selected_jars$passage, selected_jars$jar), ] %>%
-  select(treat, log_ch4, passage) %>% 
-  group_by(treat, passage) %>% 
-  summarize(selected = mean(log_ch4), .groups = 'drop')
-
-parents <- 
-  fluxes %>%
-  select(treat, log_ch4, passage) %>% 
-  group_by(treat, passage) %>% 
-  summarize(parental = mean(log_ch4), .groups = 'drop')
-
-offspring <- 
-  fluxes %>% 
-  select(passage, treat, log_ch4) %>% 
-  filter(passage != 1) %>% 
-  mutate(passage = passage - 1) %>% 
-  group_by(treat, passage) %>% 
-  summarize(offspring = mean(log_ch4), .groups = 'drop')
-
-heritability <- 
-  offspring %>% 
-  left_join(selected, by = c('passage', 'treat')) %>% 
-  left_join(parents, by = c('passage', 'treat'))
-
-# Matt's approach: h2 for each generation
-h2_per_gen  <- 
-heritability %>% 
-  mutate(R = offspring - parental) %>% 
-  mutate(S = selected - parental) %>% 
-  mutate(h2 = breeders(R, S))
-
-write_tsv(h2_per_gen, 'Output/h2_per_gen.tsv')
-
-ggplot(h2_per_gen, aes(x = passage, color = treat, y = h2)) +
-  geom_point()
-
-h2_per_gen %>% 
-  select(treat, passage, h2)
-
-h2_per_gen %>% 
-  select(treat, passage, h2) %>% 
-  group_by(treat) %>% 
-  summarize(mean_h2 = mean(h2), sd_h2 = sd(h2), .groups = 'drop')
-
-# Calculate R using regression of deviance over time.
-# Calculate S using math.
-
-dev_fit <- readRDS('../Output/dev_fit.rds')
-R <- dev_fit$beta[2]
-
-S <-
-  heritability %>%
-  filter(treat == "p") %>%
-  mutate(S = abs(selected - parental)) %>%
-  summarize(S = mean(S))  %>% 
-  pull()
-
-h2_reg_dev <- breeders(R, S)
-
-lm_fit <- 
-  fluxes %>% 
-  filter(treat == 'p') %>% 
-  select(treat, passage, ch4) %>% 
-  lm(ch4 ~ passage, data = .)
-
-R_not_dev <- pull(tidy(lm_fit)[2, 2])
-h2_reg_abs <- breeders(R_not_dev, S)
-
-# Calculate R as regression of offspring on parents
-
-ggplot(heritability, aes(x = parental, y = offspring)) +
-  geom_point() + 
-  stat_smooth(method = 'lm', se = FALSE)
-summary(lm(offspring ~ parental, data = heritability, subset = treat == 'p'))
-# Calculate h2 (heritability) using math, not regression
-# For each generation and then calculate the mean/se
-h2_calc_method <-
-  heritability %>%
-  filter(treat == "p") %>%
-  mutate(S = abs(selected - parental)) %>%
-  mutate(R = offspring - parental) %>%
-  mutate(h2 = breeders(R, S)) %>%
-  summarize(S = mean(S),
-            R = mean(R),
-            se = se(h2),
-            h2 = mean(h2))
-
-h2_calc_method_no_2 <-
-  heritability %>%
-  filter(treat == "p") %>%  
-  filter(passage != 2) %>%
-  mutate(S = abs(selected - parental)) %>%
-  mutate(R = offspring - parental) %>%
-  mutate(h2 = breeders(R, S)) %>%
-  summarize(S = mean(S),
-            R = mean(R),
-            se = se(h2),
-            h2 = mean(h2))
-
-h2 <- 
-  h2_calc_method %>% 
-  mutate(method = "calc p")
-
-h2_no2 <- 
-  h2_calc_method_no_2 %>% 
-  mutate(method = "calc p no 2")
-
-h2 <- rbind(h2, h2_no2)
-
-# Cumulative response to selection
-
-parent <- heritability[[1, 3]]
-offspr <- heritability[[8, 4]]
-select <- heritability[[8, 5]]
-R = offspr - parent
-S = select - parent
-h2 = breeders(R, S)
-
-heritability[,'cum_res'] <- NA
-heritability$cum_sel <- NA
-herit_p <- filter(heritability, treat == 'p')
-for (i in seq(nrow(herit_p))) {
-  off <- sum(herit_p$offspring[1:i])
-  par <- sum(herit_p$parental[1:i])
-  sel <- sum(herit_p$selected[1:i])
-  herit_p$cum_res[i] <- off - par
-  herit_p$cum_sel[i] <- sel - par
-}
-herit_p$cum_h2 <- herit_p$cum_res / herit_p$cum_sel
-
-ggplot(herit_p, aes(x = cum_sel, y = cum_res)) + 
-  theme_classic() +
-  labs(x = 'Cumulative selection differential',
-       y = 'Cumulative response to selection') +
-  geom_point() + 
-  stat_smooth(method = 'lm', se = FALSE)
-summary(lm(cum_res ~ cum_sel, data = herit_p))
-
-# Regression of offspring on mid-parent
-
-selected_jars <- read_csv('../Data/selected.csv')
-fluxes <- read_tsv('../Output/fluxes.tsv')
-
-# Pull out the selected parental jars
-selected <- 
-  fluxes[paste0(fluxes$passage, fluxes$jar) %in% paste0(selected_jars$passage, selected_jars$jar), ] %>%
-  select(treat, ch4, passage) %>% 
-  group_by(treat, passage) %>% 
-  summarize(parental = mean(ch4), .groups = 'drop')
-
-offspring <- 
-  fluxes %>% 
-  select(passage, treat, ch4) %>% 
-  filter(passage != 1) %>% 
-  mutate(passage = passage - 1) %>% 
-  rename(offspring = ch4)
-
-heritability <- 
-  offspring %>% 
-  left_join(selected, by = c('passage', 'treat'))
-
-heritability
-
-ggplot(heritability, aes(parental, offspring)) +
-  theme_classic() +
-  geom_point() + 
-  stat_smooth(method = 'lm', se = FALSE)
-
-summary(lm(offspring ~ parental, data = heritability))
-
-ggplot(heritability, aes(parental, offspring, color = treat)) +
-  theme_classic() +
-  geom_point() + 
-  stat_smooth(method = 'lm', se = FALSE)
-summary(lm(offspring ~ parental * treat, data = heritability))
-```

### 2022-04-26 R/07-generate-physeq.Rmd
Cleanup directory for publication

--- 2022-04-07 R/07-generate_physeq.Rmd
@@
 ```{r import_data}
-load(file = "Output/dada2_output.RData")
+load(file = "HPC_Output/dada2_output.RData")
 
@@
 
-# Create final phyloseq
-
 ```{r create_physeq}
@@
 
-# 
-# 
-# seqtab.nochim.filtered <-
-#   seqtab.nochim[rownames(seqtab.nochim) %in% rownames(fluxes_filtered),]
-# 
-# ps <- phyloseq(otu_table(seqtab.nochim.filtered, taxa_are_rows=FALSE), 
-#                                      sample_data(fluxes_filtered),
-# 
-#                tax_table(taxa))
-# 
-# 
-# 
-# dna <- Biostrings::DNAStringSet(taxa_names(ps))
-# names(dna) <- taxa_names(ps)
-# ps <- merge_phyloseq(ps, dna)
-# taxa_names(ps) <- paste0("ASV", seq(ntaxa(ps)))
-
 ```

### 2022-04-26 R/05-create-barcode-key.R

@@ -1,13 +1,14 @@
-# Script to create sample ID/barcode sequence pairs to submit to GC3F
+# Script to create sample ID/barcode sequence pairs to submit to the GC3F
+# genomics core facility at the University of Oregon
 
 # Takes one .tsv with Sample IDs and primer plate well positions for their
 # respective primers (in the form A1, A2, ... A12)
 # Takes a second table with barcode sequences and well row or column id for that
 # primer (e.g., TATGGCAC  A, ATAACGCC B, ..., CTTCACTG	12)
 
-# Outputs a "master" barcode table with the gc3f ids, my sample ids,
+# Outputs a "master" barcode table with the GC3F ids, my sample ids,
 # and all of the barcodes and well positions
-# Also outputs a table for gC3f with just the gc3f ids and the barcode
+# Also outputs a table for GC3F with just the GC3F ids and the barcode
 # sequences with the appropriate column headers
 
 library(tidyverse)
@@ -28,4 +29,4 @@ read_tsv('../Output/barcode_master.tsv') %>%
          `Index 1 (i7) sequence` = r_barcode, 
          `Index 2 (i5) sequence` = f_barcode) %>% 
   write_tsv('../Output/barcode_gc3f.tsv')
-  
+

### 2022-04-26 R/06-dada2.R

@@ -3,15 +3,15 @@ library(tidyverse)
 
 multithread = 28
 
-path <- "00_raw_seqs" # CHANGE ME to the directory containing the fastq files after unzipping.
+path <- "raw_seqs" # CHANGE ME to the directory containing the fastq files after unzipping.
 #list.files(path)
 
 # Forward and reverse fastq filenames have format: SAMPLENAME_R1_001.fastq and SAMPLENAME_R2_001.fastq
 fnFs <- sort(list.files(path, pattern="_R1_001.fastq.gz", full.names = TRUE))
 fnRs <- sort(list.files(path, pattern="_R2_001.fastq.gz", full.names = TRUE))
 # Extract sample names, assuming filenames have format: SAMPLENAME_XXX.fastq
-fnFs <- fnFs[grepl('am', fnFs)]
-fnRs <- fnRs[grepl('am', fnRs)]
+# fnFs <- fnFs[grepl('am', fnFs)]
+# fnRs <- fnRs[grepl('am', fnRs)]
 
 sample.names <- sapply(strsplit(basename(fnFs), "_"), `[`, 1)
 
@@ -25,8 +25,8 @@ sample.names <- sapply(strsplit(basename(fnFs), "_"), `[`, 1)
 #plotQualityProfile(fnRs[1:2])
 
 # Place filtered files in filtered/ subdirectory
-filtFs <- file.path("01_filtered_seqs", paste0(sample.names, "_F_filt.fastq.gz"))
-filtRs <- file.path("01_filtered_seqs", paste0(sample.names, "_R_filt.fastq.gz"))
+filtFs <- file.path("filtered_seqs", paste0(sample.names, "_F_filt.fastq.gz"))
+filtRs <- file.path("filtered_seqs", paste0(sample.names, "_R_filt.fastq.gz"))
 names(filtFs) <- sample.names
 names(filtRs) <- sample.names
 
@@ -118,4 +118,5 @@ taxa.print <- taxa # Removing sequence rownames for display only
 rownames(taxa.print) <- NULL
 head(taxa.print)
 
-save(seqtab.nochim, taxa, file = "Output/dada2_output.RData")
+save(seqtab.nochim, taxa, file = "HPC_Output/dada2_output.RData")
+

### 2022-04-27 Manuscript/manuscript.Rmd
update data availability statement

@@
 
-The 16S rRNA sequencing data generated during the current study are available in the NCBI Sequence Read Archive (SRA).
+The 16S rRNA sequencing data generated during the current study are available in the NCBI Sequence Read Archive (SRA) under BioProject accession number PRJNA832314, https://www.ncbi.nlm.nih.gov/sra/PRJNA832314.
+The metadata generated during the current study as well as the scripts to recreate the analysis are available on Github, https://github.com/amorris28/artificial_ecosystem_selection.
 

### 2022-05-25 R/07-process-community.Rmd
remove phyloseq and replace with tibbles

---
title: "Sequence data processing"
author: "Andrew Morris"
date: "9/13/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
knitr::opts_knit$set(root.dir=normalizePath('../'))
```

```{r initialize-packages}
library(tidyverse)
#library(phyloseq)
library(decontam)
#library(microViz)
library(knitr)
library(vegan)
```

```{r import-data}
load(file = "HPC_Output/dada2_output.RData")

samples <- read_tsv('Output/barcode_master.tsv')
fluxes <- read_tsv('Output/fluxes.tsv')
qubit <- read_tsv('Data/post_pcr_qubit.tsv')

```

```{r data-munging}

metadata <-
  samples %>% 
  rename(sample = sample_id) %>% 
  left_join(qubit, by = c("pcr_well" = "well")) %>% 
  mutate(treat = (Vectorize(function (x) {if(x %in% c('PCRP', 'PCRN')) x
    else substr(x, 1, 3)}))(sample)) %>% 
  mutate(col = as.character(col)) %>% 
  select(sample, row, col, treat, qubit_ngml)

rm(qubit)
rm(samples)

# Check if taxa names match between the otu and taxa tables
sum(colnames(seqtab.nochim) != rownames(taxa))
# They match if = 0

colnames(seqtab.nochim) <- paste0("ASV", seq(ncol(seqtab.nochim)))
rownames(seqtab.nochim) <- metadata$sample
rownames(taxa) <- colnames(seqtab.nochim)

asvs <-
  seqtab.nochim %>% 
  as_tibble(rownames = "sample")

rm(seqtab.nochim)

taxa <-
  taxa %>% 
  as_tibble(rownames = "asv")

aes_data <-
  asvs %>% 
  pivot_longer(-sample, names_to = "asv", values_to = "count") %>% 
  inner_join(metadata, by = "sample") %>% 
  inner_join(taxa, by = "asv")

```

# Alpha Diversity for All Data

Here is alpha diversity for each treatment, positive/negative control, etc.
In addition, there is alpha diversity by row and column of the PCR plate. I don't see any relationship between PCR plate location and alpha diversity, which is good. It looks like the PCR positive control, which was a pure culture of E. Coli, has particularly low diversity, which is good. Let's pull that sample out and look at it.

```{r alpha-diversity-treatment}

aes_data %>%
  select(sample, treat, asv, count) %>% 
  filter(count > 0) %>% 
  count(sample, treat, count) %>%
  nest(data = -c(sample,treat)) %>%
  mutate(n_seqs = map_dbl(data, ~sum(.x$count * .x$n)),
         sobs = map_dbl(data, ~sum(.x$n))) %>%
  select(-data) %>%
  select(-sample) %>% 
  group_by(treat) %>% 
  rename(Reads = n_seqs, Richness = sobs) %>% 
  pivot_longer(-c(treat)) %>%
  ggplot(aes(x=treat, y=value)) +
  facet_wrap(~ name, nrow = 3, scales = "free") +
  geom_boxplot() +
  geom_jitter() +
  theme_classic() +
  labs(x = "Treatment", y = "Value")

```

```{r alpha-diversity-row-col}

aes_data %>%
  select(row, col, asv, count) %>% 
  filter(count > 0) %>% 
  mutate(row = as.numeric(factor(row)),
         col = as.numeric(col)) %>% 
  count(row, col, count) %>%
  nest(data = -c(row, col)) %>%
  mutate(n_seqs = map_dbl(data, ~sum(.x$count * .x$n)),
         sobs = map_dbl(data, ~sum(.x$n))) %>%
  select(-data) %>%
  rename(Column = col, Row = row, Richness = sobs, Sequences = n_seqs) %>% 
  pivot_longer(c(Row, Column), names_to = "row_or_column", values_to = "position") %>%
  pivot_longer(c(Sequences, Richness), names_to = "name", values_to = "count") %>%
  ggplot(aes(x=position, y=count)) +
  facet_grid(name ~ row_or_column, scales = "free") +
  geom_point() +
  theme_classic() +
  geom_smooth() + 
  scale_x_continuous(breaks = 1:9)

```

## Top 20 taxa in controls

```{r top20-function}
top20_taxa_in_treatment <- function(x, treatment) {
x %>% 
  filter(treat == treatment) %>% 
  select(-sample, -row, -col, -qubit_ngml) %>% 
  group_by(asv) %>% 
  mutate(count = sum(count)) %>%  
  ungroup() %>% 
  distinct() %>% 
  arrange(desc(count)) %>% 
  slice_head(n = 20)
}
```

### Potting Mix

```{r top20-potting-mix}
top20_taxa_in_treatment(aes_data, "PM0")

```

### Negative Control

```{r top20-negative-control}
top20_taxa_in_treatment(aes_data, "NC0")

```

### PCR Positive Control

```{r top20-pcr-pos}
top20_taxa_in_treatment(aes_data, "PCRP")

```

### PCR Negative Control

```{r top20-pcr-neg}
top20_taxa_in_treatment(aes_data, "PCRN")

```

### Soil Core

```{r top20-soil-core}
top20_taxa_in_treatment(aes_data, "SC0")

```

# Sequence Depth Distribution + Sample Composition

```{r}

# Look at distribution of sequencing depths
aes_data %>%
  select(sample, asv, count) %>% 
  group_by(sample) %>% 
  summarize(total = sum(count)) %>% 
  ggplot(aes(x = total)) +
  geom_histogram() +
  labs(x = "Sequences", y = "Count")

```

```{r}

aes_df <- 
  aes_data %>%
  select(sample, asv, count) %>% 
  pivot_wider(names_from=asv, values_from=count, values_fill = 0) %>%
  as.data.frame()
rownames(aes_df) <- aes_df$sample
aes_df <- aes_df[,-1]

dist <- vegdist(aes_df, method="bray")
pcoa <- cmdscale(dist, eig = TRUE, add = TRUE)
positions = pcoa$points
colnames(positions) <- paste0("pcoa", 1:ncol(positions))

percent_explained <- 100 * pcoa$eig / sum(pcoa$eig)

tibble(pe = percent_explained, 
       axis=1:length(percent_explained)) %>% 
  ggplot(aes(x = axis, y = pe)) +
  geom_line() +
  coord_cartesian(xlim = c(1,10)) +
  labs(title = "Bray PCoA Scree Plot", y = "Precent Explained", x = "PCo Axis")

labs <- c(paste0("PCo 1 (", format(round(percent_explained[1], 1), nsmall = 1), "%)"),
          paste0("PCo 2 (", format(round(percent_explained[2], 1), nsmall = 1), "%)"))

positions <- 
  positions %>% 
  as_tibble(rownames = "sample")

bray_data <-
  aes_data %>% 
  select(sample, treat, row, col) %>% 
  distinct %>% 
  inner_join(positions, by = "sample")

bray_data %>% 
  ggplot(aes(x = pcoa1, y = pcoa2, color = treat)) +
  geom_point() +
  labs(x = labs[1], y = labs[2], title = "Bray PCoA by Treatment") +
  coord_fixed(percent_explained[2]/percent_explained[1], clip="off")

bray_data %>% 
  ggplot(aes(x = pcoa1, y = pcoa2, color = col)) +
  geom_point() +
  labs(x = labs[1], y = labs[2], title = "Bray PCoA by PCR plate column") +
  coord_fixed(percent_explained[2]/percent_explained[1], clip="off")

bray_data %>% 
  ggplot(aes(x = pcoa1, y = pcoa2, color = row)) +
  geom_point() +
  labs(x = labs[1], y = labs[2], title = "Bray PCoA by PCR plate row") +
  coord_fixed(percent_explained[2]/percent_explained[1], clip="off")

```

# DECONTAM

```{r decontam-n-seqs}

aes_data %>% 
  filter(!treat %in% c('PM0', 'SC0', 'PCRN', 'PCRP')) %>% 
  mutate(ctrl_smpl = case_when(treat == "NC0" ~ "control", 
                               TRUE ~ "sample")) %>% 
  add_count(sample, count) %>%
  nest(data = -c(sample, ctrl_smpl)) %>%
  mutate(n_seqs = map_dbl(data, ~sum(.x$count * .x$n))) %>% 
  arrange(n_seqs) %>% 
  mutate(index = seq(nrow(.))) %>% 
  ggplot(aes(x=index, y=n_seqs, color=ctrl_smpl)) + 
  geom_point() +
  labs(y = "Library Size", x = "Index")

```

My negative controls have a lower library size than my samples.

## Decontam Combined Prevalence and Frequency

Here, I use DNA concentration values from Qubit and prevalence in negative
controls to identify potential contaminants. Several negative controls
had DNA concentration below the detection limit of the instrument and these
were set to 0.50 ng/mL, which is close to the DNA concentration of the
negative controls that were within the detection limit and is at the limit of the 
Qubit's detection limit.

In the plots below, ASV1 is a true sequence. The others are the ASVs identified
by decontam as potential contaminants.

```{r decontam-prep}

otu_table <-
  aes_data %>% 
  filter(!treat %in% c('PM0', 'SC0', 'PCRN', 'PCRP')) %>% 
    mutate(control = case_when(treat == "NC0" ~ TRUE, 
                               TRUE ~ FALSE)) %>% 
  select(qubit_ngml, control, sample, asv, count) %>% 
  pivot_wider(names_from = asv, values_from = count) %>% 
  mutate(qubit_ngml = case_when(qubit_ngml == 0 ~ 0.5, 
                               TRUE ~ qubit_ngml)) %>% 
  as.data.frame
concentration <- otu_table$qubit_ngml
otu_table <- otu_table[, -1]
control <- otu_table$control
otu_table <- otu_table[, -1]
rownames(otu_table) <- otu_table$sample
otu_table <- otu_table[, -1]
otu_mat <- as.matrix(otu_table)

```

```{r decontam}
contamdf <- isContaminant(otu_mat, conc = concentration, neg=control)

alleged_contaminants <- which(contamdf$contaminant)

quant_test <- 
  otu_mat[,c(1,alleged_contaminants)] %>% 
  as.data.frame %>% 
  as_tibble(rownames = "sample") %>% 
  mutate(conc = concentration) %>% 
  mutate(cont = control) %>% 
  pivot_longer(starts_with("ASV"))

ggplot(quant_test, aes(x = conc, y = value, color = cont)) +
  facet_wrap(~ name, scales = "free") +
  geom_point()

contaminants <- 
  which(contamdf$contaminant)[which(contamdf$contaminant) %in% 
  c(570, 810, 1283, 1859, 2233, 2777, 2926, 5687, 6696)]

```

Decontam identified `r length(which(contamdf$contaminant))` potential contaminants based
on prevalence and frequency. Visual inspection of abundance ~ concentration
plots indicated that `r length(contaminants)` were likely contaminants. These
ASVs were removed.

```{r remove-contaminants}
aes_data_no_contam <-
 aes_data %>%
  nest(data = -asv) %>% 
  filter(!asv %in% paste0("ASV", contaminants)) %>% 
  unnest(data)

```

# Combine ASV table and flux data

```{r combine-data}

flux_sd <- 
  fluxes %>% 
  mutate(sample = paste0('G', passage, toupper(treat), rep)) %>% 
  select(sample, flux_date:selected) %>% 
  mutate(passage = as.character(passage))

flux_sd <-
  flux_sd %>% 
  select(sample, ch4, selected)

full_data <-
  aes_data %>% 
  left_join(flux_sd, by = "sample")

write_tsv(full_data, 'Output/communty_data.tsv')

```

```{r old-stuff, eval = FALSE}

samples_filtered <- 
filter(samples, sample_id %in% flux_sd$sample)

fluxes_filtered <- 
  filter(flux_sd, sample_id %in% samples$sample_id) %>% 
  as.data.frame()

```

### 2022-05-25 R/07-process-community.Rmd
remove phyloseq and replace with tibbles
revise processing script and shrink otu table

@@
   select(sample, flux_date:selected) %>% 
-  mutate(passage = as.character(passage))
-
-flux_sd <-
-  flux_sd %>% 
+  mutate(passage = as.character(passage)) %>% 
   select(sample, ch4, selected)
@@
   aes_data %>% 
+  filter(!treat %in% c("NC0",  "PM0",  "SC0",  "PCRP", "PCRN")) %>% 
+  select(-c(row, col, qubit_ngml)) %>% 
   left_join(flux_sd, by = "sample")
 
-write_tsv(full_data, 'Output/communty_data.tsv')
+# Remove zero-abundance ASVs
+full_data <-
+  full_data %>% 
+  group_by(asv) %>% 
+  mutate(total = sum(count)) %>% 
+  filter(total != 0) %>% 
+  ungroup() %>% 
+  select(-total) %>% 
+  filter(count > 0)
+
+write_tsv(full_data, 'Output/community_data.tsv')
 
 ```
-
-```{r old-stuff, eval = FALSE}
-
-samples_filtered <- 
-filter(samples, sample_id %in% flux_sd$sample)
-
-fluxes_filtered <- 
-  filter(flux_sd, sample_id %in% samples$sample_id) %>% 
-  as.data.frame()
-
-```

### 2022-05-25 R/08-compare-methods.R
run comparison of alpha and beta diversity analyses

@@ -1,0 +1,176 @@
+library(vegan)
+library(breakaway)
+library(magrittr)
+library(zCompositions)
+library(tidyverse)
+source('R/functions.R')
+
+set.seed(19760620)
+
+aes_data <- read_tsv('Output/community_data.tsv')
+
+# Look at distribution of sequencing depths
+aes_data %>% 
+  group_by(sample) %>% 
+  summarize(total = sum(count)) %>% 
+  ggplot(aes(x = total)) +
+  geom_histogram()
+
+
+seq_range <- 
+  aes_data %>%
+  group_by(sample) %>% 
+  summarize(total = sum(count)) %>% 
+  summarize(seq_range = range(total)) %$%
+  seq_range
+
+# Generate a null dataset
+
+rand_aes <- randomize_asv_table(aes_data)
+
+#######################
+# Alpha diversity
+#####################
+
+
+analysis <- rand_aes %>%
+  count(sample, value) %>%
+  nest(data = -sample) %>%
+  mutate(n_seqs = map_dbl(data, ~sum(.x$value * .x$n)),
+         sobs = map_dbl(data, ~sum(.x$n)),
+         rare = map_dbl(data, ~rarefy(rep(.x$value, .x$n), sample=seq_range[1])),
+         chao = map_dbl(data, ~get_chao(.x)),
+         ba = map(data, ~get_breakaway(.x))) %>%
+  select(-data) %>%
+  unnest(ba)
+
+
+analysis %>%
+  select(sample, n_seqs, sobs, chao, rare, est) %>%
+  pivot_longer(-c(sample, n_seqs)) %>%
+  ggplot(aes(x=n_seqs, y=value, color=name)) +
+  geom_point() +
+  geom_smooth(se=FALSE) +
+  coord_cartesian(ylim=c(0, 40000)) +
+  theme_classic() +
+  labs(title = "AES", x = "Sequencing Depth", y = "Alpha Diversity")
+
+analysis %>%
+  select(sample, n_seqs, sobs, chao, rare, est) %>%
+  pivot_longer(-c(sample, n_seqs)) %>%
+  ggplot(aes(x=n_seqs, y=value, color=name)) +
+  geom_point() +
+  geom_smooth(se=FALSE) +
+  coord_cartesian(ylim=c(0, 10000)) +
+  theme_classic() +
+  labs(title = "AES", x = "Sequencing Depth", y = "Alpha Diversity")
+
+##### Evenness
+
[…]

### 2022-05-25 R/functions.R

@@ -156,3 +156,59 @@ dt_plot <- function(dt_data) {
     coord_cartesian(clip="off") +
     theme(panel.border = element_blank())
 }
+
+randomize_asv_table <- function(x) {
+  x %>%
+    uncount(count) %>%
+    mutate(rand_asv = sample(asv)) %>%
+    select(-asv) %>%
+    count(sample, rand_asv, name="value")
+}
+
+
+gm <- function(x){
+  
+  exp(mean(log(x[x>0])))
+  
+}
+
+get_breakaway <- function(x){
+  
+  ba <- breakaway(x)
+  tibble(est= ba$estimate,
+         lci=ba$interval[1], uci=ba$interval[2],
+         model=ba$model)
+  
+}
+
+get_chao <- function(x){
+  
+  sobs <- sum(x$n)
+  sing <- x[x$value == 1, "n"] %>% pull(n)
+  doub <- x[x$value == 2, "n"] %>% pull(n)
+  
+  sobs + sing^2 / (2*doub)
+  
+}
+richness <- function(x){
+  
+  # r <- sum(x > 0)
+  # return(r)
+  
+  sum(x>0)
+}
+shannon <- function(x){
+  
+  rabund <- x[x>0]/sum(x)
+  -sum(rabund * log(rabund))
+  
+}
+
+simpson <- function(x){
+  
+  n <- sum(x)
+  
+  # sum(x * (x-1) / (n * (n-1)))
+  1 - sum((x/n)^2)
+}
+

### 2022-06-24 R/09-community-analysis.Rmd
rename communty analysis file

--- 2022-04-09 R/08-community.Rmd
@@
 library(ggpubr)
-# library(SpiecEasi)
 library(ggthemes)
@@
 ```
+
 There is no systematic difference in sequencing depth/library size between treatments. It does appear that sequencing depth decreases between passage 2 and 5.
@@
 
-```{r, eval=FALSE}
+```{r richness}
+
+# Test effect of passage number on richness using all samples
 richness_soil <- ps %>% breakaway
@@
 
-save(bt_passage, file = 'Output/richness_model_passage.Rdata')
+# Test effect of treatment on richness in passage 5
 
@@
 
-save(bt_treat5, file = 'Output/richness_model_treat5.Rdata')
+# Test effect of CH4 on richness
 
@@
 
-save(bt_ch4, file = 'Output/richness_model_ch4.Rdata')
+save(bt_passage, bt_treat5, bt_ch4, file = 'Output/richness_models.RData')
 
@@
 and Passage 5. No difference in richness by treatment in passage 5 or by
-methane.
+methane flux.
 
@@
 
+# Replace unknown and missing taxa
 ps_fixed <-
@@
 
+# Recode treatments as binary 1/0
 ps_fixed <- 
@@
 ps_fixed
-```
-
-## Beta diversity test passage by treatment
-
-```{r}
-plot_beta(ps_fixed)
 ```
@@
 
-```{r plot-beta, results=FALSE, fig.width = 6, fig.height = 2}
+## Bray-Curtis and Jaccard
+
+```{r plot-beta, results=FALSE, fig.width = 8, fig.height = 3}
 bray <-
@@
                        labels = c('Neutral', 'Positive'),
-                       values = c('gray40', 'darkorange2'))
+                       values = c('gray40', 'darkorange2')) +
+  labs(title = "Bray-Curtis")
 
@@
                        labels = c('Neutral', 'Positive'),
-                       values = c('gray40', 'darkorange2'))
+                       values = c('gray40', 'darkorange2')) +
+  labs(title = "Jaccard")
 
@@
 
-```{r plot-aitch, fig.width = 4, fig.height = 4}
+## Aitchison
 
-ord_calc_aitch %>% 
-  ord_plot(color = "treat",
-           auto_caption = FALSE) +
-  theme_tufte() +
-  geom_rangeframe() +
-  coord_fixed(5.5/43.6, clip="off") +
-  scale_color_manual(name = "Treatment", 
-                       labels = c('Neutral', 'Positive'),
-                       values = c('gray40', 'darkorange2'))
-
+```{r}
+plot_beta(ps_fixed)
 ```
 
-```{r permanova-aitch, eval=FALSE}
-dist_aitch <-
-  ps %>% 
-  dist_calc(dist = "aitchison")
-
-dist_permanova(dist_aitch, c("passage", "treat"), "passage * treat")
-
-passage_groups <- factor(c(rep(1,22), rep(2,24)), labels = c("2","5"))
-
-dist_aitch %>% 
-  dist_get %>% 
-  vegan::betadisper(passage_groups) %>% 
-  permutest()
-
-treat_groups <- factor(c(rep(1,12), rep(2,10), rep(1,12), rep(2,12)), labels = c("n","p"))
-
-dist_aitch %>% 
-  dist_get %>% 
-  vegan::betadisper(treat_groups) %>% 
-  permutest()
-
-ps %>%
-  ps_filter(passage == 2) %>% 
-  dist_calc(dist = "aitchison") %>% 
-  dist_permanova("treat")
-
-ps %>%
-  ps_filter(passage == 5) %>% 
-  dist_calc(dist = "aitchison") %>% 
-  dist_permanova("treat")
-
-adonis2(dist_get(dist_aitch) ~ passage * treat, data = as(sample_data(ps_fixed), 'data.frame'))
-
-dist_aitch2 <-
-  ps %>% 
-  ps_filter(passage == 2) %>% 
-  dist_calc(dist = "aitchison")
-dist_permanova(dist_aitch2, "treat")
-groups <- factor(c(rep(1,12), rep(2,10)), labels = c("N","P"))
-
-dist_aitch2 %>% 
-  dist_get %>% 
-  vegan::betadisper(groups) %>% 
-  permutest()
-dist_aitch5 <-
-  ps %>% 
-  ps_filter(passage == 5) %>% 
-  dist_calc(dist = "aitchison")
-dist_permanova(dist_aitch5, "treat")
-groups <- factor(c(rep(1,12), rep(2,12)), labels = c("N","P"))
-
-dist_aitch5 %>% 
-  dist_get %>% 
-  vegan::betadisper(groups) %>% 
-  permutest()
-
-```
-
-There is a significant main effect and interaction of passage and treatment on
-taxonomic dissimilarity of the soil microbiome. In addition, I reject the null
-hypothesis of homogeneity of group dispersions for passage, but not for treatment.
-Inspecting the NMDS plot, from this I conclude that passaging between Passage 2
-and 5 both altered the soil microbiome taxonomic composition and reduced its 
-variability. In addition, there is a difference in centroid between the two
-treatments in both Passage 2 and Passage 5, but no difference in dispersion.
+## Beta diversity test passage by treatment
 
 ```{r beta-test}
-# ps_prop <- transform_sample_counts(ps, function(otu) otu/sum(otu))
-# 
-# fit_bray <- adonis2(otu_table(ps_prop) ~ passage * treat, data = as(sample_data(ps_fixed), 'data.frame'), method = "bray")
-# 
-# fit_bray
-# 
-# fit_jacc <- adonis2(otu_table(ps_prop) ~ passage * treat, data = as(sample_data(ps_fixed), 'data.frame'), method = "jacc", binary = TRUE)
-# 
-# fit_jacc
+
 ps_clr <- microbiome::transform(ps, "clr")
@@
       'Passage', 'Treatment', 'Pass:Treat', 'Residual', 'Total')
-save(fit_adonis, file = 'Output/beta_model.Rdata')
+save(fit_adonis, file = 'Output/beta_model.RData')
 
@@
 
-###########3
+```
 
-# adonis2(otu_table(ps_filter(ps_prop, passage == 5)) ~ treat, 
-#         data = 
-#           as(
-#             sample_data(
-#               ps_filter(
-#                 ps_prop, 
-#                 passage == 5)
-#             ), 
-#             'data.frame'), 
-#         method = "bray")
-# adonis2(otu_table(ps_filter(ps_prop, passage == 5)) ~ treat, 
-#         data = 
-#           as(
-#             sample_data(
-#               ps_filter(
-#                 ps_prop, 
-#                 passage == 5)
-#             ), 
-#             'data.frame'), 
-#         method = "jaccard",
-#         binary = TRUE)
-
-```
+There is a significant main effect and interaction of passage and treatment on
+taxonomic dissimilarity of the soil microbiome. In addition, there is no difference
+in dispersion between treatments, but there is between passage2 and 5.
 
@@
 This ASV had a prevalence of 10/12 and a median abundance of ~2 reads. The other
-prevalent family was a member of the Silvanigrellaceae, a newly described family places in
+prevalent family was a member of the Silvanigrellaceae, a newly described family placed in
 its own order. This family was present in
@@
 
-```{r, eval=FALSE}
-
-g5<- 
-  ps_fixed %>% 
-  ps_filter(passage == 5) %>% 
-  tax_agg('Family') %>% 
-  ps_get() %>% 
-  ps_melt()
-
-g2<- 
-  ps_fixed %>% 
-  ps_filter(passage == 2) %>% 
-  tax_agg('Family') %>% 
-  ps_get() %>% 
-  ps_melt()
-
-unique(g5$Family)[!(unique(g5$Family) %in% unique(g2$Family))]
-unique(g5$Family)[unique(g5$Family) %in% unique(g2$Family)]
-
-unique(g2$Family)[!(unique(g2$Family) %in% unique(g5$Family))]
-unique(g2$Family)[unique(g2$Family) %in% unique(g5$Family)]
-
-```
-
-## Differential Abundance Treatment in passage 5
-
-```{r corncob_treatment, results=FALSE, fig.hight=40}
-ps_fixed_family <- speedyseq::tax_glom(ps_fixed, taxrank="Family")
-
-ps_fam_pas5 <- microViz::ps_filter(ps_fixed_family, passage == 5)
-
-ps_fam_pas5_prev <- microViz::tax_filter(ps_fam_pas5, min_prevalence = 0.10)
-
-sample_data(ps_fam_pas5_prev) =
-  sample_data(ps_fam_pas5_prev)[,c('sample_id', 'treat')] %>% 
-  sample_data()
-ps = ps_fam_pas5_prev
-
-save(ps, file = 'ps.RData')
-# Differential abundance test
-corncob <- bbdml(formula = ASV2954 ~ treat,
-                                 phi.formula = ~ 1,
-                                 formula_null = ~ 1,
-                                 phi.formula_null = ~ 1,
-                                 test = "lrt", boot = TRUE,
-                                 data = ps_fam_pas5_prev,
-                                 fdr_cutoff = 0.05)
-da_analysis <- differentialTest(formula = ~ treat,
-                                 phi.formula = ~ 1,
-                                 formula_null = ~ 1,
-                                 phi.formula_null = ~ 1,
-                                 test = "Wald", boot = FALSE,
-                                 data = ps_fam_pas5_prev,
-                                 fdr_cutoff = 0.05, B = 50)
-plot(da_analysis, level = c("Family"))
-da_analysis$significant_models
-dt_data <- plot(da_analysis, level = c("Family"), data_only = TRUE)
-
-sig_taxa_treat <- tax_select(ps_fam_pas5_prev, da_analysis$significant_taxa)
-ntaxa(sig_taxa_treat)
-save(da_analysis, sig_taxa_treat, file = 'Output/da_corncob.Rdata')
-
-save(dt_data, file = 'Output/dt_data.Rdata')
-
-```
-
-```{r corncob-plot, fig.width=6, fig.height=6}
-dt_plot(dt_data)
-
-```
-
-```{r stahp}
-knitr::opts_chunk$set(eval = FALSE)
-
-```
-
-```{r stacked_bar, eval=FALSE}
-level = "Family"
-ps_phylum <- speedyseq::tax_glom(ps, taxrank=level)
-top10 = names(sort(taxa_sums(ps_phylum), TRUE)[1:10])
-ps_phylum10 <- prune_taxa(top10, ps_phylum)
-ps_phylum10 %>% 
-  ps_filter(passage == 5) %>% 
-  speedyseq::plot_bar(facet_grid = 'treat', fill = level) + 
-  geom_bar(aes_string( color=level, fill=level), stat="identity", position="stack") +
-  geom_bar(aes(x = reorder(rep, rank_ch4)))
-
-```
-
-## ALDex2
-
-```{r, eval = FALSE}
-library(ALDEx2)
-data(selex)
-#subset only the last 400 features for efficiency
-selex.sub <- selex[1:400,]
-conds <- c(rep("NS", 7), rep("S", 7))
-x <- aldex.clr(selex.sub, conds, mc.samples=16, denom="iqlr", verbose=F)
-x.tt <- aldex.ttest(x, paired.test=FALSE, verbose=FALSE)
-x.effect <- aldex.effect(x, CI=T, verbose=FALSE)
-head(x.effect)
-
-```
-
-```{r, eval=FALSE}
-ps_fixed_family <- speedyseq::tax_glom(ps_fixed, taxrank="Family")
-
-ps_fam_pas5 <- microViz::ps_filter(ps_fixed_family, passage == 5)
-
-ps_fam_pas5_prev <- microViz::tax_filter(ps_fam_pas5, min_prevalence = 0.10)
-ps_pass_5 <- microViz::ps_filter(ps_fixed, passage == 5)
-otu_table <- as(t(otu_table(ps_fam_pas5_prev)), 'matrix')
-conds <- c(rep("n", 12), rep("p", 12))
-
-# Based on effect confidence interval
-
-x <- aldex.clr(otu_table, conds, mc.samples=128, denom="all", verbose=F)
-x.tt <- aldex.ttest(x, paired.test=FALSE, verbose=FALSE)
-x.effect <- aldex.effect(x, CI=T, verbose=FALSE)
-
-sgn <- sign(x.effect$effect.low) == sign(x.effect$effect.high)
-par(mfrow=c(1,2))
-plot(x.effect$rab.all, x.effect$diff.btw, pch=19, cex=0.3, col="grey", xlab="Abundance", ylab="Difference", main="Bland-Altman")
-points(x.effect$rab.all[abs(x.effect$effect) >=2], x.effect$diff.btw[abs(x.effect$effect) >=2], pch=19, cex=0.5, col="red")
-points(x.effect$rab.all[sgn], x.effect$diff.btw[sgn], cex=0.7, col="blue")
-
-plot(x.effect$diff.win, x.effect$diff.btw, pch=19, cex=0.3, col="grey",xlab="Dispersion", ylab="Difference", main="Effect")
-points(x.effect$diff.win[abs(x.effect$effect) >=2], x.effect$diff.btw[abs(x.effect$effect) >=2], pch=19, cex=0.5, col="red")
-points(x.effect$diff.win[sgn], x.effect$diff.btw[sgn], cex=0.7, col="blue")
-abline(0,2, lty=2, col="grey")
-abline(0,-2, lty=2, col="grey")
-
-merge(as(tax_table(ps_fam_pas5_prev)[rownames(x.effect[sgn,]), ], 'data.frame'), x.effect[sgn,], by = 0)
-```
-
-```{r, eval=FALSE}
-
-# Based on effect size
-
-x.all <- aldex(otu_table, conds, mc.samples=128, test="t", effect=TRUE,
-     include.sample.summary=FALSE, denom="all", verbose=FALSE)
-effect_gt_2 <- merge(as(tax_table(ps_pass_5)[rownames(x.all[abs(x.all$effect)>2,]), ], 'data.frame'), x.all[abs(x.all$effect)>2,], by = 0)
-
-effect_gt_2 %>% 
-  ggplot(aes(x = effect, y = reorder(Family,effect))) +
-  geom_point()
-
-par(mfrow=c(1,2))
-aldex.plot(x.all, type="MA", test="welch", xlab="Log-ratio abundance",
-    ylab="Difference")
-aldex.plot(x.all, type="MW", test="welch", xlab="Dispersion",
-    ylab="Difference")
-
-par(mfrow=c(1,2))
-plot(x.all$effect, x.all$we.ep, log="y", cex=0.7, col=rgb(0,0,1,0.2),
-  pch=19, xlab="Effect size", ylab="P value", main="Effect size plot")
-points(x.all$effect, x.all$we.eBH, cex=0.7, col=rgb(1,0,0,0.2),
-  pch=19)
-abline(h=0.05, lty=2, col="grey")
-legend(15,1, legend=c("P value", "BH-adjusted"), pch=19, col=c("blue", "red"))
-
-plot(x.all$diff.btw, x.all$we.ep, log="y", cex=0.7, col=rgb(0,0,1,0.2),
-  pch=19, xlab="Difference", ylab="P value", main="Volcano plot")
-points(x.all$diff.btw, x.all$we.eBH, cex=0.7, col=rgb(1,0,0,0.2),
-  pch=19)
-abline(h=0.05, lty=2, col="grey")
-
-# identify which values are significant in both the t-test and glm tests
-found.by.all <- which(x.all$we.eBH < 0.05 & x.all$wi.eBH < 0.05)
-
-# identify which values are significant in fewer than all tests
-found.by.one <- which(x.all$we.eBH < 0.05 | x.all$wi.eBH < 0.05)
-
-# plot the within and between variation of the data
-plot(x.all$diff.win, x.all$diff.btw, pch=19, cex=0.3, col=rgb(0,0,0,0.3),
- xlab="Dispersion", ylab="Difference")
-points(x.all$diff.win[found.by.one], x.all$diff.btw[found.by.one], pch=19,
- cex=0.7, col=rgb(0,0,1,0.5))
-points(x.all$diff.win[found.by.all], x.all$diff.btw[found.by.all], pch=19,
- cex=0.7, col=rgb(1,0,0,1))
-abline(0,1,lty=2)
-abline(0,-1,lty=2)
-
-#View(tax_table(ps_fam_pas5_prev)[rownames(otu_table[found.by.all, ]), ])
-```
-
-```{r, eval=FALSE}
-boxData <-
-  ps_fixed %>% 
-  ps_filter(passage ==5) %>% 
-  tax_agg('Family') %>% 
-  ps_get %>% 
-  ps_melt()
-
-Family_order<-
-  boxData %>% 
-  group_by(Family) %>% 
-  summarize(median = median(Abundance)) %>% 
-  arrange(median)
-boxData$Family <- factor(boxData$Family, levels = Family_order$Family)
-boxData$treat <- factor(boxData$treat, levels = c('p', 'n'))
-```
-
-```{r, eval=FALSE, fig.height=3, fig.width=12}
-
-boxData %>% 
-  ggplot(aes(x = Family, 
-             y = Abundance)) +
-  facet_grid(treat ~ .) +
-  theme_tufte() +
-  geom_rangeframe(color = "black") +
-  # scale_x_reordered() +
-  scale_y_continuous(trans = "pseudo_log") +
-  geom_tufteboxplot() +
-  theme(axis.text.x = element_text(angle = 90))
-  # geom_boxplot() +
-  # scale_color_discrete(guide = 'none') +
-  # coord_flip()
-
-```
-
-```{r, eval=FALSE}
+```{r}
+# families in to Passage 5, Treatment P
 g5p<- 
@@
   ps_melt()
+
+# families in to Passage 5, Treatment N
 g5n<- 
@@
 
+# Plot families unique to P5 in order of median abundance
 g5p_unique %>% 
@@
   theme(axis.text.x = element_text(angle = 90)) +
-  scale_y_continuous(breaks = c(0, 5, 10, 15, 20, 24))
+  scale_y_continuous(breaks = c(0, 5, 10, 15, 20, 24)) +
+  labs(title = "Unique to P5 by abundance")
 
@@
 
+# Plot families unique to P5 in order of prevalence
 g5p_unique %>% 
@@
   theme(axis.text.x = element_text(angle = 90)) +
-  scale_y_continuous(breaks = c(1, 2, 3, 5, 10, 12))
+  scale_y_continuous(breaks = c(1, 2, 3, 5, 10, 12)) +
+  labs(title = "Unique to P5 by prevalence")
 
@@
 
+# Plot families unique to N5 in order of abundance
 g5n_unique %>% 
@@
   theme(axis.text.x = element_text(angle = 90)) +
-  scale_y_continuous(breaks = c(0, 10, 20, 30, 37))
+  scale_y_continuous(breaks = c(0, 10, 20, 30, 37)) +
+  labs(title = "Unique to N5 by abundance")
 
@@
 
+# Plot families unique to N5 in order of prevalence
+
 g5n_unique %>% 
@@
   theme(axis.text.x = element_text(angle = 90)) +
-  scale_y_continuous(breaks = c(1, 2, 6))
-```
-
-## Faprotax on DA taxa
-
-```{r export_to_faprotax, eval=FALSE}
-faprotax_taxonomy <-
-  sig_taxa_treat %>% 
-  tax_table %>% 
-  as_tibble %>% 
-  mutate(.taxonomy = 
-           faprotax_taxonomy %>% 
-           select(Kingdom:Species) %>% 
-           apply(1 , paste , collapse = "; " )) %>% 
-  select(.otu, .taxonomy)
-
-otu<-t(as(otu_table(sig_taxa_treat),"matrix")) 
-  # 't' to transform if taxa_are_rows=FALSE
-  #if taxa_are_rows=TRUE
-  #otu<-as(otu_table(GlobalPatterns),"matrix"))
-rownames(otu) <- faprotax_taxonomy$.taxonomy
-otu_biom <- biomformat::make_biom(data=otu)
-biomformat::write_biom(otu_biom,"otu_biom.biom")
+  scale_y_continuous(breaks = c(1, 2, 6)) +
+  labs(title = "Unique to N5 by prevalence")
 
@@
 
-```{bash faprotax, eval=FALSE}
-conda activate faprotax
+## Differential Abundance Treatment in passage 5
 
-rm Output/func_table_from_biom.biom \
-    Output/faprotax_report_from_biom.txt
+```{r corncob_treatment, results=FALSE, fig.hight=40}
+ps_fixed_family <- speedyseq::tax_glom(ps_fixed, taxrank="Family")
 
-python python/FAPROTAX_1.2.4/collapse_table.py \
-    -i otu_biom.biom \
-    -g python/FAPROTAX_1.2.4/FAPROTAX.txt \
-    -o Output/func_table_from_biom.biom \
-    -r Output/faprotax_report_from_biom.txt \
-    -v
+ps_fam_pas5 <- microViz::ps_filter(ps_fixed_family, passage == 5)
 
-biom convert \
-    -i Output/func_table_from_biom.biom \
-    -o Output/func_table_from_biom_json.biom \
-    --table-type="OTU table" \
-    --to-json
+ps_fam_pas5_prev <- microViz::tax_filter(ps_fam_pas5, min_prevalence = 0.10)
 
-conda deactivate
-```
+sample_data(ps_fam_pas5_prev) =
+  sample_data(ps_fam_pas5_prev)[,c('sample_id', 'treat')] %>% 
+  sample_data()
+ps = ps_fam_pas5_prev
 
-```{r, eval=FALSE}
-biom_file <- "Output/func_table_from_biom_json.biom"
+# Differential abundance test
+da_analysis <- differentialTest(formula = ~ treat,
+                                 phi.formula = ~ 1,
+                                 formula_null = ~ 1,
+                                 phi.formula_null = ~ 1,
+                                 test = "Wald", boot = FALSE,
+                                 data = ps_fam_pas5_prev,
+                                 fdr_cutoff = 0.05, B = 50)
 
-ps_faprotax <-
-  biom_file %>% 
-  biomformat::read_biom() %>% 
-  biomformat::biom_data() %>% 
-  as("matrix") %>% 
-  otu_table(taxa_are_rows=TRUE) %>% 
-  phyloseq(sample_data(ps_gen_pas5_prev))
+#plot(da_analysis, level = c("Family"))
+da_analysis$significant_models
+dt_data <- plot(da_analysis, level = c("Family"), data_only = TRUE)
 
-da_faprotax <- differentialTest(formula = ~ treat,
-                                 phi.formula = ~ treat,
-                                 formula_null = ~ 1,
-                                 phi.formula_null = ~ treat,
-                                 test = "Wald", boot = FALSE,
-                                 data = ps_faprotax,
-                                 fdr_cutoff = 0.05)
-plot(da_faprotax)
+sig_taxa_treat <- tax_select(ps_fam_pas5_prev, da_analysis$significant_taxa)
+ntaxa(sig_taxa_treat)
+save(da_analysis, sig_taxa_treat, dt_data, file = 'Output/da_corncob.RData')
 
@@
 
-## Picrust on DA taxa
-
-```{r}
-x = NULL
-for (i in seq_along(da_analysis$significant_models)) {
-  x <- c(x, da_analysis$significant_models[[i]]$coefficients[2, 1])
-}
-x > 0
-sig_taxa_pos_neg <-
-  da_analysis$significant_taxa %>% 
-  as_tibble_col(column_name = ".otu") %>% 
-  mutate(estimate = x) %>% 
-  mutate(pos_or_neg = if_else(x > 0, "pos", "neg"))
-
-top_da_taxa <- 
-  sig_taxa_treat %>% 
-  otu_table %>% 
-  as_tibble %>% 
-  filter(.abundance > 2000) %>% 
-  distinct(.otu) %$% .otu
-
-taxonomy <-
-  sig_taxa_treat %>% 
-  tax_table %>% 
-  as_tibble %>%
-  filter(.otu %in% top_da_taxa)
-
-taxonomy$.taxonomy <- taxonomy %>% 
-           select(Kingdom:Species) %>% 
-           apply(1 , paste , collapse = "; " )
-
-taxonomy <-
-  taxonomy %>% 
-  select(.otu, .taxonomy)
-
-df_top_da_taxa <-
-  sig_taxa_treat %>% 
-  otu_table %>% 
-  as_tibble %>%
-  left_join(as_tibble(sample_data(sig_taxa_treat))) %>% 
-  left_join(sig_taxa_pos_neg, by = ".otu") %>% 
-  filter(.otu != 'ASV4174') %>% 
-  filter(.otu %in% top_da_taxa) %>% 
-  left_join(taxonomy)
-
-df_top_da_taxa %>% 
-  ggplot(., aes(y = Genus, x = .abundance, color = treat)) +
-  geom_jitter(height = 0.1) + 
-  #theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  facet_wrap(~ pos_or_neg)
-ggsave('Output/dif_abund_results.png', width = 6, height = 4)
-
-major_otus <- 
-sig_taxa_treat %>% 
-  otu_table %>% 
-  as_tibble %>%
-  left_join(as_tibble(sample_data(sig_taxa_treat))) %>% 
-  left_join(sig_taxa_pos_neg, by = ".otu") %>% 
-  filter(.otu != 'ASV4174') %>% 
-  filter(.abundance > 300) %>% 
-  select(.otu, pos_or_neg) %>% 
-  distinct %>% 
-  mutate(tax = otu_to_taxonomy(.otu, sig_taxa_treat))
-
-otu_to_taxonomy(major_otus$.otu, sig_taxa_treat)
-
-otu_to_taxonomy('ASV192', sig_taxa_treat)
-
-sig_taxa_treat %>% 
-  otu_table %>% 
-  as_tibble %>%
-  left_join(as_tibble(sample_data(sig_taxa_treat))) %>% 
-  left_join(sig_taxa_pos_neg, by = ".otu") %>% 
-  filter(.otu == 'ASV192') %>% 
-  ggplot(., aes(x = .otu, y = .abundance, color = treat)) +
-  geom_point() + 
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  facet_wrap(~ pos_or_neg)
-
-beijer_otu <-
-  ps %>% 
-  otu_table %>% 
-  as_tibble %>% 
-  filter(.otu == 'ASV192')
-
-beijer_metadata <-
-  ps %>% 
-  sample_data %>% 
-  as_tibble %>% 
-    left_join(beijer_otu)
-
-ggplot(beijer_metadata, aes(x = .abundance, y = log_ch4, color = treat)) +
-  geom_point() +
-  facet_wrap(~ passage)
-
-df_top_da_taxa %>% 
-filter(pos_or_neg == "pos") %>% 
-select(Phylum:Genus) %>% 
-distinct %>% 
-  write_tsv(., 'top_positive_taxa.txt')
-```
-
-```{bash, eval=FALSE}
-mkdir picrust2
-```
-
-```{r export_to_picrust}
-
-top_da_taxa_pos <-
-  df_top_da_taxa %>% 
-  select(.otu, pos_or_neg) %>% 
-  distinct(.otu, pos_or_neg) %>% 
-  filter(pos_or_neg == "pos") %$%
-  .otu
-sig_taxa_treat %>% 
-  tax_select(top_da_taxa_pos) %>% 
-  otu_table %>% 
-  as("matrix") %>% 
-  t %>% 
-  biomformat::make_biom() %>% 
-  biomformat::write_biom("Output/study_seqs.biom")
-
-sequences <-
-  sig_taxa_treat %>% 
-  tax_select(top_da_taxa_pos) %>% 
-  refseq %>% 
-  as.vector %>% 
-  as.list
-seqinr::write.fasta(sequences, names(sequences), file.out = "Output/study_seqs.fna")
+```{r corncob-plot, fig.width=6, fig.height=6}
+dt_plot(dt_data)
 
@@
 
-```{bash picrust, eval=FALSE}
-conda activate picrust2
-cd picrust3c
-
-place_seqs.py -s ../Output/study_seqs.fna \
-    -o out.tre \
-    -p 1 \
-    --intermediate intermediate/place_seqs \
-    --chunk_size 500
-hsp.py -i 16S -t out.tre -o marker_predicted_and_nsti.tsv.gz -p 1 -n
-
-hsp.py -i EC -t out.tre -o EC_predicted.tsv.gz -p 1
-
-metagenome_pipeline.py -i ../Output/study_seqs.biom -m marker_predicted_and_nsti.tsv.gz -f EC_predicted.tsv.gz \
-                       -o EC_metagenome_out --strat_out
-
-pathway_pipeline.py -i EC_metagenome_out/pred_metagenome_contrib.tsv.gz \
-                    -o pathways_out -p 1
-
-add_descriptions.py -i EC_metagenome_out/pred_metagenome_unstrat.tsv.gz -m EC \
-                    -o EC_metagenome_out/pred_metagenome_unstrat_descrip.tsv.gz
-
-add_descriptions.py -i pathways_out/path_abun_unstrat.tsv.gz -m METACYC \
-                    -o pathways_out/path_abun_unstrat_descrip.tsv.gz
-
-#####
-
-picrust2_pipeline.py -s Output/study_seqs.fna -i Output/study_seqs.biom -o picrust2_out_pipeline -p 4 --stratified
-
-conda deactivate
-```
-
-```{r}
-se.mb.amgut <- spiec.easi(amgut1.filt, method='mb', lambda.min.ratio=1e-2,
-                          nlambda=20, pulsar.params=list(rep.num=50))
-se.gl.amgut <- spiec.easi(amgut1.filt, method='glasso', lambda.min.ratio=1e-2,
-                          nlambda=20, pulsar.params=list(rep.num=50))
-sparcc.amgut <- sparcc(amgut1.filt)
-## Define arbitrary threshold for SparCC correlation matrix for the graph
-sparcc.graph <- abs(sparcc.amgut$Cor) >= 0.3
-diag(sparcc.graph) <- 0
-library(Matrix)
-sparcc.graph <- Matrix(sparcc.graph, sparse=TRUE)
-## Create igraph objects
-ig.mb     <- adj2igraph(getRefit(se.mb.amgut))
-ig.gl     <- adj2igraph(getRefit(se.gl.amgut))
-ig.sparcc <- adj2igraph(sparcc.graph)
+```{r stop}
+knitr::opts_chunk$set(eval = FALSE)
 
@@
 
-```{r}
-library(igraph)
-## set size of vertex proportional to clr-mean
-vsize    <- rowMeans(clr(amgut1.filt, 1))+6
-am.coord <- layout.fruchterman.reingold(ig.mb)
-
-par(mfrow=c(1,3))
-plot(ig.mb, layout=am.coord, vertex.size=vsize, vertex.label=NA, main="MB")
-plot(ig.gl, layout=am.coord, vertex.size=vsize, vertex.label=NA, main="glasso")
-plot(ig.sparcc, layout=am.coord, vertex.size=vsize, vertex.label=NA, main="sparcc")
-
-```
-
-```{r}
-
-library(phyloseq)
-## Load round 2 of American gut project
-data('amgut2.filt.phy')
-se.mb.amgut2 <- spiec.easi(amgut2.filt.phy, method='mb', lambda.min.ratio=1e-2,
-                           nlambda=20, pulsar.params=list(rep.num=50))
-ig2.mb <- adj2igraph(getRefit(se.mb.amgut2),  vertex.attr=list(name=taxa_names(amgut2.filt.phy)))
-plot_network(ig2.mb, amgut2.filt.phy, type='taxa', color="Rank3")
-
-ps_p <-
-  test %>% 
-  speedyseq::tax_glom("Family") %>% 
-  ps_filter(passage == "5") %>% 
-  ps_filter(treat == "p")
-
-se.mb.ps_p <- spiec.easi(ps_p, method='mb', lambda.min.ratio=1e-2,
-                           nlambda=200, pulsar.params=list(rep.num=50))
-ig2.mb.ps_p <- adj2igraph(getRefit(se.mb.ps_p),  vertex.attr=list(name=taxa_names(ps_p)))
-plot_network(ig2.mb.ps_p, ps_p, type='taxa', color = "Class")
-
-ps_n <-
-  test %>% 
-  speedyseq::tax_glom("Family") %>% 
-  ps_filter(passage == "5") %>% 
-  ps_filter(treat == "n")
-
-se.mb.ps_n <- spiec.easi(ps_n, method='mb', lambda.min.ratio=1e-2,
-                           nlambda=200, pulsar.params=list(rep.num=50))
-ig2.mb.ps_n <- adj2igraph(getRefit(se.mb.ps_n),  vertex.attr=list(name=taxa_names(ps_n)))
-plot_network(ig2.mb.ps_n, ps_n, type='taxa', color="Class")
-getOptMerge(se.mb.ps_p)
-getOptMerge(se.mb.ps_n)
-getStability(se.mb.ps_n)
-###
-
-se1.mb.amgut <- spiec.easi(amgut1.filt, method='mb', lambda.min.ratio=1e-2,
-                          nlambda=20, pulsar.params=list(rep.num=20, ncores=4))
-se2.mb.amgut <- spiec.easi(amgut2.filt.phy, method='mb', lambda.min.ratio=1e-2,
-                          nlambda=20, pulsar.params=list(rep.num=20, ncores=4))
-
-se1.slr.amgut <- spiec.easi(amgut1.filt, method='slr', r=10, lambda.min.ratio=1e-2,
-                          nlambda=20, pulsar.params=list(rep.num=20, ncores=4))
-se2.slr.amgut <- spiec.easi(amgut2.filt.phy, method='slr', r=10, lambda.min.ratio=1e-2,
-                          nlambda=20, pulsar.params=list(rep.num=20, ncores=4))
-
-otu1 <- colnames(amgut1.filt)
-otu2 <- taxa_names(amgut2.filt.phy)
-edge.diss(getRefit(se1.mb.amgut), getRefit(se2.mb.amgut), 'jaccard', otu1, otu2)
-edge.diss(getRefit(se1.slr.amgut), getRefit(se2.slr.amgut), 'jaccard', otu1, otu2)
-
-X <- se2.slr.amgut$est$data
-L <- se2.slr.amgut$est$resid[[getOptInd(se2.slr.amgut)]]
-pca <- robustPCA(X, L)
-age <- as.numeric(as.character(sample_data(amgut2.filt.phy)$AGE))
-bmi <- as.numeric(as.character(sample_data(amgut2.filt.phy)$BMI))
-depth <- colSums(otu_table(amgut2.filt.phy))
-
-cor(age, pca$scores, use='pairwise')
-cor(bmi, pca$scores, use='pairwise')
-cor(depth, pca$scores, use='pairwise')
-
-```
-
-## Differential Abundance Methane
-
 ```{r corncob_ch4}
 
-ps_gen_pas5_trep <- ps_filter(ps_gen_pas5, treat == "p")
+ps_fam_pas5_trep <- ps_filter(ps_fam_pas5, treat == "p")
 
-ps_gen_prev_trep_prev <- tax_filter(ps_gen_pas5_trep, min_prevalence = 0.10)
+ps_fam_prev_trep_prev <- tax_filter(ps_fam_pas5_trep, min_prevalence = 0.10)
 
@@
                                  test = "Wald", boot = FALSE,
-                                 data = ps_gen_prev_trep_prev,
+                                 data = ps_fam_prev_trep_prev,
                                  fdr_cutoff = 0.05)
-# otu_to_taxonomy(OTU = da_analysis$significant_taxa, data = ps_gen_prev)
+otu_to_taxonomy(OTU = da_analysis$significant_taxa, data = ps_fam_prev_trep_prev)
 
-plot(da_analysis, level = c("Family", "Genus"))
+plot(da_analysis, level = c("Family"))
 
@@
 ```
-
-## Alpha Diversity Methane
-
-```{r diversity_ch4, eval = FALSE}
-ps_phylum <- speedyseq::tax_glom(ps, taxrank="Phylum")
-
-divnet_phylum_ch4 <- ps_phylum %>%
-  divnet(ncores = 4, tuning = "careful", formula = ~ rank_ch4)
-
-divnet_phylum_ch4$shannon %>% 
-  plot(ps_phylum, color = "rank_ch4") +
-  xlab("Rank Methane") +
-  ylab("Shannon diversity estimate\n(phylum level)") #+
-  #coord_cartesian(ylim = c(0,2))
-
-estimates <- divnet_phylum_ch4$shannon %>% summary %$% estimate
-ses <- sqrt(divnet_phylum_ch4$`shannon-variance`)
-X <- breakaway::make_design_matrix(ps_phylum, "rank_ch4")
-betta(estimates, ses, X)$table
-
-betta(estimates, ses, X)$global[2]
-
-```
-
-## Alpha Diversity Treatment in Passage 2
-
-```{r, results=FALSE, eval = FALSE}
-ps_class <- speedyseq::tax_glom(ps, taxrank="Class")
-ps_class
-
-ps_c_2 <- ps_filter(ps_class, passage == 2)
-
-divnet_c_2 <- ps_c_2 %>%
-  divnet(ncores = 4, tuning = "careful", formula = ~ treat)
-```
-
-```{r, eval = FALSE}
-divnet_c_2$shannon %>% 
-  plot(ps_c_2, color = "treat") +
-  xlab("Selection treatment") +
-  ylab("Shannon diversity estimate\n(class level)") +
-  labs(title="Test alpha diversity between P and N treatments within Passage 2")
-
-estimates <- divnet_c_2$shannon %>% summary %$% estimate
-ses <- sqrt(divnet_c_2$`shannon-variance`)
-X <- breakaway::make_design_matrix(ps_c_2, "treat")
-betta(estimates, ses, X)$table
-betta(estimates, ses, X)$global
-
-divnet_c_2$simpson %>% 
-  plot(ps_c_2, color = "treat") +
-  xlab("Selection treatment") +
-  ylab("Simpson diversity estimate\n(class level)") +
-  labs(title="Test alpha diversity between P and N treatments within Passage 2")
-
-estimates <- divnet_c_2$simpson %>% summary %$% estimate
-ses <- sqrt(divnet_c_2$`simpson-variance`)
-X <- breakaway::make_design_matrix(ps_c_2, "treat")
-betta(estimates, ses, X)$table
-betta(estimates, ses, X)$global
-
-```
-
-## Alpha Diversity Passage 2 vs 5 (all treatment)
-
-```{r, results=FALSE, eval=FALSE}
-ps_phylum <- speedyseq::tax_glom(ps, taxrank="Phylum")
-
-divnet_p <- ps_phylum %>%
-  divnet(ncores = 4, tuning = "careful", formula = ~ passage)
-```
-
-```{r, eval = FALSE}
-divnet_p$shannon %>% 
-  plot(ps_phylum, color = "passage") +
-  xlab("Selection treatment") +
-  ylab("Shannon diversity estimate\n(phylum level)") +
-  labs(title="Test alpha diversity between Passage 2 and 5 all treatments")
-
-estimates <- divnet_p$shannon %>% summary %$% estimate
-ses <- sqrt(divnet_p$`shannon-variance`)
-X <- breakaway::make_design_matrix(ps_phylum, "passage")
-betta(estimates, ses, X)$table
-betta(estimates, ses, X)$global[2]
-
-divnet_p$simpson %>% 
-  plot(ps_phylum, color = "passage") +
-  xlab("Selection treatment") +
-  ylab("Simpson diversity estimate\n(phylum level)") +
-  labs(title="Test alpha diversity between Passage 2 and 5 all treatments")
-
-estimates <- divnet_p$simpson %>% summary %$% estimate
-ses <- sqrt(divnet_p$`simpson-variance`)
-X <- breakaway::make_design_matrix(ps_phylum, "passage")
-betta(estimates, ses, X)$table
-betta(estimates, ses, X)$global[2]
-
-```
-
-## Alpha Diversity Treatment in Passage 5
-
-```{r, results=FALSE, eval = FALSE}
-
-ps_class <- speedyseq::tax_glom(ps, taxrank="Class")
-
-ps_c_5 <- ps_filter(ps_class, passage == 5)
-
-ps_5 <- ps_filter(ps_phylum, passage == 5)
-
-divnet_p_5 <- ps_5 %>%
-  divnet(ncores = 4, tuning = "careful", formula = ~ treat)
-```
-
-```{r, eval = FALSE}
-divnet_p_5$shannon %>% 
-  plot(ps_5, color = "treat") +
-  xlab("Selection treatment") +
-  ylab("Shannon diversity estimate\n(phylum level)") +
-  labs(title="Test alpha diversity between P and N treatments within Passage 5")
-
-estimates <- divnet_p_5$shannon %>% summary %$% estimate
-ses <- sqrt(divnet_p_5$`shannon-variance`)
-X <- breakaway::make_design_matrix(ps_5, "treat")
-betta(estimates, ses, X)$table
-betta(estimates, ses, X)$global
-#betta_pic(estimates, ses)
-
-divnet_p_5$simpson %>% 
-  plot(ps_5, color = "treat") +
-  xlab("Selection treatment") +
-  ylab("Simpson diversity estimate\n(phylum level)") +
-  labs(title="Test alpha diversity between P and N treatments within Passage 5")
-
-estimates <- divnet_p_5$simpson %>% summary %$% estimate
-ses <- sqrt(divnet_p_5$`simpson-variance`)
-X <- breakaway::make_design_matrix(ps_5, "treat")
-betta(estimates, ses, X)$table
-betta(estimates, ses, X)$global
-#betta_pic(estimates, ses)
-
-```
-
-## Beta Diversity passage + treatment
-
-```{r, eval=FALSE}
-
-ps_f_c <- speedyseq::tax_glom(ps_fixed, "Class")
-
-X <- diag(nrow(sample_data(ps_f_c)))
-colnames(X) <- rownames(X) <-  as.character(1:ncol(X))
-
-divnet_class <-  divnet(W = ps_f_c,
-                         X = X,
-                         variance = "none")
-
-# Median composition
-set.seed(454)
-bc_test <- testBetaDiversity(divnet_class, 
-                             h0 = "bray-curtis",
-                             n_boot = 1000, # set this to 10000
-                             sample_specimen_matrix = divnet_class$X,
-                             groups = sample_data(ps_f_c)$treat_pas)
-bc_test$p_value
-
-# Euclidean distance
-set.seed(3532)
-euc_test <- testBetaDiversity(dv = divnet_class , 
-                              h0 = "euclidean",
-                              n_boot = 1000,
-                              sample_specimen_matrix = divnet_class$X,
-                              groups = sample_data(ps_f_c)$treat_pas)
-euc_test$p_value
-
-# Aitchison Distance
-set.seed(3423)
-ait_test <- testBetaDiversity(dv = divnet_class, 
-                              sample_specimen_matrix = divnet_class$X,
-                              h0 = "aitchison",
-                              n_boot = 1000,
-                    groups = sample_data(ps_f_c)$treat_pas)
-
-ait_test$p_value
-```
-
-## Beta Diversity gen 2 treatment
-
-```{r, results=FALSE, eval = FALSE}
-ps_f_2 <- ps_filter(ps_fixed, passage == "2")
-ps_f_2_c <- speedyseq::tax_glom(ps_f_2, "Class")
-X <- diag(nrow(sample_data(ps_f_2_c)))
-colnames(X) <- rownames(X) <-  as.character(1:ncol(X))
-
-divnet_c_2 <-  divnet(W = ps_f_2_c,
-                         X = X,
-                         variance = "none")
-
-```
-
-```{r, eval = FALSE}
-
-# Median composition
-set.seed(454)
-bc_test <- testBetaDiversity(dv = divnet_c_2, 
-                             h0 = "bray-curtis",
-                             n_boot = 1000, # set this to 10000
-                             sample_specimen_matrix = divnet_c_2$X,
-                             groups = sample_data(ps_f_2_c)$treat_bin)
-bc_test$p_value
-set.seed(3532)
-# Euclidean distance
-euc_test <- testBetaDiversity(dv = divnet_c_2 , 
-                              h0 = "euclidean",
-                              n_boot = 1000,
-                              sample_specimen_matrix = divnet_c_2$X,
-                              groups = sample_data(ps_f_2_c)$treat_bin)
-euc_test$p_value
-
-# Aitchison Distance
-set.seed(3423)
-ait_test <- testBetaDiversity(dv = divnet_c_2, 
-                              sample_specimen_matrix = divnet_c_2$X,
-                              h0 = "aitchison",
-                              n_boot = 1000,
-                              groups = sample_data(ps_f_2_c)$treat_bin)
-
-ait_test$p_value
-
-```
-
-## Beta Diversity gen 5 treatment
-
-```{r, eval = FALSE}
-ps_f_5 <- ps_filter(ps_fixed, passage == "5")
-ps_f_5_c <- speedyseq::tax_glom(ps_f_5, "Class")
-X <- diag(nrow(sample_data(ps_f_5_c)))
-colnames(X) <- rownames(X) <-  as.character(1:ncol(X))
-
-divnet_c_5 <-  divnet(W = ps_f_5_c,
-                         X = X,
-                         variance = "none")
-
-# Median composition
-set.seed(454)
-bc_test <- testBetaDiversity(dv = divnet_c_5, 
-                             h0 = "bray-curtis",
-                             n_boot = 1000, # set this to 10000
-                             sample_specimen_matrix = divnet_c_5$X,
-                             groups = sample_data(ps_f_5_c)$treat_bin)
-bc_test$p_value
-set.seed(3532)
-# Euclidean distance
-euc_test <- testBetaDiversity(dv = divnet_c_5 , 
-                              h0 = "euclidean",
-                              n_boot = 1000,
-                              sample_specimen_matrix = divnet_c_5$X,
-                              groups = sample_data(ps_f_5_c)$treat_bin)
-euc_test$p_value
-
-# Aitchison Distance
-set.seed(3423)
-ait_test <- testBetaDiversity(dv = divnet_c_5, 
-                              sample_specimen_matrix = divnet_c_5$X,
-                              h0 = "aitchison",
-                              n_boot = 1000,
-                              groups = sample_data(ps_f_5_c)$treat_bin)
-
-ait_test$p_value
-
-```

### 2022-06-24 R/07-process-community.Rmd
add seqs to output community data file

@@
 # They match if = 0 
-
+seqs <- tibble(asv = paste0("ASV", seq(ncol(seqtab.nochim))),
+               seq = colnames(seqtab.nochim))
 colnames(seqtab.nochim) <- paste0("ASV", seq(ncol(seqtab.nochim)))
@@
 
-# Sequence Depth Distribution + Sample Composition
+# Sequence Depth Distribution
 
@@
 ```
+
+# Beta Diversity all samples
 
@@
 
+full_data <-
+  full_data %>% 
+  left_join(seqs, by = "asv")
+
 write_tsv(full_data, 'Output/community_data.tsv')

### 2022-06-24 R/functions.R
add quartile function and log methane function

@@ -4,9 +4,9 @@ der_dir <- '../Output/' # Derived/Modified Data Directory
 
 # theme_set(theme_tufte()+
 #             theme(axis.title.y = element_text(angle = 0, vjust = 0.5)))
-theme_set(theme_classic() +
-            theme(panel.border = element_rect(fill = NA, size = 1.0625),
-                  axis.line = element_blank()))
+# theme_set(theme_classic() +
+#             theme(panel.border = element_rect(fill = NA, size = 1.0625),
+#                   axis.line = element_blank()))
 
 plot_flux <- function(fluxes, passage, estimate, ratio = NULL, log10 = FALSE) {
   p <- ggplot(fluxes, aes(x = {{passage}}, y = {{estimate}}, color = treat)) +
@@ -212,3 +212,20 @@ simpson <- function(x){
   1 - sum((x/n)^2)
 }
 
+
+log_methane <- function(x) {
+  log10( x + abs(min(x, na.rm = TRUE)) + 1)
+}
+
+
+quart <- function(x) {
+  x <- sort(x)
+  n <- length(x)
+  m <- (n+1)/2
+  if (floor(m) != m) {
+    l <- m-1/2; u <- m+1/2
+  } else {
+    l <- m-1; u <- m+1
+  }
+  c(Q1=median(x[1:l]), Q3=median(x[u:n]))
+}

### 2022-07-05 Manuscript/manuscript.Rmd
combine all edits to incorporate the figures and analyses into the manuscript

@@
 load(file = '../Output/response_model.RData')
+load(file = '../Output/response_plot.RData')
 load(file = '../Output/herit_model.RData')
+load(file = '../Output/herit_plot.RData')
 
@@
 load(file = '../Output/beta_model.RData')
+load(file = '../Output/beta_fig.RData')
 load(file = '../Output/da_corncob.RData')
+load(file = '../Output/da_fig.RData')
 
@@
 
+## Response to selection on methane oxidation rate
+
 We observed a response to selection on whole-ecosystem soil CH~4~ oxidation
-rate (`r figures("response", display = "cite")`). In addition, the
-response to selection varied with treatment 
-`r print_lrt(compare_response_mods, 2, 'LRT of nested models with and without treatment')`. At the start of the experiment, the Positive
+rate (`r figures("response", display = "cite")`; `r print_lrt(compare_response_mods, 2, 'LRT of nested models with and without treatment')`. At the start of the experiment, the Positive
 treatment had a mean CH~4~ oxidation rate that was
@@
 
+## Heritability of methane oxidation
+
 We estimated heritability as the regression of mid-offspring on mid-parent (`r figures("herit", display = "cite")`).
@@
 
-Sequencing of 16S rRNA amplicons revealed that there were `r ntaxa(ps)` unique
-ASVs across all `r nsamples(ps)` jars sampled.
-Richness of ASVs decreased among all of the jars between Passage 2 and 5
-(Richness difference between Passage 5 and Passage 2 = `r myround(bt_passage$table[2, 1], 1)`, SE =
-`r myround(bt_passage$table[2, 2], 2)`, p < 0.001). However, there was no
+## Taxonomic richness
+
+Richness of ASVs decreased between Passage 2 (median = `r myround(rare_median[1, 2], 1)`, IQR = `r myround(rare_median[1, 3], 1)`) and 5 (median = `r myround(rare_median[2, 2], 1)`, IQR = `r myround(rare_median[2, 3], 1)`; Kruskal-Wallace test: chi-squared =
+`r myround(kt[1,1], 1)`, df = `r kt[1,3]`, p < 0.001; Pairwise Wilcoxon test: p < 0.001). However, there was no
 difference in richness between the Positive and the Neutral treatment in
-Passage 5 (Difference = `r myround(bt_treat5$table[2, 1], 1)`, SE =
-`r myround(bt_treat5$table[2, 2], 2)`, p =
-`r myround(bt_treat5$table[2, 3], 3)`).  In addition, there was no
+Passage 2 or 5 (Pairwise Wilcoxon test: p = `r myround(wt[1, 3], 1)`).  In addition, there was no
 correlation between richness and CH~4~ oxidation rate across all passages
-and treatments.  (slope = `r myround(bt_ch4$table[2, 1], 1)`, SE =
-`r myround(bt_ch4$table[2, 2], 2)`, p = `r myround(bt_ch4$table[2, 3], 3)`).
+and treatments (Spearman's rho = `r myround(s[1,1], 1)`, p = `r myround(s[1,3], 1)`).
 
-Taxonomic dissimilarity of the soil microbiome varied strongly by passage and
-weakly by treatment (`r figures("beta", display = "cite")`). Permutational
-analysis of variance demonstrated an effect of passage, treatment,
-and their interaction on Aitchison dissimilarity (`r tables("adonis", display = "cite")`). 
-Passage explained `r myround(100*fit_adonis[1, 4])`\% of the
-variation in Aitchison dissimilarity and `r myround(100*fit_adonis[6, 4])`\% in
-Jaccard dissimilarity.  Treatment explains `r myround(100*fit_adonis[2, 4])`\%
-for Aitchison and `r myround(100*fit_adonis[7, 4])`\% for Jaccard.  Variation
-in microbiome composition decreased as a result of passaging in jars (F = 80.5,
-p < 0.001), but did not differ between selection treatments (F = 0.40, p =
-0.54).
+## Community dissimilarity
+
+Bray-Curtis dissimilarity of the soil microbiome varied strongly by passage and
+weakly by treatment and the interaction between passage and treatment (`r figures("beta", display = "cite")`; `r tables("tidy_permanova", display = "cite")`). 
+Passage explained `r myround(100*tidy_permanova[1, 4])`\% of the
+variation in Bray-Curtis dissimilarity.  Treatment explained 
+`r myround(100*tidy_permanova[2, 4])`\%
+of the variation. There was also an interaction
+between treatment and passage, which explained 
+`r myround(100*tidy_permanova[3, 4])`\% of the variation. We found there was 
+no difference in dispersion between treatments or passages (F~3,44~ = 0.91, p = 0.45).
+
+## Taxa that responded to selection
 
@@
 abundance of ~2 reads. The other prevalent family was a member of the
-Silvanigrellaceae, a newly described family placed in its own order. This family
+*Silvanigrellaceae*, a newly described family placed in its own order. This family
 was present in all 12 samples and had a median abundance of 8 reads.
-Silvanigrella is the only cultivated member of
-Silvanigrellaceae and was isolated from a temperate fresh water lake
+*Silvanigrella* is the only cultivated member of
+*Silvanigrellaceae* and was isolated from a temperate fresh water lake
 [@hahn2017]. Of the families unique to the Neutral treatment, only one had a
-prevalence greater than 2/12. This family, Armatimonadaceae, was present in half
-of the Neutral samples (prevalence = 6/12). The type strain for Armatimonadaceae was isolated from
+prevalence greater than 2/12. This family, *Armatimonadaceae*, was present in half
+of the Neutral samples (prevalence = 6/12). The type strain for *Armatimonadaceae* was isolated from
 the rhizosphere of *Phragmites australis* [@tamaki2011].
@@
 abundance of families in the Positive jars relative to the Neutral jars within
-Passage 5 using a beta-binomial model. We identified `r ntaxa(sig_taxa_treat)`
-families that were enriched or depleted in the Positive treatmeant relative to
+Passage 5 using three methods and then plotted the taxa identified by all three methods. We identified `r length(da_fams)`
+families that were enriched or depleted in the Positive treatment relative to
 the Neutral treatment (`r figures("da", display = "cite")`).
@@
 This class is one of the groups containing methanotrophic families such as
-Methylococcaceae. However, gammaproteobacteria is among the most diverse 
+*Methylococcaceae.* However, Gammaproteobacteria is among the most diverse 
 groups in the Prokaryotes, so this is not strong evidence for a selection response
@@
 The next family
-was Puniceicoccaceae, which is a member of the phylum Verrucomicrobia and
+was *Puniceicoccaceae*, which is a member of the phylum Verrucomicrobia and
 contains marine microbes. The Verrucomicrobia is a diverse group that contain
@@
 An ASV from the order Armatimonadales was also enriched in the Positive treatment.
-This order also contains the family Armatimonadaceae, which was
-found to be unique to the Neutral treatment. Cytophagaceae was also enriched in
+This order also contains the family *Armatimonadaceae*, which was
+found to be unique to the Neutral treatment. *Cytophagaceae* was also enriched in
 the Positive treatment and contains a number of mainly aerobic heterotrophs
 that can digest a variety of macromolecules [@mcbride2014]. Notably, the family
-Gemmatimonadaceae has a single bacterial bin, Candidatus '*Methylotropicum kingii*',
+*Gemmatimonadaceae* has a single bacterial bin, Candidatus '*Methylotropicum kingii*',
 that contains methanotrophy genes, but this family was depleted in the Positive
@@
 ```{r response, fig.width=4, fig.height=3}
-ratio_log10 <- calc_plot_ratios(fluxes$passage, fluxes$log_ch4)
 
-p <- plot_flux(fluxes, passage, log_ch4, ratio_log10, log10 = TRUE)
-p
-ggsave('fig1.jpg', p)
+p_response
 
@@
 
-p <- plot_herit(heritability)
-p
-ggsave('fig2.jpg', p)
+p_herit
 
@@
 
-p <- plot_beta(ps)
-p
-ggsave('fig3.jpg', p)
+p_beta
 
@@
 
-p <- dt_plot(dt_data)
-p
-ggsave('fig4.jpg', p)
+p_log10
 
@@
 
-`r tables("adonis", "PERMANOVA on Aitchison and Jaccard dissimilarities.")`
+`r tables("adonis", "PERMANOVA on Bray-Curtis dissimilarities.")`
 
@@
 
-write.table(fit_adonis, 'table_1.tsv')
-fit_adonis %>%
-  mutate(metric = c(rep("aitchison", 5), rep("jaccard", 5))) %>% 
-  select(metric, term:p.value) %>% 
+tidy_permanova %>%
 	flextable()
-# kable(col.names = c('Term', 'df', 'SS', 'R2', 'F', 'p'), 
-#       digits = c(2,2,2,2,2,3),
-## 	format = "Table",
-## 	booktabs = T,
-# 	caption = "PERMANOVA on Bray-Curtis and Jaccard dissimilarities.",
-#   	label = "beta") %>%
-#   pack_rows("Bray-Curtis", 1, 5) %>%
-#   pack_rows("Jaccard", 6, 10)
 

### 2022-07-05 R/04-heritability.Rmd
combine all edits to incorporate the figures and analyses into the manuscript

@@
 ---
+
+```{r setup, include=FALSE}
+
+knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = TRUE)
+knitr::opts_knit$set(root.dir=normalizePath('../'))
+
+```
 
@@
 ```{r initialize_variables}
-source('functions.R')
+source('R/functions.R')
 ```
@@
 
-plot_herit(heritability)
+p_herit <- plot_herit(heritability)
+
+p_herit
+
+save(p_herit, file = 'Output/herit_plot.RData')
+
+ggsave('Manuscript/fig2.jpg', p_herit, width = 4, height = 3)
 
@@
 ```{r save_heritability}
-save(herit_model, file = '../Output/herit_model.RData')
+save(herit_model, file = 'Output/herit_model.RData')
 ```

### 2022-07-05 R/09-community-analysis.Rmd
combine all edits to incorporate the figures and analyses into the manuscript

@@
 author: "Andrew Morris"
-date: "9/15/2021"
-output: html_document
+date: "6/28/2022"
+output: 
+  html_document:
+    toc: true
 editor_options: 
-  chunk_output_type: inline
+  chunk_output_type: console
 ---
@@
 ```{r setup, include=FALSE}
+
 knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = TRUE)
 knitr::opts_knit$set(root.dir=normalizePath('../'))
+set.seed(1337)
 
@@
 ```{r initialize_packages}
-library(tidyverse)
+
 library(phyloseq)
-library(decontam)
-library(microViz)
 library(knitr)
-library(corncob)
-library(DivNet)
 library(magrittr)
@@
 library(broom)
+library(ggtext)
+library(ALDEx2)
+library(corncob)
+library(ANCOMBC)
+library(tidyverse)
 
@@
 ```{r import-data}
-load(file = "Output/phyloseq.RData")
-ps <-
-  ps %>% 
-  ps_mutate(pas_by_treat = paste0(passage, treat))
-ps_melt <-
-  ps %>% 
-  ps_melt()
+
+aes_data <- read_tsv('Output/community_data.tsv')
+
+```
+
+```{r fix-treatment}
+
+aes_data <-
+  aes_data %>% 
+  mutate(passage_treat = treat) %>% 
+  separate(passage_treat, c('passage', "treat"), sep = 2, remove = FALSE) %>% 
+  select(sample, passage_treat, passage, treat, selected, ch4, asv, count, Kingdom:Species, seq)
+
+```
+
+```{r fix-taxonomy}
+
+# There are "Unknown Family" taxa with Genus and Class, but Order is Incertae Sedis.
+# I'm just going to remove everything below Class and call it "Unclassified"
+
+fill_taxonomy <- function(tab) {
+    if(is.na(tab[[1]])) tab[[1]] = "Unclassified"
+    for(i in 2:length(tab)) {
+      if(is.na(tab[[i]])){
+        if(grepl("unclassified", tab[[i-1]])) tab[[i]] = tab[[i-1]]
+        else tab[[i]] = paste(tab[[i-1]], "unclassified", sep = "_")
+      }
+    }
+    tab
+}
+
+aes_data <-
+  aes_data %>% 
+  nest(data = -c(asv, Kingdom:Species)) %>% 
+  nest(taxa = Kingdom:Species) %>% 
+  mutate(taxa = map(taxa, ~fill_taxonomy(.x))) %>% 
+  unnest(taxa) %>% 
+  unnest(data)
+
 ```
@@
 
-```{r plot-depth}
+```{r n-seqs}
 
-depth_data <-
-  ps %>% 
-  ps_melt %>% 
-  group_by(Sample, passage, treat) %>% 
-  summarize(depth = sum(Abundance))
+aes_data %>% 
+  group_by(sample) %>% 
+  summarize(total = sum(count)) %>% 
+  ggplot(aes(x = total)) +
+  geom_histogram() +
+  theme_classic() +
+  labs(title = "Distribution sequencing depths",
+       x = "Library size",
+       y = "Frequency")
 
-summary(aov(depth ~ passage * treat, data = depth_data))
-depth_data %>% 
-  ggplot(aes(x = interaction(treat, passage), y = depth)) +
-  theme_tufte() +
-  geom_rangeframe() +
-  # coord_cartesian(clip="off") +
-  geom_tufteboxplot()
+seq_range <-
+  aes_data %>% 
+  group_by(sample) %>% 
+  summarize(total = sum(count)) %>% 
+  summarize(seq_range = range(total)) %$% 
+  seq_range
+
+a_analysis <- 
+  aes_data %>%
+  count(sample, count) %>%
+  nest(data = -sample) %>%
+  mutate(n_seqs = map_dbl(data, ~sum(.x$count * .x$n)),
+         sobs = map_dbl(data, ~sum(.x$n)),
+         rare = map_dbl(data, ~rarefy(rep(.x$count, .x$n), sample=seq_range[1]))) %>%
+  select(-data)
+
+a_analysis %>%
+  select(sample, n_seqs, sobs, rare) %>%
+  pivot_longer(-c(sample, n_seqs)) %>%
+  ggplot(aes(x=n_seqs, y=value, color=name)) +
+  geom_point() +
+  geom_smooth(se=FALSE) +
+  coord_cartesian(ylim=c(0, 6000)) +
+  theme_classic() +
+  labs(x = "Sequencing Depth", y = "Alpha Diversity")
+
+a_analysis %>%
+  select(sample, n_seqs, sobs, rare) %>%
+  mutate(treat = substr(sample, 1, 3)) %>% 
+  select(-sample) %>% 
+  group_by(treat) %>% 
+  pivot_longer(-c(treat)) %>%
+  ggplot(aes(x=treat, y=value)) +
+  facet_wrap(~ name, nrow = 3, scales = "free") +
+  geom_boxplot() +
+  geom_jitter() +
+  theme_classic() +
+  labs(title = "AES", x = "Treatment", y = "Value")
 
@@
 
-There is no systematic difference in sequencing depth/library size between treatments. It does appear that sequencing depth decreases between passage 2 and 5.
+This is the number of reads (n_seqs), the rarefied richness (rare), and 
+the observed richness (sobs for "species observed") for the four treatments:
+generations 2 and 5 and selection treatments neutral and positive. As you can
+see, there were more sequences from G2 than G5 and also greater richness
+in G2 than G5. However, that richness difference is robust to rarefying
+to the level of the shallowest sample. There doesn't seem to be a systemic
+difference in sequencing depth or richness between the neutral and positive 
+treatments in either generation.
 
-```{r plot-richness}
+## Richness
 
-plot_richness(ps, x = "pas_by_treat", color="treat", measures=c("Observed", "Shannon", "Simpson"), title = "Alpha Diversity ASV") +
+### Plot
+
+```{r alpha-diversity-plot}
+
+# aes_data %>% 
+#   group_by(sample) %>% 
+#   summarize(n_seqs = sum(count)) %>% 
+#   summarize(min = min(n_seqs))
+
+aes_df <-
+  aes_data %>% 
+  select(sample, asv, count) %>%
+  pivot_wider(names_from = "asv", values_from = "count", values_fill = 0) %>% 
+  as.data.frame()
+
+rownames(aes_df) <- aes_df$sample
+aes_df <- aes_df[, -1]
+
+rare <- 
+  rarefy(aes_df, seq_range[1]) %>% 
+  as_tibble(rownames = "sample") %>% 
+  rename(rare = value)
+
+aes_data %>% 
+  inner_join(rare, by = 'sample') %>% 
+  select(sample, passage_treat, rare) %>% 
+  distinct %>% 
+  ggplot(aes(x = passage_treat, y = rare)) +
+  geom_boxplot() +
+  geom_jitter() +
+  labs(x = "Passage x Treatment", y = "Rarefied Richness") +
+  theme_bw() +
+  theme(
+    plot.background = element_blank(),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank())
+
+```
+
+### Hypothesis Test
+
+```{r alpha-diversity-test}
+
+rare_data <-
+  aes_data %>% 
+  select(sample, passage_treat, passage, treat, ch4) %>% 
+  distinct %>% 
+  inner_join(rare, by = "sample")
+
+kt <- kruskal.test(rare ~ passage_treat, data = rare_data)
+kt <- tidy(kt)
+
+wt <- pairwise.wilcox.test(rare_data$rare, 
+                           rare_data$passage_treat, 
+                           p.adjust.method = "BH")
+wt <- tidy(wt)
+
+rare_median <-
+  rare_data %>% 
+  group_by(passage) %>% 
+  summarize(median = median(rare),
+            IQR = quart(rare)[2] - quart(rare)[1])
+
+```
+
+There is an effect of passage x treatment on richness (Kruskal-Wallace test: chi-squared = 
+`r kt[1,1]`, df = `r kt[1,3]`, p = `r kt[1, 2]`). The pairwise test shows that
+G2 is different from G5, but N and P are not different in either generation
+(Wilcoxon/Mann-Whitney Test).
+
+```{r}
+
+kable(wt)
+
+```
+
+```{r alpha-diversity-ch4-test}
+
+s <- cor.test(~ ch4 + rare, data = rare_data, method = "spearman", alternative = "two.sided") %>% 
+  tidy
+
+s5 <-
+  rare_data %>% 
+  filter(passage == "G5") %>% 
+  cor.test(~ ch4 + rare, data = ., method = "spearman", alternative = "two.sided")
+save(kt, wt, rare_median, s, file = "Output/richness_models.RData")
+
+```
+
+A Spearman's rank correlation test shows no correlation between
+methane oxidation rate and richness (Spearman's rho = `r s[1,1]`, p = `r s[1,3]`).
+
+# Beta Diversity
+
+```{r bray-curtis-nmds, messages=FALSE}
+
+dist <- vegdist(aes_df, method="bray")
+
+if(file.exists("Output/rare_dist_bray.rds")) {
+  rare_dist_bray <- 
+    readRDS("Output/rare_dist_bray.rds")
+} else {
+  rare_dist_bray <- avgdist(aes_df, dmethod="bray", sample=seq_range[1])
+  rare_dist_bray %>% 
+    saveRDS("Output/rare_dist_bray.rds")
+}
+
+nmds <- metaMDS(rare_dist_bray)
+
+nmds_scores <-
+  scores(nmds, display = "sites") %>% 
+  as_tibble(rownames = "sample")
+
+nmds_data <-
+  aes_data %>% 
+  select(sample, passage, treat) %>% 
+  distinct %>% 
+  inner_join(nmds_scores, by = "sample")
+
+centroids <-
+  nmds_data %>% 
+  group_by(passage, treat) %>% 
+  summarize(NMDS1 = mean(NMDS1),
+            NMDS2 = mean(NMDS2),
+            .groups = "drop")
+p_beta <-
+  nmds_data %>%   
+  ggplot(aes(x = NMDS1, y = NMDS2, shape = passage, color = treat)) +
+  geom_point() +
+  stat_ellipse(show.legend = FALSE) +
+  coord_fixed() +
+  scale_shape_discrete(name = "Passage") +
   scale_color_manual(name = "Treatment",
+                       labels = c('Neutral', 'Positive'),
+                       values = c('gray40', 'darkorange2')) #+
+  # theme_bw() +
+  # theme(
+  #   plot.background = element_blank(),
+  #   panel.grid.major = element_blank(),
+  #   panel.grid.minor = element_blank())
+
+save(p_beta, file = "Output/beta_fig.RData")
+
+ggsave('Manuscript/fig3.jpg', p_beta, height = 2.5, width = 4)
+
+```
+
+## Beta diversity test passage by treatment
+
+```{r permanova-passage-treatment, warnings=FALSE}
+
+dist_df <-
+  rare_dist_bray %>% 
+  as.matrix %>% 
+  as_tibble(rownames = "sample")
+
+meta_dist <-
+  aes_data %>% 
+  select(sample, passage_treat, passage, treat, ch4) %>% 
+  distinct %>% 
+  inner_join(dist_df, by = "sample")
+
+permanova <- adonis2(rare_dist_bray ~ passage * treat, data = meta_dist, permutations = 999)
+
+betadisper(rare_dist_bray, meta_dist$passage_treat) %>% 
+  permutest()
+
+tidy_permanova <- 
+  tidy(permanova)
+
+save(tidy_permanova, file = "Output/beta_model.RData")
+
+```
+
+There is a significant difference of centroid between passage, treatment, and
+their interaction. Testing a difference of dispersion using betadisper
+shows no difference in variance between the four groups.
+
+# Distance-based RDA Methane
+
+```{r full-dbrda}
+
+meta_dist %>% 
+  select(passage_treat, ch4) %>% 
+  mutate(log_ch4 = log_methane(ch4)) %>% 
+  ggplot(aes(x = passage_treat, y = log_ch4)) +
+  geom_boxplot() +
+  theme_classic() +
+  labs(x = "Passage x Treatment", y = "Log10 Methane")
+
+dbrda_ch4 <- dbrda(rare_dist_bray ~ rank(ch4), 
+                      data = meta_dist, 
+                      na.action = "na.omit",
+                      add = TRUE)
+
+dbrda_ch4 %>% 
+  permutest(permutations = 9999)
+
+percent_explained <- 100 * eigenvals(dbrda_ch4) / sum(eigenvals(dbrda_ch4))
+
+dbrda_scores <-
+  scores(dbrda_ch4, display = "sites") %>% 
+  as_tibble(rownames = "sample")
+
+dbrda_data <-
+  aes_data %>% 
+  select(sample, passage, treat, ch4) %>% 
+  distinct %>% 
+  inner_join(dbrda_scores, by = "sample")
+
+dbrda_data %>%   
+  ggplot(aes(x = dbRDA1, y = MDS1, shape = passage, color = treat)) +
+  geom_point() +
+  stat_ellipse(show.legend = FALSE) +
+  coord_fixed() +
+  scale_shape_discrete(name = "Passage") +
+  scale_color_manual(name = "Treatment",
+                       labels = c('Neutral', 'Positive'),
+                       values = c('gray40', 'darkorange2')) +
+  theme_bw() +
+  theme(
+    plot.background = element_blank(),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank())
+
+```
+
+```{r pre-dbrda-data-subsets}
+
+# aes_data %>% 
+#   filter(treat == "P") %>% 
+#   group_by(asv) %>% 
+#   mutate(total = sum(count)) %>% 
+#   filter(total != 0) %>% 
+#   ungroup() %>% 
+#   select(-total) %>% 
+#   group_by(sample) %>% 
+#   summarize(n_seqs = sum(count)) %>% 
+#   summarize(min = min(n_seqs))
+
+treatP_df <-
+  aes_data %>%  
+  filter(treat == "P") %>% 
+  group_by(asv) %>% 
+  mutate(total = sum(count)) %>% 
+  filter(total != 0) %>% 
+  ungroup() %>% 
+  select(-total) %>% 
+  select(sample, asv, count) %>%
+  pivot_wider(names_from = "asv", values_from = "count", values_fill = 0) %>% 
+  as.data.frame()
+
+rownames(treatP_df) <- treatP_df$sample
+treatP_df <- treatP_df[, -1]
+
+# aes_data %>% 
+#   filter(passage == "G5") %>% 
+#   group_by(asv) %>% 
+#   mutate(total = sum(count)) %>% 
+#   filter(total != 0) %>% 
+#   ungroup() %>% 
+#   select(-total) %>% 
+#   group_by(sample) %>% 
+#   summarize(n_seqs = sum(count)) %>% 
+#   summarize(min = min(n_seqs))
+
+passage5_df <-
+  aes_data %>%  
+  filter(passage == "G5") %>% 
+  group_by(asv) %>% 
+  mutate(total = sum(count)) %>% 
+  filter(total != 0) %>% 
+  ungroup() %>% 
+  select(-total) %>% 
+  select(sample, asv, count) %>%
+  pivot_wider(names_from = "asv", values_from = "count", values_fill = 0) %>% 
+  as.data.frame()
+
+rownames(passage5_df) <- passage5_df$sample
+passage5_df <- passage5_df[, -1]
+
+```
+
+```{r treat-dbrda}
+
+if(file.exists("Output/treat_rare_dist_bray.rds")) {
+  treat_rare_dist_bray <- 
+    readRDS("Output/treat_rare_dist_bray.rds")
+} else {
+treat_rare_dist_bray <- avgdist(treatP_df, dmethod="bray", sample=seq_range[1])
+  treat_rare_dist_bray %>% 
+    saveRDS("Output/treat_rare_dist_bray.rds")
+}
+
+treat_dist_df <-
+  treat_rare_dist_bray %>% 
+  as.matrix %>% 
+  as_tibble(rownames = "sample")
+
+treat_meta_dist <-
+  aes_data %>% 
+  select(sample, passage_treat, passage, treat, ch4) %>% 
+  distinct %>% 
+  inner_join(treat_dist_df, by = "sample")
+
+treat_dbrda_ch4 <- dbrda(treat_rare_dist_bray ~ rank(ch4), 
+                      data = treat_meta_dist, 
+                      na.action = "na.omit",
+                      add = TRUE)
+
+treat_dbrda_ch4 %>% 
+  permutest(permutations = 9999)
+
+percent_explained <- 100 * eigenvals(treat_dbrda_ch4) / sum(eigenvals(treat_dbrda_ch4))
+
+treat_dbrda_scores <-
+  scores(treat_dbrda_ch4, display = "sites") %>% 
+  as_tibble(rownames = "sample")
+
+treat_dbrda_data <-
+  aes_data %>% 
+  select(sample, passage, treat, ch4) %>% 
+  distinct %>% 
+  inner_join(treat_dbrda_scores, by = "sample")
+
+treat_dbrda_data %>%   
+  ggplot(aes(x = dbRDA1, y = MDS1, shape = passage, color = treat)) +
+  geom_point() +
+  stat_ellipse(show.legend = FALSE) +
+  coord_fixed() +
+  scale_shape_discrete(name = "Passage") +
+  scale_color_manual(name = "Treatment",
+                       labels = c('Positive'),
+                       values = c('darkorange2')) +
+  theme_bw() +
+  theme(
+    plot.background = element_blank(),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank())
+
+```
+
+```{r passage-dbrda}
+if(file.exists("Output/passage_rare_dist_bray.rds")) {
+  passage_rare_dist_bray <- 
+    readRDS("Output/passage_rare_dist_bray.rds")
+} else {
+passage_rare_dist_bray <- avgdist(passage5_df, dmethod="bray", sample=seq_range[1])
+  passage_rare_dist_bray %>% 
+    saveRDS("Output/passage_rare_dist_bray.rds")
+}
+
+passage_dist_df <-
+  passage_rare_dist_bray %>% 
+  as.matrix %>% 
+  as_tibble(rownames = "sample")
+
+passage_meta_dist <-
+  aes_data %>% 
+  select(sample, passage_treat, passage, treat, ch4) %>% 
+  distinct %>% 
+  inner_join(passage_dist_df, by = "sample")
+
+passage_dbrda_ch4 <- dbrda(passage_rare_dist_bray ~ rank(ch4), 
+                      data = passage_meta_dist, 
+                      na.action = "na.omit",
+                      add = TRUE)
+
+passage_dbrda_ch4 %>% 
+  permutest(permutations = 9999)
+
+passage_percent_explained <- 100 * eigenvals(passage_dbrda_ch4) / sum(eigenvals(passage_dbrda_ch4))
+
+passage_dbrda_scores <-
+  scores(passage_dbrda_ch4, display = "sites") %>% 
+  as_tibble(rownames = "sample")
+
+passage_dbrda_data <-
+  aes_data %>% 
+  select(sample, passage, treat, ch4) %>% 
+  distinct %>% 
+  inner_join(passage_dbrda_scores, by = "sample")
+
+passage_dbrda_data %>%   
+  ggplot(aes(x = dbRDA1, y = MDS1, shape = passage, color = treat)) +
+  geom_point() +
+  stat_ellipse(show.legend = FALSE) +
+  coord_fixed() +
+  scale_shape_discrete(name = "Passage") +
+  scale_color_manual(name = "Treatment",
+                       labels = c('Neutral', 'Positive'),
+                       values = c('gray40', 'darkorange2')) +
+  theme_bw() +
+  theme(
+    plot.background = element_blank(),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank())
+
+```
+
+The overall dbRDA is significant and rank of methane explains 5% of the variation.
+Subsetting the Positive treatment in both G2 and G5, it is not significant: no
+correlation between dissimilarity and rank of methane between G2 and G5. This
+makes sense, because most of that variation would be due to selection in the jar.
+However, subsetting passage G5 and comparing N and P, 10% of dissimilarity is
+explained by rank of CH4.
+
+# Differential Abundance
+
+## Unique to each treatment
+
+```{r}
+g5n <- 
+  aes_data %>% 
+  filter(passage_treat == "G5N") %>% 
+  select(Family) %>% 
+  distinct() %$%
+  Family
+
+g5p <- 
+  aes_data %>% 
+  filter(passage_treat == "G5P") %>% 
+  select(Family) %>% 
+  distinct() %$%
+  Family
+
+p_not_n <- g5p[!g5p %in% g5n]
+
+n_not_p <- g5n[!g5n %in% g5p]
+
+n_not_p_prevalence <-
+  aes_data %>% 
+  filter(passage_treat == "G5N") %>% 
+  filter(Family %in% n_not_p) %>% 
+  select(sample, Family, count) %>% 
+  mutate(prevalence = as.numeric(count > 0)) %>% 
+  select(-count) %>% 
+  group_by(Family) %>% 
+  summarize(prevalence = sum(prevalence))
+
+n_not_p_prevalence %>% 
+  ggplot(aes(x = prevalence)) +
+  geom_histogram() +
+  scale_x_continuous(breaks = 1:12) +
+  theme_classic() +
+  labs(title = "Unique to Positive")
+
+# n_not_p_prevalence %>%
+#   filter(prevalence == 1) %>% 
+#   nrow
+
+# n_not_p_prevalence %>%
+#   filter(prevalence == 2) %>% 
+#   nrow
+# n_not_p_prevalence %>%
+#   filter(prevalence > 2) %$%
+#   Family
+
+p_not_n_prevalence <-
+  aes_data %>% 
+  filter(passage_treat == "G5P") %>% 
+  filter(Family %in% p_not_n) %>% 
+  select(sample, Family, count) %>% 
+  mutate(prevalence = as.numeric(count > 0)) %>% 
+  select(-count) %>% 
+  group_by(Family) %>% 
+  summarize(prevalence = sum(prevalence))
+
+p_not_n_prevalence %>% 
+  ggplot(aes(x = prevalence)) +
+  geom_histogram() +
+  scale_x_continuous(breaks = 1:12) +
+  theme_classic() +
+  labs(title = "Unique to Neutral")
+# 
+# p_not_n_prevalence %>%
+#   filter(prevalence == 1) %>% 
+#   nrow
+# 
+# p_not_n_prevalence %>%
+#   filter(prevalence == 2) %>% 
+#   nrow
+p_not_n_over_6 <- p_not_n_prevalence %>%
+  filter(prevalence > 6)
+
+```
+
+First, I looked at which Families are unique to each treatment in passage 5.
+There were `r length(n_not_p)` families unique to N and `r length(p_not_n)`
+unique to P. However, of the  `r length(n_not_p)` unique to P, 
+`r n_not_p_prevalence %>% filter(prevalence == 1) %>% nrow` had a prevalence
+of 1 and `r n_not_p_prevalence %>% filter(prevalence == 2) %>% nrow` had
+a prevalence of 2 leaving only a single family, which had a prevalence of
+`r n_not_p_prevalence %>% filter(prevalence > 2) %$% prevalence`, which was `r n_not_p_prevalence %>% filter(prevalence > 2) %$% Family`.
+Of the `r length(p_not_n)` Families unique to P, 2 had prevalences greater than
+6, which were `r p_not_n_over_6$Family[1]` with a prevalence of
+`r p_not_n_over_6$prevalence[1]` and `r p_not_n_over_6$Family[2]` with
+a prevalence of `r p_not_n_over_6$prevalence[2]`.
+
+## Setup Differential Abundance Data
+
+```{r setup-differential-abundance-data}
+
+aes_G5 <-
+  aes_data %>% 
+  filter(passage == "G5")
+
+aes_family_G5 <-
+  aes_G5 %>% 
+  select(asv, sample, treat, count, Family) %>%
+  group_by(sample, Family) %>% 
+  mutate(total = sum(count)) %>% 
+  ungroup() %>% 
+  select(sample, treat, Family, total) %>% 
+  distinct()
+
+prevalent_families <- 
+  aes_family_G5 %>% 
+  group_by(Family, treat) %>% 
+  summarize(prev = sum(total > 0), .groups = "drop") %>% 
+  pivot_wider(names_from = "treat", values_from = "prev", values_fill = 0) %>% 
+  filter(N/12 > 0.1 & P/12 > 0.1) %$%
+  Family
+
+otu_table <- 
+  aes_family_G5 %>% 
+  filter(Family %in% prevalent_families) %>% 
+  select(sample, Family, total) %>% 
+  pivot_wider(names_from = sample, values_from = total, values_fill = 0) %>% 
+  as.data.frame()
+
+rownames(otu_table) <- otu_table[,1]
+otu_table <- otu_table[,-1]
+
+sample_data <- aes_family_G5 %>% 
+  select(sample, treat) %>% 
+  distinct() %>% 
+  as.data.frame()
+
+rownames(sample_data) <- sample_data[,1]
+
+```
+
+## Corncob
+
+```{r}
+
+da_analysis <- differentialTest(formula = ~ treat,
+                                phi.formula = ~ 1,
+                                formula_null = ~ 1,
+                                phi.formula_null = ~ 1,
+                                test = "Wald", 
+                                boot = FALSE,
+                                data = otu_table,
+                                sample_data = sample_data,
+                                fdr_cutoff = 0.05, B = 100,
+                                taxa_are_rows = TRUE)
+
+dt_data <- plot(da_analysis, 
+                # level = c("Family"), 
+                data_only = TRUE)
+
+dt_plot <- function(dt_data) {
+  ggplot(dt_data, aes(x = x, xmin = xmin, xmax = xmax, y = reorder(taxa, x))) +
+    geom_pointrange() +
+    geom_vline(xintercept = 0) +
+    labs(x = "Estimate", y = NULL) +
+    coord_cartesian(clip="off") +
+    # theme(panel.border = element_blank()) +
+    theme_classic() +
+      theme(axis.text.y = element_markdown())
+
+}
+
+dt_data %>%
+  mutate(taxa = str_replace(taxa, "(.*)", "*\\1*"),
+         taxa = str_replace(taxa, "\\*(.*)_unclassified\\*",
+                                "Unclassified *\\1*")) %>% 
+  dt_plot()
+
+```
+
+## ANCOM
+
+```{r}
+
+# Setup phyloseq object
+aes_ps <- phyloseq(otu_table(otu_table, taxa_are_rows=TRUE), sample_data(sample_data))
+
+if(file.exists("Output/ancom_out.rds")) {
+  out <- 
+    readRDS("Output/ancom_out.rds")
+} else {
+  # Run Ancom
+  out = ancom(phyloseq = aes_ps,  
+              p_adj_method = "BH",
+              prv_cut = 0.1, 
+              lib_cut = 0, 
+              main_var = "treat",
+              rand_formula = NULL, 
+              lme_control = NULL,
+              struc_zero = TRUE, 
+              neg_lb = FALSE, 
+              alpha = 0.05, 
+              n_cl = 1)  
+  out %>% 
+    saveRDS("Output/ancom_out.rds")
+}
+
+res = out$res
+q_val = out$q_data
+beta_val = out$beta_data
+# Only consider the effect sizes with the corresponding q-value less than alpha
+beta_val = beta_val * (q_val < 0.05) 
+# Choose the maximum of beta's as the effect size
+beta_pos = apply(abs(beta_val), 2, which.max) 
+beta_max = vapply(seq_along(beta_pos), function(i) beta_val[beta_pos[i], i],
+                  FUN.VALUE = double(1))
+# Number of taxa except structural zeros
+n_taxa = ifelse(is.null(out$zero_ind), 
+                nrow(feature_table), 
+                sum(apply(out$zero_ind, 1, sum) == 0))
+# Cutoff values for declaring differentially abundant taxa
+cut_off = 0.7 * (n_taxa - 1)
+
+df_fig_w = res %>%
+  dplyr::mutate(beta = beta_max,
+                direct = case_when(
+                  detected_0.7 == TRUE & beta > 0 ~ "Positive",
+                  detected_0.7 == TRUE & beta <= 0 ~ "Negative",
+                  TRUE ~ "Not Significant"
+                  )) %>%
+  dplyr::arrange(W)
+df_fig_w$taxon_id = factor(df_fig_w$taxon_id, levels = df_fig_w$taxon_id)
+df_fig_w$W = replace(df_fig_w$W, is.infinite(df_fig_w$W), n_taxa - 1)
+df_fig_w$direct = factor(df_fig_w$direct, 
+                     levels = c("Negative", "Positive", "Not Significant"))
+
+p_w = df_fig_w %>%
+  ggplot(aes(x = taxon_id, y = W, color = direct)) +
+  geom_point(size = 2, alpha = 0.6) +
+  labs(x = "Taxon", y = "W") +
+  scale_color_discrete(name = NULL) + 
+  geom_hline(yintercept = cut_off, linetype = "dotted", 
+             color = "blue", size = 1.5) +
+  geom_text(aes(x = 2, y = cut_off + 0.5, label = "W[0.7]"), 
+            size = 5, vjust = -0.5, hjust = 0, color = "orange", parse = TRUE) +
+  theme_bw() +
+  theme(axis.text.x = element_blank(),
+        axis.ticks.x = element_blank(),
+        panel.grid.major = element_blank())
+p_w
+
+```
+
+## Aldex2
+
+```{r}
+
+conds <- colnames(otu_table) %>% 
+  substr(3, 3)
+
+if(file.exists("Output/aldex_x_all.rds")) {
+  x.all <- 
+    readRDS("Output/aldex_x_all.rds")
+} else {
+  x.all <- aldex(otu_table, 
+                 conds, mc.samples=1000, test="t", effect=TRUE,
+                 include.sample.summary=FALSE, denom="all", verbose=FALSE, paired.test=FALSE)
+  x.all %>% 
+    saveRDS("Output/aldex_x_all.rds")
+}
+
+par(mfrow=c(1,2))
+aldex.plot(x.all, type="MA", test="welch", xlab="Log-ratio abundance",
+    ylab="Difference")
+aldex.plot(x.all, type="MW", test="welch", xlab="Dispersion",
+    ylab="Difference")
+
+par(mfrow=c(1,2))
+plot(x.all$effect, x.all$we.ep, log="y", cex=0.7, col=rgb(0,0,1,0.2),
+  pch=19, xlab="Effect size", ylab="P value", main="Effect size plot")
+points(x.all$effect, x.all$we.eBH, cex=0.7, col=rgb(1,0,0,0.2),
+  pch=19)
+abline(h=0.05, lty=2, col="grey")
+legend(15,1, legend=c("P value", "BH-adjusted"), pch=19, col=c("blue", "red"))
+
+plot(x.all$diff.btw, x.all$we.ep, log="y", cex=0.7, col=rgb(0,0,1,0.2),
+  pch=19, xlab="Difference", ylab="P value", main="Volcano plot")
+points(x.all$diff.btw, x.all$we.eBH, cex=0.7, col=rgb(1,0,0,0.2),
+  pch=19)
+abline(h=0.05, lty=2, col="grey")
+
+x.all_effect_1 <-
+  x.all %>% 
+  filter(abs(effect) > 1) %>% 
+  as_tibble(rownames = "Family")
+
+# x.all_effect_1 %>% 
+#   # inner_join(taxa_table, by = "asv")
+#   ggplot(aes(y = reorder(Family, effect), x = effect)) +
+#   geom_point()
+
+# x.all_effect_1 %>% 
+#   select(Family, effect)
+
+```
+
+## Comparison
+
+### All Significant
+
+```{r}
+
+# CORNCOB
+corncob_taxa <- dt_data$taxa
+
+# ANCOM
+ancom_taxa <- as.character(df_fig_w[df_fig_w$detected_0.7, ]$taxon_id)
+
+# ALDEX2
+aldex_taxa <- x.all_effect_1$Family
+
+at_least_one_taxa <-
+	aes_family_G5 %>%
+	group_by(sample) %>%
+	mutate(rel_abund = (total / sum(total)) * 100) %>%
+	filter(Family %in% ancom_taxa | Family %in% aldex_taxa | Family %in% corncob_taxa) %>%
+	mutate(Family = str_replace(Family, "(.*)", "*\\1*"),
+	       Family = str_replace(Family, "\\*(.*)_unclassified\\*",
+				    "Unclassified *\\1*"))
+
+at_least_one_iqm <- 
+	at_least_one_taxa %>%
+	group_by(treat, Family) %>%
+	summarize(median = median(rel_abund), 
+		  first_quartile = quart(rel_abund)[1],
+		  third_quartile = quart(rel_abund)[2], .groups = "drop")
+
+sort_order <-
+	at_least_one_taxa %>%
+	group_by(Family) %>%
+	summarize(median = median(rel_abund)) %>%
+	arrange(median)
+
+at_least_one_taxa$Family <- factor(at_least_one_taxa$Family, levels = sort_order$Family)
+
+at_least_one_taxa %>%
+	ggplot(aes(x = rel_abund, y = Family, color = treat)) +
+	geom_point(position=position_jitterdodge()) +
+	geom_pointrange(aes(x = median, xmax = third_quartile, xmin = first_quartile, fill = treat), 
+			data = at_least_one_iqm, 
+			position = position_dodge(width = 0.75), 
+			shape = 21, 
+			color = "black") +
+	theme_bw() +
+	theme(axis.text.y = element_markdown()) +
+	labs(x = "Relative Abundance (%)", y = NULL) +
+	scale_color_manual(name = "Treatment",
+			   labels = c('Neutral', 'Positive'),
+			   values = c('gray40', 'darkorange2')) +
+	scale_fill_manual(name = "Treatment",
 			   labels = c('Neutral', 'Positive'),
@@
 
-Richness decreases with passage, but seems unchanged between 5n and 5p.
-Initial richness appears similar between the treatments.
+### Consensus Taxa
 
-# Hypothesis Testing
+```{r}
+aldex_effect <-
+  x.all_effect_1 %>% 
+  select(Family, effect)
 
-## Richness
+all_taxa <-
+	aes_family_G5 %>%
+	group_by(sample) %>%
+	mutate(rel_abund = (total / sum(total)) * 100) %>%
+  left_join(aldex_effect, by = "Family") %>% 
+	filter(Family %in% ancom_taxa & Family %in% aldex_taxa & Family %in% corncob_taxa) %>% 
+  ungroup()
 
-```{r richness}
+da_fams <-
+  all_taxa %>% 
+  select(Family) %>% 
+  distinct() %$%
+  Family
 
-# Test effect of passage number on richness using all samples
-richness_soil <- ps %>% breakaway
-plot(richness_soil, physeq=ps, color="passage")
-summary(richness_soil) %>% as_tibble
+all_taxa <-
+  all_taxa %>%
+	mutate(Family = str_replace(Family, "(.*)", "*\\1*"),
+	       Family = str_replace(Family, "\\*(.*)_unclassified\\*",
+				    "Unclassified *\\1*"))
+all_taxa$Family[all_taxa$Family == "Unclassified *0319-6G20*"] <- "Uncultured family 0319-6G20"
+all_taxa$Family[all_taxa$Family == "Unclassified *Chloroplast*"] <- "Chloroplast"
+all_taxa$Family[all_taxa$Family == "Unclassified *Armatimonadales*"] <- "Unclassified Armatimonadales"
+all_taxa$Family[all_taxa$Family == "Unclassified *Gammaproteobacteria*"] <- "Unclassified Gammaproteobacteria"
+all_taxa$Family[all_taxa$Family == "Unclassified *Kapabacteriales*"] <- "Unclassified Kapabacteriales"
 
-meta <- ps %>%
-  sample_data %>%
-  as_tibble %>%
-  mutate("sample_names" = ps %>% sample_names )
+all_iqm <- 
+	all_taxa %>%
+	group_by(treat, Family) %>%
+	summarize(median = median(rel_abund), 
+		  first_quartile = quart(rel_abund)[1],
+		  third_quartile = quart(rel_abund)[2], .groups = "drop")
 
-combined_richness <- meta %>%
-  left_join(summary(richness_soil),
-            by = "sample_names")
+sort_order <-
+	all_taxa %>%
+  select(Family, effect) %>% 
+  distinct() %>% 
+  arrange(effect)
+	# group_by(Family) %>%
+	# summarize(median = median(rel_abund)) %>%
+	# arrange(median)
 
-bt_passage <- betta(chats = combined_richness$estimate,
-                      ses = combined_richness$error,
-                      X = model.matrix(~passage, data = combined_richness))
-bt_passage$table
+all_taxa$Family <- factor(all_taxa$Family, levels = sort_order$Family)
+all_iqm$Family <- factor(all_iqm$Family, levels = sort_order$Family)
 
-# Test effect of treatment on richness in passage 5
+p <-
+	all_taxa %>%
+	ggplot(aes(x = rel_abund, y = Family, color = treat)) +
+	geom_point(position=position_jitterdodge()) +
+	geom_pointrange(aes(x = median, xmax = third_quartile, xmin = first_quartile, fill = treat), 
+			data = all_iqm, 
+			position = position_dodge(width = 0.75), 
+			shape = 21, 
+			color = "black") +
+	theme_bw() +
+	theme(axis.text.y = element_markdown()) +
+	labs(x = "Relative Abundance (%)", y = NULL) +
+	scale_color_manual(name = "Treatment",
+			   labels = c('Neutral', 'Positive'),
+			   values = c('gray40', 'darkorange2')) +
+	scale_fill_manual(name = "Treatment",
+			   labels = c('Neutral', 'Positive'),
+			   values = c('gray40', 'darkorange2'))
 
-ps_5 <- ps_filter(ps, passage == "5")
-richness_soil <- ps_5 %>% breakaway
-plot(richness_soil, physeq=ps_5, color="treat")
-summary(richness_soil) %>% as_tibble
+p_log10 <- 
+	p +
+	scale_x_log10(breaks = c(0.001, 0.01, 0.1, 1),
+		      labels = c(0.001, 0.01, 0.1, 1))
 
-meta <- ps_5 %>%
-  sample_data %>%
-  as_tibble %>%
-  mutate("sample_names" = ps_5 %>% sample_names )
+p
+p_log10
 
-combined_richness <- meta %>%
-  left_join(summary(richness_soil),
-            by = "sample_names")
+save(p_log10, da_fams, file = "Output/da_fig.RData")
 
-bt_treat5 <- betta(chats = combined_richness$estimate,
-                      ses = combined_richness$error,
-                      X = model.matrix(~treat, data = combined_richness))
-bt_treat5$table
+ggsave('Manuscript/fig4.jpg', p_log10, width = 7, height = 7)
 
-# Test effect of CH4 on richness
+## All
+# ggsave("shared_taxa.png", p)
+ggsave("shared_taxa_log.png", p_log10, width = 7, height = 7)
 
-bt_ch4 <- betta(chats = combined_richness$estimate,
-                      ses = combined_richness$error,
-                      X = model.matrix(~ch4, data = combined_richness))
-bt_ch4$table
+# aes_family_rclr_plot <-
+#   aes_family_rclr %>% 
+# 	filter(Family %in% ancom_taxa & Family %in% aldex_taxa & Family %in% corncob_taxa) %>% 
+#   pivot_longer(G2:G5, names_to = "passage", values_to = "rclr") %>% 
+#   select(Family, treat, rep, passage, rclr) %>% 
+#   distinct() %>% 	
+#   mutate(Family = str_replace(Family, "(.*)", "*\\1*"),
+# 	       Family = str_replace(Family, "\\*(.*)_unclassified\\*",
+# 				    "Unclassified *\\1*"))
+# 
+# aes_family_rclr_plot$Family[aes_family_rclr_plot$Family == "Unclassified *0319-6G20*"] <- "Uncultured family 0319-6G20"
+# aes_family_rclr_plot$Family[aes_family_rclr_plot$Family == "Unclassified *Chloroplast*"] <- "Chloroplast"
+# aes_family_rclr_plot$Family <- factor(aes_family_rclr_plot$Family, levels = sort_order$Family)
+# ggplot(aes_family_rclr_plot, aes(x = rclr, y = Family, color = treat)) +
+#   facet_grid( ~ passage) +
+#   geom_point(position = position_jitterdodge())
 
-save(bt_passage, bt_treat5, bt_ch4, file = 'Output/richness_models.RData')
+p_effect <-
+	all_taxa %>%
+	ggplot(aes(x = effect, y = Family)) +
+	geom_point() +
+
+	theme_bw() +
+	theme(axis.text.y = element_markdown()) +
+	labs(x = "Aldex2 Effect Size", y = NULL) +
+	scale_color_manual(name = "Treatment",
+			   labels = c('Neutral', 'Positive'),
+			   values = c('gray40', 'darkorange2')) +
+	scale_fill_manual(name = "Treatment",
+			   labels = c('Neutral', 'Positive'),
+			   values = c('gray40', 'darkorange2'))
+p_effect
 
@@
 
-Test confirms the above observation: Richness decreases between Passage 2
-and Passage 5. No difference in richness by treatment in passage 5 or by
-methane flux.
+## Methanogens and Methanotrophs
 
-# Beta Diversity
+Methanotrophy is (mostly) conserved at the family level.
+The Type I and Type X methanotrophs are in the Gammaproteobacteria in the families _Methylococcaceae_ and _Methylothermaceae_.
+There are also methanotrophs in the Verrumicrobiota in the family _Methylacidiphilaceae_.
+The Type II methanotrophs are in the _Alphaproteobacteria_ and are found in the families _Methylocystaceae_ and _Beijerinckiaceae_.
+However, the _Beijerinckiaceae_ also contain non-methanotrophs including N2-fixers and photoautotrophs.
+
+```{r methanotroph-lists}
+
+methanotrophs <- c('Methylococcaceae', 
+                   'Methylothermaceae', 
+                   'Methylocystaceae', 
+                   'Beijerinckiaceae', 
+                   'Methylacidiphilaceae')
+methanotrophs <- c(methanotrophs, paste(methanotrophs, "unclassified", sep = "_"))
+
+troph_orders <- c('Methylococcales')
+
+methanogen_classes <- c("Methanococci", "Methanomicrobia", "Methanobacteria")
+
+methanogen_classes <- c(methanogen_classes, paste(methanogen_classes, "unclassified", sep = "_"))
+
+```
 
@@
 
-# Replace unknown and missing taxa
-ps_fixed <-
-ps %>%
- tax_fix(
-  min_length = 4,
-  unknowns = c(""),
-  sep = " ", anon_unique = TRUE,
-  suffix_rank = "classified"
- )
+aes_data %>% 
+  select(Family, Genus, count, sample) %>% 
+  group_by(sample, Genus, Family) %>% 
+  summarize(total = sum(count), .groups = "drop") %>% 
+  group_by(sample) %>% 
+  mutate(rel_abund = total / sum(total)) %>% 
+  mutate(rclr = log(total/gm(total))) %>% 
+  filter(Family %in% methanotrophs) %>% 
+  select(-Family) %>% 
+  mutate(treat = substr(sample, 3, 3), passage = substr(sample, 1, 2)) %>% 
+  pivot_longer(rel_abund:rclr, names_to = "method", values_to = "value") %>% 
+  ggplot(aes(x = value, y = Genus, color = treat)) +
+  geom_jitter() +
+  facet_grid(passage ~ method, scales = "free") +
+  theme_bw()
 
-# Recode treatments as binary 1/0
-ps_fixed <- 
-  ps_fixed %>% 
-  ps_mutate(treat_bin = factor(ifelse(treat=='p', 1,0))) %>% 
-  ps_mutate(passage_bin = factor(ifelse(passage==5, 1,0))) %>% 
-  ps_mutate(treat_pas = interaction(treat_bin, passage_bin))
+aes_data %>% 
+  select(Family, count, sample) %>% 
+  group_by(sample, Family) %>% 
+  summarize(total = sum(count), .groups = "drop") %>% 
+  group_by(sample) %>% 
+  mutate(rel_abund = total / sum(total)) %>% 
+  mutate(rclr = log(total/gm(total))) %>% 
+  filter(Family %in% methanotrophs) %>% 
+  mutate(treat = substr(sample, 3, 3), passage = substr(sample, 1, 2)) %>% 
+  pivot_longer(rel_abund:rclr, names_to = "method", values_to = "value") %>% 
+  ggplot(aes(x = value, y = Family, color = treat)) +
+  geom_jitter() +
+  facet_grid(passage ~ method, scales = "free") +
+  theme_bw()
 
-ps_fixed
-```
-
-```{r calc-beta, results=FALSE}
-
-ord_calc_aitch <-
-  ps_fixed %>% 
-  tax_transform(trans = 'clr') %>% 
-  ord_calc(method = "PCA")
-
-ord_calc_bray <-
-  ps_fixed %>%
-  dist_calc("bray") %>%
-  ord_calc(method = "NMDS")
-
-ord_calc_jacc <-
-  ps_fixed %>%
-  dist_calc("jaccard", binary = TRUE) %>%
-  ord_calc(method = "PCoA")
+troph_asvs <-
+  aes_data %>% 
+  select(Kingdom:Species, asv) %>% 
+  distinct() %>% 
+  filter(Family %in% methanotrophs) %$%
+  asv
 
@@
 
-## Bray-Curtis and Jaccard
+Of the five methanotroph families, two are represented in this dataset:
+_Methylacidiphilaceae_ and _Beijerinckiaceae_. From these two families,
+there are eleven genera represented. These are spread across
+`r length(troph_asvs)` ASVs
 
-```{r plot-beta, results=FALSE, fig.width = 8, fig.height = 3}
-bray <-
-  ord_calc_bray %>% 
-  ord_plot(color = "treat") +
-  theme_tufte() +
-  geom_rangeframe() +
-  coord_cartesian(clip="off") +
-  scale_color_manual(name = "Treatment", 
-                       labels = c('Neutral', 'Positive'),
-                       values = c('gray40', 'darkorange2')) +
-  labs(title = "Bray-Curtis")
+```{r}
 
-jacc <- 
-  ord_calc_jacc %>% 
-  ord_plot(color = "treat") +
-  theme_tufte() +
-  geom_rangeframe() +
-  coord_fixed(3.9/48.3) +
-  # coord_cartesian(clip="off") +
-  scale_color_manual(name = "Treatment", 
-                       labels = c('Neutral', 'Positive'),
-                       values = c('gray40', 'darkorange2')) +
-  labs(title = "Jaccard")
-
-ggarrange(bray, jacc)
+aes_data %>% 
+  select(Class, count, sample) %>% 
+  group_by(sample, Class) %>% 
+  mutate(total = sum(count)) %>% 
+  select(-count) %>% 
+  distinct() %>% 
+  filter(Class %in% methanogen_classes)
 
@@
 
-## Aitchison
+A grand total of 10 reads map to methanogen classes across the entire dataset,
+including 2 reads in Methanobacteria (sample G2N02) and 8 reads of
+Methanomicrobia (sample G2N07).
+
+### Methanotroph differential abundance
 
 ```{r}
-plot_beta(ps_fixed)
-```
 
-## Beta diversity test passage by treatment
+troph_data <-
+  aes_data %>% 
+  select(sample, asv, treat, passage, count, Genus, Family) %>% 
+  filter(passage == "G5") %>% 
+  select(-passage) %>% 
+  filter(asv %in% troph_asvs)
 
-```{r beta-test}
+troph_data_family <-
+  troph_data %>% 
+  select(sample, Family, count) %>% 
+  group_by(sample, Family) %>% 
+  summarize(total = sum(count), .groups = "drop") %>% 
+  pivot_wider(names_from = sample, values_from = total, values_fill = 0) %>% 
+  as.data.frame()
 
-ps_clr <- microbiome::transform(ps, "clr")
+rownames(troph_data_family) <- troph_data_family[, 1]
+troph_data_family <- troph_data_family[, -1]
 
-fit_bin_jacc <- adonis2(otu_table(ps) ~ passage * treat, data = as(sample_data(ps), 'data.frame'), method = "jaccard", binary = TRUE)
+troph_data_genus <-
+  troph_data %>% 
+  select(sample, Genus, count) %>% 
+  group_by(sample, Genus) %>% 
+  summarize(total = sum(count), .groups = "drop") %>% 
+  pivot_wider(names_from = sample, values_from = total, values_fill = 0) %>% 
+  as.data.frame()
 
-fit_aitch <- adonis2(otu_table(ps_clr) ~ passage * treat, data = as(sample_data(ps_clr), 'data.frame'), method = "euclidean")
+rownames(troph_data_genus) <- troph_data_genus[, 1]
+troph_data_genus <- troph_data_genus[, -1]
 
-fit_adonis <- rbind(tidy(fit_aitch), tidy(fit_bin_jacc))
-fit_adonis$term <- c('Passage', 'Treatment', 'Pass:Treat', 'Residual', 'Total',
-      'Passage', 'Treatment', 'Pass:Treat', 'Residual', 'Total')
-save(fit_adonis, file = 'Output/beta_model.RData')
-
-############
-
-clr_dist_matrix <- phyloseq::distance(ps_clr, method = "euclidean")
-
-dispr <- vegan::betadisper(clr_dist_matrix, phyloseq::sample_data(ps_clr)$passage)
-disper_test <- permutest(dispr)
-
-dispr <- vegan::betadisper(clr_dist_matrix, phyloseq::sample_data(ps_clr)$treat)
-disper_test <- permutest(dispr)
-
-dist_aitch5 <-
-  ps %>% 
-  ps_filter(passage == 5) %>% 
-  dist_calc(dist = "aitchison")
-dist_permanova(dist_aitch5, "treat")
-groups <- factor(c(rep(1,12), rep(2,12)), labels = c("N","P"))
-
-dist_aitch5 %>% 
-  dist_get %>% 
-  vegan::betadisper(groups) %>% 
-  permutest()
-
-dist_aitch2 <-
-  ps %>% 
-  ps_filter(passage == 2) %>% 
-  dist_calc(dist = "aitchison")
-dist_permanova(dist_aitch5, "treat")
-groups <- factor(c(rep(1,12), rep(2,10)), labels = c("N","P"))
-
-dist_aitch2 %>% 
-  dist_get %>% 
-  vegan::betadisper(groups) %>% 
-  permutest()
+sample_data <- 
+  troph_data %>% 
+  select(sample, treat) %>% 
+  distinct() %>% 
+  as.data.frame()
+conds <- sample_data[, 2]
 
@@
 
-There is a significant main effect and interaction of passage and treatment on
-taxonomic dissimilarity of the soil microbiome. In addition, there is no difference
-in dispersion between treatments, but there is between passage2 and 5.
+```{r corncob-methanotrophs}
+if(file.exists("Output/cc_family.rds")) {
+  cc_family <- 
+    readRDS("Output/cc_family.rds")
+} else {
+cc_family <- differentialTest(formula = ~ treat,
+                                phi.formula = ~ 1,
+                                formula_null = ~ 1,
+                                phi.formula_null = ~ 1,
+                                test = "Wald", 
+                                boot = TRUE,
+                                data = troph_data_family,
+                                sample_data = sample_data,
+                                fdr_cutoff = 0.05, B = 1000)  
+cc_family %>% 
+    saveRDS("Output/cc_family.rds")
+}
 
-# Differential Abundance
+plot(cc_family)
+
+if(file.exists("Output/cc_genus.rds")) {
+  cc_genus <- 
+    readRDS("Output/cc_genus.rds")
+} else {
+cc_genus <- differentialTest(formula = ~ treat,
+                                phi.formula = ~ 1,
+                                formula_null = ~ 1,
+                                phi.formula_null = ~ 1,
+                                test = "Wald", 
+                                boot = TRUE,
+                                data = troph_data_genus,
+                                sample_data = sample_data,
+                                fdr_cutoff = 0.05, B = 1000) 
+cc_genus %>% 
+    saveRDS("Output/cc_genus.rds")
+}
+
+plot(cc_genus)
+
+```
+
+```{r stop}
+
+knitr::opts_chunk$set(eval = FALSE)
+
+```
+
+```{r}
+
+troph_genus_data <-
+	aes_data %>%
+	filter(Family %in% methanotrophs) %>%
+	select(sample, count, Genus, passage, treat) %>%
+	group_by(sample, Genus) %>%
+	mutate(total = sum(count)) %>%
+	select(-count) %>%
+	ungroup() %>%
+	group_by(sample) %>%
+	mutate(rel_abund = total / sum(total)) %>%
+	ungroup() %>%
+	distinct() 
+troph_genus_iqm <- 
+	troph_genus_data %>%
+	group_by(passage, treat, Genus) %>%
+	summarize(median = median(rel_abund), 
+		  first_quartile = quart(rel_abund)[1],
+		  third_quartile = quart(rel_abund)[2], .groups = "drop")
+troph_genus_data %>%
+	ggplot(aes(x = rel_abund, y = Genus, color = treat)) +
+	geom_point(position=position_jitterdodge()) +
+	facet_grid( ~ passage) +
+	geom_pointrange(aes(x = median, xmax = third_quartile, xmin = first_quartile, fill = treat), data = troph_genus_iqm, position = position_dodge(width = 0.75), shape = 21, color = "black") +
+	theme_bw()
+ggsave("troph_genus.png")
+
+aes_data %>%
+	select(sample, count, Genus, passage, treat) %>%
+	group_by(sample, Genus) %>%
+	mutate(total = sum(count)) %>%
+	select(-count) %>%
+	ungroup() %>%
+	group_by(sample) %>%
+	mutate(rel_abund = total / sum(total)) %>%
+	ungroup() %>%
+	filter(Genus %in% c("Methylocella", "Microvirga", "Methylorosula")) %>%
+	distinct() %>%
+	ggplot(aes(x = rel_abund, y = Genus)) +
+	geom_jitter() +
+	facet_grid(treat ~ passage)
+
+troph_family_data <-
+	aes_data %>%
+	select(sample, count, Family, passage, treat) %>%
+	group_by(sample, Family) %>%
+	mutate(total = sum(count)) %>%
+	select(-count) %>%
+	ungroup() %>%
+	group_by(sample) %>%
+	mutate(rel_abund = total / sum(total)) %>%
+	ungroup() %>%
+	filter(Family %in% methanotrophs) %>%
+	distinct()
+troph_family_iqm <- 
+	troph_family_data %>%
+	group_by(passage, treat, Family) %>%
+	summarize(median = median(rel_abund), 
+		  first_quartile = quart(rel_abund)[1],
+		  third_quartile = quart(rel_abund)[2], .groups = "drop")
+troph_family_data %>%
+	ggplot(aes(x = rel_abund, y = Family, color = treat)) +
+	geom_point(position=position_jitterdodge()) +
+	facet_grid( ~ passage) +
+	geom_pointrange(aes(x = median, xmax = third_quartile, xmin = first_quartile, fill = treat), data = troph_family_iqm, position = position_dodge(width = 0.75), shape = 21, color = "black") +
+	theme_bw()
+ggsave("troph_family.png")
+
+troph_species_data <-
+	aes_data %>%
+	filter(Family %in% methanotrophs) %>%
+	select(sample, count, Species, passage, treat) %>%
+	group_by(sample, Species) %>%
+	mutate(total = sum(count)) %>%
+	select(-count) %>%
+	ungroup() %>%
+	group_by(sample) %>%
+	mutate(rel_abund = total / sum(total)) %>%
+	ungroup() %>%
+	distinct()
+troph_species_iqm <- 
+	troph_species_data %>%
+	group_by(passage, treat, Species) %>%
+	summarize(median = median(rel_abund), 
+		  first_quartile = quart(rel_abund)[1],
+		  third_quartile = quart(rel_abund)[2], .groups = "drop")
+troph_species_data %>%
+	ggplot(aes(x = rel_abund, y = Species, color = treat)) +
+	geom_point(position=position_jitterdodge()) +
+	facet_grid( ~ passage) +
+	geom_pointrange(aes(x = median, xmax = third_quartile, xmin = first_quartile, fill = treat), data = troph_species_iqm, position = position_dodge(width = 0.75), shape = 21, color = "black") +
+	theme_bw()
+ggsave("troph_species.png")
+
+```
+
+```{r aldex-methanotrophs-family}
+
+x.all <- aldex(troph_data_family, 
+               conds, mc.samples=1000, test="t", effect=TRUE,
+     include.sample.summary=FALSE, denom="all", verbose=FALSE, paired.test=FALSE)
+
+par(mfrow=c(1,2))
+aldex.plot(x.all, type="MA", test="welch", xlab="Log-ratio abundance",
+    ylab="Difference")
+aldex.plot(x.all, type="MW", test="welch", xlab="Dispersion",
+    ylab="Difference")
+
+par(mfrow=c(1,2))
+plot(x.all$effect, x.all$we.ep, log="y", cex=0.7, col=rgb(0,0,1,0.2),
+  pch=19, xlab="Effect size", ylab="P value", main="Effect size plot")
+points(x.all$effect, x.all$we.eBH, cex=0.7, col=rgb(1,0,0,0.2),
+  pch=19)
+abline(h=0.05, lty=2, col="grey")
+legend(15,1, legend=c("P value", "BH-adjusted"), pch=19, col=c("blue", "red"))
+
+plot(x.all$diff.btw, x.all$we.ep, log="y", cex=0.7, col=rgb(0,0,1,0.2),
+  pch=19, xlab="Difference", ylab="P value", main="Volcano plot")
+points(x.all$diff.btw, x.all$we.eBH, cex=0.7, col=rgb(1,0,0,0.2),
+  pch=19)
+abline(h=0.05, lty=2, col="grey")
+
+x.all
+
+```
+
+```{r aldex-methanotrophs-genus}
+
+x.all <- aldex(troph_data_genus, 
+               conds, mc.samples=1000, test="t", effect=TRUE,
+     include.sample.summary=FALSE, denom="all", verbose=FALSE, paired.test=FALSE)
+
+par(mfrow=c(1,2))
+aldex.plot(x.all, type="MA", test="welch", xlab="Log-ratio abundance",
+    ylab="Difference")
+aldex.plot(x.all, type="MW", test="welch", xlab="Dispersion",
+    ylab="Difference")
+
+par(mfrow=c(1,2))
+plot(x.all$effect, x.all$we.ep, log="y", cex=0.7, col=rgb(0,0,1,0.2),
+  pch=19, xlab="Effect size", ylab="P value", main="Effect size plot")
+points(x.all$effect, x.all$we.eBH, cex=0.7, col=rgb(1,0,0,0.2),
+  pch=19)
+abline(h=0.05, lty=2, col="grey")
+legend(15,1, legend=c("P value", "BH-adjusted"), pch=19, col=c("blue", "red"))
+
+plot(x.all$diff.btw, x.all$we.ep, log="y", cex=0.7, col=rgb(0,0,1,0.2),
+  pch=19, xlab="Difference", ylab="P value", main="Volcano plot")
+points(x.all$diff.btw, x.all$we.eBH, cex=0.7, col=rgb(1,0,0,0.2),
+  pch=19)
+abline(h=0.05, lty=2, col="grey")
+
+x.all
+
+```
+
+```{r}
+## ANCOM
+```
+```{r}
+
+# Setup phyloseq objects
+
+ps_family <- phyloseq(otu_table(troph_data_family,
+                                taxa_are_rows = TRUE),
+                      sample_data(sample_data))
+ps_genus <- phyloseq(otu_table(troph_data_genus,
+                                taxa_are_rows = TRUE),
+                      sample_data(sample_data))
+
+# Run Ancom
+
+out_family = ancom(phyloseq = ps_family,  
+                  p_adj_method = "BH",
+                  prv_cut = 0.1, 
+                  lib_cut = 0, 
+                  main_var = "treat",
+                  rand_formula = NULL, 
+                  lme_control = NULL,
+                  struc_zero = TRUE, 
+                  neg_lb = FALSE, 
+                  alpha = 0.05, 
+                  n_cl = 1)
+
+out = ancom(phyloseq = ps_genus,  
+                  p_adj_method = "BH",
+                  prv_cut = 0.1, 
+                  lib_cut = 0, 
+                  main_var = "treat",
+                  rand_formula = NULL, 
+                  lme_control = NULL,
+                  struc_zero = TRUE, 
+                  neg_lb = FALSE, 
+                  alpha = 0.05, 
+                  n_cl = 1)
+
+```
+
+```{r}
+
+q_val = out$q_data
+beta_val = out$beta_data
+# Only consider the effect sizes with the corresponding q-value less than alpha
+beta_val = beta_val * (q_val < 0.05) 
+# Choose the maximum of beta's as the effect size
+beta_pos = apply(abs(beta_val), 2, which.max) 
+beta_max = vapply(seq_along(beta_pos), function(i) beta_val[beta_pos[i], i],
+                  FUN.VALUE = double(1))
+
+# Number of taxa except structural zeros
+n_taxa = ifelse(is.null(out$zero_ind), 
+                nrow(feature_table), 
+                sum(apply(out$zero_ind, 1, sum) == 0))
+
+# Cutoff values for declaring differentially abundant taxa
+cut_off = 0.7 * (n_taxa - 1)
+
+df_fig_w = out$res %>%
+  dplyr::mutate(beta = beta_max,
+                direct = case_when(
+                  detected_0.7 == TRUE & beta > 0 ~ "Positive",
+                  detected_0.7 == TRUE & beta <= 0 ~ "Negative",
+                  TRUE ~ "Not Significant"
+                  )) %>%
+  dplyr::arrange(W)
+df_fig_w$taxon_id = factor(df_fig_w$taxon_id, levels = df_fig_w$taxon_id)
+df_fig_w$W = replace(df_fig_w$W, is.infinite(df_fig_w$W), n_taxa - 1)
+df_fig_w$direct = factor(df_fig_w$direct, 
+                     levels = c("Negative", "Positive", "Not Significant"))
+
+p_w = df_fig_w %>%
+  ggplot(aes(x = taxon_id, y = W, color = direct)) +
+  geom_point(size = 2, alpha = 0.6) +
+  labs(x = "Taxon", y = "W") +
+  scale_color_discrete(name = NULL) + 
+  geom_hline(yintercept = cut_off, linetype = "dotted", 
+             color = "blue", size = 1.5) +
+  geom_text(aes(x = 2, y = cut_off + 0.5, label = "W[0.7]"), 
+            size = 5, vjust = -0.5, hjust = 0, color = "orange", parse = TRUE) +
+  theme_bw() +
+  theme(axis.text.x = element_blank(),
+        axis.ticks.x = element_blank(),
+        panel.grid.major = element_blank())
+p_w
+
+```
+
+```{r}
+
+otu_table <-  aes_data %>% 
+  filter(asv %in% troph_asvs) %>% 
+  filter(passage == "G5") %>% 
+  select(sample, count, Genus) %>% 
+  group_by(Genus, sample) %>% 
+  summarize(total = sum(count), .groups = "drop") %>% 
+  mutate(treat = substr(sample, 3, 3)) %>% 
+  group_by(Genus, treat) %>% 
+  mutate(prev = sum(total > 0)) %>% 
+  ungroup() %>% 
+  filter(prev > 2) %>% 
+  select(-treat, -prev) %>% 
+  pivot_wider(names_from = sample, values_from = total, values_fill = 0) %>% 
+  as.data.frame()
+
+rownames(otu_table) <- otu_table[, 1]
+otu_table <- otu_table[, -1]
+
+conds <- 
+  otu_table %>% 
+  colnames() %>% 
+  substr(3, 3)
+
+x.all <- aldex(otu_table, 
+               conds, mc.samples=1000, test="t", effect=TRUE,
+     include.sample.summary=FALSE, denom="all", verbose=FALSE, paired.test=FALSE)
+
+par(mfrow=c(1,2))
+aldex.plot(x.all, type="MA", test="welch", xlab="Log-ratio abundance",
+    ylab="Difference")
+aldex.plot(x.all, type="MW", test="welch", xlab="Dispersion",
+    ylab="Difference")
+
+par(mfrow=c(1,2))
+plot(x.all$effect, x.all$we.ep, log="y", cex=0.7, col=rgb(0,0,1,0.2),
+  pch=19, xlab="Effect size", ylab="P value", main="Effect size plot")
+points(x.all$effect, x.all$we.eBH, cex=0.7, col=rgb(1,0,0,0.2),
+  pch=19)
+abline(h=0.05, lty=2, col="grey")
+legend(15,1, legend=c("P value", "BH-adjusted"), pch=19, col=c("blue", "red"))
+
+plot(x.all$diff.btw, x.all$we.ep, log="y", cex=0.7, col=rgb(0,0,1,0.2),
+  pch=19, xlab="Difference", ylab="P value", main="Volcano plot")
+points(x.all$diff.btw, x.all$we.eBH, cex=0.7, col=rgb(1,0,0,0.2),
+  pch=19)
+abline(h=0.05, lty=2, col="grey")
+
+x.all_effect_1 <-
+  x.all %>% 
+  filter(abs(effect) > 0.5) %>% 
+  as_tibble(rownames = "asv")
+
+x.all_effect_1 %>% 
+  ggplot(aes(y = reorder(asv, effect), x = effect)) +
+  geom_point()
+
+```
+
+```{r aldex_genus_everything}
+
+otu_table <- 
+  aes_data %>% 
+  filter(passage == "G5") %>% 
+  select(Genus, count, sample) %>% 
+  group_by(Genus, sample) %>% 
+  summarize(total = sum(count), .groups = "drop") %>% 
+  mutate(treat = substr(sample, 3, 3)) %>% 
+  group_by(Genus, treat) %>% 
+  mutate(prev = sum(total > 0)) %>% 
+  ungroup() %>% 
+  filter(prev > 2) %>% 
+  select(-treat, -prev) %>% 
+  pivot_wider(names_from = "sample", values_from = "total", values_fill = 0) %>% 
+  as.data.frame()
+
+rownames(otu_table) <- otu_table[, 1]
+
+otu_table <- otu_table[, -1]
+
+x.all <- aldex(otu_table, 
+               conds, mc.samples=1000, test="t", effect=TRUE,
+     include.sample.summary=FALSE, denom="all", verbose=FALSE, paired.test=FALSE)
+
+par(mfrow=c(1,2))
+aldex.plot(x.all, type="MA", test="welch", xlab="Log-ratio abundance",
+    ylab="Difference")
+aldex.plot(x.all, type="MW", test="welch", xlab="Dispersion",
+    ylab="Difference")
+
+par(mfrow=c(1,2))
+plot(x.all$effect, x.all$we.ep, log="y", cex=0.7, col=rgb(0,0,1,0.2),
+  pch=19, xlab="Effect size", ylab="P value", main="Effect size plot")
+points(x.all$effect, x.all$we.eBH, cex=0.7, col=rgb(1,0,0,0.2),
+  pch=19)
+abline(h=0.05, lty=2, col="grey")
+legend(15,1, legend=c("P value", "BH-adjusted"), pch=19, col=c("blue", "red"))
+
+plot(x.all$diff.btw, x.all$we.ep, log="y", cex=0.7, col=rgb(0,0,1,0.2),
+  pch=19, xlab="Difference", ylab="P value", main="Volcano plot")
+points(x.all$diff.btw, x.all$we.eBH, cex=0.7, col=rgb(1,0,0,0.2),
+  pch=19)
+abline(h=0.05, lty=2, col="grey")
+
+x.all_effect_1 <-
+  x.all %>% 
+  filter(abs(effect) > 1) %>% 
+  as_tibble(rownames = "Genus")
+
+x.all_effect_1 %>% 
+  ggplot(aes(y = reorder(Genus, effect), x = effect)) +
+  geom_point()
+
+```
+
+```{r ancova-clr}
+
+gm <- function(x){
+
+  exp(mean(log(x[x>0])))
+
+}
+
+rclr <- function(x) {
+  log(x/gm(x))
+}
+
+aes_family <-
+  aes_data %>%
+  select(Family, count, sample, passage, treat) %>%
+  group_by(Family, sample) %>%
+  mutate(total = sum(count)) %>%
+  select(-count) %>%
+  ungroup() %>%
+  distinct() %>%
+  mutate(rep = as.numeric(substr(sample, 4, 5))) %>%
+  select(-sample)
+
+prevalent_families <- 
+  aes_family %>% 
+  mutate(passagetreat = paste0(passage, treat)) %>%
+  group_by(Family, passagetreat) %>% 
+  summarize(prev = sum(total > 0), .groups = "drop") %>% 
+  pivot_wider(names_from = "passagetreat", values_from = "prev", values_fill = 0) %>% 
+  filter(G2N/12 > 0.1 & G2P/12 > 0.1 & G5N/12 > 0.1 & G5P/12 > 0.1) %$%
+  Family
+
+aes_family_rclr <-
+  aes_family %>% 
+  filter(Family %in% prevalent_families) %>% 
+  group_by(passage, treat, rep) %>%
+  mutate(rclr = log(total/gm(total))) %>%
+  select(-total) %>%
+  ungroup() %>%
+  pivot_wider(names_from = passage, values_from = rclr)
+
+ancova <- function(df) {
+  lm(G5 ~ G2 + treat, data = df, na.action = )
+}
+
+model_results <-
+aes_family_rclr %>%
+  nest(data = -Family) %>%
+  mutate(model = map(data, ancova)) %>%
+  mutate(tidy = map(model, tidy)) %>% 
+  unnest(tidy) %>% 
+  filter(term == "treatP")
+
+model_results$p.adjust <- p.adjust(model_results$p.value, method = "BH")
+
+model_results %>%
+  filter(p.adjust < 0.05) %>% 
+  arrange(estimate) %>% 
+  print(n = 22)
+
+model_results %>% 
+  filter(p.adjust < 0.05) %>% 
+  arrange(estimate) %>% 
+  unnest(data) %>% 
+  ggplot(aes(x = estimate, xmax = estimate + std.error, xmin = estimate - std.error, y = reorder(Family, estimate))) +
+  geom_pointrange() +
+  labs(x = "Estimate \n(log-ratio increase in P)", y = NULL) +
+  theme_bw()
+ggsave("ancova-plot.png", width = 5, height = 5)
+
+model_results %>% 
+  filter(p.adjust < 0.05) %>% 
+  unnest(data) %>% 
+  pivot_longer(G2:G5, names_to = "passage", values_to = "rclr") %>% 
+  select(Family, treat, rep, passage, rclr) %>% 
+  distinct() %>% 
+  ggplot(aes(x = rclr, y = Family, color = treat)) +
+  facet_grid( ~ passage) +
+  geom_point(position = position_jitterdodge())
+
+```
+
+```{r}
+
+aes_family_data <-
+  aes_data %>% 
+  select(asv, sample, passage, treat, count, Family) %>%
+  group_by(sample, Family) %>% 
+  mutate(total = sum(count)) %>% 
+  ungroup() %>% 
+  select(sample, treat, passage, Family, total) %>% 
+  distinct()
+
+prevalent_families <- 
+  aes_family_data %>% 
+  group_by(Family, treat, passage) %>% 
+  summarize(prev = sum(total > 0), .groups = "drop") %>% 
+  pivot_wider(names_from = "treat", values_from = "prev", values_fill = 0) %>% 
+  filter(N/12 > 0.1 & P/12 > 0.1) %$%
+  Family
+
+aes_family_data <-
+  aes_family_data %>% 
+  filter(Family %in% prevalent_families) %>% 
+  mutate(jar = substr(sample, 3, 5)) %>%
+  select(-sample) %>%
+  pivot_wider(names_from = "passage", values_from = "total", values_fill = 0)
+
+ancova_model <- function(df) {
+	lm(G5 ~ G2 + treat, data = df)
+}
+
+aes_family_nested <-
+	aes_family_data %>%
+	group_by(Family) %>%
+	nest() %>%
+	mutate(model = map(data, ancova_model)) %>%
+	mutate(tidy = map(model, tidy)) %>%
+	ungroup()
+
+aes_ancova <- aes_family_nested %>%
+	unnest(tidy) %>%
+	filter(term == "treatP")
+
+aes_ancova$p.adjust <- p.adjust(aes_ancova$p.value, method = "BH")
+
+aes_ancova %>%
+	arrange(estimate) %>%
+	filter(p.adjust < 0.05) %>%
+	print(n = 22)
+
+```
+
+```{r}
+
+aes_family_data <-
+  aes_data %>% 
+  select(asv, sample, passage, treat, count, Family) %>%
+  group_by(sample, Family) %>% 
+  mutate(total = sum(count)) %>% 
+  ungroup() %>% 
+  select(sample, treat, passage, Family, total) %>% 
+  distinct()
+
+prevalent_families <- 
+  aes_family_data %>% 
+  group_by(Family, treat, passage) %>% 
+  summarize(prev = sum(total > 0), .groups = "drop") %>% 
+  pivot_wider(names_from = "treat", values_from = "prev", values_fill = 0) %>% 
+  filter(N/12 > 0.1 & P/12 > 0.1) %$%
+  Family
+
+otu_table <- 
+  aes_family_data %>% 
+  filter(Family %in% prevalent_families) %>% 
+  select(sample, Family, total) %>% 
+  pivot_wider(names_from = sample, values_from = total, values_fill = 0) %>% 
+  as.data.frame()
+
+rownames(otu_table) <- otu_table[,1]
+otu_table <- otu_table[,-1]
+
+sample_data <- aes_family_data %>% 
+  select(sample, treat, passage) %>% 
+  distinct() %>% 
+  as.data.frame()
+
+rownames(sample_data) <- sample_data[,1]
+sample_data <- sample_data[, -1]
+sample_data <-
+	sample_data %>%
+	mutate(passage12 = if_else(passage == "G2", 1, 2))
+
+# Setup phyloseq object
+aes_ps <- phyloseq(otu_table(otu_table, taxa_are_rows=TRUE), sample_data(sample_data))
+
+set.seed(123)
+out2 = ancom(phyloseq = aes_ps,  p_adj_method = "holm",
+             prv_cut = 0.10, lib_cut = 1000, main_var = "treat", 
+             adj_formula = "passage12", 
+             rand_formula = "~ passage12 | 1",
+             lme_control = list(maxIter = 100, msMaxIter = 100, opt = "optim"), 
+             struc_zero = TRUE, neg_lb = FALSE, alpha = 0.05, n_cl = 2)
+
+res2 = out2$res
+
+out2 = ancom(phyloseq = family_data2,  p_adj_method = "holm",
+             prv_cut = 0.10, lib_cut = 1000, main_var = "group", 
+             adj_formula = "nationality + timepoint + bmi_group", 
+             rand_formula = "~ timepoint | subject",
+             lme_control = list(maxIter = 100, msMaxIter = 100, opt = "optim"), 
+             struc_zero = TRUE, neg_lb = TRUE, alpha = 0.05, n_cl = 2)
+
+res2 = out2$res
+
+```
+
+```{r}
+
+x.all_effect_1$Family
+
+dt_data$taxa
+
+```
+
+```{r}
+combined_da <-
+  dt_data %>% 
+  as_tibble %>% 
+  full_join(sig_families, by = c("taxa" = "Family")) %>% 
+  full_join(x.all_effect_1, by = c("taxa" = "Family")) %>% 
+  full_join(df_fig_w[df_fig_w$detected_0.7,], by = c("taxa" = "taxon_id"))
+
+shared_da <-
+  dt_data %>% 
+  as_tibble %>% 
+  inner_join(sig_families, by = c("taxa" = "Family")) %>% 
+  inner_join(x.all_effect_1, by = c("taxa" = "Family")) %>% 
+  inner_join(df_fig_w[df_fig_w$detected_0.7,], by = c("taxa" = "taxon_id"))
+
+shared_da %>% 
+  ggplot(aes(y = reorder(taxa, effect), x = effect)) +
+  geom_point()
+
+aes_data %>% 
+  filter(passage == "G5") %>% 
+  select(sample, Family, count) %>% 
+  group_by(sample, Family) %>% 
+  summarize(total = sum(count)) %>% 
+  filter(Family %in% shared_da$taxa) %>% 
+  mutate(treat = substr(sample, 3, 3)) %>% 
+  ggplot(aes(x = total, y = Family, color = treat)) +
+  geom_jitter() +
+  scale_x_log10()
+
+rubini_seqs <-
+  aes_data %>% 
+  filter(Family == "Rubinisphaeraceae") %>% 
+  select(asv, seq, Kingdom:Species) %>% 
+  distinct()
+
+write(rubini_seqs$seq, file = 'Output/rubini_seqs.txt')
+
+```
+
+```{r}
+
+##############
 
 First, I looked at which taxa are unique to treatments N and P. Several families
-present in P and absent in N. Most of them had low prevalance (present in 3 or
+present in P and absent in N. Most of them had low prevalence (present in 3 or
 fewer samples) and low abundance (median < 5 reads across samples). There were two families unique
@@
 Armatimonas was isolated from the rhizoplane of Phragmites australis (https://www.microbiologyresearch.org/content/journal/ijsem/10.1099/ijs.0.025643-0)
+```
 
@@
 
-## Differential Abundance Treatment in passage 5
+```{r}
+# Differential Abundance Treatment in passage 5
+```
 
 ```{r corncob_treatment, results=FALSE, fig.hight=40}
+
 ps_fixed_family <- speedyseq::tax_glom(ps_fixed, taxrank="Family")
@@
 dt_plot(dt_data)
-
-```
-
-```{r stop}
-knitr::opts_chunk$set(eval = FALSE)
 

### 2022-07-05 R/functions.R

@@ -1,12 +1,12 @@
 
-raw_dir <- '../Data/' # Raw Data Directory
-der_dir <- '../Output/' # Derived/Modified Data Directory
+raw_dir <- 'Data/' # Raw Data Directory
+der_dir <- 'Output/' # Derived/Modified Data Directory
 
 # theme_set(theme_tufte()+
 #             theme(axis.title.y = element_text(angle = 0, vjust = 0.5)))
-# theme_set(theme_classic() +
-#             theme(panel.border = element_rect(fill = NA, size = 1.0625),
-#                   axis.line = element_blank()))
+theme_set(theme_classic() +
+            theme(panel.border = element_rect(fill = NA, size = 1.0625),
+                  axis.line = element_blank()))
 
 plot_flux <- function(fluxes, passage, estimate, ratio = NULL, log10 = FALSE) {
   p <- ggplot(fluxes, aes(x = {{passage}}, y = {{estimate}}, color = treat)) +
@@ -229,3 +229,4 @@ quart <- function(x) {
   }
   c(Q1=median(x[1:l]), Q3=median(x[u:n]))
 }
+

### 2022-07-06 R/03-response.Rmd

@@
 
-plot_flux(fluxes, passage, log_ch4, ratio_log10, log10 = TRUE)
+p_response <- plot_flux(fluxes, passage, log_ch4, ratio_log10, log10 = TRUE)
 
-#ggsave('fluxes_log.pdf', width = 5, height = 4)
+save(p_response, file = 'Output/response_plot.RData')
+
+ggsave('Manuscript/fig1.jpg', p_response, width = 4, height = 3)
+
 ```
@@
 ```{r save_model}
-save(response_model, compare_response_mods, per_slope, file = '../Output/response_model.RData')
+save(response_model, compare_response_mods, per_slope, file = 'Output/response_model.RData')
 ```

### 2022-07-06 Manuscript/manuscript.Rmd
start to update text and figures with new methods, results

@@
 library(flextable)
+library(ggtext)
 figures <- captioner()
@@
 transformed to better meet the assumptions of a linear model and to make
-figures easier to interpret. First, we tested the effect of treatment by
-fitting two nested models with and without treatment using 'lm'. We compared these
-models using the likelihood ratio test with the 'anova' function. We
-then fit the full model to estimate the slope of the positive line, which
-represents the change in CH~4~ oxidation rate per passage as a response to
+figures easier to interpret. First, we tested the difference of slopes between
+the positive and neutral treatments by
+fitting two nested models with and without the interaction between treatment
+and passage using 'lm'. We compared these
+models using an F-test with the 'anova' function. We
+then fit the full model with the interaction term to estimate the slopes for each treatment, which
+represented the change in CH~4~ oxidation rate per passage as a response to
 selection.
 
-To estimate the proportion of variance in ecosystem function due to variation in
+To estimate the proportion of variance in CH~4~ oxidation rate due to variation in
 the microbiome, we estimated narrow-sense heritability ($h^2$) as the
@@
 produced by those parents. First, we tested if there was an effect of treatment
-on the heritability estimate. We compared these models using the likelihood ratio
+on the heritability estimate. We compared these models using an F-test
 test with the 'anova' function. We then fit the full model to estimate the
@@
 
-Richness estimates and tests were performed using 'breakaway' [@willis2015]. We
+Richness was estimated using the method from @hurlbert1971 with a subsample size of `r min_seqs` using the 'rarefy'
+function in 'vegan' [@oksanen2019]. We
 tested a difference in richness between Passages 2 and 5 with both treatments
 combined. Then we subset the samples from Passage 5 and compared richness between
-the Positive and Neutral treatment. Next, we compared beta-diversity as both
-Aitchison dissimilarity, which is the Euclidean distance after center-log ratio
-transformation, and the binary Jaccard distance for presence or absence [@aitchison1982]. 
+the Positive and Neutral treatment. Next, we compared beta-diversity as the 
+Bray-Curtis dissimilarity by averaging random subsets with a subsample size of `r min_seqs` using the 'avgdist' function in 'vegan' [@bray1957; @oksanen2019].
 We tested a difference in centroid and dispersion of beta
@@
 To identify taxa that responded to selection on CH~4~ oxidation rate, we tested
-differential abundance using a beta-binomial model [@martin2020]. We first
+differential abundance between the two treatments in passage 5. We first
 grouped ASVs at the family level keeping ASVs that lacked a family-level taxonomic
-assignment. We
+assignment by grouping them at a higher level. We
 then subset the samples in Passage 5 and removed all families with a prevalence
-of less than 10\%. We compared families in passage 5 between the Positive and
-Neutral treatment to identify taxa that were enriched or depleted as a response
-to selection. we used the likelihood ratio test to estimate p-values with a
-significance threshold of false-discovery rate < 0.05 [@martin2020]. Estimates
-are presented as the expected relative abundance in the positive treatment
-relative to the neutral treatment.
+of less than 10\% in either treatment. We used three methods for testing differential abundance: ANCOM-II, ALDEx2, and CORNCOB. We then
+pulled out the consensus taxa that were significant
+with all three tests and plotted their relative abundances.
+For ANCOM-II, we used the 'ancom' function in the 'ANCOM-BC' package with a cutoff of W = 0.7 [@kaul2017; @lin2020].
+For ALDEx2, we used the 'aldex' function in the 'ALDEx2' package with the Welch's t test and we used an effect size of
+1 as our significance threshold [@fernandes2014]. Finally, we used CORNCOB with the 'differentialTest' function in the 'corncob' package with the Wald test and without bootstrapping [@martin2020].
 
@@
 We observed a response to selection on whole-ecosystem soil CH~4~ oxidation
-rate (`r figures("response", display = "cite")`; `r print_lrt(compare_response_mods, 2, 'LRT of nested models with and without treatment')`. At the start of the experiment, the Positive
+rate (`r figures("response", display = "cite")`; difference of slopes:
+F~`r compare_response_mods[2,3]`,`r compare_response_mods[2,1]`~ = `r myround(compare_response_mods[2,5], 2)`, p = `r myround(compare_response_mods[2,6], 2)`).
+At the start of the experiment, the Positive
 treatment had a mean CH~4~ oxidation rate that was
@@
 
-Richness of ASVs decreased between Passage 2 (median = `r myround(rare_median[1, 2], 1)`, IQR = `r myround(rare_median[1, 3], 1)`) and 5 (median = `r myround(rare_median[2, 2], 1)`, IQR = `r myround(rare_median[2, 3], 1)`; Kruskal-Wallace test: chi-squared =
-`r myround(kt[1,1], 1)`, df = `r kt[1,3]`, p < 0.001; Pairwise Wilcoxon test: p < 0.001). However, there was no
+Richness of ASVs decreased between Passage 2 (median = `r myround(rare_median[1, 2], 1)`, IQR = `r myround(rare_median[1, 3], 1)`) and 5 (median = `r myround(rare_median[2, 2], 1)`, IQR = `r myround(rare_median[2, 3], 1)`; Kruskal-Wallace test: $\chi^2$ =
+`r myround(kt[1,1], 1)`, df = `r kt[1,3]`, p < 0.001; pairwise Wilcoxon test: p < 0.001). However, there was no
 difference in richness between the Positive and the Neutral treatment in
-Passage 2 or 5 (Pairwise Wilcoxon test: p = `r myround(wt[1, 3], 1)`).  In addition, there was no
+Passage 2 or 5 (pairwise Wilcoxon test: p = `r myround(wt[1, 3], 1)`).  In addition, there was no
 correlation between richness and CH~4~ oxidation rate across all passages
@@
 `r figures("beta", "Principal component analysis plot of beta diversity for all jars.
-Dissimilarities are based on Aitchison dissimilarity, which is Euclidean distance after center-log ratio transformation.
-Colors represent treatment and shapes represent passage number.
-The ratio of the axes is proportional to the variance explained by each principal component in order to accurately represent the distance between the points.")`
+Dissimilarities are based on rarefied Bray-Curtis dissimilarity.
+Colors represent treatment and shapes represent passage number.")`
 
@@
 
-`r figures("da", "Differentially abundant family-level taxa based on a beta-binomial model.
-Values on the x-axis are estimates and standard errors of the expected difference in the logit-transformed relative abundance between the two treatments.
-Positive values are enriched in the Positive treatment and negative values are enriched in the Neutral treatment.
-Taxa presented here are significant at FDR < 0.5")`
+`r figures("da", "Differentially abundant family-level taxa identified by ANCOM-II, ALDEx2, and CORNCOB.
+Values on the x-axis are relative abundances on a log~10~ scale.
+Points are colored by treatment.
+Large points with error bars are the median relative abundance and interquartile range.
+Taxa in the top panel are enriched in the Positive treatment and taxa in the bottom panel are depleted in the Positive treatment.
+Taxa are sorted by their effect size with taxa at the top having the largest positive effect size and taxa at the bottom with the largest negative effect size.")`
 

### 2022-07-06 R/02-process-herit.Rmd
start to update text and figures with new methods, results

@@
 ---
+```{r setup, include=FALSE}
 
+knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = TRUE)
+knitr::opts_knit$set(root.dir=normalizePath('../'))
+
+```
 This script takes in the flux data from script 01 and the data on which jars

### 2022-07-06 R/03-response.Rmd
start to update text and figures with new methods, results

@@
 # Compare interaction to no interaction
-compare_response_mods <- anova(no_treat_model, response_model, test = 'LRT')
-kable(tidy(compare_response_mods))
+compare_response_mods <- anova(no_treat_model, response_model)
+compare_response_mods <- tidy(compare_response_mods)
+kable(compare_response_mods)
 ```

### 2022-07-06 R/08-community-analysis.Rmd
start to update text and figures with new methods, results

--- 2022-07-05 R/09-community-analysis.Rmd
@@
 
+min_seqs <- seq_range[1]
+
 a_analysis <- 
@@
 
+```{r compare_richness_methods}
+
+aes_df_long <-
+  aes_df %>% 
+  as_tibble(rownames = "sample") %>% 
+  pivot_longer(starts_with("ASV")) %>%     
+  group_by(sample) %>% 
+  uncount(value)
+
+subsample <- function(data, sample_size) {
+  data %>% 
+    sample_n(sample_size) %>% 
+    summarize(s = n_distinct(name))
+}
+
+empirical <- map_dfr(1:10, ~subsample(aes_df_long, seq_range[1]), .id = "iters")
+empirical <-
+empirical %>% 
+  group_by(sample) %>% 
+  summarize(s = mean(s))
+
+inner_join(rare, empirical, by = "sample") %>% 
+  ggplot(aes(x = rare, y = s)) +
+  geom_abline(intercept = 0, slope = 1, color = "gray") +
+  geom_point()
+```
+
+This plot shows the correlation between Hurlbert's (1971) approach to
+richness estimation (rare) and empirically rarefying 10 times and taking
+the mean richness (s).
+
 ### Hypothesis Test
@@
   cor.test(~ ch4 + rare, data = ., method = "spearman", alternative = "two.sided")
-save(kt, wt, rare_median, s, file = "Output/richness_models.RData")
+
+min_seqs <- paste(substr(as.character(min_seqs), 1, 3), substr(as.character(min_seqs), 4, 6), sep = ",")
+save(kt, wt, rare_median, s, min_seqs, file = "Output/richness_models.RData")
 
@@
 all_iqm$Family <- factor(all_iqm$Family, levels = sort_order$Family)
-
-p <-
+all_taxa <-
+  all_taxa %>% 
+  mutate(direction = if_else(effect > 0, 0, 1))
+all_iqm <-
+all_taxa %>% 
+  select(Family, direction) %>% 
+  distinct() %>% 
+  full_join(all_iqm, by = "Family")
+(p <-
 	all_taxa %>%
@@
 			   labels = c('Neutral', 'Positive'),
-			   values = c('gray40', 'darkorange2'))
-
-p_log10 <- 
+			   values = c('gray40', 'darkorange2')) +
+    facet_grid(direction ~ ., scale = "free_y", space = "free")  + 
+theme(
+  strip.background = element_blank(),
+  strip.text.y = element_blank()
+))
+
+(p_log10 <- 
 	p +
 	scale_x_log10(breaks = c(0.001, 0.01, 0.1, 1),
-		      labels = c(0.001, 0.01, 0.1, 1))
-
-p
-p_log10
+		      labels = c(0.001, 0.01, 0.1, 1)))
 
@@
 
-ggsave('Manuscript/fig4.jpg', p_log10, width = 7, height = 7)
+ggsave('Manuscript/fig4.jpg', p_log10, width = 5, height = 5)
 
@@
 # ggsave("shared_taxa.png", p)
-ggsave("shared_taxa_log.png", p_log10, width = 7, height = 7)
+ggsave("shared_taxa_log.png", p_log10, width = 5, height = 5)
 

### 2022-07-06 R/05-create-barcode-key.R

@@ -13,8 +13,8 @@
 
 library(tidyverse)
 
-sample_key <- read_tsv('../Data/sample_key.tsv')
-primer_key <- read_tsv('../Data/primer_sequence_key.tsv')
+sample_key <- read_tsv('Data/sample_key.tsv')
+primer_key <- read_tsv('Data/primer_sequence_key.tsv')
 
 sample_key %>% 
   separate(primer_well, c('row', 'col'), 1) %>%  
@@ -22,11 +22,11 @@ sample_key %>%
   rename(r_barcode = sequence) %>% 
   left_join(primer_key, by = c('row' = 'key')) %>% 
   rename(f_barcode = sequence) %>%
-  write_tsv('../Output/barcode_master.tsv')
+  write_tsv('Output/barcode_master.tsv')
 
-read_tsv('../Output/barcode_master.tsv') %>% 
+read_tsv('Output/barcode_master.tsv') %>% 
   select(`Library Name / ID` = gc3f_id, 
          `Index 1 (i7) sequence` = r_barcode, 
          `Index 2 (i5) sequence` = f_barcode) %>% 
-  write_tsv('../Output/barcode_gc3f.tsv')
+  write_tsv('Output/barcode_gc3f.tsv')
 

### 2022-07-06 R/functions.R

@@ -49,9 +49,9 @@ print_lm <- function(fit, term = 1, estimate = 'slope') {
 }
 
 print_lrt <- function(fit, term = 2, test = "Likelihood ratio test") {
-  paste0('(', test, ': df = ', round(tidy(fit)[term, 3]), 
+  paste0(test, ': df = ', round(tidy(fit)[term, 3]), 
          ', ss = ', myround(tidy(fit)[term, 4], 2), 
-         ', p = ', myround(tidy(fit)[term, 5], 2), ')')
+         ', p = ', myround(tidy(fit)[term, 5], 2))
 }
 
 

### 2022-07-14 Manuscript/manuscript.Rmd
Add new decontam results, add dbrda results, add new discussion draft, update methods

@@
 
+load(file = '../Output/decontam.RData')
+
 # CH4 models
@@
 load(file = '../Output/da_fig.RData')
+load(file = '../Output/dbrda.RData')
 
@@
 For the neutral treatment, three jars were randomly selected to inoculate the
-next generation. For each treatment, these three jars were homogenized and 5 g 
-of this homogenized soil was used to inoculate the next set of jars and which
+next generation. For each treatment, the selected jars were homogenized and 5 g 
+of the homogenized soil was used to inoculate the next set of jars, which
 represents a 10% subsampling. The
@@
 contaminants was evaluated using both the prevalence and frequency methods from
-'DECONTAM' by comparing samples to negative controls [@davis2018]. We did not identify
-any obvious contamination using these methods.
+'DECONTAM' by comparing samples to negative controls [@davis2018].
+Decontam identified `r length(which(contamdf$contaminant))` potential contaminants based
+on prevalence and frequency. Visual inspection of abundance ~ concentration
+plots indicated that `r length(contaminants)` of these were likely contaminants and these
+ASVs were removed.
 
@@
 
-Richness of ASVs decreased between Passage 2 (median = `r myround(rare_median[1, 2], 1)`, IQR = `r myround(rare_median[1, 3], 1)`) and 5 (median = `r myround(rare_median[2, 2], 1)`, IQR = `r myround(rare_median[2, 3], 1)`; Kruskal-Wallace test: $\chi^2$ =
+Richness of ASVs decreased from `r myround(rare_median[1, 2], 1)` (`r myround(rare_median[1, 3], 1)`) in Passage 2 to `r myround(rare_median[2, 2], 1)` (`r myround(rare_median[2, 3], 1)`) in Passage 5 (Kruskal-Wallace test: $\chi^2$ =
 `r myround(kt[1,1], 1)`, df = `r kt[1,3]`, p < 0.001; pairwise Wilcoxon test: p < 0.001). However, there was no
@@
 Bray-Curtis dissimilarity of the soil microbiome varied strongly by passage and
-weakly by treatment and the interaction between passage and treatment (`r figures("beta", display = "cite")`; `r tables("tidy_permanova", display = "cite")`). 
+weakly by treatment with an interaction between passage and treatment (`r figures("beta", display = "cite")`; `r tables("adonis", display = "cite")`). 
 Passage explained `r myround(100*tidy_permanova[1, 4])`\% of the
-variation in Bray-Curtis dissimilarity.  Treatment explained 
+variation in Bray-Curtis dissimilarity, treatment explained 
 `r myround(100*tidy_permanova[2, 4])`\%
-of the variation. There was also an interaction
-between treatment and passage, which explained 
+of the variation, and the interaction
+between treatment and passage explained 
 `r myround(100*tidy_permanova[3, 4])`\% of the variation. We found there was 
 no difference in dispersion between treatments or passages (F~3,44~ = 0.91, p = 0.45).
+
+Community dissimilarity also correlated with CH~4~ oxidation rate. In Passage 2, there was no correlation between CH~4~ oxidation rate and Bray-Curtis dissimilarity 
+(dbRDA: F~`r passage2_dbrda_permtest[1, 2]`,`r passage2_dbrda_permtest[2, 2]`~ = `r myround(passage2_dbrda_permtest[1, 4], 2)`, p = `r myround(passage2_dbrda_permtest[1, 5], 2)`).
+However, in Passage 5, CH~4~ oxidation rate explained `r myround(passage5_percent_explained[1], 1)`\% of the variation in Bray-Curtis dissimilarity (dbRDA: F~`r passage5_dbrda_permtest[1, 2]`,`r passage5_dbrda_permtest[2, 2]`~ = `r myround(passage5_dbrda_permtest[1, 4], 2)`, p = `r myround(passage5_dbrda_permtest[1, 5], 2)`).
 
@@
 selection treatment [@bay2021].
+
+# New Discussion
+
+We observed a response to selection on whole-ecosystem soil methane oxidation rate, which demonstrates that variation in the microbiome generates variation in the rate of ecosystem function independent of the underlying abiotic conditions.
+While it is well understood that the microbiome mediates both host functions and environmental functions, a fundamental question in microbial ecology is whether we need to understand variation in the composition of microbiomes to understand variation in the rate of ecosystem functions.
+To address that question, we performed artificial ecosystem selection on soil microbiomes by selecting for ecosystems with high methane oxidation rates.
+We found that there was a significant increase in methane oxidation rate as a response to selection.
+Each generation of soil ecosystems was generated from the same pool of sterile soil substrate and a small inoculum from the previous generation.
+Therefore, it is unlikely that this response to selection was due to changes in the abiotic conditions of the environment.
+In addition, the response was only observed in the positive selection treatment and not in the neutral selection treatment, which shows that this was not a side effect of passaging in jars.
+Therefore, we conclude that variation in the microbiome can generate variation in ecosystem functions independent of the abiotic environmental conditions.
+
+In addition, we observed a large heritability for methane oxidation rate, which suggests that microbiome variation explains a large portion of the variation in methane oxidation rate relative to environmental variation in this system.
+To estimate how much of the variation in methane oxidation rate can be explained by the presence, absence, and relative abundances of specific microbial taxa, we calculated the narrow-sense heritability of methane oxidation rate in this experiment.
+We found that the heritability was 1.08 (95\% CI
+[`r myround(confint(herit_model)[4, ], 2)`]).
+Heritability is theoretically bounded by 0 and 1 and, in a genomic context, a heritability of 1 would indicate that a trait is entirely genetically determined and thus offspring would have the exact same value of that trait as their parents.
+A heritability greater than 1 would indicate that offspring exceed the value of their parents.
+This is unlikely in the context of a genomic trait, but microbiomes are not limited by the constraints of genomes because microbes are not alleles and communities are not chromosomes.
+Therefore, it is entirely possible that complementarity within a microbial community could lead the shuffling of microbes to generate communities that exceed the trait values of their "parent" ecosystems.
+This could be considered an "emergent property" where a particular combination of microbes can lead to ecosystem-level trait values that exceed the sum of the trait value of each individual microbe.
+In any case, the large heritability we observed indicates that microbiome variation is a large contributor to the variation in methane oxidation in this experimental system.
+
+Notably, the Neutral selection treatment displayed a negative heritability of -0.58 (0.20).
+Negative heratibility estimates are common in artificial selection experiments with a random selection treatment.
+One explanation for this phenomenon is that traits that are favored experimental conditions are negatively correlated with the trait under study.
+In the case of this experiment, methane oxidation rate could be negatively correlated with traits that favor persistence in jars.
+This tradeoff would result in a negative heritability
+
+Next we wanted to address what aspects of the microbiome, in terms of alpha diversity, beta diversity, or individual taxa, explain this contribution of the microbiome to variation in methane oxidation rate.
+There are three ways that the microbiome could respond to selection in this experiment: gain or loss of species, changes in the relative abundances of species, or changes within the genomes of the constituent species.
+This of course assumes that there is no change in the absolute abundance or biomass of the entire microbiome, but since we are unable to measure absolute microbial biomass in soil we must make the simplifying assumption that biomass is constant across the jars.
+This is not an unreasonable assumption, given that most microbes in soil are heterotrophs and that those heterotrophs would drive overall biomass in the soil microbiome whereas autotrophs involved in the cycling of nutrients and the metabolism of C1 compounds would make up a relatively small fraction of the overall biomass.
+
+Richness at the ASV level did not vary between the two treatments and there were relatively few taxa gained or lost in the Positive selection treatment and none of these were prevalent across the 12 jars in passage 5.
+Therefore, the gain or loss of species is unlikely to explain the increase in methane oxidation rate.
+There were 13 taxa that were differentially abundant between the Positive and Neutral treatments; 9 were enriched in the Positive treatment and 4 were depleted in the Positive treatment.
+Of these taxa, only one is a clade known to contain methanotrophs.
+This group did not have a Family-level taxonomic assignment, but was placed within the Gammaproteobacteria.
+The Gammaproteobacteria include the type ???
+methanotrophs in the family ???.
+However, the Gammaproteobacteria are one of the most diverse clades of bacteria and so this is not strong evidence for an increase in methanotrophs.
+We are unable to address whether taxa in this experiment could have evolved changes in their genomes as a result of selection because 16S rRNA gene sequences are unable to resolve these changes.
+
+Caveats and future directions.
+
+One caveat of this experimental design is that we intentionally eliminated as much environmental variation as possible.
+This was intentional, because we wanted to determine whether it was possible for variation in the microbiome to generate variation in methane oxidation rate independent of the environment.
+As a result, our heritability estimates are likely to be greater in this system than they would be in natural ecosystems with greater environmental variation.
+Future work could estimate heritability in experimental systems that manipulate environmental variation to evaluate how much of the variation in ecosystem function can be explained by microbiome variation outside the lab.
+This experiment also had a relatively small sample size which resulted in heritability estimates with wide error bars.
+Increasing replication would help us more accuaretly estimate this parameter.
+Lastly, using 16S rRNA gene sequencing we are unable to resolve genomic changes within taxa that may have undergone evolution.
+Future work could analyze an artificial ecosystem selection experiment using genome-resolved metagenomics to try to detect strain-level changes in the constituent taxa.
+This would help evaluate whether certain metabolic functions are gained or lost or whether existing functions changed in their efficiency.
 

### 2022-07-14 R/08-community-analysis.Rmd
Add new decontam results, add dbrda results, add new discussion draft, update methods

@@
 
+passage2_df <-
+  aes_data %>%  
+  filter(passage == "G2") %>% 
+  group_by(asv) %>% 
+  mutate(total = sum(count)) %>% 
+  filter(total != 0) %>% 
+  ungroup() %>% 
+  select(-total) %>% 
+  select(sample, asv, count) %>%
+  pivot_wider(names_from = "asv", values_from = "count", values_fill = 0) %>% 
+  as.data.frame()
+
+rownames(passage2_df) <- passage2_df$sample
+passage2_df <- passage2_df[, -1]
+
 passage5_df <-
@@
 
-```{r passage-dbrda}
-if(file.exists("Output/passage_rare_dist_bray.rds")) {
-  passage_rare_dist_bray <- 
-    readRDS("Output/passage_rare_dist_bray.rds")
+```{r passage2-dbrda}
+if(file.exists("Output/passage2_rare_dist_bray.rds")) {
+  passage2_rare_dist_bray <- 
+    readRDS("Output/passage2_rare_dist_bray.rds")
 } else {
-passage_rare_dist_bray <- avgdist(passage5_df, dmethod="bray", sample=seq_range[1])
-  passage_rare_dist_bray %>% 
-    saveRDS("Output/passage_rare_dist_bray.rds")
+passage2_rare_dist_bray <- avgdist(passage2_df, dmethod="bray", sample=seq_range[1])
+  passage2_rare_dist_bray %>% 
+    saveRDS("Output/passage2_rare_dist_bray.rds")
 }
 
-passage_dist_df <-
-  passage_rare_dist_bray %>% 
+passage2_dist_df <-
+  passage2_rare_dist_bray %>% 
   as.matrix %>% 
@@
 
-passage_meta_dist <-
+passage2_meta_dist <-
   aes_data %>% 
@@
   distinct %>% 
-  inner_join(passage_dist_df, by = "sample")
-
-passage_dbrda_ch4 <- dbrda(passage_rare_dist_bray ~ rank(ch4), 
-                      data = passage_meta_dist, 
+  inner_join(passage2_dist_df, by = "sample")
+
+passage2_dbrda_ch4 <- dbrda(passage2_rare_dist_bray ~ rank(ch4), 
+                      data = passage2_meta_dist, 
                       na.action = "na.omit",
@@
 
-passage_dbrda_ch4 %>% 
+passage2_dbrda_permtest <- 
+passage2_dbrda_ch4 %>% 
   permutest(permutations = 9999)
 
-passage_percent_explained <- 100 * eigenvals(passage_dbrda_ch4) / sum(eigenvals(passage_dbrda_ch4))
-
-passage_dbrda_scores <-
-  scores(passage_dbrda_ch4, display = "sites") %>% 
+passage2_dbrda_permtest <- 
+passage2_dbrda_ch4 %>% 
+  anova(permutations = 9999) %>% 
+  tidy()
+
+passage2_percent_explained <- 100 * eigenvals(passage2_dbrda_ch4) / sum(eigenvals(passage2_dbrda_ch4))
+
+passage2_dbrda_scores <-
+  scores(passage2_dbrda_ch4, display = "sites") %>% 
   as_tibble(rownames = "sample")
 
-passage_dbrda_data <-
+passage2_dbrda_data <-
   aes_data %>% 
@@
   distinct %>% 
-  inner_join(passage_dbrda_scores, by = "sample")
-
-passage_dbrda_data %>%   
+  inner_join(passage2_dbrda_scores, by = "sample")
+
+passage2_dbrda_data %>%   
   ggplot(aes(x = dbRDA1, y = MDS1, shape = passage, color = treat)) +
@@
 
+```{r passage5-dbrda}
+if(file.exists("Output/passage5_rare_dist_bray.rds")) {
+  passage5_rare_dist_bray <- 
+    readRDS("Output/passage5_rare_dist_bray.rds")
+} else {
+passage5_rare_dist_bray <- avgdist(passage5_df, dmethod="bray", sample=seq_range[1])
+  passage5_rare_dist_bray %>% 
+    saveRDS("Output/passage5_rare_dist_bray.rds")
+}
+
+passage5_dist_df <-
+  passage5_rare_dist_bray %>% 
+  as.matrix %>% 
+  as_tibble(rownames = "sample")
+
+passage5_meta_dist <-
+  aes_data %>% 
+  select(sample, passage_treat, passage, treat, ch4) %>% 
+  distinct %>% 
+  inner_join(passage5_dist_df, by = "sample")
+
+passage5_dbrda_ch4 <- dbrda(passage5_rare_dist_bray ~ rank(ch4), 
+                      data = passage5_meta_dist, 
+                      na.action = "na.omit",
+                      add = TRUE)
+
+passage5_dbrda_permtest <-
+passage5_dbrda_ch4 %>% 
+  anova(permutations = 9999) %>% 
+  tidy
+
+passage5_percent_explained <- 100 * eigenvals(passage5_dbrda_ch4) / sum(eigenvals(passage5_dbrda_ch4))
+
+passage5_dbrda_scores <-
+  scores(passage5_dbrda_ch4, display = "sites") %>% 
+  as_tibble(rownames = "sample")
+
+passage5_dbrda_data <-
+  aes_data %>% 
+  select(sample, passage, treat, ch4) %>% 
+  distinct %>% 
+  inner_join(passage5_dbrda_scores, by = "sample")
+
+passage5_dbrda_data %>%   
+  ggplot(aes(x = dbRDA1, y = MDS1, shape = passage, color = treat)) +
+  geom_point() +
+  stat_ellipse(show.legend = FALSE) +
+  coord_fixed() +
+  scale_shape_discrete(name = "Passage") +
+  scale_color_manual(name = "Treatment",
+                       labels = c('Neutral', 'Positive'),
+                       values = c('gray40', 'darkorange2')) +
+  theme_bw() +
+  theme(
+    plot.background = element_blank(),
+    panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank())
+
+```
+
+```{r}
+save(passage2_dbrda_permtest, 
+     passage5_dbrda_permtest, 
+     passage5_percent_explained,
+     file = "Output/dbrda.RData")
+```
+
 The overall dbRDA is significant and rank of methane explains 5% of the variation.
@@
 ## Unique to each treatment
+
+### Family
 
@@
 
+### Genus
+
+```{r}
+g5n <- 
+  aes_data %>% 
+  filter(passage_treat == "G5N") %>% 
+  select(Genus) %>% 
+  distinct() %$%
+  Genus
+
+g5p <- 
+  aes_data %>% 
+  filter(passage_treat == "G5P") %>% 
+  select(Genus) %>% 
+  distinct() %$%
+  Genus
+
+p_not_n <- g5p[!g5p %in% g5n]
+
+n_not_p <- g5n[!g5n %in% g5p]
+
+n_not_p_prevalence <-
+  aes_data %>% 
+  filter(passage_treat == "G5N") %>% 
+  filter(Genus %in% n_not_p) %>% 
+  select(sample, Genus, count) %>% 
+  mutate(prevalence = as.numeric(count > 0)) %>% 
+  select(-count) %>% 
+  group_by(Genus) %>% 
+  summarize(prevalence = sum(prevalence))
+
+n_not_p_prevalence %>% 
+  ggplot(aes(x = prevalence)) +
+  geom_histogram() +
+  scale_x_continuous(breaks = 1:12) +
+  theme_classic() +
+  labs(title = "Unique to Positive")
+
+p_not_n_prevalence <-
+  aes_data %>% 
+  filter(passage_treat == "G5P") %>% 
+  filter(Genus %in% p_not_n) %>% 
+  select(sample, Genus, count) %>% 
+  mutate(prevalence = as.numeric(count > 0)) %>% 
+  select(-count) %>% 
+  group_by(Genus) %>% 
+  summarize(prevalence = sum(prevalence))
+
+p_not_n_prevalence %>% 
+  ggplot(aes(x = prevalence)) +
+  geom_histogram() +
+  scale_x_continuous(breaks = 1:12) +
+  theme_classic() +
+  labs(title = "Unique to Neutral")
+
+p_not_n_over_6 <- p_not_n_prevalence %>%
+  filter(prevalence > 6)
+
+n_not_p_over_6 <- n_not_p_prevalence %>%
+  filter(prevalence > 6)
+
+```
+
+### Species
+
+```{r}
+g5n <- 
+  aes_data %>% 
+  filter(passage_treat == "G5N") %>% 
+  select(Species) %>% 
+  distinct() %$%
+  Species
+
+g5p <- 
+  aes_data %>% 
+  filter(passage_treat == "G5P") %>% 
+  select(Species) %>% 
+  distinct() %$%
+  Species
+
+p_not_n <- g5p[!g5p %in% g5n]
+
+n_not_p <- g5n[!g5n %in% g5p]
+
+n_not_p_prevalence <-
+  aes_data %>% 
+  filter(passage_treat == "G5N") %>% 
+  filter(Species %in% n_not_p) %>% 
+  select(sample, Species, count) %>% 
+  mutate(prevalence = as.numeric(count > 0)) %>% 
+  select(-count) %>% 
+  group_by(Species) %>% 
+  summarize(prevalence = sum(prevalence))
+
+n_not_p_prevalence %>% 
+  ggplot(aes(x = prevalence)) +
+  geom_histogram() +
+  scale_x_continuous(breaks = 1:12) +
+  theme_classic() +
+  labs(title = "Unique to Positive")
+
+p_not_n_prevalence <-
+  aes_data %>% 
+  filter(passage_treat == "G5P") %>% 
+  filter(Species %in% p_not_n) %>% 
+  select(sample, Species, count) %>% 
+  mutate(prevalence = as.numeric(count > 0)) %>% 
+  select(-count) %>% 
+  group_by(Species) %>% 
+  summarize(prevalence = sum(prevalence))
+
+p_not_n_prevalence %>% 
+  ggplot(aes(x = prevalence)) +
+  geom_histogram() +
+  scale_x_continuous(breaks = 1:12) +
+  theme_classic() +
+  labs(title = "Unique to Neutral")
+
+p_not_n_over_6 <- p_not_n_prevalence %>%
+  filter(prevalence > 3)
+
+n_not_p_over_6 <- n_not_p_prevalence %>%
+  filter(prevalence > 3)
+
+```
+
 ## Setup Differential Abundance Data
 
-```{r setup-differential-abundance-data}
+### Family
+
+```{r setup-differential-abundance-data-family}
 
@@
 
-## Corncob
+#### Corncob
 
@@
 
-## ANCOM
+#### ANCOM
 
@@
 
-## Aldex2
+#### Aldex2
 
@@
 
-## Comparison
-
-### All Significant
+#### Comparison
+
+##### All Significant
 
@@
 
-### Consensus Taxa
+##### Consensus Taxa
 
@@
 
+```{r}
+
+knitr::opts_chunk$set(eval = FALSE)
+
+```
+
+### Genus
+
+```{r setup-differential-abundance-data-genus}
+
+aes_G5 <-
+  aes_data %>% 
+  filter(passage == "G5")
+
+aes_Genus_G5 <-
+  aes_G5 %>% 
+  select(asv, sample, treat, count, Genus) %>%
+  group_by(sample, Genus) %>% 
+  mutate(total = sum(count)) %>% 
+  ungroup() %>% 
+  select(sample, treat, Genus, total) %>% 
+  distinct()
+
+prevalent_families <- 
+  aes_Genus_G5 %>% 
+  group_by(Genus, treat) %>% 
+  summarize(prev = sum(total > 0), .groups = "drop") %>% 
+  pivot_wider(names_from = "treat", values_from = "prev", values_fill = 0) %>% 
+  filter(N/12 > 0.1 & P/12 > 0.1) %$%
+  Genus
+
+otu_table <- 
+  aes_Genus_G5 %>% 
+  filter(Genus %in% prevalent_families) %>% 
+  select(sample, Genus, total) %>% 
+  pivot_wider(names_from = sample, values_from = total, values_fill = 0) %>% 
+  as.data.frame()
+
+rownames(otu_table) <- otu_table[,1]
+otu_table <- otu_table[,-1]
+
+sample_data <- aes_Genus_G5 %>% 
+  select(sample, treat) %>% 
+  distinct() %>% 
+  as.data.frame()
+
+rownames(sample_data) <- sample_data[,1]
+
+```
+
+#### Corncob
+
+```{r}
+
+da_analysis <- differentialTest(formula = ~ treat,
+                                phi.formula = ~ 1,
+                                formula_null = ~ 1,
+                                phi.formula_null = ~ 1,
+                                test = "Wald", 
+                                boot = FALSE,
+                                data = otu_table,
+                                sample_data = sample_data,
+                                fdr_cutoff = 0.05, B = 100,
+                                taxa_are_rows = TRUE)
+
+dt_data <- plot(da_analysis, 
+                # level = c("Genus"), 
+                data_only = TRUE)
+
+dt_plot <- function(dt_data) {
+  ggplot(dt_data, aes(x = x, xmin = xmin, xmax = xmax, y = reorder(taxa, x))) +
+    geom_pointrange() +
+    geom_vline(xintercept = 0) +
+    labs(x = "Estimate", y = NULL) +
+    coord_cartesian(clip="off") +
+    # theme(panel.border = element_blank()) +
+    theme_classic() +
+      theme(axis.text.y = element_markdown())
+
+}
+
+dt_data %>%
+  mutate(taxa = str_replace(taxa, "(.*)", "*\\1*"),
+         taxa = str_replace(taxa, "\\*(.*)_unclassified\\*",
+                                "Unclassified *\\1*")) %>% 
+  dt_plot()
+
+```
+
+#### ANCOM
+
+```{r}
+
+# Setup phyloseq object
+aes_ps <- phyloseq(otu_table(otu_table, taxa_are_rows=TRUE), sample_data(sample_data))
+
+if(file.exists("Output/ancom_out.rds")) {
+  out <- 
+    readRDS("Output/ancom_out.rds")
+} else {
+  # Run Ancom
+  out = ancom(phyloseq = aes_ps,  
+              p_adj_method = "BH",
+              prv_cut = 0.1, 
+              lib_cut = 0, 
+              main_var = "treat",
+              rand_formula = NULL, 
+              lme_control = NULL,
+              struc_zero = TRUE, 
+              neg_lb = FALSE, 
+              alpha = 0.05, 
+              n_cl = 1)  
+  out %>% 
+    saveRDS("Output/ancom_out.rds")
+}
+
+res = out$res
+q_val = out$q_data
+beta_val = out$beta_data
+# Only consider the effect sizes with the corresponding q-value less than alpha
+beta_val = beta_val * (q_val < 0.05) 
+# Choose the maximum of beta's as the effect size
+beta_pos = apply(abs(beta_val), 2, which.max) 
+beta_max = vapply(seq_along(beta_pos), function(i) beta_val[beta_pos[i], i],
+                  FUN.VALUE = double(1))
+# Number of taxa except structural zeros
+n_taxa = ifelse(is.null(out$zero_ind), 
+                nrow(feature_table), 
+                sum(apply(out$zero_ind, 1, sum) == 0))
+# Cutoff values for declaring differentially abundant taxa
+cut_off = 0.7 * (n_taxa - 1)
+
+df_fig_w = res %>%
+  dplyr::mutate(beta = beta_max,
+                direct = case_when(
+                  detected_0.7 == TRUE & beta > 0 ~ "Positive",
+                  detected_0.7 == TRUE & beta <= 0 ~ "Negative",
+                  TRUE ~ "Not Significant"
+                  )) %>%
+  dplyr::arrange(W)
+df_fig_w$taxon_id = factor(df_fig_w$taxon_id, levels = df_fig_w$taxon_id)
+df_fig_w$W = replace(df_fig_w$W, is.infinite(df_fig_w$W), n_taxa - 1)
+df_fig_w$direct = factor(df_fig_w$direct, 
+                     levels = c("Negative", "Positive", "Not Significant"))
+
+p_w = df_fig_w %>%
+  ggplot(aes(x = taxon_id, y = W, color = direct)) +
+  geom_point(size = 2, alpha = 0.6) +
+  labs(x = "Taxon", y = "W") +
+  scale_color_discrete(name = NULL) + 
+  geom_hline(yintercept = cut_off, linetype = "dotted", 
+             color = "blue", size = 1.5) +
+  geom_text(aes(x = 2, y = cut_off + 0.5, label = "W[0.7]"), 
+            size = 5, vjust = -0.5, hjust = 0, color = "orange", parse = TRUE) +
+  theme_bw() +
+  theme(axis.text.x = element_blank(),
+        axis.ticks.x = element_blank(),
+        panel.grid.major = element_blank())
+p_w
+
+```
+
+#### Aldex2
+
+```{r}
+
+conds <- colnames(otu_table) %>% 
+  substr(3, 3)
+
+if(file.exists("Output/aldex_x_all.rds")) {
+  x.all <- 
+    readRDS("Output/aldex_x_all.rds")
+} else {
+  x.all <- aldex(otu_table, 
+                 conds, mc.samples=1000, test="t", effect=TRUE,
+                 include.sample.summary=FALSE, denom="all", verbose=FALSE, paired.test=FALSE)
+  x.all %>% 
+    saveRDS("Output/aldex_x_all.rds")
+}
+
+par(mfrow=c(1,2))
+aldex.plot(x.all, type="MA", test="welch", xlab="Log-ratio abundance",
+    ylab="Difference")
+aldex.plot(x.all, type="MW", test="welch", xlab="Dispersion",
+    ylab="Difference")
+
+par(mfrow=c(1,2))
+plot(x.all$effect, x.all$we.ep, log="y", cex=0.7, col=rgb(0,0,1,0.2),
+  pch=19, xlab="Effect size", ylab="P value", main="Effect size plot")
+points(x.all$effect, x.all$we.eBH, cex=0.7, col=rgb(1,0,0,0.2),
+  pch=19)
+abline(h=0.05, lty=2, col="grey")
+legend(15,1, legend=c("P value", "BH-adjusted"), pch=19, col=c("blue", "red"))
+
+plot(x.all$diff.btw, x.all$we.ep, log="y", cex=0.7, col=rgb(0,0,1,0.2),
+  pch=19, xlab="Difference", ylab="P value", main="Volcano plot")
+points(x.all$diff.btw, x.all$we.eBH, cex=0.7, col=rgb(1,0,0,0.2),
+  pch=19)
+abline(h=0.05, lty=2, col="grey")
+
+x.all_effect_1 <-
+  x.all %>% 
+  filter(abs(effect) > 1) %>% 
+  as_tibble(rownames = "Genus")
+
+# x.all_effect_1 %>% 
+#   # inner_join(taxa_table, by = "asv")
+#   ggplot(aes(y = reorder(Genus, effect), x = effect)) +
+#   geom_point()
+
+# x.all_effect_1 %>% 
+#   select(Genus, effect)
+
+```
+
+#### Comparison
+
+##### All Significant
+
+```{r}
+
+# CORNCOB
+corncob_taxa <- dt_data$taxa
+
+# ANCOM
+ancom_taxa <- as.character(df_fig_w[df_fig_w$detected_0.7, ]$taxon_id)
+
+# ALDEX2
+aldex_taxa <- x.all_effect_1$Genus
+
+at_least_one_taxa <-
+	aes_Genus_G5 %>%
+	group_by(sample) %>%
+	mutate(rel_abund = (total / sum(total)) * 100) %>%
+	filter(Genus %in% ancom_taxa | Genus %in% aldex_taxa | Genus %in% corncob_taxa) %>%
+	mutate(Genus = str_replace(Genus, "(.*)", "*\\1*"),
+	       Genus = str_replace(Genus, "\\*(.*)_unclassified\\*",
+				    "Unclassified *\\1*"))
+
+at_least_one_iqm <- 
+	at_least_one_taxa %>%
+	group_by(treat, Genus) %>%
+	summarize(median = median(rel_abund), 
+		  first_quartile = quart(rel_abund)[1],
+		  third_quartile = quart(rel_abund)[2], .groups = "drop")
+
+sort_order <-
+	at_least_one_taxa %>%
+	group_by(Genus) %>%
+	summarize(median = median(rel_abund)) %>%
+	arrange(median)
+
+at_least_one_taxa$Genus <- factor(at_least_one_taxa$Genus, levels = sort_order$Genus)
+
+at_least_one_taxa %>%
+	ggplot(aes(x = rel_abund, y = Genus, color = treat)) +
+	geom_point(position=position_jitterdodge()) +
+	geom_pointrange(aes(x = median, xmax = third_quartile, xmin = first_quartile, fill = treat), 
+			data = at_least_one_iqm, 
+			position = position_dodge(width = 0.75), 
+			shape = 21, 
+			color = "black") +
+	theme_bw() +
+	theme(axis.text.y = element_markdown()) +
+	labs(x = "Relative Abundance (%)", y = NULL) +
+	scale_color_manual(name = "Treatment",
+			   labels = c('Neutral', 'Positive'),
+			   values = c('gray40', 'darkorange2')) +
+	scale_fill_manual(name = "Treatment",
+			   labels = c('Neutral', 'Positive'),
+			   values = c('gray40', 'darkorange2'))
+
+```
+
+##### Consensus Taxa
+
+```{r}
+aldex_effect <-
+  x.all_effect_1 %>% 
+  select(Genus, effect)
+
+all_taxa <-
+	aes_Genus_G5 %>%
+	group_by(sample) %>%
+	mutate(rel_abund = (total / sum(total)) * 100) %>%
+  left_join(aldex_effect, by = "Genus") %>% 
+	filter(Genus %in% ancom_taxa & Genus %in% aldex_taxa & Genus %in% corncob_taxa) %>% 
+  ungroup()
+
+da_fams <-
+  all_taxa %>% 
+  select(Genus) %>% 
+  distinct() %$%
+  Genus
+
+all_taxa <-
+  all_taxa %>%
+	mutate(Genus = str_replace(Genus, "(.*)", "*\\1*"),
+	       Genus = str_replace(Genus, "\\*(.*)_unclassified\\*",
+				    "Unclassified *\\1*"))
+all_taxa$Genus[all_taxa$Genus == "Unclassified *0319-6G20*"] <- "Uncultured Genus 0319-6G20"
+all_taxa$Genus[all_taxa$Genus == "Unclassified *Chloroplast*"] <- "Chloroplast"
+all_taxa$Genus[all_taxa$Genus == "Unclassified *Armatimonadales*"] <- "Unclassified Armatimonadales"
+all_taxa$Genus[all_taxa$Genus == "Unclassified *Gammaproteobacteria*"] <- "Unclassified Gammaproteobacteria"
+all_taxa$Genus[all_taxa$Genus == "Unclassified *Kapabacteriales*"] <- "Unclassified Kapabacteriales"
+
+all_iqm <- 
+	all_taxa %>%
+	group_by(treat, Genus) %>%
+	summarize(median = median(rel_abund), 
+		  first_quartile = quart(rel_abund)[1],
+		  third_quartile = quart(rel_abund)[2], .groups = "drop")
+
+sort_order <-
+	all_taxa %>%
+  select(Genus, effect) %>% 
+  distinct() %>% 
+  arrange(effect)
+	# group_by(Genus) %>%
+	# summarize(median = median(rel_abund)) %>%
+	# arrange(median)
+
+all_taxa$Genus <- factor(all_taxa$Genus, levels = sort_order$Genus)
+all_iqm$Genus <- factor(all_iqm$Genus, levels = sort_order$Genus)
+all_taxa <-
+  all_taxa %>% 
+  mutate(direction = if_else(effect > 0, 0, 1))
+all_iqm <-
+all_taxa %>% 
+  select(Genus, direction) %>% 
+  distinct() %>% 
+  full_join(all_iqm, by = "Genus")
+(p <-
+	all_taxa %>%
+	ggplot(aes(x = rel_abund, y = Genus, color = treat)) +
+	geom_point(position=position_jitterdodge()) +
+	geom_pointrange(aes(x = median, xmax = third_quartile, xmin = first_quartile, fill = treat), 
+			data = all_iqm, 
+			position = position_dodge(width = 0.75), 
+			shape = 21, 
+			color = "black") +
+	theme_bw() +
+	theme(axis.text.y = element_markdown()) +
+	labs(x = "Relative Abundance (%)", y = NULL) +
+	scale_color_manual(name = "Treatment",
+			   labels = c('Neutral', 'Positive'),
+			   values = c('gray40', 'darkorange2')) +
+	scale_fill_manual(name = "Treatment",
+			   labels = c('Neutral', 'Positive'),
+			   values = c('gray40', 'darkorange2')) +
+    facet_grid(direction ~ ., scale = "free_y", space = "free")  + 
+theme(
+  strip.background = element_blank(),
+  strip.text.y = element_blank()
+))
+
+(p_log10 <- 
+	p +
+	scale_x_log10(breaks = c(0.001, 0.01, 0.1, 1),
+		      labels = c(0.001, 0.01, 0.1, 1)))
+
+save(p_log10, da_fams, file = "Output/da_fig.RData")
+
+ggsave('Manuscript/fig4.jpg', p_log10, width = 5, height = 5)
+
+## All
+# ggsave("shared_taxa.png", p)
+ggsave("shared_taxa_log.png", p_log10, width = 5, height = 5)
+
+# aes_Genus_rclr_plot <-
+#   aes_Genus_rclr %>% 
+# 	filter(Genus %in% ancom_taxa & Genus %in% aldex_taxa & Genus %in% corncob_taxa) %>% 
+#   pivot_longer(G2:G5, names_to = "passage", values_to = "rclr") %>% 
+#   select(Genus, treat, rep, passage, rclr) %>% 
+#   distinct() %>% 	
+#   mutate(Genus = str_replace(Genus, "(.*)", "*\\1*"),
+# 	       Genus = str_replace(Genus, "\\*(.*)_unclassified\\*",
+# 				    "Unclassified *\\1*"))
+# 
+# aes_Genus_rclr_plot$Genus[aes_Genus_rclr_plot$Genus == "Unclassified *0319-6G20*"] <- "Uncultured Genus 0319-6G20"
+# aes_Genus_rclr_plot$Genus[aes_Genus_rclr_plot$Genus == "Unclassified *Chloroplast*"] <- "Chloroplast"
+# aes_Genus_rclr_plot$Genus <- factor(aes_Genus_rclr_plot$Genus, levels = sort_order$Genus)
+# ggplot(aes_Genus_rclr_plot, aes(x = rclr, y = Genus, color = treat)) +
+#   facet_grid( ~ passage) +
+#   geom_point(position = position_jitterdodge())
+
+p_effect <-
+	all_taxa %>%
+	ggplot(aes(x = effect, y = Genus)) +
+	geom_point() +
+
+	theme_bw() +
+	theme(axis.text.y = element_markdown()) +
+	labs(x = "Aldex2 Effect Size", y = NULL) +
+	scale_color_manual(name = "Treatment",
+			   labels = c('Neutral', 'Positive'),
+			   values = c('gray40', 'darkorange2')) +
+	scale_fill_manual(name = "Treatment",
+			   labels = c('Neutral', 'Positive'),
+			   values = c('gray40', 'darkorange2'))
+p_effect
+
+```
+
+### Species
+
+```{r setup-differential-abundance-data-species}
+
+aes_G5 <-
+  aes_data %>% 
+  filter(passage == "G5")
+
+aes_Species_G5 <-
+  aes_G5 %>% 
+  select(asv, sample, treat, count, Species) %>%
+  group_by(sample, Species) %>% 
+  mutate(total = sum(count)) %>% 
+  ungroup() %>% 
+  select(sample, treat, Species, total) %>% 
+  distinct()
+
+prevalent_families <- 
+  aes_Species_G5 %>% 
+  group_by(Species, treat) %>% 
+  summarize(prev = sum(total > 0), .groups = "drop") %>% 
+  pivot_wider(names_from = "treat", values_from = "prev", values_fill = 0) %>% 
+  filter(N/12 > 0.1 & P/12 > 0.1) %$%
+  Species
+
+otu_table <- 
+  aes_Species_G5 %>% 
+  filter(Species %in% prevalent_families) %>% 
+  select(sample, Species, total) %>% 
+  pivot_wider(names_from = sample, values_from = total, values_fill = 0) %>% 
+  as.data.frame()
+
+rownames(otu_table) <- otu_table[,1]
+otu_table <- otu_table[,-1]
+
+sample_data <- aes_Species_G5 %>% 
+  select(sample, treat) %>% 
+  distinct() %>% 
+  as.data.frame()
+
+rownames(sample_data) <- sample_data[,1]
+
+```
+
+#### Corncob
+
+```{r}
+
+da_analysis <- differentialTest(formula = ~ treat,
+                                phi.formula = ~ 1,
+                                formula_null = ~ 1,
+                                phi.formula_null = ~ 1,
+                                test = "Wald", 
+                                boot = FALSE,
+                                data = otu_table,
+                                sample_data = sample_data,
+                                fdr_cutoff = 0.05, B = 100,
+                                taxa_are_rows = TRUE)
+
+dt_data <- plot(da_analysis, 
+                # level = c("Species"), 
+                data_only = TRUE)
+
+dt_plot <- function(dt_data) {
+  ggplot(dt_data, aes(x = x, xmin = xmin, xmax = xmax, y = reorder(taxa, x))) +
+    geom_pointrange() +
+    geom_vline(xintercept = 0) +
+    labs(x = "Estimate", y = NULL) +
+    coord_cartesian(clip="off") +
+    # theme(panel.border = element_blank()) +
+    theme_classic() +
+      theme(axis.text.y = element_markdown())
+
+}
+
+dt_data %>%
+  mutate(taxa = str_replace(taxa, "(.*)", "*\\1*"),
+         taxa = str_replace(taxa, "\\*(.*)_unclassified\\*",
+                                "Unclassified *\\1*")) %>% 
+  dt_plot()
+
+```
+
+#### ANCOM
+
+```{r}
+
+# Setup phyloseq object
+aes_ps <- phyloseq(otu_table(otu_table, taxa_are_rows=TRUE), sample_data(sample_data))
+
+if(file.exists("Output/ancom_out.rds")) {
+  out <- 
+    readRDS("Output/ancom_out.rds")
+} else {
+  # Run Ancom
+  out = ancom(phyloseq = aes_ps,  
+              p_adj_method = "BH",
+              prv_cut = 0.1, 
+              lib_cut = 0, 
+              main_var = "treat",
+              rand_formula = NULL, 
+              lme_control = NULL,
+              struc_zero = TRUE, 
+              neg_lb = FALSE, 
+              alpha = 0.05, 
+              n_cl = 1)  
+  out %>% 
+    saveRDS("Output/ancom_out.rds")
+}
+
+res = out$res
+q_val = out$q_data
+beta_val = out$beta_data
+# Only consider the effect sizes with the corresponding q-value less than alpha
+beta_val = beta_val * (q_val < 0.05) 
+# Choose the maximum of beta's as the effect size
+beta_pos = apply(abs(beta_val), 2, which.max) 
+beta_max = vapply(seq_along(beta_pos), function(i) beta_val[beta_pos[i], i],
+                  FUN.VALUE = double(1))
+# Number of taxa except structural zeros
+n_taxa = ifelse(is.null(out$zero_ind), 
+                nrow(feature_table), 
+                sum(apply(out$zero_ind, 1, sum) == 0))
+# Cutoff values for declaring differentially abundant taxa
+cut_off = 0.7 * (n_taxa - 1)
+
+df_fig_w = res %>%
+  dplyr::mutate(beta = beta_max,
+                direct = case_when(
+                  detected_0.7 == TRUE & beta > 0 ~ "Positive",
+                  detected_0.7 == TRUE & beta <= 0 ~ "Negative",
+                  TRUE ~ "Not Significant"
+                  )) %>%
+  dplyr::arrange(W)
+df_fig_w$taxon_id = factor(df_fig_w$taxon_id, levels = df_fig_w$taxon_id)
+df_fig_w$W = replace(df_fig_w$W, is.infinite(df_fig_w$W), n_taxa - 1)
+df_fig_w$direct = factor(df_fig_w$direct, 
+                     levels = c("Negative", "Positive", "Not Significant"))
+
+p_w = df_fig_w %>%
+  ggplot(aes(x = taxon_id, y = W, color = direct)) +
+  geom_point(size = 2, alpha = 0.6) +
+  labs(x = "Taxon", y = "W") +
+  scale_color_discrete(name = NULL) + 
+  geom_hline(yintercept = cut_off, linetype = "dotted", 
+             color = "blue", size = 1.5) +
+  geom_text(aes(x = 2, y = cut_off + 0.5, label = "W[0.7]"), 
+            size = 5, vjust = -0.5, hjust = 0, color = "orange", parse = TRUE) +
+  theme_bw() +
+  theme(axis.text.x = element_blank(),
+        axis.ticks.x = element_blank(),
+        panel.grid.major = element_blank())
+p_w
+
+```
+
+#### Aldex2
+
+```{r}
+
+conds <- colnames(otu_table) %>% 
+  substr(3, 3)
+
+if(file.exists("Output/aldex_x_all.rds")) {
+  x.all <- 
+    readRDS("Output/aldex_x_all.rds")
+} else {
+  x.all <- aldex(otu_table, 
+                 conds, mc.samples=1000, test="t", effect=TRUE,
+                 include.sample.summary=FALSE, denom="all", verbose=FALSE, paired.test=FALSE)
+  x.all %>% 
+    saveRDS("Output/aldex_x_all.rds")
+}
+
+par(mfrow=c(1,2))
+aldex.plot(x.all, type="MA", test="welch", xlab="Log-ratio abundance",
+    ylab="Difference")
+aldex.plot(x.all, type="MW", test="welch", xlab="Dispersion",
+    ylab="Difference")
+
+par(mfrow=c(1,2))
+plot(x.all$effect, x.all$we.ep, log="y", cex=0.7, col=rgb(0,0,1,0.2),
+  pch=19, xlab="Effect size", ylab="P value", main="Effect size plot")
+points(x.all$effect, x.all$we.eBH, cex=0.7, col=rgb(1,0,0,0.2),
+  pch=19)
+abline(h=0.05, lty=2, col="grey")
+legend(15,1, legend=c("P value", "BH-adjusted"), pch=19, col=c("blue", "red"))
+
+plot(x.all$diff.btw, x.all$we.ep, log="y", cex=0.7, col=rgb(0,0,1,0.2),
+  pch=19, xlab="Difference", ylab="P value", main="Volcano plot")
+points(x.all$diff.btw, x.all$we.eBH, cex=0.7, col=rgb(1,0,0,0.2),
+  pch=19)
+abline(h=0.05, lty=2, col="grey")
+
+x.all_effect_1 <-
+  x.all %>% 
+  filter(abs(effect) > 1) %>% 
+  as_tibble(rownames = "Species")
+
+# x.all_effect_1 %>% 
+#   # inner_join(taxa_table, by = "asv")
+#   ggplot(aes(y = reorder(Species, effect), x = effect)) +
+#   geom_point()
+
+# x.all_effect_1 %>% 
+#   select(Species, effect)
+
+```
+
+#### Comparison
+
+##### All Significant
+
+```{r}
+
+# CORNCOB
+corncob_taxa <- dt_data$taxa
+
+# ANCOM
+ancom_taxa <- as.character(df_fig_w[df_fig_w$detected_0.7, ]$taxon_id)
+
+# ALDEX2
+aldex_taxa <- x.all_effect_1$Species
+
+at_least_one_taxa <-
+	aes_Species_G5 %>%
+	group_by(sample) %>%
+	mutate(rel_abund = (total / sum(total)) * 100) %>%
+	filter(Species %in% ancom_taxa | Species %in% aldex_taxa | Species %in% corncob_taxa) %>%
+	mutate(Species = str_replace(Species, "(.*)", "*\\1*"),
+	       Species = str_replace(Species, "\\*(.*)_unclassified\\*",
+				    "Unclassified *\\1*"))
+
+at_least_one_iqm <- 
+	at_least_one_taxa %>%
+	group_by(treat, Species) %>%
+	summarize(median = median(rel_abund), 
+		  first_quartile = quart(rel_abund)[1],
+		  third_quartile = quart(rel_abund)[2], .groups = "drop")
+
+sort_order <-
+	at_least_one_taxa %>%
+	group_by(Species) %>%
+	summarize(median = median(rel_abund)) %>%
+	arrange(median)
+
+at_least_one_taxa$Species <- factor(at_least_one_taxa$Species, levels = sort_order$Species)
+
+at_least_one_taxa %>%
+	ggplot(aes(x = rel_abund, y = Species, color = treat)) +
+	geom_point(position=position_jitterdodge()) +
+	geom_pointrange(aes(x = median, xmax = third_quartile, xmin = first_quartile, fill = treat), 
+			data = at_least_one_iqm, 
+			position = position_dodge(width = 0.75), 
+			shape = 21, 
+			color = "black") +
+	theme_bw() +
+	theme(axis.text.y = element_markdown()) +
+	labs(x = "Relative Abundance (%)", y = NULL) +
+	scale_color_manual(name = "Treatment",
+			   labels = c('Neutral', 'Positive'),
+			   values = c('gray40', 'darkorange2')) +
+	scale_fill_manual(name = "Treatment",
+			   labels = c('Neutral', 'Positive'),
+			   values = c('gray40', 'darkorange2'))
+
+```
+
+##### Consensus Taxa
+
+```{r}
+aldex_effect <-
+  x.all_effect_1 %>% 
+  select(Species, effect)
+
+all_taxa <-
+	aes_Species_G5 %>%
+	group_by(sample) %>%
+	mutate(rel_abund = (total / sum(total)) * 100) %>%
+  left_join(aldex_effect, by = "Species") %>% 
+	filter(Species %in% ancom_taxa & Species %in% aldex_taxa & Species %in% corncob_taxa) %>% 
+  ungroup()
+
+da_fams <-
+  all_taxa %>% 
+  select(Species) %>% 
+  distinct() %$%
+  Species
+
+all_taxa <-
+  all_taxa %>%
+	mutate(Species = str_replace(Species, "(.*)", "*\\1*"),
+	       Species = str_replace(Species, "\\*(.*)_unclassified\\*",
+				    "Unclassified *\\1*"))
+all_taxa$Species[all_taxa$Species == "Unclassified *0319-6G20*"] <- "Uncultured Species 0319-6G20"
+all_taxa$Species[all_taxa$Species == "Unclassified *Chloroplast*"] <- "Chloroplast"
+all_taxa$Species[all_taxa$Species == "Unclassified *Armatimonadales*"] <- "Unclassified Armatimonadales"
+all_taxa$Species[all_taxa$Species == "Unclassified *Gammaproteobacteria*"] <- "Unclassified Gammaproteobacteria"
+all_taxa$Species[all_taxa$Species == "Unclassified *Kapabacteriales*"] <- "Unclassified Kapabacteriales"
+
+all_iqm <- 
+	all_taxa %>%
+	group_by(treat, Species) %>%
+	summarize(median = median(rel_abund), 
+		  first_quartile = quart(rel_abund)[1],
+		  third_quartile = quart(rel_abund)[2], .groups = "drop")
+
+sort_order <-
+	all_taxa %>%
+  select(Species, effect) %>% 
+  distinct() %>% 
+  arrange(effect)
+	# group_by(Species) %>%
+	# summarize(median = median(rel_abund)) %>%
+	# arrange(median)
+
+all_taxa$Species <- factor(all_taxa$Species, levels = sort_order$Species)
+all_iqm$Species <- factor(all_iqm$Species, levels = sort_order$Species)
+all_taxa <-
+  all_taxa %>% 
+  mutate(direction = if_else(effect > 0, 0, 1))
+all_iqm <-
+all_taxa %>% 
+  select(Species, direction) %>% 
+  distinct() %>% 
+  full_join(all_iqm, by = "Species")
+(p <-
+	all_taxa %>%
+	ggplot(aes(x = rel_abund, y = Species, color = treat)) +
+	geom_point(position=position_jitterdodge()) +
+	geom_pointrange(aes(x = median, xmax = third_quartile, xmin = first_quartile, fill = treat), 
+			data = all_iqm, 
+			position = position_dodge(width = 0.75), 
+			shape = 21, 
+			color = "black") +
+	theme_bw() +
+	theme(axis.text.y = element_markdown()) +
+	labs(x = "Relative Abundance (%)", y = NULL) +
+	scale_color_manual(name = "Treatment",
+			   labels = c('Neutral', 'Positive'),
+			   values = c('gray40', 'darkorange2')) +
+	scale_fill_manual(name = "Treatment",
+			   labels = c('Neutral', 'Positive'),
+			   values = c('gray40', 'darkorange2')) +
+    facet_grid(direction ~ ., scale = "free_y", space = "free")  + 
+theme(
+  strip.background = element_blank(),
+  strip.text.y = element_blank()
+))
+
+(p_log10 <- 
+	p +
+	scale_x_log10(breaks = c(0.001, 0.01, 0.1, 1),
+		      labels = c(0.001, 0.01, 0.1, 1)))
+
+save(p_log10, da_fams, file = "Output/da_fig.RData")
+
+ggsave('Manuscript/fig4.jpg', p_log10, width = 5, height = 5)
+
+## All
+# ggsave("shared_taxa.png", p)
+ggsave("shared_taxa_log.png", p_log10, width = 5, height = 5)
+
+# aes_Species_rclr_plot <-
+#   aes_Species_rclr %>% 
+# 	filter(Species %in% ancom_taxa & Species %in% aldex_taxa & Species %in% corncob_taxa) %>% 
+#   pivot_longer(G2:G5, names_to = "passage", values_to = "rclr") %>% 
+#   select(Species, treat, rep, passage, rclr) %>% 
+#   distinct() %>% 	
+#   mutate(Species = str_replace(Species, "(.*)", "*\\1*"),
+# 	       Species = str_replace(Species, "\\*(.*)_unclassified\\*",
+# 				    "Unclassified *\\1*"))
+# 
+# aes_Species_rclr_plot$Species[aes_Species_rclr_plot$Species == "Unclassified *0319-6G20*"] <- "Uncultured Species 0319-6G20"
+# aes_Species_rclr_plot$Species[aes_Species_rclr_plot$Species == "Unclassified *Chloroplast*"] <- "Chloroplast"
+# aes_Species_rclr_plot$Species <- factor(aes_Species_rclr_plot$Species, levels = sort_order$Species)
+# ggplot(aes_Species_rclr_plot, aes(x = rclr, y = Species, color = treat)) +
+#   facet_grid( ~ passage) +
+#   geom_point(position = position_jitterdodge())
+
+p_effect <-
+	all_taxa %>%
+	ggplot(aes(x = effect, y = Species)) +
+	geom_point() +
+
+	theme_bw() +
+	theme(axis.text.y = element_markdown()) +
+	labs(x = "Aldex2 Effect Size", y = NULL) +
+	scale_color_manual(name = "Treatment",
+			   labels = c('Neutral', 'Positive'),
+			   values = c('gray40', 'darkorange2')) +
+	scale_fill_manual(name = "Treatment",
+			   labels = c('Neutral', 'Positive'),
+			   values = c('gray40', 'darkorange2'))
+p_effect
+
+```
+
 ```{r stop}

### 2022-07-15 R/08-community-analysis.Rmd
add dbrda and start updating da

@@
 
-if(file.exists("Output/ancom_out.rds")) {
+if(file.exists("Output/ancom_out_genus.rds")) {
   out <- 
-    readRDS("Output/ancom_out.rds")
+    readRDS("Output/ancom_out_genus.rds")
 } else {
@@
   out %>% 
-    saveRDS("Output/ancom_out.rds")
+    saveRDS("Output/ancom_out_genus.rds")
 }
@@
 
-if(file.exists("Output/aldex_x_all.rds")) {
+if(file.exists("Output/aldex_x_all_genus.rds")) {
   x.all <- 
-    readRDS("Output/aldex_x_all.rds")
+    readRDS("Output/aldex_x_all_genus.rds")
 } else {
@@
   x.all %>% 
-    saveRDS("Output/aldex_x_all.rds")
+    saveRDS("Output/aldex_x_all_genus.rds")
 }
@@
 				    "Unclassified *\\1*"))
-all_taxa$Genus[all_taxa$Genus == "Unclassified *0319-6G20*"] <- "Uncultured Genus 0319-6G20"
-all_taxa$Genus[all_taxa$Genus == "Unclassified *Chloroplast*"] <- "Chloroplast"
-all_taxa$Genus[all_taxa$Genus == "Unclassified *Armatimonadales*"] <- "Unclassified Armatimonadales"
-all_taxa$Genus[all_taxa$Genus == "Unclassified *Gammaproteobacteria*"] <- "Unclassified Gammaproteobacteria"
-all_taxa$Genus[all_taxa$Genus == "Unclassified *Kapabacteriales*"] <- "Unclassified Kapabacteriales"
+# all_taxa$Genus[all_taxa$Genus == "Unclassified *0319-6G20*"] <- "Uncultured Genus 0319-6G20"
+# all_taxa$Genus[all_taxa$Genus == "Unclassified *Chloroplast*"] <- "Chloroplast"
+# all_taxa$Genus[all_taxa$Genus == "Unclassified *Armatimonadales*"] <- "Unclassified Armatimonadales"
+# all_taxa$Genus[all_taxa$Genus == "Unclassified *Gammaproteobacteria*"] <- "Unclassified Gammaproteobacteria"
+# all_taxa$Genus[all_taxa$Genus == "Unclassified *Kapabacteriales*"] <- "Unclassified Kapabacteriales"
 

### 2022-07-26 Manuscript/manuscript.Rmd
update analysis

--- 2022-07-06 Manuscript/manuscript.Rmd
@@
 library(ggthemes)
-library(microViz)
+#library(microViz)
 library(captioner)
@@
 fluxes <- read_tsv('../Output/fluxes.tsv')
+
+load(file = '../Output/decontam.RData')
 
@@
 to determine the relationship between microbiome composition and the rate of
-any ecosystem function. This challenge remains because microbiomes are often
-complex, it is difficult to know a priori which microbiome members limit the
-rate of an ecosystem function, and it is usually challenging to manipulate
-microbiome composition directly. We propose that artificial ecosystem selection
+any ecosystem function. This challenge remains because it is difficult to
+manipulate microbiomes composition directly, we cannot know a priori which members of the
+microbiome limit the rate of an ecosystem function, and microbiomes covary with
+the environment. We propose that artificial ecosystem selection
 can overcome these challenges. Artificial ecosystem selection involves creating
@@
 soil microcosms to select for increased soil methane oxidation rate. We
-observed a strong response of methane oxidation rate to artificial ecosystem
-selection with a 51% increase in methane oxidation rate per round over five
+observed a strong response to selection with a 51% increase in methane oxidation rate per round over five
 rounds of selection. This suggests that a large portion of the variation in
@@
 variation. We also found that selection did not enrich for known methane
-oxidizers; rather microbes that were enriched after selection included 24
-families such as Puniceicoccaceae, Elsteraceae, and Cytophagaceae. This result
+oxidizers; rather microbes that responded to selection included `r length(da_fams)`
+families such as *Fimbriimonadaceae*, *Cytophagaceae*, and *Diplorickettsiaceae*. This result
 is in contrast to the typical assumption that the rate of an ecosystem function
 is limited by the final step in the associated microbial pathway. Our study
-demonstrates that artificial ecosystem selection is a useful strategy for
-revealing relationships between microbiome composition and ecosystem function.
+demonstrates that variation in microbiome composition can contribute to
+variation in the rate of ecosystem function.
 
@@
 exclusive, nor are they likely unique to microbes, but methods to determine the
-relative important of these two scenarios have not until recently been employed.
+relative importance of these two scenarios have not until recently been employed.
 
@@
 approach could be used to identify microbiome attributes, such as genes or taxa,
-associated with the rate of an ecosystem function, although, to my knowledge
+associated with the rate of an ecosystem function, although, to our knowledge
 this has not previously been accomplished [@arias-sanchez2019; @morris2020]. 
@@
 microbial ecosystem-function relationships more accurately and allow us to
-manage ecosystems for particular outcomes [@fuhrman2009; @schimel1998]. Here I
+manage ecosystems for particular outcomes [@fuhrman2009; @schimel1998]. Here, we
 use this approach to estimate how much of the variation in ecosystem function
@@
 
-For the current project, we applied artificial ecosystem selection on soil
-microbiomes by selecting on the ecosystem-scale CH~4~ oxidation rate in order
-to estimate the amount of microbiome variation underlying variation in soil CH~4~ oxidation rate
-[@swenson2000]. To identify which microorganisms could be markers of CH~4~
-oxidation rate, we compared the composition of the artificially selected
-microbiomes to a control set of microbiomes without selection. Artificial
-ecosystem selection has a similar effect to enrichment culturing by amplifying
-the population of interest [@swenson2000]. This will reduce the diversity of
-the soil microbiome and allow for greater power in detecting significant
-markers of CH~4~ oxidation rate. We then evaluated whether these markers meet
-our underlying assumptions about which taxa limit the rate of ecosystem
-function.
+We applied artificial ecosystem selection on soil microbiomes by selecting on
+the ecosystem-scale CH~4~ oxidation rate.  Using this approach, we were able to
+estimate how much of the variation in CH~4~ oxidation rate was associated with
+variation in the microbiome.  To identify which microorganisms could be markers
+of CH~4~ oxidation rate, we compared the composition of the artificially
+selected microbiomes to a control set of microbiomes without selection.  We
+then evaluated whether these markers meet our underlying assumptions about
+which taxa limit the rate of ecosystem function.
 
@@
 the bacterial and archaeal taxa in our soil ecosystems, we sequenced the V4
-region of the 16S rRNA gene using the using the 515F - 806R primer combination
+region of the 16S rRNA gene using the 515F - 806R primer combination
 [@caporaso2011]. PCR mixtures were: 10 $\mu$l NEBNext Q5 Hot Start HiFi PCR
@@
 of amplicon sequence variants (ASVs)
-[@callahan2016]. Taxonomic assignment for the 16S reads was performed
+[@callahan2016]. Taxonomic assignment was performed
 using the Ribosomal Database Project naive Bayesian classifier [@wang2007]. The presence of
 contaminants was evaluated using both the prevalence and frequency methods from
-'DECONTAM' by comparing samples to negative controls [@davis2018]. We did not identify
-any obvious contamination using these methods.
+'DECONTAM' [@davis2018].
+Decontam identified `r length(which(contamdf$contaminant))` potential contaminants based
+on prevalence and frequency. Visual inspection of abundance-concentration
+plots indicated that `r length(contaminants)` of these were likely contaminants and these
+ASVs were removed.
 
@@
 
-Statistical analyses were performed in 'R' [@rcoreteam2018]. To test
-whether there was a significant change in CH~4~ oxidation rate ($k$) as a
+Statistical analyses were performed in 'R' [@rcoreteam2018]. To determine
+whether there was a significant change in CH~4~ oxidation rate as a
 response to selection, we tested a difference in slopes between the positive
-and neutral selection lines. The CH~4~ oxidation rates were strongly 
-right-skewed with most values close to zero and few large, positive values. This
-resulted in residuals that did not meet the assumptions of constant variance
+and neutral selection lines. Residuals did not meet the assumptions of constant variance
 and normal distribution. Therefore, CH~4~ oxidation rates were log~10~
-transformed to better meet the assumptions of a linear model and to make
-figures easier to interpret. First, we tested the difference of slopes between
+transformed prior to analysis. First, we tested the difference of slopes between
 the positive and neutral treatments by
@@
 produced by those parents. First, we tested if there was an effect of treatment
-on the heritability estimate. We compared these models using an F-test
+on the heritability estimate. We compared nested models with and without treatment using an F-test
 test with the 'anova' function. We then fit the full model to estimate the
@@
 
-Richness was estimated using the method from @hurlbert1971 with a subsample size of `r min_seqs` using the 'rarefy'
-function in 'vegan' [@oksanen2019]. We
-tested a difference in richness between Passages 2 and 5 with both treatments
-combined. Then we subset the samples from Passage 5 and compared richness between
-the Positive and Neutral treatment. Next, we compared beta-diversity as the 
-Bray-Curtis dissimilarity by averaging random subsets with a subsample size of `r min_seqs` using the 'avgdist' function in 'vegan' [@bray1957; @oksanen2019].
-We tested a difference in centroid and dispersion of beta
-diversity between the Positive and Neutral treatment and Passage 2 and 5 with
-permutational analysis of variance (PERMANOVA) using the 'adonis2' function from
-'vegan' [@mcardle2001; @oksanen2019].
+Richness was estimated using the method from @hurlbert1971 with a subsample
+size of `r min_seqs` using the 'rarefy' function in 'vegan' [@oksanen2019]. We
+tested a difference in richness by both passage and treatment with a
+Kruskal-Wallace test followed by a pairwise Wilcoxon test. Next, we estimated
+beta-diversity as the Bray-Curtis dissimilarity by averaging 100 random subsets
+with a subsample size of `r min_seqs` using the 'avgdist' function in 'vegan'
+[@bray1957; @oksanen2019].  We tested a difference in centroid and dispersion
+of beta diversity by passage and treatment using a permutational analysis of
+variance (PERMANOVA) with 999 permutations using the 'adonis2' function from
+'vegan' and tested a difference of group dispersions using 'betadisper' and
+'anova' with 999 permutations [@mcardle2001; @oksanen2019].  Lastly, we tested
+the correlation between CH~4~ oxidation rate and Bray-Curtis dissimilarity in
+Passage 5 with a distance-based redundancy analysis (dbRDA) using the 'dbrda'
+function in 'vegan' and estimated the p-value using a permutation F-test with
+999 permutations [@mcardle2001; @oksanen2019]
 
@@
 For ANCOM-II, we used the 'ancom' function in the 'ANCOM-BC' package with a cutoff of W = 0.7 [@kaul2017; @lin2020].
-For ALDEx2, we used the 'aldex' function in the 'ALDEx2' package with the Welch's t test and we used an effect size of
-1 as our significance threshold [@fernandes2014]. Finally, we used CORNCOB with the 'differentialTest' function in the 'corncob' package with the Wald test and without bootstrapping [@martin2020].
+For ALDEx2, we used the 'aldex' function in the 'ALDEx2' package with Welch's t-test and we used an effect size of
+1 as our significance threshold [@fernandes2014].
+Finally, we used CORNCOB with the 'differentialTest' function in the 'corncob' package with the Wald test and without bootstrapping [@martin2020].
+Lastly, to test differentially abundant methanotrophs, we subset all ASVs within methanotrophic families and tested their differential abundance aggregated at the family and genus level using 'corncob'.
 
@@
 
-Richness of ASVs decreased between Passage 2 (median = `r myround(rare_median[1, 2], 1)`, IQR = `r myround(rare_median[1, 3], 1)`) and 5 (median = `r myround(rare_median[2, 2], 1)`, IQR = `r myround(rare_median[2, 3], 1)`; Kruskal-Wallace test: $\chi^2$ =
+Median ASV richness decreased from `r myround(rare_median[1, 2], 1)` (`r myround(rare_median[1, 3], 1)`) in Passage 2 to `r myround(rare_median[2, 2], 1)` (`r myround(rare_median[2, 3], 1)`) in Passage 5 (Kruskal-Wallace test: $\chi^2$ =
 `r myround(kt[1,1], 1)`, df = `r kt[1,3]`, p < 0.001; pairwise Wilcoxon test: p < 0.001). However, there was no
@@
 Passage 2 or 5 (pairwise Wilcoxon test: p = `r myround(wt[1, 3], 1)`).  In addition, there was no
-correlation between richness and CH~4~ oxidation rate across all passages
-and treatments (Spearman's rho = `r myround(s[1,1], 1)`, p = `r myround(s[1,3], 1)`).
+correlation between richness and CH~4~ oxidation rate across the two treatments in Passage 5 (Spearman's rho = `r myround(s5[1,1], 1)`, p = `r myround(s5[1,3], 1)`).
 
@@
 Bray-Curtis dissimilarity of the soil microbiome varied strongly by passage and
-weakly by treatment and the interaction between passage and treatment (`r figures("beta", display = "cite")`; `r tables("tidy_permanova", display = "cite")`). 
+weakly by treatment with an interaction between passage and treatment (`r figures("beta", display = "cite")`). 
 Passage explained `r myround(100*tidy_permanova[1, 4])`\% of the
-variation in Bray-Curtis dissimilarity.  Treatment explained 
+variation in Bray-Curtis dissimilarity (F~`r tidy_permanova[1, 2]`,`r tidy_permanova[4, 2]`~ = `r myround(tidy_permanova[1, 5], 1)`, p = `r myround(tidy_permanova[1, 6], 3)`), treatment explained 
 `r myround(100*tidy_permanova[2, 4])`\%
-of the variation. There was also an interaction
-between treatment and passage, which explained 
-`r myround(100*tidy_permanova[3, 4])`\% of the variation. We found there was 
+of the variation (F~`r tidy_permanova[2, 2]`,`r tidy_permanova[4, 2]`~ = `r myround(tidy_permanova[2, 5], 1)`, p = `r myround(tidy_permanova[2, 6], 3)`), and the interaction
+between treatment and passage explained 
+`r myround(100*tidy_permanova[3, 4])`\% of the variation (F~`r tidy_permanova[3, 2]`,`r tidy_permanova[4, 2]`~ = `r myround(tidy_permanova[3, 5], 1)`, p = `r myround(tidy_permanova[3, 6], 3)`).
+There was
 no difference in dispersion between treatments or passages (F~3,44~ = 0.91, p = 0.45). 
+Finally, CH~4~ oxidation rate was correlated with Bray-Curtis dissimilarity across both treatments in Passage 5 and explained `r myround(dbrda_percent_explained[1], 1)`\% of the variation in Bray-Curtis dissimilarity (dbRDA: F~`r dbrda_aov[1, 1]`,`r dbrda_aov[2, 1]`~ = `r myround(dbrda_aov[1, 3], 2)`, p = `r myround(dbrda_aov[1, 4], 3)`)
 
@@
 prevalence. This included an ASV that was a member of the Bacteroidia Class with
-no lower taxonomic designation. This ASV had a prevalence of 10/12 and a median
+no lower taxonomic assignment. This ASV had a prevalence of 10/12 and a median
 abundance of ~2 reads. The other prevalent family was a member of the
@@
 Overall, none of the families enriched in the Positive selection treatment
-are known methanotrophs. Several taxa identified had a 
+contain known methanotrophs. Several taxa identified had a 
 higher taxonomic designation that contains methanotrophs, for example, the
 Gammaproteobacteria class had a large effect size.
-This class is one of the groups containing methanotrophic families such as
-*Methylococcaceae.* However, Gammaproteobacteria is among the most diverse 
+The Gammaproteobacteria include the type I and type X methanotrophs in the families *Methylococcaceae* and *Methylothermaceae* [@stein2012].
+However, the Gammaproteobacteria is among the most diverse 
 groups in the Prokaryotes, so this is not strong evidence for a selection response
 by methanotrophs [@garrity2005].
-The next family
-was *Puniceicoccaceae*, which is a member of the phylum Verrucomicrobia and
-contains marine microbes. The Verrucomicrobia is a diverse group that contain
+In addition, the *Puniceicoccaceae* is a member of the phylum Verrucomicrobia. The Verrucomicrobia is a diverse group that contain
 known methanotrophs as well as ammonia-oxidizing bacteria [@freitag2003].
-An ASV from the order Armatimonadales was also enriched in the Positive treatment.
-This order also contains the family *Armatimonadaceae*, which was
-found to be unique to the Neutral treatment. *Cytophagaceae* was also enriched in
+Other than these two groups, none of the other taxa enriched in the Positive treatment are known to be related to methanotrophs.
+Two groups in the Armatimonadales were enriched in the Positive treatment including the family *Fimbriimonadaceae* and an unclassified ASV from the order Armatimonadales [@im2012].
+*Cytophagaceae* was also enriched in
 the Positive treatment and contains a number of mainly aerobic heterotrophs
-that can digest a variety of macromolecules [@mcbride2014]. Notably, the family
-*Gemmatimonadaceae* has a single bacterial bin, Candidatus '*Methylotropicum kingii*',
-that contains methanotrophy genes, but this family was depleted in the Positive
-selection treatment [@bay2021].
+that can digest a variety of macromolecules [@mcbride2014]. 
+The remaining families include the uncultured family 0319-6G20, *Diplorickettsiaceae*, *Rhodospirillaceae*, and an unclassified Kapabacteriales.
 
-# Discussion
+We did not identify any methanotrophic families in the overall differential abundance analysis. 
+However, we wanted to look more closely at the known methanotrophs in our dataset to be sure they did not have an effect.
+To do this, we subset all of the ASVs in our dataset that were in families that contained methanotrophs.
+Only two families were represented: *Methylacidiphilaceae* and *Beijerinckiaceae*. 
+Aggregating reads at the family level, neither family was differentially abundant between the two treatments.
+However, aggregated at the genus level, a group of unclassified genera in the *Beijerinckiaceae* were depleted in the Positive treatment and the genus
+*Rhodoblastus*, a member of the *Beijerinckiaceae*, was enriched in the Positive treatment.
+While many *Beijerinckiaceae* are methanotrophs, several taxa in this family have lost the the ability to oxidize CH~4~ and it appears that *Rhodoblastus* species are not able to grow on CH~4~, though they can grow on methanol [@tamas2014, @dedysh2016].
+Therefore, based on this 16S analysis, it appears that no methanotrophs were enriched in the Positive treatment.
+
+# New Discussion
+
+We observed a response to selection on whole-ecosystem soil methane oxidation rate, which demonstrates that variation in the microbiome generates variation in the rate of ecosystem function independent of the underlying environmental conditions.
+While it is well understood that the microbiome mediates a variety of ecosystem functions, a fundamental question in microbial ecology is whether we need to understand variation in the composition of microbiomes to understand variation in the rate of ecosystem functions.
+To address that question, we performed artificial ecosystem selection on soil microbiomes by selecting for ecosystems with high methane oxidation rates.
+We found that there was a significant increase in methane oxidation rate as a response to selection.
+Each generation of soil ecosystems was created from the same pool of sterile soil substrate and a small inoculum from the previous generation.
+Therefore, it is unlikely that this response to selection was due to changes in the abiotic conditions of the ecosystem.
+In addition, the response was only observed in the Positive treatment and not in the Neutral treatment, which shows that this was not a side effect of passaging in jars.
+Therefore, we conclude that variation in the microbiome can generate variation in ecosystem functions independent of the environment.
+
+In addition, we observed a large heritability for methane oxidation rate, which suggests that microbiome variation explains a large portion of the variation in methane oxidation rate relative to environmental variation in this system.
+To estimate how much of the variation in methane oxidation rate can be explained by the presence, absence, and relative abundances of specific microbial taxa, we calculated the narrow-sense heritability of methane oxidation rate, which was `r myround(coefficients(herit_model)[4], 2)` for the Positive treatment.
+A heritability of 1 would indicate that a trait is entirely genetically determined and thus offspring would have the exact same value of that trait as their parents.
+A heritability greater than 1 indicates that offspring exceed the value of their parents.
+This is unlikely in the context of a genomic trait, but microbiomes are not limited by the constraints of genomes because microbes are not alleles and communities are not chromosomes.
+Therefore, complementarity within a microbial community could result in offspring communities that exceed the trait values of their "parent" ecosystems.
+The large heritability we observed indicates that microbiome variation is a large contributor to the variation in methane oxidation in this experimental system.
+
+Notably, the Neutral selection treatment displayed a negative heritability of `r myround(coefficients(herit_model)[2], 2)`.
+Negative heritability estimates can arise in artificial selection experiments.
+One explanation for this phenomenon is that traits that are favored under the experimental conditions are negatively correlated with the trait under selection.
+In this experiment, CH~4~ oxidation rate could be negatively correlated with traits that favor persistence in jars [@conner2003].
+This tradeoff would result in a negative heritability because, in the absence of an imposed selection pressure, passaging in jars would favor microbes with lower CH4 oxidation rates.
+
+Next, we wanted to address what aspects of the microbiome, in terms of alpha diversity, beta diversity, or individual taxa, explain this contribution of the microbiome to variation in methane oxidation rate.
+There are three ways that the microbiome could respond to selection in this experiment: gain or loss of species, changes in the relative abundances of species, or changes within the genomes of the constituent species.
+Using 16S rRNA gene sequences, we are unable to address the third possibility of whether taxa in this experiment could have evolved changes in their genomes as a result of selection.
+Therefore, we will focus on the first two possibilities.
+
+Richness at the ASV level did not vary between the two treatments and there were relatively few taxa gained or lost in the Positive selection treatment and none of these were prevalent across the 12 jars in passage 5.
+Therefore, the gain or loss of species is unlikely to explain the increase in methane oxidation rate.
+However, we found that Bray-Curtis dissimilarity was greater between the two treatments in Passage 5 than within each treatment and was correlated with methane oxidation rate, which suggests that changes in the relative abundance of taxa could explain the response to selection.
+A difference in composition between the two treatments could be a direct result of selection for organisms that contribute to CH~4~ oxidation or a result of ecological drift over the course of the experiment.
+In the former case, that would suggest that changes in the relative abundance of specific microbial taxa could explain the response to selection on CH~4~ oxidation rate.
+
+Even though we observed an increase in methane oxidation rate in the Positive treatment and a difference in composition between the two treatments, we did not observe an increase in the relative abundance of methanotrophs.
+This contrasts with the usual assumption that the final step in a metabolic pathway is the rate-limiting step (citation?).
+In certain ecosystems, methane oxidation rate is correlated with the abundance of methanotrophs as estimated from marker genes (Cite.
+I know that one methanogenesis paper, not sure about methanotrophy).
+However, it does not appear generally true that the rate of an ecosystem function is limited by the abundance of the gene that encodes the final step in that pathway (cite rocca, maybe some of the methane work.) Since we did not observe an increase in the relative abundance of methanotrophs in the treatment that displayed in increase in methane oxidation rate, we may have to look at other microbial functional groups within an ecosystem to understand what drives changes in the rate of ecosystem functions.
+
+One of the goals of this study was to determine how much of the variation in ecosystem function is due to variation in the microbiome.
+In this experiment, we eliminated as much environmental variation as possible.
+This was intentional, because we wanted to determine whether it was possible for variation in the microbiome to generate variation in methane oxidation rate independent of the environment.
+As a result, our heritability estimates are likely to be greater than they would be in natural ecosystems with greater environmental variation.
+Future work could estimate heritability in experimental systems that manipulate environmental variation or across natural ecosystems to evaluate how much of the variation in ecosystem function can be explained by microbiome variation outside the lab.
+This experiment also had a relatively small sample size which resulted in heritability estimates with wide error bars.
+Increasing replication would help us more accurately estimate this parameter and give us a better understanding of the relative importance of environmental variation and microbiome variation in the rate of microbiome-mediated ecosystem functions.
+
+Another important question is whether selection at the ecosystem scale resulted in within-strain evolution.
+Using 16S rRNA gene sequencing we are unable to resolve genomic changes within taxa that may have undergone evolution.
+Future work could analyze an artificial ecosystem selection experiment using genome-resolved metagenomics to try to detect strain-level changes in the constituent taxa.
+This would help evaluate whether certain metabolic pathways were enriched in the positive selection treatment or if functional genes changed in frequency within the microbiome.
+
+As more is understood about the role of the microbiome in regulating ecosystem functions and the specific taxa involved, more detailed experiments could selectively add or remove important taxa or manipulate their relative abundance to start to estimate the quantitative relationship between specific taxa and the rate of an ecosystem function.
+Both our differential abundance analysis and this reductionist approach to manipulating individual taxa assumes that taxa act individually on ecosystem function.
+However, it is much more likely that the microbiome affects the rate of ecosystem function through the interactions of multiple taxa and future analyses with greater replication should test the role of multiple taxa simultaneously using machine learning methods [@topçuoğlu2020]
+
+A fundamental question in microbial ecology is whether variation in the microbiome contributes to variation in ecosystem function.
+We performed artificial ecosystem selection on soil methane oxidation rate to test the relationship between microbiome variation and soil methane oxidation rate.
+We observed a response to selection on soil methane oxidation rate and found a high heritability for methane oxidation, which suggests that variation in the microbiome contributes to variation in soil methane oxidation rate.
+Surprisingly, we did not observe an increase in the relative of abundance of methanotrophs as a response to selection.
+Future research could estimate heritability under greater environmental variation to understand how much of the variation in methane oxidation rate is due to variation in the microbiome in a natural ecosystem.
+In addition, future selection experiments could use genome-resolved metagenomics to observe within-strain evolution and test changes in the functional pathways of the microbiome as a response to selection to determine what aspects of the microbiome explain this increase in oxidation rate.
+This study demonstrates that soil methane oxidation rate can vary greatly with variation in the microbiome and opens up future research opportunities to establish a more detailed understanding of the relationship between the microbiome and ecosystem function.
+
+# Old Discussion
 
@@
 
-# Conclusion
-
 We performed an artificial ecosystem selection experiment on whole-ecosystem
@@
 Taxa are sorted by their effect size with taxa at the top having the largest positive effect size and taxa at the bottom with the largest negative effect size.")`
-
-`r tables("adonis", "PERMANOVA on Bray-Curtis dissimilarities.")`
-
-```{r adonis}
-
-tidy_permanova %>%
-	flextable()
-
-```

### 2022-07-26 R/08-community-analysis.Rmd
update analysis

@@
   filter(passage == "G5") %>% 
-  cor.test(~ ch4 + rare, data = ., method = "spearman", alternative = "two.sided")
+  cor.test(~ ch4 + rare, data = ., method = "spearman", alternative = "two.sided") %>%
+  tidy
 
 min_seqs <- paste(substr(as.character(min_seqs), 1, 3), substr(as.character(min_seqs), 4, 6), sep = ",")
-save(kt, wt, rare_median, s, min_seqs, file = "Output/richness_models.RData")
+save(kt, wt, rare_median, s5, min_seqs, file = "Output/richness_models.RData")
 
@@
 
-## Beta diversity test passage by treatment
-
-```{r permanova-passage-treatment, warnings=FALSE}
-
+```{r}
 dist_df <-
@@
 betadisper(rare_dist_bray, meta_dist$passage_treat) %>% 
-  permutest()
+  anova()
 
@@
 
-# Distance-based RDA Methane
-
-```{r full-dbrda}
-
-meta_dist %>% 
-  select(passage_treat, ch4) %>% 
-  mutate(log_ch4 = log_methane(ch4)) %>% 
-  ggplot(aes(x = passage_treat, y = log_ch4)) +
-  geom_boxplot() +
-  theme_classic() +
-  labs(x = "Passage x Treatment", y = "Log10 Methane")
-
-dbrda_ch4 <- dbrda(rare_dist_bray ~ rank(ch4), 
-                      data = meta_dist, 
+## dbrda methane in passage 5
+
+```{r dbrda, warnings=FALSE}
+passage5_df <-
+  aes_data %>%  
+  filter(passage == "G5") %>% 
+  group_by(asv) %>% 
+  mutate(total = sum(count)) %>% 
+  filter(total != 0) %>% 
+  ungroup() %>% 
+  select(-total) %>% 
+  select(sample, asv, count) %>%
+  pivot_wider(names_from = "asv", values_from = "count", values_fill = 0) %>% 
+  as.data.frame()
+
+rownames(passage5_df) <- passage5_df$sample
+passage5_df <- passage5_df[, -1]
+
+if(file.exists("Output/rare_dist_bray_p5.rds")) {
+  rare_dist_bray_p5 <- 
+    readRDS("Output/rare_dist_bray_p5.rds")
+} else {
+rare_dist_bray_p5 <- avgdist(passage5_df, dmethod="bray", sample=seq_range[1])
+  rare_dist_bray_p5 %>% 
+    saveRDS("Output/rare_dist_bray_p5.rds")
+}
+
+rare_dist_df <-
+  rare_dist_bray_p5 %>% 
+  as.matrix %>% 
+  as_tibble(rownames = "sample")
+
+dbrda_meta_dist <-
+  aes_data %>% 
+  select(sample, passage_treat, passage, treat, ch4) %>% 
+  distinct %>% 
+  inner_join(rare_dist_df, by = "sample")
+
+dbrda_ch4 <- dbrda(rare_dist_bray_p5 ~ rank(ch4), 
+                      data = dbrda_meta_dist, 
                       na.action = "na.omit",
@@
 
+dbrda_aov <-
   dbrda_ch4 %>% 
-  permutest(permutations = 9999)
-
-percent_explained <- 100 * eigenvals(dbrda_ch4) / sum(eigenvals(dbrda_ch4))
+  anova()
+
+dbrda_percent_explained <- 100 * eigenvals(dbrda_ch4) / sum(eigenvals(dbrda_ch4))
 
@@
 
-```
-
-```{r pre-dbrda-data-subsets}
-
-# aes_data %>% 
-#   filter(treat == "P") %>% 
-#   group_by(asv) %>% 
-#   mutate(total = sum(count)) %>% 
-#   filter(total != 0) %>% 
-#   ungroup() %>% 
-#   select(-total) %>% 
-#   group_by(sample) %>% 
-#   summarize(n_seqs = sum(count)) %>% 
-#   summarize(min = min(n_seqs))
-
-treatP_df <-
-  aes_data %>%  
-  filter(treat == "P") %>% 
-  group_by(asv) %>% 
-  mutate(total = sum(count)) %>% 
-  filter(total != 0) %>% 
-  ungroup() %>% 
-  select(-total) %>% 
-  select(sample, asv, count) %>%
-  pivot_wider(names_from = "asv", values_from = "count", values_fill = 0) %>% 
-  as.data.frame()
-
-rownames(treatP_df) <- treatP_df$sample
-treatP_df <- treatP_df[, -1]
-
-# aes_data %>% 
-#   filter(passage == "G5") %>% 
-#   group_by(asv) %>% 
-#   mutate(total = sum(count)) %>% 
-#   filter(total != 0) %>% 
-#   ungroup() %>% 
-#   select(-total) %>% 
-#   group_by(sample) %>% 
-#   summarize(n_seqs = sum(count)) %>% 
-#   summarize(min = min(n_seqs))
-
-passage2_df <-
-  aes_data %>%  
-  filter(passage == "G2") %>% 
-  group_by(asv) %>% 
-  mutate(total = sum(count)) %>% 
-  filter(total != 0) %>% 
-  ungroup() %>% 
-  select(-total) %>% 
-  select(sample, asv, count) %>%
-  pivot_wider(names_from = "asv", values_from = "count", values_fill = 0) %>% 
-  as.data.frame()
-
-rownames(passage2_df) <- passage2_df$sample
-passage2_df <- passage2_df[, -1]
-
-passage5_df <-
-  aes_data %>%  
-  filter(passage == "G5") %>% 
-  group_by(asv) %>% 
-  mutate(total = sum(count)) %>% 
-  filter(total != 0) %>% 
-  ungroup() %>% 
-  select(-total) %>% 
-  select(sample, asv, count) %>%
-  pivot_wider(names_from = "asv", values_from = "count", values_fill = 0) %>% 
-  as.data.frame()
-
-rownames(passage5_df) <- passage5_df$sample
-passage5_df <- passage5_df[, -1]
-
-```
-
-```{r treat-dbrda}
-
-if(file.exists("Output/treat_rare_dist_bray.rds")) {
-  treat_rare_dist_bray <- 
-    readRDS("Output/treat_rare_dist_bray.rds")
-} else {
-treat_rare_dist_bray <- avgdist(treatP_df, dmethod="bray", sample=seq_range[1])
-  treat_rare_dist_bray %>% 
-    saveRDS("Output/treat_rare_dist_bray.rds")
-}
-
-treat_dist_df <-
-  treat_rare_dist_bray %>% 
-  as.matrix %>% 
-  as_tibble(rownames = "sample")
-
-treat_meta_dist <-
-  aes_data %>% 
-  select(sample, passage_treat, passage, treat, ch4) %>% 
-  distinct %>% 
-  inner_join(treat_dist_df, by = "sample")
-
-treat_dbrda_ch4 <- dbrda(treat_rare_dist_bray ~ rank(ch4), 
-                      data = treat_meta_dist, 
-                      na.action = "na.omit",
-                      add = TRUE)
-
-treat_dbrda_ch4 %>% 
-  permutest(permutations = 9999)
-
-percent_explained <- 100 * eigenvals(treat_dbrda_ch4) / sum(eigenvals(treat_dbrda_ch4))
-
-treat_dbrda_scores <-
-  scores(treat_dbrda_ch4, display = "sites") %>% 
-  as_tibble(rownames = "sample")
-
-treat_dbrda_data <-
-  aes_data %>% 
-  select(sample, passage, treat, ch4) %>% 
-  distinct %>% 
-  inner_join(treat_dbrda_scores, by = "sample")
-
-treat_dbrda_data %>%   
-  ggplot(aes(x = dbRDA1, y = MDS1, shape = passage, color = treat)) +
-  geom_point() +
-  stat_ellipse(show.legend = FALSE) +
-  coord_fixed() +
-  scale_shape_discrete(name = "Passage") +
-  scale_color_manual(name = "Treatment",
-                       labels = c('Positive'),
-                       values = c('darkorange2')) +
-  theme_bw() +
-  theme(
-    plot.background = element_blank(),
-    panel.grid.major = element_blank(),
-    panel.grid.minor = element_blank())
-
-```
-
-```{r passage2-dbrda}
-if(file.exists("Output/passage2_rare_dist_bray.rds")) {
-  passage2_rare_dist_bray <- 
-    readRDS("Output/passage2_rare_dist_bray.rds")
-} else {
-passage2_rare_dist_bray <- avgdist(passage2_df, dmethod="bray", sample=seq_range[1])
-  passage2_rare_dist_bray %>% 
-    saveRDS("Output/passage2_rare_dist_bray.rds")
-}
-
-passage2_dist_df <-
-  passage2_rare_dist_bray %>% 
-  as.matrix %>% 
-  as_tibble(rownames = "sample")
-
-passage2_meta_dist <-
-  aes_data %>% 
-  select(sample, passage_treat, passage, treat, ch4) %>% 
-  distinct %>% 
-  inner_join(passage2_dist_df, by = "sample")
-
-passage2_dbrda_ch4 <- dbrda(passage2_rare_dist_bray ~ rank(ch4), 
-                      data = passage2_meta_dist, 
-                      na.action = "na.omit",
-                      add = TRUE)
-
-passage2_dbrda_permtest <- 
-passage2_dbrda_ch4 %>% 
-  permutest(permutations = 9999)
-
-passage2_dbrda_permtest <- 
-passage2_dbrda_ch4 %>% 
-  anova(permutations = 9999) %>% 
-  tidy()
-
-passage2_percent_explained <- 100 * eigenvals(passage2_dbrda_ch4) / sum(eigenvals(passage2_dbrda_ch4))
-
-passage2_dbrda_scores <-
-  scores(passage2_dbrda_ch4, display = "sites") %>% 
-  as_tibble(rownames = "sample")
-
-passage2_dbrda_data <-
-  aes_data %>% 
-  select(sample, passage, treat, ch4) %>% 
-  distinct %>% 
-  inner_join(passage2_dbrda_scores, by = "sample")
-
-passage2_dbrda_data %>%   
-  ggplot(aes(x = dbRDA1, y = MDS1, shape = passage, color = treat)) +
-  geom_point() +
-  stat_ellipse(show.legend = FALSE) +
-  coord_fixed() +
-  scale_shape_discrete(name = "Passage") +
-  scale_color_manual(name = "Treatment",
-                       labels = c('Neutral', 'Positive'),
-                       values = c('gray40', 'darkorange2')) +
-  theme_bw() +
-  theme(
-    plot.background = element_blank(),
-    panel.grid.major = element_blank(),
-    panel.grid.minor = element_blank())
-
-```
-
-```{r passage5-dbrda}
-if(file.exists("Output/passage5_rare_dist_bray.rds")) {
-  passage5_rare_dist_bray <- 
-    readRDS("Output/passage5_rare_dist_bray.rds")
-} else {
-passage5_rare_dist_bray <- avgdist(passage5_df, dmethod="bray", sample=seq_range[1])
-  passage5_rare_dist_bray %>% 
-    saveRDS("Output/passage5_rare_dist_bray.rds")
-}
-
-passage5_dist_df <-
-  passage5_rare_dist_bray %>% 
-  as.matrix %>% 
-  as_tibble(rownames = "sample")
-
-passage5_meta_dist <-
-  aes_data %>% 
-  select(sample, passage_treat, passage, treat, ch4) %>% 
-  distinct %>% 
-  inner_join(passage5_dist_df, by = "sample")
-
-passage5_dbrda_ch4 <- dbrda(passage5_rare_dist_bray ~ rank(ch4), 
-                      data = passage5_meta_dist, 
-                      na.action = "na.omit",
-                      add = TRUE)
-
-passage5_dbrda_permtest <-
-passage5_dbrda_ch4 %>% 
-  anova(permutations = 9999) %>% 
-  tidy
-
-passage5_percent_explained <- 100 * eigenvals(passage5_dbrda_ch4) / sum(eigenvals(passage5_dbrda_ch4))
-
-passage5_dbrda_scores <-
-  scores(passage5_dbrda_ch4, display = "sites") %>% 
-  as_tibble(rownames = "sample")
-
-passage5_dbrda_data <-
-  aes_data %>% 
-  select(sample, passage, treat, ch4) %>% 
-  distinct %>% 
-  inner_join(passage5_dbrda_scores, by = "sample")
-
-passage5_dbrda_data %>%   
-  ggplot(aes(x = dbRDA1, y = MDS1, shape = passage, color = treat)) +
-  geom_point() +
-  stat_ellipse(show.legend = FALSE) +
-  coord_fixed() +
-  scale_shape_discrete(name = "Passage") +
-  scale_color_manual(name = "Treatment",
-                       labels = c('Neutral', 'Positive'),
-                       values = c('gray40', 'darkorange2')) +
-  theme_bw() +
-  theme(
-    plot.background = element_blank(),
-    panel.grid.major = element_blank(),
-    panel.grid.minor = element_blank())
-
-```
-
-```{r}
-save(passage2_dbrda_permtest, 
-     passage5_dbrda_permtest, 
-     passage5_percent_explained,
-     file = "Output/dbrda.RData")
-```
-
-The overall dbRDA is significant and rank of methane explains 5% of the variation.
-Subsetting the Positive treatment in both G2 and G5, it is not significant: no
-correlation between dissimilarity and rank of methane between G2 and G5. This
-makes sense, because most of that variation would be due to selection in the jar.
-However, subsetting passage G5 and comparing N and P, 10% of dissimilarity is
-explained by rank of CH4.
+save(dbrda_aov, dbrda_percent_explained, file = "Output/dbrda.RData")
+
+```
 
@@
 ```{r methanotroph-lists}
+# List of methanotrophs from Table 1 of Stein 2012 
+troph_genera <- c('Methylococcus',
+                  'Methylocaldum',
+                  'Methylobacter',
+                  'Methylomicrobium',
+                  'Methylomonas',
+                  'Methylosarcina',
+                  'Methylosoma',
+                  'Crenothrixa',
+                  'Clonothrixa',
+                  'Methylosphaera',
+                  'Methylohalobiusa',
+                  'Methylothermusa',
+                  'Methylocystis',
+                  'Methylosinus',
+                  'Methylocella',
+                  'Methylocapsa',
+                  'Methyloferula',
+                  'Methylomirabilis',
+                  'Methylacidiphilum')
+troph_genera <- c(troph_genera, paste(troph_genera, "unclassified", sep = "_"))
 
@@
                    'Beijerinckiaceae', 
+                   'NC10',
                    'Methylacidiphilaceae')
@@
   theme_bw()
+
+troph_asvs <-
+  aes_data %>% 
+  select(Kingdom:Species, asv) %>% 
+  distinct() %>% 
+  filter(Genus %in% troph_genera) %$%
+  asv
 
@@
 
-```{r stop}
-
-knitr::opts_chunk$set(eval = FALSE)
-
-```
-
 ```{r}

### 2022-07-26 packages.R

@@ -1,0 +1,5 @@
+library(devtools)
+
+# Install specific commit/version from github
+install_github("amorris28/morris", dependencies = FALSE)
+install_github("kbroman/broman", dependencies = FALSE)

### 2022-07-27 Manuscript/manuscript.Rmd
update manuscript with edits for brendan

@@
 any ecosystem function. This challenge remains because it is difficult to
-manipulate microbiomes composition directly, we cannot know a priori which members of the
+manipulate microbiome composition directly, we cannot know a priori which members of the
 microbiome limit the rate of an ecosystem function, and microbiomes covary with
@@
 selected microbiomes to a control set of microbiomes without selection.  We
-then evaluated whether these markers meet our underlying assumptions about
+then evaluated whether these markers met our underlying assumptions about
 which taxa limit the rate of ecosystem function.
@@
 rates in both the Positive treatment `r print_lm(herit_model, 4)` and the
-Neutral treatment `r print_lm(herit_model, 2)`. Notably, the sign of the effect
-was reversed with a positive heritability for the Positive treatment and a negative heritability for the Neutral treatment.
+Neutral treatment `r print_lm(herit_model, 2)`.
 
@@
 
-To investigate which taxa responded to selection on soil CH~4~ oxidation rate, we
-first looked at which taxa were unique to the Positive and Neutral treatments
-aggregated at the family level. Several taxa were present in the Positive
-treatment and absent in the Neutral treatment. Most of these families had low
-prevalence (present in 3 or fewer samples) and low abundance (median < 5 reads).
-There were two families unique to the Positive treatment with relatively high
-prevalence. This included an ASV that was a member of the Bacteroidia Class with
-no lower taxonomic assignment. This ASV had a prevalence of 10/12 and a median
-abundance of ~2 reads. The other prevalent family was a member of the
-*Silvanigrellaceae*, a newly described family placed in its own order. This family
-was present in all 12 samples and had a median abundance of 8 reads.
-*Silvanigrella* is the only cultivated member of
-*Silvanigrellaceae* and was isolated from a temperate fresh water lake
-[@hahn2017]. Of the families unique to the Neutral treatment, only one had a
-prevalence greater than 2/12. This family, *Armatimonadaceae*, was present in half
-of the Neutral samples (prevalence = 6/12). The type strain for *Armatimonadaceae* was isolated from
-the rhizosphere of *Phragmites australis* [@tamaki2011].
-
-The remaining families were shared between the Positive and Neutral treatment,
-but some differed in their relative abundance. To identify taxa that responded
+To identify taxa that responded
 to selection on soil CH~4~ oxidation rate, we tested the differential relative
@@
 However, aggregated at the genus level, a group of unclassified genera in the *Beijerinckiaceae* were depleted in the Positive treatment and the genus
-*Rhodoblastus*, a member of the *Beijerinckiaceae*, was enriched in the Positive treatment.
-While many *Beijerinckiaceae* are methanotrophs, several taxa in this family have lost the the ability to oxidize CH~4~ and it appears that *Rhodoblastus* species are not able to grow on CH~4~, though they can grow on methanol [@tamas2014, @dedysh2016].
-Therefore, based on this 16S analysis, it appears that no methanotrophs were enriched in the Positive treatment.
+*Rhodoblastus*, a member of the *Beijerinckiaceae*, was enriched in the
+Positive treatment.  While many *Beijerinckiaceae* are methanotrophs, several
+taxa in this family have lost the the ability to oxidize CH~4~ and it appears
+that *Rhodoblastus* species are not able to grow on CH~4~, though they can grow
+on methanol [@tamas2014, @dedysh2016]. Based on this analysis,
+it appears that no methanotrophs were enriched in the Positive treatment.
 
@@
 
-We observed a response to selection on whole-ecosystem soil methane oxidation rate, which demonstrates that variation in the microbiome generates variation in the rate of ecosystem function independent of the underlying environmental conditions.
+We observed a response to selection on whole-ecosystem soil methane oxidation rate, which demonstrates that variation in the microbiome can generate variation in the rate of ecosystem function independent of the underlying environmental conditions.
 While it is well understood that the microbiome mediates a variety of ecosystem functions, a fundamental question in microbial ecology is whether we need to understand variation in the composition of microbiomes to understand variation in the rate of ecosystem functions.
@@
 
-Notably, the Neutral selection treatment displayed a negative heritability of `r myround(coefficients(herit_model)[2], 2)`.
-Negative heritability estimates can arise in artificial selection experiments.
-One explanation for this phenomenon is that traits that are favored under the experimental conditions are negatively correlated with the trait under selection.
-In this experiment, CH~4~ oxidation rate could be negatively correlated with traits that favor persistence in jars [@conner2003].
-This tradeoff would result in a negative heritability because, in the absence of an imposed selection pressure, passaging in jars would favor microbes with lower CH4 oxidation rates.
-
 Next, we wanted to address what aspects of the microbiome, in terms of alpha diversity, beta diversity, or individual taxa, explain this contribution of the microbiome to variation in methane oxidation rate.
@@
 I know that one methanogenesis paper, not sure about methanotrophy).
-However, it does not appear generally true that the rate of an ecosystem function is limited by the abundance of the gene that encodes the final step in that pathway (cite rocca, maybe some of the methane work.) Since we did not observe an increase in the relative abundance of methanotrophs in the treatment that displayed in increase in methane oxidation rate, we may have to look at other microbial functional groups within an ecosystem to understand what drives changes in the rate of ecosystem functions.
+However, it does not appear generally true that the rate of an ecosystem function is limited by the abundance of the gene that encodes the final step in that pathway (cite rocca, maybe some of the methane work.) Since we did not observe an increase in the relative abundance of methanotrophs in the treatment that displayed an increase in methane oxidation rate, we may have to consider other microbial functional groups within an ecosystem to understand what drives changes in the rate of ecosystem functions.
 
@@
 Both our differential abundance analysis and this reductionist approach to manipulating individual taxa assumes that taxa act individually on ecosystem function.
-However, it is much more likely that the microbiome affects the rate of ecosystem function through the interactions of multiple taxa and future analyses with greater replication should test the role of multiple taxa simultaneously using machine learning methods [@topçuoğlu2020]
+However, it is much more likely that the microbiome affects the rate of ecosystem function through the interactions of multiple taxa and future analyses with greater replication should test the role of multiple taxa simultaneously by, for example, using machine learning methods [@topçuoğlu2020]
 

### 2022-07-30 R/08-community-analysis.Rmd
figure and test updates for poster and draft to brendan

@@
   coord_fixed() +
-  scale_shape_discrete(name = "Passage") +
+  scale_shape_discrete(name = "Passage", labels = c(2, 5)) +
   scale_color_manual(name = "Treatment",

### 2022-08-09 R/08-community-analysis.Rmd
convert pngs to svgs for poster and update some figure details

@@
 
-ggsave('Manuscript/fig3.jpg', p_beta, height = 2.5, width = 4)
+ggsave('Manuscript/fig3.png', p_beta, height = 2.5, width = 4)
+
+ggsave('Manuscript/fig3.svg', p_beta, height = 2.5, width = 4)
 
@@
 ggsave('Manuscript/fig4.jpg', p_log10, width = 5, height = 5) 
+ggsave('Manuscript/fig4.svg', p_log10, width = 5, height = 5)
 
@@
 # ggsave("shared_taxa.png", p)
-ggsave("shared_taxa_log.png", p_log10, width = 5, height = 5)
+# ggsave("shared_taxa_log.png", p_log10, width = 5, height = 5)
 

### 2022-09-14 R/02-process-herit.Rmd
update analyses and intermediate data files

@@
 
+```{r}
+offspring_each <- 
+  offspring %>% 
+  group_by(passage, treat) %>% 
+  mutate(offspring = log_ch4)
+
+heritability_each <- 
+  selected_mean %>% 
+  left_join(offspring_each, by = c("passage", "treat")) %>% 
+  filter(passage != 5) %>% # Remove passage 5, which has no offspring
+  ungroup()
+
+write_tsv(heritability_each, paste0(der_dir, 'heritability_each.tsv'))
+
+```
+
 ```{r print_data}

### 2022-09-14 R/04-heritability.Rmd
convert pngs to svgs for poster and update some figure details
update analyses and intermediate data files

@@
 
-p_herit <- plot_herit(heritability)
+  ratio <- calc_plot_ratios(heritability$selected, heritability$offspring)
+p_herit <- 
+  ggplot(heritability, aes(selected, offspring, color = treat)) +
+    geom_point() + 
+    stat_smooth(method = 'lm', se = FALSE, formula = 'y ~ x') +
+    labs(x = expression(Parental ~ log[10] ~ CH[4] ~ "(-k)"), 
+         y = expression(atop(Offspring, log[10] ~ CH[4] ~ "(-k)"))) +
+    scale_color_manual(name = "Treatment", 
+                       labels = c('Neutral', 'Positive'), 
+                       values = c('gray40', 'darkorange2')) +
+    coord_fixed(ratio)
 
@@
 ggsave('Manuscript/fig2.jpg', p_herit, width = 4, height = 3) 
+ggsave('Manuscript/fig2.svg', p_herit, width = 4, height = 3)
 
@@
 ```
+
+```{r}
+heritability_each <- read_tsv(paste0(der_dir, 'heritability_each.tsv'))
+
+fit_no_treat <- lm(offspring ~ selected, data = heritability_each)
+kable(tidy(fit_no_treat))
+herit_model <- lm(offspring ~ selected * treat, data = heritability_each)
+kable(tidy(herit_model))
+kable(tidy(anova(fit_no_treat, herit_model)))
+
+ratio <- calc_plot_ratios(heritability_each$selected, heritability_each$offspring)
+p_herit_each <- 
+ggplot(heritability_each, aes(selected, offspring, color = treat)) +
+    # theme_classic() +
+    # geom_rangeframe(color = 'black') +
+    geom_point() + 
+    stat_smooth(method = 'lm', se = FALSE, formula = 'y ~ x') +
+    labs(x = expression(atop(Mean ~ Parental, log[10] ~ CH[4] ~ "(-k)")), 
+         y = expression(atop(Offspring, log[10] ~ CH[4] ~ "(-k)"))) +
+    scale_color_manual(name = "Treatment", 
+                       labels = c('Neutral', 'Positive'), 
+                       values = c('gray40', 'darkorange2')) +
+    coord_fixed(ratio)
+
+save(p_herit_each, file = 'Output/herit_plot.RData')
+
+ggsave('Manuscript/fig2_each.jpg', p_herit_each, width = 4, height = 3) 
+ggsave('Manuscript/fig2_each.svg', p_herit_each, width = 4, height = 3) 
+```

### 2022-09-14 R/08-community-analysis.Rmd
update analyses and intermediate data files

@@
   unnest(data)
+
+```
+
+```{r remove_chloroplasts, eval=FALSE}
+
+chloroplast_seqs <- 
+  aes_data %>% 
+  filter(Order == 'Chloroplast') %>% 
+  select(asv, seq) %>% 
+  distinct() %$%
+  seq
+# 10 ASVs were identified as chloroplast in their taxonomy.
+# seq matches complete genomes of chloroplasts on NCBI so I will remove them.
+
+aes_data <- 
+  aes_data %>% 
+  filter(Order != 'Chloroplast')
 
@@
 all_taxa$Family[all_taxa$Family == "Unclassified *0319-6G20*"] <- "Uncultured family 0319-6G20"
-all_taxa$Family[all_taxa$Family == "Unclassified *Chloroplast*"] <- "Chloroplast"
+# all_taxa$Family[all_taxa$Family == "Unclassified *Chloroplast*"] <- "Chloroplast"
 all_taxa$Family[all_taxa$Family == "Unclassified *Armatimonadales*"] <- "Unclassified Armatimonadales"
@@
 ```
-
-```{r}
-
-knitr::opts_chunk$set(eval = FALSE)
-
-```
-
-### Genus
-
-```{r setup-differential-abundance-data-genus}
-
-aes_G5 <-
-  aes_data %>% 
-  filter(passage == "G5")
-
-aes_Genus_G5 <-
-  aes_G5 %>% 
-  select(asv, sample, treat, count, Genus) %>%
-  group_by(sample, Genus) %>% 
-  mutate(total = sum(count)) %>% 
-  ungroup() %>% 
-  select(sample, treat, Genus, total) %>% 
-  distinct()
-
-prevalent_families <- 
-  aes_Genus_G5 %>% 
-  group_by(Genus, treat) %>% 
-  summarize(prev = sum(total > 0), .groups = "drop") %>% 
-  pivot_wider(names_from = "treat", values_from = "prev", values_fill = 0) %>% 
-  filter(N/12 > 0.1 & P/12 > 0.1) %$%
-  Genus
-
-otu_table <- 
-  aes_Genus_G5 %>% 
-  filter(Genus %in% prevalent_families) %>% 
-  select(sample, Genus, total) %>% 
-  pivot_wider(names_from = sample, values_from = total, values_fill = 0) %>% 
-  as.data.frame()
-
-rownames(otu_table) <- otu_table[,1]
-otu_table <- otu_table[,-1]
-
-sample_data <- aes_Genus_G5 %>% 
-  select(sample, treat) %>% 
-  distinct() %>% 
-  as.data.frame()
-
-rownames(sample_data) <- sample_data[,1]
-
-```
-
-#### Corncob
-
-```{r}
-
-da_analysis <- differentialTest(formula = ~ treat,
-                                phi.formula = ~ 1,
-                                formula_null = ~ 1,
-                                phi.formula_null = ~ 1,
-                                test = "Wald", 
-                                boot = FALSE,
-                                data = otu_table,
-                                sample_data = sample_data,
-                                fdr_cutoff = 0.05, B = 100,
-                                taxa_are_rows = TRUE)
-
-dt_data <- plot(da_analysis, 
-                # level = c("Genus"), 
-                data_only = TRUE)
-
-dt_plot <- function(dt_data) {
-  ggplot(dt_data, aes(x = x, xmin = xmin, xmax = xmax, y = reorder(taxa, x))) +
-    geom_pointrange() +
-    geom_vline(xintercept = 0) +
-    labs(x = "Estimate", y = NULL) +
-    coord_cartesian(clip="off") +
-    # theme(panel.border = element_blank()) +
-    theme_classic() +
-      theme(axis.text.y = element_markdown())
-
-}
-
-dt_data %>%
-  mutate(taxa = str_replace(taxa, "(.*)", "*\\1*"),
-         taxa = str_replace(taxa, "\\*(.*)_unclassified\\*",
-                                "Unclassified *\\1*")) %>% 
-  dt_plot()
-
-```
-
-#### ANCOM
-
-```{r}
-
-# Setup phyloseq object
-aes_ps <- phyloseq(otu_table(otu_table, taxa_are_rows=TRUE), sample_data(sample_data))
-
-if(file.exists("Output/ancom_out_genus.rds")) {
-  out <- 
-    readRDS("Output/ancom_out_genus.rds")
-} else {
-  # Run Ancom
-  out = ancom(phyloseq = aes_ps,  
-              p_adj_method = "BH",
-              prv_cut = 0.1, 
-              lib_cut = 0, 
-              main_var = "treat",
-              rand_formula = NULL, 
-              lme_control = NULL,
-              struc_zero = TRUE, 
-              neg_lb = FALSE, 
-              alpha = 0.05, 
-              n_cl = 1)  
-  out %>% 
-    saveRDS("Output/ancom_out_genus.rds")
-}
-
-res = out$res
-q_val = out$q_data
-beta_val = out$beta_data
-# Only consider the effect sizes with the corresponding q-value less than alpha
-beta_val = beta_val * (q_val < 0.05) 
-# Choose the maximum of beta's as the effect size
-beta_pos = apply(abs(beta_val), 2, which.max) 
-beta_max = vapply(seq_along(beta_pos), function(i) beta_val[beta_pos[i], i],
-                  FUN.VALUE = double(1))
-# Number of taxa except structural zeros
-n_taxa = ifelse(is.null(out$zero_ind), 
-                nrow(feature_table), 
-                sum(apply(out$zero_ind, 1, sum) == 0))
-# Cutoff values for declaring differentially abundant taxa
-cut_off = 0.7 * (n_taxa - 1)
-
-df_fig_w = res %>%
-  dplyr::mutate(beta = beta_max,
-                direct = case_when(
-                  detected_0.7 == TRUE & beta > 0 ~ "Positive",
-                  detected_0.7 == TRUE & beta <= 0 ~ "Negative",
-                  TRUE ~ "Not Significant"
-                  )) %>%
-  dplyr::arrange(W)
-df_fig_w$taxon_id = factor(df_fig_w$taxon_id, levels = df_fig_w$taxon_id)
-df_fig_w$W = replace(df_fig_w$W, is.infinite(df_fig_w$W), n_taxa - 1)
-df_fig_w$direct = factor(df_fig_w$direct, 
-                     levels = c("Negative", "Positive", "Not Significant"))
-
-p_w = df_fig_w %>%
-  ggplot(aes(x = taxon_id, y = W, color = direct)) +
-  geom_point(size = 2, alpha = 0.6) +
-  labs(x = "Taxon", y = "W") +
-  scale_color_discrete(name = NULL) + 
-  geom_hline(yintercept = cut_off, linetype = "dotted", 
-             color = "blue", size = 1.5) +
-  geom_text(aes(x = 2, y = cut_off + 0.5, label = "W[0.7]"), 
-            size = 5, vjust = -0.5, hjust = 0, color = "orange", parse = TRUE) +
-  theme_bw() +
-  theme(axis.text.x = element_blank(),
-        axis.ticks.x = element_blank(),
-        panel.grid.major = element_blank())
-p_w
-
-```
-
-#### Aldex2
-
-```{r}
-
-conds <- colnames(otu_table) %>% 
-  substr(3, 3)
-
-if(file.exists("Output/aldex_x_all_genus.rds")) {
-  x.all <- 
-    readRDS("Output/aldex_x_all_genus.rds")
-} else {
-  x.all <- aldex(otu_table, 
-                 conds, mc.samples=1000, test="t", effect=TRUE,
-                 include.sample.summary=FALSE, denom="all", verbose=FALSE, paired.test=FALSE)
-  x.all %>% 
-    saveRDS("Output/aldex_x_all_genus.rds")
-}
-
-par(mfrow=c(1,2))
-aldex.plot(x.all, type="MA", test="welch", xlab="Log-ratio abundance",
-    ylab="Difference")
-aldex.plot(x.all, type="MW", test="welch", xlab="Dispersion",
-    ylab="Difference")
-
-par(mfrow=c(1,2))
-plot(x.all$effect, x.all$we.ep, log="y", cex=0.7, col=rgb(0,0,1,0.2),
-  pch=19, xlab="Effect size", ylab="P value", main="Effect size plot")
-points(x.all$effect, x.all$we.eBH, cex=0.7, col=rgb(1,0,0,0.2),
-  pch=19)
-abline(h=0.05, lty=2, col="grey")
-legend(15,1, legend=c("P value", "BH-adjusted"), pch=19, col=c("blue", "red"))
-
-plot(x.all$diff.btw, x.all$we.ep, log="y", cex=0.7, col=rgb(0,0,1,0.2),
-  pch=19, xlab="Difference", ylab="P value", main="Volcano plot")
-points(x.all$diff.btw, x.all$we.eBH, cex=0.7, col=rgb(1,0,0,0.2),
-  pch=19)
-abline(h=0.05, lty=2, col="grey")
-
-x.all_effect_1 <-
-  x.all %>% 
-  filter(abs(effect) > 1) %>% 
-  as_tibble(rownames = "Genus")
-
-# x.all_effect_1 %>% 
-#   # inner_join(taxa_table, by = "asv")
-#   ggplot(aes(y = reorder(Genus, effect), x = effect)) +
-#   geom_point()
-
-# x.all_effect_1 %>% 
-#   select(Genus, effect)
-
-```
-
-#### Comparison
-
-##### All Significant
-
-```{r}
-
-# CORNCOB
-corncob_taxa <- dt_data$taxa
-
-# ANCOM
-ancom_taxa <- as.character(df_fig_w[df_fig_w$detected_0.7, ]$taxon_id)
-
-# ALDEX2
-aldex_taxa <- x.all_effect_1$Genus
-
-at_least_one_taxa <-
-	aes_Genus_G5 %>%
-	group_by(sample) %>%
-	mutate(rel_abund = (total / sum(total)) * 100) %>%
-	filter(Genus %in% ancom_taxa | Genus %in% aldex_taxa | Genus %in% corncob_taxa) %>%
-	mutate(Genus = str_replace(Genus, "(.*)", "*\\1*"),
-	       Genus = str_replace(Genus, "\\*(.*)_unclassified\\*",
-				    "Unclassified *\\1*"))
-
-at_least_one_iqm <- 
-	at_least_one_taxa %>%
-	group_by(treat, Genus) %>%
-	summarize(median = median(rel_abund), 
-		  first_quartile = quart(rel_abund)[1],
-		  third_quartile = quart(rel_abund)[2], .groups = "drop")
-
-sort_order <-
-	at_least_one_taxa %>%
-	group_by(Genus) %>%
-	summarize(median = median(rel_abund)) %>%
-	arrange(median)
-
-at_least_one_taxa$Genus <- factor(at_least_one_taxa$Genus, levels = sort_order$Genus)
-
-at_least_one_taxa %>%
-	ggplot(aes(x = rel_abund, y = Genus, color = treat)) +
-	geom_point(position=position_jitterdodge()) +
-	geom_pointrange(aes(x = median, xmax = third_quartile, xmin = first_quartile, fill = treat), 
-			data = at_least_one_iqm, 
-			position = position_dodge(width = 0.75), 
-			shape = 21, 
-			color = "black") +
-	theme_bw() +
-	theme(axis.text.y = element_markdown()) +
-	labs(x = "Relative Abundance (%)", y = NULL) +
-	scale_color_manual(name = "Treatment",
-			   labels = c('Neutral', 'Positive'),
-			   values = c('gray40', 'darkorange2')) +
-	scale_fill_manual(name = "Treatment",
-			   labels = c('Neutral', 'Positive'),
-			   values = c('gray40', 'darkorange2'))
-
-```
-
-##### Consensus Taxa
-
-```{r}
-aldex_effect <-
-  x.all_effect_1 %>% 
-  select(Genus, effect)
-
-all_taxa <-
-	aes_Genus_G5 %>%
-	group_by(sample) %>%
-	mutate(rel_abund = (total / sum(total)) * 100) %>%
-  left_join(aldex_effect, by = "Genus") %>% 
-	filter(Genus %in% ancom_taxa & Genus %in% aldex_taxa & Genus %in% corncob_taxa) %>% 
-  ungroup()
-
-da_fams <-
-  all_taxa %>% 
-  select(Genus) %>% 
-  distinct() %$%
-  Genus
-
-all_taxa <-
-  all_taxa %>%
-	mutate(Genus = str_replace(Genus, "(.*)", "*\\1*"),
-	       Genus = str_replace(Genus, "\\*(.*)_unclassified\\*",
-				    "Unclassified *\\1*"))
-# all_taxa$Genus[all_taxa$Genus == "Unclassified *0319-6G20*"] <- "Uncultured Genus 0319-6G20"
-# all_taxa$Genus[all_taxa$Genus == "Unclassified *Chloroplast*"] <- "Chloroplast"
-# all_taxa$Genus[all_taxa$Genus == "Unclassified *Armatimonadales*"] <- "Unclassified Armatimonadales"
-# all_taxa$Genus[all_taxa$Genus == "Unclassified *Gammaproteobacteria*"] <- "Unclassified Gammaproteobacteria"
-# all_taxa$Genus[all_taxa$Genus == "Unclassified *Kapabacteriales*"] <- "Unclassified Kapabacteriales"
-
-all_iqm <- 
-	all_taxa %>%
-	group_by(treat, Genus) %>%
-	summarize(median = median(rel_abund), 
-		  first_quartile = quart(rel_abund)[1],
-		  third_quartile = quart(rel_abund)[2], .groups = "drop")
-
-sort_order <-
-	all_taxa %>%
-  select(Genus, effect) %>% 
-  distinct() %>% 
-  arrange(effect)
-	# group_by(Genus) %>%
-	# summarize(median = median(rel_abund)) %>%
-	# arrange(median)
-
-all_taxa$Genus <- factor(all_taxa$Genus, levels = sort_order$Genus)
-all_iqm$Genus <- factor(all_iqm$Genus, levels = sort_order$Genus)
-all_taxa <-
-  all_taxa %>% 
-  mutate(direction = if_else(effect > 0, 0, 1))
-all_iqm <-
-all_taxa %>% 
-  select(Genus, direction) %>% 
-  distinct() %>% 
-  full_join(all_iqm, by = "Genus")
-(p <-
-	all_taxa %>%
-	ggplot(aes(x = rel_abund, y = Genus, color = treat)) +
-	geom_point(position=position_jitterdodge()) +
-	geom_pointrange(aes(x = median, xmax = third_quartile, xmin = first_quartile, fill = treat), 
-			data = all_iqm, 
-			position = position_dodge(width = 0.75), 
-			shape = 21, 
-			color = "black") +
-	theme_bw() +
-	theme(axis.text.y = element_markdown()) +
-	labs(x = "Relative Abundance (%)", y = NULL) +
-	scale_color_manual(name = "Treatment",
-			   labels = c('Neutral', 'Positive'),
-			   values = c('gray40', 'darkorange2')) +
-	scale_fill_manual(name = "Treatment",
-			   labels = c('Neutral', 'Positive'),
-			   values = c('gray40', 'darkorange2')) +
-    facet_grid(direction ~ ., scale = "free_y", space = "free")  + 
-theme(
-  strip.background = element_blank(),
-  strip.text.y = element_blank()
-))
-
-(p_log10 <- 
-	p +
-	scale_x_log10(breaks = c(0.001, 0.01, 0.1, 1),
-		      labels = c(0.001, 0.01, 0.1, 1)))
-
-save(p_log10, da_fams, file = "Output/da_fig.RData")
-
-ggsave('Manuscript/fig4.jpg', p_log10, width = 5, height = 5)
-
-## All
-# ggsave("shared_taxa.png", p)
-ggsave("shared_taxa_log.png", p_log10, width = 5, height = 5)
-
-# aes_Genus_rclr_plot <-
-#   aes_Genus_rclr %>% 
-# 	filter(Genus %in% ancom_taxa & Genus %in% aldex_taxa & Genus %in% corncob_taxa) %>% 
-#   pivot_longer(G2:G5, names_to = "passage", values_to = "rclr") %>% 
-#   select(Genus, treat, rep, passage, rclr) %>% 
-#   distinct() %>% 	
-#   mutate(Genus = str_replace(Genus, "(.*)", "*\\1*"),
-# 	       Genus = str_replace(Genus, "\\*(.*)_unclassified\\*",
-# 				    "Unclassified *\\1*"))
-# 
-# aes_Genus_rclr_plot$Genus[aes_Genus_rclr_plot$Genus == "Unclassified *0319-6G20*"] <- "Uncultured Genus 0319-6G20"
-# aes_Genus_rclr_plot$Genus[aes_Genus_rclr_plot$Genus == "Unclassified *Chloroplast*"] <- "Chloroplast"
-# aes_Genus_rclr_plot$Genus <- factor(aes_Genus_rclr_plot$Genus, levels = sort_order$Genus)
-# ggplot(aes_Genus_rclr_plot, aes(x = rclr, y = Genus, color = treat)) +
-#   facet_grid( ~ passage) +
-#   geom_point(position = position_jitterdodge())
-
-p_effect <-
-	all_taxa %>%
-	ggplot(aes(x = effect, y = Genus)) +
-	geom_point() +
-
-	theme_bw() +
-	theme(axis.text.y = element_markdown()) +
-	labs(x = "Aldex2 Effect Size", y = NULL) +
-	scale_color_manual(name = "Treatment",
-			   labels = c('Neutral', 'Positive'),
-			   values = c('gray40', 'darkorange2')) +
-	scale_fill_manual(name = "Treatment",
-			   labels = c('Neutral', 'Positive'),
-			   values = c('gray40', 'darkorange2'))
-p_effect
-
-```
-
-### Species
-
-```{r setup-differential-abundance-data-species}
-
-aes_G5 <-
-  aes_data %>% 
-  filter(passage == "G5")
-
-aes_Species_G5 <-
-  aes_G5 %>% 
-  select(asv, sample, treat, count, Species) %>%
-  group_by(sample, Species) %>% 
-  mutate(total = sum(count)) %>% 
-  ungroup() %>% 
-  select(sample, treat, Species, total) %>% 
-  distinct()
-
-prevalent_families <- 
-  aes_Species_G5 %>% 
-  group_by(Species, treat) %>% 
-  summarize(prev = sum(total > 0), .groups = "drop") %>% 
-  pivot_wider(names_from = "treat", values_from = "prev", values_fill = 0) %>% 
-  filter(N/12 > 0.1 & P/12 > 0.1) %$%
-  Species
-
-otu_table <- 
-  aes_Species_G5 %>% 
-  filter(Species %in% prevalent_families) %>% 
-  select(sample, Species, total) %>% 
-  pivot_wider(names_from = sample, values_from = total, values_fill = 0) %>% 
-  as.data.frame()
-
-rownames(otu_table) <- otu_table[,1]
-otu_table <- otu_table[,-1]
-
-sample_data <- aes_Species_G5 %>% 
-  select(sample, treat) %>% 
-  distinct() %>% 
-  as.data.frame()
-
-rownames(sample_data) <- sample_data[,1]
-
-```
-
-#### Corncob
-
-```{r}
-
-da_analysis <- differentialTest(formula = ~ treat,
-                                phi.formula = ~ 1,
-                                formula_null = ~ 1,
-                                phi.formula_null = ~ 1,
-                                test = "Wald", 
-                                boot = FALSE,
-                                data = otu_table,
-                                sample_data = sample_data,
-                                fdr_cutoff = 0.05, B = 100,
-                                taxa_are_rows = TRUE)
-
-dt_data <- plot(da_analysis, 
-                # level = c("Species"), 
-                data_only = TRUE)
-
-dt_plot <- function(dt_data) {
-  ggplot(dt_data, aes(x = x, xmin = xmin, xmax = xmax, y = reorder(taxa, x))) +
-    geom_pointrange() +
-    geom_vline(xintercept = 0) +
-    labs(x = "Estimate", y = NULL) +
-    coord_cartesian(clip="off") +
-    # theme(panel.border = element_blank()) +
-    theme_classic() +
-      theme(axis.text.y = element_markdown())
-
-}
-
-dt_data %>%
-  mutate(taxa = str_replace(taxa, "(.*)", "*\\1*"),
-         taxa = str_replace(taxa, "\\*(.*)_unclassified\\*",
-                                "Unclassified *\\1*")) %>% 
-  dt_plot()
-
-```
-
-#### ANCOM
-
-```{r}
-
-# Setup phyloseq object
-aes_ps <- phyloseq(otu_table(otu_table, taxa_are_rows=TRUE), sample_data(sample_data))
-
-if(file.exists("Output/ancom_out.rds")) {
-  out <- 
-    readRDS("Output/ancom_out.rds")
-} else {
-  # Run Ancom
-  out = ancom(phyloseq = aes_ps,  
-              p_adj_method = "BH",
-              prv_cut = 0.1, 
-              lib_cut = 0, 
-              main_var = "treat",
-              rand_formula = NULL, 
-              lme_control = NULL,
-              struc_zero = TRUE, 
-              neg_lb = FALSE, 
-              alpha = 0.05, 
-              n_cl = 1)  
-  out %>% 
-    saveRDS("Output/ancom_out.rds")
-}
-
-res = out$res
-q_val = out$q_data
-beta_val = out$beta_data
-# Only consider the effect sizes with the corresponding q-value less than alpha
-beta_val = beta_val * (q_val < 0.05) 
-# Choose the maximum of beta's as the effect size
-beta_pos = apply(abs(beta_val), 2, which.max) 
-beta_max = vapply(seq_along(beta_pos), function(i) beta_val[beta_pos[i], i],
-                  FUN.VALUE = double(1))
-# Number of taxa except structural zeros
-n_taxa = ifelse(is.null(out$zero_ind), 
-                nrow(feature_table), 
-                sum(apply(out$zero_ind, 1, sum) == 0))
-# Cutoff values for declaring differentially abundant taxa
-cut_off = 0.7 * (n_taxa - 1)
-
-df_fig_w = res %>%
-  dplyr::mutate(beta = beta_max,
-                direct = case_when(
-                  detected_0.7 == TRUE & beta > 0 ~ "Positive",
-                  detected_0.7 == TRUE & beta <= 0 ~ "Negative",
-                  TRUE ~ "Not Significant"
-                  )) %>%
-  dplyr::arrange(W)
-df_fig_w$taxon_id = factor(df_fig_w$taxon_id, levels = df_fig_w$taxon_id)
-df_fig_w$W = replace(df_fig_w$W, is.infinite(df_fig_w$W), n_taxa - 1)
-df_fig_w$direct = factor(df_fig_w$direct, 
-                     levels = c("Negative", "Positive", "Not Significant"))
-
-p_w = df_fig_w %>%
-  ggplot(aes(x = taxon_id, y = W, color = direct)) +
-  geom_point(size = 2, alpha = 0.6) +
-  labs(x = "Taxon", y = "W") +
-  scale_color_discrete(name = NULL) + 
-  geom_hline(yintercept = cut_off, linetype = "dotted", 
-             color = "blue", size = 1.5) +
-  geom_text(aes(x = 2, y = cut_off + 0.5, label = "W[0.7]"), 
-            size = 5, vjust = -0.5, hjust = 0, color = "orange", parse = TRUE) +
-  theme_bw() +
-  theme(axis.text.x = element_blank(),
-        axis.ticks.x = element_blank(),
-        panel.grid.major = element_blank())
-p_w
-
-```
-
-#### Aldex2
-
-```{r}
-
-conds <- colnames(otu_table) %>% 
-  substr(3, 3)
-
-if(file.exists("Output/aldex_x_all.rds")) {
-  x.all <- 
-    readRDS("Output/aldex_x_all.rds")
-} else {
-  x.all <- aldex(otu_table, 
-                 conds, mc.samples=1000, test="t", effect=TRUE,
-                 include.sample.summary=FALSE, denom="all", verbose=FALSE, paired.test=FALSE)
-  x.all %>% 
-    saveRDS("Output/aldex_x_all.rds")
-}
-
-par(mfrow=c(1,2))
-aldex.plot(x.all, type="MA", test="welch", xlab="Log-ratio abundance",
-    ylab="Difference")
-aldex.plot(x.all, type="MW", test="welch", xlab="Dispersion",
-    ylab="Difference")
-
-par(mfrow=c(1,2))
-plot(x.all$effect, x.all$we.ep, log="y", cex=0.7, col=rgb(0,0,1,0.2),
-  pch=19, xlab="Effect size", ylab="P value", main="Effect size plot")
-points(x.all$effect, x.all$we.eBH, cex=0.7, col=rgb(1,0,0,0.2),
-  pch=19)
-abline(h=0.05, lty=2, col="grey")
-legend(15,1, legend=c("P value", "BH-adjusted"), pch=19, col=c("blue", "red"))
-
-plot(x.all$diff.btw, x.all$we.ep, log="y", cex=0.7, col=rgb(0,0,1,0.2),
-  pch=19, xlab="Difference", ylab="P value", main="Volcano plot")
-points(x.all$diff.btw, x.all$we.eBH, cex=0.7, col=rgb(1,0,0,0.2),
-  pch=19)
-abline(h=0.05, lty=2, col="grey")
-
-x.all_effect_1 <-
-  x.all %>% 
-  filter(abs(effect) > 1) %>% 
-  as_tibble(rownames = "Species")
-
-# x.all_effect_1 %>% 
-#   # inner_join(taxa_table, by = "asv")
-#   ggplot(aes(y = reorder(Species, effect), x = effect)) +
-#   geom_point()
-
-# x.all_effect_1 %>% 
-#   select(Species, effect)
-
-```
-
-#### Comparison
-
-##### All Significant
-
-```{r}
-
-# CORNCOB
-corncob_taxa <- dt_data$taxa
-
-# ANCOM
-ancom_taxa <- as.character(df_fig_w[df_fig_w$detected_0.7, ]$taxon_id)
-
-# ALDEX2
-aldex_taxa <- x.all_effect_1$Species
-
-at_least_one_taxa <-
-	aes_Species_G5 %>%
-	group_by(sample) %>%
-	mutate(rel_abund = (total / sum(total)) * 100) %>%
-	filter(Species %in% ancom_taxa | Species %in% aldex_taxa | Species %in% corncob_taxa) %>%
-	mutate(Species = str_replace(Species, "(.*)", "*\\1*"),
-	       Species = str_replace(Species, "\\*(.*)_unclassified\\*",
-				    "Unclassified *\\1*"))
-
-at_least_one_iqm <- 
-	at_least_one_taxa %>%
-	group_by(treat, Species) %>%
-	summarize(median = median(rel_abund), 
-		  first_quartile = quart(rel_abund)[1],
-		  third_quartile = quart(rel_abund)[2], .groups = "drop")
-
-sort_order <-
-	at_least_one_taxa %>%
-	group_by(Species) %>%
-	summarize(median = median(rel_abund)) %>%
-	arrange(median)
-
-at_least_one_taxa$Species <- factor(at_least_one_taxa$Species, levels = sort_order$Species)
-
-at_least_one_taxa %>%
-	ggplot(aes(x = rel_abund, y = Species, color = treat)) +
-	geom_point(position=position_jitterdodge()) +
-	geom_pointrange(aes(x = median, xmax = third_quartile, xmin = first_quartile, fill = treat), 
-			data = at_least_one_iqm, 
-			position = position_dodge(width = 0.75), 
-			shape = 21, 
-			color = "black") +
-	theme_bw() +
-	theme(axis.text.y = element_markdown()) +
-	labs(x = "Relative Abundance (%)", y = NULL) +
-	scale_color_manual(name = "Treatment",
-			   labels = c('Neutral', 'Positive'),
-			   values = c('gray40', 'darkorange2')) +
-	scale_fill_manual(name = "Treatment",
-			   labels = c('Neutral', 'Positive'),
-			   values = c('gray40', 'darkorange2'))
-
-```
-
-##### Consensus Taxa
-
-```{r}
-aldex_effect <-
-  x.all_effect_1 %>% 
-  select(Species, effect)
-
-all_taxa <-
-	aes_Species_G5 %>%
-	group_by(sample) %>%
-	mutate(rel_abund = (total / sum(total)) * 100) %>%
-  left_join(aldex_effect, by = "Species") %>% 
-	filter(Species %in% ancom_taxa & Species %in% aldex_taxa & Species %in% corncob_taxa) %>% 
-  ungroup()
-
-da_fams <-
-  all_taxa %>% 
-  select(Species) %>% 
-  distinct() %$%
-  Species
-
-all_taxa <-
-  all_taxa %>%
-	mutate(Species = str_replace(Species, "(.*)", "*\\1*"),
-	       Species = str_replace(Species, "\\*(.*)_unclassified\\*",
-				    "Unclassified *\\1*"))
-all_taxa$Species[all_taxa$Species == "Unclassified *0319-6G20*"] <- "Uncultured Species 0319-6G20"
-all_taxa$Species[all_taxa$Species == "Unclassified *Chloroplast*"] <- "Chloroplast"
-all_taxa$Species[all_taxa$Species == "Unclassified *Armatimonadales*"] <- "Unclassified Armatimonadales"
-all_taxa$Species[all_taxa$Species == "Unclassified *Gammaproteobacteria*"] <- "Unclassified Gammaproteobacteria"
-all_taxa$Species[all_taxa$Species == "Unclassified *Kapabacteriales*"] <- "Unclassified Kapabacteriales"
-
-all_iqm <- 
-	all_taxa %>%
-	group_by(treat, Species) %>%
-	summarize(median = median(rel_abund), 
-		  first_quartile = quart(rel_abund)[1],
-		  third_quartile = quart(rel_abund)[2], .groups = "drop")
-
-sort_order <-
-	all_taxa %>%
-  select(Species, effect) %>% 
-  distinct() %>% 
-  arrange(effect)
-	# group_by(Species) %>%
-	# summarize(median = median(rel_abund)) %>%
-	# arrange(median)
-
-all_taxa$Species <- factor(all_taxa$Species, levels = sort_order$Species)
-all_iqm$Species <- factor(all_iqm$Species, levels = sort_order$Species)
-all_taxa <-
-  all_taxa %>% 
-  mutate(direction = if_else(effect > 0, 0, 1))
-all_iqm <-
-all_taxa %>% 
-  select(Species, direction) %>% 
-  distinct() %>% 
-  full_join(all_iqm, by = "Species")
-(p <-
-	all_taxa %>%
-	ggplot(aes(x = rel_abund, y = Species, color = treat)) +
-	geom_point(position=position_jitterdodge()) +
-	geom_pointrange(aes(x = median, xmax = third_quartile, xmin = first_quartile, fill = treat), 
-			data = all_iqm, 
-			position = position_dodge(width = 0.75), 
-			shape = 21, 
-			color = "black") +
-	theme_bw() +
-	theme(axis.text.y = element_markdown()) +
-	labs(x = "Relative Abundance (%)", y = NULL) +
-	scale_color_manual(name = "Treatment",
-			   labels = c('Neutral', 'Positive'),
-			   values = c('gray40', 'darkorange2')) +
-	scale_fill_manual(name = "Treatment",
-			   labels = c('Neutral', 'Positive'),
-			   values = c('gray40', 'darkorange2')) +
-    facet_grid(direction ~ ., scale = "free_y", space = "free")  + 
-theme(
-  strip.background = element_blank(),
-  strip.text.y = element_blank()
-))
-
-(p_log10 <- 
-	p +
-	scale_x_log10(breaks = c(0.001, 0.01, 0.1, 1),
-		      labels = c(0.001, 0.01, 0.1, 1)))
-
-save(p_log10, da_fams, file = "Output/da_fig.RData")
-
-ggsave('Manuscript/fig4.jpg', p_log10, width = 5, height = 5)
-
-## All
-# ggsave("shared_taxa.png", p)
-ggsave("shared_taxa_log.png", p_log10, width = 5, height = 5)
-
-# aes_Species_rclr_plot <-
-#   aes_Species_rclr %>% 
-# 	filter(Species %in% ancom_taxa & Species %in% aldex_taxa & Species %in% corncob_taxa) %>% 
-#   pivot_longer(G2:G5, names_to = "passage", values_to = "rclr") %>% 
-#   select(Species, treat, rep, passage, rclr) %>% 
-#   distinct() %>% 	
-#   mutate(Species = str_replace(Species, "(.*)", "*\\1*"),
-# 	       Species = str_replace(Species, "\\*(.*)_unclassified\\*",
-# 				    "Unclassified *\\1*"))
-# 
-# aes_Species_rclr_plot$Species[aes_Species_rclr_plot$Species == "Unclassified *0319-6G20*"] <- "Uncultured Species 0319-6G20"
-# aes_Species_rclr_plot$Species[aes_Species_rclr_plot$Species == "Unclassified *Chloroplast*"] <- "Chloroplast"
-# aes_Species_rclr_plot$Species <- factor(aes_Species_rclr_plot$Species, levels = sort_order$Species)
-# ggplot(aes_Species_rclr_plot, aes(x = rclr, y = Species, color = treat)) +
-#   facet_grid( ~ passage) +
-#   geom_point(position = position_jitterdodge())
-
-p_effect <-
-	all_taxa %>%
-	ggplot(aes(x = effect, y = Species)) +
-	geom_point() +
-
-	theme_bw() +
-	theme(axis.text.y = element_markdown()) +
-	labs(x = "Aldex2 Effect Size", y = NULL) +
-	scale_color_manual(name = "Treatment",
-			   labels = c('Neutral', 'Positive'),
-			   values = c('gray40', 'darkorange2')) +
-	scale_fill_manual(name = "Treatment",
-			   labels = c('Neutral', 'Positive'),
-			   values = c('gray40', 'darkorange2'))
-p_effect
-
-```
-
-```{r}
-
-troph_genus_data <-
-	aes_data %>%
-	filter(Family %in% methanotrophs) %>%
-	select(sample, count, Genus, passage, treat) %>%
-	group_by(sample, Genus) %>%
-	mutate(total = sum(count)) %>%
-	select(-count) %>%
-	ungroup() %>%
-	group_by(sample) %>%
-	mutate(rel_abund = total / sum(total)) %>%
-	ungroup() %>%
-	distinct() 
-troph_genus_iqm <- 
-	troph_genus_data %>%
-	group_by(passage, treat, Genus) %>%
-	summarize(median = median(rel_abund), 
-		  first_quartile = quart(rel_abund)[1],
-		  third_quartile = quart(rel_abund)[2], .groups = "drop")
-troph_genus_data %>%
-	ggplot(aes(x = rel_abund, y = Genus, color = treat)) +
-	geom_point(position=position_jitterdodge()) +
-	facet_grid( ~ passage) +
-	geom_pointrange(aes(x = median, xmax = third_quartile, xmin = first_quartile, fill = treat), data = troph_genus_iqm, position = position_dodge(width = 0.75), shape = 21, color = "black") +
-	theme_bw()
-ggsave("troph_genus.png")
-
-aes_data %>%
-	select(sample, count, Genus, passage, treat) %>%
-	group_by(sample, Genus) %>%
-	mutate(total = sum(count)) %>%
-	select(-count) %>%
-	ungroup() %>%
-	group_by(sample) %>%
-	mutate(rel_abund = total / sum(total)) %>%
-	ungroup() %>%
-	filter(Genus %in% c("Methylocella", "Microvirga", "Methylorosula")) %>%
-	distinct() %>%
-	ggplot(aes(x = rel_abund, y = Genus)) +
-	geom_jitter() +
-	facet_grid(treat ~ passage)
-
-troph_family_data <-
-	aes_data %>%
-	select(sample, count, Family, passage, treat) %>%
-	group_by(sample, Family) %>%
-	mutate(total = sum(count)) %>%
-	select(-count) %>%
-	ungroup() %>%
-	group_by(sample) %>%
-	mutate(rel_abund = total / sum(total)) %>%
-	ungroup() %>%
-	filter(Family %in% methanotrophs) %>%
-	distinct()
-troph_family_iqm <- 
-	troph_family_data %>%
-	group_by(passage, treat, Family) %>%
-	summarize(median = median(rel_abund), 
-		  first_quartile = quart(rel_abund)[1],
-		  third_quartile = quart(rel_abund)[2], .groups = "drop")
-troph_family_data %>%
-	ggplot(aes(x = rel_abund, y = Family, color = treat)) +
-	geom_point(position=position_jitterdodge()) +
-	facet_grid( ~ passage) +
-	geom_pointrange(aes(x = median, xmax = third_quartile, xmin = first_quartile, fill = treat), data = troph_family_iqm, position = position_dodge(width = 0.75), shape = 21, color = "black") +
-	theme_bw()
-ggsave("troph_family.png")
-
-troph_species_data <-
-	aes_data %>%
-	filter(Family %in% methanotrophs) %>%
-	select(sample, count, Species, passage, treat) %>%
-	group_by(sample, Species) %>%
-	mutate(total = sum(count)) %>%
-	select(-count) %>%
-	ungroup() %>%
-	group_by(sample) %>%
-	mutate(rel_abund = total / sum(total)) %>%
-	ungroup() %>%
-	distinct()
-troph_species_iqm <- 
-	troph_species_data %>%
-	group_by(passage, treat, Species) %>%
-	summarize(median = median(rel_abund), 
-		  first_quartile = quart(rel_abund)[1],
-		  third_quartile = quart(rel_abund)[2], .groups = "drop")
-troph_species_data %>%
-	ggplot(aes(x = rel_abund, y = Species, color = treat)) +
-	geom_point(position=position_jitterdodge()) +
-	facet_grid( ~ passage) +
-	geom_pointrange(aes(x = median, xmax = third_quartile, xmin = first_quartile, fill = treat), data = troph_species_iqm, position = position_dodge(width = 0.75), shape = 21, color = "black") +
-	theme_bw()
-ggsave("troph_species.png")
-
-```
-
-```{r aldex-methanotrophs-family}
-
-x.all <- aldex(troph_data_family, 
-               conds, mc.samples=1000, test="t", effect=TRUE,
-     include.sample.summary=FALSE, denom="all", verbose=FALSE, paired.test=FALSE)
-
-par(mfrow=c(1,2))
-aldex.plot(x.all, type="MA", test="welch", xlab="Log-ratio abundance",
-    ylab="Difference")
-aldex.plot(x.all, type="MW", test="welch", xlab="Dispersion",
-    ylab="Difference")
-
-par(mfrow=c(1,2))
-plot(x.all$effect, x.all$we.ep, log="y", cex=0.7, col=rgb(0,0,1,0.2),
-  pch=19, xlab="Effect size", ylab="P value", main="Effect size plot")
-points(x.all$effect, x.all$we.eBH, cex=0.7, col=rgb(1,0,0,0.2),
-  pch=19)
-abline(h=0.05, lty=2, col="grey")
-legend(15,1, legend=c("P value", "BH-adjusted"), pch=19, col=c("blue", "red"))
-
-plot(x.all$diff.btw, x.all$we.ep, log="y", cex=0.7, col=rgb(0,0,1,0.2),
-  pch=19, xlab="Difference", ylab="P value", main="Volcano plot")
-points(x.all$diff.btw, x.all$we.eBH, cex=0.7, col=rgb(1,0,0,0.2),
-  pch=19)
-abline(h=0.05, lty=2, col="grey")
-
-x.all
-
-```
-
-```{r aldex-methanotrophs-genus}
-
-x.all <- aldex(troph_data_genus, 
-               conds, mc.samples=1000, test="t", effect=TRUE,
-     include.sample.summary=FALSE, denom="all", verbose=FALSE, paired.test=FALSE)
-
-par(mfrow=c(1,2))
-aldex.plot(x.all, type="MA", test="welch", xlab="Log-ratio abundance",
-    ylab="Difference")
-aldex.plot(x.all, type="MW", test="welch", xlab="Dispersion",
-    ylab="Difference")
-
-par(mfrow=c(1,2))
-plot(x.all$effect, x.all$we.ep, log="y", cex=0.7, col=rgb(0,0,1,0.2),
-  pch=19, xlab="Effect size", ylab="P value", main="Effect size plot")
-points(x.all$effect, x.all$we.eBH, cex=0.7, col=rgb(1,0,0,0.2),
-  pch=19)
-abline(h=0.05, lty=2, col="grey")
-legend(15,1, legend=c("P value", "BH-adjusted"), pch=19, col=c("blue", "red"))
-
-plot(x.all$diff.btw, x.all$we.ep, log="y", cex=0.7, col=rgb(0,0,1,0.2),
-  pch=19, xlab="Difference", ylab="P value", main="Volcano plot")
-points(x.all$diff.btw, x.all$we.eBH, cex=0.7, col=rgb(1,0,0,0.2),
-  pch=19)
-abline(h=0.05, lty=2, col="grey")
-
-x.all
-
-```
-
-```{r}
-## ANCOM
-```
-```{r}
-
-# Setup phyloseq objects
-
-ps_family <- phyloseq(otu_table(troph_data_family,
-                                taxa_are_rows = TRUE),
-                      sample_data(sample_data))
-ps_genus <- phyloseq(otu_table(troph_data_genus,
-                                taxa_are_rows = TRUE),
-                      sample_data(sample_data))
-
-# Run Ancom
-
-out_family = ancom(phyloseq = ps_family,  
-                  p_adj_method = "BH",
-                  prv_cut = 0.1, 
-                  lib_cut = 0, 
-                  main_var = "treat",
-                  rand_formula = NULL, 
-                  lme_control = NULL,
-                  struc_zero = TRUE, 
-                  neg_lb = FALSE, 
-                  alpha = 0.05, 
-                  n_cl = 1)
-
-out = ancom(phyloseq = ps_genus,  
-                  p_adj_method = "BH",
-                  prv_cut = 0.1, 
-                  lib_cut = 0, 
-                  main_var = "treat",
-                  rand_formula = NULL, 
-                  lme_control = NULL,
-                  struc_zero = TRUE, 
-                  neg_lb = FALSE, 
-                  alpha = 0.05, 
-                  n_cl = 1)
-
-```
-
-```{r}
-
-q_val = out$q_data
-beta_val = out$beta_data
-# Only consider the effect sizes with the corresponding q-value less than alpha
-beta_val = beta_val * (q_val < 0.05) 
-# Choose the maximum of beta's as the effect size
-beta_pos = apply(abs(beta_val), 2, which.max) 
-beta_max = vapply(seq_along(beta_pos), function(i) beta_val[beta_pos[i], i],
-                  FUN.VALUE = double(1))
-
-# Number of taxa except structural zeros
-n_taxa = ifelse(is.null(out$zero_ind), 
-                nrow(feature_table), 
-                sum(apply(out$zero_ind, 1, sum) == 0))
-
-# Cutoff values for declaring differentially abundant taxa
-cut_off = 0.7 * (n_taxa - 1)
-
-df_fig_w = out$res %>%
-  dplyr::mutate(beta = beta_max,
-                direct = case_when(
-                  detected_0.7 == TRUE & beta > 0 ~ "Positive",
-                  detected_0.7 == TRUE & beta <= 0 ~ "Negative",
-                  TRUE ~ "Not Significant"
-                  )) %>%
-  dplyr::arrange(W)
-df_fig_w$taxon_id = factor(df_fig_w$taxon_id, levels = df_fig_w$taxon_id)
-df_fig_w$W = replace(df_fig_w$W, is.infinite(df_fig_w$W), n_taxa - 1)
-df_fig_w$direct = factor(df_fig_w$direct, 
-                     levels = c("Negative", "Positive", "Not Significant"))
-
-p_w = df_fig_w %>%
-  ggplot(aes(x = taxon_id, y = W, color = direct)) +
-  geom_point(size = 2, alpha = 0.6) +
-  labs(x = "Taxon", y = "W") +
-  scale_color_discrete(name = NULL) + 
-  geom_hline(yintercept = cut_off, linetype = "dotted", 
-             color = "blue", size = 1.5) +
-  geom_text(aes(x = 2, y = cut_off + 0.5, label = "W[0.7]"), 
-            size = 5, vjust = -0.5, hjust = 0, color = "orange", parse = TRUE) +
-  theme_bw() +
-  theme(axis.text.x = element_blank(),
-        axis.ticks.x = element_blank(),
-        panel.grid.major = element_blank())
-p_w
-
-```
-
-```{r}
-
-otu_table <-  aes_data %>% 
-  filter(asv %in% troph_asvs) %>% 
-  filter(passage == "G5") %>% 
-  select(sample, count, Genus) %>% 
-  group_by(Genus, sample) %>% 
-  summarize(total = sum(count), .groups = "drop") %>% 
-  mutate(treat = substr(sample, 3, 3)) %>% 
-  group_by(Genus, treat) %>% 
-  mutate(prev = sum(total > 0)) %>% 
-  ungroup() %>% 
-  filter(prev > 2) %>% 
-  select(-treat, -prev) %>% 
-  pivot_wider(names_from = sample, values_from = total, values_fill = 0) %>% 
-  as.data.frame()
-
-rownames(otu_table) <- otu_table[, 1]
-otu_table <- otu_table[, -1]
-
-conds <- 
-  otu_table %>% 
-  colnames() %>% 
-  substr(3, 3)
-
-x.all <- aldex(otu_table, 
-               conds, mc.samples=1000, test="t", effect=TRUE,
-     include.sample.summary=FALSE, denom="all", verbose=FALSE, paired.test=FALSE)
-
-par(mfrow=c(1,2))
-aldex.plot(x.all, type="MA", test="welch", xlab="Log-ratio abundance",
-    ylab="Difference")
-aldex.plot(x.all, type="MW", test="welch", xlab="Dispersion",
-    ylab="Difference")
-
-par(mfrow=c(1,2))
-plot(x.all$effect, x.all$we.ep, log="y", cex=0.7, col=rgb(0,0,1,0.2),
-  pch=19, xlab="Effect size", ylab="P value", main="Effect size plot")
-points(x.all$effect, x.all$we.eBH, cex=0.7, col=rgb(1,0,0,0.2),
-  pch=19)
-abline(h=0.05, lty=2, col="grey")
-legend(15,1, legend=c("P value", "BH-adjusted"), pch=19, col=c("blue", "red"))
-
-plot(x.all$diff.btw, x.all$we.ep, log="y", cex=0.7, col=rgb(0,0,1,0.2),
-  pch=19, xlab="Difference", ylab="P value", main="Volcano plot")
-points(x.all$diff.btw, x.all$we.eBH, cex=0.7, col=rgb(1,0,0,0.2),
-  pch=19)
-abline(h=0.05, lty=2, col="grey")
-
-x.all_effect_1 <-
-  x.all %>% 
-  filter(abs(effect) > 0.5) %>% 
-  as_tibble(rownames = "asv")
-
-x.all_effect_1 %>% 
-  ggplot(aes(y = reorder(asv, effect), x = effect)) +
-  geom_point()
-
-```
-
-```{r aldex_genus_everything}
-
-otu_table <- 
-  aes_data %>% 
-  filter(passage == "G5") %>% 
-  select(Genus, count, sample) %>% 
-  group_by(Genus, sample) %>% 
-  summarize(total = sum(count), .groups = "drop") %>% 
-  mutate(treat = substr(sample, 3, 3)) %>% 
-  group_by(Genus, treat) %>% 
-  mutate(prev = sum(total > 0)) %>% 
-  ungroup() %>% 
-  filter(prev > 2) %>% 
-  select(-treat, -prev) %>% 
-  pivot_wider(names_from = "sample", values_from = "total", values_fill = 0) %>% 
-  as.data.frame()
-
-rownames(otu_table) <- otu_table[, 1]
-
-otu_table <- otu_table[, -1]
-
-x.all <- aldex(otu_table, 
-               conds, mc.samples=1000, test="t", effect=TRUE,
-     include.sample.summary=FALSE, denom="all", verbose=FALSE, paired.test=FALSE)
-
-par(mfrow=c(1,2))
-aldex.plot(x.all, type="MA", test="welch", xlab="Log-ratio abundance",
-    ylab="Difference")
-aldex.plot(x.all, type="MW", test="welch", xlab="Dispersion",
-    ylab="Difference")
-
-par(mfrow=c(1,2))
-plot(x.all$effect, x.all$we.ep, log="y", cex=0.7, col=rgb(0,0,1,0.2),
-  pch=19, xlab="Effect size", ylab="P value", main="Effect size plot")
-points(x.all$effect, x.all$we.eBH, cex=0.7, col=rgb(1,0,0,0.2),
-  pch=19)
-abline(h=0.05, lty=2, col="grey")
-legend(15,1, legend=c("P value", "BH-adjusted"), pch=19, col=c("blue", "red"))
-
-plot(x.all$diff.btw, x.all$we.ep, log="y", cex=0.7, col=rgb(0,0,1,0.2),
-  pch=19, xlab="Difference", ylab="P value", main="Volcano plot")
-points(x.all$diff.btw, x.all$we.eBH, cex=0.7, col=rgb(1,0,0,0.2),
-  pch=19)
-abline(h=0.05, lty=2, col="grey")
-
-x.all_effect_1 <-
-  x.all %>% 
-  filter(abs(effect) > 1) %>% 
-  as_tibble(rownames = "Genus")
-
-x.all_effect_1 %>% 
-  ggplot(aes(y = reorder(Genus, effect), x = effect)) +
-  geom_point()
-
-```
-
-```{r ancova-clr}
-
-gm <- function(x){
-
-  exp(mean(log(x[x>0])))
-
-}
-
-rclr <- function(x) {
-  log(x/gm(x))
-}
-
-aes_family <-
-  aes_data %>%
-  select(Family, count, sample, passage, treat) %>%
-  group_by(Family, sample) %>%
-  mutate(total = sum(count)) %>%
-  select(-count) %>%
-  ungroup() %>%
-  distinct() %>%
-  mutate(rep = as.numeric(substr(sample, 4, 5))) %>%
-  select(-sample)
-
-prevalent_families <- 
-  aes_family %>% 
-  mutate(passagetreat = paste0(passage, treat)) %>%
-  group_by(Family, passagetreat) %>% 
-  summarize(prev = sum(total > 0), .groups = "drop") %>% 
-  pivot_wider(names_from = "passagetreat", values_from = "prev", values_fill = 0) %>% 
-  filter(G2N/12 > 0.1 & G2P/12 > 0.1 & G5N/12 > 0.1 & G5P/12 > 0.1) %$%
-  Family
-
-aes_family_rclr <-
-  aes_family %>% 
-  filter(Family %in% prevalent_families) %>% 
-  group_by(passage, treat, rep) %>%
-  mutate(rclr = log(total/gm(total))) %>%
-  select(-total) %>%
-  ungroup() %>%
-  pivot_wider(names_from = passage, values_from = rclr)
-
-ancova <- function(df) {
-  lm(G5 ~ G2 + treat, data = df, na.action = )
-}
-
-model_results <-
-aes_family_rclr %>%
-  nest(data = -Family) %>%
-  mutate(model = map(data, ancova)) %>%
-  mutate(tidy = map(model, tidy)) %>% 
-  unnest(tidy) %>% 
-  filter(term == "treatP")
-
-model_results$p.adjust <- p.adjust(model_results$p.value, method = "BH")
-
-model_results %>%
-  filter(p.adjust < 0.05) %>% 
-  arrange(estimate) %>% 
-  print(n = 22)
-
-model_results %>% 
-  filter(p.adjust < 0.05) %>% 
-  arrange(estimate) %>% 
-  unnest(data) %>% 
-  ggplot(aes(x = estimate, xmax = estimate + std.error, xmin = estimate - std.error, y = reorder(Family, estimate))) +
-  geom_pointrange() +
-  labs(x = "Estimate \n(log-ratio increase in P)", y = NULL) +
-  theme_bw()
-ggsave("ancova-plot.png", width = 5, height = 5)
-
-model_results %>% 
-  filter(p.adjust < 0.05) %>% 
-  unnest(data) %>% 
-  pivot_longer(G2:G5, names_to = "passage", values_to = "rclr") %>% 
-  select(Family, treat, rep, passage, rclr) %>% 
-  distinct() %>% 
-  ggplot(aes(x = rclr, y = Family, color = treat)) +
-  facet_grid( ~ passage) +
-  geom_point(position = position_jitterdodge())
-
-```
-
-```{r}
-
-aes_family_data <-
-  aes_data %>% 
-  select(asv, sample, passage, treat, count, Family) %>%
-  group_by(sample, Family) %>% 
-  mutate(total = sum(count)) %>% 
-  ungroup() %>% 
-  select(sample, treat, passage, Family, total) %>% 
-  distinct()
-
-prevalent_families <- 
-  aes_family_data %>% 
-  group_by(Family, treat, passage) %>% 
-  summarize(prev = sum(total > 0), .groups = "drop") %>% 
-  pivot_wider(names_from = "treat", values_from = "prev", values_fill = 0) %>% 
-  filter(N/12 > 0.1 & P/12 > 0.1) %$%
-  Family
-
-aes_family_data <-
-  aes_family_data %>% 
-  filter(Family %in% prevalent_families) %>% 
-  mutate(jar = substr(sample, 3, 5)) %>%
-  select(-sample) %>%
-  pivot_wider(names_from = "passage", values_from = "total", values_fill = 0)
-
-ancova_model <- function(df) {
-	lm(G5 ~ G2 + treat, data = df)
-}
-
-aes_family_nested <-
-	aes_family_data %>%
-	group_by(Family) %>%
-	nest() %>%
-	mutate(model = map(data, ancova_model)) %>%
-	mutate(tidy = map(model, tidy)) %>%
-	ungroup()
-
-aes_ancova <- aes_family_nested %>%
-	unnest(tidy) %>%
-	filter(term == "treatP")
-
-aes_ancova$p.adjust <- p.adjust(aes_ancova$p.value, method = "BH")
-
-aes_ancova %>%
-	arrange(estimate) %>%
-	filter(p.adjust < 0.05) %>%
-	print(n = 22)
-
-```
-
-```{r}
-
-aes_family_data <-
-  aes_data %>% 
-  select(asv, sample, passage, treat, count, Family) %>%
-  group_by(sample, Family) %>% 
-  mutate(total = sum(count)) %>% 
-  ungroup() %>% 
-  select(sample, treat, passage, Family, total) %>% 
-  distinct()
-
-prevalent_families <- 
-  aes_family_data %>% 
-  group_by(Family, treat, passage) %>% 
-  summarize(prev = sum(total > 0), .groups = "drop") %>% 
-  pivot_wider(names_from = "treat", values_from = "prev", values_fill = 0) %>% 
-  filter(N/12 > 0.1 & P/12 > 0.1) %$%
-  Family
-
-otu_table <- 
-  aes_family_data %>% 
-  filter(Family %in% prevalent_families) %>% 
-  select(sample, Family, total) %>% 
-  pivot_wider(names_from = sample, values_from = total, values_fill = 0) %>% 
-  as.data.frame()
-
-rownames(otu_table) <- otu_table[,1]
-otu_table <- otu_table[,-1]
-
-sample_data <- aes_family_data %>% 
-  select(sample, treat, passage) %>% 
-  distinct() %>% 
-  as.data.frame()
-
-rownames(sample_data) <- sample_data[,1]
-sample_data <- sample_data[, -1]
-sample_data <-
-	sample_data %>%
-	mutate(passage12 = if_else(passage == "G2", 1, 2))
-
-# Setup phyloseq object
-aes_ps <- phyloseq(otu_table(otu_table, taxa_are_rows=TRUE), sample_data(sample_data))
-
-set.seed(123)
-out2 = ancom(phyloseq = aes_ps,  p_adj_method = "holm",
-             prv_cut = 0.10, lib_cut = 1000, main_var = "treat", 
-             adj_formula = "passage12", 
-             rand_formula = "~ passage12 | 1",
-             lme_control = list(maxIter = 100, msMaxIter = 100, opt = "optim"), 
-             struc_zero = TRUE, neg_lb = FALSE, alpha = 0.05, n_cl = 2)
-
-res2 = out2$res
-
-out2 = ancom(phyloseq = family_data2,  p_adj_method = "holm",
-             prv_cut = 0.10, lib_cut = 1000, main_var = "group", 
-             adj_formula = "nationality + timepoint + bmi_group", 
-             rand_formula = "~ timepoint | subject",
-             lme_control = list(maxIter = 100, msMaxIter = 100, opt = "optim"), 
-             struc_zero = TRUE, neg_lb = TRUE, alpha = 0.05, n_cl = 2)
-
-res2 = out2$res
-
-```
-
-```{r}
-
-x.all_effect_1$Family
-
-dt_data$taxa
-
-```
-
-```{r}
-combined_da <-
-  dt_data %>% 
-  as_tibble %>% 
-  full_join(sig_families, by = c("taxa" = "Family")) %>% 
-  full_join(x.all_effect_1, by = c("taxa" = "Family")) %>% 
-  full_join(df_fig_w[df_fig_w$detected_0.7,], by = c("taxa" = "taxon_id"))
-
-shared_da <-
-  dt_data %>% 
-  as_tibble %>% 
-  inner_join(sig_families, by = c("taxa" = "Family")) %>% 
-  inner_join(x.all_effect_1, by = c("taxa" = "Family")) %>% 
-  inner_join(df_fig_w[df_fig_w$detected_0.7,], by = c("taxa" = "taxon_id"))
-
-shared_da %>% 
-  ggplot(aes(y = reorder(taxa, effect), x = effect)) +
-  geom_point()
-
-aes_data %>% 
-  filter(passage == "G5") %>% 
-  select(sample, Family, count) %>% 
-  group_by(sample, Family) %>% 
-  summarize(total = sum(count)) %>% 
-  filter(Family %in% shared_da$taxa) %>% 
-  mutate(treat = substr(sample, 3, 3)) %>% 
-  ggplot(aes(x = total, y = Family, color = treat)) +
-  geom_jitter() +
-  scale_x_log10()
-
-rubini_seqs <-
-  aes_data %>% 
-  filter(Family == "Rubinisphaeraceae") %>% 
-  select(asv, seq, Kingdom:Species) %>% 
-  distinct()
-
-write(rubini_seqs$seq, file = 'Output/rubini_seqs.txt')
-
-```
-
-```{r}
-
-##############
-
-First, I looked at which taxa are unique to treatments N and P. Several families
-present in P and absent in N. Most of them had low prevalence (present in 3 or
-fewer samples) and low abundance (median < 5 reads across samples). There were two families unique
-to the Positive treatment with relatively high prevalence. This included an ASV
-that was a member of the Bacteroidia Class with no lower taxonomic designation. 
-This ASV had a prevalence of 10/12 and a median abundance of ~2 reads. The other
-prevalent family was a member of the Silvanigrellaceae, a newly described family placed in
-its own order. This family was present in
-all 12 samples and was represented by three OTUs: one with no lower designation, one from the genus
-Silvanigrella and one from the genus Candidatus Turbobacter. Silvanigrella was
-isolated from a temperate fresh water lake (https://www.microbiologyresearch.org/content/journal/ijsem/10.1099/ijsem.0.001965).
-
-Of the families unique to the Neutral treatment, only one had a prevalence greater than 2/12. 
-This family, Armatimonadaceae, was present in half of the Neutral samples (6/12). This
-family was represented by two ASVs in the genus Armatimonas. The type strain for
-Armatimonas was isolated from the rhizoplane of Phragmites australis (https://www.microbiologyresearch.org/content/journal/ijsem/10.1099/ijs.0.025643-0)
-```
-
-```{r}
-# families in to Passage 5, Treatment P
-g5p<- 
-  ps_fixed %>% 
-  ps_filter(treat == "p", passage == 5) %>% 
-  tax_agg('Family') %>% 
-  ps_get() %>% 
-  ps_melt()
-
-# families in to Passage 5, Treatment N
-g5n<- 
-  ps_fixed %>% 
-  ps_filter(treat == "n", passage == 5) %>% 
-  tax_agg('Family') %>% 
-  ps_get() %>% 
-  ps_melt()
-
-# Families unique to p in passage 5
-g5p_unique <-
-  g5p %>% 
-  filter(Family %in% unique(g5p$Family)[!(unique(g5p$Family) %in% unique(g5n$Family))]) %>% 
-  mutate(Prevalence = as.numeric(Abundance > 0))
-
-# Abundance
-Family_order<-
-  g5p_unique %>% 
-  group_by(OTU) %>% 
-  summarize(median = median(Abundance), max = max(Abundance)) %>% 
-  arrange(median, max)
-g5p_unique$OTU <- factor(g5p_unique$OTU, levels = Family_order$OTU)
-
-# Plot families unique to P5 in order of median abundance
-g5p_unique %>% 
-  ggplot(aes(y = Abundance, x = OTU)) +
-  geom_tufteboxplot() +
-  theme_tufte() +
-  geom_rangeframe() +
-  theme(axis.text.x = element_text(angle = 90)) +
-  scale_y_continuous(breaks = c(0, 5, 10, 15, 20, 24)) +
-  labs(title = "Unique to P5 by abundance")
-
-# Prevalence
-Family_order<-
-  g5p_unique %>% 
-  group_by(OTU) %>% 
-  summarize(sum = sum(Prevalence)) %>% 
-  arrange(sum)
-g5p_unique$OTU <- factor(g5p_unique$OTU, levels = Family_order$OTU)
-
-# Plot families unique to P5 in order of prevalence
-g5p_unique %>% 
-  group_by(OTU) %>% 
-  summarize(Prevalence = sum(Prevalence)) %>% 
-  ggplot(aes(y = Prevalence, x = OTU)) +
-  geom_point() +
-  theme_tufte() +
-  geom_rangeframe() +
-  theme(axis.text.x = element_text(angle = 90)) +
-  scale_y_continuous(breaks = c(1, 2, 3, 5, 10, 12)) +
-  labs(title = "Unique to P5 by prevalence")
-
-# Families unique to n in passage 5
-g5n_unique <-
-  g5n %>% 
-  filter(Family %in% unique(g5n$Family)[!(unique(g5n$Family) %in% unique(g5p$Family))]) %>% 
-  mutate(Prevalence = as.numeric(Abundance > 0))
-
-# Abundance
-Family_order<-
-  g5n_unique %>% 
-  group_by(OTU) %>% 
-  summarize(median = median(Abundance), max = max(Abundance)) %>% 
-  arrange(median, max)
-g5n_unique$OTU <- factor(g5n_unique$OTU, levels = Family_order$OTU)
-
-# Plot families unique to N5 in order of abundance
-g5n_unique %>% 
-  ggplot(aes(y = Abundance, x = OTU)) +
-  geom_tufteboxplot() +
-  theme_tufte() +
-  geom_rangeframe() +
-  theme(axis.text.x = element_text(angle = 90)) +
-  scale_y_continuous(breaks = c(0, 10, 20, 30, 37)) +
-  labs(title = "Unique to N5 by abundance")
-
-# Prevalence
-Family_order<-
-  g5n_unique %>% 
-  group_by(OTU) %>% 
-  summarize(sum = sum(Prevalence)) %>% 
-  arrange(sum)
-g5n_unique$OTU <- factor(g5n_unique$OTU, levels = Family_order$OTU)
-
-# Plot families unique to N5 in order of prevalence
-
-g5n_unique %>% 
-  group_by(OTU) %>% 
-  summarize(Prevalence = sum(Prevalence)) %>% 
-  ggplot(aes(y = Prevalence, x = OTU)) +
-  geom_point() +
-  theme_tufte() +
-  geom_rangeframe() +
-  theme(axis.text.x = element_text(angle = 90)) +
-  scale_y_continuous(breaks = c(1, 2, 6)) +
-  labs(title = "Unique to N5 by prevalence")
-
-```
-
-```{r}
-# Differential Abundance Treatment in passage 5
-```
-
-```{r corncob_treatment, results=FALSE, fig.hight=40}
-
-ps_fixed_family <- speedyseq::tax_glom(ps_fixed, taxrank="Family")
-
-ps_fam_pas5 <- microViz::ps_filter(ps_fixed_family, passage == 5)
-
-ps_fam_pas5_prev <- microViz::tax_filter(ps_fam_pas5, min_prevalence = 0.10)
-
-sample_data(ps_fam_pas5_prev) =
-  sample_data(ps_fam_pas5_prev)[,c('sample_id', 'treat')] %>% 
-  sample_data()
-ps = ps_fam_pas5_prev
-
-# Differential abundance test
-da_analysis <- differentialTest(formula = ~ treat,
-                                 phi.formula = ~ 1,
-                                 formula_null = ~ 1,
-                                 phi.formula_null = ~ 1,
-                                 test = "Wald", boot = FALSE,
-                                 data = ps_fam_pas5_prev,
-                                 fdr_cutoff = 0.05, B = 50)
-
-#plot(da_analysis, level = c("Family"))
-da_analysis$significant_models
-dt_data <- plot(da_analysis, level = c("Family"), data_only = TRUE)
-
-sig_taxa_treat <- tax_select(ps_fam_pas5_prev, da_analysis$significant_taxa)
-ntaxa(sig_taxa_treat)
-save(da_analysis, sig_taxa_treat, dt_data, file = 'Output/da_corncob.RData')
-
-```
-
-```{r corncob-plot, fig.width=6, fig.height=6}
-dt_plot(dt_data)
-
-```
-
-```{r corncob_ch4}
-
-ps_fam_pas5_trep <- ps_filter(ps_fam_pas5, treat == "p")
-
-ps_fam_prev_trep_prev <- tax_filter(ps_fam_pas5_trep, min_prevalence = 0.10)
-
-da_analysis <- differentialTest(formula = ~ ch4,
-                                 phi.formula = ~ ch4,
-                                 formula_null = ~ 1,
-                                 phi.formula_null = ~ ch4,
-                                 test = "Wald", boot = FALSE,
-                                 data = ps_fam_prev_trep_prev,
-                                 fdr_cutoff = 0.05)
-otu_to_taxonomy(OTU = da_analysis$significant_taxa, data = ps_fam_prev_trep_prev)
-
-plot(da_analysis, level = c("Family"))
-
-# Check unfitted taxa
-# otu_to_taxonomy(which(is.na(da_analysis$p)) %>% names, data = ps_gen_prev)
-# otu_table(ps_gen_prev)[,which(is.na(da_analysis$p))]
-
-# Differential variability test
-
-dv_analysis <- differentialTest(formula = ~ passage + treat + ch4,
-                                 phi.formula = ~ passage + treat + ch4,
-                                 formula_null = ~ passage + treat + ch4,
-                                 phi.formula_null = ~ passage + treat,
-                                 data = ps_gen_prev,
-                                 test = "Wald", boot = FALSE,
-                                 fdr_cutoff = 0.05)
-# otu_to_taxonomy(OTU = dv_analysis$significant_taxa, data = ps_gen_prev)
-plot(dv_analysis, level = c("Family", "Genus"))
-
-# Check unfitted taxa
-
-# otu_to_taxonomy(which(is.na(dv_analysis$p)) %>% names, data = ps_gen_prev)
-# otu_table(ps_gen_prev)[,which(is.na(dv_analysis$p))]
-
-```

### 2022-09-14 packages.R

@@ -3,3 +3,14 @@ library(devtools)
 # Install specific commit/version from github
 install_github("amorris28/morris", dependencies = FALSE)
 install_github("kbroman/broman", dependencies = FALSE)
+install_github("FrederickHuangLin/ANCOMBC", dependencies = FALSE)
+
+# if (!requireNamespace("BiocManager", quietly = TRUE))
+#   install.packages("BiocManager")
+
+renv::install("bioc::ALDEx2")
+renv::install("bioc::phyloseq")
+renv::install("bioc::ANCOMBC")
+renv::install("bioc::dada2")
+renv::install("bioc::decontam")
+

### 2022-09-16 Manuscript/manuscript.Rmd
remove packages no longer used and update renv lockfile

@@
 library(morris)
-library(phyloseq)
+# library(phyloseq)
 library(ggpubr)
@@
 library(corncob)
-library(ggthemes)
-#library(microViz)
+# library(ggthemes)
 library(captioner)
 library(knitr)
-library(kableExtra)
-library(flextable)
 library(ggtext)

### 2022-09-16 R/08-community-analysis.Rmd
remove packages no longer used and update renv lockfile

@@
 library(ggpubr)
-library(ggthemes)
 library(broom)

### 2022-09-16 R/functions.R

@@ -2,16 +2,12 @@
 raw_dir <- 'Data/' # Raw Data Directory
 der_dir <- 'Output/' # Derived/Modified Data Directory
 
-# theme_set(theme_tufte()+
-#             theme(axis.title.y = element_text(angle = 0, vjust = 0.5)))
 theme_set(theme_classic() +
             theme(panel.border = element_rect(fill = NA, size = 1.0625),
                   axis.line = element_blank()))
 
 plot_flux <- function(fluxes, passage, estimate, ratio = NULL, log10 = FALSE) {
   p <- ggplot(fluxes, aes(x = {{passage}}, y = {{estimate}}, color = treat)) +
-    # theme_classic() +
-    # geom_rangeframe(color = 'black') +
     stat_smooth(method = 'lm', se = FALSE, formula = 'y ~ x') +
     scale_color_manual(
       name = "Treatment", 
@@ -58,8 +54,6 @@ print_lrt <- function(fit, term = 2, test = "Likelihood ratio test") {
 plot_herit <- function(heritability, ratio) {
   ratio <- calc_plot_ratios(heritability$selected, heritability$offspring)
   ggplot(heritability, aes(selected, offspring, color = treat)) +
-    # theme_classic() +
-    # geom_rangeframe(color = 'black') +
     geom_jitter(width = 0.01) + 
     stat_smooth(method = 'lm', se = FALSE, formula = 'y ~ x') +
     labs(x = expression(Parental ~ log[10] ~ CH[4] ~ "(-k)"), 
@@ -85,8 +79,6 @@ plot_beta <- function(ps) {
     ord_plot(color = "treat",
              shape = "passage",
              auto_caption = NA) +
-    # theme_tufte() +
-    # geom_rangeframe(color = 'black') +
     coord_fixed(5.5/43.6, clip="off") +
     scale_shape_discrete(name = "Passage") +
     scale_color_manual(name = "Treatment",
@@ -95,8 +87,6 @@ plot_beta <- function(ps) {
     theme_classic() +
     theme(panel.border = element_rect(fill = NA, size = 0.5),
           axis.line = element_blank())  
-  # theme_tufte() +
-  #   theme(axis.title.y = element_text(angle = 0, vjust = 0.5))
   
 
   # ps.prop <- transform_sample_counts(ps, function(otu) otu/sum(otu))

### 2022-10-19 Manuscript/manuscript.Rmd
update everything for manuscript draft to lab group

@@
 to determine the relationship between microbiome composition and the rate of
-any ecosystem function. This challenge remains because it is difficult to
-manipulate microbiome composition directly, we cannot know a priori which members of the
-microbiome limit the rate of an ecosystem function, and microbiomes covary with
-the environment. We propose that artificial ecosystem selection
-can overcome these challenges. Artificial ecosystem selection involves creating
-replicate ecosystems, selecting replicates with a desired ecosystem-level
-trait, and using the microbiomes from the selected ecosystems to inoculate a
-new set of ecosystems. We performed artificial ecosystem selection on replicate
-soil microcosms to select for increased soil methane oxidation rate. We
-observed a strong response to selection with a 51% increase in methane oxidation rate per round over five
-rounds of selection. This suggests that a large portion of the variation in
-methane oxidation rate in these soils can be attributed to microbiome
-variation. We also found that selection did not enrich for known methane
-oxidizers; rather microbes that responded to selection included `r length(da_fams)`
-families such as *Fimbriimonadaceae*, *Cytophagaceae*, and *Diplorickettsiaceae*. This result
-is in contrast to the typical assumption that the rate of an ecosystem function
-is limited by the final step in the associated microbial pathway. Our study
-demonstrates that variation in microbiome composition can contribute to
-variation in the rate of ecosystem function.
+ecosystem functions. This challenge remains because it is difficult to
+manipulate microbiome composition directly, we often cannot know a priori which
+members of a microbiome influence the rate of an ecosystem function, and
+microbiomes can covary strongly with other drivers of ecosystem function, such
+as the environment. We propose that artificial ecosystem selection can overcome
+these challenges. Artificial ecosystem selection involves creating replicate
+ecosystems, selecting replicates with a desired ecosystem-level trait, and
+using the microbiomes from the selected ecosystems to inoculate a new set of
+ecosystems. We performed artificial ecosystem selection on replicate soil
+microcosms to select for increased soil methane oxidation rate. We observed a
+strong response to selection with a 51% increase in methane oxidation rate per
+round over five rounds of selection. This suggests that a large portion of the
+variation in methane oxidation rate in these soils can be attributed to
+microbiome variation. We also found that selection did not enrich for known
+methane oxidizers; rather microbes that responded to selection included members
+of `r length(da_fams)` families such as *Fimbriimonadaceae*, *Cytophagaceae*,
+and *Diplorickettsiaceae*. This result is in contrast to the typical assumption
+that the rate of an ecosystem function is limited by the final step in the
+associated microbial pathway. Our study demonstrates that variation in
+microbiome composition can contribute to variation in the rate of ecosystem
+function and that this relationship can be revealed via artificial ecosystem
+selection.
 
@@
 is great interest in understanding how attributes of microbiomes may influence
-variation in ecosystem functions [@crowther2019]. Biodiversity-ecosystem
+variation in ecosystem functions [@conrad1996; @schimel1998; @crowther2019]. Biodiversity-ecosystem
 function relationships have been described for a variety of macroorganismal
@@
 consumption, and nutrient release
-[@cardinale2006; @hooper2012; @gamfeldt2015]. While there is evidence that
-variation in the microbiome contributes to variation in ecosystem functions,
-describing this relationship has been elusive [@rocca2015; @graham2016].
+[@cardinale2006; @hooper2012; @gamfeldt2015]. While there is some evidence that
+variation in microbiomes can contribute to variation in ecosystem functions,
+documenting this relationship has been elusive [@conrad1996; @rocca2015; @graham2016].
 
-One reason for this difficulty is that it is challenging to directly manipulate
-microbiome composition and diversity the way we can for macroorganismal
-communities. Instead, we must use genetics through marker gene or metagenomic
-sequencing to characterize the microbiome.  However, studies that measure the
-taxonomic or functional diversity of a microbiome through these approaches
-rarely find a correlation between diversity and function [@graham2016]. Another
-approach is to quantify cells or genes using  quantitative
-PCR. Studies such as this still rarely find a correlation between the abundance
-of a functional group and the rate of the corresponding process [@rocca2015].
-As a result, it remains unclear how and when microbiome variation contributes
-to variation in the rate of ecosystem functions.
+There are several reasons for this difficulty. One reason is that it is
+challenging to manipulate microbiome composition and diversity directly.
+Experimental manipulations of community composition and diversity have been
+central to documenting the relationships between ecosystem functions and the
+biodiversity of plants and animals [@hooper2012], but this is seldom possible
+with microbiomes. Most microorganisms cannot yet be cultured and most
+available methods for manipulation of microbiomes (e.g., antibiotics, dilution,
+etc.) are not sufficiently specific to manipulate individual microbial
+lineages. Microbial ecologists are thus often restricted to using comparative
+methods, in which microbiome composition and diversity is surveyed via
+environmental DNA and then correlated with the rate of an ecosystem function
+{cite examples}. However, such studies are rarely successful [@graham2016]. In an
+attempt to increase the likelihood of detecting such a correlation, microbial
+ecologists may focus their survey efforts on a particular group of microbes
+thought to be involved in the ecosystem function of interest. For example, many
+studies have looked for relationships between an ecosystem function and the
+abundance or diversity of a microbial gene that codes for an enzyme or
+enzyme-subunit known to be involved in a particular function. But this approach
+assumes that one can identify the “right” marker gene a priori which, given how
+little is understood about microbial diversity, seems highly unlikely. And in
+fact these studies are rarely successful [@rocca2015].
 
-Another likely reason that microbiome-ecosystem function relationships have been
-difficult to document is that the most common approaches for looking for a
-relationship requires some prior knowledge of the likely causal links between
-microbiomes and ecosystem functions. For example, many studies have looked for
-relationships with the abundance or diversity of a "marker gene" for a
-particular function. Marker genes are microbial genes that code for an enzyme or
-enzyme-subunit known to be involved in a particular function. But this approach
-assumes that one can identify the "right" marker gene a priori which, given
-how little is understood about microbial diversity, seems highly unlikely. What
-is needed are approaches to microbiome-ecosystem
-function mapping that do not require this degree of prior knowledge.
-
-Yet another likely reason that microbiome-ecosystem function relationships have
-been difficult to document is that microbiomes can be related to ecosystem
-functions in two related but distinct ways that historically have been difficult
-to separate. One way is that microbiomes may simply be conduits through which
-the environment drives function. That is to say that the environmental
-conditions completely regulate the attributes of the microbiome and a shift in
-the environment causes a shift in microbiome attributes, such as the abundance
-of a microbial functional group. This shift results in a change in the rate of
-an ecosystem function. In this case, microbiome attributes would be
-statistically associated with variation in ecosystem function, but also strongly
-covary with environmental conditions. Therefore, there would be no association
-between microbiomes and ecosystem functions independent of the environmental
-conditions. Under this scenario, knowing anything about the microbiome would not
-necessarily provide you with information regarding the causes of the change in
-ecosystem function or improve one's ability to predict changes in ecosystem
-function. Furthermore, attempting to alter ecosystem function by altering the
-microbiome without changing the underlying environmental conditions would likely
-be fruitless.
-
-Another possibility is that the microbiome itself alters the rate of ecosystem
-function independent of the environment. In this case, one cannot predict the
-rate of ecosystem function without understanding changes in the microbiome. In
-addition, managing the microbiome through inoculation or selective antibiotics
-would potentially be an effective approach for enhancing or mitigating certain
-changes to ecosystem functions. These two scenarios are not mutually
-exclusive, nor are they likely unique to microbes, but methods to determine the
-relative importance of these two scenarios have not until recently been employed.
+Another likely reason that microbiome-ecosystem function relationships have
+been difficult to document is that microbiomes often covary with other drivers
+of variation in ecosystem functions, such as environmental conditions or
+resources. Such covariation can make it impossible to determine whether
+microbiomes directly contribute to functional variation. Understanding the
+degree to which microbiomes directly influence functional variation is crucial
+if scientists are to develop microbiome-focused methods for altering or
+maintaining ecosystem functions.
 
@@
 phenotypic traits in organisms, and solutions to this problem may be found by
-exploring this analogy [@morris2020]. For example, one approach to the problem of
-genotype-phenotype mapping is to perform artificial selection on a trait of
+exploring this analogy [@morris2020]. For example, one approach to the problem
+of genotype-phenotype mapping is to perform artificial selection on a trait of
 interest and identify genes that respond to selection. Such genes are likely to
@@
 exploring microbiome-ecosystem function relationships is gaining acceptance
-among microbiologists [@swenson2000; @panke-buisse2015]. Microbiomes have been
-shown to respond to selection at the ecosystem-trait level [@swenson2000],
-and some ecosystem-scale microbiome traits have been shown to be transferable
-through microbiome inoculations [@panke-buisse2015]. Applying artificial
-selection to whole microbiomes could be a novel approach for quantifying how
-much of the variation in the rate of an ecosystem function could be attributed
-independently to microbiome variation [@goodnight2000]. In addition, this
-approach could be used to identify microbiome attributes, such as genes or taxa,
-associated with the rate of an ecosystem function, although, to our knowledge
-this has not previously been accomplished [@arias-sanchez2019; @morris2020]. 
-Such an approach has the potential to expand our ability to model
-microbial ecosystem-function relationships more accurately and allow us to
-manage ecosystems for particular outcomes [@fuhrman2009; @schimel1998]. Here, we
-use this approach to estimate how much of the variation in ecosystem function
-(the flux of CH~4~ gas between the soil and atmosphere) can be attributed to
-variation in the microbiome and to identify microbial markers associated with
-this ecosystem function.
+among microbiologists [@swenson2000; @panke-buisse2015; @sanchez2021].
+Microbiomes have been shown to respond to selection at the ecosystem-trait
+level [@swenson2000], and some ecosystem-scale microbiome traits have been
+shown to be transferable through microbiome inoculations [@panke-buisse2015].
+Applying artificial selection to whole microbiomes could be a novel approach
+for quantifying how much of the variation in the rate of an ecosystem function
+could be attributed independently to microbiome variation [@goodnight2000]. In
+addition, this approach could be used to identify microbiome attributes, such
+as genes or taxa, associated with the rate of an ecosystem function, although,
+to our knowledge this has not previously been accomplished [@arias-sanchez2019;
+@morris2020]. Such an approach has the potential to expand our ability to more
+accurately model microbial ecosystem-function relationships and allow us to
+manage ecosystems for particular outcomes [@fuhrman2009; @schimel1998].
 
-Soil CH~4~ oxidation is a suitable function for this study because there is
-evidence that it may vary with microbiome attributes. For example, variation in
-CH~4~ emissions in arctic permafrost is correlated with the transcriptional
-activity of certain methanogens [@freitag2009; @freitag2010]. In addition,
-soil microbiome phylogenetic variation is a strong predictor of CH~4~ oxidation in
-forests and pastures of the Brazilian Amazon [@meyer2020]. Finally,
-methanogenesis and methanotrophy are two of the most deeply conserved microbial
-physiologies and are represented in a narrow range of taxa and so the taxonomic
-composition of the microbiome is more likely to be associated with the rate of
-CH~4~ flux than other broader or more shallowly conserved functions
-[@schimel1998; @martiny2013]. This suggests that microbiome variation might
-drive variation in CH~4~ oxidation in soils.
+We use this approach to estimate how much of the variation in an ecosystem
+function - the flux of CH~4~ gas between the soil and atmosphere - can be
+attributed to variation in the microbiome and to identify microbial markers
+associated with this ecosystem function.  Soil CH~4~ oxidation is a suitable
+function for this study because there is evidence that it may vary with
+microbiome attributes. For example, variation in CH~4~ emissions in arctic
+permafrost is correlated with the transcriptional activity of certain
+methanogens [@freitag2009; @freitag2010]. In addition, soil microbiome
+phylogenetic variation is a strong predictor of CH~4~ oxidation in forests and
+pastures of the Brazilian Amazon and the Congo Basin [@meyer2020; @meyer2020a]. Finally, methanogenesis and
+methanotrophy are two of the most deeply conserved microbial physiologies and
+are represented in a narrow range of taxa and so the taxonomic composition of
+the microbiome is more likely to be associated with the rate of CH~4~ flux than
+other broader or more shallowly conserved functions [@schimel1998;
+@martiny2013]. This suggests that microbiome variation might drive variation in
+CH~4~ oxidation in soils.
 
@@
 of CH~4~ oxidation rate, we compared the composition of the artificially
-selected microbiomes to a control set of microbiomes without selection.  We
+selected microbiomes to a control set of microbiomes with random selection.  We
 then evaluated whether these markers met our underlying assumptions about
@@
 of the homogenized soil was used to inoculate the next set of jars, which
-represents a 10% subsampling. The
-next set of jars were created in an identical manner to the first generation
-with fresh autoclaved potting mix and the same moisture and CH~4~ content.
+represents a 10% subsampling. The next set of jars were created in an identical
+manner to the first generation with fresh autoclaved potting mix and the same
+moisture and CH~4~ content.
 
@@
 Methane oxidation rates were determined after flushing and spiking jars to
-approximately 1000 ppm CH~4~. Headspace samples of 1 mL were collected from each
-jar immediately after spiking and then at time points 3, 6, 24, and 48 hours for
-a 5-point curve. Samples were immediately injected into a SRI model 8610C gas
-chromatograph equipped with a flame ionization detector (SRI Instruments, Torrance, CA, USA) to determine the headspace CH~4~
-concentration. Fluxes were calculated from a first-order exponential decay
-function as decay constant $k$ with units day^-1^. Oxidation rates are presented
-as the additive inverse of $k$ (i.e., $-k$) so that a more positive value
-represents a greater oxidation rate. The jars selected for the positive
-treatment in passage 2 had the lowest CH~4~ oxidation rate of the twelve jars
-due to a calculation error in the CH~4~ oxidation rate. All other passages
-correctly used the top three jars.
+approximately 1000 ppm CH~4~. Headspace samples of 1 mL were collected from
+each jar immediately after spiking and then at time points 3, 6, 24, and 48
+hours for a 5-point curve. Samples were immediately injected into a SRI model
+8610C gas chromatograph equipped with a flame ionization detector (SRI
+Instruments, Torrance, CA, USA) to determine the headspace CH~4~ concentration.
+Fluxes were calculated from a first-order exponential decay function as decay
+constant $k$ with units day^-1^. Oxidation rates are presented as the additive
+inverse of $k$ (i.e., $-k$) so that a more positive value represents a greater
+oxidation rate. The jars selected for the positive treatment in passage 2 had
+the lowest CH~4~ oxidation rate of the twelve jars due to a calculation error
+in the CH~4~ oxidation rate. Despite this error, we still observed a strong
+response to selection demonstrating the power of this technique. In addition,
+this likely reintroduced diversity lost during the selection process,
+potentially increasing the variation available to be selected upon, as
+is recommended for microbiome selection experiments [@sanchez2021].
 
@@
 plots indicated that `r length(contaminants)` of these were likely contaminants and these
-ASVs were removed.
+ASVs were removed. Amplicon sequence variants that were assigned chloroplast taxonomy were removed prior to analysis.
 
@@
 the microbiome, we estimated narrow-sense heritability ($h^2$) as the
-regression of mid-offspring on mid-parent [@falconer1996]. The mid-parent was the
-mean for all three selected jars and the mid-offspring was the mean for all jars
-produced by those parents. First, we tested if there was an effect of treatment
+regression of offspring on mid-parent [@falconer1996]. The mid-parent was the
+mean for all three selected jars and the offspring was the CH~4~ oxidation rate for each jar produced by those parents. First, we tested if there was an effect of treatment
 on the heritability estimate. We compared nested models with and without treatment using an F-test
@@
 of less than 10\% in either treatment. We used three methods for testing differential abundance: ANCOM-II, ALDEx2, and CORNCOB. We then
-pulled out the consensus taxa that were significant
+identified the consensus taxa that were significant
 with all three tests and plotted their relative abundances.
@@
 rate (`r figures("response", display = "cite")`; difference of slopes:
-F~`r compare_response_mods[2,3]`,`r compare_response_mods[2,1]`~ = `r myround(compare_response_mods[2,5], 2)`, p = `r myround(compare_response_mods[2,6], 2)`).
+F~`r compare_response_mods[2,4]`,`r compare_response_mods[2,2]`~ = `r myround(compare_response_mods[2,5], 2)`, p = `r myround(compare_response_mods[2,6], 2)`).
 At the start of the experiment, the Positive
@@
 
-## Heritability of methane oxidation
-
-We estimated heritability as the regression of mid-offspring on mid-parent (`r figures("herit", display = "cite")`).
-Offspring CH~4~ oxidation rates were correlated with parental CH~4~ oxidation
-rates in both the Positive treatment `r print_lm(herit_model, 4)` and the
-Neutral treatment `r print_lm(herit_model, 2)`.
+We estimated the proportion of variation in methane oxidation due to variation
+in the microbiome as the regression of offspring on mid-parent 
+(`r figures("herit", display = "cite")`).  Offspring CH~4~ oxidation rates were
+correlated with parental CH~4~ oxidation rates in both the Positive treatment
+`r print_lm(herit_model, 4)` and the Neutral treatment `r print_lm(herit_model,
+2)`.
 
@@
 
-# New Discussion
+# Discussion
 
-We observed a response to selection on whole-ecosystem soil methane oxidation rate, which demonstrates that variation in the microbiome can generate variation in the rate of ecosystem function independent of the underlying environmental conditions.
-While it is well understood that the microbiome mediates a variety of ecosystem functions, a fundamental question in microbial ecology is whether we need to understand variation in the composition of microbiomes to understand variation in the rate of ecosystem functions.
-To address that question, we performed artificial ecosystem selection on soil microbiomes by selecting for ecosystems with high methane oxidation rates.
-We found that there was a significant increase in methane oxidation rate as a response to selection.
-Each generation of soil ecosystems was created from the same pool of sterile soil substrate and a small inoculum from the previous generation.
-Therefore, it is unlikely that this response to selection was due to changes in the abiotic conditions of the ecosystem.
-In addition, the response was only observed in the Positive treatment and not in the Neutral treatment, which shows that this was not a side effect of passaging in jars.
-Therefore, we conclude that variation in the microbiome can generate variation in ecosystem functions independent of the environment.
+While it is well understood that the microbiome mediates a variety of ecosystem
+functions, a fundamental question in microbial ecology is whether we need to
+understand variation in the composition of microbiomes to understand variation
+in the rate of ecosystem functions [@conrad1996; @schimel1998]. To address this question, we performed
+artificial ecosystem selection on soil microbiomes by selecting for ecosystems
+with high methane oxidation rates. We found that there was a significant
+increase in methane oxidation rate as a response to selection. Each generation
+of soil ecosystems was created from the same pool of sterile soil substrate and
+a small inoculum from the previous generation. Therefore, it is unlikely that
+this response to selection was due to changes in the abiotic conditions of the
+ecosystem. In addition, the response was only observed in the Positive
+treatment and not in the Neutral treatment, which shows that this was not a
+side effect of passaging in jars.  Therefore, we conclude that variation in the
+microbiome can generate variation in ecosystem functions independent of the
+environment.
 
-In addition, we observed a large heritability for methane oxidation rate, which suggests that microbiome variation explains a large portion of the variation in methane oxidation rate relative to environmental variation in this system.
-To estimate how much of the variation in methane oxidation rate can be explained by the presence, absence, and relative abundances of specific microbial taxa, we calculated the narrow-sense heritability of methane oxidation rate, which was `r myround(coefficients(herit_model)[4], 2)` for the Positive treatment.
-A heritability of 1 would indicate that a trait is entirely genetically determined and thus offspring would have the exact same value of that trait as their parents.
-A heritability greater than 1 indicates that offspring exceed the value of their parents.
-This is unlikely in the context of a genomic trait, but microbiomes are not limited by the constraints of genomes because microbes are not alleles and communities are not chromosomes.
-Therefore, complementarity within a microbial community could result in offspring communities that exceed the trait values of their "parent" ecosystems.
-The large heritability we observed indicates that microbiome variation is a large contributor to the variation in methane oxidation in this experimental system.
+In quantitative genetics, heritability quantifies the degree to which variation
+in a trait is due to variation in genotype [@falconer1996]. This is often estimated as the
+regression of offspring phenotype on parental phenotype. Thus, a significant
+correlation between the trait values of parents and offspring indicates that
+the trait is at least partially genetically determined. To estimate the
+variation in an ecosystem trait due to variation in the microbiome, we can
+perform a similar estimation by comparing the trait values of the "offspring"
+jars in one generation to the trait value of the "parental" jars chosen for
+selection in the previous generation. This "ecosystem heritability" then
+quantifies how much of the variation in ecosystem function is due to variation
+in the microbiome [@goodnight2000; @sanchez2021; @henry2021].
 
-Next, we wanted to address what aspects of the microbiome, in terms of alpha diversity, beta diversity, or individual taxa, explain this contribution of the microbiome to variation in methane oxidation rate.
-There are three ways that the microbiome could respond to selection in this experiment: gain or loss of species, changes in the relative abundances of species, or changes within the genomes of the constituent species.
-Using 16S rRNA gene sequences, we are unable to address the third possibility of whether taxa in this experiment could have evolved changes in their genomes as a result of selection.
+We quantified how much of the variation in methane oxidation
+rate was due to variation in the microbiome by regressing the methane oxidation
+rates of the 12 offspring jars against the mean methane oxidation rate of the
+three jars selected in the previous generation 
+We observed a heritability of `r myround(coefficients(herit_model)[4], 2)` (add SE/CI) for the
+Positive selection treatment. 
+A heritability of 1 would indicate that a trait is entirely
+genetically determined and thus offspring would have the exact same value of
+that trait as their parents. A heritability greater than 1 indicates that
+offspring exceed the value of their parents. This is unlikely in the context of
+a genomic trait, but microbiomes are not limited by the constraints of genomes
+because microbes are not alleles and communities are not chromosomes.
+Therefore, complementarity within a microbial community could result in
+offspring communities that exceed the trait values of their "parent"
+ecosystems. The large heritability we observed indicates that microbiome
+variation is a large contributor to variation in methane oxidation in this
+experimental system.
+
+Next, we wanted to address what aspects of the microbiome, in terms of alpha
+diversity, beta diversity, or individual taxa, explain this contribution of the
+microbiome to variation in methane oxidation rate. There are three ways that
+the microbiome could respond to selection in this experiment: gain or loss of
+species, changes in the relative abundances of species, or changes within the
+genomes of the constituent species. Using 16S rRNA gene sequences, we are
+unable to address the third possibility of whether taxa in this experiment
+could have evolved changes in their genomes as a result of selection.
 Therefore, we will focus on the first two possibilities.
 
-Richness at the ASV level did not vary between the two treatments and there were relatively few taxa gained or lost in the Positive selection treatment and none of these were prevalent across the 12 jars in passage 5.
-Therefore, the gain or loss of species is unlikely to explain the increase in methane oxidation rate.
-However, we found that Bray-Curtis dissimilarity was greater between the two treatments in Passage 5 than within each treatment and was correlated with methane oxidation rate, which suggests that changes in the relative abundance of taxa could explain the response to selection.
-A difference in composition between the two treatments could be a direct result of selection for organisms that contribute to CH~4~ oxidation or a result of ecological drift over the course of the experiment.
-In the former case, that would suggest that changes in the relative abundance of specific microbial taxa could explain the response to selection on CH~4~ oxidation rate.
+Richness at the ASV level did not vary between the two treatments and there
+were relatively few taxa gained or lost in the Positive selection treatment and
+none of these were prevalent across the 12 jars in passage 5. Therefore, the
+gain or loss of species is unlikely to explain the increase in methane
+oxidation rate. However, we found that Bray-Curtis dissimilarity was greater
+between the two treatments in Passage 5 than within each treatment and was
+correlated with methane oxidation rate, which suggests that changes in the
+relative abundance of taxa could explain the response to selection.
 
-Even though we observed an increase in methane oxidation rate in the Positive treatment and a difference in composition between the two treatments, we did not observe an increase in the relative abundance of methanotrophs.
-This contrasts with the usual assumption that the final step in a metabolic pathway is the rate-limiting step (citation?).
-In certain ecosystems, methane oxidation rate is correlated with the abundance of methanotrophs as estimated from marker genes (Cite.
-I know that one methanogenesis paper, not sure about methanotrophy).
-However, it does not appear generally true that the rate of an ecosystem function is limited by the abundance of the gene that encodes the final step in that pathway (cite rocca, maybe some of the methane work.) Since we did not observe an increase in the relative abundance of methanotrophs in the treatment that displayed an increase in methane oxidation rate, we may have to consider other microbial functional groups within an ecosystem to understand what drives changes in the rate of ecosystem functions.
+Even though we observed an increase in methane oxidation rate in the Positive
+treatment and a difference in composition between the two treatments, we did
+not observe an increase in the relative abundance of methanotrophs. This was
+surprising given that methane consumption is not a common trait among microbes
+and that it is often assumed that the rate of an ecosystem function is
+limited by the final enzymatic step in a metabolic pathway
+[@rocca2015]. In certain ecosystems, methane production and consumption are 
+correlated with the abundance of methanogens and methanotrophs as estimated from marker
+genes [@meyer2020, @meyer2020a]. However, it does not appear generally true that the rate of an
+ecosystem function is limited by the abundance of the gene that encodes the
+final step in that pathway [@rocca2015]. Our results suggest that the rate of methane oxidation can be altered by
+non-methanotrophs, perhaps through ecological interactions with methanotrophic
+species.
 
-One of the goals of this study was to determine how much of the variation in ecosystem function is due to variation in the microbiome.
-In this experiment, we eliminated as much environmental variation as possible.
-This was intentional, because we wanted to determine whether it was possible for variation in the microbiome to generate variation in methane oxidation rate independent of the environment.
-As a result, our heritability estimates are likely to be greater than they would be in natural ecosystems with greater environmental variation.
-Future work could estimate heritability in experimental systems that manipulate environmental variation or across natural ecosystems to evaluate how much of the variation in ecosystem function can be explained by microbiome variation outside the lab.
-This experiment also had a relatively small sample size which resulted in heritability estimates with wide error bars.
-Increasing replication would help us more accurately estimate this parameter and give us a better understanding of the relative importance of environmental variation and microbiome variation in the rate of microbiome-mediated ecosystem functions.
+Another important question is whether selection at the ecosystem scale resulted
+in within-strain evolution. Using 16S rRNA gene sequencing we are unable to
+resolve genomic changes within taxa that may have undergone evolution. 
+Therefore, it is possible that the response to selection was driven by changes 
+within the genomes of methanotrophs. But given
+the slow growth rate of methanotrophs, often requiring weeks to years of
+serial batch culturing for isolation, this seems like an unlikely explanation 
+[@whittenbury1970; @hirayama2014; @kim2018]. Future
+work could analyze an artificial ecosystem selection experiment using
+genome-resolved metagenomics to try to detect strain-level changes in the
+constituent taxa. However, reconstructing genomes from in situ soil samples
+has been difficult to achieve and may require enrichment techniques, particularly for methanotrophs which are a relatively rare component of the soil microbiome [@delmont2015]. 
+If these challenges could be overcome for soil metagenomics, we could begin to evaluate whether ecosystem selection results in strain-level evolution or enriches for certain metabolic pathways within the microbiome.
 
-Another important question is whether selection at the ecosystem scale resulted in within-strain evolution.
-Using 16S rRNA gene sequencing we are unable to resolve genomic changes within taxa that may have undergone evolution.
-Future work could analyze an artificial ecosystem selection experiment using genome-resolved metagenomics to try to detect strain-level changes in the constituent taxa.
-This would help evaluate whether certain metabolic pathways were enriched in the positive selection treatment or if functional genes changed in frequency within the microbiome.
+As more is understood about the role of the microbiome in regulating ecosystem
+functions and the specific taxa involved, more detailed experiments could
+selectively add or remove important taxa or manipulate their relative abundance
+to start to estimate the quantitative relationship between specific taxa and
+the rate of an ecosystem function. Both our differential abundance analysis and
+this reductionist approach to manipulating individual taxa assumes that taxa
+act individually on ecosystem function. However, it is much more likely that
+the microbiome affects the rate of ecosystem function through the interactions
+of multiple taxa. We are unable to evaluate this possibility given our small 
+sample size, but future analyses with greater replication should test the
+role of multiple taxa simultaneously by, for example, using machine learning
+methods [@topcuoglu2020].
 
-As more is understood about the role of the microbiome in regulating ecosystem functions and the specific taxa involved, more detailed experiments could selectively add or remove important taxa or manipulate their relative abundance to start to estimate the quantitative relationship between specific taxa and the rate of an ecosystem function.
-Both our differential abundance analysis and this reductionist approach to manipulating individual taxa assumes that taxa act individually on ecosystem function.
-However, it is much more likely that the microbiome affects the rate of ecosystem function through the interactions of multiple taxa and future analyses with greater replication should test the role of multiple taxa simultaneously by, for example, using machine learning methods [@topçuoğlu2020]
-
-A fundamental question in microbial ecology is whether variation in the microbiome contributes to variation in ecosystem function.
-We performed artificial ecosystem selection on soil methane oxidation rate to test the relationship between microbiome variation and soil methane oxidation rate.
-We observed a response to selection on soil methane oxidation rate and found a high heritability for methane oxidation, which suggests that variation in the microbiome contributes to variation in soil methane oxidation rate.
-Surprisingly, we did not observe an increase in the relative of abundance of methanotrophs as a response to selection.
-Future research could estimate heritability under greater environmental variation to understand how much of the variation in methane oxidation rate is due to variation in the microbiome in a natural ecosystem.
-In addition, future selection experiments could use genome-resolved metagenomics to observe within-strain evolution and test changes in the functional pathways of the microbiome as a response to selection to determine what aspects of the microbiome explain this increase in oxidation rate.
-This study demonstrates that soil methane oxidation rate can vary greatly with variation in the microbiome and opens up future research opportunities to establish a more detailed understanding of the relationship between the microbiome and ecosystem function.
-
-# Old Discussion
-
-We observed a response to
-selection on whole-ecosystem soil CH~4~ oxidation rate. This type of response
-has been observed for other functions, such as chloroaniline degradation in
-water [@swenson2000], but this is the first instance, to our knowledge, of
-selection being performed directly on soil biogeochemical cycling. With this
-experimental design, we attempted to create multiple ecosystems that were very
-similar with respect to the type and quantity of soil, soil moisture content,
-and the headspace concentration of CH~4~. we further controlled for environmental
-variation across replicates by having twelve replicates per treatment, using
-controlled laboratory conditions, and by having a random selection treatment as
-a control. Therefore, the observed response to selection on
-CH~4~ oxidation rate is likely due to changes in the microbiome.
-
-To investigate the relationship between microbiome structure and
-ecosystem function, we wanted to determine how much of the variation in soil CH4
-oxidation was attributable to variation in the microbiome. This ecological
-question is analogous to the problem of estimating heritability in quantitative
-genetics [@morris2020]. Variation in an organismal trait is
-determined by the sum of genetic variation, environmental variation, and the
-interaction between the two [@falconer1996]. One goal of quantitative
-genetics is to determine the proportion of phenotypic variation attributable to
-genetic variation. This is commonly estimated as the narrow-sense heritability
-defined as the additive genetic variance. This goal can be achieved through
-artificial selection experiments with selection on the character of interest
-[@falconer1996].
-
-The heritability we observed in the Positive treatment was large ($h^2$ =
-`r myround(tidy(herit_model)[4, 2], 2)`) relative to other studies of biological
-heritability [@visscher2008]. However, given the small sample
-size and wide confidence intervals (95\% CI: 
-`r myround(confint(herit_model)[4, ], 2)`) it is likely that a larger study could
-more precisely estimate the true heritability. We intentionally eliminated much of the
-environmental variation that would be present in natural ecosystems, but it is
-unclear how much jar-to-jar variation there is in the composition of the soil
-substrate. Given these data, though, we conclude that variation within
-the soil microbiome has a very strong effect on soil CH~4~
-oxidation rates independent of any environmental variation.
-
-The negative heritability observed in the Neutral treatment could indicate a
-negative genetic correlation between CH~4~ oxidation rate and traits
-associated with persistence in a jar [@conner2003]. The conditions in the
-jar favor organisms that can grow in potting mix, reproduce within a three week
-incubation, and survive passaging between jars. These traits might be
-negatively correlated with a community's ability to oxidize CH~4~ indicating
-a potential trade-off [@conner2003]. This trade-off would result in a
-negative heritability of CH~4~ oxidation rate in the Neutral treatment,
-which did not experience selection on CH~4~ oxidation rate.
-
-Since there was a response to selection, we wanted to identify
-which taxa responded to selection. That is, the taxa that
-were enriched in the Positive treatment relative to the Neutral treatment. 
-Studies of microbiome structure-function
-relationships typically quantify marker genes for the final enzyme in a pathway
-to try to predict the rate of flux through that pathway. However, of the 
-`r ntaxa(sig_taxa_treat)` taxa identified by the differential abundance test, none
-were members of a family known to oxidize CH~4~. This suggests
-that, in this system, the taxa that perform the final step in the CH~4~ oxidation pathway
-do not
-limit the rate of flux through the pathway. Traits that we assume are important
-for regulating function (such as the diversity or relative abundance of
-methanotrophs estimated from phylogenetic marker genes) may not be important for
-determining variation in function.
-Perhaps once a sufficient number of methanotrophs are present, the
-process itself is limited by other metabolic processes in the ecosystem, such
-as nitrogen cycling mediated by non-methanotrophic organisms.
-
-We found that `r nrow(dt_data[dt_data$x > 0,])` taxa were enriched in the Positive treatment,
-which might indicate that the soil CH~4~ oxidation rate in these jars is controlled by
-multiple taxa. 
-This conclusion is analogous to the conclusion in quantitative genetics that
-most traits at the organismal level are the result of the interaction between
-many independent genes. Indeed, very few traits or genetic diseases are the
-result of a single gene or mutation [@kerem1989; @macdonald1992]. This results in the use of polygenic risk scores for predicting
-phenotype or disease risk in humans [@reich2001; @hindorff2009]. An analogous conclusion can be drawn for predicting
-the rate of ecosystem function from the presence or relative abundance of 
-microbial taxa. The rate of an ecosystem function at the whole-ecosystem level is 
-the result of interactions among a variety of disparate taxa with different traits.
-To better understand microbiome structure-function relationships, our results suggest that we
-should move away from single marker genes and instead investigate the role of
-multiple functional groups in determining the function of ecosystems whether it
-is CH~4~ flux from soil or host-microbiome health.
-
-Taxonomic richness dropped precipitously
-between Passage 2 and Passage 5. There are several explanations for this drop
-in diversity. The initial community was a diverse microbiome sampled from a natural soil.
-With the initial inoculation and at
-each subsequent passaging, the ecosystem was subsampled to 10\% (5 g of living
-soil was combined with 45 g of sterilized substrate). This subsampling likely
-explains a large part of the drop in diversity. In addition, beyond the imposed
-selection regime based on CH~4~ oxidation rate, these soil microbiomes were under
-a variety of selection pressures such as persistence in a jar, survival during the transfer
-from one jar to the next, and the ability to colonize a new jar in a couple of
-weeks, to name a few. These selection pressures would have further filtered
-the microbiome for taxa that could persist in this laboratory environment.
-
-In addition to the drop in richness, there was also a decrease in beta
-diversity among the jars as a result of passaging. This could be due to the
-homogenization step between each passage. Once the jars were selected to generate
-the next set of jars, the soil from those jars was homogenized and this homogenate
-was used to inoculate the next set of jars. The goal of this step was to "shuffle"
-membership in the microbiomes among the jars in order to test different
-combinations and relative abundances of taxa in terms of their effect on CH~4~
-oxidation rate. This likely also made the communities more similar in their
-membership by increasing the likelihood that each taxon was represented in each
-jar. Another explanation for this biotic homogenization is the aforementioned
-selection for persistence in a jar. These selection pressures would have 
-favored the subset of taxa suited to these laboratory conditions regardless
-of whether they were in the Positive or Neutral selection line. This can be
-seen by the relative similarity of the Positive and Neutral jars in Passage 5.
-Despite the Positive treatment undergoing selection on CH~4~ oxidation rate,
-both the Positive and Neutral jars were experiencing selection on a variety
-of other traits that made the overall soil microbiomes appear similar.
-
-One caveat that is important to keep in mind is that 16S rRNA genes are not
-markers of functional characteristics of microorganisms. They are simply
-phylogenetic markers useful for determining the relative position of taxa in a
-phylogenetic tree. In addition, assigning functions to taxa based solely on
-their taxonomy is a fraught exercise because many microbial functions are not
-very deeply conserved, are spread across disparate lineages of the tree of life,
-or are easily transferred between distantly related taxa. However,
-methanotrophy and methanogenesis are two of the most deeply conserved microbial
-traits [@martiny2013]. This results in part from the fact that
-these functions require multiple genes to perform and are therefore not easy to
-evolve independently or to transfer between lineages through horizontal gene
-transfer. However, in order to be confident that the response we observed was not
-simply due to a relative increase in methanotrophs, we would need to apply
-a functional approach to characterizing the microbiome, for example, by sequencing
-whole metagenomes.
-
-Future research in this area should use methods that are able to detect
-functions and metabolic pathways within the community that respond to
-selection. For example, genome-resolved metagenomics could identify pathways
-that are enriched in the Positive selection treatment. This would further
-advance our understanding of the exact traits that are enriched in high-flux
-ecosystems and therefore could be markers of ecosystem function. In addition,
-future studies could build on experimental results such as these to sample a
-natural gradient of ecosystem CH~4~ fluxes to determine whether the taxa identified
-in our selection experiment could be useful indicators of CH~4~ flux in the
-field. These studies would allow us to build a picture of the multiple traits
-that drive variation in ecosystem function at the whole-ecosystem level.
-
-We performed an artificial ecosystem selection experiment on whole-ecosystem
-CH~4~ oxidation rate to determine whether variation in microbiome
-composition can contribute to variation in ecosystem function. We observed
-a significant response to selection on CH~4~ oxidation rate - the first
-example of an experiment performing selection on soil biogeochemical cycling.
-In addition, we observed strong heritability of CH~4~ flux between
-passages, suggesting that variation in microbiome composition could be
-a major source of variation in CH~4~ flux in ecosystems. Surprisingly, 
-the taxonomic groups that experienced selection on CH~4~
-oxidation rate were not enriched in methanotrophs. This suggests that
-understanding a single functional group is insufficient for predicting the
-effect of microbiome composition on the rate of ecosystem function.
-Instead, we may need to investigate alternative functional groups or multiple
-interacting taxa to understand the role of microbiomes in ecosystem function. While it has often been assumed that
-microorganisms play a major role in determining variation in ecosystem function, 
-this has not previously been shown empirically. This experiment demonstrates that
-variation in microbiome composition can contribute to considerable variation
-in ecosystem function.
+A fundamental question in microbial ecology is whether variation in the
+microbiome contributes to variation in ecosystem function. We performed
+artificial ecosystem selection on soil methane oxidation rate to test the
+relationship between microbiome variation and soil methane oxidation rate. We
+observed a response to selection on soil methane oxidation rate and found a
+high heritability for methane oxidation, which suggests that variation in the
+microbiome contributes to variation in soil methane oxidation rate.
+Surprisingly, we did not observe an increase in the relative of abundance of
+methanotrophs as a response to selection. This suggests that the rate of methane 
+oxidation could be altered by
+non-methanotrophs. Future research could estimate
+heritability under greater environmental variation to understand how much of
+the variation in methane oxidation rate is due to variation in the microbiome
+in a natural ecosystem. In addition, future selection experiments could use
+genome-resolved metagenomics to observe within-strain evolution and test
+changes in the functional pathways of the microbiome as a response to selection
+to determine what aspects of the microbiome explain this increase in oxidation
+rate. This study demonstrates that soil methane oxidation rate can vary greatly
+with variation in the microbiome and opens up future research opportunities to
+establish a more detailed understanding of the relationship between the
+microbiome and ecosystem function.
 
@@
 
-This project was supported by the National Science Foundation Graduate Research Fellowship Program (grant no. DGE 1255832) and the ARCS Foundation Florence and Mike Nudelman Scholarship.
+This project was supported by the National Science Foundation Graduate Research
+Fellowship Program (grant no. DGE 1255832) and the ARCS Foundation Florence and
+Mike Nudelman Scholarship.
 
@@
 
-The 16S rRNA sequencing data generated during the current study are available in the NCBI Sequence Read Archive (SRA) under BioProject accession number PRJNA832314, https://www.ncbi.nlm.nih.gov/sra/PRJNA832314.
-The metadata generated during the current study as well as the scripts to recreate the analysis are available on Github, https://github.com/amorris28/artificial_ecosystem_selection.
+The 16S rRNA sequencing data generated during the current study are available
+in the NCBI Sequence Read Archive (SRA) under BioProject accession number
+PRJNA832314, https://www.ncbi.nlm.nih.gov/sra/PRJNA832314. The metadata
+generated during the current study as well as the scripts to recreate the
+analysis are available on Github,
+https://github.com/amorris28/artificial_ecosystem_selection.
 
@@
 
-p_herit
+p_herit_each
 
@@
 
-`r figures("herit", "Ordinary least squares regression of mid-offspring CH~4~ flux on mid-parent CH~4~ flux ($-k$ with units day^-1^).
-Mid-parent is the mean of the jars selected to inoculate the next passage.
-Mid-offspring is the mean of all twelve jars produced in one passage.
+`r figures("herit", "Ordinary least squares regression of offspring CH~4~ flux on mid-parent CH~4~ flux ($-k$ with units day^-1^).
+Mean parental is the mean of the jars selected to inoculate the next passage.
+Offspring is the oxidation rate of each jar in the next passage.
 The orange lines and points are the Positive treatment and the gray lines and points are the Neutral treatment")`

### 2022-11-09 Manuscript/manuscript.Rmd
save manuscript progress

@@
 ecosystems. We performed artificial ecosystem selection on replicate soil
-microcosms to select for increased soil methane oxidation rate. We observed a
-strong response to selection with a 51% increase in methane oxidation rate per
-round over five rounds of selection. This suggests that a large portion of the
-variation in methane oxidation rate in these soils can be attributed to
-microbiome variation. We also found that selection did not enrich for known
-methane oxidizers; rather microbes that responded to selection included members
-of `r length(da_fams)` families such as *Fimbriimonadaceae*, *Cytophagaceae*,
-and *Diplorickettsiaceae*. This result is in contrast to the typical assumption
-that the rate of an ecosystem function is limited by the final step in the
-associated microbial pathway. Our study demonstrates that variation in
-microbiome composition can contribute to variation in the rate of ecosystem
-function and that this relationship can be revealed via artificial ecosystem
-selection.
+microbiomes to select for increased soil methane oxidation rate. We observed a
+strong response to selection with a `r myround(per_slope, 1)`\% increase in
+methane oxidation rate per round over five rounds of selection. In addition, we
+estimated that `r myround((coefficients(herit_model)[2] +
+coefficients(herit_model)[4])*100, 1)`\% of the variation in methane oxidation
+rate in these soils can be attributed to microbiome variation. We also found
+that selection did not enrich for known methane oxidizers; rather microbes that
+responded to selection included members of `r length(da_fams)` families not
+known to oxidize methane, such as
+*Fimbriimonadaceae*, *Cytophagaceae*, and *Diplorickettsiaceae*. This result is
+in contrast to the typical assumption that the rate of an ecosystem function is
+limited by the final step in the associated microbial pathway. Our study
+demonstrates that variation in microbiome composition can contribute to
+variation in the rate of ecosystem function and that manipulating microbiome composition
+could be a viable strategy for managing ecosystem functions.
 
@@
 is great interest in understanding how attributes of microbiomes may influence
-variation in ecosystem functions [@conrad1996; @schimel1998; @crowther2019]. Biodiversity-ecosystem
-function relationships have been described for a variety of macroorganismal
-communities. For example, plant species richness correlates with productivity
-and marine community diversity correlates with functions such as production,
-consumption, and nutrient release
+variation in ecosystem functions [@conrad1996; @schimel1998; @crowther2019].
+Biodiversity-ecosystem function relationships have been documented for a
+variety of macroorganismal communities. For example, plant species richness
+correlates with productivity and marine community diversity correlates with
+functions such as production, consumption, and nutrient release
 [@cardinale2006; @hooper2012; @gamfeldt2015]. While there is some evidence that
 variation in microbiomes can contribute to variation in ecosystem functions,
-documenting this relationship has been elusive [@conrad1996; @rocca2015; @graham2016].
+describing a more detailed mapping between the microbiome and ecosystem
+function has been elusive [@conrad1996; @rocca2015; @graham2016].
 
@@
 methods, in which microbiome composition and diversity is surveyed via
-environmental DNA and then correlated with the rate of an ecosystem function
-{cite examples}. However, such studies are rarely successful [@graham2016]. In an
+environmental DNA and then correlated with the rate of an ecosystem function. 
+However, microbiome attributes add an average of only 8% of variance explained
+to ecosystem models [@graham2016].
+
+In an
 attempt to increase the likelihood of detecting such a correlation, microbial
@@
 assumes that one can identify the “right” marker gene a priori which, given how
-little is understood about microbial diversity, seems highly unlikely. And in
-fact these studies are rarely successful [@rocca2015].
+little is understood about microbial diversity, seems highly unlikely. And for
+most functions in most ecosystems there is no correlation between the abundance
+of a marker gene and the rate of the corresponding process [@rocca2015].
 
@@
 exploring microbiome-ecosystem function relationships is gaining acceptance
-among microbiologists [@swenson2000; @panke-buisse2015; @sanchez2021].
+among microbiologists [@swenson2000; @blouin2015; @panke-buisse2015; @sanchez2021].
 Microbiomes have been shown to respond to selection at the ecosystem-trait
@@
 for quantifying how much of the variation in the rate of an ecosystem function
-could be attributed independently to microbiome variation [@goodnight2000]. In
+could be attributed independently to microbiome variation [@goodnight2000; @blouin2015]. 
+By "microbiome variation," we mean aspects of microbial diversity that are
+routinely measured by microbial ecologists, such as the relative
+abundance of taxonomic groups inferred from amplicon sequence variants. 
+In
 addition, this approach could be used to identify microbiome attributes, such
 as genes or taxa, associated with the rate of an ecosystem function, although,
-to our knowledge this has not previously been accomplished [@arias-sanchez2019;
-@morris2020]. Such an approach has the potential to expand our ability to more
+to our knowledge this has not previously been accomplished [@arias-sanchez2019].
+Such an approach has the potential to expand our ability to more
 accurately model microbial ecosystem-function relationships and allow us to
@@
 
-We applied artificial ecosystem selection on soil microbiomes by selecting on
-the ecosystem-scale CH~4~ oxidation rate.  Using this approach, we were able to
-estimate how much of the variation in CH~4~ oxidation rate was associated with
-variation in the microbiome.  To identify which microorganisms could be markers
-of CH~4~ oxidation rate, we compared the composition of the artificially
-selected microbiomes to a control set of microbiomes with random selection.  We
-then evaluated whether these markers met our underlying assumptions about
-which taxa limit the rate of ecosystem function.
+In this study, we wanted to evaluate whether variation in the microbiome
+contributes to variation in the rate of ecosystem functions and to identify
+microbiome attributes that might explain the mapping between microbiome and
+ecosystem function. To begin to address this problem, we used artificial
+ecosystem selection on soil microbiomes as a model system.  Using this
+experimental model, we answered the following questions: Does variation in the
+relative abundance of microbial taxa contribute to variation in soil methane
+oxidation rate in the lab? How much of the variation in soil methane oxidation
+rate can be attributed to the microbiome? What attributes of the microbiome
+regulate variation in methane oxidation rate? And do these attributes match our
+assumptions about what regulates methane oxidation rate?  We applied artificial
+ecosystem selection on soil microbiomes by selecting on the ecosystem-scale
+CH~4~ oxidation rate. We estimated how much of the variation in CH~4~ oxidation
+rate was associated with variation in the microbiome by calculating the slope
+of the regression between donor and recipient communities. To identify which
+microorganisms could be markers of CH~4~ oxidation rate, we compared the
+composition of the artificially selected microbiomes to a control set of
+microbiomes with random selection. We then evaluated whether these markers met
+our underlying assumptions about which taxa limit the rate of ecosystem
+function.
 
@@
 
-We performed an artificial ecosystem selection experiment [sensu @swenson2000]
+We performed an artificial ecosystem selection experiment [*sensu* @swenson2000]
 by passaging replicate soil microbiomes. The trait we selected on was
@@
 plots indicated that `r length(contaminants)` of these were likely contaminants and these
-ASVs were removed. Amplicon sequence variants that were assigned chloroplast taxonomy were removed prior to analysis.
+ASVs were removed. Amplicon sequence variants that were assigned chloroplast or mitochondria taxonomy were removed prior to analysis.
 
@@
 
-To estimate the proportion of variance in CH~4~ oxidation rate due to variation in
-the microbiome, we estimated narrow-sense heritability ($h^2$) as the
-regression of offspring on mid-parent [@falconer1996]. The mid-parent was the
-mean for all three selected jars and the offspring was the CH~4~ oxidation rate for each jar produced by those parents. First, we tested if there was an effect of treatment
-on the heritability estimate. We compared nested models with and without treatment using an F-test
-test with the 'anova' function. We then fit the full model to estimate the
-heritability of the Positive and Neutral treatments.
+We estimated the proportion of variance in CH~4~ oxidation rate due to variation in
+the microbiome as the
+regression of recipient on donor [@falconer1996]. The mid-donor was the
+mean for all three selected jars and the mid-recipient was the mean methane oxidation 
+rate for the twelve jars produced by those three donors.
+First, we tested if there was an effect of treatment by comparing nested models with and without treatment using an F-test with the 'anova' function.
+We then fit the full model to estimate the
+slope of the relationship between recipient and donor for the Positive and Neutral treatments.
 
@@
 treatment over the five passages `r print_lm(response_model, 2)`. By contrast,
-the Positive treatment had a `r round(per_slope)`\% increase in CH~4~
+the Positive treatment had a `r myround(per_slope, 1)`\% increase in CH~4~
 oxidation rate per passage `r print_lm(response_model, 4)`.
@@
 We estimated the proportion of variation in methane oxidation due to variation
-in the microbiome as the regression of offspring on mid-parent 
-(`r figures("herit", display = "cite")`).  Offspring CH~4~ oxidation rates were
-correlated with parental CH~4~ oxidation rates in both the Positive treatment
-`r print_lm(herit_model, 4)` and the Neutral treatment `r print_lm(herit_model,
-2)`.
+in the microbiome as the regression of mid-recipient on mid-donor 
+(`r figures("herit", display = "cite")`).  Recipient CH~4~ oxidation rates were
+correlated with donor CH~4~ oxidation rates in both the Positive treatment
+`r paste0('(', 'slope', ' = ', myround(tidy(herit_model)[2, 2] + tidy(herit_model)[4, 2], 2), ', SE = ', myround(tidy(herit_model)[4, 3], 2), ', t = ', myround(tidy(herit_model)[4, 4], 2), ', p = ', myround(tidy(herit_model)[4, 5], 2), ')')` and
+the Neutral treatment `r print_lm(herit_model, 2)`.
 
@@
 understand variation in the composition of microbiomes to understand variation
-in the rate of ecosystem functions [@conrad1996; @schimel1998]. To address this question, we performed
-artificial ecosystem selection on soil microbiomes by selecting for ecosystems
-with high methane oxidation rates. We found that there was a significant
-increase in methane oxidation rate as a response to selection. Each generation
-of soil ecosystems was created from the same pool of sterile soil substrate and
-a small inoculum from the previous generation. Therefore, it is unlikely that
-this response to selection was due to changes in the abiotic conditions of the
-ecosystem. In addition, the response was only observed in the Positive
-treatment and not in the Neutral treatment, which shows that this was not a
-side effect of passaging in jars.  Therefore, we conclude that variation in the
-microbiome can generate variation in ecosystem functions independent of the
-environment.
+in the rate of ecosystem functions [@conrad1996; @schimel1998]. To address this
+question, we performed artificial ecosystem selection on soil microbiomes by
+selecting for ecosystems with high methane oxidation rates. We found that there
+was a significant increase in methane oxidation rate in the Positive selection
+treatment relative to the Neutral selection treatment. This suggests that 
+variation in the microbiome can generate variation in ecosystem
+functions independent of the environment.
 
-In quantitative genetics, heritability quantifies the degree to which variation
-in a trait is due to variation in genotype [@falconer1996]. This is often estimated as the
-regression of offspring phenotype on parental phenotype. Thus, a significant
-correlation between the trait values of parents and offspring indicates that
-the trait is at least partially genetically determined. To estimate the
-variation in an ecosystem trait due to variation in the microbiome, we can
-perform a similar estimation by comparing the trait values of the "offspring"
-jars in one generation to the trait value of the "parental" jars chosen for
-selection in the previous generation. This "ecosystem heritability" then
-quantifies how much of the variation in ecosystem function is due to variation
-in the microbiome [@goodnight2000; @sanchez2021; @henry2021].
+We wanted to estimate how much of the variation in ecosystem function can be
+attributed to variation in the microbiome. This problem is analogous to the
+problem in quantitative genetics of estimating how much of the variation in an
+organismal trait is due variation in genotype, i.e., heritability. Organismal
+biologists often estimate the percent heritability of a trait as the slope of the
+regression of offspring phenotype
+on parental phenotype [@falconer1996]. A correlation between the trait values 
+of parents and offspring indicates that the trait is at least partially 
+genetically determined. Using artificial
+ecosystem selection we can estimate the variation in an ecosystem trait due to
+variation in the microbiome by comparing the trait values of the "offspring" or
+recipient jars in one generation to the trait value of the "parental" or donor
+jars chosen for selection in the previous generation. This "ecosystem
+heritability" then quantifies what percentage of the variation in ecosystem function
+is due to variation in the microbiome [@goodnight2000; @blouin2015].
 
-We quantified how much of the variation in methane oxidation
-rate was due to variation in the microbiome by regressing the methane oxidation
-rates of the 12 offspring jars against the mean methane oxidation rate of the
-three jars selected in the previous generation 
-We observed a heritability of `r myround(coefficients(herit_model)[4], 2)` (add SE/CI) for the
-Positive selection treatment. 
-A heritability of 1 would indicate that a trait is entirely
-genetically determined and thus offspring would have the exact same value of
-that trait as their parents. A heritability greater than 1 indicates that
-offspring exceed the value of their parents. This is unlikely in the context of
-a genomic trait, but microbiomes are not limited by the constraints of genomes
-because microbes are not alleles and communities are not chromosomes.
-Therefore, complementarity within a microbial community could result in
-offspring communities that exceed the trait values of their "parent"
-ecosystems. The large heritability we observed indicates that microbiome
-variation is a large contributor to variation in methane oxidation in this
-experimental system.
+We quantified how much of the variation in methane oxidation rate was due to
+variation in the microbiome by regressing the mean methane oxidation rates of
+the 12 recipient jars against the mean methane oxidation rate of the three
+donor jars selected to inoculate the recipients. We observed a heritability of
+`r myround((coefficients(herit_model)[2] + coefficients(herit_model)[4])*100,
+1)`\% (SE = `r myround(tidy(herit_model)$std.error[4] * 100, 1)`\%) for the
+Positive selection treatment. This estimate is comparable to the heritability
+of carbon dioxide emissions from water [@blouin2015]. Notably, we observed a
+negative heritability in the Neutral treatment 
+(`r myround(coefficients(herit_model)[2]*100, 1)`\%, 
+SE = `r myround(tidy(herit_model)$std.error[2] * 100, 1)`\%). Negative heritability
+estimates are not uncommon in quantitative genetics [@steinsaltz2020]. This can
+occur for random selection controls in artificial selection experiments and may
+indicate that methane oxidation rate is negatively correlated with other traits
+associated with persistence in jars. 
+The heritability we observed in the Positive treatment indicates that
+roughly half of the variation in methane oxidation rate is due to variation in
+the microbiome in our experimental system.
 
-Next, we wanted to address what aspects of the microbiome, in terms of alpha
-diversity, beta diversity, or individual taxa, explain this contribution of the
+Given that variation in the microbiome is driving the observed variation in
+methane oxidation rate, we next wanted to address what aspects of the microbiome, in
+terms of alpha diversity, beta diversity, or the relative abundance of taxa,
+explain this contribution of the
 microbiome to variation in methane oxidation rate. There are three ways that
@@
 surprising given that methane consumption is not a common trait among microbes
-and that it is often assumed that the rate of an ecosystem function is
-limited by the final enzymatic step in a metabolic pathway
-[@rocca2015]. In certain ecosystems, methane production and consumption are 
-correlated with the abundance of methanogens and methanotrophs as estimated from marker
-genes [@meyer2020, @meyer2020a]. However, it does not appear generally true that the rate of an
-ecosystem function is limited by the abundance of the gene that encodes the
-final step in that pathway [@rocca2015]. Our results suggest that the rate of methane oxidation can be altered by
-non-methanotrophs, perhaps through ecological interactions with methanotrophic
-species.
+and that it is often assumed that the rate of an ecosystem function is limited
+by the final enzymatic step in a metabolic pathway [@rocca2015]. In certain
+ecosystems, methane production and consumption are correlated with the
+abundance of methanogens and methanotrophs as estimated from marker genes
+[@meyer2020; @meyer2020a]. However, it does not appear generally true that the
+rate of an ecosystem function is limited by the abundance of the gene that
+encodes the final step in that pathway [@rocca2015]. Our results suggest that
+ecosystem-scale methane oxidation rates can be altered by non-methanotrophs, perhaps
+through ecological interactions with methanotrophic species.
 
@@
 Therefore, it is possible that the response to selection was driven by changes
-within the genomes of methanotrophs. But given
-the slow growth rate of methanotrophs, often requiring weeks to years of
-serial batch culturing for isolation, this seems like an unlikely explanation 
-[@whittenbury1970; @hirayama2014; @kim2018]. Future
-work could analyze an artificial ecosystem selection experiment using
-genome-resolved metagenomics to try to detect strain-level changes in the
-constituent taxa. However, reconstructing genomes from in situ soil samples
-has been difficult to achieve and may require enrichment techniques, particularly for methanotrophs which are a relatively rare component of the soil microbiome [@delmont2015]. 
-If these challenges could be overcome for soil metagenomics, we could begin to evaluate whether ecosystem selection results in strain-level evolution or enriches for certain metabolic pathways within the microbiome.
-
-As more is understood about the role of the microbiome in regulating ecosystem
-functions and the specific taxa involved, more detailed experiments could
-selectively add or remove important taxa or manipulate their relative abundance
-to start to estimate the quantitative relationship between specific taxa and
-the rate of an ecosystem function. Both our differential abundance analysis and
-this reductionist approach to manipulating individual taxa assumes that taxa
-act individually on ecosystem function. However, it is much more likely that
-the microbiome affects the rate of ecosystem function through the interactions
-of multiple taxa. We are unable to evaluate this possibility given our small 
-sample size, but future analyses with greater replication should test the
-role of multiple taxa simultaneously by, for example, using machine learning
-methods [@topcuoglu2020].
+within the genomes of methanotrophs. But given the slow growth rate of
+methanotrophs, often requiring weeks to years of serial batch culturing for
+isolation, this seems like an unlikely explanation [@whittenbury1970;
+@hirayama2014; @kim2018]. Future work could analyze an artificial ecosystem
+selection experiment using genome-resolved metagenomics to try to detect
+strain-level changes in the constituent taxa. However, reconstructing genomes
+from in situ soil samples has been difficult to achieve and may require
+enrichment techniques, particularly for methanotrophs which are a relatively
+rare component of the soil microbiome [@delmont2015].  If these challenges
+could be overcome for soil metagenomics, we could begin to evaluate whether
+ecosystem selection results in strain-level evolution or enriches for certain
+metabolic pathways within the microbiome.
 
@@
 high heritability for methane oxidation, which suggests that variation in the
-microbiome contributes to variation in soil methane oxidation rate.
-Surprisingly, we did not observe an increase in the relative of abundance of
-methanotrophs as a response to selection. This suggests that the rate of methane 
-oxidation could be altered by
-non-methanotrophs. Future research could estimate
-heritability under greater environmental variation to understand how much of
-the variation in methane oxidation rate is due to variation in the microbiome
-in a natural ecosystem. In addition, future selection experiments could use
-genome-resolved metagenomics to observe within-strain evolution and test
-changes in the functional pathways of the microbiome as a response to selection
-to determine what aspects of the microbiome explain this increase in oxidation
-rate. This study demonstrates that soil methane oxidation rate can vary greatly
-with variation in the microbiome and opens up future research opportunities to
+microbiome contributes to variation in soil methane oxidation rate
+independent of the environment. This shows that manipulating the microbiome
+could be a viable strategy for mitigating or enhancing certain
+ecosystem functions.  Surprisingly, we did not observe an increase in the
+relative abundance of methanotrophs as a response to selection. This
+suggests that the rate of methane oxidation could be altered by
+non-methanotrophs, possibly through ecological interactions with methanotrophs.
+Future research should investigate other aspects of microbiome variation, such
+as strain-level evolution or functional gene composition, to begin to describe
+a more detailed mapping between the microbiome and ecosystem function. This
+study demonstrates that soil methane oxidation rate can vary greatly with
+variation in the microbiome and opens up future research opportunities to
 establish a more detailed understanding of the relationship between the
@@
 
-`r figures("response", "Response to selection on soil CH~4~ oxidation rate fit by ordinary least squares regression.
+`r figures("response", "Response to selection on soil CH~4~ oxidation rate.
 The y-axis is the log~10~ of the additive inverse of the first-order exponential decay constant $k$ (i.e., $-k$) with units day^-1^.
@@
 
-p_herit_each
+p_herit
 
@@
 
-`r figures("herit", "Ordinary least squares regression of offspring CH~4~ flux on mid-parent CH~4~ flux ($-k$ with units day^-1^).
-Mean parental is the mean of the jars selected to inoculate the next passage.
-Offspring is the oxidation rate of each jar in the next passage.
-The orange lines and points are the Positive treatment and the gray lines and points are the Neutral treatment")`
+`r figures("herit", paste0("Regression of mid-recipient methane flux on mid-donor methane flux ($-k$ with units day^-1^).
+Donor is the mean oxidation rate of the jars selected to inoculate the next passage.
+Recipient is the mean oxidation rate of the twelve jars in the next passage.
+The orange lines and points are the Positive treatment and the gray lines and points are the Neutral treatment.
+Slope of the regression line estimates the proportion of variation in methane oxidation rate due to variation in the microbiome.
+For the Positive treatment, the variance explained by the microbiome was ", myround((coefficients(herit_model)[2] + coefficients(herit_model)[4])*100, 1), "% (SE = ", myround(tidy(herit_model)$std.error[4] * 100, 1), "%) and for the Neutral treatment the variance explained by the microbiome was ", myround(coefficients(herit_model)[2]*100, 1), "% (SE = ", myround(tidy(herit_model)$std.error[2] * 100, 1), "%)."))`
 

### 2022-11-21 04-heritability.Rmd
move manuscript location

--- 2022-09-14 R/04-heritability.Rmd
@@
 
-save(p_herit, file = 'Output/herit_plot.RData')
-
-ggsave('Manuscript/fig2.jpg', p_herit, width = 4, height = 3) 
-ggsave('Manuscript/fig2.svg', p_herit, width = 4, height = 3)
+# save(p_herit, file = 'Output/herit_plot.RData')
+# 
+# ggsave('Manuscript/fig2.jpg', p_herit, width = 4, height = 3) 
+# ggsave('Manuscript/fig2.svg', p_herit, width = 4, height = 3)
 
@@
 ```{r save_heritability}
-save(herit_model, file = 'Output/herit_model.RData')
+# save(herit_model, file = 'Output/herit_model.RData')
 ```
@@
 save(p_herit_each, file = 'Output/herit_plot.RData')
+save(herit_model, file = 'Output/herit_model.RData')
 

### 2022-11-21 manuscript/manuscript.Rmd
rename and reorganize stuff

--- 2022-11-09 Manuscript/manuscript.Rmd
@@
 ^1^ Institute of Ecology and Evolution, University of Oregon, Eugene, OR, USA  
-
-\* Correspondence: Andrew H. Morris, amorris3@uoregon.edu
-
+\* Correspondence: Andrew H. Morris, amorris3\@uoregon.edu  
 Competing Interests: We declare we have no competing interests.
@@
 knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
+knitr::opts_knit$set(root.dir=normalizePath('../'))
 
@@
 tables <- captioner(prefix = "Table")
+
+source('R/functions.R')
 ```
@@
 ```{r load-data}
-heritability <- read_tsv('../Output/heritability.tsv')
-fluxes <- read_tsv('../Output/fluxes.tsv')
+heritability <- read_tsv(paste0(der_dir, 'heritability.tsv'))
+fluxes <- read_tsv(paste0(der_dir, 'fluxes.tsv'))
 
-load(file = '../Output/decontam.RData')
+load(file = paste0(der_dir, 'decontam.RData'))
 
 # CH4 models
-load(file = '../Output/response_model.RData')
-load(file = '../Output/response_plot.RData')
-load(file = '../Output/herit_model.RData')
-load(file = '../Output/herit_plot.RData')
+load(file = paste0(der_dir, 'response_model.RData'))
+load(file = paste0(der_dir, 'response_plot.RData'))
+load(file = paste0(der_dir, 'herit_model.RData'))
+load(file = paste0(der_dir, 'herit_plot.RData'))
 
 # 16S models
-load(file = '../Output/richness_models.RData')
-load(file = '../Output/phyloseq.RData')
-load(file = '../Output/beta_model.RData')
-load(file = '../Output/beta_fig.RData')
-load(file = '../Output/da_corncob.RData')
-load(file = '../Output/da_fig.RData')
-load(file = '../Output/dbrda.RData')
-
-source('../R/functions.R')
+load(file = paste0(der_dir, 'richness_models.RData'))
+load(file = paste0(der_dir, 'phyloseq.RData'))
+load(file = paste0(der_dir, 'beta_model.RData'))
+load(file = paste0(der_dir, 'beta_fig.RData'))
+load(file = paste0(der_dir, 'da_corncob.RData'))
+load(file = paste0(der_dir, 'da_fig.RData'))
+load(file = paste0(der_dir, 'dbrda.RData'))
 
@@
 strong response to selection with a `r myround(per_slope, 1)`\% increase in
-methane oxidation rate per round over five rounds of selection. In addition, we
+methane oxidation rate per passage. In addition, we
 estimated that `r myround((coefficients(herit_model)[2] +
@@
 correlates with productivity and marine community diversity correlates with
-functions such as production, consumption, and nutrient release
-[@cardinale2006; @hooper2012; @gamfeldt2015]. While there is some evidence that
-variation in microbiomes can contribute to variation in ecosystem functions,
-describing a more detailed mapping between the microbiome and ecosystem
-function has been elusive [@conrad1996; @rocca2015; @graham2016].
+production, consumption, and nutrient release 
+[@cardinale2006; @hooper2012; @gamfeldt2015]. However, we have been unsuccessful
+in documenting similar relationships for microbial communities. While there is
+some evidence that microbiomes from different ecosystems display different
+functional rates [@balser2005; @strickland2009], adding microbiome attributes to
+ecosystem models adds very little to total variance explained [@graham2016].
+Therefore, we still lack a detailed understanding of what
+functions are altered by microbiome composition and what attributes of the
+microbiome explain this relationship.
 
@@
 biodiversity of plants and animals [@hooper2012], but this is seldom possible
-with microbiomes. Most microorganisms cannot yet be cultured and most
-available methods for manipulation of microbiomes (e.g., antibiotics, dilution,
-etc.) are not sufficiently specific to manipulate individual microbial
-lineages. Microbial ecologists are thus often restricted to using comparative
-methods, in which microbiome composition and diversity is surveyed via
-environmental DNA and then correlated with the rate of an ecosystem function. 
-However, microbiome attributes add an average of only 8% of variance explained
-to ecosystem models [@graham2016].
+with microbiomes. Most microorganisms cannot yet be cultured and most available
+methods for manipulation of microbiomes (e.g., antibiotics, dilution, etc.) are
+not sufficiently specific to manipulate individual microbial lineages. In
+addition, many studies have looked for relationships between an ecosystem
+function and the abundance or diversity of a microbial gene that codes for an
+enzyme or enzyme-subunit known to be involved in that function. However, it is
+difficult to identify the “right” marker gene a priori and for most functions
+in most ecosystems there is no correlation between the abundance of a marker
+gene and the rate of the corresponding process [@rocca2015]. Lastly,
+microbiome-ecosystem function relationships have been difficult to document
+because microbiomes often covary with other drivers of variation in ecosystem
+functions, such as environmental conditions or resources. Such covariation can
+make it impossible to determine whether microbiomes directly contribute to
+functional variation. Understanding the degree to which microbiomes directly
+influence functional variation is crucial if scientists are to develop
+microbiome-focused methods for altering or maintaining ecosystem functions.
 
-In an
-attempt to increase the likelihood of detecting such a correlation, microbial
-ecologists may focus their survey efforts on a particular group of microbes
-thought to be involved in the ecosystem function of interest. For example, many
-studies have looked for relationships between an ecosystem function and the
-abundance or diversity of a microbial gene that codes for an enzyme or
-enzyme-subunit known to be involved in a particular function. But this approach
-assumes that one can identify the “right” marker gene a priori which, given how
-little is understood about microbial diversity, seems highly unlikely. And for
-most functions in most ecosystems there is no correlation between the abundance
-of a marker gene and the rate of the corresponding process [@rocca2015].
-
-Another likely reason that microbiome-ecosystem function relationships have
-been difficult to document is that microbiomes often covary with other drivers
-of variation in ecosystem functions, such as environmental conditions or
-resources. Such covariation can make it impossible to determine whether
-microbiomes directly contribute to functional variation. Understanding the
-degree to which microbiomes directly influence functional variation is crucial
-if scientists are to develop microbiome-focused methods for altering or
-maintaining ecosystem functions.
-
-The problem of connecting microbiome attributes to the rate of an ecosystem
-function is analogous to the problem of connecting genomic variation to
-phenotypic traits in organisms, and solutions to this problem may be found by
-exploring this analogy [@morris2020]. For example, one approach to the problem
-of genotype-phenotype mapping is to perform artificial selection on a trait of
-interest and identify genes that respond to selection. Such genes are likely to
-be causally linked to the trait under selection. An analogous approach for
-exploring microbiome-ecosystem function relationships is gaining acceptance
-among microbiologists [@swenson2000; @blouin2015; @panke-buisse2015; @sanchez2021].
+One way to overcome these challenges is to manipulate the relationship between
+microbiome biodiversity and ecosystem function using artificial ecosystem selection 
+[@swenson2000; @blouin2015; @panke-buisse2015; @sanchez2021]. 
 Microbiomes have been shown to respond to selection at the ecosystem-trait
 level [@swenson2000], and some ecosystem-scale microbiome traits have been
-shown to be transferable through microbiome inoculations [@panke-buisse2015].
-Applying artificial selection to whole microbiomes could be a novel approach
-for quantifying how much of the variation in the rate of an ecosystem function
-could be attributed independently to microbiome variation [@goodnight2000; @blouin2015]. 
-By "microbiome variation," we mean aspects of microbial diversity that are
-routinely measured by microbial ecologists, such as the relative
-abundance of taxonomic groups inferred from amplicon sequence variants. 
-In
-addition, this approach could be used to identify microbiome attributes, such
-as genes or taxa, associated with the rate of an ecosystem function, although,
-to our knowledge this has not previously been accomplished [@arias-sanchez2019].
-Such an approach has the potential to expand our ability to more
-accurately model microbial ecosystem-function relationships and allow us to
-manage ecosystems for particular outcomes [@fuhrman2009; @schimel1998].
+shown to be transferrable through microbiome inoculations [@panke-buisse2015].
+Therefore, applying artificial selection to whole microbiomes could be a novel
+approach for describing a more detailed microbiome-ecosystem function
+relationship. This approach works by subsampling a microbiome to inoculate
+multiple replicate ecosystems that vary in microbiome composition and then
+measuring the rate of ecosystem function for each replicate. The investigator
+then selects the replicates with the highest (or lowest) rates and subsamples
+those ecosystems to passage them to a new set of ecosystems. By holding
+environmental conditions relatively constant among the replicates and from
+passage to passage (or by using isogenic hosts, in the case of host-associated
+microbiomes), investigators can determine whether variation in the microbiome
+contributes to variation in the rate of ecosystem function independent of the
+environment.
 
-We use this approach to estimate how much of the variation in an ecosystem
-function - the flux of CH~4~ gas between the soil and atmosphere - can be
-attributed to variation in the microbiome and to identify microbial markers
-associated with this ecosystem function.  Soil CH~4~ oxidation is a suitable
-function for this study because there is evidence that it may vary with
-microbiome attributes. For example, variation in CH~4~ emissions in arctic
-permafrost is correlated with the transcriptional activity of certain
-methanogens [@freitag2009; @freitag2010]. In addition, soil microbiome
-phylogenetic variation is a strong predictor of CH~4~ oxidation in forests and
-pastures of the Brazilian Amazon and the Congo Basin [@meyer2020; @meyer2020a]. Finally, methanogenesis and
-methanotrophy are two of the most deeply conserved microbial physiologies and
-are represented in a narrow range of taxa and so the taxonomic composition of
-the microbiome is more likely to be associated with the rate of CH~4~ flux than
-other broader or more shallowly conserved functions [@schimel1998;
-@martiny2013]. This suggests that microbiome variation might drive variation in
-CH~4~ oxidation in soils.
+A valuable feature of this approach is that it can be used to estimate how much
+of the variation in ecosystem function is due to microbiome variation.  By
+calculating the slope of the regression between the "donor" ecosystems selected
+in one passage and the "recipient" ecosystems inoculated by those donors in the
+next passage, we can quantify an estimate of "ecosystem heritability"
+[@goodnight2000; @blouin2015]. The concept of heritability has been borrowed
+from quantitative genetics where heritability is defined as *the ratio of
+additive genetic variance to the total phenotypic variance* and is commonly
+estimated as *the slope of the linear regression between offspring phenotype
+and parental phenotype* [@bell1977; @falconer1996]. Here we apply this concept
+to donor and recipient communities in an artificial ecosystem selection
+experiment to estimate *the variance in ecosystem function due to microbiome
+variance*. We do not assume that environmental variation is zero, instead we
+quantify the proportion of variation due to microbiome effects. If the slope of
+the regression is zero, we would conclude that environmental variation among
+the replicates and technical error are the primary drivers of variation in
+ecosystem function. On the other hand, a significant slope between donors and
+recipients would demonstrate that microbiome variation generates variation in
+the rate of an ecosystem function. We then use our artificially selected
+microbiomes to investigate what aspects of microbiome variation explain this
+response. By "microbiome variation" we mean aspects of microbial diversity
+commonly measured by microbial ecologists, such as the relative abundance of
+taxonomic groups inferred from amplicon sequence variants. With this approach,
+we can not only determine whether an ecosystem function is influenced by
+microbiome variation, but also estimate how much of the variation and which
+taxa are associated with higher rates of ecosystem function.
 
-In this study, we wanted to evaluate whether variation in the microbiome
-contributes to variation in the rate of ecosystem functions and to identify
-microbiome attributes that might explain the mapping between microbiome and
-ecosystem function. To begin to address this problem, we used artificial
-ecosystem selection on soil microbiomes as a model system.  Using this
-experimental model, we answered the following questions: Does variation in the
-relative abundance of microbial taxa contribute to variation in soil methane
-oxidation rate in the lab? How much of the variation in soil methane oxidation
-rate can be attributed to the microbiome? What attributes of the microbiome
-regulate variation in methane oxidation rate? And do these attributes match our
-assumptions about what regulates methane oxidation rate?  We applied artificial
-ecosystem selection on soil microbiomes by selecting on the ecosystem-scale
-CH~4~ oxidation rate. We estimated how much of the variation in CH~4~ oxidation
-rate was associated with variation in the microbiome by calculating the slope
-of the regression between donor and recipient communities. To identify which
-microorganisms could be markers of CH~4~ oxidation rate, we compared the
-composition of the artificially selected microbiomes to a control set of
-microbiomes with random selection. We then evaluated whether these markers met
-our underlying assumptions about which taxa limit the rate of ecosystem
-function.
+We applied artificial ecosystem selection to soil microbial communities by
+selecting on soil methane oxidation rate. We chose this function because
+methane is a globally important greenhouse gas and methane oxidation by soil
+bacteria is the primary biological sink for atmospheric methane
+[@kirschke2013]. In addition, there is evidence that soil methane oxidation
+rate may vary with microbiome composition. For example, soil microbiome
+phylogenetic variation is a strong predictor of methane oxidation in forests
+and pastures of the Brazilian Amazon as well as soils of the Congo Basin
+[@meyer2020; @meyer2020a]. Finally, methanotrophy is one of the most deeply
+conserved microbial physiologies and is represented in a narrow range of taxa
+and so the taxonomic composition of the microbiome is more likely to be
+associated with the rate of methane oxidation than other broader or more
+shallowly conserved functions [@schimel1998; @martiny2013]. This suggests that
+microbiome variation might drive variation in methane oxidation in soils. One
+other study has estimated ecosystem heritability for an environmental
+microbiome. Blouin et al. [@blouin2015] selected for water microbiomes with
+low carbon dioxide respiration rates and estimated a heritability of 52%.
+However, this appeared to be driven primarily by biomass, not microbiome
+composition. Here we want to build on this work by selecting for higher rates
+of a more narrow function that is less likely to be driven by total biomass.
+That way we can begin to assess whether variation in microbiome composition is
+important for understanding variation in ecosystem functions.
+
+In this study, we evaluated whether variation in the microbiome contributes to
+variation in the rate of ecosystem functions. To address this problem, we
+applied artificial ecosystem selection to methane oxidation rate of the soil
+microbiome and answered the following questions: Does variation in the relative
+abundance of microbial taxa contribute to variation in soil methane oxidation
+rate in the lab? How much of the variation in soil methane oxidation rate can
+be attributed to the microbiome? What attributes of the microbiome regulate
+variation in methane oxidation rate? And do these attributes match our
+assumptions about what regulates methane oxidation rate in nature?
 
@@
 
-We performed an artificial ecosystem selection experiment [*sensu* @swenson2000]
-by passaging replicate soil microbiomes. The trait we selected on was
-CH~4~ oxidation rate. This experiment had two selection lines with twelve jars
-each for a total of twenty-four jars per passage. One line was a positive
-selection line where the two or three jars with the highest CH~4~ oxidation rate
-were chosen to inoculate the next set of positive jars. The other line was a
-neutral selection line where an equal number of jars as the positive line were
-chosen at random to inoculate the next set of neutral jars. The number of jars
-chosen was based on the distribution of fluxes among the positive jars, i.e., we
-chose the top three jars unless only two jars had considerably greater CH~4~ oxidation
-rates based on visual inspection of histograms. The experiment was carried out
-over five passages until a significant divergence in functional rates was 
-observed between the two selection treatments.
+We performed an artificial ecosystem selection experiment 
+[*sensu* @swenson2000] by passaging replicate soil microbiomes. The trait we
+selected on was CH~4~ oxidation rate. This experiment had two selection lines
+with twelve jars each for a total of twenty-four jars per passage. One line was
+a positive selection line where the two or three jars with the highest CH~4~
+oxidation rate were homogenized to inoculate the next set of positive jars. The
+other line was a neutral selection line where an equal number of jars as the
+positive line were chosen at random to inoculate the next set of neutral jars.
+The number of jars chosen was based on the distribution of fluxes among the
+positive jars, i.e., we chose the top three jars unless only two jars had
+considerably greater CH~4~ oxidation rates based on visual inspection of
+histograms. The experiment was carried out over five passages until a
+significant divergence in functional rates was observed between the two
+selection treatments based on a difference of slopes.
 
@@
 was added 45 g of autoclaved potting mix, 5 g of living soil, and 3.5 mL of
-sterile deionized water to bring them to 60\% of field capacity. The potting mix was
-Lane Potting Mix from Lane Forest Products (Eugene, OR, USA). The jars were
-then homogenized, capped, and injected with 4.3 mL of 99\% CH~4~ to bring
-the headspace concentration to approximately 1000 ppm CH~4~. To create the
-two treatment groups, twenty-four jars were created in an identical manner and
-then randomly assigned to either the positive or neutral selection line. Jars
-were flushed and respiked twice per week to maintain aerobic conditions and
-elevated CH~4~ concentrations and were incubated at ambient temperature for
+sterile deionized water to bring them to 60\% of field capacity. The potting
+mix was Lane Potting Mix from Lane Forest Products (Eugene, OR, USA). Each jar
+was then capped and injected with 4.3 mL of 99\% CH~4~ to bring
+the headspace concentration to approximately 1000 ppm CH~4~. To create the two
+treatment groups, twenty-four jars were created in an identical manner and then
+randomly assigned to either the positive or neutral selection line. Jars were
+flushed and respiked with methane twice per week to maintain aerobic conditions and elevated
+CH~4~ concentrations and were incubated at ambient temperature for
 approximately three weeks. Methane oxidation rates were determined at the end
@@
 Instruments, Torrance, CA, USA) to determine the headspace CH~4~ concentration.
-Fluxes were calculated from a first-order exponential decay function as decay
-constant $k$ with units day^-1^. Oxidation rates are presented as the additive
-inverse of $k$ (i.e., $-k$) so that a more positive value represents a greater
-oxidation rate. The jars selected for the positive treatment in passage 2 had
+We applied a first-order exponential decay function to determine the rate
+constant (k, units = d^−1^; i.e., dCH~4~/dt = k[CH~4~]) of the exponential
+decrease in methane. Oxidation rates are presented as the additive inverse of
+$k$ (i.e., $-k$) so that a more positive value represents a greater
+oxidation rate. The jars selected in passage two for the positive treatment had
 the lowest CH~4~ oxidation rate of the twelve jars due to a calculation error
-in the CH~4~ oxidation rate. Despite this error, we still observed a strong
+in the rate constant. Despite this error, we still observed a strong
 response to selection demonstrating the power of this technique. In addition,
@@
 A subsample of soil from the starting inoculum and from every jar in passages 2
-and 5 was collected and stored at $-80^\circ$C. Soil DNA was extracted from 0.25
-g soil. Negative controls were extracted from autoclaved potting mix and
+and 5 was collected and stored at $-80^\circ$C. Soil DNA was extracted from
+0.25 g soil. Negative controls were extracted from autoclaved potting mix and
 DNase-free water. Extractions were performed using the DNeasy PowerSoil kit
-(QIAGEN, Düsseldorf, Germany) and quantified using Qubit dsDNA HS Assay Kit (Thermo Fisher
-Scientific, Inc., Waltham, MA, USA). To estimate the diversity and relative abundance of
-the bacterial and archaeal taxa in our soil ecosystems, we sequenced the V4
-region of the 16S rRNA gene using the 515F - 806R primer combination
-[@caporaso2011]. PCR mixtures were: 10 $\mu$l NEBNext Q5 Hot Start HiFi PCR
-master mix, 9.2 $\mu$l primer mixture (1.09 $\mu$M concentration), and 0.8
-$\mu$l of DNA template. Reaction conditions were: $98^\circ$C for 30 s
-(initialization); 35 cycles of $98^\circ$C for 10 s (denaturation), $61^\circ$C
-for 20 s (annealing), and $72^\circ$C for 20 s (extension); and $72^\circ$C for
-2 m (final extension). Reactions were performed in triplicate and then combined.
-Amplicons were purified twice using 0.8x ratio Mag-Bind
-RxnPure Plus isolation beads (Omega Bio-Tek, Norcross, GA, USA). Sequencing libraries were
-prepared using a dual-indexing approach [@fadrosh2014; @kozich2013]. Amplicon
-concentrations were quantified using Qubit and multiplexed at equimolar concentration. Sequencing
-was performed at the University of Oregon Genomics Core Facility on the Illumina
-NovaSeq 6000 with paired-end 150 bp reads (Illumina, Inc., San Diego, CA, USA).
+(QIAGEN, Düsseldorf, Germany) and quantified using Qubit dsDNA HS Assay Kit
+(Thermo Fisher Scientific, Inc., Waltham, MA, USA). To estimate the diversity
+and relative abundance of the bacterial and archaeal taxa in our soil
+ecosystems, we sequenced the V4 region of the 16S rRNA gene using the 515F -
+806R primer combination [@caporaso2011]. PCR mixtures were: 10 $\mu$l NEBNext
+Q5 Hot Start HiFi PCR master mix, 9.2 $\mu$l primer mixture (1.09 $\mu$M
+concentration), and 0.8 $\mu$l of DNA template. Reaction conditions were:
+$98^\circ$C for 30 s (initialization); 35 cycles of $98^\circ$C for 10 s
+(denaturation), $61^\circ$C for 20 s (annealing), and $72^\circ$C for 20 s
+(extension); and $72^\circ$C for 2 m (final extension). Reactions were
+performed in triplicate and then combined.  Amplicons were purified twice using
+0.8x ratio Mag-Bind RxnPure Plus isolation beads (Omega Bio-Tek, Norcross, GA,
+USA). Sequencing libraries were prepared using a dual-indexing approach
+[@fadrosh2014; @kozich2013]. Amplicon concentrations were quantified using
+Qubit and multiplexed at equimolar concentration. Sequencing was performed at
+the University of Oregon Genomics Core Facility on the Illumina NovaSeq 6000
+with paired-end 150 bp reads (Illumina, Inc., San Diego, CA, USA).
 
@@
 
-Bioinformatics processing was performed in 'R' [@rcoreteam2018].
-Demultiplexed sequencing reads were denoised using 'DADA2' to generate a table
-of amplicon sequence variants (ASVs)
-[@callahan2016]. Taxonomic assignment was performed
-using the Ribosomal Database Project naive Bayesian classifier [@wang2007]. The presence of
-contaminants was evaluated using both the prevalence and frequency methods from
-'DECONTAM' [@davis2018].
-Decontam identified `r length(which(contamdf$contaminant))` potential contaminants based
-on prevalence and frequency. Visual inspection of abundance-concentration
-plots indicated that `r length(contaminants)` of these were likely contaminants and these
-ASVs were removed. Amplicon sequence variants that were assigned chloroplast or mitochondria taxonomy were removed prior to analysis.
+Bioinformatics processing was performed in 'R' [@rcoreteam2018].  Demultiplexed
+sequencing reads were denoised using 'DADA2' to generate a table of amplicon
+sequence variants (ASVs) [@callahan2016]. Taxonomic assignment was performed
+using the Ribosomal Database Project naive Bayesian classifier [@wang2007]. The
+presence of contaminants was evaluated using both the prevalence and frequency
+methods from 'DECONTAM' [@davis2018]. Decontam identified 
+`r length(which(contamdf$contaminant))` potential contaminants based on
+prevalence and frequency. Visual inspection of abundance-concentration plots
+indicated that `r length(contaminants)` of these were likely contaminants and
+these ASVs were removed. Amplicon sequence variants that were assigned
+chloroplast or mitochondria taxonomy were removed prior to analysis.
 
@@
 Statistical analyses were performed in 'R' [@rcoreteam2018]. To determine
-whether there was a significant change in CH~4~ oxidation rate as a
-response to selection, we tested a difference in slopes between the positive
-and neutral selection lines. Residuals did not meet the assumptions of constant variance
+whether there was a significant change in CH~4~ oxidation rate as a response to
+selection, we tested a difference in slopes between the positive and neutral
+selection lines. Residuals did not meet the assumptions of constant variance
 and normal distribution. Therefore, CH~4~ oxidation rates were log~10~
-transformed prior to analysis. First, we tested the difference of slopes between
-the positive and neutral treatments by
-fitting two nested models with and without the interaction between treatment
-and passage using 'lm'. We compared these
-models using an F-test with the 'anova' function. We
-then fit the full model with the interaction term to estimate the slopes for each treatment, which
+transformed prior to analysis. First, we tested if there was a difference of
+slopes between the positive and neutral treatments based on the interaction
+between passage and treatment. To test the interaction, we fit two nested models
+with and without the interaction term and compared them using an F-test with the
+'anova' function. We then present the slopes for each treatment, which
 represented the change in CH~4~ oxidation rate per passage as a response to
@@
 
-We estimated the proportion of variance in CH~4~ oxidation rate due to variation in
-the microbiome as the
-regression of recipient on donor [@falconer1996]. The mid-donor was the
-mean for all three selected jars and the mid-recipient was the mean methane oxidation 
-rate for the twelve jars produced by those three donors.
-First, we tested if there was an effect of treatment by comparing nested models with and without treatment using an F-test with the 'anova' function.
-We then fit the full model to estimate the
-slope of the relationship between recipient and donor for the Positive and Neutral treatments.
+We estimated the ecosystem heritability of methane oxidation rate, which is the
+proportion of variance in CH~4~ oxidation rate due to variation in the
+microbiome, as the regression of mid-recipient on mid-donor [@falconer1996].
+The mid-donor was the mean for all three selected jars and the mid-recipient was
+the mean methane oxidation rate for the twelve jars produced by those three
+donors. First, we tested if there was a difference of slopes between the
+positive and neutral treatments. To test this, we fit two nested models with and
+without the treatment term and compared them using an F-test with the 'anova'
+function. We then present the slopes for each treatment, which represented the
+ecosystem heritability or the proprotion of total variance in methane oxidation
+rate due to microbiome variance.
+
+First, we tested if there was an effect of treatment by comparing
+nested models with and without treatment using an F-test with the 'anova'
+function. We then fit the full model to estimate the slope of the relationship
+between recipient and donor for the Positive and Neutral treatments.
 
@@
 differential abundance between the two treatments in passage 5. We first
-grouped ASVs at the family level keeping ASVs that lacked a family-level taxonomic
-assignment by grouping them at a higher level. We
-then subset the samples in Passage 5 and removed all families with a prevalence
-of less than 10\% in either treatment. We used three methods for testing differential abundance: ANCOM-II, ALDEx2, and CORNCOB. We then
-identified the consensus taxa that were significant
-with all three tests and plotted their relative abundances.
-For ANCOM-II, we used the 'ancom' function in the 'ANCOM-BC' package with a cutoff of W = 0.7 [@kaul2017; @lin2020].
-For ALDEx2, we used the 'aldex' function in the 'ALDEx2' package with Welch's t-test and we used an effect size of
-1 as our significance threshold [@fernandes2014].
-Finally, we used CORNCOB with the 'differentialTest' function in the 'corncob' package with the Wald test and without bootstrapping [@martin2020].
-Lastly, to test differentially abundant methanotrophs, we subset all ASVs within methanotrophic families and tested their differential abundance aggregated at the family and genus level using 'corncob'.
+grouped ASVs at the family level. Any ASVs that lacked a family-level
+taxonomic assignment were grouped at a higher taxonomic level. We then subset the
+samples in Passage 5 and removed all families with a prevalence of less than
+10\% in either treatment. We used three methods for testing differential
+abundance: ANCOM-II, ALDEx2, and CORNCOB 
+[@kaul2017; @lin2020; @fernandes2014; @martin2020]. We then identified the
+consensus taxa that were significant with all three tests and plotted their
+relative abundances. For ANCOM-II, we used the 'ancom' function in the
+'ANCOM-BC' package with a cutoff of W = 0.7 [@kaul2017; @lin2020]. For ALDEx2,
+we used the 'aldex' function in the 'ALDEx2' package with Welch's t-test and we
+used an effect size of 1 as our significance threshold [@fernandes2014].
+Finally, we used CORNCOB with the 'differentialTest' function in the 'corncob'
+package with the Wald test and without bootstrapping [@martin2020]. Lastly, to
+test differentially abundant methanotrophs, we subset all ASVs within
+methanotrophic families and tested their differential abundance aggregated at
+the family and genus level using 'corncob'.
 
@@
 
-We observed a response to selection on whole-ecosystem soil CH~4~ oxidation
+We observed a response to artificial selection on whole-ecosystem soil CH~4~ oxidation
 rate (`r figures("response", display = "cite")`; difference of slopes:
@@
 
-We estimated the proportion of variation in methane oxidation due to variation
+We estimated the ecosystem heritability defined as the 
+proportion of variation in methane oxidation due rate to variation
 in the microbiome as the regression of mid-recipient on mid-donor 
@@
 was a significant increase in methane oxidation rate in the Positive selection
-treatment relative to the Neutral selection treatment. This suggests that 
-variation in the microbiome can generate variation in ecosystem
-functions independent of the environment.
-
-We wanted to estimate how much of the variation in ecosystem function can be
-attributed to variation in the microbiome. This problem is analogous to the
-problem in quantitative genetics of estimating how much of the variation in an
-organismal trait is due variation in genotype, i.e., heritability. Organismal
-biologists often estimate the percent heritability of a trait as the slope of the
-regression of offspring phenotype
-on parental phenotype [@falconer1996]. A correlation between the trait values 
-of parents and offspring indicates that the trait is at least partially 
-genetically determined. Using artificial
-ecosystem selection we can estimate the variation in an ecosystem trait due to
-variation in the microbiome by comparing the trait values of the "offspring" or
-recipient jars in one generation to the trait value of the "parental" or donor
-jars chosen for selection in the previous generation. This "ecosystem
-heritability" then quantifies what percentage of the variation in ecosystem function
-is due to variation in the microbiome [@goodnight2000; @blouin2015].
+treatment relative to the Neutral selection treatment. This demonstrates that
+passaging the microbiome through artificial ecosystem selection resulted in a
+shift in the mean methane oxidation rate of the soil microbiome. 
+This suggests that variation in the microbiome can generate variation in
+ecosystem functions independent of the environment.
 
 We quantified how much of the variation in methane oxidation rate was due to
-variation in the microbiome by regressing the mean methane oxidation rates of
+variation in the microbiome by regressing the mean methane oxidation rate of
 the 12 recipient jars against the mean methane oxidation rate of the three
 donor jars selected to inoculate the recipients. We observed a heritability of
-`r myround((coefficients(herit_model)[2] + coefficients(herit_model)[4])*100,
-1)`\% (SE = `r myround(tidy(herit_model)$std.error[4] * 100, 1)`\%) for the
-Positive selection treatment. This estimate is comparable to the heritability
-of carbon dioxide emissions from water [@blouin2015]. Notably, we observed a
-negative heritability in the Neutral treatment 
-(`r myround(coefficients(herit_model)[2]*100, 1)`\%, 
-SE = `r myround(tidy(herit_model)$std.error[2] * 100, 1)`\%). Negative heritability
-estimates are not uncommon in quantitative genetics [@steinsaltz2020]. This can
-occur for random selection controls in artificial selection experiments and may
-indicate that methane oxidation rate is negatively correlated with other traits
-associated with persistence in jars. 
-The heritability we observed in the Positive treatment indicates that
-roughly half of the variation in methane oxidation rate is due to variation in
-the microbiome in our experimental system.
+`r myround((coefficients(herit_model)[2] + coefficients(herit_model)[4])*100,1)`\% 
+for the Positive selection treatment. This indicates that 
+`r myround((coefficients(herit_model)[2] + coefficients(herit_model)[4])*100,1)`\% 
+of the variation in methane oxidation rate can be attributed to variation in the
+microbiome. This is comparable to the ecosystem heritability of carbon dioxide
+emissions from water [@blouin2015]. Notably, we observed a negative heritability
+in the Neutral treatment (`r myround(coefficients(herit_model)[2]*100, 1)`\%).
+Negative heritability estimates are not uncommon in quantitative genetics
+[@steinsaltz2020]. This can occur for random selection controls in artificial
+selection experiments and may indicate that methane oxidation rate is negatively
+correlated with other traits associated with persistence in jars. The
+heritability we observed in the Positive treatment indicates that a large
+portion of the variation in methane oxidation rate is due to variation in the
+microbiome in our experimental system.
 
 Given that variation in the microbiome is driving the observed variation in
-methane oxidation rate, we next wanted to address what aspects of the microbiome, in
-terms of alpha diversity, beta diversity, or the relative abundance of taxa,
-explain this contribution of the
-microbiome to variation in methane oxidation rate. There are three ways that
-the microbiome could respond to selection in this experiment: gain or loss of
-species, changes in the relative abundances of species, or changes within the
-genomes of the constituent species. Using 16S rRNA gene sequences, we are
-unable to address the third possibility of whether taxa in this experiment
-could have evolved changes in their genomes as a result of selection.
-Therefore, we will focus on the first two possibilities.
+methane oxidation rate, we next wanted to address what aspects of the
+microbiome, in terms of alpha diversity, beta diversity, or the relative
+abundance of taxa, explain this contribution of the microbiome to variation in
+methane oxidation rate. There are three ways that the microbiome could respond
+to selection in this experiment: gain or loss of species, changes in the
+relative abundances of species, or changes within the genomes of the constituent
+species. Using 16S rRNA gene sequences, we are unable to address the third
+possibility of whether taxa in this experiment could have evolved changes in
+their genomes as a result of selection.  Therefore, we will focus on the first
+two possibilities.
 
@@
 encodes the final step in that pathway [@rocca2015]. Our results suggest that
-ecosystem-scale methane oxidation rates can be altered by non-methanotrophs, perhaps
-through ecological interactions with methanotrophic species.
+ecosystem-scale methane oxidation rates can be altered by non-methanotrophs,
+perhaps through ecological interactions with methanotrophic species.
 
@@
 methanotrophs, often requiring weeks to years of serial batch culturing for
-isolation, this seems like an unlikely explanation [@whittenbury1970;
-@hirayama2014; @kim2018]. Future work could analyze an artificial ecosystem
-selection experiment using genome-resolved metagenomics to try to detect
-strain-level changes in the constituent taxa. However, reconstructing genomes
-from in situ soil samples has been difficult to achieve and may require
-enrichment techniques, particularly for methanotrophs which are a relatively
-rare component of the soil microbiome [@delmont2015].  If these challenges
-could be overcome for soil metagenomics, we could begin to evaluate whether
-ecosystem selection results in strain-level evolution or enriches for certain
-metabolic pathways within the microbiome.
+isolation, this seems like an unlikely explanation 
+[@whittenbury1970; @hirayama2014; @kim2018]. Future work could analyze an
+artificial ecosystem selection experiment using genome-resolved metagenomics to
+try to detect strain-level changes in the constituent taxa. However,
+reconstructing genomes from in situ soil samples has been difficult to achieve
+and may require enrichment techniques, particularly for methanotrophs which are
+a relatively rare component of the soil microbiome [@delmont2015]. If these
+challenges could be overcome for soil metagenomics, we could begin to evaluate
+whether ecosystem selection results in strain-level evolution or enriches for
+certain metabolic pathways within the microbiome.
 
@@
 
-`r figures("herit", paste0("Regression of mid-recipient methane flux on mid-donor methane flux ($-k$ with units day^-1^).
-Donor is the mean oxidation rate of the jars selected to inoculate the next passage.
-Recipient is the mean oxidation rate of the twelve jars in the next passage.
+`r figures("herit", paste0("Ecosystem heritability as the regression of mid-recipient methane flux on mid-donor methane flux ($-k$ with units day^-1^).
+Donor is the mean oxidation rate of the three jars selected to inoculate the next passage.
+Recipient is the mean oxidation rate of the 12 jars in the next passage.
 The orange lines and points are the Positive treatment and the gray lines and points are the Neutral treatment.
 Slope of the regression line estimates the proportion of variation in methane oxidation rate due to variation in the microbiome.
-For the Positive treatment, the variance explained by the microbiome was ", myround((coefficients(herit_model)[2] + coefficients(herit_model)[4])*100, 1), "% (SE = ", myround(tidy(herit_model)$std.error[4] * 100, 1), "%) and for the Neutral treatment the variance explained by the microbiome was ", myround(coefficients(herit_model)[2]*100, 1), "% (SE = ", myround(tidy(herit_model)$std.error[2] * 100, 1), "%)."))`
+For the Positive treatment, the variance explained by the microbiome was ", myround((coefficients(herit_model)[2] + coefficients(herit_model)[4])*100, 1), "% (slope = ", myround(coefficients(herit_model)[2] + coefficients(herit_model)[4], 3), ", SE = ", myround(tidy(herit_model)$std.error[4], 2), ") and for the Neutral treatment the variance explained by the microbiome was ", myround(coefficients(herit_model)[2]*100, 1), "% (slope = ", myround(coefficients(herit_model)[2], 3), ", SE = ", myround(tidy(herit_model)$std.error[2], 2), ")."))`
 

### 2022-12-06 analysis/analysis.Rmd
reorganize working directory and add new analysis and text

--- 2022-09-16 R/08-community-analysis.Rmd
@@
 ---
-title: "Community Analysis"
+title: "Analysis"
 author: "Andrew Morris"
-date: "6/28/2022"
+date: "2022-11-21"
 output: 
@@
 ```{r setup, include=FALSE}
-
-knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = TRUE)
-knitr::opts_knit$set(root.dir=normalizePath('../'))
-set.seed(1337)
-
-```
-
-```{r initialize_packages}
+here::i_am("analysis/analysis.Rmd")
+
+knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = TRUE)
+
+```
+
+```{r initialize_packages, message=FALSE}
+library(morris)
+library(broom)
+library(knitr)
+library(broman)
 
@@
 library(ggpubr)
-library(broom)
 library(ggtext)
@@
 library(tidyverse)
-
-source('R/functions.R')
-
-```
-
-```{r import-data}
-
-aes_data <- read_tsv('Output/community_data.tsv')
-
-```
+library(here)
+
+source(here('R/', 'functions.R'))
+
+set.seed(1337)
+```
+
+```{r import_data, message=FALSE}
+fluxes <- read_tsv(here(der_dir, 'fluxes.tsv'))
+heritability <- read_tsv(here(der_dir, 'heritability.tsv'))
+aes_data <- read_tsv(here(der_dir, 'community_data.tsv'))
+
+# Shift the passage numbers so the first passage becomes "Passage 0"
+fluxes <-
+  fluxes %>% 
+  mutate(passage = passage - 1)
+```
+
+# Overview
+
+The final dataset includes `r length(unique(fluxes$treat))` treatments across `r length(unique(fluxes$passage))` passages (generations).
+This includes a total of n = `r nrow(fluxes)` samples.
+The samples are broken down across passages and treatments as in the following table:
+
+```{r}
+fluxes %>% 
+  group_by(flux_date, treat) %>% 
+  summarize(n = n_distinct(rep), .groups='drop') %>% 
+  kable()
+```
+
+# Reponse to Selection
+
+This script takes in the flux data. It then plots the raw data
+and the log-transformed data. Finally, it fits a linear model to the flux data
+over time per treatment and tests a difference of slope between the positive
+and the neutral treatment.
+
+## Plot the raw data
+
+```{r plot_raw}
+ratio <- calc_plot_ratios(fluxes$passage, fluxes$ch4)
+
+ggplot(fluxes, aes(x = passage, y = ch4, color = treat)) +
+    stat_smooth(method = 'lm', se = FALSE, formula = 'y ~ x') +
+    scale_color_manual(
+        name = "Treatment", 
+        labels = c('Neutral', 'Positive'), 
+        values = c('gray40', 'darkorange2')) +
+    geom_jitter(width = 0.1) + 
+    labs(x = "Passage Number", 
+         y = "Methane Oxidation Rate (-k)") +
+    coord_fixed(ratio) +
+  scale_x_continuous(labels = 1:5)
+```
+
+## Plot the log-transformed data
+
+```{r plot_log10}
+ratio_log10 <- calc_plot_ratios(fluxes$passage, fluxes$log_ch4)
+
+(p_response <- ggplot(fluxes, aes(x = passage, y = log_ch4, color = treat)) +
+    stat_smooth(method = 'lm', se = FALSE, formula = 'y ~ x') +
+    scale_color_manual(
+      name = "Treatment", 
+      labels = c('Neutral', 'Positive'), 
+      values = c('gray40', 'darkorange2')) +
+    geom_jitter(#aes(alpha = selected), 
+      width = 0.1) + 
+        labs(x = "Passage Number", 
+                  y = expression(log[10] ~ CH[4] ~ "(-k)")) +
+        coord_fixed(ratio_log10) +
+  scale_x_continuous(labels = 1:5))
+
+save(p_response, file = here(der_dir, 'response_plot.RData'))
+
+# ggsave(here(fig_dir, 'fig1.jpg'), p_response, width = 4, height = 3) 
+# ggsave(here(fig_dir, 'fig1.svg'), p_response, width = 4, height = 3)
+
+```
+
+## Test a difference of slopes
+
+```{r raw_model}
+raw_model <- lm(ch4 ~ passage * treat, data = fluxes)
+
+ggplot(raw_model, aes(x = .fitted, y = .resid)) + geom_point()
+
+```
+
+```{r response_model}
+
+response_model <- lm(log_ch4 ~ passage * treat, data = fluxes)
+
+ggplot(response_model, aes(x = .fitted, y = .resid)) + geom_point()
+
+# Percent increase in CH4 per passage
+# Back calculate the slope of the positive treatment
+
+pos_slope <- 
+    response_model$coef[["passage:treatp"]]
+
+per_slope <-
+    (10^pos_slope - 1) * 100
+
+```
+
+```{r no_treat_model}
+# No treatment model
+no_treat_model <- lm(log_ch4 ~ passage, data = fluxes)
+
+```
+
+Here I fit two models.
+The first is the full model which includes passage number, treatment, and the interaction between the two.
+The second model includes only passage with no effect of treatment.
+Comparing the models using `anova`
+
+```{r compare_models}
+# Compare interaction to no interaction
+compare_response_mods <- anova(no_treat_model, response_model)
+compare_response_mods <- tidy(compare_response_mods)
+kable(compare_response_mods)
+```
+
+Here is the full interaction model:
+
+```{r print_model}
+kable(tidy(response_model))
+```
+
+```{r save_model}
+save(response_model, compare_response_mods, per_slope, file = here(der_dir, 'response_model.RData'))
+```
+
+The intercept for the Neutral treatment was significantly different from zero
+at `r myround(tidy(response_model)[[1, 2]], 2)` 
+(SE = `r myround(tidy(response_model)[[1, 3]], 2)`, 
+t = `r myround(tidy(response_model)[[1, 4]], 2)`, 
+p = `r myround(tidy(response_model)[[1, 5]], 2)`). 
+The Positive
+treatment started at a significantly lower methane oxidation rate than the
+Neutral treatment with a y-intercept of 
+`r myround(tidy(response_model)[[3, 2]], 2)` relative to Neutral 
+(SE = `r myround(tidy(response_model)[[3, 3]], 2)`, 
+t = `r myround(tidy(response_model)[[3, 4]], 2)`, 
+p = `r myround(tidy(response_model)[[3, 5]], 2)`). 
+There was no change in methane oxidation rate of the Neutral
+treatment over the five passages 
+(slope = `r myround(tidy(response_model)[[2, 2]], 2)`, 
+SE = `r myround(tidy(response_model)[[2, 3]], 2)`, 
+t = `r myround(tidy(response_model)[[2, 4]], 2)`, 
+p = `r myround(tidy(response_model)[[2, 5]], 2)`). 
+By contrast, there was a significant increase in the Positive
+treatment of `r myround(tidy(response_model)[[4, 2]], 2)` per passage which is
+approximately a `r round((pos_slope - 1) * 100, 0)`% increase in methane
+oxidation rate per passage 
+(SE = `r myround(tidy(response_model)[[4, 3]], 2)`, 
+t = `r myround(tidy(response_model)[[4, 4]], 2)`, 
+p = `r myround(tidy(response_model)[[4, 5]], 2)`).
+
+# Heritability
+
+This script imports the heritability data. This dataset includes the mean flux
+within treatment by generation for 1) all parents, 2) selected parents, and 3)
+offspring. Then, it plots offspring on the selected parents with the two
+treatments combined and then separately for the two treatments.
+
+## Test heritability as mid-offspring on mid-parent
+
+```{r test_treatment}
+
+fit_no_treat <- lm(offspring ~ selected, data = heritability)
+kable(tidy(fit_no_treat))
+herit_model <- lm(offspring ~ selected * treat, data = heritability)
+kable(tidy(herit_model))
+kable(tidy(anova(fit_no_treat, herit_model)))
+summary(herit_model)
+
+herit_model_n <- 
+heritability %>% 
+  filter(treat == "n") %>% 
+  lm(offspring ~ selected, data = .) %>% 
+  tidy()
+herit_model_n
+herit_model_p <- 
+heritability %>% 
+  filter(treat == "p") %>% 
+  lm(offspring ~ selected, data = .)
+summary(herit_model_p)
+```
+
+Here, I fit two models.
+The first is simply offspring flux on selected parent flux, ignoring treatment.
+There is no correlation `r print_lm(fit_no_treat, 2)`.
+Next, I added treatment and the interaction between treatment and parental flux.
+Here, there is a significant correlation between parent and offspring for both the Neutral treatment `r print_lm(herit_model, 2)` and the Positive treatment `r print_lm(herit_model, 4)`.
+Notably, the sign of the effect is reversed: negative for the Neutral treatment and positive for the Positive treatment.
+
+```{r plot_herit}
+
+n_label <- paste0('Neutral slope = ', myround(tidy(herit_model)[2, 2], 2))
+p_label <- paste0('Positive slope = ', myround(tidy(herit_model)[2, 2] + tidy(herit_model)[4, 2], 2))
+  ratio <- calc_plot_ratios(heritability$selected, heritability$offspring)
+(p_herit <- 
+  ggplot(heritability, aes(selected, offspring, color = treat)) +
+    geom_point() + 
+    # stat_poly_eq() +
+    stat_smooth(method = 'lm', se = F) +
+    # annotate("text", x = min(heritability$parental), y = max(heritability$offspring), label = p_label, hjust = 0, color = "darkorange2") +
+    # annotate("text", x = min(heritability$parental), y = max(heritability$offspring) - 0.04, label = n_label, hjust = 0, color = "gray40") +
+    labs(x = expression(Donor ~ log[10] ~ CH[4] ~ "(-k)"), 
+         y = expression(atop(Recipient, log[10] ~ CH[4] ~ "(-k)"))) +
+    scale_color_manual(name = "Treatment", 
+                       labels = c('Neutral', 'Positive'), 
+                       values = c('gray40', 'darkorange2')) +
+    coord_fixed(1))
+
+save(p_herit, file = here(der_dir, 'herit_plot.RData'))
+
+# ggsave(here(fig_dir, 'fig2.jpg'), p_herit, width = 4, height = 3)
+# ggsave(here(fig_dir, 'fig2.svg'), p_herit, width = 4, height = 3)
+
+```
+
+```{r save_heritability}
+save(herit_model, file = here(der_dir, 'herit_model.RData'))
+```
+
+```{r}
+heritability_each <- read_tsv(here(der_dir, 'heritability_each.tsv'))
+
+fit_no_treat <- lm(offspring ~ selected, data = heritability_each)
+kable(tidy(fit_no_treat))
+herit_model <- lm(offspring ~ selected * treat, data = heritability_each)
+kable(tidy(herit_model))
+kable(tidy(anova(fit_no_treat, herit_model)))
+
+herit_model_n_each <- 
+heritability_each %>% 
+  filter(treat == "n") %>% 
+  lm(offspring ~ selected, data = .) %>% 
+  tidy()
+herit_model_p_each <- 
+heritability_each %>% 
+  filter(treat == "p") %>% 
+  lm(offspring ~ selected, data = .) %>% 
+  tidy()
+
+ratio <- calc_plot_ratios(heritability_each$selected, heritability_each$offspring)
+(p_herit_each <- 
+ggplot(heritability_each, aes(selected, offspring, color = treat)) +
+    # theme_classic() +
+    # geom_rangeframe(color = 'black') +
+    geom_point() + 
+    stat_smooth(method = 'lm', se = F) +
+    # stat_poly_line(se = T) +
+    # stat_poly_eq(aes(label = paste(after_stat(eq.label),
+                                 # after_stat(rr.label), sep = "*\", \"*"))) +
+    labs(x = expression(atop(Mean ~ Donor, log[10] ~ CH[4] ~ "(-k)")), 
+         y = expression(atop(Recipient, log[10] ~ CH[4] ~ "(-k)"))) +
+    scale_color_manual(name = "Treatment", 
+                       labels = c('Neutral', 'Positive'), 
+                       values = c('gray40', 'darkorange2')) +
+    coord_fixed(1))
+
+# save(p_herit_each, file = here(der_dir, 'herit_plot.RData'))
+# save(herit_model, file = here(der_dir 'herit_model.RData'))
+# 
+# ggsave(here(fig_dir, 'fig2_each.jpg'), p_herit_each, width = 4, height = 3) 
+# ggsave(here(fig_dir, 'fig2_each.svg'), p_herit_each, width = 4, height = 3) 
+```
+
+# Community Analysis
 
@@
 
-```{r remove_chloroplasts, eval=FALSE}
+```{r remove_chloroplasts_mitochondria}
 
@@
 
-```
-
-# Exploratory
+mitochondria_seqs <-
+  aes_data %>% 
+  filter(Family == 'Mitochondria') %>%
+  select(asv, seq) %>% 
+  distinct() %$%
+  seq
+
+# 99 ASVs were identified as Mitochondria
+
+aes_data <- 
+  aes_data %>% 
+  filter(Family != 'Mitochondria')
+
+```
+
+## Exploratory
 
@@
   select(-data)
-
-a_analysis %>%
-  select(sample, n_seqs, sobs, rare) %>%
-  pivot_longer(-c(sample, n_seqs)) %>%
-  ggplot(aes(x=n_seqs, y=value, color=name)) +
-  geom_point() +
-  geom_smooth(se=FALSE) +
-  coord_cartesian(ylim=c(0, 6000)) +
-  theme_classic() +
-  labs(x = "Sequencing Depth", y = "Alpha Diversity")
 
@@
 
-## Richness
-
-### Plot
+### Richness
+
+#### Plot
 
@@
 
-### Hypothesis Test
+#### Hypothesis Test
 
@@
 min_seqs <- paste(substr(as.character(min_seqs), 1, 3), substr(as.character(min_seqs), 4, 6), sep = ",")
-save(kt, wt, rare_median, s5, min_seqs, file = "Output/richness_models.RData")
+save(kt, wt, rare_median, s5, min_seqs, file = here(der_dir, "richness_models.RData"))
 
@@
 
-# Beta Diversity
+## Beta Diversity
 
@@
 
-if(file.exists("Output/rare_dist_bray.rds")) {
+if(file.exists(here(der_dir, "rare_dist_bray.rds"))) {
   rare_dist_bray <- 
-    readRDS("Output/rare_dist_bray.rds")
+    readRDS(here(der_dir, "rare_dist_bray.rds"))
 } else {
@@
   rare_dist_bray %>% 
-    saveRDS("Output/rare_dist_bray.rds")
+    saveRDS(here(der_dir, "rare_dist_bray.rds"))
 }
@@
 
-save(p_beta, file = "Output/beta_fig.RData")
-
-ggsave('Manuscript/fig3.png', p_beta, height = 2.5, width = 4)
-
-ggsave('Manuscript/fig3.svg', p_beta, height = 2.5, width = 4)
+save(p_beta, file = here(der_dir, "beta_fig.RData"))
+
+ggsave(here(fig_dir, 'fig3.png'), p_beta, height = 2.5, width = 4)
+
+ggsave(here(fig_dir, 'fig3.svg'), p_beta, height = 2.5, width = 4)
 
@@
 
-save(tidy_permanova, file = "Output/beta_model.RData")
+save(tidy_permanova, file = here(der_dir, "beta_model.RData"))
 
@@
 
-## dbrda methane in passage 5
+### dbrda methane in passage 5
 
@@
 
-if(file.exists("Output/rare_dist_bray_p5.rds")) {
+if(file.exists(here(der_dir, "rare_dist_bray_p5.rds"))) {
   rare_dist_bray_p5 <- 
-    readRDS("Output/rare_dist_bray_p5.rds")
+    readRDS(here(der_dir, "rare_dist_bray_p5.rds"))
 } else {
@@
   rare_dist_bray_p5 %>% 
-    saveRDS("Output/rare_dist_bray_p5.rds")
+    saveRDS(here(der_dir, "rare_dist_bray_p5.rds"))
 }
@@
 
-save(dbrda_aov, dbrda_percent_explained, file = "Output/dbrda.RData")
-
-```
-
-# Differential Abundance
-
-## Unique to each treatment
-
-### Family
+save(dbrda_aov, dbrda_percent_explained, file = here(der_dir, "dbrda.RData"))
+
+```
+
+## Differential Abundance
+
+### Unique to each treatment
+
+#### Family
 
@@
 
-### Genus
+#### Genus
 
@@
 
-### Species
+#### Species
 
@@
 ```{r}
+# 
+# # Install ANCOMBC
+# if (!requireNamespace("BiocManager", quietly=TRUE))
+#     install.packages("BiocManager")
+# BiocManager::install("ANCOMBC")
+# 
+# # Load ANCOM
+# library(ANCOMBC)
 
 # Setup phyloseq object
-aes_ps <- phyloseq(otu_table(otu_table, taxa_are_rows=TRUE), sample_data(sample_data))
-
-if(file.exists("Output/ancom_out.rds")) {
+aes_ps <- phyloseq(otu_table(otu_table, taxa_are_rows=TRUE),
+                   sample_data(sample_data))
+
+# Save/Load output so you don't have to rerun ANCOM
+if(file.exists(here(der_dir, "ancom_out.rds"))) {
   out <- 
-    readRDS("Output/ancom_out.rds")
+    readRDS(here(der_dir, "ancom_out.rds"))
 } else {
@@
   out %>% 
-    saveRDS("Output/ancom_out.rds")
+    saveRDS(here(der_dir, "ancom_out.rds"))
 }
@@
 
-if(file.exists("Output/aldex_x_all.rds")) {
+if(file.exists(here(der_dir, "aldex_x_all.rds"))) {
   x.all <- 
-    readRDS("Output/aldex_x_all.rds")
+    readRDS(here(der_dir, "aldex_x_all.rds"))
 } else {
@@
   x.all %>% 
-    saveRDS("Output/aldex_x_all.rds")
+    saveRDS(here(der_dir, "aldex_x_all.rds"))
 }
@@
 
-save(p_log10, da_fams, file = "Output/da_fig.RData")
-
-ggsave('Manuscript/fig4.jpg', p_log10, width = 5, height = 5) 
-ggsave('Manuscript/fig4.svg', p_log10, width = 5, height = 5)
+save(p_log10, da_fams, file = here(der_dir, "da_fig.RData"))
+
+ggsave(here(fig_dir, 'fig4.jpg'), p_log10, width = 5, height = 5) 
+ggsave(here(fig_dir, 'fig4.svg'), p_log10, width = 5, height = 5)
 
 ## All
-# ggsave("shared_taxa.png", p)
-# ggsave("shared_taxa_log.png", p_log10, width = 5, height = 5)
+# ggsave(here(fig_dir, "shared_taxa.png"), p)
+# ggsave(here(fig_dir, "shared_taxa_log.png"), p_log10, width = 5, height = 5)
 
@@
 ```{r corncob-methanotrophs}
-if(file.exists("Output/cc_family.rds")) {
-  cc_family <- 
-    readRDS("Output/cc_family.rds")
+if(file.exists(here(der_dir, "cc_family.rds"))) {
+  cc_family <- readRDS(here(der_dir, "cc_family.rds"))
 } else {
@@
                                 fdr_cutoff = 0.05, B = 1000)  
-cc_family %>% 
-    saveRDS("Output/cc_family.rds")
+cc_family %>% saveRDS(here(der_dir, "cc_family.rds"))
 }
@@
 
-if(file.exists("Output/cc_genus.rds")) {
+if(file.exists(here(der_dir, "cc_genus.rds"))) {
   cc_genus <- 
-    readRDS("Output/cc_genus.rds")
+    readRDS(here(der_dir, "cc_genus.rds"))
 } else {
@@
 cc_genus %>% 
-    saveRDS("Output/cc_genus.rds")
+    saveRDS(here(der_dir, "cc_genus.rds"))
 }

### 2022-12-06 manuscript/freewrite.Rmd
reorganize working directory and add new analysis and text

---
title: Artificial ecosystem selection reveals relationships between microbiome composition and ecosystem function
author: "Andrew H. Morris^1^* and Brendan J. M. Bohannan^1^"
output:
  word_document:
    reference_docx: reference.docx
bibliography: bibliography.bib
csl: vancouver.csl
---

^1^ Institute of Ecology and Evolution, University of Oregon, Eugene, OR, USA  
\* Correspondence: Andrew H. Morris, amorris3\@uoregon.edu  
Competing Interests: We declare we have no competing interests.

```{r options, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
#knitr::opts_knit$set(root.dir=normalizePath('../'))

```

# Questions

- What should be the order of the Intro?

    1. What are we interested in and why
    2. Past approaches and their limitations
    3. Our approach and why it's better
    4. Broader conceptual framework (Figure 1, Equation 1)

- Or:

    1. What are we interested in and why
    2. Past approaches and their limitations
    3. Broader conceptual framework (Figure 1, Equation 1)
    4. Our approach and why it's better

- Or:

    1. What are we interested in and why
    2. Broader conceptual framework (Figure 1, Equation 1)
    3. Past approaches and their limitations
    4. Our approach and why it's better

# Introduction

## What is the knowledge gap?

### What are we interested in, why are we interested in it, how has it been and studied and what are the shortfalls, and why precisely is our approach better?

Microbiomes mediate a variety of important ecosystem functions. This suggests
that understanding variation in microbiome composition might be important for
predicting the rates of ecosystem functions. Many studies have attempted to
document such a relationship. However, it is difficult to isolate the effect of
microbiome composition from other drivers of variation in ecosystem function.
For example, microbiome manipulations often confound "diversity" with broad
phylogenetic patterns or "composition" with abundance. Likewise, comparative
studies must control for abiotic variation statistically to identify the direct
effect of the microbiome, which relies on measuring the correct abiotic
variables and cannot account for unmeasured environmental variation. Here, we
overcome these limitations by using a selection approach to estimate the degree
to which an ecosystem function varies with microbiome composition. We want to
understand what is the relationship between _variation in the microbiome_ and
_variation in the rate of ecosystem function_?

There have been two general categories of approaches that investigators have
applied to address this question: comparative and manipulative. Comparative
studies sample natural variation in an ecosystem function across different
habitats and simultaneously measure variation in community composition.
Investigators can then correlate ecosystem function with community structure
while trying to control for environmental variation. These approaches have
documented important relationships between microbiomes and ecosystem
functions. For example, a meta-analysis of these studies observed a small but
significant unique contribution of the microbiome to variation in ecosystem
function after controlling for environmental variation (Graham 2016). In
addition, studies focusing on the correlation between the rate of an ecosystem
function and the abundance of an associated functional marker gene observe a
significant correlation, though this relationship is rare and contingent upon
both the function and the ecosystem sampled (Rocca 2015). Comparative studies
come with some unique challenges and limitations. One issues is the covariance
problem where microbiome attributes tend to covary with the abiotic conditions
within an environment and whether and how the investigator controls for these
variables affects the conclusions. In addition, it is difficult to know a
priori which environmental variables or community attributes to measure.
Finally, while these approaches can establish a potential magnitude and
direction for these relationships, they fail to establish a mechanism
underlying the relationship.

The other broad category of approaches to address this question are
manipulative approaches. Manipulative experiments try to alter microbial
community structure and observe the effect on function. For example,
reciprocal transplant and common garden experiments have shown that
microbiomes originating from different ecosystems inoculated onto the same
substrate or introduced into a common environment display distinct functional
rates (Cavigeli 2000, Balser 2005, Waldrop 2006, Strickland 2009). In
addition, manipulating diversity by filtering communities by cell size or
through dilution alters ecosystem function (Wagg 2014, 2020, dilution
citation?). However, manipulating the microbiome directly is challenging. Each
of the manipulative approaches applied confound community structure with other
factors such as abiotic conditions or phylogenetic relationships. For example,
reciprocal transplant and common garden experiments confound community
structure with the abiotic conditions carried on the inoculum (Reed and
Martiny 2007). In addition, manipulating microbial diversity through filtering
confounds diversity with broad phylogenetic categories such as nematodes,
fungi, and bacteria while manipulating diversity through dilution confounds
diversity with biomass and composition.

In this study, we sought to build on the observations of comparative and
manipulative studies by applying a different approach to the question of
whether variation in the microbiome contributes to variation in the rate of an
ecosystem function. We used artificial ecosystem selection to select for
microbiomes that performed a greater rate of ecosysem function. We then tested
whether variation in the microbiome contributed to variation in the rate of
ecosystem function and identified microbiome attributes that might explain
this relationship. There are several potential advantages to this approach for
documenting BEF relationships and for investigating potential mechanisms
underlying those relationships. In particular, by passaging microbiomes over
multiple rounds using a common substrate (or isogenic host), we can weaken the
covariance between microbes and the environment by repeatedly diluting the
abiotic effect. In addition, by inoculating our ecosystems with equally sized
subsamples, we eliminate the need to generate microbiome variation through
methods that are confounded with diversity, biomass, or cell size. Lastly, by
comparing our artificially selected community to a control community without
selection or with random selection, we can both control for changes in the
environment over time and identify significant genes or taxa associated with our
function of interest to establish a potential mechanism that explains this
variation.

One important conceptual distinction that we should make when investigating
biodiversity-ecosystem function relationships is to be clear that we are
interested in _variation_. In particular, we are interested in what drives
variation in the rate of an ecosystem function across ecosystems. For example,
we want to know whether variation in the rate of methane oxidation rate in soils
is driven by variation in microbial community structure. This question is
distinct from the question "Which microbes consume methane?" To illustrate this
distinction, we can draw an analogy with the genetic basis of traits in
organisms. If we are interested in whether variation in human height is
associated with variation in genotype, we can sample a population of humans,
measure their height, and look at the correlation between phenotype and genotype
using gene markers or proxies for genotypic similarity, such as kinship.
However, this does not address the question of "Which genes control human
height?" Likewise, we could go into the lab and perform an experiment on mice
where we knockout or over-express a gene thought to be associated with height
and observe height differences among the treatment groups. This would
demonstrate that a particular gene affects height, but it does _not_ demonstrate
that variation in genotype drives variation in height in a population.
___WHY DOES THIS MATTER?___

To understand variation in the rate of an ecosystem function, we can think about
different drivers of variation as illustrated in Figure 1 and Equation 1. In
this model, variation in the rate of an ecosystem function ($V_F$) is the sum of
the variation in the environmental conditions ($V_E$), variation in the
microbiome ($V_M$), and $\epsilon$, which represents variation unexplained by
the other two factors, technical error, and unmeasured environmental or
microbial variation. In addition, we can consider another variable,
$cov(V_E:V_M)$, which is the covariance between the microbiome and the
environment. We can conceptualize this as a situation in which a particular
environmental condition, such as pH or moisture content, selects for a particular
microorganism that is associated with variation in function. For simplicity, we
have omitted a final parameter, which is the interaction between the microbiome
and the environment, which would represent a change in the effect of the
microbiome under different environmental conditions. Here, we assume we're
talking about a single ecosystem and so assume this parameter to be zero.

![Figure 1. Conceptual diagram of the potential relationship between environmental variation, microbiome variation, and variation in the rate of an ecosystem function. We omitted the reverse arrows for illustration, but in theory those relationships also exist but we make the assumption that they are small. In addition, this diagram represents the relationships for a single environment, but one could imagine the same microbiome exerting a different influence on variation in function in a second environment, which would be illustrated by the interaction between microbiome and environment ($V_E:V_M$)`](Figure1.png)

$$
V_F = V_E + V_M + COV(V_E:V_M) + \epsilon
$$

How can we use this conceptual framework to address our original question: Does
variation in the microbiome contribute to variation in the rate of ecosystem
function? Under the null hypothesis that we can ignore the microbiome and
simply use abiotic environmental variables to predict function, we would assume
that $V_M$ is essentially zero and that all of the variation in function
($V_F$) is due to variation in the environment and the covariance between the
microbiome and the environment ($V_E + COV(V_E:V_M)$). In this case, it is
unnecessary to understand anything about the microbiome. However, if we consider
the results from Graham (2016), this hypothesis does not appear to match the
data. Graham observed that across the studies that they analyzed, environmental
variables explained an average of 56\% of the variation in ecosystem function
while microbial community structure explained an average of 31\% of the
variation. Models that included both explained an average of 65\% of the
variation. Putting these numbers together, we see that $V_M$ = 11\% (65 - 56),
$cov(V_E:V_M)$ = 20\% (31 - $V_M$ = 31 - 11), and $V_E$ = 36\% (56 -
$cov(V_E:V_M)$ = 56 - 20). Therefore, the environment explains 36% of the
variation directly
and 20% via covariation with the microbiome and the microbiome explains 11%
directly. This means at least 1/6 (11/65) of the explained variation can be
uniquely attributed to the microbiome according to the conclusions of Graham.

Of course, these numbers should be taken with a grain of salt. As discussed
earlier, comparative studies such as these are unable to break the covariance
between microbes and the environment. In this case, $V_M$ and $\epsilon$ may
represent the direct or indirect effect (via covariance) of unmeasured
environmental variables rather than a unique contribution of the microbiome. In
addition, the proportion of variation depends upon the magnitude of the
variation in $V_F$ and $V_E$. For example, if we were to include variation in
methane emissions of wetland and upland ecosystems, we might find that virtually
100\% of the variation in function was due to variation in the environment,
since methane production requires anaerobic conditions. However, if we looked
across wetlands, we might observe a smaller $V_F$ and $V_E$ and an equivalent
$V_M$. Under this scenario, even though $V_M$ was the same, $\frac{V_M}{V_F}$
would be much greater and therefore the environment would no longer explain
100\% of the variation. Therefore, any study that wishes to address this
question must consider the variation in the population that they are studying.

We performed artificial ecosystem selection on soil microbiomes by selecting on
the methane oxidation rate. To perform this experiment, we created 24 jars in an
identical manner using sterile potting mix inoculated with living soil. Each jar
was randomly assigned to either a positive selection treatment that underwent
selection on high soil methane oxidation rate or a random selection treatment.
At the end of each incubation, we selected jars with a high methane oxidation
rate (for the positive treatment), homogenized them, and used this homogenate to
inoculate a new set of sterile soil jars. We repeated this process over multiple
passges. The idea is that by shuffling microbiome composition through
subsampling and homogenizing the environment we are exploring the parameter
space for the bidoversity-ecosystem function relationship. In addition,
subsampling at each step would dilute away the effect of the abiotic
environment. The only way for there to be a response to selection would be for
microbes to actually grow in the medium.

Here are some hypothetical results we could observe. If soil methane oxidation
rate is entirely environmentally determined, i.e., in the parlance of our
conceptual framework that would mean that $V_E$ and $cov(V_E:V_M)$ are large and
that $V_M$ is relatively small, then we would expect to observe no response to
selection. This is because the effect of the abiotic environment would be
diluted at eaching passaging step. If variation in methane oxidation rate was
independent of microbes and the environment in our system then the varation
should be random and $\epsilon$ would be large relative to the other parameters
and again we would observe no response to selection. If variation in methane
oxidation rate is at least partially driven by variation in the microbiome then
we would expect manipulating microbiome composition through selection would
result in an increase in methane oxidation rate as a response to selection. This
would manifest as a correlation between donor and recipient jars in consecutive
passages.

(Aside: What we seem to be saying in the above paragraph is that passaging
ecosystem propagules is somehow intrinsically tied to the magnitude of Vm. This
seems to be based on the argument that Vm reproduces itself with high fidelity,
but Ve does not. Is this true? Adding a low pH 5g of soil to a neutral pH 45g of
soil doesn't shift the pH, but adding a 5g soil microbial inoculum to 45g of
autoclaved soil does shift the community? Is this any different than the
manipulative experiments we describe above? Or perhaps the argument is that
_repeated_ passaging dilutes away the environmental effect so by the time you
get to passage 5 it's 1/1e5 of the original soil inoculum, but if the microbes
have reproduced then they have rebuilt their population at each passage. I guess
this is the idea behind serial batch culturing??? IDK if a reviewer will be
convinced by this argument.)

### 2022-12-06 manuscript/manuscript.Rmd
reorganize working directory and add new analysis and text

@@
 CH~4~ concentrations and were incubated at ambient temperature for
-approximately three weeks. Methane oxidation rates were determined at the end
+approximately four weeks. Methane oxidation rates were determined at the end
 of the incubation period. For the positive treatment, the three jars with the

### 2022-12-06 R/functions.R

@@ -1,9 +1,15 @@
 
-raw_dir <- 'Data/' # Raw Data Directory
-der_dir <- 'Output/' # Derived/Modified Data Directory
+# Directory Definitions
+
+raw_dir <- 'data/'      # Raw Data Directory
+der_dir <- 'analysis/output/'    # Derived/Modified Data Directory
+man_dir <- 'manuscript/'       # Manuscript directory
+fig_dir <- 'analysis/figs/'      # Figures directory
+
+# Set global ggplot theme
 
 theme_set(theme_classic() +
-            theme(panel.border = element_rect(fill = NA, size = 1.0625),
+            theme(panel.border = element_rect(fill = NA, linewidth = 1.0625),
                   axis.line = element_blank()))
 
 plot_flux <- function(fluxes, passage, estimate, ratio = NULL, log10 = FALSE) {
@@ -32,16 +38,15 @@ plot_flux <- function(fluxes, passage, estimate, ratio = NULL, log10 = FALSE) {
   return(p)
 }
 
-calc_plot_ratios <- function(x, y) {
-  (max(x)-min(x))/(max(y)-min(y))
-}
+
 
 
 print_lm <- function(fit, term = 1, estimate = 'slope') {
-  paste0('(', estimate, ' = ', myround(tidy(fit)[term, 2], 2), 
-        ', SE = ', myround(tidy(fit)[term, 3], 2), 
-        ', t = ', myround(tidy(fit)[term, 4], 2), 
-        ', p = ', myround(tidy(fit)[term, 5], 2), ')')
+  fit <- tidy(fit)
+  paste0('(', estimate, ' = ', myround(fit[term, 2], 2), 
+        ', SE = ', myround(fit[term, 3], 2), 
+        ', t = ', myround(fit[term, 4], 2), 
+        ', p = ', myround(fit[term, 5], 2), ')')
 }
 
 print_lrt <- function(fit, term = 2, test = "Likelihood ratio test") {
@@ -50,19 +55,6 @@ print_lrt <- function(fit, term = 2, test = "Likelihood ratio test") {
          ', p = ', myround(tidy(fit)[term, 5], 2))
 }
 
-
-plot_herit <- function(heritability, ratio) {
-  ratio <- calc_plot_ratios(heritability$selected, heritability$offspring)
-  ggplot(heritability, aes(selected, offspring, color = treat)) +
-    geom_jitter(width = 0.01) + 
[…]

### 2022-12-06 analysis/new_analysis.R

@@ -1,0 +1,90 @@
+# Response to selection
+0.1779718
+# Standard Error
+0.0643900
+ratio_log10 <- calc_plot_ratios(fluxes$passage, fluxes$log_ch4)
+
+ggplot(fluxes, aes(x = passage, y = log_ch4, color = treat)) +
+    stat_smooth(method = 'lm', se = FALSE, formula = 'y ~ x') +
+    scale_color_manual(
+        name = "Treatment", 
+        labels = c('Neutral', 'Positive'), 
+        values = c('gray40', 'darkorange2')) +
+    geom_jitter(#aes(alpha = selected), 
+        width = 0.1) + 
+    labs(x = "Passage Number", 
+         y = expression(log[10] ~ CH[4] ~ "(-k)")) +
+    coord_fixed(1) +
+    scale_x_continuous(labels = 1:5)
+
+model1 <- lm(log_ch4 ~ passage * treat, data = fluxes)
+summary(model1)
+
+divergence <-
+    fluxes %>% 
+    select(passage, treat, log_ch4) %>% 
+    group_by(passage, treat) %>% 
+    summarize(mean = mean(log_ch4), .groups = "drop") %>% 
+    mutate(treat = factor(treat)) %>% 
+    pivot_wider(names_from = "treat", values_from = "mean") %>% 
+    mutate(divergence = p - n)
+
+ratio_div <- calc_plot_ratios(divergence$passage, divergence$divergence)
+
+ggplot(divergence, 
+       aes(x = passage, 
+           y = divergence)) +
+    stat_smooth(method = 'lm', 
+                se = FALSE, 
+                formula = 'y ~ x',
+                color = "black",
+                linewidth = 0.5) +
+    geom_point() + 
+    labs(x = "Passage Number", 
+         y = expression(atop("Divergence", log[10] ~ CH[4] ~ "(-k)"))) +
+    coord_fixed(ratio_div) +
+    scale_x_continuous(labels = 1:5)
+
+model2 <- lm(divergence ~ passage, data = divergence)
+summary(model2)
+
+sd <- 
+    heritability %>% 
+    mutate(sd = selected - parental) %>% 
+    select(passage, treat, sd) %>% 
+    pivot_wider(names_from = 'treat', values_from = 'sd') %>% 
+    mutate(sd = p - n) %>% 
+    select(-n, -p)
+
+sd$cumulative_sd <- 
[…]

### 2022-12-06 analysis/scripts/00_packages.R

@@ -1,12 +1,11 @@
-library(devtools)
 
 # Install specific commit/version from github
-install_github("amorris28/morris", dependencies = FALSE)
-install_github("kbroman/broman", dependencies = FALSE)
-install_github("FrederickHuangLin/ANCOMBC", dependencies = FALSE)
+devtools::install_github("amorris28/morris", dependencies = FALSE)
+devtools::install_github("kbroman/broman", dependencies = FALSE)
+devtools::install_github("FrederickHuangLin/ANCOMBC", dependencies = FALSE)
 
-# if (!requireNamespace("BiocManager", quietly = TRUE))
-#   install.packages("BiocManager")
+if (!requireNamespace("BiocManager", quietly = TRUE))
+  install.packages("BiocManager")
 
 renv::install("bioc::ALDEx2")
 renv::install("bioc::phyloseq")

### 2022-12-06 analysis/scripts/01-calculate-flux.R

@@ -1,119 +1,46 @@
----
-title: "Process Flux Data"
-author: "Andrew Morris"
-date: "`r Sys.Date()`"
-output: html_document
----
-```{r setup, include=FALSE}
 
-knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = TRUE)
-knitr::opts_knit$set(root.dir=normalizePath('../'))
+# This script processes the standard curve data from the gas chromatograph on
+# each sample date. It then converts area-under-the-curve to methane
+# concentration using the standard curves for every sample. Finally, it computes
+# fluxes (change in concentration over time) as a first-order exponential decay
+# function.
 
-```
-
-This script processes the standard curve data from the gas chromatograph on
-each sample date. It then converts area-under-the-curve to methane
-concentration using the standard curves for every sample. Finally, it
-computes fluxes (change in concentration over time) as a first-order
-exponential decay function.
-
-```{r initialize_packages, message=FALSE}
 library(tidyverse)
 library(broom)
 library(lubridate)
-library(knitr)
-```
-
-```{r initialize_variables}
 source('R/functions.R')
-```
 
-
-```{r import_data}
+# import_data
 sc <- read.csv(paste0(raw_dir, 'standard_curve.csv'))
 sc_dates <- read.csv(paste0(raw_dir, 'sc_dates.csv'))
 time_data <- data.table::fread(paste0(raw_dir, 'time_data.csv'))
 conc_data <- data.table::fread(paste0(raw_dir, 'conc_data.csv'))
 selected <- read.csv(paste0(raw_dir, 'selected.csv'))
-```
-
-This dataset includes `r length(unique(substr(conc_data$jar, 1, 1)))` treatments across `r length(unique(conc_data$flux_date))` passages (generations). This includes a total of n = `r nrow(conc_data)` samples. However, passage 6 will not be included in the analyzed dataset, because that passage was sampled much later due to covid and used different potting mix so has too many confounding factors. The samples are broken down across passages and treatments as in the following table:
-
-
-```{r echo=FALSE}
-conc_data %>% 
-  mutate(treat = substr(conc_data$jar, 1, 1),
[…]

### 2022-12-06 analysis/scripts/02-process-herit.R

@@ -1,43 +1,26 @@
----
-title: "Process Heritability Data"
-author: "Andrew Morris"
-date: "`r Sys.Date()`"
-output: html_document
----
-```{r setup, include=FALSE}
 
-knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = TRUE)
-knitr::opts_knit$set(root.dir=normalizePath('../'))
+# This script takes in the flux data from script 01 and the data on which jars
+# were selected in each passage. It then aligns these data so that for each
+# passage there is a mean flux for all parents, the selected parents, and all
+# offspring.
 
-```
-This script takes in the flux data from script 01 and the data on which jars
-were selected in each passage. It then aligns these data so that for each
-passage there is a mean flux for all parents, the selected parents, and all
-offspring. 
-
-```{r initialize_packages, message=FALSE}
 library(tidyverse)
 library(knitr)
-```
 
-```{r initialize_variables}
 source('R/functions.R')
-```
 
-```{r import_data, message=FALSE}
+# import data
+
 selected_jars <- read_csv(paste0(raw_dir, 'selected.csv'))
 fluxes <- read_tsv(paste0(der_dir, 'fluxes.tsv'))
-```
 
 ## Organize parent/offspring values
 
-```{r process_herit_data}
 # Pull out the selected parental jars
 selected <- 
   fluxes %>% 
   filter(selected) %>% 
   select(passage, treat, log_ch4)
-  
 
 # Pull out all parental jars
 parental <-
@@ -79,9 +62,9 @@ heritability <-
   filter(passage != 5) %>% # Remove passage 5, which has no offspring,
   # because it's the last generation
   ungroup()
-```
 
-```{r}
+
+
 offspring_each <- 
   offspring %>% 
   group_by(passage, treat) %>% 
@@ -95,13 +78,5 @@ heritability_each <-
 
 write_tsv(heritability_each, paste0(der_dir, 'heritability_each.tsv'))
 
-```
 
-
-```{r print_data}
-kable(heritability)
-```
-
-```{r export_data}
 write_tsv(heritability, paste0(der_dir, 'heritability.tsv'))
-```

### 2022-12-06 analysis/scripts/03-create-barcode-key.R

@@ -13,8 +13,8 @@
 
 library(tidyverse)
 
-sample_key <- read_tsv('Data/sample_key.tsv')
-primer_key <- read_tsv('Data/primer_sequence_key.tsv')
+sample_key <- read_tsv('data/sample_key.tsv')
+primer_key <- read_tsv('data/primer_sequence_key.tsv')
 
 sample_key %>% 
   separate(primer_well, c('row', 'col'), 1) %>%  
@@ -22,11 +22,11 @@ sample_key %>%
   rename(r_barcode = sequence) %>% 
   left_join(primer_key, by = c('row' = 'key')) %>% 
   rename(f_barcode = sequence) %>%
-  write_tsv('Output/barcode_master.tsv')
+  write_tsv('analysis/output/barcode_master.tsv')
 
-read_tsv('Output/barcode_master.tsv') %>% 
+read_tsv('analysis/output/barcode_master.tsv') %>% 
   select(`Library Name / ID` = gc3f_id, 
          `Index 1 (i7) sequence` = r_barcode, 
          `Index 2 (i5) sequence` = f_barcode) %>% 
-  write_tsv('Output/barcode_gc3f.tsv')
+  write_tsv('analysis/output/barcode_gc3f.tsv')
 

### 2022-12-06 analysis/scripts/04-create-asvs.R

@@ -118,5 +118,5 @@ taxa.print <- taxa # Removing sequence rownames for display only
 rownames(taxa.print) <- NULL
 head(taxa.print)
 
-save(seqtab.nochim, taxa, file = "HPC_Output/dada2_output.RData")
+save(seqtab.nochim, taxa, file = "output/dada2_output.RData")
 

### 2022-12-06 analysis/scripts/05-combine-asvs-and-fluxes.R

@@ -1,37 +1,18 @@
----
-title: "Sequence data processing"
-author: "Andrew Morris"
-date: "9/13/2021"
-output: html_document
-editor_options: 
-  chunk_output_type: console
----
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
-knitr::opts_knit$set(root.dir=normalizePath('../'))
-```
-
-
-```{r initialize-packages}
 library(tidyverse)
 #library(phyloseq)
 library(decontam)
 #library(microViz)
 library(knitr)
 library(vegan)
-```
+source('R/functions.R')
 
-```{r import-data}
-load(file = "HPC_Output/dada2_output.RData")
+load(file = paste0(der_dir, "dada2_output.RData"))
 
-samples <- read_tsv('Output/barcode_master.tsv')
-fluxes <- read_tsv('Output/fluxes.tsv')
-qubit <- read_tsv('Data/post_pcr_qubit.tsv')
+samples <- read_tsv(paste0(der_dir, "barcode_master.tsv"))
+fluxes <- read_tsv(paste0(der_dir, "fluxes.tsv"))
+qubit <- read_tsv(paste0(raw_dir, "post_pcr_qubit.tsv"))
 
-```
-
-```{r data-munging}
 
 metadata <-
   samples %>% 
@@ -70,14 +51,12 @@ aes_data <-
   inner_join(metadata, by = "sample") %>% 
   inner_join(taxa, by = "asv")
 
-```
 
 # Alpha Diversity for All Data
 
-Here is alpha diversity for each treatment, positive/negative control, etc.
-In addition, there is alpha diversity by row and column of the PCR plate. I don't see any relationship between PCR plate location and alpha diversity, which is good. It looks like the PCR positive control, which was a pure culture of E. Coli, has particularly low diversity, which is good. Let's pull that sample out and look at it.
+# Here is alpha diversity for each treatment, positive/negative control, etc.
+# In addition, there is alpha diversity by row and column of the PCR plate. I don't see any relationship between PCR plate location and alpha diversity, which is good. It looks like the PCR positive control, which was a pure culture of E. Coli, has particularly low diversity, which is good. Let's pull that sample out and look at it.
 
-```{r alpha-diversity-treatment}
 
 aes_data %>%
[…]

### 2022-12-06 analysis/scripts/compare-rare-methods.R

@@ -1,0 +1,197 @@
+library(vegan)
+library(breakaway)
+library(magrittr)
+library(zCompositions)
+library(tidyverse)
+source('R/functions.R')
+
+set.seed(19760620)
+
+aes_data <- read_tsv('analysis/output/community_data.tsv')
+
+# Look at distribution of sequencing depths
+aes_data %>% 
+  group_by(sample) %>% 
+  summarize(total = sum(count)) %>% 
+  ggplot(aes(x = total)) +
+  geom_histogram()
+
+
+seq_range <- 
+  aes_data %>%
+  group_by(sample) %>% 
+  summarize(total = sum(count)) %>% 
+  summarize(seq_range = range(total)) %$%
+  seq_range
+
+# Generate a null dataset
+randomize_asv_table <- function(x) {
+    x %>%
+        uncount(count) %>%
+        mutate(rand_asv = sample(asv)) %>%
+        dplyr::select(-asv) %>%
+        count(sample, rand_asv, name="value")
+}
+rand_aes <- randomize_asv_table(aes_data)
+
+#######################
+# Alpha diversity
+#####################
+
+
+analysis <- rand_aes %>%
+  count(sample, value) %>%
+  nest(data = -sample) %>%
+  mutate(n_seqs = map_dbl(data, ~sum(.x$value * .x$n)),
+         sobs = map_dbl(data, ~sum(.x$n)),
+         rare = map_dbl(data, ~rarefy(rep(.x$value, .x$n), sample=seq_range[1])),
+         chao = map_dbl(data, ~get_chao(.x)),
+         ba = map(data, ~get_breakaway(.x))) %>%
+  select(-data) %>%
+  unnest(ba)
+
+
+analysis %>%
+  select(sample, n_seqs, sobs, chao, rare, est) %>%
+  pivot_longer(-c(sample, n_seqs)) %>%
+  ggplot(aes(x=n_seqs, y=value, color=name)) +
+  geom_point() +
+  geom_smooth(se=FALSE) +
+  coord_cartesian(ylim=c(0, 40000)) +
+  theme_classic() +
+  labs(title = "AES", x = "Sequencing Depth", y = "Alpha Diversity")
+
+analysis %>%
+  select(sample, n_seqs, sobs, chao, rare, est) %>%
+  pivot_longer(-c(sample, n_seqs)) %>%
[…]

### 2022-12-06 analysis/scripts/dna-yields.R

@@ -1,0 +1,192 @@
+library(tidyverse)
+library(broom)
+library(lubridate)
+library(knitr)
+source('R/functions.R')
+
+qubit <- read_csv('Data/extraction_qubit.csv')
+selected <- read_csv('Data/selected.csv') %>% 
+  mutate(passage = as.character(passage)) %>% 
+  mutate(selected = 1)
+
+qubit <- 
+  qubit %>% 
+  mutate(passage = substr(sample, 2, 2)) %>% 
+  mutate(treat = substr(sample, 3, 3)) %>% 
+  mutate(jar = substr(sample, 3, 5)) %>% 
+  left_join(selected) %>% 
+  mutate(selected = as.character(if_else(is.na(selected), 0, selected))) %>% 
+  mutate(sample_ugml = qubit_ngml * 200 / vol_ul)
+
+ggplot(qubit, aes(x = selected, y = qubit_ngml)) +
+  geom_boxplot() +
+  geom_jitter()
+ggplot(qubit, aes(x = treat, y = qubit_ngml, color = selected)) +
+  geom_boxplot() +
+  geom_jitter(aes(shape = passage))
+qubit %>% 
+  filter(treat == 'p', passage == '2') %>% 
+t.test(qubit_ngml ~ selected, .)
+
+
+####################
+
+sc <- read.csv(paste0(raw_dir, 'standard_curve.csv'))
+sc_dates <- read.csv(paste0(raw_dir, 'sc_dates.csv'))
+time_data <- data.table::fread(paste0(raw_dir, 'time_data.csv'))
+conc_data <- data.table::fread(paste0(raw_dir, 'conc_data.csv'))
+selected <- read.csv(paste0(raw_dir, 'selected.csv'))
+
+fluxes <- read_tsv(paste0(der_dir, 'fluxes.tsv'))
+
+sc_model <- function(df) {lm(injection_ppm ~ area, data = df)}
+
+# fit a separate model for each date
+standard_curves <- 
+  sc %>% 
+  filter(molecule == 'co2') %>% 
+  group_by(flux_date, sc_date) %>% 
+  nest() %>% 
+  mutate(model = map(data, sc_model),
+         glance = map(model, glance),
+         tidy = map(model, tidy)) 
+
+sc_r2 <- 
+  standard_curves %>% 
+  unnest(glance) %>% 
+  select(flux_date, sc_date, r.squared)
+
+sc_equation <- 
[…]

### 2023-01-04 analysis/analysis.Rmd
revice manuscript

@@
 aes_data <- read_tsv(here(der_dir, 'community_data.tsv'))
-
+```
+
+# Clean data
+
+```{r}
 # Shift the passage numbers so the first passage becomes "Passage 0"
@@
   mutate(passage = passage - 1)
+```
+
+```{r fix-treatment}
+
+aes_data <-
+  aes_data %>% 
+  mutate(passage_treat = treat) %>% 
+  separate(passage_treat, c('passage', "treat"), sep = 2, remove = FALSE) %>% 
+  select(sample, passage_treat, passage, treat, selected, ch4, asv, count, Kingdom:Species, seq)
+
+```
+
+```{r fix-taxonomy}
+
+# There are "Unknown Family" taxa with Genus and Class, but Order is Incertae Sedis.
+# I'm just going to remove everything below Class and call it "Unclassified"
+
+fill_taxonomy <- function(tab) {
+    if(is.na(tab[[1]])) tab[[1]] = "Unclassified"
+    for(i in 2:length(tab)) {
+      if(is.na(tab[[i]])){
+        if(grepl("unclassified", tab[[i-1]])) tab[[i]] = tab[[i-1]]
+        else tab[[i]] = paste(tab[[i-1]], "unclassified", sep = "_")
+      }
+    }
+    tab
+}
+
+aes_data <-
+  aes_data %>% 
+  nest(data = -c(asv, Kingdom:Species)) %>% 
+  nest(taxa = Kingdom:Species) %>% 
+  mutate(taxa = map(taxa, ~fill_taxonomy(.x))) %>% 
+  unnest(taxa) %>% 
+  unnest(data)
+
+```
+
+```{r remove_chloroplasts_mitochondria}
+
+chloroplast_seqs <- 
+  aes_data %>% 
+  filter(Order == 'Chloroplast') %>% 
+  select(asv, seq) %>% 
+  distinct() %$%
+  seq
+# 10 ASVs were identified as chloroplast in their taxonomy.
+# seq matches complete genomes of chloroplasts on NCBI so I will remove them.
+
+aes_data <- 
+  aes_data %>% 
+  filter(Order != 'Chloroplast')
+
+mitochondria_seqs <-
+  aes_data %>% 
+  filter(Family == 'Mitochondria') %>%
+  select(asv, seq) %>% 
+  distinct() %$%
+  seq
+
+# 99 ASVs were identified as Mitochondria
+
+aes_data <- 
+  aes_data %>% 
+  filter(Family != 'Mitochondria')
+
 ```
@@
 # Community Analysis
-
-```{r fix-treatment}
-
-aes_data <-
-  aes_data %>% 
-  mutate(passage_treat = treat) %>% 
-  separate(passage_treat, c('passage', "treat"), sep = 2, remove = FALSE) %>% 
-  select(sample, passage_treat, passage, treat, selected, ch4, asv, count, Kingdom:Species, seq)
-
-```
-
-```{r fix-taxonomy}
-
-# There are "Unknown Family" taxa with Genus and Class, but Order is Incertae Sedis.
-# I'm just going to remove everything below Class and call it "Unclassified"
-
-fill_taxonomy <- function(tab) {
-    if(is.na(tab[[1]])) tab[[1]] = "Unclassified"
-    for(i in 2:length(tab)) {
-      if(is.na(tab[[i]])){
-        if(grepl("unclassified", tab[[i-1]])) tab[[i]] = tab[[i-1]]
-        else tab[[i]] = paste(tab[[i-1]], "unclassified", sep = "_")
-      }
-    }
-    tab
-}
-
-aes_data <-
-  aes_data %>% 
-  nest(data = -c(asv, Kingdom:Species)) %>% 
-  nest(taxa = Kingdom:Species) %>% 
-  mutate(taxa = map(taxa, ~fill_taxonomy(.x))) %>% 
-  unnest(taxa) %>% 
-  unnest(data)
-
-```
-
-```{r remove_chloroplasts_mitochondria}
-
-chloroplast_seqs <- 
-  aes_data %>% 
-  filter(Order == 'Chloroplast') %>% 
-  select(asv, seq) %>% 
-  distinct() %$%
-  seq
-# 10 ASVs were identified as chloroplast in their taxonomy.
-# seq matches complete genomes of chloroplasts on NCBI so I will remove them.
-
-aes_data <- 
-  aes_data %>% 
-  filter(Order != 'Chloroplast')
-
-mitochondria_seqs <-
-  aes_data %>% 
-  filter(Family == 'Mitochondria') %>%
-  select(asv, seq) %>% 
-  distinct() %$%
-  seq
-
-# 99 ASVs were identified as Mitochondria
-
-aes_data <- 
-  aes_data %>% 
-  filter(Family != 'Mitochondria')
-
-```
 

### 2023-01-04 manuscript/freewrite.Rmd
revice manuscript

@@
 ^1^ Institute of Ecology and Evolution, University of Oregon, Eugene, OR, USA  
-\* Correspondence: Andrew H. Morris, amorris3\@uoregon.edu  
-Competing Interests: We declare we have no competing interests.
-
-```{r options, echo=FALSE, include=FALSE}
-knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
-#knitr::opts_knit$set(root.dir=normalizePath('../'))
-
-```
-
-# Questions
-
-- What should be the order of the Intro?
-
-    1. What are we interested in and why
-    2. Past approaches and their limitations
-    3. Our approach and why it's better
-    4. Broader conceptual framework (Figure 1, Equation 1)
-
-- Or:
-
-    1. What are we interested in and why
-    2. Past approaches and their limitations
-    3. Broader conceptual framework (Figure 1, Equation 1)
-    4. Our approach and why it's better
-
-- Or:
-
-    1. What are we interested in and why
-    2. Broader conceptual framework (Figure 1, Equation 1)
-    3. Past approaches and their limitations
-    4. Our approach and why it's better
+\* correspondence: Andrew H. Morris, amorris3\@uoregon.edu  
+Competing interests: We declare we have no competing interests.
 
@@
 
-## What is the knowledge gap?
-
-### What are we interested in, why are we interested in it, how has it been and studied and what are the shortfalls, and why precisely is our approach better?
-
 Microbiomes mediate a variety of important ecosystem functions. This suggests
 that understanding variation in microbiome composition might be important for
-predicting the rates of ecosystem functions. Many studies have attempted to
-document such a relationship. However, it is difficult to isolate the effect of
-microbiome composition from other drivers of variation in ecosystem function.
-For example, microbiome manipulations often confound "diversity" with broad
-phylogenetic patterns or "composition" with abundance. Likewise, comparative
-studies must control for abiotic variation statistically to identify the direct
-effect of the microbiome, which relies on measuring the correct abiotic
-variables and cannot account for unmeasured environmental variation. Here, we
-overcome these limitations by using a selection approach to estimate the degree
-to which an ecosystem function varies with microbiome composition. We want to
-understand what is the relationship between _variation in the microbiome_ and
-_variation in the rate of ecosystem function_?
+predicting the rates of ecosystem functions 
+[@conrad1996; @schimel1998; @crowther2019]. 
+Many studies have attempted to document such a relationship
+[@cavigeli2000; @balser2005; @waldrop2006; @strickland2009; wagg2014; wagg2020].
+However, it is difficult to isolate the effect of variation in microbiome
+composition from other drivers of variation in ecosystem function. For example,
+microbiome manipulations often confound variation in diversity, composition, and
+abundance.  Likewise, comparative studies must control for abiotic variation
+statistically to identify the direct effect of microbiome variation. This
+approach relies on measuring the relevant abiotic variables and cannot account
+for unmeasured environmental variation. Here, we overcome these limitations by
+using a selection approach to estimate the degree to which an ecosystem function
+varies with microbiome composition.
 
 There have been two general categories of approaches that investigators have
-applied to address this question: comparative and manipulative. Comparative
+used to estimate the degree to which an ecosystem function varies with
+microbiome composition: comparative and manipulative. Comparative
 studies sample natural variation in an ecosystem function across different
@@
 functions. For example, a meta-analysis of these studies observed a small but
-significant unique contribution of the microbiome to variation in ecosystem
-function after controlling for environmental variation (Graham 2016). In
+significant contribution of the microbiome to variation in ecosystem
+function after controlling for environmental variation [@graham2016]. In
 addition, studies focusing on the correlation between the rate of an ecosystem
@@
 significant correlation, though this relationship is rare and contingent upon
-both the function and the ecosystem sampled (Rocca 2015). Comparative studies
-come with some unique challenges and limitations. One issues is the covariance
-problem where microbiome attributes tend to covary with the abiotic conditions
-within an environment and whether and how the investigator controls for these
-variables affects the conclusions. In addition, it is difficult to know a
-priori which environmental variables or community attributes to measure.
-Finally, while these approaches can establish a potential magnitude and
-direction for these relationships, they fail to establish a mechanism
-underlying the relationship.
+both the function and the ecosystem sampled [@rocca2015]. However, comparative
+studies come with some unique challenges and limitations. One issue is that
+microbiome attributes tend to covary with the abiotic conditions within an
+environment, and whether and how the investigator controls for these variables
+can affect the conclusions of such a comparative study. In addition, it is
+difficult to know a priori which environmental variables or community attributes
+to measure. Finally, while these approaches can establish a potential magnitude
+and direction for these relationships, it is often difficult to identify the taxa or
+genes that explain the connection between composition and function.
 
-The other broad category of approaches to address this question are
+The other broad category of approaches used to address this question are
 manipulative approaches. Manipulative experiments try to alter microbial
-community structure and observe the effect on function. For example,
-reciprocal transplant and common garden experiments have shown that
-microbiomes originating from different ecosystems inoculated onto the same
-substrate or introduced into a common environment display distinct functional
-rates (Cavigeli 2000, Balser 2005, Waldrop 2006, Strickland 2009). In
-addition, manipulating diversity by filtering communities by cell size or
-through dilution alters ecosystem function (Wagg 2014, 2020, dilution
-citation?). However, manipulating the microbiome directly is challenging. Each
-of the manipulative approaches applied confound community structure with other
-factors such as abiotic conditions or phylogenetic relationships. For example,
-reciprocal transplant and common garden experiments confound community
-structure with the abiotic conditions carried on the inoculum (Reed and
-Martiny 2007). In addition, manipulating microbial diversity through filtering
-confounds diversity with broad phylogenetic categories such as nematodes,
-fungi, and bacteria while manipulating diversity through dilution confounds
-diversity with biomass and composition.
+community composition and observe the effect on function. For example,
+reciprocal transplant and common garden experiments have shown that microbiomes
+originating from different ecosystems inoculated into the same substrate or
+introduced into a common environment display distinct functional rates 
+[@cavigeli2000; @balser2005; @waldrop2006; @strickland2009]. In addition,
+manipulating
+diversity by filtering communities by cell size or through dilution has been
+shown to alter the rate of ecosystem functions [@wagg2014; @wagg2020]. However,
+manipulating the microbiome directly is challenging. Each
+of the manipulative approaches applied confound community composition with other
+factors. For example, reciprocal transplant and common garden experiments
+confound community composition with the abiotic conditions carried on the
+inoculum while manipulating composition through dilution
+may confound composition with biomass [@reed2007].
 
 In this study, we sought to build on the observations of comparative and
-manipulative studies by applying a different approach to the question of
-whether variation in the microbiome contributes to variation in the rate of an
-ecosystem function. We used artificial ecosystem selection to select for
-microbiomes that performed a greater rate of ecosysem function. We then tested
-whether variation in the microbiome contributed to variation in the rate of
-ecosystem function and identified microbiome attributes that might explain
-this relationship. There are several potential advantages to this approach for
-documenting BEF relationships and for investigating potential mechanisms
-underlying those relationships. In particular, by passaging microbiomes over
-multiple rounds using a common substrate (or isogenic host), we can weaken the
-covariance between microbes and the environment by repeatedly diluting the
-abiotic effect. In addition, by inoculating our ecosystems with equally sized
-subsamples, we eliminate the need to generate microbiome variation through
-methods that are confounded with diversity, biomass, or cell size. Lastly, by
-comparing our artificially selected community to a control community without
-selection or with random selection, we can both control for changes in the
-environment over time and identify significant genes or taxa associated with our
-function of interest to establish a potential mechanism that explains this
-variation.
+manipulative studies by applying a different approach to the question of whether
+variation in the microbiome contributes to variation in the rate of an ecosystem
+function. We used artificial ecosystem selection to select for microbiomes that
+performed a greater rate of ecosystem function [@swenson2000; @blouin2015; @panke-buisse2015]. We then tested whether variation
+in the microbiome contributed to variation in the rate of ecosystem function and
+identified microbiome attributes that might explain this relationship. There are
+several potential advantages to this approach for documenting
+microbiome-function relationships and for investigating potential mechanisms
+underlying those relationships. By passaging microbiomes over multiple rounds
+using a common environment, we can weaken the covariance between microbes and
+the environment by repeatedly diluting abiotic effects. In addition, by
+inoculating our ecosystems with equally sized subsamples, we eliminate the need
+to generate microbiome variation through methods that are confounded with
+biomass or cell size. Lastly, by comparing our artificially selected community
+to a control community resulting from random selection, we can both control for
+changes in the environment over time and identify significant genes or taxa
+associated with our selection treatment to establish a potential mechanism that
+explains this variation.
 
-One important conceptual distinction that we should make when investigating
-biodiversity-ecosystem function relationships is to be clear that we are
-interested in _variation_. In particular, we are interested in what drives
-variation in the rate of an ecosystem function across ecosystems. For example,
-we want to know whether variation in the rate of methane oxidation rate in soils
-is driven by variation in microbial community structure. This question is
-distinct from the question "Which microbes consume methane?" To illustrate this
-distinction, we can draw an analogy with the genetic basis of traits in
-organisms. If we are interested in whether variation in human height is
-associated with variation in genotype, we can sample a population of humans,
-measure their height, and look at the correlation between phenotype and genotype
-using gene markers or proxies for genotypic similarity, such as kinship.
-However, this does not address the question of "Which genes control human
-height?" Likewise, we could go into the lab and perform an experiment on mice
-where we knockout or over-express a gene thought to be associated with height
-and observe height differences among the treatment groups. This would
-demonstrate that a particular gene affects height, but it does _not_ demonstrate
-that variation in genotype drives variation in height in a population.
-___WHY DOES THIS MATTER?___
+We applied artificial ecosystem selection to soil microbiomes by selecting on
+soil methane oxidation rate. We chose this function because methane is a
+globally important greenhouse gas and methane oxidation by soil bacteria is the
+primary biological sink for atmospheric methane as well as possibly the main driver
+of temporal methane dynamics [@kirschke2013]. In addition, there is evidence that
+soil methane oxidation rate may vary with microbiome composition based on
+comparative studies in a variety of arctic and tropical ecosystems 
+[@meyer2020; @meyer2020a; @freitag2009; @freitag2010]. Finally, methanotrophy is one of
+the most deeply conserved
+microbial physiologies and is represented in a narrow range of taxa, which
+suggests that the taxonomic composition of the microbiome is more
+likely to be associated with the rate of methane oxidation than other broader or
+more shallowly conserved functions [@schimel1998; @martiny2013].
 
-To understand variation in the rate of an ecosystem function, we can think about
-different drivers of variation as illustrated in Figure 1 and Equation 1. In
-this model, variation in the rate of an ecosystem function ($V_F$) is the sum of
-the variation in the environmental conditions ($V_E$), variation in the
-microbiome ($V_M$), and $\epsilon$, which represents variation unexplained by
-the other two factors, technical error, and unmeasured environmental or
-microbial variation. In addition, we can consider another variable,
-$cov(V_E:V_M)$, which is the covariance between the microbiome and the
-environment. We can conceptualize this as a situation in which a particular
-environmental condition, such as pH or moisture content, selects for a particular
-microorganism that is associated with variation in function. For simplicity, we
-have omitted a final parameter, which is the interaction between the microbiome
-and the environment, which would represent a change in the effect of the
-microbiome under different environmental conditions. Here, we assume we're
-talking about a single ecosystem and so assume this parameter to be zero.
+In this study, we evaluated whether variation in the microbiome contributes to
+variation in the rate of ecosystem functions. To address this problem, we
+applied artificial ecosystem selection to methane oxidation rate of the soil
+microbiome and answered the following questions: Does variation in the relative
+abundance of microbial taxa contribute to variation in soil methane oxidation
+rate in the lab? How much of the variation in soil methane oxidation rate can
+be attributed to the microbiome? What attributes of the microbiome regulate
+variation in methane oxidation rate? And do these attributes match our
+assumptions about what regulates methane oxidation rate in nature?
+
+The ultimate goal of this study could be summarized as "do microbes matter to
+ecosystem function?" And there are perhaps three ways we could address that
+question. One would be to ask whether microorganisms mediate the function. This
+question is perhaps trivial. In the case of methane emissions from soil, we know
+that microorganisms produce and consume methane. We could also ask what would
+happen if we removed the microorganisms. Again, this question is trivial. We
+don't know of any abiotic reactions in soil that consume methane (outside of the
+troposphere, the stratosphere, and the marine boundary layer; kirschke).
+Therefore, removing the microbiome would likely eliminate methane production and
+consumption. The third question we could ask is whether *variation* in the
+microbiome is associated with *variation* in the rate of ecosystem function. The
+answer to this question is not so obvious.
 
@@
 Conceptual diagram of the potential relationship between environmental variation, microbiome variation, and variation in the rate of an ecosystem function.
-We omitted the reverse arrows for illustration, but in theory those relationships also exist but we make the assumption that they are small.
-In addition, this diagram represents the relationships for a single environment, but one could imagine the same microbiome exerting a different influence on variation in function in a second environment, which would be illustrated by the interaction between microbiome and environment ($V_E:V_M$)`](Figure1.png)
+Without experimentally or statistically controlling for the underlying environmental variation, it can be difficult to distinguish between the direct effect of the microbiome and the effect of the microbiome due to covariance with the environment.](Figure1.png)
 
-$$
-V_F = V_E + V_M + COV(V_E:V_M) + \epsilon
-$$
+If we look at variation in the rate of a microbially-mediated ecosystem function
+across a set of habitats, that variation could be associated with variation in
+the abiotic environmental conditions, variation in the composition of
+microbiomes, or some interaction or covariance between the two (see Figure 1). To address the
+question of whether variation in the microbiome matters for ecosystem function,
+we need to identify the component of variation uniquely associated with the
+microbiome. If we sample natural variation, as in the comparative approach, it
+can be difficult to distinguish between the direct effect of the microbiome and
+the effect of the microbiome due to covariance with the environment. Likewise,
+with the manipulative approach even though we are placing ecosystem propagules
+into a common environment or onto a common substrate, the propagule itself
+carries with it environmental conditions that cannot be removed. Generating
+variation in microbiome composition through filtering or other similar means
+often confounds composition with other factors, for example, by extinguishing
+all fungi while sparing the bacteria, in which case it is uncertain whether this
+represents the effect of composition or simply the gross effects of mass
+extinction.
 
-How can we use this conceptual framework to address our original question: Does
-variation in the microbiome contribute to variation in the rate of ecosystem
-function? Under the null hypothesis that we can ignore the microbiome and
-simply use abiotic environmental variables to predict function, we would assume
-that $V_M$ is essentially zero and that all of the variation in function
-($V_F$) is due to variation in the environment and the covariance between the
-microbiome and the environment ($V_E + COV(V_E:V_M)$). In this case, it is
-unnecessary to understand anything about the microbiome. However, if we consider
-the results from Graham (2016), this hypothesis does not appear to match the
-data. Graham observed that across the studies that they analyzed, environmental
-variables explained an average of 56\% of the variation in ecosystem function
-while microbial community structure explained an average of 31\% of the
-variation. Models that included both explained an average of 65\% of the
-variation. Putting these numbers together, we see that $V_M$ = 11\% (65 - 56),
-$cov(V_E:V_M)$ = 20\% (31 - $V_M$ = 31 - 11), and $V_E$ = 36\% (56 -
-$cov(V_E:V_M)$ = 56 - 20). Therefore, the environment explains 36% of the
-variation directly
-and 20% via covariation with the microbiome and the microbiome explains 11%
-directly. This means at least 1/6 (11/65) of the explained variation can be
-uniquely attributed to the microbiome according to the conclusions of Graham.
-
-Of course, these numbers should be taken with a grain of salt. As discussed
-earlier, comparative studies such as these are unable to break the covariance
-between microbes and the environment. In this case, $V_M$ and $\epsilon$ may
-represent the direct or indirect effect (via covariance) of unmeasured
-environmental variables rather than a unique contribution of the microbiome. In
-addition, the proportion of variation depends upon the magnitude of the
-variation in $V_F$ and $V_E$. For example, if we were to include variation in
-methane emissions of wetland and upland ecosystems, we might find that virtually
-100\% of the variation in function was due to variation in the environment,
-since methane production requires anaerobic conditions. However, if we looked
-across wetlands, we might observe a smaller $V_F$ and $V_E$ and an equivalent
-$V_M$. Under this scenario, even though $V_M$ was the same, $\frac{V_M}{V_F}$
-would be much greater and therefore the environment would no longer explain
-100\% of the variation. Therefore, any study that wishes to address this
-question must consider the variation in the population that they are studying.
-
-We performed artificial ecosystem selection on soil microbiomes by selecting on
-the methane oxidation rate. To perform this experiment, we created 24 jars in an
-identical manner using sterile potting mix inoculated with living soil. Each jar
-was randomly assigned to either a positive selection treatment that underwent
-selection on high soil methane oxidation rate or a random selection treatment.
-At the end of each incubation, we selected jars with a high methane oxidation
-rate (for the positive treatment), homogenized them, and used this homogenate to
-inoculate a new set of sterile soil jars. We repeated this process over multiple
-passges. The idea is that by shuffling microbiome composition through
-subsampling and homogenizing the environment we are exploring the parameter
-space for the bidoversity-ecosystem function relationship. In addition,
-subsampling at each step would dilute away the effect of the abiotic
-environment. The only way for there to be a response to selection would be for
-microbes to actually grow in the medium.
-
-Here are some hypothetical results we could observe. If soil methane oxidation
-rate is entirely environmentally determined, i.e., in the parlance of our
-conceptual framework that would mean that $V_E$ and $cov(V_E:V_M)$ are large and
-that $V_M$ is relatively small, then we would expect to observe no response to
-selection. This is because the effect of the abiotic environment would be
-diluted at eaching passaging step. If variation in methane oxidation rate was
-independent of microbes and the environment in our system then the varation
-should be random and $\epsilon$ would be large relative to the other parameters
-and again we would observe no response to selection. If variation in methane
-oxidation rate is at least partially driven by variation in the microbiome then
-we would expect manipulating microbiome composition through selection would
-result in an increase in methane oxidation rate as a response to selection. This
-would manifest as a correlation between donor and recipient jars in consecutive
-passages.
-
-(Aside: What we seem to be saying in the above paragraph is that passaging
-ecosystem propagules is somehow intrinsically tied to the magnitude of Vm. This
-seems to be based on the argument that Vm reproduces itself with high fidelity,
-but Ve does not. Is this true? Adding a low pH 5g of soil to a neutral pH 45g of
-soil doesn't shift the pH, but adding a 5g soil microbial inoculum to 45g of
-autoclaved soil does shift the community? Is this any different than the
-manipulative experiments we describe above? Or perhaps the argument is that
-_repeated_ passaging dilutes away the environmental effect so by the time you
-get to passage 5 it's 1/1e5 of the original soil inoculum, but if the microbes
-have reproduced then they have rebuilt their population at each passage. I guess
-this is the idea behind serial batch culturing??? IDK if a reviewer will be
-convinced by this argument.)
+Using artificial ecosystem selection, we can generate a set of ecosystems in
+bottles using a common, sterile substrate and inoculate them with a propagule
+containing a diverse microbiome that is a subsample of an existing microbiome.
+This generates variation in microbiome composition among the ecosystems. We can
+then measure the rate of an ecosystem function for each bottle and select the
+bottles with the highest (or lowest) rate of ecosystem function. The selected
+ecosystems are then homogenized and used as the propagule to inoculate a new set of
+ecosystems containing the same sterile substrate as the initial set. By
+passaging the microbiome in this way over multiple generations we can observe
+whether varying the microbiome can generate variation in the rate of our
+ecosystem function. If we observe no change in ecosystem function from
+generation to generation, we would conclude that this ecosystem function in this
+set of ecosystems is driven primarily by environmental variation or technical
+error among the replicates. However, if we observe a response to selection, that
+is, a shift in the mean ecosystem function between successive passages, we would
+conclude variation in the microbiome can generate variation in the rate of
+ecosystem function.

### 2023-01-04 manuscript/manuscript.Rmd
revice manuscript

@@
 responded to selection included members of `r length(da_fams)` families not
-known to oxidize methane, such as
-*Fimbriimonadaceae*, *Cytophagaceae*, and *Diplorickettsiaceae*. This result is
-in contrast to the typical assumption that the rate of an ecosystem function is
-limited by the final step in the associated microbial pathway. Our study
-demonstrates that variation in microbiome composition can contribute to
-variation in the rate of ecosystem function and that manipulating microbiome composition
-could be a viable strategy for managing ecosystem functions.
+known to oxidize methane, such as *Fimbriimonadaceae*, *Cytophagaceae*, and
+*Diplorickettsiaceae*. This result is in contrast to the typical assumption that
+the rate of an ecosystem function is limited by the final step in the associated
+microbial pathway. Our study demonstrates that variation in microbiome
+composition can contribute to variation in the rate of ecosystem function and
+that manipulating microbiome composition could be a viable strategy for managing
+ecosystem functions.
 
@@
 
-Microbiomes mediate a variety of important functions in ecosystems, and there
-is great interest in understanding how attributes of microbiomes may influence
-variation in ecosystem functions [@conrad1996; @schimel1998; @crowther2019].
-Biodiversity-ecosystem function relationships have been documented for a
-variety of macroorganismal communities. For example, plant species richness
-correlates with productivity and marine community diversity correlates with
-production, consumption, and nutrient release 
-[@cardinale2006; @hooper2012; @gamfeldt2015]. However, we have been unsuccessful
-in documenting similar relationships for microbial communities. While there is
-some evidence that microbiomes from different ecosystems display different
-functional rates [@balser2005; @strickland2009], adding microbiome attributes to
-ecosystem models adds very little to total variance explained [@graham2016].
-Therefore, we still lack a detailed understanding of what
-functions are altered by microbiome composition and what attributes of the
-microbiome explain this relationship.
+Microbiomes mediate a variety of important ecosystem functions. This suggests
+that understanding variation in microbiome composition might be important for
+predicting the rates of ecosystem functions 
+[@conrad1996; @schimel1998; @crowther2019]. 
+Many studies have attempted to document such a relationship
+[@cavigelli2000; @balser2005; @waldrop2006; @strickland2009; @wagg2014; @wagg2019].
+However, it is difficult to isolate the effect of variation in microbiome
+composition from other drivers of variation in ecosystem function. For example,
+microbiome manipulations often confound variation in diversity, composition, and
+abundance. Likewise, comparative studies must control for abiotic variation
+statistically to identify the direct effect of microbiome variation. This
+approach relies on measuring the relevant abiotic variables and cannot account
+for unmeasured environmental variation. Here, we overcome these limitations by
+using a selection approach to estimate the degree to which an ecosystem function
+varies with microbiome composition.
 
-There are several reasons for this difficulty. One reason is that it is
-challenging to manipulate microbiome composition and diversity directly.
-Experimental manipulations of community composition and diversity have been
-central to documenting the relationships between ecosystem functions and the
-biodiversity of plants and animals [@hooper2012], but this is seldom possible
-with microbiomes. Most microorganisms cannot yet be cultured and most available
-methods for manipulation of microbiomes (e.g., antibiotics, dilution, etc.) are
-not sufficiently specific to manipulate individual microbial lineages. In
-addition, many studies have looked for relationships between an ecosystem
-function and the abundance or diversity of a microbial gene that codes for an
-enzyme or enzyme-subunit known to be involved in that function. However, it is
-difficult to identify the “right” marker gene a priori and for most functions
-in most ecosystems there is no correlation between the abundance of a marker
-gene and the rate of the corresponding process [@rocca2015]. Lastly,
-microbiome-ecosystem function relationships have been difficult to document
-because microbiomes often covary with other drivers of variation in ecosystem
-functions, such as environmental conditions or resources. Such covariation can
-make it impossible to determine whether microbiomes directly contribute to
-functional variation. Understanding the degree to which microbiomes directly
-influence functional variation is crucial if scientists are to develop
-microbiome-focused methods for altering or maintaining ecosystem functions.
+There have been two general categories of approaches that investigators have
+used to estimate the degree to which an ecosystem function varies with
+microbiome composition: comparative and manipulative. Comparative
+studies sample natural variation in an ecosystem function across different
+habitats and simultaneously measure variation in community composition.
+Investigators can then correlate ecosystem function with aspects of community
+composition while attempting to control for environmental variation. These
+approaches have
+documented important relationships between microbiomes and ecosystem
+functions. For example, a meta-analysis of these studies observed a small but
+significant contribution of the microbiome to variation in ecosystem
+function after controlling for environmental variation [@graham2016]. In
+addition, studies focusing on the correlation between the rate of an ecosystem
+function and the abundance of an associated functional marker gene observe a
+significant correlation, though this relationship is rare and contingent upon
+both the function and the ecosystem sampled [@rocca2015]. However, comparative
+studies come with some unique challenges and limitations. One issue is that
+microbiome attributes tend to covary with the abiotic conditions within an
+environment, and whether and how the investigator controls for these variables
+can affect the conclusions of such a comparative study. In addition, it is
+difficult to know a priori which environmental variables or community attributes
+to measure. Finally, while these approaches can establish a potential magnitude
+and direction for these relationships, it is often difficult to identify the
+taxa or genes that explain the connection between composition and function.
 
-One way to overcome these challenges is to manipulate the relationship between
-microbiome biodiversity and ecosystem function using artificial ecosystem selection 
-[@swenson2000; @blouin2015; @panke-buisse2015; @sanchez2021]. 
-Microbiomes have been shown to respond to selection at the ecosystem-trait
-level [@swenson2000], and some ecosystem-scale microbiome traits have been
-shown to be transferrable through microbiome inoculations [@panke-buisse2015].
-Therefore, applying artificial selection to whole microbiomes could be a novel
-approach for describing a more detailed microbiome-ecosystem function
-relationship. This approach works by subsampling a microbiome to inoculate
-multiple replicate ecosystems that vary in microbiome composition and then
-measuring the rate of ecosystem function for each replicate. The investigator
-then selects the replicates with the highest (or lowest) rates and subsamples
-those ecosystems to passage them to a new set of ecosystems. By holding
-environmental conditions relatively constant among the replicates and from
-passage to passage (or by using isogenic hosts, in the case of host-associated
-microbiomes), investigators can determine whether variation in the microbiome
-contributes to variation in the rate of ecosystem function independent of the
-environment.
+The other broad category of approaches used to address this question are
+manipulative approaches. Manipulative experiments try to alter microbial
+community composition and observe the effect on function. For example,
+reciprocal transplant and common garden experiments have shown that microbiomes
+originating from different ecosystems inoculated into the same substrate or
+introduced into a common environment display distinct functional rates 
+[@cavigelli2000; @balser2005; @waldrop2006; @strickland2009]. In addition,
+manipulating diversity by filtering communities by cell size or through dilution
+has been shown to alter the rate of ecosystem functions 
+[@wagg2014; @wagg2019; @griffiths2001]. However, manipulating the microbiome
+directly is challenging. Each of the manipulative approaches applied confound
+community composition with other factors. For example, reciprocal transplant and
+common garden experiments confound community composition with the abiotic
+conditions carried on the inoculum while manipulating composition through
+dilution may confound composition with biomass [@reed2007].
 
-A valuable feature of this approach is that it can be used to estimate how much
-of the variation in ecosystem function is due to microbiome variation.  By
-calculating the slope of the regression between the "donor" ecosystems selected
-in one passage and the "recipient" ecosystems inoculated by those donors in the
-next passage, we can quantify an estimate of "ecosystem heritability"
-[@goodnight2000; @blouin2015]. The concept of heritability has been borrowed
-from quantitative genetics where heritability is defined as *the ratio of
-additive genetic variance to the total phenotypic variance* and is commonly
-estimated as *the slope of the linear regression between offspring phenotype
-and parental phenotype* [@bell1977; @falconer1996]. Here we apply this concept
-to donor and recipient communities in an artificial ecosystem selection
-experiment to estimate *the variance in ecosystem function due to microbiome
-variance*. We do not assume that environmental variation is zero, instead we
-quantify the proportion of variation due to microbiome effects. If the slope of
-the regression is zero, we would conclude that environmental variation among
-the replicates and technical error are the primary drivers of variation in
-ecosystem function. On the other hand, a significant slope between donors and
-recipients would demonstrate that microbiome variation generates variation in
-the rate of an ecosystem function. We then use our artificially selected
-microbiomes to investigate what aspects of microbiome variation explain this
-response. By "microbiome variation" we mean aspects of microbial diversity
-commonly measured by microbial ecologists, such as the relative abundance of
-taxonomic groups inferred from amplicon sequence variants. With this approach,
-we can not only determine whether an ecosystem function is influenced by
-microbiome variation, but also estimate how much of the variation and which
-taxa are associated with higher rates of ecosystem function.
+(Side note: papers like the Strickland paper were actually asking a subtly
+different question from the one that we are asking. Strickland was asking
+whether communites can be functionally dissimilar as opposed to functionally
+redundant. That is, that historical contingency plays a role in the functioning
+of microbial communities. They demonstrated quite well that communities
+originating from different environments could have unique functional rates.
+However, is that due to microbiome-environment covariance where a different
+environment (i.e., a different litter substrate) selects for a litter-degrading
+community or is that a result of the unique effect of variation in the
+community? They actually point this out in their Discussion, they suggest future
+research should determine whether the differences between communities are driven
+by taxonomic composition or is function the result of the selection environment
+(I think of this is analogous to our "covariation" argument)? They also suggest
+trying to adapt communities to a new environment through continuous inputs of
+the same resource (just like in our experiment!))
 
-We applied artificial ecosystem selection to soil microbial communities by
-selecting on soil methane oxidation rate. We chose this function because
-methane is a globally important greenhouse gas and methane oxidation by soil
-bacteria is the primary biological sink for atmospheric methane
-[@kirschke2013]. In addition, there is evidence that soil methane oxidation
-rate may vary with microbiome composition. For example, soil microbiome
-phylogenetic variation is a strong predictor of methane oxidation in forests
-and pastures of the Brazilian Amazon as well as soils of the Congo Basin
-[@meyer2020; @meyer2020a]. Finally, methanotrophy is one of the most deeply
-conserved microbial physiologies and is represented in a narrow range of taxa
-and so the taxonomic composition of the microbiome is more likely to be
-associated with the rate of methane oxidation than other broader or more
-shallowly conserved functions [@schimel1998; @martiny2013]. This suggests that
-microbiome variation might drive variation in methane oxidation in soils. One
-other study has estimated ecosystem heritability for an environmental
-microbiome. Blouin et al. [@blouin2015] selected for water microbiomes with
-low carbon dioxide respiration rates and estimated a heritability of 52%.
-However, this appeared to be driven primarily by biomass, not microbiome
-composition. Here we want to build on this work by selecting for higher rates
-of a more narrow function that is less likely to be driven by total biomass.
-That way we can begin to assess whether variation in microbiome composition is
-important for understanding variation in ecosystem functions.
+In this study, we sought to build on the observations of comparative and
+manipulative studies by applying a different approach to the question of whether
+variation in the microbiome contributes to variation in the rate of an ecosystem
+function. We used artificial ecosystem selection to select for microbiomes that
+performed a greater rate of ecosystem function
+[@swenson2000; @blouin2015; @panke-buisse2015]. We then tested whether variation
+in the microbiome contributed to variation in the rate of ecosystem function and
+identified microbiome attributes that might explain this relationship. There are
+several potential advantages to this approach for documenting
+microbiome-function relationships and for investigating potential mechanisms
+underlying those relationships. By passaging microbiomes over multiple rounds
+using a common environment, we can weaken the covariance between microbes and
+the environment by repeatedly diluting abiotic effects. In addition, by
+inoculating our ecosystems with equally sized subsamples, we eliminate the need
+to generate microbiome variation through methods that are confounded with
+biomass or cell size. Lastly, by comparing our artificially selected community
+to a control community resulting from random selection, we can both control for
+changes in the environment over time and identify significant genes or taxa
+associated with our selection treatment to establish a potential mechanism that
+explains this variation.
+
+We applied artificial ecosystem selection to soil microbiomes by selecting on
+soil methane oxidation rate. We chose this function because methane is a
+globally important greenhouse gas and methane oxidation by soil bacteria is the
+primary biological sink for atmospheric methane as well as possibly the main driver
+of temporal methane dynamics [@kirschke2013]. In addition, there is evidence that
+soil methane oxidation rate may vary with microbiome composition based on
+comparative studies in a variety of arctic and tropical ecosystems 
+[@meyer2020; @meyer2020a; @freitag2009; @freitag2010] as well as studies that
+manipulate methanotroph richness [@schnyder2018]. Finally, methanotrophy is one
+of the most deeply conserved microbial physiologies and is represented in a
+narrow range of taxa, which suggests that the taxonomic composition of the
+microbiome is more likely to be associated with the rate of methane oxidation
+than other broader or more shallowly conserved functions
+[@schimel1998; @martiny2013].
 
@@
 
-We estimated the ecosystem heritability of methane oxidation rate, which is the
-proportion of variance in CH~4~ oxidation rate due to variation in the
-microbiome, as the regression of mid-recipient on mid-donor [@falconer1996].
-The mid-donor was the mean for all three selected jars and the mid-recipient was
-the mean methane oxidation rate for the twelve jars produced by those three
-donors. First, we tested if there was a difference of slopes between the
-positive and neutral treatments. To test this, we fit two nested models with and
-without the treatment term and compared them using an F-test with the 'anova'
-function. We then present the slopes for each treatment, which represented the
-ecosystem heritability or the proprotion of total variance in methane oxidation
-rate due to microbiome variance.
-
-First, we tested if there was an effect of treatment by comparing
-nested models with and without treatment using an F-test with the 'anova'
-function. We then fit the full model to estimate the slope of the relationship
-between recipient and donor for the Positive and Neutral treatments.
+We estimated the proportion of variation in methane oxidation rate due to
+variation in the microbiome as the regression of divergence between the positive
+and neutral treatment on the cumulative selection differential [@falconer1996].
+Divergence was calculated as the mean methane oxidation rate of the positive
+treatment minus the mean methane oxidation rate of the neutral treatment in each
+passage. The selection differential was calculated as the difference between
+the mean of the three selected jars and the mean of all twelve jars in a passage.
+Cumulative selection differential was then the sum of the selection differential
+from all preceding selection events. We test this as the regression of
+divergence on cumulative selection differential using the 'lm' function.
 
@@
 rate (`r figures("response", display = "cite")`; difference of slopes:
-F~`r compare_response_mods[2,4]`,`r compare_response_mods[2,2]`~ = `r myround(compare_response_mods[2,5], 2)`, p = `r myround(compare_response_mods[2,6], 2)`).
-At the start of the experiment, the Positive
-treatment had a mean CH~4~ oxidation rate that was
-`r percent_less(initial = tidy(response_model)[[1, 2]], final = tidy(response_model)[[3, 2]] + tidy(response_model)[[1, 2]])`\% lower than the
-Neutral treatment `r print_lm(response_model, 3, 'difference of y-intercepts')`. 
-There was no change in CH~4~ oxidation rate of the Neutral
-treatment over the five passages `r print_lm(response_model, 2)`. By contrast,
-the Positive treatment had a `r myround(per_slope, 1)`\% increase in CH~4~
-oxidation rate per passage `r print_lm(response_model, 4)`.
+F~`r compare_response_mods[2,4]`,`r compare_response_mods[2,2]`~ = 
+`r myround(compare_response_mods[2,5], 2)`,
+p = `r myround(compare_response_mods[2,6], 2)`). At the start of the experiment,
+the Positive treatment had a mean CH~4~ oxidation rate that was
+`r percent_less(initial = tidy(response_model)[[1, 2]], final = tidy(response_model)[[3, 2]] + tidy(response_model)[[1, 2]])`\%
+lower than the Neutral treatment
+`r print_lm(response_model, 3, 'difference of y-intercepts')`. 
+There was no change in CH~4~ oxidation rate of the Neutral treatment over the
+five passages `r print_lm(response_model, 2)`. By contrast, the Positive
+treatment had a `r myround(per_slope, 1)`\% increase in CH~4~ oxidation rate per
+passage `r print_lm(response_model, 4)`.
 
-We estimated the ecosystem heritability defined as the 
-proportion of variation in methane oxidation due rate to variation
-in the microbiome as the regression of mid-recipient on mid-donor 
-(`r figures("herit", display = "cite")`).  Recipient CH~4~ oxidation rates were
-correlated with donor CH~4~ oxidation rates in both the Positive treatment
+We estimated the proportion of variation in methane oxidation rate due to
+variation in the microbiome by regressing the divergence between the positive
+and neutral treatments against the cumulative selection differential. We
+observed a slope of 0.40, which would indicate that 40\% of the variation is due
+to the microbiome, but this was not significant (statistics).
+
+We estimated the ecosystem heritability defined as the proportion of variation
+in methane oxidation due rate to variation in the microbiome as the regression
+of mid-recipient on mid-donor (`r figures("herit", display = "cite")`).
+Recipient CH~4~ oxidation rates were correlated with donor CH~4~ oxidation rates
+in both the Positive treatment
 `r paste0('(', 'slope', ' = ', myround(tidy(herit_model)[2, 2] + tidy(herit_model)[4, 2], 2), ', SE = ', myround(tidy(herit_model)[4, 3], 2), ', t = ', myround(tidy(herit_model)[4, 4], 2), ', p = ', myround(tidy(herit_model)[4, 5], 2), ')')` and
@@
 
-While it is well understood that the microbiome mediates a variety of ecosystem
-functions, a fundamental question in microbial ecology is whether we need to
-understand variation in the composition of microbiomes to understand variation
-in the rate of ecosystem functions [@conrad1996; @schimel1998]. To address this
-question, we performed artificial ecosystem selection on soil microbiomes by
-selecting for ecosystems with high methane oxidation rates. We found that there
-was a significant increase in methane oxidation rate in the Positive selection
-treatment relative to the Neutral selection treatment. This demonstrates that
-passaging the microbiome through artificial ecosystem selection resulted in a
-shift in the mean methane oxidation rate of the soil microbiome. 
-This suggests that variation in the microbiome can generate variation in
-ecosystem functions independent of the environment.
+In this study, we addressed whether variation in the microbiome contributes to
+variation in the rate of ecosystem function. To test this, we applied artificial
+ecosystem selection to soil microbiomes by selecting on the whole-ecosystem soil
+methane oxidation rate. We observed an increase in methane oxidation rate in the
+selection treatment relative to the control, which demonstrates that there was a
+response to selection. Given that we observed a response to selection, we
+conclude that variation in the microbiome contributes to variation in the
+methane oxidation rate independent of the environment.
 
-We quantified how much of the variation in methane oxidation rate was due to
-variation in the microbiome by regressing the mean methane oxidation rate of
-the 12 recipient jars against the mean methane oxidation rate of the three
-donor jars selected to inoculate the recipients. We observed a heritability of
-`r myround((coefficients(herit_model)[2] + coefficients(herit_model)[4])*100,1)`\% 
-for the Positive selection treatment. This indicates that 
-`r myround((coefficients(herit_model)[2] + coefficients(herit_model)[4])*100,1)`\% 
-of the variation in methane oxidation rate can be attributed to variation in the
-microbiome. This is comparable to the ecosystem heritability of carbon dioxide
-emissions from water [@blouin2015]. Notably, we observed a negative heritability
-in the Neutral treatment (`r myround(coefficients(herit_model)[2]*100, 1)`\%).
-Negative heritability estimates are not uncommon in quantitative genetics
-[@steinsaltz2020]. This can occur for random selection controls in artificial
-selection experiments and may indicate that methane oxidation rate is negatively
-correlated with other traits associated with persistence in jars. The
-heritability we observed in the Positive treatment indicates that a large
-portion of the variation in methane oxidation rate is due to variation in the
-microbiome in our experimental system.
+The ultimate goal of this study (and much of microbial ecology) could be
+summarized as "Do microbes matter to ecosystem function (or health)?" and there
+are at least three ways we could address that question. One would be to ask
+whether microorganisms mediate function. This question is perhaps trivial. In
+the case of methane emissions from soil, we know that microorganisms produce and
+consume methane. We could also ask what would happen if we removed the
+microorganisms. Again, this question is trivial. There are no known abiotic
+reactions in soil that consume methane (outside of the troposphere, the
+stratosphere, and the marine boundary layer @kirschke2013). Therefore, removing
+the microbiome would likely eliminate methane production and consumption. Then,
+the third question we could ask is whether *variation* in the microbiome is
+associated with *variation* in the rate of ecosystem function. The answer to
+this question is not so obvious.
+
+If we look at variation in the rate of a microbially-mediated ecosystem function
+across a set of habitats, that variation could be associated with variation in
+the abiotic environmental conditions, variation in the composition of
+microbiomes, or some interaction or covariance between the two (see Figure 5).
+To address the question of whether variation in the microbiome matters for
+ecosystem function, we need to identify the component of variation uniquely
+associated with the microbiome. If we sample natural variation, as in the
+comparative approach, it can be difficult to distinguish between the direct
+effect of the microbiome and the effect of the microbiome due to covariance with
+the environment. Likewise, with the manipulative approach even though we are
+placing ecosystem propagules into a common environment or substrate, the
+propagule itself carries with it environmental conditions that
+cannot be removed. Generating variation in microbiome composition through
+filtering or other similar means often confounds composition with other factors,
+such as biomass or cell size. To overcome these limitations and to isolate
+the direct effect of the microbiome on ecosystem function, we used
+artificial ecosystem selection.
+
+To perform artificial ecosystem selection, we generate a set of ecosystems in
+bottles using a common, sterile substrate and inoculate them with a propagule
+containing a diverse microbiome that is a subsample of an existing microbiome.
+This generates variation in microbiome composition among the ecosystems. We can
+then measure the rate of an ecosystem function for each bottle and select the
+bottles with the highest (or lowest) rate of ecosystem function. The selected
+ecosystems are then homogenized and used as the propagule to inoculate a new set of
+ecosystems containing the same sterile substrate as the initial set. By
+subsampling the microbiome, we can generate variation in the composition of the
+microbiome that is independent of cell size or biomass. And by passaging the
+microbiome through the same sterile substrate over multiple rounds of selection,
+we can progressively minimize the effect of the abiotic conditions from the
+donor environment on both the rate of function and the composition of the
+microbiome. In this way, over multiple rounds of selection, we can observe
+whether varying the microbiome can generate variation in the rate of our
+ecosystem function. If we observe no change in ecosystem function from
+passage to passage despite our imposed selection, we would conclude that this
+ecosystem function in this set of ecosystems is driven primarily by
+environmental variation or technical error among the replicates. However, if we
+observe a response to selection, that is, a shift in the mean ecosystem function
+between successive passages, we would conclude that variation in the microbiome can
+generate variation in the rate of ecosystem function. And that is, in fact, what
+we observed.
+
+Given that variation in the microbiome is associated with variation in ecosystem
+function, a reasonable follow-up question is "How much variation?" This is a
+challenging question to answer since the value depends on the population of
+ecosystems that is sampled and the estimate for one set of ecosystems may not
+apply to another set. For example, we cannot estimate a percent of variation
+explained in tropical rainforests and then assume that number applies in polar
+wetlands. In addition, the magnitude of the estimate will depend on how much
+variation is sampled. A large $\frac{V_M}{V_F}$ is not very interesting if $V_F$
+is small.  Keeping these caveats in mind, one way to estimate this number is to
+determine how much the recipient jars resemble the selected donor jars that were
+used to inoculate them. We can calculate the response to selection as the
+difference between two successive passages in their mean methane oxidation rate.
+We will denote this as $R$. We can also calculate the strength of selection as
+the difference in mean methane oxidation rate between the twelve jars in one
+generation and the 3 jars chosen for selection in that generation, which we will
+call the selection differential and denote as $S$. If we plot the cumulative
+$R$ against the cumulative $S$, the slope of this relationship will equal the
+proportion of variation explained by the microbiome.
+
+Using this approach, we observed a relatively large percent of variation
+explained by the microbiome, but this number was not significantly different
+from zero.
+
+The concept that we have just described is analogous to the concept of
+heritability from quantitative genetics. Heritability for given a population is
+defined as the proportion of variation in a phenotype associated with variation
+in genotype [@lynch1998]. This concept has been applied above the population
+level for groups within a species [@goodnight2000], as well as communities and
+ecosystems [@whitham2003; @whitham2006]. One study previously estimated the
+heritability of a community trait after artificial ecosystem selection using
+parent-offspring regression and observed a relatively large heritability of 50\%
+[@blouin2015]. However, parent-offspring regression assumes a randomly mating
+population, which, by definition, does not apply to a selection experiment. In
+the present study, we estimated the variation explained by the microbiome
+somewhat differently using the breeder's equation ($R = S h^2$) and regressing
+the observed response to selection against the imposed selection differential
+(Figure 1; @falconer1996). We also observed a relatively large response to selection
+(40\%), though this estimate had large error bars and was not significant so we
+did not have the power to estimate an accurate percent of variation explained
+for this experiment.
+
+Asking how much of the variation in ecosystem function is due to variation in
+the microbiome is a silly question. We cannot answer this questions broadly
+for all communities or all ecosystems, because the answer depends in large part
+on the magnitude of variation sampled.
+
+What do I want to get across here:
+
+This estimate is analogous to the concept of heritability from quantitative
+genetics
+Define heritability
+Point out that it's been applied to communities and ecosystems
+Mention the different ways of estimating heritability
+Mention the assumptions inherent in some of these estimates (non-assortative
+mating; particularly numbers of shared alleles based on kinship)
+Point out that while heritability is a useful framework to borrow from in order
+to describe the components of variation, there are limitations in its
+application
+Propose that we pursue new theory about how to understand what drives variation
+in microbiome functions and microbiome-encoded traits
+
+The challenge with applying the concept of heritability outside its original
+scope is that it is unclear if the assumptions underlying these estimates apply
+to microbiomes. For a genomic trait, we can rely on the assumption that
+offspring share, on average, 50\% of the alleles of each of their
+parents. It is less clear if offspring share 50\% of the microbes of each of
+their parents or if it is sufficient for 50\% of the microbes to be _present_ or
+if they must be in a similar relative abundance. Estimates of heritability, such
+as parent-offspring regression and the ACE model rely on specific estimates of
+the similarity of relatives based on their relatedness. It is unclear if these
+assumptions apply to the microbiomes of host ecosystems. It is even less clear
+whether they apply to non-host ecosystems that do not reproduce through sexual
+recombination.
 
@@
 Taxa are sorted by their effect size with taxa at the top having the largest positive effect size and taxa at the bottom with the largest negative effect size.")`
+
+![Figure 5.
+Variation in the rate of an ecosystem function is potentially the result of at least three components: variation in the abiotic environmental conditions, variation in microbiome composition, and the covariance between microbiomes and the environment.
+It is difficult with traditional study designs to isolate the direct effect of the microbiome from the effect of the environment via covariance with the microbiome.
+Here, we attempt to isolate the effect of the microbiome through artificial selection on microbiome composition.
+For simplicity, we omitted the reverse arrows as well as the interaction between microbiome and the environment, though these relationshps may also exist.](Figure1.png)

### 2023-01-04 analysis/new_analysis.R

@@ -86,5 +86,79 @@ ggplot(rh,
 
 model3 <- lm(divergence ~ 1, rh)
 model4 <- lm(divergence ~ cumulative_sd, rh)
+fitdata <- fitted(model3)
+fitdata4 <- fitted(model4)
+ggplot(rh, 
+       aes(x = cumulative_sd, 
+           y = divergence)) +
+    stat_smooth(method = 'lm', 
+                se = TRUE, 
+                formula = 'y ~ x',
+                color = "red",
+                linetype = 2,
+                linewidth = 0.5) +
+    geom_point() + 
+    labs(x = "Cumulative Selection Differential", 
+         y = expression(atop("Divergence", log[10] ~ CH[4] ~ "(-k)"))) +
+    coord_fixed(ratio_rh) +
+    geom_line(aes(y = fitdata), linetype = 2) +
+    geom_abline(intercept = 0, slope = 1, linetype = 3) +
+    geom_abline(intercept = 0, slope = 0, linetype = 3) +
+    
+    xlim(-1, 2) +
+    ylim(-1, 2)
+
 anova(model3, model4)
 summary(model4)
+
+
+############################
+
+library(glmmTMB)
+aes_data %>% 
+    filter(!is.na(ch4)) ->
+    aes_data_nona
+aes_data_nona %>% 
+    select(asv, sample, count) %>% 
+    pivot_wider(names_from = asv, values_from = count, values_fill = 0) %>% 
+    as.data.frame() ->
+    aes_com_df
+rownames(aes_com_df) <- aes_com_df[, 1]
+aes_com_df <- aes_com_df[, -1]
+rowSums(aes_com_df) %>% 
+    min() ->
+    min_seqs
+beta_dis <- avgdist(aes_com_df, iterations = 1, sample = min_seqs)
+beta_dis %>% 
+    as.matrix() ->
+    beta_dis_mat
+beta_dis_mat %>% 
+    as_tibble(rownames = 'sample1') %>% 
+    pivot_longer(-sample1, names_to = 'sample2', values_to = 'beta') %>% 
+    mutate(samples = paste(sample1, sample2, sep = "-")) ->
+    beta_df
+aes_data_nona %>% 
+    select(sample, passage, treat, ch4) %>% 
+    distinct() ->
+    model_df
+
+model_df %>% 
+    select(sample, ch4) %>% 
+    as.data.frame() ->
+    ch4_mat
+rownames(ch4_mat) <- ch4_mat[, 1]
[…]

### 2023-01-04 analysis/scripts/05-combine-asvs-and-fluxes.R

@@ -1,8 +1,6 @@
 
 library(tidyverse)
-#library(phyloseq)
 library(decontam)
-#library(microViz)
 library(knitr)
 library(vegan)
 source('R/functions.R')

### 2023-01-04 manuscript/render.R

@@ -1,0 +1,1 @@
+rmarkdown::render('manuscript/manuscript.Rmd')

### 2023-01-04 my_render.sh

@@ -1,0 +1,2 @@
+#!/bin/bash
+Rscript manuscript/render.R

### 2023-02-20 analysis/analysis.Rmd
edit readme, recreate makefile, simplify analysis scripts

@@
 library(broman)
-
 library(phyloseq)
@@
 library(here)
+library(patchwork)
 
@@
   select(sample, passage_treat, passage, treat, selected, ch4, asv, count, Kingdom:Species, seq)
+
+```
+
+```{r}
+aes_data %>% 
+    select(sample, count, Phylum, Family) %>% 
+    group_by(sample) %>% 
+    mutate(sum_count = sum(count)) %>% 
+    mutate(ra = count / sum_count) %>% 
+    filter(is.na(Family)) %>%
+    # filter(Family == "") %>%
+    summarize(sum = sum(ra))
 
@@
 ```{r plot_raw}
-ratio <- calc_plot_ratios(fluxes$passage, fluxes$ch4)
 
@@
          y = "Methane Oxidation Rate (-k)") +
-    coord_fixed(ratio) +
   scale_x_continuous(labels = 1:5)
@@
 ```{r plot_log10}
-ratio_log10 <- calc_plot_ratios(fluxes$passage, fluxes$log_ch4)
+
+ratio = calc_plot_ratios(fluxes$passage, fluxes$log_ch4)
 
@@
                   y = expression(log[10] ~ CH[4] ~ "(-k)")) +
-        coord_fixed(ratio_log10) +
-  scale_x_continuous(labels = 1:5))
-
-save(p_response, file = here(der_dir, 'response_plot.RData'))
+  scale_x_continuous(labels = 1:5) +
+    coord_fixed(ratio))
+
+# save(p_response, file = here(der_dir, 'response_plot.RData'))
 
@@
 ```
+
+Since the residuals are not normally distributed, we will use the log-transformed data.
 
@@
 ```{r save_model}
-save(response_model, compare_response_mods, per_slope, file = here(der_dir, 'response_model.RData'))
+# save(response_model, compare_response_mods, per_slope, file = here(der_dir, 'response_model.RData'))
 ```
@@
 
-## Test heritability as mid-offspring on mid-parent
-
-```{r test_treatment}
-
-fit_no_treat <- lm(offspring ~ selected, data = heritability)
-kable(tidy(fit_no_treat))
-herit_model <- lm(offspring ~ selected * treat, data = heritability)
-kable(tidy(herit_model))
-kable(tidy(anova(fit_no_treat, herit_model)))
-summary(herit_model)
-
-herit_model_n <- 
+## Test heritability as divergence on cumulative selection differential
+
+```{r heritability}
+
+divergence <-
+    fluxes %>% 
+    select(passage, treat, log_ch4) %>% 
+    group_by(passage, treat) %>% 
+    summarize(mean = mean(log_ch4), .groups = "drop") %>% 
+    pivot_wider(names_from = "treat", values_from = "mean") %>% 
+    mutate(divergence = p - n, .keep = 'unused')
+
+# Selection Differential
+sd <- 
     heritability %>% 
-  filter(treat == "n") %>% 
-  lm(offspring ~ selected, data = .) %>% 
-  tidy()
-herit_model_n
-herit_model_p <- 
-heritability %>% 
-  filter(treat == "p") %>% 
-  lm(offspring ~ selected, data = .)
-summary(herit_model_p)
-```
-
-Here, I fit two models.
-The first is simply offspring flux on selected parent flux, ignoring treatment.
-There is no correlation `r print_lm(fit_no_treat, 2)`.
-Next, I added treatment and the interaction between treatment and parental flux.
-Here, there is a significant correlation between parent and offspring for both the Neutral treatment `r print_lm(herit_model, 2)` and the Positive treatment `r print_lm(herit_model, 4)`.
-Notably, the sign of the effect is reversed: negative for the Neutral treatment and positive for the Positive treatment.
-
-```{r plot_herit}
-
-n_label <- paste0('Neutral slope = ', myround(tidy(herit_model)[2, 2], 2))
-p_label <- paste0('Positive slope = ', myround(tidy(herit_model)[2, 2] + tidy(herit_model)[4, 2], 2))
-  ratio <- calc_plot_ratios(heritability$selected, heritability$offspring)
-(p_herit <- 
-  ggplot(heritability, aes(selected, offspring, color = treat)) +
+    mutate(sd = selected - parental) %>% 
+    select(passage, treat, sd) %>% 
+    pivot_wider(names_from = 'treat', values_from = 'sd') %>% 
+    mutate(sd_sum = p + n, .keep = 'unused')
+
+sd %>% 
+    arrange(passage) %>% 
+    mutate(sd_csum = cumsum(sd_sum)) ->
+    sd_csum
+
+# Realized Heritability
+rh <-
+    divergence %>% 
+    left_join(sd_csum, by = "passage") %>% 
+    filter(passage != 0 )
+```
+
+```{r}
+
+ratio_div <- calc_plot_ratios(divergence$passage, divergence$divergence)
+(
+p_div <-
+ggplot(divergence, 
+       aes(x = passage, 
+           y = divergence)) +
+    stat_smooth(method = 'lm', 
+                se = FALSE, 
+                formula = 'y ~ x',
+                color = "black",
+                linewidth = 0.5) +
     geom_point() + 
-    # stat_poly_eq() +
-    stat_smooth(method = 'lm', se = F) +
-    # annotate("text", x = min(heritability$parental), y = max(heritability$offspring), label = p_label, hjust = 0, color = "darkorange2") +
-    # annotate("text", x = min(heritability$parental), y = max(heritability$offspring) - 0.04, label = n_label, hjust = 0, color = "gray40") +
-    labs(x = expression(Donor ~ log[10] ~ CH[4] ~ "(-k)"), 
-         y = expression(atop(Recipient, log[10] ~ CH[4] ~ "(-k)"))) +
-    scale_color_manual(name = "Treatment", 
-                       labels = c('Neutral', 'Positive'), 
-                       values = c('gray40', 'darkorange2')) +
-    coord_fixed(1))
-
-save(p_herit, file = here(der_dir, 'herit_plot.RData'))
-
-# ggsave(here(fig_dir, 'fig2.jpg'), p_herit, width = 4, height = 3)
-# ggsave(here(fig_dir, 'fig2.svg'), p_herit, width = 4, height = 3)
-
-```
-
-```{r save_heritability}
-save(herit_model, file = here(der_dir, 'herit_model.RData'))
-```
-
-```{r}
-heritability_each <- read_tsv(here(der_dir, 'heritability_each.tsv'))
-
-fit_no_treat <- lm(offspring ~ selected, data = heritability_each)
-kable(tidy(fit_no_treat))
-herit_model <- lm(offspring ~ selected * treat, data = heritability_each)
-kable(tidy(herit_model))
-kable(tidy(anova(fit_no_treat, herit_model)))
-
-herit_model_n_each <- 
-heritability_each %>% 
-  filter(treat == "n") %>% 
-  lm(offspring ~ selected, data = .) %>% 
-  tidy()
-herit_model_p_each <- 
-heritability_each %>% 
-  filter(treat == "p") %>% 
-  lm(offspring ~ selected, data = .) %>% 
-  tidy()
-
-ratio <- calc_plot_ratios(heritability_each$selected, heritability_each$offspring)
-(p_herit_each <- 
-ggplot(heritability_each, aes(selected, offspring, color = treat)) +
-    # theme_classic() +
-    # geom_rangeframe(color = 'black') +
+    labs(x = expression(atop("Passage Number", " ")), 
+         y = expression(atop("Divergence", log[10] ~ CH[4] ~ "(-k)"))) +
+    coord_fixed(ratio_div) +
+    scale_x_continuous(labels = 1:5) + 
+    scale_y_continuous(n.breaks = 4) +
+    theme(plot.margin = margin(0, 0.25, 0, 0.5, "cm"))
+)
+ratio_rh <- calc_plot_ratios(rh$sd_csum, rh$divergence)
+
+(
+p_rh <-
+rh %>% 
+ggplot(aes(x = sd_csum, 
+           y = divergence)) +
+    stat_smooth(method = 'lm',
+                se = F,
+                formula = 'y ~ x',
+                color = "black",
+                linewidth = 0.5) +
     geom_point() + 
-    stat_smooth(method = 'lm', se = F) +
-    # stat_poly_line(se = T) +
-    # stat_poly_eq(aes(label = paste(after_stat(eq.label),
-                                 # after_stat(rr.label), sep = "*\", \"*"))) +
-    labs(x = expression(atop(Mean ~ Donor, log[10] ~ CH[4] ~ "(-k)")), 
-         y = expression(atop(Recipient, log[10] ~ CH[4] ~ "(-k)"))) +
-    scale_color_manual(name = "Treatment", 
-                       labels = c('Neutral', 'Positive'), 
-                       values = c('gray40', 'darkorange2')) +
-    coord_fixed(1))
-
-# save(p_herit_each, file = here(der_dir, 'herit_plot.RData'))
-# save(herit_model, file = here(der_dir 'herit_model.RData'))
-# 
-# ggsave(here(fig_dir, 'fig2_each.jpg'), p_herit_each, width = 4, height = 3) 
-# ggsave(here(fig_dir, 'fig2_each.svg'), p_herit_each, width = 4, height = 3) 
+    labs(x = expression(atop("Cumulative Selection Differential", log[10] ~ CH[4] ~ "(-k)")), 
+         y = expression(atop("Divergence", log[10] ~ CH[4] ~ "(-k)"))) +
+    scale_y_continuous(n.breaks = 4) +
+    # theme(plot.margin = margin(0, 0.5, 0, 0.25, "cm")) +
+    coord_fixed()
+)
+
+remove_y <- theme(
+    axis.text.y = element_blank(),
+    axis.ticks.y = element_blank(),
+    axis.title.y = element_blank()
+)
+p <- list(p_div, p_rh)
+herit_plot <- wrap_plots(p, nrow = 1) + plot_layout(guides = "collect")
+
+rh_model<-lm(divergence~sd_csum,data = rh)
+summary(rh_model)
+
+# save(herit_plot, file = here(der_dir, "herit_plot.RData"))
+
+# ggsave(here(fig_dir, "fig2.png"), width = 8*.9, height = 4*.9)
+# ggsave(here(fig_dir, "fig2.svg"), width = 8*.9, height = 4*.9)
+
+# save(rh_model, file = here(der_dir, 'herit_model.RData'))
+
 ```
@@
 min_seqs <- paste(substr(as.character(min_seqs), 1, 3), substr(as.character(min_seqs), 4, 6), sep = ",")
-save(kt, wt, rare_median, s5, min_seqs, file = here(der_dir, "richness_models.RData"))
+# save(kt, wt, rare_median, s5, min_seqs, file = here(der_dir, "richness_models.RData"))
 
@@
   #   panel.grid.minor = element_blank())
-
-save(p_beta, file = here(der_dir, "beta_fig.RData"))
-
-ggsave(here(fig_dir, 'fig3.png'), p_beta, height = 2.5, width = 4)
-
-ggsave(here(fig_dir, 'fig3.svg'), p_beta, height = 2.5, width = 4)
+  #   
+pcoa <- cmdscale(rare_dist_bray, eig = TRUE, add = TRUE)
+round(pcoa$eig*100/sum(pcoa$eig),1)
+pcoa_per_exp <- 100 * eigenvals(pcoa) / sum(eigenvals(pcoa))
+
+ratio = pcoa_per_exp[2]/pcoa_per_exp[1]
+
+pcoa_scores <-
+  scores(pcoa, display = "sites") %>% 
+  as_tibble(rownames = "sample")
+
+pcoa_data <-
+  aes_data %>% 
+  select(sample, passage, treat) %>% 
+  distinct %>% 
+  inner_join(pcoa_scores, by = "sample")
+# p_beta <-
+  pcoa_data %>%   
+  ggplot(aes(x = Dim1, y = Dim2, shape = passage, color = treat)) +
+  geom_point() +
+  stat_ellipse(show.legend = FALSE) +
+  coord_fixed(ratio) +
+  scale_shape_discrete(name = "Passage", labels = c(2, 5)) +
+  scale_color_manual(name = "Treatment",
+                       labels = c('Neutral', 'Positive'),
+                       values = c('gray40', 'darkorange2')) #+
+  # theme_bw() +
+  # theme(
+  #   plot.background = element_blank(),
+  #   panel.grid.major = element_blank(),
+  #   panel.grid.minor = element_blank())
+
+# save(p_beta, file = here(der_dir, "beta_fig.RData"))
+
+# ggsave(here(fig_dir, 'fig3.png'), p_beta, height = 2.5, width = 4) 
+# ggsave(here(fig_dir, 'fig3.svg'), p_beta, height = 2.5, width = 4)
 
@@
 
-save(tidy_permanova, file = here(der_dir, "beta_model.RData"))
+# save(tidy_permanova, file = here(der_dir, "beta_model.RData"))
 
@@
 
-save(dbrda_aov, dbrda_percent_explained, file = here(der_dir, "dbrda.RData"))
+# save(dbrda_aov, dbrda_percent_explained, file = here(der_dir, "dbrda.RData"))
 
@@
 
-save(p_log10, da_fams, file = here(der_dir, "da_fig.RData"))
-
-ggsave(here(fig_dir, 'fig4.jpg'), p_log10, width = 5, height = 5) 
-ggsave(here(fig_dir, 'fig4.svg'), p_log10, width = 5, height = 5)
+# save(p_log10, da_fams, file = here(der_dir, "da_fig.RData"))
+
+# ggsave(here(fig_dir, 'fig4.jpg'), p_log10, width = 5, height = 5) 
+# ggsave(here(fig_dir, 'fig4.svg'), p_log10, width = 5, height = 5)
 
@@
 ```
+
+```{r save_outputs}
+
+save(p_response, herit_plot, p_beta, p_log10, file = here(der_dir, 'plots.RData'))
+save(response_model, compare_response_mods, per_slope, rh_model, kt, wt, rare_median, s5, min_seqs, tidy_permanova, dbrda_aov, dbrda_percent_explained, da_fams, file = here(der_dir, 'models.RData'))
+
+# save(p_response, file = here(der_dir, 'response_plot.RData'))
+# save(response_model, compare_response_mods, per_slope, file = here(der_dir, 'response_model.RData'))
+# 
+# save(herit_plot, file = here(der_dir, "herit_plot.RData"))
+
+# save(rh_model, file = here(der_dir, 'herit_model.RData'))
+# save(kt, wt, rare_median, s5, min_seqs, file = here(der_dir, "richness_models.RData"))
+# save(p_beta, file = here(der_dir, "beta_fig.RData"))
+# save(tidy_permanova, file = here(der_dir, "beta_model.RData"))
+# save(dbrda_aov, dbrda_percent_explained, file = here(der_dir, "dbrda.RData"))
+# save(p_log10, da_fams, file = here(der_dir, "da_fig.RData"))
+
+```

### 2023-02-20 manuscript/manuscript.Rmd
edit readme, recreate makefile, simplify analysis scripts

@@
 ---
+
+{title max characters 150; current = 106 w/spaces}
 
@@
 library(ggtext)
+library(patchwork)
 figures <- captioner()
@@
 
-# CH4 models
-load(file = paste0(der_dir, 'response_model.RData'))
-load(file = paste0(der_dir, 'response_plot.RData'))
-load(file = paste0(der_dir, 'herit_model.RData'))
-load(file = paste0(der_dir, 'herit_plot.RData'))
+load(file = paste0(der_dir, 'plots.RData'))
+load(file = paste0(der_dir, 'models.RData'))
 
-# 16S models
-load(file = paste0(der_dir, 'richness_models.RData'))
-load(file = paste0(der_dir, 'phyloseq.RData'))
-load(file = paste0(der_dir, 'beta_model.RData'))
-load(file = paste0(der_dir, 'beta_fig.RData'))
-load(file = paste0(der_dir, 'da_corncob.RData'))
-load(file = paste0(der_dir, 'da_fig.RData'))
-load(file = paste0(der_dir, 'dbrda.RData'))
+# # CH4 models
+# load(file = paste0(der_dir, 'response_model.RData'))
+# load(file = paste0(der_dir, 'response_plot.RData'))
+# load(file = paste0(der_dir, 'herit_model.RData'))
+# load(file = paste0(der_dir, 'herit_plot.RData'))
+# 
+# # 16S models
+# load(file = paste0(der_dir, 'richness_models.RData'))
+# load(file = paste0(der_dir, 'phyloseq.RData'))
+# load(file = paste0(der_dir, 'beta_model.RData'))
+# load(file = paste0(der_dir, 'beta_fig.RData'))
+# load(file = paste0(der_dir, 'da_corncob.RData'))
+# load(file = paste0(der_dir, 'da_fig.RData'))
+# load(file = paste0(der_dir, 'dbrda.RData'))
 
@@
 
+{max 250 words; current = 249}
+
 Microbiomes mediate important ecosystem functions, yet it has proven difficult
@@
 ecosystem functions. This challenge remains because it is difficult to
-manipulate microbiome composition directly, we often cannot know a priori which
-members of a microbiome influence the rate of an ecosystem function, and
+manipulate microbiome composition directly, we often cannot know _a priori_ which
+microbiome members influence the rate of an ecosystem function, and
 microbiomes can covary strongly with other drivers of ecosystem function, such
@@
 strong response to selection with a `r myround(per_slope, 1)`\% increase in
-methane oxidation rate per passage. In addition, we
-estimated that `r myround((coefficients(herit_model)[2] +
-coefficients(herit_model)[4])*100, 1)`\% of the variation in methane oxidation
+methane oxidation rate per passage. We
+estimated that 40\% of the variation in methane oxidation
 rate in these soils can be attributed to microbiome variation. We also found
-that selection did not enrich for known methane oxidizers; rather microbes that
-responded to selection included members of `r length(da_fams)` families not
-known to oxidize methane, such as *Fimbriimonadaceae*, *Cytophagaceae*, and
-*Diplorickettsiaceae*. This result is in contrast to the typical assumption that
+that selection did not enrich for known methane oxidizers; instead, `r length(da_fams)` families not
+known to oxidize methane, including *Fimbriimonadaceae*, *Cytophagaceae*, and
+*Diplorickettsiaceae*, were enriched by selection. This result is in contrast to the typical assumption that
 the rate of an ecosystem function is limited by the final step in the associated
@@
 
+{max text 5000 words; current text = 4549}
+
 # Introduction
 
-Microbiomes mediate a variety of important ecosystem functions. This suggests
-that understanding variation in microbiome composition might be important for
-predicting the rates of ecosystem functions 
-[@conrad1996; @schimel1998; @crowther2019]. 
-Many studies have attempted to document such a relationship
+Microbiomes mediate a variety of important ecosystem functions relevant to human
+health, agriculture, and global change. As a result, there is great interest in
+understanding how to manipulate the microbiome to achieve desirable outcomes
+within these domains [@conrad1996; @schimel1998; @crowther2019]. However, for
+microbiome manipulations to be successful, variation in the microbiome must
+contribute directly to variation in the magnitude of the function of interest
+independent of other factors. Many studies have attempted to document such a
+relationship
 [@cavigelli2000; @balser2005; @waldrop2006; @strickland2009; @wagg2014; @wagg2019].
-However, it is difficult to isolate the effect of variation in microbiome
-composition from other drivers of variation in ecosystem function. For example,
-microbiome manipulations often confound variation in diversity, composition, and
-abundance. Likewise, comparative studies must control for abiotic variation
-statistically to identify the direct effect of microbiome variation. This
-approach relies on measuring the relevant abiotic variables and cannot account
-for unmeasured environmental variation. Here, we overcome these limitations by
-using a selection approach to estimate the degree to which an ecosystem function
-varies with microbiome composition.
+However, it is difficult to isolate the direct effect of variation in microbiome
+composition from other drivers of variation in ecosystem function, such as the
+indirect effect of the environment on function through microbiome assembly.
+Here, we overcome these limitations by using a selection approach to estimate
+the degree to which an ecosystem function varies with microbiome composition.
 
@@
 addition, studies focusing on the correlation between the rate of an ecosystem
-function and the abundance of an associated functional marker gene observe a
+function and the abundance of an associated marker gene (i.e., a gene that codes
+for a protein assumed to be involved in the ecosystem function) sometimes observe a
 significant correlation, though this relationship is rare and contingent upon
@@
 can affect the conclusions of such a comparative study. In addition, it is
-difficult to know a priori which environmental variables or community attributes
+difficult to know _a priori_ which environmental variables or community attributes
 to measure. Finally, while these approaches can establish a potential magnitude
@@
 community composition with other factors. For example, reciprocal transplant and
-common garden experiments confound community composition with the abiotic
-conditions carried on the inoculum while manipulating composition through
+common garden experiments can confound community composition with the abiotic
+conditions introduced with the inoculum, while manipulating composition through
 dilution may confound composition with biomass [@reed2007].
-
-(Side note: papers like the Strickland paper were actually asking a subtly
-different question from the one that we are asking. Strickland was asking
-whether communites can be functionally dissimilar as opposed to functionally
-redundant. That is, that historical contingency plays a role in the functioning
-of microbial communities. They demonstrated quite well that communities
-originating from different environments could have unique functional rates.
-However, is that due to microbiome-environment covariance where a different
-environment (i.e., a different litter substrate) selects for a litter-degrading
-community or is that a result of the unique effect of variation in the
-community? They actually point this out in their Discussion, they suggest future
-research should determine whether the differences between communities are driven
-by taxonomic composition or is function the result of the selection environment
-(I think of this is analogous to our "covariation" argument)? They also suggest
-trying to adapt communities to a new environment through continuous inputs of
-the same resource (just like in our experiment!))
 
@@
 manipulative studies by applying a different approach to the question of whether
-variation in the microbiome contributes to variation in the rate of an ecosystem
+microbiome variation contributes to variation in the rate of an ecosystem
 function. We used artificial ecosystem selection to select for microbiomes that
@@
 several potential advantages to this approach for documenting
-microbiome-function relationships and for investigating potential mechanisms
+microbiome-function relationships and for investigating the mechanisms
 underlying those relationships. By passaging microbiomes over multiple rounds
@@
 the environment by repeatedly diluting abiotic effects. In addition, by
-inoculating our ecosystems with equally sized subsamples, we eliminate the need
+inoculating our ecosystems with equally-sized subsamples, we eliminate the need
 to generate microbiome variation through methods that are confounded with
@@
 to a control community resulting from random selection, we can both control for
-changes in the environment over time and identify significant genes or taxa
-associated with our selection treatment to establish a potential mechanism that
-explains this variation.
+changes in the environment over time and identify genes or taxa that are
+associated with the ecosystem function under selection.
 
@@
 globally important greenhouse gas and methane oxidation by soil bacteria is the
-primary biological sink for atmospheric methane as well as possibly the main driver
-of temporal methane dynamics [@kirschke2013]. In addition, there is evidence that
-soil methane oxidation rate may vary with microbiome composition based on
-comparative studies in a variety of arctic and tropical ecosystems 
-[@meyer2020; @meyer2020a; @freitag2009; @freitag2010] as well as studies that
-manipulate methanotroph richness [@schnyder2018]. Finally, methanotrophy is one
-of the most deeply conserved microbial physiologies and is represented in a
-narrow range of taxa, which suggests that the taxonomic composition of the
-microbiome is more likely to be associated with the rate of methane oxidation
-than other broader or more shallowly conserved functions
+primary biological sink for atmospheric methane [@kirschke2013]. In addition,
+there is evidence that soil methane oxidation rate may vary with microbiome
+composition based on comparative studies in a variety of arctic and tropical
+ecosystems [@meyer2020; @meyer2020a; @freitag2009; @freitag2010] as well as
+studies that manipulate methanotroph richness [@schnyder2018]. Finally,
+methanotrophy is one of the most deeply conserved microbial physiologies and is
+represented in a narrow range of taxa, which suggests that the taxonomic
+composition of the microbiome is more likely to be associated with the rate of
+methane oxidation than other broader or more shallowly conserved functions
 [@schimel1998; @martiny2013].
 
-In this study, we evaluated whether variation in the microbiome contributes to
-variation in the rate of ecosystem functions. To address this problem, we
-applied artificial ecosystem selection to methane oxidation rate of the soil
-microbiome and answered the following questions: Does variation in the relative
-abundance of microbial taxa contribute to variation in soil methane oxidation
-rate in the lab? How much of the variation in soil methane oxidation rate can
-be attributed to the microbiome? What attributes of the microbiome regulate
-variation in methane oxidation rate? And do these attributes match our
-assumptions about what regulates methane oxidation rate in nature?
+In this study, we used artificial ecosystem selection on methane oxidation rate
+to address the following questions: Does variation in the relative abundance of
+microbial taxa contribute to variation in soil methane oxidation rate in our
+system? Which attributes of the microbiome are associated with variation in
+methane oxidation rate, and do these attributes match our assumptions about the
+factors that regulate methane oxidation rate in nature?
 
@@
 [*sensu* @swenson2000] by passaging replicate soil microbiomes. The trait we
-selected on was CH~4~ oxidation rate. This experiment had two selection lines
+selected on was methane (CH~4~)-oxidation rate. This experiment had two selection lines
 with twelve jars each for a total of twenty-four jars per passage. One line was
@@
 randomly assigned to either the positive or neutral selection line. Jars were
-flushed and respiked with methane twice per week to maintain aerobic conditions and elevated
-CH~4~ concentrations and were incubated at ambient temperature for
-approximately four weeks. Methane oxidation rates were determined at the end
-of the incubation period. For the positive treatment, the three jars with the
+flushed and respiked with CH~4~ twice per week to maintain aerobic conditions
+and elevated CH~4~ concentrations and were incubated at ambient temperature for
+approximately four weeks. Methane oxidation rates were determined at the end of
+the incubation period. For the positive treatment, the three jars with the
 greatest CH~4~ oxidation rates were chosen to inoculate the next generation.
@@
 the mean of the three selected jars and the mean of all twelve jars in a passage.
-Cumulative selection differential was then the sum of the selection differential
-from all preceding selection events. We test this as the regression of
+Cumulative selection differential was calculated as the sum of the selection
+differential from all preceding selection events. We then regressed cumulative
 divergence on cumulative selection differential using the 'lm' function.
@@
 Richness was estimated using the method from @hurlbert1971 with a subsample
-size of `r min_seqs` using the 'rarefy' function in 'vegan' [@oksanen2019]. We
+size of `r min_seqs` calculated via the 'rarefy' function in 'vegan' [@oksanen2019]. We
 tested a difference in richness by both passage and treatment with a
@@
 
-We estimated the proportion of variation in methane oxidation rate due to
-variation in the microbiome by regressing the divergence between the positive
-and neutral treatments against the cumulative selection differential. We
-observed a slope of 0.40, which would indicate that 40\% of the variation is due
-to the microbiome, but this was not significant (statistics).
-
-We estimated the ecosystem heritability defined as the proportion of variation
-in methane oxidation due rate to variation in the microbiome as the regression
-of mid-recipient on mid-donor (`r figures("herit", display = "cite")`).
-Recipient CH~4~ oxidation rates were correlated with donor CH~4~ oxidation rates
-in both the Positive treatment
-`r paste0('(', 'slope', ' = ', myround(tidy(herit_model)[2, 2] + tidy(herit_model)[4, 2], 2), ', SE = ', myround(tidy(herit_model)[4, 3], 2), ', t = ', myround(tidy(herit_model)[4, 4], 2), ', p = ', myround(tidy(herit_model)[4, 5], 2), ')')` and
-the Neutral treatment `r print_lm(herit_model, 2)`.
+To estimate the proportion of variation in CH~4~ oxidation rate due to
+variation in microbiome composition, we regressed divergence between the
+positive selection treatment and the control against the cumulative selection
+differential (`r figures("herit", display = "cite")`). This estimate is analogous
+to estimates of "microbiability" from the animal breeding literature, which
+quantifies the
+variation in a host trait that is due to microbiome variation [@difford2016]. 
+The slope of the regression of divergence on cumulative selection differential
+provides an estimate of realized microbiability (h^2^ &pm; SE), which in our
+experiment was `r myround(tidy(rh_model)[[2, 2]], 2)` &pm;
+`r myround(tidy(rh_model)[[2, 3]], 2)`, though this was not significant (F~`r glance(rh_model)$df`,`r glance(rh_model)$df.residual`~ = 
+`r myround(glance(rh_model)$statistic, 2)`, 
+p = `r myround(glance(rh_model)$p.value, 2)`).
 
@@
 We did not identify any methanotrophic families in the overall differential abundance analysis. 
-However, we wanted to look more closely at the known methanotrophs in our dataset to be sure they did not have an effect.
-To do this, we subset all of the ASVs in our dataset that were in families that contained methanotrophs.
+However, we wanted to look more closely at the known methanotrophs in our dataset to be sure that they did not have an effect.
+To do this, we subset the ASVs in our dataset that were in families that contained methanotrophs.
 Only two families were represented: *Methylacidiphilaceae* and *Beijerinckiaceae*. 
@@
 Positive treatment.  While many *Beijerinckiaceae* are methanotrophs, several
-taxa in this family have lost the the ability to oxidize CH~4~ and it appears
+taxa in this family have lost the ability to oxidize CH~4~ and it appears
 that *Rhodoblastus* species are not able to grow on CH~4~, though they can grow
-on methanol [@tamas2014, @dedysh2016]. Based on this analysis,
+on methanol [@tamas2014; @dedysh2016]. Based on this analysis,
 it appears that no methanotrophs were enriched in the Positive treatment.
@@
 
-In this study, we addressed whether variation in the microbiome contributes to
-variation in the rate of ecosystem function. To test this, we applied artificial
-ecosystem selection to soil microbiomes by selecting on the whole-ecosystem soil
-methane oxidation rate. We observed an increase in methane oxidation rate in the
+We used artificial ecosystem selection to estimate the contribution of variation
+in microbiome composition to variation in the rate of an ecosystem function,
+methane oxidation in soil, independent of environmental variation. Understanding
+how and to what degree microbiome variation contributes to variation in
+ecosystem function is important for many reasons. For example, successful
+microbiome manipulations require that the manipulated microbiome contribute to
+variation in ecosystem function independent of other drivers of ecosystem
+variation (such as variation in environmental conditions). This is because the
+drivers of variation in ecosystem functions can interact in complicated ways
+(Figure 5);
+for example, environmental variation can directly contribute by providing
+conditions that select for microbial groups that in turn alter the rates of
+ecosystem functions. Determining the independent contribution of microbiome
+variation to ecosystem function is crucial because if microbiome composition is
+driven primarily by environmental conditions, then introducing a desirable taxon
+through microbiome manipulation without altering the environment will likely be
+unsuccessful at shifting the targeted ecosystem function. 
+The artificial selection approach is different from the comparative and
+manipulative approaches used in past attempts at answering this question,
+because it can control for both the direct effect of environment on function as
+well as the indirect effect on environment via its impact on microbiome
+assembly.
+
+In our study, we observed an increase in methane oxidation rate in the
 selection treatment relative to the control, which demonstrates that there was a
@@
 conclude that variation in the microbiome contributes to variation in the
-methane oxidation rate independent of the environment.
+methane oxidation rate independent of the environment. This suggests that
+microbiome manipulations could be an effective approach for altering the rate of
+methane oxidation in this soil, and that the artificial selection approach may
+be useful in determining the potential for microbiome manipulations for other
+functions in other ecosystems.
 
-The ultimate goal of this study (and much of microbial ecology) could be
-summarized as "Do microbes matter to ecosystem function (or health)?" and there
-are at least three ways we could address that question. One would be to ask
-whether microorganisms mediate function. This question is perhaps trivial. In
-the case of methane emissions from soil, we know that microorganisms produce and
-consume methane. We could also ask what would happen if we removed the
-microorganisms. Again, this question is trivial. There are no known abiotic
-reactions in soil that consume methane (outside of the troposphere, the
-stratosphere, and the marine boundary layer @kirschke2013). Therefore, removing
-the microbiome would likely eliminate methane production and consumption. Then,
-the third question we could ask is whether *variation* in the microbiome is
-associated with *variation* in the rate of ecosystem function. The answer to
-this question is not so obvious.
+Given that variation in the microbiome is associated with variation in the rate
+of an ecosystem function in our system, a reasonable follow-up question is "how
+much variation in ecosystem function is associated with microbiome variation in
+this system?" One way to estimate this is to determine how much the recipient
+jars resemble the selected donor jars that were used to inoculate them
+[@falconer1996]. We can
+calculate the response to selection as the difference between two successive
+passages in their mean methane oxidation rate. We will denote this as $R$. We
+can also calculate the strength of selection as the difference in mean methane
+oxidation rate between the twelve jars in one generation and the three jars chosen
+for selection in that generation, which we will call the selection differential
+and denote as $S$. If we plot the cumulative $R$ against the cumulative $S$, the
+slope of this relationship will equal the proportion of variation explained by
+the microbiome. If the change in mean function from passage one to passage two
+($R$) is equal to the difference in mean function between the twelve jars in
+passage one and the three jars selected to inoculate passage two ($S$), then we
+would conclude that 100% of the variation is due to variation in the microbiome.
+Likewise, if recipients do not resemble the donors in their mean methane
+oxidation rate and simply wander randomly, then we would conclude that all of
+the variation is due to the environment or technical variation.
 
-If we look at variation in the rate of a microbially-mediated ecosystem function
-across a set of habitats, that variation could be associated with variation in
-the abiotic environmental conditions, variation in the composition of
-microbiomes, or some interaction or covariance between the two (see Figure 5).
-To address the question of whether variation in the microbiome matters for
-ecosystem function, we need to identify the component of variation uniquely
-associated with the microbiome. If we sample natural variation, as in the
-comparative approach, it can be difficult to distinguish between the direct
-effect of the microbiome and the effect of the microbiome due to covariance with
-the environment. Likewise, with the manipulative approach even though we are
-placing ecosystem propagules into a common environment or substrate, the
-propagule itself carries with it environmental conditions that
-cannot be removed. Generating variation in microbiome composition through
-filtering or other similar means often confounds composition with other factors,
-such as biomass or cell size. To overcome these limitations and to isolate
-the direct effect of the microbiome on ecosystem function, we used
-artificial ecosystem selection.
+This relationship between microbiome variation and ecosystem function variation
+is analogous to the concept of "heritability" [@lynch1998] used by quantitative
+geneticists, or more precisely the concept of "microbiability" [@difford2016] proposed
+by microbiome scientists who study host-associated microbiomes. Although rarely
+used in the study of environmental microbiomes, this concept could be very
+useful for understanding and manipulating microbially-mediated functions in a
+variety of ecosystems. In our experiment, variation in microbiome taxonomic
+composition statistically explained (i.e., was associated with) 40% of the
+variation we observed in the rate of methane oxidation.  This suggests that
+there is substantial potential for altering this ecosystem function through
+microbiome manipulation in this soil. It is very likely that the
+“environmental microbiability” will be different for other ecosystem functions
+in this soil and for methane oxidation in other soils. However, our experiment
+demonstrates that this relationship is measurable and provides an example of how
+this can be accomplished.
 
-To perform artificial ecosystem selection, we generate a set of ecosystems in
-bottles using a common, sterile substrate and inoculate them with a propagule
-containing a diverse microbiome that is a subsample of an existing microbiome.
-This generates variation in microbiome composition among the ecosystems. We can
-then measure the rate of an ecosystem function for each bottle and select the
-bottles with the highest (or lowest) rate of ecosystem function. The selected
-ecosystems are then homogenized and used as the propagule to inoculate a new set of
-ecosystems containing the same sterile substrate as the initial set. By
-subsampling the microbiome, we can generate variation in the composition of the
-microbiome that is independent of cell size or biomass. And by passaging the
-microbiome through the same sterile substrate over multiple rounds of selection,
-we can progressively minimize the effect of the abiotic conditions from the
-donor environment on both the rate of function and the composition of the
-microbiome. In this way, over multiple rounds of selection, we can observe
-whether varying the microbiome can generate variation in the rate of our
-ecosystem function. If we observe no change in ecosystem function from
-passage to passage despite our imposed selection, we would conclude that this
-ecosystem function in this set of ecosystems is driven primarily by
-environmental variation or technical error among the replicates. However, if we
-observe a response to selection, that is, a shift in the mean ecosystem function
-between successive passages, we would conclude that variation in the microbiome can
-generate variation in the rate of ecosystem function. And that is, in fact, what
-we observed.
-
-Given that variation in the microbiome is associated with variation in ecosystem
-function, a reasonable follow-up question is "How much variation?" This is a
-challenging question to answer since the value depends on the population of
-ecosystems that is sampled and the estimate for one set of ecosystems may not
-apply to another set. For example, we cannot estimate a percent of variation
-explained in tropical rainforests and then assume that number applies in polar
-wetlands. In addition, the magnitude of the estimate will depend on how much
-variation is sampled. A large $\frac{V_M}{V_F}$ is not very interesting if $V_F$
-is small.  Keeping these caveats in mind, one way to estimate this number is to
-determine how much the recipient jars resemble the selected donor jars that were
-used to inoculate them. We can calculate the response to selection as the
-difference between two successive passages in their mean methane oxidation rate.
-We will denote this as $R$. We can also calculate the strength of selection as
-the difference in mean methane oxidation rate between the twelve jars in one
-generation and the 3 jars chosen for selection in that generation, which we will
-call the selection differential and denote as $S$. If we plot the cumulative
-$R$ against the cumulative $S$, the slope of this relationship will equal the
-proportion of variation explained by the microbiome.
-
-Using this approach, we observed a relatively large percent of variation
-explained by the microbiome, but this number was not significantly different
-from zero.
-
-The concept that we have just described is analogous to the concept of
-heritability from quantitative genetics. Heritability for given a population is
-defined as the proportion of variation in a phenotype associated with variation
-in genotype [@lynch1998]. This concept has been applied above the population
-level for groups within a species [@goodnight2000], as well as communities and
-ecosystems [@whitham2003; @whitham2006]. One study previously estimated the
-heritability of a community trait after artificial ecosystem selection using
-parent-offspring regression and observed a relatively large heritability of 50\%
-[@blouin2015]. However, parent-offspring regression assumes a randomly mating
-population, which, by definition, does not apply to a selection experiment. In
-the present study, we estimated the variation explained by the microbiome
-somewhat differently using the breeder's equation ($R = S h^2$) and regressing
-the observed response to selection against the imposed selection differential
-(Figure 1; @falconer1996). We also observed a relatively large response to selection
-(40\%), though this estimate had large error bars and was not significant so we
-did not have the power to estimate an accurate percent of variation explained
-for this experiment.
-
-Asking how much of the variation in ecosystem function is due to variation in
-the microbiome is a silly question. We cannot answer this questions broadly
-for all communities or all ecosystems, because the answer depends in large part
-on the magnitude of variation sampled.
-
-What do I want to get across here:
-
-This estimate is analogous to the concept of heritability from quantitative
-genetics
-Define heritability
-Point out that it's been applied to communities and ecosystems
-Mention the different ways of estimating heritability
-Mention the assumptions inherent in some of these estimates (non-assortative
-mating; particularly numbers of shared alleles based on kinship)
-Point out that while heritability is a useful framework to borrow from in order
-to describe the components of variation, there are limitations in its
-application
-Propose that we pursue new theory about how to understand what drives variation
-in microbiome functions and microbiome-encoded traits
-
-The challenge with applying the concept of heritability outside its original
-scope is that it is unclear if the assumptions underlying these estimates apply
-to microbiomes. For a genomic trait, we can rely on the assumption that
-offspring share, on average, 50\% of the alleles of each of their
-parents. It is less clear if offspring share 50\% of the microbes of each of
-their parents or if it is sufficient for 50\% of the microbes to be _present_ or
-if they must be in a similar relative abundance. Estimates of heritability, such
-as parent-offspring regression and the ACE model rely on specific estimates of
-the similarity of relatives based on their relatedness. It is unclear if these
-assumptions apply to the microbiomes of host ecosystems. It is even less clear
-whether they apply to non-host ecosystems that do not reproduce through sexual
-recombination.
-
-Given that variation in the microbiome is driving the observed variation in
-methane oxidation rate, we next wanted to address what aspects of the
-microbiome, in terms of alpha diversity, beta diversity, or the relative
-abundance of taxa, explain this contribution of the microbiome to variation in
-methane oxidation rate. There are three ways that the microbiome could respond
-to selection in this experiment: gain or loss of species, changes in the
-relative abundances of species, or changes within the genomes of the constituent
-species. Using 16S rRNA gene sequences, we are unable to address the third
-possibility of whether taxa in this experiment could have evolved changes in
-their genomes as a result of selection.  Therefore, we will focus on the first
+Given that variation in the microbiome is driving substantial variation in
+methane oxidation rate in our system, we next wanted to determine which aspects of the
+microbiome best explain this contribution of microbiome variation to
+methane oxidation rate. There are three inter-related ways that microbiomes could have responded
+to selection in this experiment: gain or loss of taxa, changes in the
+relative abundances of taxa, or changes within the genomes of the constituent
+taxa. We surveyed microbiome variation via 16s rRNA ribotyping in our
+experiment, which allowed us to deeply sample taxonomic diversity but did not
+allow us to directly address whether taxa in this experiment evolved genomic changes
+as a result of selection. However, if such genomic changes resulted in increased
+persistence or abundance of the population with these changes, this would be
+detectable. Therefore, we will focus on the first
 two possibilities.
@@
 and that it is often assumed that the rate of an ecosystem function is limited
-by the final enzymatic step in a metabolic pathway [@rocca2015]. In certain
+by the final enzymatic step in the underlying metabolic pathway [@rocca2015]. In certain
 ecosystems, methane production and consumption are correlated with the
 abundance of methanogens and methanotrophs as estimated from marker genes
-[@meyer2020; @meyer2020a]. However, it does not appear generally true that the
-rate of an ecosystem function is limited by the abundance of the gene that
-encodes the final step in that pathway [@rocca2015]. Our results suggest that
+[@meyer2020; @meyer2020a]. However, our results suggest that
 ecosystem-scale methane oxidation rates can be altered by non-methanotrophs,
-perhaps through ecological interactions with methanotrophic species.
+perhaps through ecological interactions with methanotrophic species. This
+suggests that simple assumptions about how microbes contribute to rate variation
+in ecosystem function may not apply universally, and it demonstrates the
+importance of using biologically "agnostic" approaches (that make few starting
+assumptions) to linking microbial taxa to ecosystem functions [@morris2020].
+Artificial ecosystem selection is an important example of such an approach.
 
-Another important question is whether selection at the ecosystem scale resulted
-in within-strain evolution. Using 16S rRNA gene sequencing we are unable to
-resolve genomic changes within taxa that may have undergone evolution.
-Therefore, it is possible that the response to selection was driven by changes
-within the genomes of methanotrophs. But given the slow growth rate of
-methanotrophs, often requiring weeks to years of serial batch culturing for
-isolation, this seems like an unlikely explanation 
-[@whittenbury1970; @hirayama2014; @kim2018]. Future work could analyze an
-artificial ecosystem selection experiment using genome-resolved metagenomics to
-try to detect strain-level changes in the constituent taxa. However,
-reconstructing genomes from in situ soil samples has been difficult to achieve
-and may require enrichment techniques, particularly for methanotrophs which are
-a relatively rare component of the soil microbiome [@delmont2015]. If these
-challenges could be overcome for soil metagenomics, we could begin to evaluate
-whether ecosystem selection results in strain-level evolution or enriches for
-certain metabolic pathways within the microbiome.
-
-A fundamental question in microbial ecology is whether variation in the
-microbiome contributes to variation in ecosystem function. We performed
-artificial ecosystem selection on soil methane oxidation rate to test the
-relationship between microbiome variation and soil methane oxidation rate. We
-observed a response to selection on soil methane oxidation rate and found a
-high heritability for methane oxidation, which suggests that variation in the
-microbiome contributes to variation in soil methane oxidation rate
-independent of the environment. This shows that manipulating the microbiome
-could be a viable strategy for mitigating or enhancing certain
-ecosystem functions.  Surprisingly, we did not observe an increase in the
-relative abundance of methanotrophs as a response to selection. This
-suggests that the rate of methane oxidation could be altered by
-non-methanotrophs, possibly through ecological interactions with methanotrophs.
-Future research should investigate other aspects of microbiome variation, such
-as strain-level evolution or functional gene composition, to begin to describe
-a more detailed mapping between the microbiome and ecosystem function. This
-study demonstrates that soil methane oxidation rate can vary greatly with
-variation in the microbiome and opens up future research opportunities to
-establish a more detailed understanding of the relationship between the
-microbiome and ecosystem function.
+There is increasing interest in using artificial selection for understanding and
+manipulating the microbiomes associated with plants and animals (aka “microbiome
+breeding”; @mueller2022). Our study demonstrates that artificial ecosystem
+selection can also be an important tool for exploring the relationship between
+microbiome composition and ecosystem function in non-host systems. This
+approach can provide unique information about the independent contribution of
+microbiomes to ecosystem functions. Such information is crucial if we are to
+successfully manipulate environmental microbiomes to alter ecosystem functions,
+whether to improve crop productivity [@chaparro2012]  or
+ameliorate the impacts of environmental change [@jansson2020].
 
@@
 `r figures("response", "Response to selection on soil CH~4~ oxidation rate.
-The y-axis is the log~10~ of the additive inverse of the first-order exponential decay constant $k$ (i.e., $-k$) with units day^-1^.
-The x-axis is the passage number.
-The orange line and points are the Positive treatment and the gray line and points are the Neutral treatment.")`
+The y-axis is the log~10~ of the additive inverse of the first-order exponential decay constant $k$ (i.e., $-k$) with units day^-1^ so that a more positive value represents a higher CH~4~ oxidation rate.")`
 
-```{r herit, fig.width=5, fig.height=3}
-
-p_herit
-
-```
-
-`r figures("herit", paste0("Ecosystem heritability as the regression of mid-recipient methane flux on mid-donor methane flux ($-k$ with units day^-1^).
+```{r herit, fig.width=7}
+# 
+herit_plot
+#`r figures("herit", paste0("Ecosystem heritability as the regression of mid-recipient methane flux on mid-donor methane flux ($-k$ with units day^-1^).
 Donor is the mean oxidation rate of the three jars selected to inoculate the next passage.
@@
 For the Positive treatment, the variance explained by the microbiome was ", myround((coefficients(herit_model)[2] + coefficients(herit_model)[4])*100, 1), "% (slope = ", myround(coefficients(herit_model)[2] + coefficients(herit_model)[4], 3), ", SE = ", myround(tidy(herit_model)$std.error[4], 2), ") and for the Neutral treatment the variance explained by the microbiome was ", myround(coefficients(herit_model)[2]*100, 1), "% (slope = ", myround(coefficients(herit_model)[2], 3), ", SE = ", myround(tidy(herit_model)$std.error[2], 2), ")."))`
+```
+
+`r figures("herit", paste0("Divergence between the positive selection line and the control line over five passages.
+Values are CH~4~ oxidation rate as the additive inverse of the first-order exponential decay constant k (day^-1^).
+Regression of divergence on cumulative selection differential provides an estimate of realized heritability (h^2^ &pm; SE) and the slope of the regression was ", myround(tidy(rh_model)[[2, 2]], 2), " &pm; ", myround(tidy(rh_model)[[2, 3]], 2), "."))`
 
@@
 
-`r figures("beta", "Principal component analysis plot of beta diversity for all jars.
+`r figures("beta", "Non-metric multidimensional scaling plot of beta diversity for all jars.
 Dissimilarities are based on rarefied Bray-Curtis dissimilarity.
@@
 Values on the x-axis are relative abundances on a log~10~ scale.
-Points are colored by treatment.
 Large points with error bars are the median relative abundance and interquartile range.

### 2023-02-20 analysis/new_analysis.R

@@ -1,164 +1,0 @@
-# Response to selection
-0.1779718
-# Standard Error
-0.0643900
-ratio_log10 <- calc_plot_ratios(fluxes$passage, fluxes$log_ch4)
-
-ggplot(fluxes, aes(x = passage, y = log_ch4, color = treat)) +
-    stat_smooth(method = 'lm', se = FALSE, formula = 'y ~ x') +
-    scale_color_manual(
-        name = "Treatment", 
-        labels = c('Neutral', 'Positive'), 
-        values = c('gray40', 'darkorange2')) +
-    geom_jitter(#aes(alpha = selected), 
-        width = 0.1) + 
-    labs(x = "Passage Number", 
-         y = expression(log[10] ~ CH[4] ~ "(-k)")) +
-    coord_fixed(1) +
-    scale_x_continuous(labels = 1:5)
-
-model1 <- lm(log_ch4 ~ passage * treat, data = fluxes)
-summary(model1)
-
-divergence <-
-    fluxes %>% 
-    select(passage, treat, log_ch4) %>% 
-    group_by(passage, treat) %>% 
-    summarize(mean = mean(log_ch4), .groups = "drop") %>% 
-    mutate(treat = factor(treat)) %>% 
-    pivot_wider(names_from = "treat", values_from = "mean") %>% 
-    mutate(divergence = p - n)
-
-ratio_div <- calc_plot_ratios(divergence$passage, divergence$divergence)
-
-ggplot(divergence, 
-       aes(x = passage, 
-           y = divergence)) +
-    stat_smooth(method = 'lm', 
-                se = FALSE, 
-                formula = 'y ~ x',
-                color = "black",
-                linewidth = 0.5) +
-    geom_point() + 
-    labs(x = "Passage Number", 
-         y = expression(atop("Divergence", log[10] ~ CH[4] ~ "(-k)"))) +
-    coord_fixed(ratio_div) +
-    scale_x_continuous(labels = 1:5)
-
-model2 <- lm(divergence ~ passage, data = divergence)
-summary(model2)
-
-sd <- 
-    heritability %>% 
-    mutate(sd = selected - parental) %>% 
-    select(passage, treat, sd) %>% 
-    pivot_wider(names_from = 'treat', values_from = 'sd') %>% 
-    mutate(sd = p - n) %>% 
-    select(-n, -p)
-
-sd$cumulative_sd <- 
[…]

### 2023-02-20 analysis/scripts/02-process-herit.R

@@ -14,6 +14,65 @@ source('R/functions.R')
 selected_jars <- read_csv(paste0(raw_dir, 'selected.csv'))
 fluxes <- read_tsv(paste0(der_dir, 'fluxes.tsv'))
 
+
+#######
+# New approach?
+
+fluxes %>% 
+    select(ch4, passage, treat) %>% 
+    group_by(passage, treat) %>% 
+    summarize(mean_flux = mean(ch4), .groups = "drop") %>% 
+    filter(treat == 'n') %>% 
+    select(-treat) ->
+    mean_fluxes
+
+fluxes %>% 
+    inner_join(mean_fluxes, by = 'passage') %>% 
+    filter(treat == 'p') %>% 
+    mutate(adjusted_ch4 = ch4 - mean_flux) %>% 
+    select(passage, selected, adjusted_ch4) ->
+    heritability
+
+heritability %>% 
+    filter(selected) %>% 
+    group_by(passage)->
+    selected
+heritability ->
+    parental
+heritability %>% 
+    filter(passage != 1) %>% 
+    mutate(passage = passage - 1) ->
+    offspring
+
+selected %>% 
+    group_by(passage) %>% 
+    summarize(selected = mean(adjusted_ch4)) ->
+    selected_mean
+
+parental %>% 
+    group_by(passage) %>% 
+    summarize(parental = mean(adjusted_ch4)) ->
+    parental_mean
+
+offspring %>% 
+    group_by(passage) %>% 
+    summarize(offspring = mean(adjusted_ch4)) ->
+    offspring_mean
+
+selected_mean %>% 
+    inner_join(parental_mean, by = 'passage') %>% 
+    inner_join(offspring_mean, by = 'passage') ->
+    heritability
+
+lm(offspring ~ selected, data = heritability) %>% 
+    summary()
+
+ggplot(heritability, aes(selected, offspring)) + 
+    geom_point() +
+    stat_smooth(method = 'lm') +
+    coord_fixed()
+#######
+
 ## Organize parent/offspring values
 
 # Pull out the selected parental jars
@@ -76,7 +135,7 @@ heritability_each <-
   filter(passage != 5) %>% # Remove passage 5, which has no offspring
   ungroup()
 
-write_tsv(heritability_each, paste0(der_dir, 'heritability_each.tsv'))
+#write_tsv(heritability_each, paste0(der_dir, 'heritability_each.tsv'))
 
 
 write_tsv(heritability, paste0(der_dir, 'heritability.tsv'))

### 2023-02-20 analysis/scripts/04-dada2.R

@@ -118,5 +118,5 @@ taxa.print <- taxa # Removing sequence rownames for display only
 rownames(taxa.print) <- NULL
 head(taxa.print)
 
-save(seqtab.nochim, taxa, file = "output/dada2_output.RData")
+save(seqtab.nochim, taxa, file = "hpc_output/dada2_output.RData")
 

### 2023-02-20 analysis/scripts/05-combine-asvs-and-fluxes.R

@@ -5,7 +5,7 @@ library(knitr)
 library(vegan)
 source('R/functions.R')
 
-load(file = paste0(der_dir, "dada2_output.RData"))
+load(file = "analysis/hpc_output/dada2_output.RData")
 
 samples <- read_tsv(paste0(der_dir, "barcode_master.tsv"))
 fluxes <- read_tsv(paste0(der_dir, "fluxes.tsv"))

### 2023-02-20 manuscript/render.R
edit readme and cleanu up old scripts

@@ -1,1 +1,0 @@
-rmarkdown::render('manuscript/manuscript.Rmd')

### 2023-02-20 my_render.sh

@@ -1,2 +1,0 @@
-#!/bin/bash
-Rscript manuscript/render.R

### 2023-02-21 manuscript/manuscript.Rmd
update conceptual figure

@@
 the degree to which an ecosystem function varies with microbiome composition.
+
+(`r figures("concept", display = "cite")`)
 
@@
 drivers of variation in ecosystem functions can interact in complicated ways
-(Figure 5);
+(`r figures("concept", display = "cite")`);
 for example, environmental variation can directly contribute by providing
@@
 
+![](Figure1.png)
+
+`r figures("concept", "Variation in the rate of an ecosystem function is potentially the result of at least three components: variation in the abiotic environmental conditions, variation in microbiome composition, and the covariance between microbiomes and the environment.
+It is difficult with traditional study designs to isolate the direct effect of the microbiome from the effect of the environment via covariance with the microbiome.
+Here, we attempt to isolate the effect of the microbiome through artificial selection on microbiome composition.
+For simplicity, we omitted the reverse arrows as well as the interaction between microbiome and the environment, though these relationships may also exist.")`
+
 ```{r response, fig.width=4, fig.height=3}
@@
 Taxa are sorted by their effect size with taxa at the top having the largest positive effect size and taxa at the bottom with the largest negative effect size.")`
-
-![Figure 5.
-Variation in the rate of an ecosystem function is potentially the result of at least three components: variation in the abiotic environmental conditions, variation in microbiome composition, and the covariance between microbiomes and the environment.
-It is difficult with traditional study designs to isolate the direct effect of the microbiome from the effect of the environment via covariance with the microbiome.
-Here, we attempt to isolate the effect of the microbiome through artificial selection on microbiome composition.
-For simplicity, we omitted the reverse arrows as well as the interaction between microbiome and the environment, though these relationshps may also exist.](Figure1.png)

### 2023-02-27 manuscript/manuscript.Rmd
clean up references and update figure captions

@@
 {title max characters 150; current = 106 w/spaces}  
-
 ^1^ Institute of Ecology and Evolution, University of Oregon, Eugene, OR, USA  
@@
 
-load(file = paste0(der_dir, 'plots.RData'))
+#load(file = paste0(der_dir, 'plots.RData'))
 load(file = paste0(der_dir, 'models.RData'))
@@
 {max 250 words; current = 249}  
-
 Microbiomes mediate important ecosystem functions, yet it has proven difficult
@@
 methane oxidation rate per passage. We
-estimated that 40\% of the variation in methane oxidation
+estimated that `r myround(tidy(rh_model)[[2, 2]]*100, 1)`\% of the variation in
+methane oxidation
 rate in these soils can be attributed to microbiome variation. We also found
@@
 variety of ecosystems. In our experiment, variation in microbiome taxonomic
-composition statistically explained (i.e., was associated with) 40% of the
+composition statistically explained (i.e., was associated with) 
+`r myround(tidy(rh_model)[[2, 2]]*100, 1)`\% of the
 variation we observed in the rate of methane oxidation.  This suggests that
@@
 
-# Figures
-
-![](Figure1.png)
+# Figure Captions
 
@@
 
-```{r response, fig.width=4, fig.height=3}
+`r figures("response", paste0("Response to selection on soil CH~4~ oxidation rate.
+The y-axis is CH~4~ oxidation rate as the log~10~ of the additive inverse of the first-order exponential decay constant $k$ (i.e., $-k$) with units day^-1^ so that a more positive value represents a higher CH~4~ oxidation rate.
+There was a significant difference of slopes between the positive selection treatment and the control (F~", compare_response_mods[2,4], ",", compare_response_mods[2,2], "~ = ", myround(compare_response_mods[2,5], 2), ", p = ", myround(compare_response_mods[2,6], 2), ")."))`
 
-p_response
+`r figures("herit", paste0("Divergence between the positive selection line and the control line over the cumulative selection differential.
+Values are CH~4~ oxidation rate as the additive inverse of the first-order exponential decay constant k (day^-1^).
+Regression of divergence on cumulative selection differential provides an estimate of realized heritability (h^2^ &pm; SE) and the slope of the regression was ", myround(tidy(rh_model)[[2, 2]], 2), " &pm; ", myround(tidy(rh_model)[[2, 3]], 2), ", though this was not significant (F~", glance(rh_model)$df,  ",", glance(rh_model)$df.residual,  "~ = ", myround(glance(rh_model)$statistic, 2),  ", p =  ", myround(glance(rh_model)$p.value, 2),  ")."))`
 
-```
-
-`r figures("response", "Response to selection on soil CH~4~ oxidation rate.
-The y-axis is the log~10~ of the additive inverse of the first-order exponential decay constant $k$ (i.e., $-k$) with units day^-1^ so that a more positive value represents a higher CH~4~ oxidation rate.")`
-
-```{r herit, fig.width=7}
-# 
-herit_plot
-#`r figures("herit", paste0("Ecosystem heritability as the regression of mid-recipient methane flux on mid-donor methane flux ($-k$ with units day^-1^).
-Donor is the mean oxidation rate of the three jars selected to inoculate the next passage.
-Recipient is the mean oxidation rate of the 12 jars in the next passage.
-The orange lines and points are the Positive treatment and the gray lines and points are the Neutral treatment.
-Slope of the regression line estimates the proportion of variation in methane oxidation rate due to variation in the microbiome.
-For the Positive treatment, the variance explained by the microbiome was ", myround((coefficients(herit_model)[2] + coefficients(herit_model)[4])*100, 1), "% (slope = ", myround(coefficients(herit_model)[2] + coefficients(herit_model)[4], 3), ", SE = ", myround(tidy(herit_model)$std.error[4], 2), ") and for the Neutral treatment the variance explained by the microbiome was ", myround(coefficients(herit_model)[2]*100, 1), "% (slope = ", myround(coefficients(herit_model)[2], 3), ", SE = ", myround(tidy(herit_model)$std.error[2], 2), ")."))`
-```
-
-`r figures("herit", paste0("Divergence between the positive selection line and the control line over five passages.
-Values are CH~4~ oxidation rate as the additive inverse of the first-order exponential decay constant k (day^-1^).
-Regression of divergence on cumulative selection differential provides an estimate of realized heritability (h^2^ &pm; SE) and the slope of the regression was ", myround(tidy(rh_model)[[2, 2]], 2), " &pm; ", myround(tidy(rh_model)[[2, 3]], 2), "."))`
-
-```{r beta, results=FALSE, fig.height=2.5}
-
-p_beta
-
-```
-
-`r figures("beta", "Non-metric multidimensional scaling plot of beta diversity for all jars.
-Dissimilarities are based on rarefied Bray-Curtis dissimilarity.
-Colors represent treatment and shapes represent passage number.")`
-
-```{r da, results=FALSE}
-
-p_log10
-
-```
+`r figures("beta", "Non-metric multidimensional scaling plot of beta diversity for all jars. Dissimilarities are rarefied Bray-Curtis dissimilarity averaged over 99 subsamples.")`
 
@@
 Values on the x-axis are relative abundances on a log~10~ scale.
-Large points with error bars are the median relative abundance and interquartile range.
-Taxa in the top panel are enriched in the Positive treatment and taxa in the bottom panel are depleted in the Positive treatment.
+Taxa in the top panel are enriched in the positive selection treatment relative to the control and taxa in the bottom panel are depleted in the positive selection treatment relative to the control.
 Taxa are sorted by their effect size with taxa at the top having the largest positive effect size and taxa at the bottom with the largest negative effect size.")`

### 2023-02-27 analysis/analysis.Rmd
update figures and analysis scripts

@@
 author: "Andrew Morris"
-date: "2022-11-21"
+date: "`r Sys.Date()`"
 output: 
@@
 library(patchwork)
+library(extrafont)
 
@@
 set.seed(1337)
+
+theme_set(theme_classic(base_size = 7) +
+            theme(panel.border = element_rect(fill = NA, linewidth = 1.0625),
+                  axis.line = element_blank(),
+                  text = element_text(family = 'Arial')))
 ```
@@
         name = "Treatment", 
-        labels = c('Neutral', 'Positive'), 
+        labels = c('Control', 'Selection'), 
         values = c('gray40', 'darkorange2')) +
@@
 
-(p_response <- ggplot(fluxes, aes(x = passage, y = log_ch4, color = treat)) +
-    stat_smooth(method = 'lm', se = FALSE, formula = 'y ~ x') +
-    scale_color_manual(
+# (p_response_cl <- ggplot(fluxes, aes(x = passage, y = log_ch4, color = treat)) +
+#     stat_smooth(method = 'lm', se = FALSE, formula = 'y ~ x') +
+#     scale_color_manual(
+#       name = "Treatment", 
+#       labels = c('Control', 'Selection'), 
+#       values = c('gray40', 'darkorange2')) +
+#     geom_jitter(#aes(alpha = selected), 
+#       width = 0.1) + 
+#         labs(x = "Passage Number", 
+#                   y = expression(log[10] ~ CH[4] ~ "(-k)")) +
+#   scale_x_continuous(labels = 1:5) +
+#     coord_fixed(ratio))
+
+# save(p_response, file = here(der_dir, 'response_plot.RData'))
+
+# ggsave(here(fig_dir, 'fig2.jpg'), p_response, width = 4, height = 3)
+# ggsave(here(fig_dir, 'fig2.svg'), p_response, width = 4, height = 3)
+# ggsave(here(fig_dir, 'fig2_cl.png'), p_response_cl, width = 4, height = 3)
+
+(p_response_bw <- ggplot(fluxes, aes(x = passage, y = log_ch4, fill = treat, linetype = treat)) +
+    stat_smooth(method = 'lm', se = FALSE, formula = 'y ~ x', color = 'black',
+                linewidth = 0.5) +
+    scale_fill_manual(
       name = "Treatment", 
-      labels = c('Neutral', 'Positive'), 
-      values = c('gray40', 'darkorange2')) +
-    geom_jitter(#aes(alpha = selected), 
-      width = 0.1) + 
+      labels = c('Control', 'Selection'), 
+      values = c('white', 'black')) +
+    scale_linetype_manual(
+      name = "Treatment", 
+      labels = c('Control', 'Selection'), 
+      values = c(3, 1)) +
+        geom_jitter(shape = 21, 
+      width = 0.1,
+      size = 1) + 
         labs(x = "Passage Number", 
@@
   scale_x_continuous(labels = 1:5) +
-    coord_fixed(ratio))
-
-# save(p_response, file = here(der_dir, 'response_plot.RData'))
-
-# ggsave(here(fig_dir, 'fig1.jpg'), p_response, width = 4, height = 3)
-# ggsave(here(fig_dir, 'fig1.svg'), p_response, width = 4, height = 3)
+    coord_fixed(ratio)
+    )
+
+ggsave(here(fig_dir, 'fig2.eps'), p_response_bw, width = 88, height = 65, units = 'mm', device = cairo_ps)
+# embed_fonts(here(fig_dir, 'fig2_bw.eps'))
+
+# (p_response_gs <- ggplot(fluxes, aes(x = passage, y = log_ch4, color = treat)) +
+#     stat_smooth(method = 'lm', se = FALSE, formula = 'y ~ x', linewidth = 0.5) +
+#     scale_color_manual(
+#       name = "Treatment", 
+#       labels = c('Control', 'Selection'), 
+#       values = c('gray50', 'black')) +
+#     geom_jitter(#aes(alpha = selected), 
+#       width = 0.1,
+#       size = 0.5) + 
+#         labs(x = "Passage Number", 
+#                   y = expression(log[10] ~ CH[4] ~ "(-k)")) +
+#   scale_x_continuous(labels = 1:5) +
+#     coord_fixed(ratio)
+# )
+# ggsave(here(fig_dir, 'fig2_gs.png'), p_response_gs, width = 88, height = 65, units = 'mm')
 
@@
            y = divergence)) +
-    stat_smooth(method = 'lm', 
-                se = FALSE, 
-                formula = 'y ~ x',
-                color = "black",
-                linewidth = 0.5) +
+    # stat_smooth(method = 'lm', 
+    #             se = FALSE, 
+    #             formula = 'y ~ x',
+    #             color = "black",
+    #             linewidth = 0.5) +
+        geom_line() +
     geom_point() + 
@@
                 linewidth = 0.5) +
-    geom_point() + 
+    geom_point(size = 1) + 
     labs(x = expression(atop("Cumulative Selection Differential", log[10] ~ CH[4] ~ "(-k)")), 
@@
 
-# ggsave(here(fig_dir, "fig2.png"), width = 8*.9, height = 4*.9)
-# ggsave(here(fig_dir, "fig2.svg"), width = 8*.9, height = 4*.9)
-
+# ggsave(here(fig_dir, "fig3.png"), width = 8*.9, height = 4*.9)
+# ggsave(here(fig_dir, "fig3.svg"), width = 8*.9, height = 4*.9)
+ggsave(here(fig_dir, "fig3.eps"),
+       p_rh,
+       width = 88,
+       height = 45, 
+       units = 'mm',
+       device = cairo_ps)
+# ggsave(here(fig_dir, "fig3.eps"), 
+#        p_rh,
+#        width = 5,
+#        height = 2.5)
 # save(rh_model, file = here(der_dir, 'herit_model.RData'))
@@
             .groups = "drop")
-p_beta <-
-  nmds_data %>%   
-  ggplot(aes(x = NMDS1, y = NMDS2, shape = passage, color = treat)) +
-  geom_point() +
-  stat_ellipse(show.legend = FALSE) +
-  coord_fixed() +
-  scale_shape_discrete(name = "Passage", labels = c(2, 5)) +
-  scale_color_manual(name = "Treatment",
-                       labels = c('Neutral', 'Positive'),
-                       values = c('gray40', 'darkorange2')) #+
+# p_beta_cl <-
+#   nmds_data %>%   
+#   ggplot(aes(x = NMDS1, y = NMDS2, shape = passage, color = treat)) +
+#   geom_point() +
+#   stat_ellipse(show.legend = FALSE) +
+#   coord_fixed() +
+#   scale_shape_discrete(name = "Passage", labels = c(2, 5)) +
+#   scale_color_manual(name = "Treatment",
+#                        labels = c('Control', 'Selection'),
+#                        values = c('gray40', 'darkorange2')) #+
   # theme_bw() +
@@
   #   
-pcoa <- cmdscale(rare_dist_bray, eig = TRUE, add = TRUE)
-round(pcoa$eig*100/sum(pcoa$eig),1)
-pcoa_per_exp <- 100 * eigenvals(pcoa) / sum(eigenvals(pcoa))
-
-ratio = pcoa_per_exp[2]/pcoa_per_exp[1]
-
-pcoa_scores <-
-  scores(pcoa, display = "sites") %>% 
-  as_tibble(rownames = "sample")
-
-pcoa_data <-
-  aes_data %>% 
-  select(sample, passage, treat) %>% 
-  distinct %>% 
-  inner_join(pcoa_scores, by = "sample")
-# p_beta <-
-  pcoa_data %>%   
-  ggplot(aes(x = Dim1, y = Dim2, shape = passage, color = treat)) +
-  geom_point() +
-  stat_ellipse(show.legend = FALSE) +
-  coord_fixed(ratio) +
-  scale_shape_discrete(name = "Passage", labels = c(2, 5)) +
-  scale_color_manual(name = "Treatment",
-                       labels = c('Neutral', 'Positive'),
-                       values = c('gray40', 'darkorange2')) #+
+# pcoa <- cmdscale(rare_dist_bray, eig = TRUE, add = TRUE)
+# round(pcoa$eig*100/sum(pcoa$eig),1)
+# pcoa_per_exp <- 100 * eigenvals(pcoa) / sum(eigenvals(pcoa))
+# 
+# ratio = pcoa_per_exp[2]/pcoa_per_exp[1]
+# 
+# pcoa_scores <-
+#   scores(pcoa, display = "sites") %>%
+#   as_tibble(rownames = "sample")
+# 
+# pcoa_data <-
+#   aes_data %>%
+#   select(sample, passage, treat) %>%
+#   distinct %>%
+#   inner_join(pcoa_scores, by = "sample")
+# 
+# # p_beta <-
+#   pcoa_data %>%
+#   ggplot(aes(x = Dim1, y = Dim2, shape = passage, color = treat)) +
+#   geom_point() +
+#   stat_ellipse(show.legend = FALSE) +
+#   coord_fixed(ratio) +
+#   scale_shape_discrete(name = "Passage", labels = c(2, 5)) +
+#   scale_color_manual(name = "Treatment",
+#                        labels = c('Neutral', 'Positive'),
+#                        values = c('gray40', 'darkorange2')) #+
   # theme_bw() +
@@
 
-# ggsave(here(fig_dir, 'fig3.png'), p_beta, height = 2.5, width = 4) 
-# ggsave(here(fig_dir, 'fig3.svg'), p_beta, height = 2.5, width = 4)
+# ggsave(here(fig_dir, 'fig4.png'), p_beta, height = 2.5, width = 4) 
+# ggsave(here(fig_dir, 'fig4.svg'), p_beta, height = 2.5, width = 4) 
+# ggsave(here(fig_dir, 'fig4_cl.png'), p_beta_cl, height = 2.5, width = 4)
+
+(p_beta_bw <-
+  nmds_data %>%   
+  ggplot(aes(x = NMDS1, y = NMDS2, shape = passage, fill = treat, 
+             # linetype = treat
+             )) +
+  geom_point() +
+  # stat_ellipse(
+  #     # show.legend = FALSE
+  #     ) +
+  #   scale_linetype_manual(
+  #     name = "Treatment", 
+  #     labels = c('Control', 'Selection'), 
+  #     values = c(3, 1)) +
+  coord_fixed() +
+  scale_shape_manual(name = "Passage", 
+                     labels = c(2, 5),
+                     values = c(24, 21)) +
+  scale_fill_manual(name = "Treatment",
+                       labels = c('Control', 'Selection'),
+                       values = c('white', 'black')) +
+  guides(fill=guide_legend(override.aes = list(shape=21)))
+  # scale_color_manual(name = "Treatment",
+  #                      labels = c('Neutral', 'Positive'),
+  #                      values = c('white', 'black')) #+
+  # theme_bw() +
+  # theme(
+  #   plot.background = element_blank(),
+  #   panel.grid.major = element_blank(),
+  #   panel.grid.minor = element_blank())
+  #   
+)
+ggsave(here(fig_dir, 'fig4.eps'), 
+       p_beta_bw, 
+       height = 60,
+       width = 180,
+       # height = 60,
+       # width = 88,
+       units = 'mm', 
+       device = cairo_ps
+       )
+
+# (p_beta_gs <-
+#   nmds_data %>%   
+#   ggplot(aes(x = NMDS1, y = NMDS2, shape = passage, color = treat)) +
+#   geom_point() +
+#   # stat_ellipse(
+#   #     # show.legend = FALSE
+#   #     ) +
+#   coord_fixed() +
+#   scale_shape_manual(name = "Passage",
+#                      labels = c(2, 5),
+#                      values = c(17, 19)) +
+#   scale_color_manual(name = "Treatment",
+#                        labels = c('Control', 'Selection'),
+#                        values = c('gray50', 'black')) +
+#   guides(fill=guide_legend(override.aes = list(shape=21)))
+# )
+# ggsave(here(fig_dir, 'fig4_gs.png'), p_beta_gs, height = 2.5*1.5, width = 4*1.5)
 
@@
   full_join(all_iqm, by = "Family")
-(p <-
+# (p_log10_cl <-
+# 	all_taxa %>%
+# 	ggplot(aes(x = rel_abund, y = Family, color = treat)) +
+# 	geom_point(position=position_jitterdodge()) +
+# 	geom_pointrange(aes(x = median, xmax = third_quartile, xmin = first_quartile, fill = treat), 
+# 			data = all_iqm, 
+# 			position = position_dodge(width = 0.75), 
+# 			shape = 21, 
+# 			color = "black") +
+# 	theme_bw() +
+# 	theme(axis.text.y = element_markdown()) +
+# 	labs(x = "Relative Abundance (%)", y = NULL) +
+# 	scale_color_manual(name = "Treatment",
+# 			   labels = c('Control', 'Selection'),
+# 			   values = c('gray40', 'darkorange2')) +
+# 	scale_fill_manual(name = "Treatment",
+# 			   labels = c('Control', 'Selection'),
+# 			   values = c('gray40', 'darkorange2')) +
+#     facet_grid(direction ~ ., scale = "free_y", space = "free")  + 
+# theme(
+#   strip.background = element_blank(),
+#   strip.text.y = element_blank()
+# ) +
+# 	scale_x_log10(breaks = c(0.001, 0.01, 0.1, 1),
+# 		      labels = c(0.001, 0.01, 0.1, 1)))
+
+# save(p_log10, da_fams, file = here(der_dir, "da_fig.RData"))
+
+# ggsave(here(fig_dir, 'fig5.jpg'), p_log10, width = 5, height = 5)
+# ggsave(here(fig_dir, 'fig5.svg'), p_log10, width = 5, height = 5)
+# ggsave(here(fig_dir, 'fig5_cl.png'), p_log10_cl, width = 5, height = 5)
+(p_log10_bw <-
 	all_taxa %>%
-	ggplot(aes(x = rel_abund, y = Family, color = treat)) +
-	geom_point(position=position_jitterdodge()) +
-	geom_pointrange(aes(x = median, xmax = third_quartile, xmin = first_quartile, fill = treat), 
-			data = all_iqm, 
-			position = position_dodge(width = 0.75), 
-			shape = 21, 
-			color = "black") +
-	theme_bw() +
+	ggplot(aes(x = rel_abund, y = Family, fill = treat)) +
+	geom_point(position=position_jitterdodge(jitter.width = 0.1,
+	                                         dodge.width = 0.4), shape = 21, size = 1) +
+        theme_bw(base_size = 7) +
+            theme(panel.border = element_rect(fill = NA, linewidth = 1.0625),
+                  axis.line = element_blank(),
+                  text = element_text(family = 'Arial')) +
+	# geom_jitter(shape = 21, height = 0.1) +
+	# geom_pointrange(aes(x = median, xmax = third_quartile, xmin = first_quartile, color = treat), 
+	# 		data = all_iqm, 
+	# 		position = position_dodge(width = 0.75), 
+	# 		shape = 21, 
+	# 		# color = "black"
+	# 		) +
+	# theme_bw() +
 	theme(axis.text.y = element_markdown()) +
 	labs(x = "Relative Abundance (%)", y = NULL) +
-	scale_color_manual(name = "Treatment",
-			   labels = c('Neutral', 'Positive'),
-			   values = c('gray40', 'darkorange2')) +
+	# scale_color_manual(name = "Treatment",
+	# 		   labels = c('Control', 'Selection'),
+	# 		   values = c('black', 'gray')) +
 	scale_fill_manual(name = "Treatment",
-			   labels = c('Neutral', 'Positive'),
-			   values = c('gray40', 'darkorange2')) +
-    facet_grid(direction ~ ., scale = "free_y", space = "free")  + 
+			   labels = c('Control', 'Selection'),
+			   values = c('white', 'black')) +
+    facet_grid(direction ~ ., 
+               scale = "free_y",
+               space = "free")  + 
 theme(
   strip.background = element_blank(),
-  strip.text.y = element_blank()
-))
-
-(p_log10 <- 
-	p +
+  strip.text.y = element_blank()) +
 	scale_x_log10(breaks = c(0.001, 0.01, 0.1, 1),
-		      labels = c(0.001, 0.01, 0.1, 1)))
-
-# save(p_log10, da_fams, file = here(der_dir, "da_fig.RData"))
-
-# ggsave(here(fig_dir, 'fig4.jpg'), p_log10, width = 5, height = 5) 
-# ggsave(here(fig_dir, 'fig4.svg'), p_log10, width = 5, height = 5)
+		      labels = c(0.001, 0.01, 0.1, 1))
+    )
+ggsave(here(fig_dir, 'fig5.eps'), 
+       p_log10_bw, 
+       width = 100, 
+       height = 88,
+       units = 'mm',
+       device = cairo_ps
+       )
+
+# (p_log10_gs <-
+# 	all_taxa %>%
+# 	ggplot(aes(x = rel_abund, y = Family, color = treat)) +
+# 	# geom_point(position=position_jitterdodge(), shape = 21) +
+# 	geom_jitter(#shape = 21, 
+# 	            height = 0.1) +
+# 	# geom_pointrange(aes(x = median, xmax = third_quartile, xmin = first_quartile, color = treat), 
+# 	# 		data = all_iqm, 
+# 	# 		position = position_dodge(width = 0.75), 
+# 	# 		shape = 21, 
+# 	# 		# color = "black"
+# 	# 		) +
+# 	theme_bw() +
+# 	theme(axis.text.y = element_markdown()) +
+# 	labs(x = "Relative Abundance (%)", y = NULL) +
+# 	# scale_color_manual(name = "Treatment",
+# 	# 		   labels = c('Control', 'Selection'),
+# 	# 		   values = c('black', 'gray')) +
+# 	scale_color_manual(name = "Treatment",
+# 			   labels = c('Control', 'Selection'),
+# 			   values = c('gray50', 'black')) +
+#     facet_grid(direction ~ ., scale = "free_y", space = "free")  + 
+# theme(
+#   strip.background = element_blank(),
+#   strip.text.y = element_blank()
+# ) +
+# 	scale_x_log10(breaks = c(0.001, 0.01, 0.1, 1),
+# 		      labels = c(0.001, 0.01, 0.1, 1)))
+# ggsave(here(fig_dir, 'fig5_gs.eps'), p_log10_gs, width = 5, height = 5)
 
@@
 
-p_effect <-
+(p_effect <-
 	all_taxa %>%
@@
 	scale_color_manual(name = "Treatment",
-			   labels = c('Neutral', 'Positive'),
+			   labels = c('Control', 'Selection'),
 			   values = c('gray40', 'darkorange2')) +
 	scale_fill_manual(name = "Treatment",
-			   labels = c('Neutral', 'Positive'),
+			   labels = c('Control', 'Selection'),
 			   values = c('gray40', 'darkorange2'))
-p_effect
+    )
 
@@
 
-save(p_response, herit_plot, p_beta, p_log10, file = here(der_dir, 'plots.RData'))
+# save(p_response, herit_plot, p_beta, p_log10, file = here(der_dir, 'plots.RData'))
 save(response_model, compare_response_mods, per_slope, rh_model, kt, wt, rare_median, s5, min_seqs, tidy_permanova, dbrda_aov, dbrda_percent_explained, da_fams, file = here(der_dir, 'models.RData'))

### 2023-02-27 R/functions.R

@@ -10,7 +10,8 @@ fig_dir <- 'analysis/figs/'      # Figures directory
 
 theme_set(theme_classic() +
             theme(panel.border = element_rect(fill = NA, linewidth = 1.0625),
-                  axis.line = element_blank()))
+                  axis.line = element_blank(),
+                  ))
 
 plot_flux <- function(fluxes, passage, estimate, ratio = NULL, log10 = FALSE) {
   p <- ggplot(fluxes, aes(x = {{passage}}, y = {{estimate}}, color = treat)) +

### 2023-03-03 analysis/analysis.Rmd
incorporate final edits

@@
 
-# save(p_response, herit_plot, p_beta, p_log10, file = here(der_dir, 'plots.RData'))
+#save(p_response, herit_plot, p_beta, p_log10, file = here(der_dir, 'plots.RData'))
 save(response_model, compare_response_mods, per_slope, rh_model, kt, wt, rare_median, s5, min_seqs, tidy_permanova, dbrda_aov, dbrda_percent_explained, da_fams, file = here(der_dir, 'models.RData'))

### 2023-03-03 manuscript/manuscript.Rmd
incorporate final edits

@@
 
-{title max characters 150; current = 106 w/spaces}  
 ^1^ Institute of Ecology and Evolution, University of Oregon, Eugene, OR, USA  
@@
 library(morris)
-# library(phyloseq)
 library(ggpubr)
@@
 library(corncob)
-# library(ggthemes)
 library(captioner)
@@
 ```{r load-data}
-heritability <- read_tsv(paste0(der_dir, 'heritability.tsv'))
-fluxes <- read_tsv(paste0(der_dir, 'fluxes.tsv'))
+#heritability <- read_tsv(paste0(der_dir, 'heritability.tsv'))
+#fluxes <- read_tsv(paste0(der_dir, 'fluxes.tsv'))
 
@@
 
-#load(file = paste0(der_dir, 'plots.RData'))
 load(file = paste0(der_dir, 'models.RData'))
-
-# # CH4 models
-# load(file = paste0(der_dir, 'response_model.RData'))
-# load(file = paste0(der_dir, 'response_plot.RData'))
-# load(file = paste0(der_dir, 'herit_model.RData'))
-# load(file = paste0(der_dir, 'herit_plot.RData'))
-# 
-# # 16S models
-# load(file = paste0(der_dir, 'richness_models.RData'))
-# load(file = paste0(der_dir, 'phyloseq.RData'))
-# load(file = paste0(der_dir, 'beta_model.RData'))
-# load(file = paste0(der_dir, 'beta_fig.RData'))
-# load(file = paste0(der_dir, 'da_corncob.RData'))
-# load(file = paste0(der_dir, 'da_fig.RData'))
-# load(file = paste0(der_dir, 'dbrda.RData'))
 
@@
 
-{max 250 words; current = 249}  
 Microbiomes mediate important ecosystem functions, yet it has proven difficult
@@
 ecosystems. We performed artificial ecosystem selection on replicate soil
-microbiomes to select for increased soil methane oxidation rate. We observed a
+microbiomes to select for increased soil CH~4~ oxidation rate. We observed a
 strong response to selection with a `r myround(per_slope, 1)`\% increase in
-methane oxidation rate per passage. We
+CH~4~ oxidation rate per passage. We
 estimated that `r myround(tidy(rh_model)[[2, 2]]*100, 1)`\% of the variation in
-methane oxidation
+CH~4~ oxidation
 rate in these soils can be attributed to microbiome variation. We also found
-that selection did not enrich for known methane oxidizers; instead, `r length(da_fams)` families not
-known to oxidize methane, including *Fimbriimonadaceae*, *Cytophagaceae*, and
+that selection did not enrich for known CH~4~ oxidizers; instead, `r length(da_fams)` families not
+known to oxidize CH~4~, including *Fimbriimonadaceae*, *Cytophagaceae*, and
 *Diplorickettsiaceae*, were enriched by selection. This result is in contrast to the typical assumption that
@@
 ecosystem functions.
-
-{max text 5000 words; current text = 4549}
 
@@
 
-(`r figures("concept", display = "cite")`)
+Altering ecosystem functions via microbiome manipulations requires that the
+microbiome contributes to variation in ecosystem function independent of other
+drivers of ecosystem variation, such as variation in environmental conditions.
+This is because the drivers of variation in ecosystem functions can interact in
+complicated ways (`r figures("concept", display = "cite")`). Variation in the
+microbiome can contribute directly to variation in ecosystem function, for
+example, if a microbial population is replaced by one with a greater enzyme
+efficiency. In addition, environmental conditions can contribute indirectly to
+ecosystem function via covariance with the microbiome, for example, by providing
+conditions that select for microbial groups that in turn alter the rates of
+ecosystem functions. In this scenario, identifying the change in microbial
+community composition without adequately controlling for the environmental
+conditions would incorrectly attribute the change in ecosystem function to the
+microbiome when it is ultimately an indirect effect of the environment.
+Determining the independent contribution of microbiome variation to ecosystem
+function is crucial because if microbiome composition is driven primarily by
+environmental conditions, then introducing a desirable taxon through microbiome
+manipulation without altering the environment will likely be unsuccessful at
+shifting the targeted ecosystem function.
 
@@
 used to estimate the degree to which an ecosystem function varies with
-microbiome composition: comparative and manipulative. Comparative
-studies sample natural variation in an ecosystem function across different
-habitats and simultaneously measure variation in community composition.
-Investigators can then correlate ecosystem function with aspects of community
-composition while attempting to control for environmental variation. These
-approaches have
-documented important relationships between microbiomes and ecosystem
-functions. For example, a meta-analysis of these studies observed a small but
-significant contribution of the microbiome to variation in ecosystem
-function after controlling for environmental variation [@graham2016]. In
-addition, studies focusing on the correlation between the rate of an ecosystem
-function and the abundance of an associated marker gene (i.e., a gene that codes
-for a protein assumed to be involved in the ecosystem function) sometimes observe a
+microbiome composition: comparative and manipulative. Comparative studies sample
+natural variation in an ecosystem function across different habitats and
+simultaneously measure variation in community composition.  Investigators can
+then correlate ecosystem function with aspects of community composition while
+attempting to control for environmental variation. These approaches have
+documented important relationships between microbiomes and ecosystem functions.
+For example, a meta-analysis of these studies observed a small but significant
+contribution of the microbiome to variation in ecosystem function after
+controlling for environmental variation [@graham2016]. In addition, studies
+focusing on the correlation between the rate of an ecosystem function and the
+abundance of an associated marker gene (i.e., a gene that codes for a protein
+assumed to be involved in the ecosystem function) sometimes observe a
 significant correlation, though this relationship is rare and contingent upon
 both the function and the ecosystem sampled [@rocca2015]. However, comparative
-studies come with some unique challenges and limitations. One issue is that
+studies come with unique challenges and limitations. One issue is that
 microbiome attributes tend to covary with the abiotic conditions within an
-environment, and whether and how the investigator controls for these variables
-can affect the conclusions of such a comparative study. In addition, it is
-difficult to know _a priori_ which environmental variables or community attributes
-to measure. Finally, while these approaches can establish a potential magnitude
-and direction for these relationships, it is often difficult to identify the
+environment, and it is difficult to control for these abiotic variables in order
+to identify the unique contribution of the microbiome to ecosystem function. In
+addition, it is difficult to know _a priori_ which environmental variables or
+community attributes to measure. Finally, while these approaches can establish a
+potential magnitude and direction for these relationships, it is often difficult
+to identify the
 taxa or genes that explain the connection between composition and function.
@@
 [@wagg2014; @wagg2019; @griffiths2001]. However, manipulating the microbiome
-directly is challenging. Each of the manipulative approaches applied confound
+directly is challenging, and manipulative approaches often confound
 community composition with other factors. For example, reciprocal transplant and
@@
 function. We used artificial ecosystem selection to select for microbiomes that
-performed a greater rate of ecosystem function
-[@swenson2000; @blouin2015; @panke-buisse2015]. We then tested whether variation
-in the microbiome contributed to variation in the rate of ecosystem function and
-identified microbiome attributes that might explain this relationship. There are
-several potential advantages to this approach for documenting
-microbiome-function relationships and for investigating the mechanisms
-underlying those relationships. By passaging microbiomes over multiple rounds
-using a common environment, we can weaken the covariance between microbes and
-the environment by repeatedly diluting abiotic effects. In addition, by
-inoculating our ecosystems with equally-sized subsamples, we eliminate the need
-to generate microbiome variation through methods that are confounded with
-biomass or cell size. Lastly, by comparing our artificially selected community
-to a control community resulting from random selection, we can both control for
-changes in the environment over time and identify genes or taxa that are
-associated with the ecosystem function under selection.
+performed a greater rate of ecosystem function [@swenson2000; @blouin2015;
+@panke-buisse2015]. We then tested whether variation in the microbiome
+contributed to variation in the rate of ecosystem function and identified
+microbiome attributes that might explain this relationship. There are several
+potential advantages to this approach for documenting the direct contribution of
+the microbiome to variation in ecosystem function and for investigating the
+mechanisms underlying those relationships. By repeated passaging of microbiomes
+in a common environment, we can weaken the covariance between microbes and the
+environment by repeatedly diluting the influence of variation in abiotic
+conditions. In addition, our approach eliminates the need to generate microbiome
+variation through methods that are confounded with biomass or cell size. Lastly,
+by comparing our artificially selected community to a control community
+resulting from random selection, we can both control for changes in the
+environment over time and identify genes or taxa that are associated with the
+ecosystem function under selection.
 
 We applied artificial ecosystem selection to soil microbiomes by selecting on
-soil methane oxidation rate. We chose this function because methane is a
-globally important greenhouse gas and methane oxidation by soil bacteria is the
-primary biological sink for atmospheric methane [@kirschke2013]. In addition,
-there is evidence that soil methane oxidation rate may vary with microbiome
+soil methane (CH~4~) oxidation rate. We chose this function because CH~4~ is a
+globally important greenhouse gas and CH~4~ oxidation by soil bacteria is the
+primary biological sink for atmospheric CH~4~ [@kirschke2013]. In addition,
+there is evidence that soil CH~4~ oxidation rate may vary with microbiome
 composition based on comparative studies in a variety of arctic and tropical
@@
 composition of the microbiome is more likely to be associated with the rate of
-methane oxidation than other broader or more shallowly conserved functions
+CH~4~ oxidation than other broader or more shallowly conserved functions
 [@schimel1998; @martiny2013].
 
-In this study, we used artificial ecosystem selection on methane oxidation rate
+In this study, we used artificial ecosystem selection on CH~4~ oxidation rate
 to address the following questions: Does variation in the relative abundance of
-microbial taxa contribute to variation in soil methane oxidation rate in our
-system? Which attributes of the microbiome are associated with variation in
-methane oxidation rate, and do these attributes match our assumptions about the
-factors that regulate methane oxidation rate in nature?
+microbial taxa contribute to variation in soil CH~4~ oxidation rate
+independent of the environment in our system? Which attributes of the microbiome
+are associated with variation in CH~4~ oxidation rate, and do these attributes
+match our assumptions about the factors that regulate CH~4~ oxidation rate in
+nature?
 
@@
 We performed an artificial ecosystem selection experiment 
-[*sensu* @swenson2000] by passaging replicate soil microbiomes. The trait we
-selected on was methane (CH~4~)-oxidation rate. This experiment had two selection lines
-with twelve jars each for a total of twenty-four jars per passage. One line was
-a positive selection line where the two or three jars with the highest CH~4~
-oxidation rate were homogenized to inoculate the next set of positive jars. The
-other line was a neutral selection line where an equal number of jars as the
-positive line were chosen at random to inoculate the next set of neutral jars.
-The number of jars chosen was based on the distribution of fluxes among the
-positive jars, i.e., we chose the top three jars unless only two jars had
+(*sensu* @swenson2000) by passaging replicate soil microbiomes. The trait we
+selected on was CH~4~ oxidation rate. This experiment had a selection line and a
+control line with twelve jars each for a total of twenty-four jars per passage.
+The selection line underwent positive selection where the two or three jars
+with the highest CH~4~ oxidation rate were homogenized to inoculate the next set
+of jars. The control line underwent random selection where an equal number
+of jars as the selection line were chosen at random to inoculate the next set of
+jars. The number of jars chosen was based on the distribution of fluxes among
+the positive jars, i.e., we chose the top three jars unless only two jars had
 considerably greater CH~4~ oxidation rates based on visual inspection of
@@
 treatment groups, twenty-four jars were created in an identical manner and then
-randomly assigned to either the positive or neutral selection line. Jars were
+randomly assigned to either the selection or control line. Jars were
 flushed and respiked with CH~4~ twice per week to maintain aerobic conditions
@@
 approximately four weeks. Methane oxidation rates were determined at the end of
-the incubation period. For the positive treatment, the three jars with the
-greatest CH~4~ oxidation rates were chosen to inoculate the next generation.
-For the neutral treatment, three jars were randomly selected to inoculate the
-next generation. For each treatment, the selected jars were homogenized and 5 g
-of the homogenized soil was used to inoculate the next set of jars, which
-represents a 10% subsampling. The next set of jars were created in an identical
-manner to the first generation with fresh autoclaved potting mix and the same
-moisture and CH~4~ content.
+the incubation period and selection was performed. For each treatment, the
+selected jars were homogenized and 5 g of the homogenized soil was used to
+inoculate the next set of jars, which represents a 10% subsampling. The next set
+of jars were created in an identical manner to the first generation with fresh
+autoclaved potting mix and the same moisture and CH~4~ content.
 
@@
 constant (k, units = d^−1^; i.e., dCH~4~/dt = k[CH~4~]) of the exponential
-decrease in methane. Oxidation rates are presented as the additive inverse of
+decrease in CH~4~ Oxidation rates are presented as the additive inverse of
 $k$ (i.e., $-k$) so that a more positive value represents a greater
-oxidation rate. The jars selected in passage two for the positive treatment had
+oxidation rate. The jars selected in passage two for the positive selection treatment had
 the lowest CH~4~ oxidation rate of the twelve jars due to a calculation error
@@
 whether there was a significant change in CH~4~ oxidation rate as a response to
-selection, we tested a difference in slopes between the positive and neutral
-selection lines. Residuals did not meet the assumptions of constant variance
+selection, we tested a difference in slopes between the selection and control
+lines. Residuals did not meet the assumptions of constant variance
 and normal distribution. Therefore, CH~4~ oxidation rates were log~10~
 transformed prior to analysis. First, we tested if there was a difference of
-slopes between the positive and neutral treatments based on the interaction
+slopes between the selection line and the control based on the interaction
 between passage and treatment. To test the interaction, we fit two nested models
@@
 
-We estimated the proportion of variation in methane oxidation rate due to
+We estimated the proportion of variation in CH~4~ oxidation rate due to
 variation in the microbiome as the regression of divergence between the positive
-and neutral treatment on the cumulative selection differential [@falconer1996].
-Divergence was calculated as the mean methane oxidation rate of the positive
-treatment minus the mean methane oxidation rate of the neutral treatment in each
+line and the control on the cumulative selection differential [@falconer1996].
+Divergence was calculated as the mean CH~4~ oxidation rate of the positive
+treatment minus the mean CH~4~ oxidation rate of the control in each
 passage. The selection differential was calculated as the difference between
@@
 rate (`r figures("response", display = "cite")`; difference of slopes:
-F~`r compare_response_mods[2,4]`,`r compare_response_mods[2,2]`~ = 
-`r myround(compare_response_mods[2,5], 2)`,
-p = `r myround(compare_response_mods[2,6], 2)`). At the start of the experiment,
-the Positive treatment had a mean CH~4~ oxidation rate that was
+F~`r compare_response_mods[[2,4]]`,`r compare_response_mods[[2,2]]`~ = 
+`r myround(compare_response_mods[[2,6]], 2)`,
+p = `r myround(compare_response_mods[[2,7]], 2)`). At the start of the experiment,
+the positive selection treatment had a mean CH~4~ oxidation rate that was
 `r percent_less(initial = tidy(response_model)[[1, 2]], final = tidy(response_model)[[3, 2]] + tidy(response_model)[[1, 2]])`\%
-lower than the Neutral treatment
+lower than the control
 `r print_lm(response_model, 3, 'difference of y-intercepts')`. 
-There was no change in CH~4~ oxidation rate of the Neutral treatment over the
-five passages `r print_lm(response_model, 2)`. By contrast, the Positive
+There was no change in CH~4~ oxidation rate in the control over the
+five passages `r print_lm(response_model, 2)`. By contrast, the selection
 treatment had a `r myround(per_slope, 1)`\% increase in CH~4~ oxidation rate per
@@
 
-Median ASV richness decreased from `r myround(rare_median[1, 2], 1)` (`r myround(rare_median[1, 3], 1)`) in Passage 2 to `r myround(rare_median[2, 2], 1)` (`r myround(rare_median[2, 3], 1)`) in Passage 5 (Kruskal-Wallace test: $\chi^2$ =
+Median ASV richness decreased from `r myround(rare_median[1, 2], 1)` (`r myround(rare_median[1, 3], 1)`) in passage 2 to `r myround(rare_median[2, 2], 1)` (`r myround(rare_median[2, 3], 1)`) in passage 5 (Kruskal-Wallace test: $\chi^2$ =
 `r myround(kt[1,1], 1)`, df = `r kt[1,3]`, p < 0.001; pairwise Wilcoxon test: p < 0.001). However, there was no
-difference in richness between the Positive and the Neutral treatment in
-Passage 2 or 5 (pairwise Wilcoxon test: p = `r myround(wt[1, 3], 1)`).  In addition, there was no
-correlation between richness and CH~4~ oxidation rate across the two treatments in Passage 5 (Spearman's rho = `r myround(s5[1,1], 1)`, p = `r myround(s5[1,3], 1)`).
+difference in richness between the selection treatment and the control in
+passage 2 or 5 (pairwise Wilcoxon test: p = `r myround(wt[1, 3], 1)`).  In addition, there was no
+correlation between richness and CH~4~ oxidation rate across the two treatments in passage 5 (Spearman's rho = `r myround(s5[1,1], 1)`, p = `r myround(s5[1,3], 1)`).
 
@@
 no difference in dispersion between treatments or passages (F~3,44~ = 0.91, p = 0.45). 
-Finally, CH~4~ oxidation rate was correlated with Bray-Curtis dissimilarity across both treatments in Passage 5 and explained `r myround(dbrda_percent_explained[1], 1)`\% of the variation in Bray-Curtis dissimilarity (dbRDA: F~`r dbrda_aov[1, 1]`,`r dbrda_aov[2, 1]`~ = `r myround(dbrda_aov[1, 3], 2)`, p = `r myround(dbrda_aov[1, 4], 3)`)
+Finally, CH~4~ oxidation rate was correlated with Bray-Curtis dissimilarity across both treatments in passage 5 and explained `r myround(dbrda_percent_explained[1], 1)`\% of the variation in Bray-Curtis dissimilarity (dbRDA: F~`r dbrda_aov[1, 1]`,`r dbrda_aov[2, 1]`~ = `r myround(dbrda_aov[1, 3], 2)`, p = `r myround(dbrda_aov[1, 4], 3)`)
 
@@
 to selection on soil CH~4~ oxidation rate, we tested the differential relative
-abundance of families in the Positive jars relative to the Neutral jars within
-Passage 5 using three methods and then plotted the taxa identified by all three methods. We identified `r length(da_fams)`
-families that were enriched or depleted in the Positive treatment relative to
-the Neutral treatment (`r figures("da", display = "cite")`).
+abundance of families in the selected jars relative to the control jars within
+passage 5 using three methods and then plotted the taxa identified by all three methods. We identified `r length(da_fams)`
+families that were enriched or depleted in the selection treatment relative to
+the control (`r figures("da", display = "cite")`).
 
-Overall, none of the families enriched in the Positive selection treatment
+Overall, none of the families enriched in the selection treatment
 contain known methanotrophs. Several taxa identified had a 
@@
 In addition, the *Puniceicoccaceae* is a member of the phylum Verrucomicrobia. The Verrucomicrobia is a diverse group that contain
-known methanotrophs as well as ammonia-oxidizing bacteria [@freitag2003].
-Other than these two groups, none of the other taxa enriched in the Positive treatment are known to be related to methanotrophs.
-Two groups in the Armatimonadales were enriched in the Positive treatment including the family *Fimbriimonadaceae* and an unclassified ASV from the order Armatimonadales [@im2012].
+known methanotrophs as well as ammonia-oxidizing bacteria [@freitag2003]. Other
+than these two groups, none of the other taxa enriched in the selection treatment
+are known to be related to methanotrophs. Two groups in the Armatimonadales were
+enriched in the selection treatment including the family *Fimbriimonadaceae* and
+an unclassified ASV from the order Armatimonadales [@im2012].
 *Cytophagaceae* was also enriched in
-the Positive treatment and contains a number of mainly aerobic heterotrophs
+the selection treatment and contains a number of mainly aerobic heterotrophs
 that can digest a variety of macromolecules [@mcbride2014]. 
@@
 Aggregating reads at the family level, neither family was differentially abundant between the two treatments.
-However, aggregated at the genus level, a group of unclassified genera in the *Beijerinckiaceae* were depleted in the Positive treatment and the genus
+However, aggregated at the genus level, a group of unclassified genera in the *Beijerinckiaceae* were depleted in the selection treatment and the genus
 *Rhodoblastus*, a member of the *Beijerinckiaceae*, was enriched in the
-Positive treatment.  While many *Beijerinckiaceae* are methanotrophs, several
+selection treatment.  While many *Beijerinckiaceae* are methanotrophs, several
 taxa in this family have lost the ability to oxidize CH~4~ and it appears
@@
 on methanol [@tamas2014; @dedysh2016]. Based on this analysis,
-it appears that no methanotrophs were enriched in the Positive treatment.
+it appears that no methanotrophs were enriched in the positive selection treatment.
 
@@
 in microbiome composition to variation in the rate of an ecosystem function,
-methane oxidation in soil, independent of environmental variation. Understanding
+CH~4~ oxidation in soil, independent of environmental variation. Understanding
 how and to what degree microbiome variation contributes to variation in
@@
 (`r figures("concept", display = "cite")`);
-for example, environmental variation can directly contribute by providing
+for example, environmental variation can indirectly contribute by providing
 conditions that select for microbial groups that in turn alter the rates of
@@
 
-In our study, we observed an increase in methane oxidation rate in the
+In our study, we observed an increase in CH~4~ oxidation rate in the
 selection treatment relative to the control, which demonstrates that there was a
@@
 conclude that variation in the microbiome contributes to variation in the
-methane oxidation rate independent of the environment. This suggests that
+CH~4~ oxidation rate independent of the environment. This suggests that
 microbiome manipulations could be an effective approach for altering the rate of
-methane oxidation in this soil, and that the artificial selection approach may
+CH~4~ oxidation in this soil, and that the artificial selection approach may
 be useful in determining the potential for microbiome manipulations for other
@@
 calculate the response to selection as the difference between two successive
-passages in their mean methane oxidation rate. We will denote this as $R$. We
-can also calculate the strength of selection as the difference in mean methane
+passages in their mean CH~4~ oxidation rate. We will denote this as $R$. We
+can also calculate the strength of selection as the difference in mean CH~4~
 oxidation rate between the twelve jars in one generation and the three jars chosen
@@
 would conclude that 100% of the variation is due to variation in the microbiome.
-Likewise, if recipients do not resemble the donors in their mean methane
+Likewise, if recipients do not resemble the donors in their mean CH~4~
 oxidation rate and simply wander randomly, then we would conclude that all of
@@
 `r myround(tidy(rh_model)[[2, 2]]*100, 1)`\% of the
-variation we observed in the rate of methane oxidation.  This suggests that
+variation we observed in the rate of CH~4~ oxidation.  This suggests that
 there is substantial potential for altering this ecosystem function through
@@
 “environmental microbiability” will be different for other ecosystem functions
-in this soil and for methane oxidation in other soils. However, our experiment
+in this soil and for CH~4~ oxidation in other soils. However, our experiment
 demonstrates that this relationship is measurable and provides an example of how
@@
 Given that variation in the microbiome is driving substantial variation in
-methane oxidation rate in our system, we next wanted to determine which aspects of the
+CH~4~ oxidation rate in our system, we next wanted to determine which aspects of the
 microbiome best explain this contribution of microbiome variation to
-methane oxidation rate. There are three inter-related ways that microbiomes could have responded
+CH~4~ oxidation rate. There are three inter-related ways that microbiomes could have responded
 to selection in this experiment: gain or loss of taxa, changes in the
@@
 Richness at the ASV level did not vary between the two treatments and there
-were relatively few taxa gained or lost in the Positive selection treatment and
+were relatively few taxa gained or lost in the selection treatment and
 none of these were prevalent across the 12 jars in passage 5. Therefore, the
-gain or loss of species is unlikely to explain the increase in methane
+gain or loss of species is unlikely to explain the increase in CH~4~
 oxidation rate. However, we found that Bray-Curtis dissimilarity was greater
 between the two treatments in Passage 5 than within each treatment and was
-correlated with methane oxidation rate, which suggests that changes in the
+correlated with CH~4~ oxidation rate, which suggests that changes in the
 relative abundance of taxa could explain the response to selection.
 
-Even though we observed an increase in methane oxidation rate in the Positive
+Even though we observed an increase in CH~4~ oxidation rate in the selection
 treatment and a difference in composition between the two treatments, we did
 not observe an increase in the relative abundance of methanotrophs. This was
-surprising given that methane consumption is not a common trait among microbes
+surprising given that CH~4~ consumption is not a common trait among microbes
 and that it is often assumed that the rate of an ecosystem function is limited
 by the final enzymatic step in the underlying metabolic pathway [@rocca2015]. In certain
-ecosystems, methane production and consumption are correlated with the
+ecosystems, CH~4~ production and consumption are correlated with the
 abundance of methanogens and methanotrophs as estimated from marker genes
 [@meyer2020; @meyer2020a]. However, our results suggest that
-ecosystem-scale methane oxidation rates can be altered by non-methanotrophs,
+ecosystem-scale CH~4~ oxidation rates can be altered by non-methanotrophs,
 perhaps through ecological interactions with methanotrophic species. This
@@
 There is increasing interest in using artificial selection for understanding and
-manipulating the microbiomes associated with plants and animals (aka “microbiome
+manipulating the microbiomes associated with plants and animals (a.k.a., “microbiome
 breeding”; @mueller2022). Our study demonstrates that artificial ecosystem
@@
 Fellowship Program (grant no. DGE 1255832) and the ARCS Foundation Florence and
-Mike Nudelman Scholarship.
+Mike Nudelman Scholarship. Figure 1 was created with BioRender.com.
 
@@
 
-`r figures("concept", "Variation in the rate of an ecosystem function is potentially the result of at least three components: variation in the abiotic environmental conditions, variation in microbiome composition, and the covariance between microbiomes and the environment.
-It is difficult with traditional study designs to isolate the direct effect of the microbiome from the effect of the environment via covariance with the microbiome.
+`r figures("concept", "Variation in the rate of an ecosystem function is the result of at least three components: variation in the abiotic environmental conditions, variation in microbiome composition, and the covariance between microbiomes and the environment.
+The arrows represent causal relationships between the components.
+It is important to isolate the direct effect of the microbiome from the effect of the environment via covariance with the microbiome.
 Here, we attempt to isolate the effect of the microbiome through artificial selection on microbiome composition.
-For simplicity, we omitted the reverse arrows as well as the interaction between microbiome and the environment, though these relationships may also exist.")`
+For simplicity, we omitted the reverse arrows as well as the interactions, though these relationships may also exist.")`
 
@@
 The y-axis is CH~4~ oxidation rate as the log~10~ of the additive inverse of the first-order exponential decay constant $k$ (i.e., $-k$) with units day^-1^ so that a more positive value represents a higher CH~4~ oxidation rate.
-There was a significant difference of slopes between the positive selection treatment and the control (F~", compare_response_mods[2,4], ",", compare_response_mods[2,2], "~ = ", myround(compare_response_mods[2,5], 2), ", p = ", myround(compare_response_mods[2,6], 2), ")."))`
+Closed points are the positive selection treatment and open points are the control.
+The solid line is the regression for the positive selection treatment and the dotted line is the regression for the control.
+There was a significant difference of slopes between the positive selection treatment and the control (F~", compare_response_mods[[2,4]], ",", compare_response_mods[[2,2]], "~ = ", myround(compare_response_mods[[2,6]], 2), ", p = ", myround(compare_response_mods[[2,7]], 2), ")."))`
 
@@
 
-`r figures("beta", "Non-metric multidimensional scaling plot of beta diversity for all jars. Dissimilarities are rarefied Bray-Curtis dissimilarity averaged over 99 subsamples.")`
+`r figures("beta", "Non-metric multidimensional scaling plot of beta diversity for all jars.
+Dissimilarities are rarefied Bray-Curtis dissimilarity averaged over 100 subsamples.
+Closed points are the positive selection treatment and open points are the control.
+Triangles are passage 2 and circles are passage 5.")`
 

### 2023-03-03 analysis/scripts/compare-rare-methods.R
remove extraneous scripts

@@ -1,197 +1,0 @@
-library(vegan)
-library(breakaway)
-library(magrittr)
-library(zCompositions)
-library(tidyverse)
-source('R/functions.R')
-
-set.seed(19760620)
-
-aes_data <- read_tsv('analysis/output/community_data.tsv')
-
-# Look at distribution of sequencing depths
-aes_data %>% 
-  group_by(sample) %>% 
-  summarize(total = sum(count)) %>% 
-  ggplot(aes(x = total)) +
-  geom_histogram()
-
-
-seq_range <- 
-  aes_data %>%
-  group_by(sample) %>% 
-  summarize(total = sum(count)) %>% 
-  summarize(seq_range = range(total)) %$%
-  seq_range
-
-# Generate a null dataset
-randomize_asv_table <- function(x) {
-    x %>%
-        uncount(count) %>%
-        mutate(rand_asv = sample(asv)) %>%
-        dplyr::select(-asv) %>%
-        count(sample, rand_asv, name="value")
-}
-rand_aes <- randomize_asv_table(aes_data)
-
-#######################
-# Alpha diversity
-#####################
-
-
-analysis <- rand_aes %>%
-  count(sample, value) %>%
-  nest(data = -sample) %>%
-  mutate(n_seqs = map_dbl(data, ~sum(.x$value * .x$n)),
-         sobs = map_dbl(data, ~sum(.x$n)),
-         rare = map_dbl(data, ~rarefy(rep(.x$value, .x$n), sample=seq_range[1])),
-         chao = map_dbl(data, ~get_chao(.x)),
-         ba = map(data, ~get_breakaway(.x))) %>%
-  select(-data) %>%
-  unnest(ba)
-
-
-analysis %>%
-  select(sample, n_seqs, sobs, chao, rare, est) %>%
-  pivot_longer(-c(sample, n_seqs)) %>%
-  ggplot(aes(x=n_seqs, y=value, color=name)) +
-  geom_point() +
-  geom_smooth(se=FALSE) +
-  coord_cartesian(ylim=c(0, 40000)) +
-  theme_classic() +
-  labs(title = "AES", x = "Sequencing Depth", y = "Alpha Diversity")
-
-analysis %>%
-  select(sample, n_seqs, sobs, chao, rare, est) %>%
-  pivot_longer(-c(sample, n_seqs)) %>%
[…]

### 2023-06-23 analysis/analysis.Rmd
update with edits for reviewers

@@
 library(ANCOMBC)
-library(tidyverse)
 library(here)
@@
 library(extrafont)
+library(tidyverse)
 
@@
 
-# (p_response_cl <- ggplot(fluxes, aes(x = passage, y = log_ch4, color = treat)) +
-#     stat_smooth(method = 'lm', se = FALSE, formula = 'y ~ x') +
-#     scale_color_manual(
-#       name = "Treatment", 
-#       labels = c('Control', 'Selection'), 
-#       values = c('gray40', 'darkorange2')) +
-#     geom_jitter(#aes(alpha = selected), 
-#       width = 0.1) + 
-#         labs(x = "Passage Number", 
-#                   y = expression(log[10] ~ CH[4] ~ "(-k)")) +
-#   scale_x_continuous(labels = 1:5) +
-#     coord_fixed(ratio))
+(p_response_cl <- ggplot(fluxes, aes(x = passage, y = log_ch4, color = treat)) +
+    stat_smooth(method = 'lm', se = FALSE, formula = 'y ~ x',
+                linewidth = 0.5) +
+    scale_color_manual(
+      name = "Treatment",
+      labels = c('Control', 'Selection'),
+      values = c('gray40', 'darkorange2')) +
+    geom_jitter(#aes(alpha = selected),
+      width = 0.1,
+      size = 0.5) +
+        labs(x = "Passage Number",
+                  y = expression(log[10] ~ CH[4] ~ "(-k)")) +
+  scale_x_continuous(labels = 1:5) +
+    coord_fixed(ratio))
 
@@
 
-# ggsave(here(fig_dir, 'fig2.jpg'), p_response, width = 4, height = 3)
+ggsave(here(fig_dir, 'fig2.png'), p_response_cl, width = 88, height = 65, units = 'mm')
 # ggsave(here(fig_dir, 'fig2.svg'), p_response, width = 4, height = 3)
@@
 
-ggsave(here(fig_dir, 'fig2.eps'), p_response_bw, width = 88, height = 65, units = 'mm', device = cairo_ps)
+# ggsave(here(fig_dir, 'fig2.eps'), p_response_bw, width = 88, height = 65, units = 'mm', device = cairo_ps)
 # embed_fonts(here(fig_dir, 'fig2_bw.eps'))
@@
 # ggsave(here(fig_dir, "fig3.svg"), width = 8*.9, height = 4*.9)
-ggsave(here(fig_dir, "fig3.eps"),
-       p_rh,
-       width = 88,
-       height = 45, 
-       units = 'mm',
-       device = cairo_ps)
+# ggsave(here(fig_dir, "fig3.eps"),
+#        p_rh,
+#        width = 88,
+#        height = 45, 
+#        units = 'mm',
+#        device = cairo_ps)
 # ggsave(here(fig_dir, "fig3.eps"), 
@@
             .groups = "drop")
-# p_beta_cl <-
-#   nmds_data %>%   
-#   ggplot(aes(x = NMDS1, y = NMDS2, shape = passage, color = treat)) +
-#   geom_point() +
+(p_beta_cl <-
+  nmds_data %>%
+  ggplot(aes(x = NMDS1, y = NMDS2, shape = passage, color = treat)) +
+  geom_point() +
   # stat_ellipse(show.legend = FALSE) +
-#   coord_fixed() +
-#   scale_shape_discrete(name = "Passage", labels = c(2, 5)) +
-#   scale_color_manual(name = "Treatment",
-#                        labels = c('Control', 'Selection'),
-#                        values = c('gray40', 'darkorange2')) #+
+  coord_fixed() +
+  scale_shape_discrete(name = "Passage", labels = c(2, 5)) +
+  scale_color_manual(name = "Treatment",
+                       labels = c('Control', 'Selection'),
+                       values = c('gray40', 'darkorange2'))) #+
 # theme_bw() +
@@
 #   panel.grid.minor = element_blank())
-  #   
+
 # pcoa <- cmdscale(rare_dist_bray, eig = TRUE, add = TRUE)
@@
 
-# ggsave(here(fig_dir, 'fig4.png'), p_beta, height = 2.5, width = 4) 
+ggsave(here(fig_dir, 'fig3.png'), p_beta_cl,
+       height = 60,
+       width = 180,
+       units = 'mm')
 # ggsave(here(fig_dir, 'fig4.svg'), p_beta, height = 2.5, width = 4) 
@@
 )
-ggsave(here(fig_dir, 'fig4.eps'), 
-       p_beta_bw, 
-       height = 60,
-       width = 180,
+# ggsave(here(fig_dir, 'fig3.eps'), 
+#        p_beta_cl, 
 #        height = 60,
-       # width = 88,
-       units = 'mm', 
-       device = cairo_ps
-       )
+#        width = 180,
+#        # height = 60,
+#        # width = 88,
+#        units = 'mm', 
+#        device = cairo_ps
+#        )
 
@@
   aes_G5 %>% 
-  select(asv, sample, treat, count, Family) %>%
+  select(asv, sample, treat, passage, count, Family) %>%
   group_by(sample, Family) %>% 
@@
   ungroup() %>% 
-  select(sample, treat, Family, total) %>% 
+  select(sample, treat, passage, Family, total) %>% 
   distinct()
+
+# prevalent_families <- 
+#   aes_family_G5 %>% 
+#   group_by(Family) %>% 
+#   summarize(prev = sum(total > 0), .groups = "drop") %>% 
+#   filter(prev / 48 > 0.1) %$%
+#   Family
 
@@
 sample_data <- aes_family_G5 %>% 
-  select(sample, treat) %>% 
+  select(sample, treat, passage) %>% 
   distinct() %>% 
@@
 
+# ps <- phyloseq(sample_data = sample_data(sample_data), 
+#                otu_table = otu_table(otu_table,
+#                taxa_are_rows = TRUE))
+
 ```
@@
 da_analysis <- differentialTest(formula = ~ treat,
-                                phi.formula = ~ 1,
+                                phi.formula = ~ treat,
                                 formula_null = ~ 1,
-                                phi.formula_null = ~ 1,
+                                phi.formula_null = ~ treat,
                                 test = "Wald", 
@@
   ggplot(dt_data, aes(x = x, xmin = xmin, xmax = xmax, y = reorder(taxa, x))) +
+        facet_wrap(~ variable) +
     geom_pointrange() +
@@
 }
+# dt_plot <- function(dt_data) {
+#   ggplot(dt_data, aes(x = x, xmin = xmin, xmax = xmax, y = reorder(taxa, x))) +
+#     geom_pointrange() +
+#     geom_vline(xintercept = 0) +
+#     labs(x = "Estimate", y = NULL) +
+#     coord_cartesian(clip="off") +
+#     # theme(panel.border = element_blank()) +
+#     theme_classic() +
+#       theme(axis.text.y = element_markdown())
+#     
+#   
+# }
 
@@
   full_join(all_iqm, by = "Family")
-# (p_log10_cl <-
-# 	all_taxa %>%
-# 	ggplot(aes(x = rel_abund, y = Family, color = treat)) +
-# 	geom_point(position=position_jitterdodge()) +
+
+(p_log10_cl <-
+	all_taxa %>%
+	ggplot(aes(x = rel_abund, y = Family, color = treat)) +
+	geom_point(position=position_jitterdodge(jitter.width = 0.1,
+	                                         dodge.width = 0.4), size = 0.5) +	
         # geom_pointrange(aes(x = median, xmax = third_quartile, xmin = first_quartile, fill = treat),
@@
 			# color = "black") +
-# 	theme_bw() +
-# 	theme(axis.text.y = element_markdown()) +
-# 	labs(x = "Relative Abundance (%)", y = NULL) +
-# 	scale_color_manual(name = "Treatment",
-# 			   labels = c('Control', 'Selection'),
-# 			   values = c('gray40', 'darkorange2')) +
-# 	scale_fill_manual(name = "Treatment",
-# 			   labels = c('Control', 'Selection'),
-# 			   values = c('gray40', 'darkorange2')) +
-#     facet_grid(direction ~ ., scale = "free_y", space = "free")  + 
-# theme(
-#   strip.background = element_blank(),
-#   strip.text.y = element_blank()
-# ) +
-# 	scale_x_log10(breaks = c(0.001, 0.01, 0.1, 1),
-# 		      labels = c(0.001, 0.01, 0.1, 1)))
+	theme_bw(base_size = 7) +
+	theme(axis.text.y = element_markdown()) +
+	labs(x = "Relative Abundance (%)", y = NULL) +
+	scale_color_manual(name = "Treatment",
+			   labels = c('Control', 'Selection'),
+			   values = c('gray40', 'darkorange2')) +
+	scale_fill_manual(name = "Treatment",
+			   labels = c('Control', 'Selection'),
+			   values = c('gray40', 'darkorange2')) +
+    facet_grid(direction ~ ., scale = "free_y", space = "free")  +
+theme(
+  strip.background = element_blank(),
+  strip.text.y = element_blank()
+) +
+	scale_x_log10(breaks = c(0.001, 0.01, 0.1, 1),
+		      labels = c(0.001, 0.01, 0.1, 1)))
 
@@
 
-# ggsave(here(fig_dir, 'fig5.jpg'), p_log10, width = 5, height = 5)
+ggsave(here(fig_dir, 'fig4.png'), p_log10_cl, 
+       width = 100, 
+       height = 88,
+       units = 'mm')
 # ggsave(here(fig_dir, 'fig5.svg'), p_log10, width = 5, height = 5)
@@
     )
-ggsave(here(fig_dir, 'fig5.eps'), 
-       p_log10_bw, 
-       width = 100, 
-       height = 88,
-       units = 'mm',
-       device = cairo_ps
-       )
+# ggsave(here(fig_dir, 'fig5.eps'), 
+#        p_log10_bw, 
+#        width = 100, 
+#        height = 88,
+#        units = 'mm',
+#        device = cairo_ps
+#        )
 
@@
 
+# Functional gene analysis with Picrust
+
+```{r, eval=FALSE}
+
+metacyc_pathways <- read_tsv(here(der_dir, 'picrust2_out_pipeline/pathways_out/path_abun_unstrat.tsv.gz'))
+metacyc_pathways <- metacyc_pathways[,1:49]
+metacyc_pathways <- as.data.frame(metacyc_pathways)
+rownames(metacyc_pathways) <- metacyc_pathways[, 1]
+metacyc_pathways <- metacyc_pathways[, -1]
+metadata <- data.frame(conds = substr(colnames(metacyc_pathways), 3, 3))
+rownames(metadata) = colnames(metacyc_pathways)
+sample_data = sample_data(metadata)
+otu_table = otu_table(metacyc_pathways, taxa_are_rows = TRUE)
+metacyc_ps <- phyloseq(otu_table, sample_data)
+taxa_names(metacyc_ps)
+# Save/Load output so you don't have to rerun ANCOM
+if(file.exists(here(der_dir, "ancom_out_metacyc.rds"))) {
+  out <- 
+    readRDS(here(der_dir, "ancom_out_metacyc.rds"))
+} else {
+  # Run Ancom
+  out = ancom(data = metacyc_ps,  
+              p_adj_method = "BH",
+              prv_cut = 0.1, 
+              lib_cut = 0, 
+              main_var = "conds",
+              rand_formula = NULL, 
+              lme_control = NULL,
+              struc_zero = TRUE, 
+              neg_lb = FALSE, 
+              alpha = 0.05, 
+              n_cl = 1)  
+  out %>% 
+    saveRDS(here(der_dir, "ancom_out_metacyc.rds"))
+}
+
+res = out$res
+q_val = out$q_data
+beta_val = out$beta_data
+# Only consider the effect sizes with the corresponding q-value less than alpha
+beta_val = beta_val * (q_val < 0.05) 
+# Choose the maximum of beta's as the effect size
+beta_pos = apply(abs(beta_val), 2, which.max) 
+beta_max = vapply(seq_along(beta_pos), function(i) beta_val[beta_pos[i], i],
+                  FUN.VALUE = double(1))
+# Number of taxa except structural zeros
+n_taxa = ifelse(is.null(out$zero_ind), 
+                ntaxa(metacyc_ps), 
+                sum(apply(out$zero_ind[, -1], 1, sum) == 0))
+# Cutoff values for declaring differentially abundant taxa
+cut_off = 0.7 * (n_taxa - 1)
+
+df_fig_w = res %>%
+  dplyr::mutate(beta = beta_max,
+                direct = case_when(
+                  detected_0.7 == TRUE & beta > 0 ~ "Positive",
+                  detected_0.7 == TRUE & beta <= 0 ~ "Negative",
+                  TRUE ~ "Not Significant"
+                  )) %>%
+  dplyr::arrange(desc(W))
+df_fig_w$taxon = factor(df_fig_w$taxon, levels = df_fig_w$taxon)
+df_fig_w$W = replace(df_fig_w$W, is.infinite(df_fig_w$W), n_taxa - 1)
+df_fig_w$direct = factor(df_fig_w$direct, 
+                     levels = c("Negative", "Positive", "Not Significant"))
+
+p_w = df_fig_w %>%
+  ggplot(aes(x = taxon, y = W, color = direct)) +
+  geom_point(size = 2, alpha = 0.6) +
+  labs(x = "Taxon", y = "W") +
+  scale_color_discrete(name = NULL) + 
+  geom_hline(yintercept = cut_off, linetype = "dotted", 
+             color = "blue", linewidth = 1.5) +
+  geom_text(aes(x = 2, y = cut_off + 0.5, label = "W[0.7]"), 
+            size = 5, vjust = -0.5, hjust = 0, color = "orange", parse = TRUE) +
+  theme_bw() +
+  theme(axis.text.x = element_blank(),
+        axis.ticks.x = element_blank(),
+        panel.grid.major = element_blank())
+p_w
+
+ancom_taxa <- as.character(df_fig_w[df_fig_w$detected_0.7, ]$taxon)
+metacyc_pathways_descrip <- read_tsv(here(der_dir, 'picrust2_out_pipeline/pathways_out/path_abun_unstrat_descrip.tsv.gz')) %>% 
+    select(pathway, description)
+metacyc_pathways_descrip[metacyc_pathways_descrip$pathway %in% ancom_taxa,] ->
+    sig_pathways
+```
+
 ```{r save_outputs}

### 2023-06-23 manuscript/manuscript.Rmd
remove header
update with edits for reviewers

@@
 microbiomes can covary strongly with other drivers of ecosystem function, such
-as the environment. We propose that artificial ecosystem selection can overcome
-these challenges. Artificial ecosystem selection involves creating replicate
-ecosystems, selecting replicates with a desired ecosystem-level trait, and
-using the microbiomes from the selected ecosystems to inoculate a new set of
-ecosystems. We performed artificial ecosystem selection on replicate soil
-microbiomes to select for increased soil CH~4~ oxidation rate. We observed a
-strong response to selection with a `r myround(per_slope, 1)`\% increase in
-CH~4~ oxidation rate per passage. We
+as the environment. To address these challenges, we imposed artificial
+selection on whole soil ecosystems over multiple generations to select for microbial communities
+with a high rate of CH~4~ oxidation. This approach
+is potentially powerful because it is biologically "agnostic" in 
+that it makes few assumptions about which taxa are important to function, and
+repeated passaging with fresh substrate weakens the covariance between microbes
+and the environment. As a response to selection, we observed a
+ `r myround(per_slope, 1)`\% increase in
+CH~4~ oxidation rate per passage relative to a control that experienced
+random selection. We
 estimated that `r myround(tidy(rh_model)[[2, 2]]*100, 1)`\% of the variation in
 CH~4~ oxidation
-rate in these soils can be attributed to microbiome variation. We also found
+rate in these soils can be attributed to microbiome variation (though this was not significant). We also found
 that selection did not enrich for known CH~4~ oxidizers; instead, `r length(da_fams)` families not
@@
 microbial pathway. Our study demonstrates that variation in microbiome
-composition can contribute to variation in the rate of ecosystem function and
-that manipulating microbiome composition could be a viable strategy for managing
+composition can contribute to variation in the rate of ecosystem function 
+independent of the environment and that this may not always be limited by the final step
+in a pathway. This suggests
+that manipulating microbiome composition directly without altering the environment 
+could be a viable strategy for managing
 ecosystem functions.
@@
 the microbiome to variation in ecosystem function and for investigating the
-mechanisms underlying those relationships. By repeated passaging of microbiomes
+mechanisms underlying those relationships. Through repeated passaging of microbiomes
 in a common environment, we can weaken the covariance between microbes and the
@@
 
-We performed an artificial ecosystem selection experiment 
-(*sensu* @swenson2000) by passaging replicate soil microbiomes. The trait we
-selected on was CH~4~ oxidation rate. This experiment had a selection line and a
-control line with twelve jars each for a total of twenty-four jars per passage.
-The selection line underwent positive selection where the two or three jars
+We performed an artificial ecosystem selection experiment (*sensu* @swenson2000)
+by passaging replicate soil microbiomes. The trait we selected on was soil CH~4~
+oxidation rate.  Soil microcosms were incubated at room temperature in sealed
+500 mL glass jars with a rubber septum for gas sampling. Each jar was sterilized
+with 70\% ethanol and was composed of 45 g of autoclaved artificial potting mix,
+5 g of living soil inoculum, and 3.5 mL of sterile deionized water to bring the
+soil to 60\% of field capacity. The potting mix consisted of bark fines, peat
+moss, pumice, sand, composted manure, and biochar (Lane Potting Mix, Lane Forest
+Products, Eugene, OR). The initial soil microbiome inoculum was sampled from the
+top 10 cm of an upland mineral soil under a deciduous forest ecosystem near the
+University of Oregon campus in Eugene, OR, USA. Each jar
+was capped and injected with 4.3 mL of 99\% CH~4~, which produced a mean
+headspace concentration of 763.9 ppm (SD = 183.1). Twice per week, jars were
+flushed in a biosafety cabinet (to avoid contamination) and respiked with CH~4~ 
+to maintain aerobic conditions and elevated CH~4~ concentrations.
+
+For the selection experiment, we created two lines of soil microcosms with 12
+jars each: a control line with random selection and an experimental line with
+directional selection for greater soil CH~4~ oxidation rate. The selection line
+underwent positive selection where the two or three jars
 with the highest CH~4~ oxidation rate were homogenized to inoculate the next set
@@
 jars. The number of jars chosen was based on the distribution of fluxes among
-the positive jars, i.e., we chose the top three jars unless only two jars had
-considerably greater CH~4~ oxidation rates based on visual inspection of
-histograms. The experiment was carried out over five passages until a
-significant divergence in functional rates was observed between the two
-selection treatments based on a difference of slopes.
-
-The initial soil microbiome was sampled from the top 10 cm of an upland mineral
-soil under a deciduous forest ecosystem near the University of Oregon campus in
-Eugene, OR, USA. Incubations were performed in 500 mL mason jars with rubber
-septa installed in the lids. Each jar was sterilized with 70\% ethanol to which
-was added 45 g of autoclaved potting mix, 5 g of living soil, and 3.5 mL of
-sterile deionized water to bring them to 60\% of field capacity. The potting
-mix was Lane Potting Mix from Lane Forest Products (Eugene, OR, USA). Each jar
-was then capped and injected with 4.3 mL of 99\% CH~4~ to bring
-the headspace concentration to approximately 1000 ppm CH~4~. To create the two
-treatment groups, twenty-four jars were created in an identical manner and then
-randomly assigned to either the selection or control line. Jars were
-flushed and respiked with CH~4~ twice per week to maintain aerobic conditions
-and elevated CH~4~ concentrations and were incubated at ambient temperature for
-approximately four weeks. Methane oxidation rates were determined at the end of
+the positive jars: three jars were chosen in all generations except for passage 3
+where two jars were selected. The experiment was carried out over five passages
+with an average
+incubation time per generation of four weeks. 
+Methane oxidation rates were determined at the end of
 the incubation period and selection was performed. For each treatment, the
-selected jars were homogenized and 5 g of the homogenized soil was used to
-inoculate the next set of jars, which represents a 10% subsampling. The next set
+selected jars were homogenized and 5 g of the homogenized soil was used as the
+living soil inoculum for the next generation. The next set
 of jars were created in an identical manner to the first generation with fresh
@@
 Methane oxidation rates were determined after flushing and spiking jars to
-approximately 1000 ppm CH~4~. Headspace samples of 1 mL were collected from
+1000 ppm CH~4~. Headspace samples of 1 mL were collected from
 each jar immediately after spiking and then at time points 3, 6, 24, and 48
@@
 constant (k, units = d^−1^; i.e., dCH~4~/dt = k[CH~4~]) of the exponential
-decrease in CH~4~ Oxidation rates are presented as the additive inverse of
+decrease in CH~4~. Oxidation rates are presented as the additive inverse of
 $k$ (i.e., $-k$) so that a more positive value represents a greater
-oxidation rate. The jars selected in passage two for the positive selection treatment had
+oxidation rate. The jars selected to inoculate passage three for the positive 
+selection treatment had
 the lowest CH~4~ oxidation rate of the twelve jars due to a calculation error
-in the rate constant. Despite this error, we still observed a strong
-response to selection demonstrating the power of this technique. In addition,
-this likely reintroduced diversity lost during the selection process,
-potentially increasing the variation available to be selected upon, as
-is recommended for microbiome selection experiments [@sanchez2021].
+in the rate constant.
 
@@
 presence of contaminants was evaluated using both the prevalence and frequency
-methods from 'DECONTAM' [@davis2018]. Decontam identified 
+methods from 'DECONTAM' by comparing samples to extraction controls of 
+water [@davis2018]. Decontam identified 
 `r length(which(contamdf$contaminant))` potential contaminants based on
@@
 and normal distribution. Therefore, CH~4~ oxidation rates were log~10~
-transformed prior to analysis. First, we tested if there was a difference of
+transformed prior to analysis. Following transformation, these assumptions
+were met. First, we tested if there was a difference of
 slopes between the selection line and the control based on the interaction
@@
 line and the control on the cumulative selection differential [@falconer1996].
+This estimate is analogous
+to estimates of "microbiability" from the animal breeding literature, which
+quantifies the
+variation in a host trait that is due to microbiome variation [@he2022]. 
+The slope of the regression of divergence on cumulative selection differential
+provides an estimate of realized microbiability (h^2^ &pm; SE).
 Divergence was calculated as the mean CH~4~ oxidation rate of the positive
@@
 differential from all preceding selection events. We then regressed cumulative
-divergence on cumulative selection differential using the 'lm' function.
+divergence on cumulative selection differential using the 'lm' function. We report
+the slope as percent change by back-calculating the percent change from the
+log-transformed data into the original units using the formula $(10^\beta - 1) * 100$
+where $\beta$ is the slope.
 
@@
 differential abundance between the two treatments in passage 5. We first
-grouped ASVs at the family level. Any ASVs that lacked a family-level
+grouped ASVs at the family level. We chose this level of agglomeration
+because CH~4~ oxidation
+is a relatively deeply conserved function [@martiny2013] and is restricted to
+a handful of bacterial and archaeal families [@kang2019]. Therefore, we are most
+likely to detect an enrichment of methanotrophs at this taxonomic scale.
+Any ASVs that lacked a family-level
 taxonomic assignment were grouped at a higher taxonomic level. We then subset the
@@
 methanotrophic families and tested their differential abundance aggregated at
-the family and genus level using 'corncob'.
+the family and genus level using 'corncob'. For each test, p-values were adjusted
+for multiple testing by controlling the false discovery rate using the 
+Banjamini-Hochberg procedure [@benjamini1995].
 
@@
 To estimate the proportion of variation in CH~4~ oxidation rate due to
-variation in microbiome composition, we regressed divergence between the
+variation in microbiome composition--i.e., microbiability (as described in the Methods; [@he2022]--we regressed divergence between the
 positive selection treatment and the control against the cumulative selection
-differential (`r figures("herit", display = "cite")`). This estimate is analogous
-to estimates of "microbiability" from the animal breeding literature, which
-quantifies the
-variation in a host trait that is due to microbiome variation [@difford2016]. 
-The slope of the regression of divergence on cumulative selection differential
-provides an estimate of realized microbiability (h^2^ &pm; SE), which in our
-experiment was `r myround(tidy(rh_model)[[2, 2]], 2)` &pm;
+differential. The microbiability was `r myround(tidy(rh_model)[[2, 2]], 2)` &pm;
 `r myround(tidy(rh_model)[[2, 3]], 2)`, though this was not significant (F~`r glance(rh_model)$df`,`r glance(rh_model)$df.residual`~ = 
@@
 difference in richness between the selection treatment and the control in
-passage 2 or 5 (pairwise Wilcoxon test: p = `r myround(wt[1, 3], 1)`).  In addition, there was no
+passage 2 or 5 (pairwise Wilcoxon test; Passage 2: p = `r myround(wt[1, 3], 2)`,
+Passage 5: p = `r myround(wt[6, 3], 2)`).  In addition, there was no
 correlation between richness and CH~4~ oxidation rate across the two treatments in passage 5 (Spearman's rho = `r myround(s5[1,1], 1)`, p = `r myround(s5[1,3], 1)`).
@@
 
-This relationship between microbiome variation and ecosystem function variation
+The relationship between microbiome variation and ecosystem function variation
 is analogous to the concept of "heritability" [@lynch1998] used by quantitative
-geneticists, or more precisely the concept of "microbiability" [@difford2016] proposed
+geneticists, or more precisely the concept of "microbiability" [@he2022] proposed
 by microbiome scientists who study host-associated microbiomes. Although rarely
@@
 `r myround(tidy(rh_model)[[2, 2]]*100, 1)`\% of the
-variation we observed in the rate of CH~4~ oxidation.  This suggests that
+variation we observed in the rate of CH~4~ oxidation, though this was not
+significant. However, we did observe a significant divergence between the 
+positive selection and control lines, which suggests that the imposed
+selection and passaging of microbiomes was sufficient to generate
+variation in soil CH~4~ oxidation rate. Future studies with
+greater replication could more precisely estimate the microbiability.
+This suggests that
 there is substantial potential for altering this ecosystem function through
@@
 
-Given that variation in the microbiome is driving substantial variation in
-CH~4~ oxidation rate in our system, we next wanted to determine which aspects of the
-microbiome best explain this contribution of microbiome variation to
-CH~4~ oxidation rate. There are three inter-related ways that microbiomes could have responded
+We next wanted to determine which aspects of the
+microbiome might explain the divergence in CH~4~ oxidation rate between the two treatments.
+There are three inter-related ways that microbiomes could have responded
 to selection in this experiment: gain or loss of taxa, changes in the
@@
 treatment and a difference in composition between the two treatments, we did
-not observe an increase in the relative abundance of methanotrophs. This was
+not observe an increase in the relative abundance of known methanotrophs. This was
 surprising given that CH~4~ consumption is not a common trait among microbes
@@
 abundance of methanogens and methanotrophs as estimated from marker genes
-[@meyer2020; @meyer2020a]. However, our results suggest that
+[@meyer2020; @meyer2020a]. However, our results suggest that in this system
 ecosystem-scale CH~4~ oxidation rates can be altered by non-methanotrophs,
-perhaps through ecological interactions with methanotrophic species. This
+perhaps through ecological interactions with methanotrophic species, or by
+unknown methanotrophs. This
 suggests that simple assumptions about how microbes contribute to rate variation
@@
 
-# Supplement
-
 # References
@@
 
-`r figures("herit", paste0("Divergence between the positive selection line and the control line over the cumulative selection differential.
-Values are CH~4~ oxidation rate as the additive inverse of the first-order exponential decay constant k (day^-1^).
-Regression of divergence on cumulative selection differential provides an estimate of realized heritability (h^2^ &pm; SE) and the slope of the regression was ", myround(tidy(rh_model)[[2, 2]], 2), " &pm; ", myround(tidy(rh_model)[[2, 3]], 2), ", though this was not significant (F~", glance(rh_model)$df,  ",", glance(rh_model)$df.residual,  "~ = ", myround(glance(rh_model)$statistic, 2),  ", p =  ", myround(glance(rh_model)$p.value, 2),  ")."))`
-
 `r figures("beta", "Non-metric multidimensional scaling plot of beta diversity for all jars.

### 2023-06-23 R/functions.R

@@ -7,7 +7,10 @@ man_dir <- 'manuscript/'       # Manuscript directory
 fig_dir <- 'analysis/figs/'      # Figures directory
 
 # Set global ggplot theme
-
+theme_set(theme_classic() +
+              theme(panel.border = element_rect(fill = NA, linewidth = 0.5),
+                    axis.line = element_blank(),
+              ))
 theme_set(theme_classic() +
             theme(panel.border = element_rect(fill = NA, linewidth = 1.0625),
                   axis.line = element_blank(),

### 2023-06-23 analysis/scripts/01-calculate-flux.R

@@ -70,6 +70,9 @@ ch4_conc_data <-
   select(flux_date, jar, t, ppm) %>% 
   spread(t, ppm)
 
+ch4_conc_data %>% 
+    summarize(mean = mean(t1), sd = sd(t1))
+
 # Finally, this step takes the time intervals and ch4 concentrations and fits a
 # first-order exponential decay function to compute methane flux as `k` which is
 # the rate-constant for the first-order exponential decay function. Here, I

### 2023-06-23 analysis/scripts/05-combine-asvs-and-fluxes.R

@@ -29,6 +29,10 @@ sum(colnames(seqtab.nochim) != rownames(taxa))
 # They match if = 0 
 seqs <- tibble(asv = paste0("ASV", seq(ncol(seqtab.nochim))),
                seq = colnames(seqtab.nochim))
+seq_out <- seqs
+names(seq_out) <- c('seq.name', 'seq.text')
+phylotools::dat2fasta(seq_out, 'analysis/output/refseqs.fna')
+
 colnames(seqtab.nochim) <- paste0("ASV", seq(ncol(seqtab.nochim)))
 rownames(seqtab.nochim) <- metadata$sample
 rownames(taxa) <- colnames(seqtab.nochim)
@@ -36,6 +40,17 @@ rownames(taxa) <- colnames(seqtab.nochim)
 asvs <-
   seqtab.nochim %>% 
   as_tibble(rownames = "sample")
+
+asvs %>% 
+    pivot_longer(-sample, names_to = 'asv', values_to = 'count') %>% 
+    pivot_wider(names_from = sample, values_from = count)%>% 
+    as.data.frame ->
+    asv_to_biom
+
+rownames(asv_to_biom) <- asv_to_biom[,1]
+asv_to_biom <- asv_to_biom[,-1]
+asv_biom <- biomformat::make_biom(asv_to_biom)
+biomformat::write_biom(asv_biom, 'analysis/output/asv.biom')
 
 rm(seqtab.nochim)
 

### 2023-06-23 analysis/scripts/compare-methods.R

@@ -1,0 +1,178 @@
+library(vegan)
+library(breakaway)
+library(magrittr)
+library(zCompositions)
+library(tidyverse)
+source('R/functions.R')
+
+set.seed(19760620)
+
+aes_data <- read_tsv('analysis/output/community_data.tsv')
+
+# Look at distribution of sequencing depths
+aes_data %>% 
+  group_by(sample) %>% 
+  summarize(total = sum(count)) %>% 
+  ggplot(aes(x = total)) +
+  geom_histogram()
+
+
+seq_range <- 
+  aes_data %>%
+  group_by(sample) %>% 
+  summarize(total = sum(count)) %>% 
+  summarize(seq_range = range(total)) %$%
+  seq_range
+
+# Generate a null dataset
+
+rand_aes <- randomize_asv_table(aes_data)
+
+#######################
+# Alpha diversity
+#####################
+
+
+analysis <- 
+  rand_aes %>%
+  count(sample, value) %>%
+  nest(data = -sample) %>%
+  mutate(n_seqs = map_dbl(data, ~sum(.x$value * .x$n)),
+         sobs = map_dbl(data, ~sum(.x$n)),
+         rare = map_dbl(data, ~rarefy(rep(.x$value, .x$n), sample=seq_range[1])),
+         chao = map_dbl(data, ~get_chao(.x)),
+         ba = map(data, ~get_breakaway(.x))) %>%
+  dplyr::select(-data) %>%
+  unnest(ba)
+
+
+analysis %>%
+  dplyr::select(sample, n_seqs, sobs, chao, rare, est) %>%
+  pivot_longer(-c(sample, n_seqs)) %>%
+  ggplot(aes(x=n_seqs, y=value, color=name)) +
+  geom_point() +
+  geom_smooth(se=FALSE) +
+  coord_cartesian(ylim=c(0, 40000)) +
+  theme_classic() +
+  labs(title = "AES", x = "Sequencing Depth", y = "Alpha Diversity")
+
+analysis %>%
+  dplyr::select(sample, n_seqs, sobs, chao, rare, est) %>%
+  pivot_longer(-c(sample, n_seqs)) %>%
+  ggplot(aes(x=n_seqs, y=value, color=name)) +
+  geom_point() +
+  geom_smooth(se=FALSE) +
+  coord_cartesian(ylim=c(0, 10000)) +
+  theme_classic() +
[…]

### 2023-06-23 analysis/scripts/picrust.sh

@@ -1,0 +1,5 @@
+#!/bin/bash
+
+picrust2_pipeline.py -s analysis/output/refseqs.fna -i analysis/output/asv.biom -o analysis/output/picrust2_out_pipeline -p 3
+add_descriptions.py -i analysis/output/picrust2_out_pipeline/pathways_out/path_abun_unstrat.tsv.gz -m METACYC \
+                    -o analysis/output/picrust2_out_pipeline/pathways_out/path_abun_unstrat_descrip.tsv.gz

### 2023-06-23 R/functions.R
cleanup for publication

@@ -4,7 +4,7 @@
 raw_dir <- 'data/'      # Raw Data Directory
 der_dir <- 'analysis/output/'    # Derived/Modified Data Directory
 man_dir <- 'manuscript/'       # Manuscript directory
-fig_dir <- 'analysis/figs/'      # Figures directory
+fig_dir <- 'manuscript/figs/'      # Figures directory
 
 # Set global ggplot theme
 theme_set(theme_classic() +

### 2023-06-23 analysis/scripts/01-calculate-flux.R

@@ -70,8 +70,8 @@ ch4_conc_data <-
   select(flux_date, jar, t, ppm) %>% 
   spread(t, ppm)
 
-ch4_conc_data %>% 
-    summarize(mean = mean(t1), sd = sd(t1))
+#ch4_conc_data %>% 
+#    summarize(mean = mean(t1), sd = sd(t1))
 
 # Finally, this step takes the time intervals and ch4 concentrations and fits a
 # first-order exponential decay function to compute methane flux as `k` which is

### 2023-06-23 analysis/scripts/05-combine-asvs-and-fluxes.R

@@ -112,42 +112,43 @@ aes_data %>%
   scale_x_continuous(breaks = 1:9)
 
 ## Top 20 taxa in controls
-
-top20_taxa_in_treatment <- function(x, treatment) {
-x %>% 
-  filter(treat == treatment) %>% 
-  select(-sample, -row, -col, -qubit_ngml) %>% 
-  group_by(asv) %>% 
-  mutate(count = sum(count)) %>%  
-  ungroup() %>% 
-  distinct() %>% 
-  arrange(desc(count)) %>% 
-  slice_head(n = 20)
-}
-
-### Potting Mix
-
-top20_taxa_in_treatment(aes_data, "PM0")
-
-
-### Negative Control
-
-top20_taxa_in_treatment(aes_data, "NC0")
-
-
-### PCR Positive Control
-
-top20_taxa_in_treatment(aes_data, "PCRP")
-
-
-### PCR Negative Control
-
-top20_taxa_in_treatment(aes_data, "PCRN")
-
-
-### Soil Core
-
-top20_taxa_in_treatment(aes_data, "SC0")
+#
+#top20_taxa_in_treatment <- function(x, treatment) {
+#    print(paste0('Top 20 taxa in ', treatment))
+#x %>% 
+#  filter(treat == treatment) %>% 
+#  select(-sample, -row, -col, -qubit_ngml) %>% 
+#  group_by(asv) %>% 
+#  mutate(count = sum(count)) %>%  
+#  ungroup() %>% 
+#  distinct() %>% 
+#  arrange(desc(count)) %>% 
+#  slice_head(n = 20)
+#}
+#
+#### Potting Mix
+#
+#top20_taxa_in_treatment(aes_data, "PM0")
+#
+#
+#### Negative Control
+#
+#top20_taxa_in_treatment(aes_data, "NC0")
+#
+#
+#### PCR Positive Control
+#
+#top20_taxa_in_treatment(aes_data, "PCRP")
+#
+#
+#### PCR Negative Control
+#
+#top20_taxa_in_treatment(aes_data, "PCRN")
+#
+#
+#### Soil Core
+#
+#top20_taxa_in_treatment(aes_data, "SC0")
 
 # Sequence Depth Distribution
 

### 2023-06-23 manuscript/manuscript.Rmd
Add license
update figure captions

@@
 ---
-title: Artificial ecosystem selection reveals relationships between microbiome composition and ecosystem function
+title: Response of soil microbiome composition to selection on methane oxidation rate
 author: "Andrew H. Morris^1^* and Brendan J. M. Bohannan^1^"
@@
 The y-axis is CH~4~ oxidation rate as the log~10~ of the additive inverse of the first-order exponential decay constant $k$ (i.e., $-k$) with units day^-1^ so that a more positive value represents a higher CH~4~ oxidation rate.
-Closed points are the positive selection treatment and open points are the control.
-The solid line is the regression for the positive selection treatment and the dotted line is the regression for the control.
+Orange points and regression line are for the positive selection treatment and gray points and regression line are for the control.
 There was a significant difference of slopes between the positive selection treatment and the control (F~", compare_response_mods[[2,4]], ",", compare_response_mods[[2,2]], "~ = ", myround(compare_response_mods[[2,6]], 2), ", p = ", myround(compare_response_mods[[2,7]], 2), ")."))`
@@
 Dissimilarities are rarefied Bray-Curtis dissimilarity averaged over 100 subsamples.
-Closed points are the positive selection treatment and open points are the control.
-Triangles are passage 2 and circles are passage 5.")`
+Orange points are the positive selection treatment and gray points are the control.
+Circles are passage 2 and triangles are passage 5.")`
 

### 2023-12-22 manuscript/manuscript.Rmd
update ms for submission to SBB

@@
     reference_docx: reference.docx
+    number_sections: true
 bibliography: bibliography.bib
-csl: vancouver.csl
+csl: soil-biology-and-biochemistry.csl
 ---
@@
 
-# Materials and Methods
+# Materials and methods
 
@@
 
-## Statistical Analysis
+## Statistical analysis
+
+### Response to selection
 
@@
 
+### Microbiability
+
 We estimated the proportion of variation in CH~4~ oxidation rate due to
-variation in the microbiome as the regression of divergence between the positive
-line and the control on the cumulative selection differential [@falconer1996].
+variation in the microbiome by comparing the response to selection to the 
+imposed selection differential [@falconer1996].
 This estimate is analogous
@@
 variation in a host trait that is due to microbiome variation [@he2022].
-The slope of the regression of divergence on cumulative selection differential
+
+For this purpose, the response to selection was calculated as the divergence between the selection treatment and the control. Divergence was the difference between the mean CH~4~ 
+oxidation rate of the positive
+treatment and the mean CH~4~ oxidation rate of the control in each
+passage. In addition, the selection differential was calculated as the difference between
+the mean of the three selected jars and the mean of all twelve jars in a passage. The slope of the divergence over the cumulative selection differential
 provides an estimate of realized microbiability (h^2^ &pm; SE).
-Divergence was calculated as the mean CH~4~ oxidation rate of the positive
-treatment minus the mean CH~4~ oxidation rate of the control in each
-passage. The selection differential was calculated as the difference between
-the mean of the three selected jars and the mean of all twelve jars in a passage.
-Cumulative selection differential was calculated as the sum of the selection
-differential from all preceding selection events. We then regressed cumulative
+We regressed cumulative
 divergence on cumulative selection differential using the 'lm' function. We report
@@
 where $\beta$ is the slope.
+
+### Alpha and beta diversity
 
@@
 999 permutations [@mcardle2001; @oksanen2019]
+
+### Differential abundance
 
@@
 To estimate the proportion of variation in CH~4~ oxidation rate due to
-variation in microbiome composition--i.e., microbiability (as described in the Methods; [@he2022]--we regressed divergence between the
+variation in microbiome composition--i.e., microbiability (as described in the Methods; [@he2022])--we regressed divergence between the
 positive selection treatment and the control against the cumulative selection
@@
 no difference in dispersion between treatments or passages (F~3,44~ = 0.91, p = 0.45). 
-Finally, CH~4~ oxidation rate was correlated with Bray-Curtis dissimilarity across both treatments in passage 5 and explained `r myround(dbrda_percent_explained[1], 1)`\% of the variation in Bray-Curtis dissimilarity (dbRDA: F~`r dbrda_aov[1, 1]`,`r dbrda_aov[2, 1]`~ = `r myround(dbrda_aov[1, 3], 2)`, p = `r myround(dbrda_aov[1, 4], 3)`)
+Finally, CH~4~ oxidation rate was correlated with Bray-Curtis dissimilarity across both treatments in passage 5 and explained `r myround(dbrda_percent_explained[1], 1)`\% of the variation in Bray-Curtis dissimilarity (dbRDA: F~`r dbrda_aov[1, 1]`,`r dbrda_aov[2, 1]`~ = `r myround(dbrda_aov[1, 3], 2)`, p = `r myround(dbrda_aov[1, 4], 2)`)
 
@@
 on methanol [@tamas2014; @dedysh2016]. Based on this analysis,
-it appears that no methanotrophs were enriched in the positive selection treatment.
+it appears that no known methanotrophs were enriched in the positive selection treatment.
 
@@
 because it can control for both the direct effect of environment on function as
-well as the indirect effect on environment via its impact on microbiome
+well as the indirect effect of environment via its impact on microbiome
 assembly.
@@
 selection treatment relative to the control, which demonstrates that there was a
-response to selection. Given that we observed a response to selection, we
+response to selection. In addition, we observed a significant correlation 
+between CH~4~ oxidation rate and overall microbiome composition. Given these observations, we
 conclude that variation in the microbiome contributes to variation in the
@@
 Given that variation in the microbiome is associated with variation in the rate
-of an ecosystem function in our system, a reasonable follow-up question is "how
-much variation in ecosystem function is associated with microbiome variation in
-this system?" One way to estimate this is to determine how much the recipient
-jars resemble the selected donor jars that were used to inoculate them
-[@falconer1996]. We can
-calculate the response to selection as the difference between two successive
-passages in their mean CH~4~ oxidation rate. We will denote this as $R$. We
-can also calculate the strength of selection as the difference in mean CH~4~
-oxidation rate between the twelve jars in one generation and the three jars chosen
-for selection in that generation, which we will call the selection differential
-and denote as $S$. If we plot the cumulative $R$ against the cumulative $S$, the
-slope of this relationship will equal the proportion of variation explained by
-the microbiome. If the change in mean function from passage one to passage two
-($R$) is equal to the difference in mean function between the twelve jars in
-passage one and the three jars selected to inoculate passage two ($S$), then we
-would conclude that 100% of the variation is due to variation in the microbiome.
-Likewise, if recipients do not resemble the donors in their mean CH~4~
-oxidation rate and simply wander randomly, then we would conclude that all of
-the variation is due to the environment or technical variation.
-
+of an ecosystem function in our system, we next wanted to estimate how
+much variation in ecosystem function is associated with microbiome variation.
 The relationship between microbiome variation and ecosystem function variation
@@
 by microbiome scientists who study host-associated microbiomes. Although rarely
-used in the study of environmental microbiomes, this concept could be very
+used in the study of environmental microbiomes, this concept could be
 useful for understanding and manipulating microbially-mediated functions in a
-variety of ecosystems. In our experiment, variation in microbiome taxonomic
-composition statistically explained (i.e., was associated with) 
-`r myround(tidy(rh_model)[[2, 2]]*100, 1)`\% of the
-variation we observed in the rate of CH~4~ oxidation, though this was not
-significant. However, we did observe a significant divergence between the 
-positive selection and control lines, which suggests that the imposed
-selection and passaging of microbiomes was sufficient to generate
-variation in soil CH~4~ oxidation rate. Future studies with
+variety of ecosystems.
+
+In our experiment, the microbiability was not statistically
+significant and, therefore, we were not able to accurately estimate
+the percentage of variation in CH~4~ oxidation rate explained by the microbiome. 
+However, we did observe a significant divergence between the 
+positive selection and control lines as a result of passaging the microbiome.
+This suggests that the imposed
+selection was sufficient to generate
+variation in soil CH~4~ oxidation rate via the microbiome. Future studies with
 greater replication could more precisely estimate the microbiability.
-This suggests that
-there is substantial potential for altering this ecosystem function through
-microbiome manipulation in this soil. It is very likely that the
+Even though we attempted to estimate the microbiability in this study, 
+it is likely that the
 “environmental microbiability” will be different for other ecosystem functions
 in this soil and for CH~4~ oxidation in other soils. However, our experiment
-demonstrates that this relationship is measurable and provides an example of how
-this can be accomplished.
+provides an example of how
+this relationship can be measured
 
@@
 
+# Conclusions
+
 There is increasing interest in using artificial selection for understanding and
@@
 approach can provide unique information about the independent contribution of
-microbiomes to ecosystem functions. Such information is crucial if we are to
+microbiomes to ecosystem functions. For example, we demonstrated that passaging
+of soil microbiomes can produce a response to selection on soil CH~4~ oxidation
+rate. This demonstrates that microbiome variation can contribute to variation in 
+ecosystem function independent of the environment.
+Such information is crucial if we are to
 successfully manipulate environmental microbiomes to alter ecosystem functions,
@@
 
-# Acknowledgments
+# Acknowledgments {.unnumbered}
 
@@
 
-# Competing Interests
+# Competing interests {.unnumbered}
 
@@
 
-# Data Availability Statement
+# Data availability statement {.unnumbered}
 
@@
 
-# References
+# References {.unnumbered}
 
@@
 
-# Figure Captions
+# Figure Captions {.unnumbered}
 

### 2024-03-08 manuscript/manuscript.Rmd
fix typo in methane spike volume
adjust spike description

@@
 University of Oregon campus in Eugene, OR, USA. Each jar
-was capped and injected with 4.3 mL of 99\% CH~4~, which produced a mean
+was capped and injected with 0.43 mL of 99\% CH~4~, which produced a mean
 headspace concentration of 763.9 ppm (SD = 183.1). Twice per week, jars were
@@
 
-Methane oxidation rates were determined after flushing and spiking jars to
-1000 ppm CH~4~. Headspace samples of 1 mL were collected from
+Methane oxidation rates were determined after flushing and spiking jars with 
+0.43 mL of 99\% CH~4~. Headspace samples of 1 mL were collected from
 each jar immediately after spiking and then at time points 3, 6, 24, and 48
